### 2020-11-13 ch1_introduction/introduction.tex
Initial commit

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                              01 - Introduction                              %
%=============================================================================%

\chapter{Introduction}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Hendrerit gravida rutrum quisque non tellus orci ac auctor augue. Eros donec ac odio tempor orci dapibus. Ultrices neque ornare aenean euismod elementum nisi. Vivamus arcu felis bibendum ut tristique et egestas quis ipsum. Elit ut aliquam purus sit amet luctus. Et malesuada fames ac turpis egestas sed. Egestas dui id ornare arcu odio ut sem nulla pharetra. Porta lorem mollis aliquam ut porttitor leo a diam. Ultricies tristique nulla aliquet enim. Urna nunc id cursus metus aliquam. Faucibus a pellentesque sit amet porttitor. Tincidunt ornare massa eget egestas purus. Venenatis tellus in metus vulputate eu scelerisque felis imperdiet. Posuere urna nec tincidunt praesent. Risus pretium quam vulputate dignissim suspendisse in.

Aliquam id diam maecenas ultricies mi eget. Mauris pharetra et ultrices neque ornare aenean euismod. Vitae suscipit tellus mauris a diam maecenas. Massa massa ultricies mi quis hendrerit. Tincidunt eget nullam non nisi est sit amet. Vitae purus faucibus ornare suspendisse sed nisi. Enim sit amet venenatis urna. Malesuada fames ac turpis egestas. Viverra nam libero justo laoreet sit amet. Ornare massa eget egestas purus viverra accumsan in nisl nisi. Ac feugiat sed lectus vestibulum mattis.

\autocite{Schlund2020}
\autocite{Schlund2020a}

### 2020-11-13 end_appendix/appendix.tex
Initial commit

--- 2020-11-13 ch1_introduction/introduction.tex
@@
 %=============================================================================%
-%                              01 - Introduction                              %
+%                                  Appendix                                   %
 %=============================================================================%
 
-\chapter{Introduction}
+\renewcommand{\thesection}{\Alph{section}}
+\addchap{Appendix}
+
+\section{Circle of fifths metric: Proof of the metric properties}
+\label{sec:app:cof_metric_proof}
 
@@
 
-Aliquam id diam maecenas ultricies mi eget.
-Mauris pharetra et ultrices neque ornare aenean euismod.
-Vitae suscipit tellus mauris a diam maecenas.
-Massa massa ultricies mi quis hendrerit.
-Tincidunt eget nullam non nisi est sit amet.
-Vitae purus faucibus ornare suspendisse sed nisi.
-Enim sit amet venenatis urna.
-Malesuada fames ac turpis egestas.
-Viverra nam libero justo laoreet sit amet.
-Ornare massa eget egestas purus viverra accumsan in nisl nisi.
-Ac feugiat sed lectus vestibulum mattis.
+\section{Source code}
 
-\autocite{Schlund2020}
-\autocite{Schlund2020a}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Hendrerit gravida rutrum quisque non tellus orci ac auctor augue.
+Eros donec ac odio tempor orci dapibus.
+Ultrices neque ornare aenean euismod elementum nisi.
+Vivamus arcu felis bibendum ut tristique et egestas quis ipsum.
+Elit ut aliquam purus sit amet luctus.
+Et malesuada fames ac turpis egestas sed.
+Egestas dui id ornare arcu odio ut sem nulla pharetra.
+Porta lorem mollis aliquam ut porttitor leo a diam.
+Ultricies tristique nulla aliquet enim.
+Urna nunc id cursus metus aliquam.
+Faucibus a pellentesque sit amet porttitor.
+Tincidunt ornare massa eget egestas purus.
+Venenatis tellus in metus vulputate eu scelerisque felis imperdiet.
+Posuere urna nec tincidunt praesent.
+Risus pretium quam vulputate dignissim suspendisse in.

### 2020-11-13 end_statutory_declaration/statutory_declaration.tex
Adapted file permissions

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                            Statutory declaration                            %
%=============================================================================%

\addchap{Statutory Declaration}

I assure that this thesis is a result of my personal work and that no other than the indicated aids have been used for its completion. Furthermore I assure that all quotations and statements that have been inferred literally or in a general manner from published or unpublished writings are marked as such. Beyond this I assure that the work has not been used, neither completely nor in parts, to pass any previous examination.

\vspace{20mm}

\noindent \thePlace{}, \theDate{}

\begin{flushright}
  $\overline{~~~~~~~~~~~~~\raisebox{0mm}[5mm]{\theAuthor{}}~~~~~~~~~~~~~}$
\end{flushright}

### 2020-11-14 02_abstract/abstract_german.tex
Adapted file permissions
Initialized chapters

--- 2020-11-13 ch1_introduction/introduction.tex
@@
 %=============================================================================%
-%                              01 - Introduction                              %
+%                              Abstract (German)                              %
 %=============================================================================%
 
-\chapter{Introduction}
+\chapter*{Abstract (German version)}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Hendrerit gravida rutrum quisque non tellus orci ac auctor augue.
-Eros donec ac odio tempor orci dapibus.
-Ultrices neque ornare aenean euismod elementum nisi.
-Vivamus arcu felis bibendum ut tristique et egestas quis ipsum.
-Elit ut aliquam purus sit amet luctus.
-Et malesuada fames ac turpis egestas sed.
-Egestas dui id ornare arcu odio ut sem nulla pharetra.
-Porta lorem mollis aliquam ut porttitor leo a diam.
-Ultricies tristique nulla aliquet enim.
-Urna nunc id cursus metus aliquam.
-Faucibus a pellentesque sit amet porttitor.
-Tincidunt ornare massa eget egestas purus.
-Venenatis tellus in metus vulputate eu scelerisque felis imperdiet.
-Posuere urna nec tincidunt praesent.
-Risus pretium quam vulputate dignissim suspendisse in.
-
-Aliquam id diam maecenas ultricies mi eget.
-Mauris pharetra et ultrices neque ornare aenean euismod.
-Vitae suscipit tellus mauris a diam maecenas.
-Massa massa ultricies mi quis hendrerit.
-Tincidunt eget nullam non nisi est sit amet.
-Vitae purus faucibus ornare suspendisse sed nisi.
-Enim sit amet venenatis urna.
-Malesuada fames ac turpis egestas.
-Viverra nam libero justo laoreet sit amet.
-Ornare massa eget egestas purus viverra accumsan in nisl nisi.
-Ac feugiat sed lectus vestibulum mattis.
-
-\autocite{Schlund2020}
-\autocite{Schlund2020a}
+\begin{otherlanguage}{ngerman}
+  TBA.
+\end{otherlanguage}

### 2020-11-14 ch01_introduction/introduction.tex
Initialized chapters

--- 2020-11-13 ch1_introduction/introduction.tex
@@
 %=============================================================================%
-%                              01 - Introduction                              %
+%                                Introduction                                 %
 %=============================================================================%
@@
 \chapter{Introduction}
+\label{ch:introduction}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Hendrerit gravida rutrum quisque non tellus orci ac auctor augue.
-Eros donec ac odio tempor orci dapibus.
-Ultrices neque ornare aenean euismod elementum nisi.
-Vivamus arcu felis bibendum ut tristique et egestas quis ipsum.
-Elit ut aliquam purus sit amet luctus.
-Et malesuada fames ac turpis egestas sed.
-Egestas dui id ornare arcu odio ut sem nulla pharetra.
-Porta lorem mollis aliquam ut porttitor leo a diam.
-Ultricies tristique nulla aliquet enim.
-Urna nunc id cursus metus aliquam.
-Faucibus a pellentesque sit amet porttitor.
-Tincidunt ornare massa eget egestas purus.
-Venenatis tellus in metus vulputate eu scelerisque felis imperdiet.
-Posuere urna nec tincidunt praesent.
-Risus pretium quam vulputate dignissim suspendisse in.
-
-Aliquam id diam maecenas ultricies mi eget.
-Mauris pharetra et ultrices neque ornare aenean euismod.
-Vitae suscipit tellus mauris a diam maecenas.
-Massa massa ultricies mi quis hendrerit.
-Tincidunt eget nullam non nisi est sit amet.
-Vitae purus faucibus ornare suspendisse sed nisi.
-Enim sit amet venenatis urna.
-Malesuada fames ac turpis egestas.
-Viverra nam libero justo laoreet sit amet.
-Ornare massa eget egestas purus viverra accumsan in nisl nisi.
-Ac feugiat sed lectus vestibulum mattis.
-
-\autocite{Schlund2020}
-\autocite{Schlund2020a}
+The \ac{ECS} is really cool. I like it very much!

### 2020-11-14 end_acronyms/acronyms.tex
Initialized chapters

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                  Acronyms                                   %
%=============================================================================%

\chapter{List of Acronyms}

\begin{acronym}
  \acro{ECS}{Effective Climate Sensitivity}
  \acro{EqCS}{Equilibrium Climate Sensitivity}
  \acro{GPP}{Gross Primary Productivity}
\end{acronym}

### 2020-11-14 end_appendix/appendix.tex
Initialized chapters

@@
 
+\chapter{Appendix}
+
 \renewcommand{\thesection}{\Alph{section}}
-\addchap{Appendix}
 
-\section{Circle of fifths metric: Proof of the metric properties}
-\label{sec:app:cof_metric_proof}
+\section{TBA}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Hendrerit gravida rutrum quisque non tellus orci ac auctor augue.
-Eros donec ac odio tempor orci dapibus.
-Ultrices neque ornare aenean euismod elementum nisi.
-Vivamus arcu felis bibendum ut tristique et egestas quis ipsum.
-Elit ut aliquam purus sit amet luctus.
-Et malesuada fames ac turpis egestas sed.
-Egestas dui id ornare arcu odio ut sem nulla pharetra.
-Porta lorem mollis aliquam ut porttitor leo a diam.
-Ultricies tristique nulla aliquet enim.
-Urna nunc id cursus metus aliquam.
-Faucibus a pellentesque sit amet porttitor.
-Tincidunt ornare massa eget egestas purus.
-Venenatis tellus in metus vulputate eu scelerisque felis imperdiet.
-Posuere urna nec tincidunt praesent.
-Risus pretium quam vulputate dignissim suspendisse in.
+TBA.
 
-\section{Source code}
+\section{TBA}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Hendrerit gravida rutrum quisque non tellus orci ac auctor augue.
-Eros donec ac odio tempor orci dapibus.
-Ultrices neque ornare aenean euismod elementum nisi.
-Vivamus arcu felis bibendum ut tristique et egestas quis ipsum.
-Elit ut aliquam purus sit amet luctus.
-Et malesuada fames ac turpis egestas sed.
-Egestas dui id ornare arcu odio ut sem nulla pharetra.
-Porta lorem mollis aliquam ut porttitor leo a diam.
-Ultricies tristique nulla aliquet enim.
-Urna nunc id cursus metus aliquam.
-Faucibus a pellentesque sit amet porttitor.
-Tincidunt ornare massa eget egestas purus.
-Venenatis tellus in metus vulputate eu scelerisque felis imperdiet.
-Posuere urna nec tincidunt praesent.
-Risus pretium quam vulputate dignissim suspendisse in.
+TBA.

### 2020-11-15 end_appendix/appendix.tex
Optimized style

@@
 
-TBA.
+\subsection{test}
+
+\subsubsection{test}
+
+\paragraph{hi}
+
+The \ac{ECS} is really cool. I like it very much!
+
+This is e.g. without an "at" and this is it with an "at" e.g.\@ difference? Test space. Real dot!
+
+These are really cool papers: \autocite{Schlund2020, Schlund2020a}
+
+And this one, too: \autocite{Lauer2020}
+
+\texttt{input <iostream>}
+
+\begin{equation}
+  c_{k_1,k_2} := 1200 \, \log_2 \left( \frac{f_1^{(k_2)}}{f_1^{(k_1)}} \right) \unit{cents}. \label{eq:2:cent_def}
+\end{equation}
+
+\begin{table}
+  \caption{The effects of treatments X and Y on the four groups studied.}
+  \label{tab:treatments}
+  \centering
+  \begin{tabular}{l l l}
+    \toprule
+    \tabhead{Groups} & \tabhead{Treatment X} & \tabhead{Treatment Y} \\
+    \midrule
+    1 & 0.2 & 0.8\\
+    2 & 0.17 & 0.7\\
+    3 & 0.24 & 0.75\\
+    4 & 0.68 & 0.3\\
+    \bottomrule\\
+  \end{tabular}
+\end{table}
+
+\begin{table}[t]
+  \centering
+  \begin{tabular}{c l r r}
+    \toprule
+    Semitones & Interval & $c/\unit{cents}$ (ET) & $c/\unit{cents}$ (JI) \\
+    \midrule
+    $0$ & Perfect unison & $0$ & $0$ \\
+    $1$ & Minor second & $100$ & $112$ \\
+    $2$ & Major second & $200$ & $204$ \\
+    $3$ & Minor third & $300$ & $316$ \\
+    $4$ & Major third & $400$ & $386$ \\
+    $5$ & Perfect fourth & $500$ & $498$ \\
+    $6$ & Augmented fourth & $600$ & $590$ \\
+    $7$ & Perfect fifth & $700$ & $702$ \\
+    $8$ & Minor sixth & $800$ & $814$ \\
+    $9$ & Major sixth & $900$ & $884$ \\
+    $10$ & Minor seventh & $1000$ & $996$ \\
+    $11$ & Major seventh & $1100$ & $1088$ \\
+    $12$ & Perfect octave & $1200$ & $1200$ \\
+    \bottomrule
+  \end{tabular}
+  \caption{Logarithmic frequency ratios $c$ of certain intervals in the equal temperament (ET) and the just intonation (JI). $x$ cents correspond to a frequency ratio of $2^{\, x/1200}$.}
+  \label{tab:2:cents}
+\end{table}
 

### 2020-11-17 ch01_introduction/introduction.tex
Optimized style
Wrote first page of introduction

@@
 
-The \ac{ECS} is really cool. I like it very much!
+\section{Structure of the thesis}
+\label{sec:introduction:structure}
+
+Parts of this thesis have already been published in scientific publications,
+which will be clearly stated at the beginning of each chapter.
+\Cref{ch:scientific_background} introduces the scientific background for this
+thesis. This includes basic principles of Earth System Modeling, sources of
+uncertainty in future projections of the climate, relevant definitions and
+state-of-the-art methods used to evaluate \ac{ESM} simulations and reduce
+associated uncertainties. \Cref{ch:esmvaltool} gives an overview over the
+contributions made to the \ac{ESMValTool}, an open-source software for the
+analysis of \acp{ESM}. These contributions helped improving the routine
+evaluation of \acp{ESM} which is useful for the whole scientific community and
+lead to co-authorship in four peer-reviewed studies
+\autocite{Righi2020, Eyring2020, Lauer2020, Weigel2020}.
+\Cref{ch:assessment_climate_metrics} covers the assessment of policy-relevant
+climate metrics like the \acl{EqCS} and the \acl{TCR} in the latest generation
+of \acp{ESM}. This work has already been published in two scientific
+publications \autocite{Bock2020, Meehl2020}. \Cref{ch:paper_ecs} describes the
+evaluation of emergent constraints (a technique to reduce uncertainties in
+climate model projections, see \vref{sec:scientific_background:techniques}) on
+the climate sensitivity in \acs{CMIP6} models. The contents of this chapter
+have been published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.
+\Cref{ch:paper_gpp} focuses on a new method to reduce climate model
+uncertainties based on \ac{ML}. As an example, the method is applied to the
+photosynthesis rate at the end of the \tfst{} century, which is already
+published in the \emph{Journal of Geophysical Research: Biogeosciences}
+\autocite{Schlund2020}. Finally, \cref{ch:summary_outlook} provides a summary
+of the results of this thesis and gives an outlook into possible future works.

### 2020-11-17 end_acronyms/acronyms.tex
Wrote first page of introduction

@@
 \begin{acronym}
+  \acro{CMIP}{Coupled Model Intercomparison Project}
+  \acro{CMIP3}{Phase 3 of the Coupled Model Intercomparison Project}
+  \acro{CMIP5}{Phase 5 of the Coupled Model Intercomparison Project}
+  \acro{CMIP6}{Phase 6 of the Coupled Model Intercomparison Project}
+  \acro{DLR}{
+    Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
   \acro{ECS}{Effective Climate Sensitivity}
+  \acro{ESM}{Earth System Model}
+  \acro{ESMValTool}{Earth System Model Evaluation Tool}
   \acro{EqCS}{Equilibrium Climate Sensitivity}
   \acro{GPP}{Gross Primary Productivity}
+  \acro{IPA}{
+    Institut für Physik der Atmosphäre (Institute of Atmospheric Physics)}
+  \acro{MIP}{Model Intercomparison Project}
+  \acro{ML}{Machine Learning}
+  \acro{TCR}{Transient Climate Response}
 \end{acronym}

### 2020-11-18 ch01_introduction/introduction.tex
Added first figure

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 \chapter{Introduction}
-\label{ch:introduction}
+\label{ch:01:introduction}
 
-The \ac{ECS} is really cool. I like it very much!
+\section{Structure of the thesis}
+\label{sec:01:structure}
+
+Parts of this thesis are published in multiple scientific publications (two as
+first author six as co-author). If applicable, this is clearly stated at the
+beginning of each chapter. \Cref{ch:02:scientific_background} introduces the
+scientific background for this thesis. This includes basic principles of Earth
+System Modeling, sources of uncertainty in future projections of the climate,
+relevant definitions and state-of-the-art methods used to evaluate \ac{ESM}
+simulations and reduce associated uncertainties. \Cref{ch:03:esmvaltool} gives
+an overview over the contributions made to the \ac{ESMValTool}, an open-source
+software for the analysis of \acp{ESM}. These contributions helped improving
+the routine evaluation of \acp{ESM} which is useful for the whole scientific
+community and lead to co-authorship in four peer-reviewed studies
+\autocite{Righi2020, Eyring2020, Lauer2020, Weigel2020}.
+\Cref{ch:04:assessment_climate_metrics} covers the assessment of
+policy-relevant climate metrics like the \ac{EqCS} and the \ac{TCR} in the
+latest generation of \acp{ESM}. This work is already published in two
+scientific publications \autocite{Bock2020, Meehl2020}. Since the \ac{EqCS} and
+\ac{TCR} are considerably higher in this new climate model generation,
+\cref{ch:05:paper_ecs} describes the assessment of emergent constraints (a
+technique to reduce uncertainties in climate model projections, see
+\vref{sec:02:techniques}) on the \ac{EqCS} for these \acp{ESM}. The contents of
+this chapter are published in \emph{Earth System Dynamics}
+\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
+reduce climate model uncertainties based on \ac{ML}. As an example, the method
+is applied to the photosynthesis rate at the end of the \nth{21} century, which
+is already published in the \emph{Journal of Geophysical Research:
+Biogeosciences} \autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook}
+provides a summary of the results of this thesis and gives an outlook of
+possible future works.

### 2020-11-18 ch02_scientific_background/scientific_background.tex
Initialized chapters
Wrote first page of introduction
Added first figure

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                            Scientific Background                            %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Scientific Background}
+\label{ch:02:scientific_background}
 
-The \ac{ECS} is really cool. I like it very much!
+\section{Earth System Models: Simulations and Analysis}
+\label{sec:02:esms}
+
+An \ac{ESM} is really cool! Yes, \acp{ESM}!!
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\figureWidth{}]{
+    ch02_scientific_background/figs/esms_historical_evolution.pdf}
+  \caption{Historical evolution of coupled climate models over the last 45
+  years. In early days, these models were \acfp{AOGCM} and only included three
+  components: the atmosphere, the land surface and the ocean. Over the time,
+  the individual components grew in complexity and included a wider range of
+  processes (illustrated by the growing cylinders). Eventually, more and more
+  components (aerosols, carbon cycle, \etc{}) were added to the coupled system,
+  forming the modern \acfp{ESM}. Taken from \textcite{Cubasch2013}.}
+  \label{fig:02:esms_historical_evolution}
+\end{figure}
+
+TBA.
+
+\section{Techniques to reduce uncertainties in climate model projections}
+\label{sec:02:techniques}

### 2020-11-18 end_acronyms/acronyms.tex
Added first figure

@@
 \begin{acronym}
+  \acro{AOGCM}{Atmosphere-Ocean General Circulation Model}
   \acro{CMIP}{Coupled Model Intercomparison Project}
@@
   \acro{CMIP6}{Phase 6 of the Coupled Model Intercomparison Project}
-  \acro{DLR}{
+  \acro{DLR}{%
     Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
@@
   \acro{GPP}{Gross Primary Productivity}
-  \acro{IPA}{
+  \acro{IPA}{%
     Institut für Physik der Atmosphäre (Institute of Atmospheric Physics)}

### 2020-11-20 ch01_introduction/introduction.tex
Added further text and figures to ch.2

@@
 
-Parts of this thesis are published in multiple scientific publications (two as
-first author six as co-author). If applicable, this is clearly stated at the
-beginning of each chapter. \Cref{ch:02:scientific_background} introduces the
-scientific background for this thesis. This includes basic principles of Earth
-System Modeling, sources of uncertainty in future projections of the climate,
-relevant definitions and state-of-the-art methods used to evaluate \ac{ESM}
-simulations and reduce associated uncertainties. \Cref{ch:03:esmvaltool} gives
-an overview over the contributions made to the \ac{ESMValTool}, an open-source
-software for the analysis of \acp{ESM}. These contributions helped improving
-the routine evaluation of \acp{ESM} which is useful for the whole scientific
-community and lead to co-authorship in four peer-reviewed studies
+Parts of this thesis are published in multiple peer-reviewed publications (two
+first-author studies and six co-author studies). If applicable, this is clearly
+stated at the beginning of each chapter. \Cref{ch:02:scientific_background}
+introduces the scientific background for this thesis. This includes relevant
+literature that is used as a baseline for this thesis. \Cref{ch:03:esmvaltool}
+gives an overview over the contributions made to the \ac{ESMValTool}, an
+open-source software for the analysis of \acp{ESM}. These contributions helped
+improving the routine evaluation of \acp{ESM} which is useful for the whole
+scientific community and lead to co-authorship in four peer-reviewed studies
 \autocite{Righi2020, Eyring2020, Lauer2020, Weigel2020}.
 \Cref{ch:04:assessment_climate_metrics} covers the assessment of
-policy-relevant climate metrics like the \ac{EqCS} and the \ac{TCR} in the
+policy-relevant climate metrics like the \ac{ECS} and the \ac{TCR} in the
 latest generation of \acp{ESM}. This work is already published in two
-scientific publications \autocite{Bock2020, Meehl2020}. Since the \ac{EqCS} and
+scientific publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and
 \ac{TCR} are considerably higher in this new climate model generation,
@@
 technique to reduce uncertainties in climate model projections, see
-\vref{sec:02:techniques}) on the \ac{EqCS} for these \acp{ESM}. The contents of
+\vref{sec:02:techniques}) on the \ac{ECS} for these \acp{ESM}. The contents of
 this chapter are published in \emph{Earth System Dynamics}

### 2020-11-20 ch02_scientific_background/scientific_background.tex
Added further text and figures to ch.2

@@
 
-\section{Earth System Models: Simulations and Analysis}
+This chapter introduces the scientific background necessary for the work
+presented in this thesis. After a brief introduction to climate model
+simulations and associated uncertainties, we present state-of-the art
+techniques used to evaluate climate model simulations and reduce uncertainties
+in projections of the future climate. These methods form the basis for the new
+techniques developed in this thesis.
+
+\section{\aclp{ESM}: Simulations and Analysis}
 \label{sec:02:esms}
 
-An \ac{ESM} is really cool! Yes, \acp{ESM}!!
+In contrast to other fields of science, researching the future evolution of the
+Earth's climate cannot be purely done by performing experiments in a
+laboratory. Due to the immense complexity of the Earth system (including
+physical, biological and chemical processes on various temporal and spatial
+scales), we do not have access to a mini version of the Earth that we can
+expose to varying external conditions and analyze its response to it
+\autocite{Flato2011}. While observing the current state of the Earth System is
+(relatively) straightforward, gaining evidence about the future climate by only
+considering present-day observations is rather difficult. A possible way out is
+given by numerical climate models, which offer the possibility to simulate the
+Earth's climate on a computer. The first numerical climate models came up in
+the 1960s and were based on weather prediction models \autocite{Flato2011}.
+Early models from the 1970s simulated only the physical components of the
+climate system: atmosphere, land surface, ocean and sea ice (see
+\cref{fig:02:esms_historical_evolution}). The main aim of these so-called
+\acp{AOGCM} \autocite{Flato2013} is to numerically solve the differential
+equations that describe the exchange of energy and matter between these
+physical components.
 
@@
   \caption{Historical evolution of coupled climate models over the last 45
-  years. In early days, these models were \acfp{AOGCM} and only included three
-  components: the atmosphere, the land surface and the ocean. Over the time,
-  the individual components grew in complexity and included a wider range of
-  processes (illustrated by the growing cylinders). Eventually, more and more
-  components (aerosols, carbon cycle, \etc{}) were added to the coupled system,
-  forming the modern \acfp{ESM}. Taken from \textcite{Cubasch2013}.}
+    years. In early days, these models were so-called \acfp{AOGCM} and only
+    included three components: the atmosphere, the land surface and the ocean.
+    Over the time, the individual components grew in complexity and included a
+    wider range of processes (illustrated by the growing cylinders). Eventually,
+    more and more components (aerosols, carbon cycle, \etc{}) were added to the
+    coupled system, forming the modern \acfp{ESM}. Taken from
+    \textcite{Cubasch2013}.}
   \label{fig:02:esms_historical_evolution}
@@
 
-TBA.
+Over the course of the years, climate models became more and more complex by
+including a wider range of processes within the components, but also by
+introducing new components to the coupled system. Examples of these are
+aerosols, the carbon cycle, a dynamic vegetation, atmospheric chemistry and
+land ice (see \cref{fig:02:esms_historical_evolution}). \acp{AOGCM} coupled to
+these additional components are called \acfp{ESM}, which are the current
+state-of-the-art models that allow the most sophisticated simulations of the
+Earth's climate. In contrast to \acp{AOGCM}, \acp{ESM} enable the simulation of
+biological and chemical processes in addition to the dynamics of the physical
+components of the Earth system. Especially in the context of anthropogenic
+climate change, these additional processes are of uttermost importance for
+realistic climate model simulations, since the anthropogenic interference with
+the Earth system directly influences the various biogeochemical cycles of the
+Earth. For example, the emission of the most prominent \ac{GHG}, \ac{CO2},
+immediately impacts the global carbon cycle by introducing an additional carbon
+source to the cycle (for details see \cref{sec:02:carbon_cycle}). Due to the
+global carbon cycle, only about \pct{50} of the emitted \ac{CO2} remains in the
+atmosphere, where it can act as \ac{GHG} by introducing an additional radiative
+forcing to the Earth System, which eventually leads to increasing surface
+temperatures. Thus, this uptake of \ac{CO2} by the global carbon cycle slows
+down global warming. Further examples include land use changes like the
+deforestation of tropical rainforests, which also directly influences several
+biogeochemical cycles (\eg{} carbon cycle, nitrogen cycle, phosphorus cycle,
+\etc{}) by altering respective sinks and sources.
 
@@
 \label{sec:02:techniques}
+
+\section{The global Carbon Cycle}
+\label{sec:02:carbon_cycle}

### 2020-11-27 ch02_scientific_background/scientific_background.tex
Added section 2.2

--- 2020-11-18 ch02_scientific_background/scientific_background.tex
@@
 
-\section{Earth System Models: Simulations and Analysis}
+This chapter introduces the scientific background of this thesis. First, basic
+concepts of climate model simulations and associated uncertainties are
+introduced. Next, the fundamental biogeochemical processes of the global carbon
+cycle and important metrics describing climate change are presented. Finally,
+state-of-the art techniques used reduce uncertainties in projections of the
+future climate are shown. These methods form the basis for the new techniques
+developed in this thesis.
+
+\section{\aclp{ESM}: Simulations and Analysis}
 \label{sec:02:esms}
 
-An \ac{ESM} is really cool! Yes, \acp{ESM}!!
+\subsection{Numerical Climate Modeling}
+\label{subsec:02:climate_modeling}
+
+In contrast to other fields of science, researching the future evolution of the
+Earth's climate cannot be purely done by performing experiments in a
+laboratory. Due to the immense complexity of the Earth system (including
+physical, biological and chemical processes on various temporal and spatial
+scales and their mutual interactions), we do not have access to a tiny replica
+of the Earth that we can analyze when exposed to different external conditions
+\autocite{Flato2011}. While observing the current state of the Earth System is
+(relatively) straightforward, gaining evidence about the future evolution of
+the climate by only considering present-day observations is rather difficult.
+
+A possible way out is given by numerical climate models, which offer the
+possibility to simulate the Earth's climate on a computer. The first numerical
+climate models came up in the 1960s and were based on weather prediction models
+\autocite{Flato2011}. Early models from the 1970s simulated only the physical
+components of the climate system: atmosphere, land surface, ocean and sea ice
+(see \cref{fig:02:esms_historical_evolution}). The basis of these so-called
+\acp{AOGCM} \autocite{Flato2013} is the numerical solving of the differential
+equations describing the exchange of energy and matter between these physical
+components.
 
@@
   \caption{Historical evolution of coupled climate models over the last 45
-  years. In early days, these models were \acfp{AOGCM} and only included three
-  components: the atmosphere, the land surface and the ocean. Over the time,
-  the individual components grew in complexity and included a wider range of
-  processes (illustrated by the growing cylinders). Eventually, more and more
-  components (aerosols, carbon cycle, \etc{}) were added to the coupled system,
-  forming the modern \acfp{ESM}. Taken from \textcite{Cubasch2013}.}
+    years. In early days, these models were so-called \acfp{AOGCM} and only
+    included three components: the atmosphere, the land surface and the ocean.
+    Over the time, the individual components grew in complexity and included a
+    wider range of processes (illustrated by the growing cylinders).
+    Eventually, more and more components (aerosols, carbon cycle, \etc{}) were
+    added to the coupled system, forming the modern \acfp{ESM}. Taken from
+    \textcite{Cubasch2013}.}
   \label{fig:02:esms_historical_evolution}
@@
 
-TBA.
+Over the course of the years, climate models became more and more complex by
+including a wider range of processes within the components, but also by
+introducing new components to the coupled system. Examples of these are
+aerosols, the carbon cycle, a dynamic vegetation, atmospheric chemistry and
+land ice (see \cref{fig:02:esms_historical_evolution}). \acp{AOGCM} coupled to
+these additional components are called \acfp{ESM}, which are the current
+state-of-the-art models that allow the most sophisticated simulations of the
+Earth's climate. In contrast to \acp{AOGCM}, \acp{ESM} enable the simulation of
+biological and chemical processes in addition to the dynamics of the physical
+components of the Earth system. Especially in the context of anthropogenic
+climate change, these additional processes are of uttermost importance for
+realistic climate model simulations, since the anthropogenic interference with
+the Earth system directly influences the various biogeochemical cycles of the
+Earth. For example, the emission of the most prominent \ac{GHG}, \ac{CO2},
+immediately impacts the global carbon cycle by inserting additional carbon into
+the system (for details see \cref{sec:02:carbon_cycle}). Further examples
+include land use changes like the deforestation of tropical rainforests, which
+also directly influences several biogeochemical cycles (\eg{} carbon cycle,
+nitrogen cycle, phosphorus cycle, \etc{}) by altering respective sinks and
+sources.
 
-\section{Techniques to reduce uncertainties in climate model projections}
+Due to the complex interactions between the different components of the Earth
+system, these changes in the biogeochemical processes also affect the physical
+properties of the climate system. For example, due to the global carbon cycle,
+only about \pct{50} of the emitted \ac{CO2} by humankind remains in the
+atmosphere \autocite{Friedlingstein2019}. The residual part is absorbed by the
+two main carbon sinks of the planet, the terrestrial biosphere and the ocean.
+Since only atmospheric \ac{CO2} can act as \ac{GHG} by introducing an
+additional radiative forcing to the Earth System leading to increasing surface
+temperatures, this uptake of \ac{CO2} by the carbon cycle slows down global
+warming.
+
+\subsection{\acs{CMIP}}
+\label{subsec:02:cmip}
+
+Due to the complex nature of the Earth system itself, numerical models of it
+consist of hundreds of thousands of lines of computer code. Thus, a
+standardization to a certain degree is crucial for the various research groups
+developing \acp{ESM} all around the world in order to obtain comparable output
+and to facilitate analysis. For this reason, the \ac{WGCM} of the \ac{WCRP}
+initiated the \ac{CMIP} in 1995, with the objective to \enquote{better
+understand past, present and future climate changes arising from natural,
+unforced variability or in response to changes in radiative forcing in a
+multi-model context} \autocite{WCRP2020}. One major element of \ac{CMIP} is to
+establish common standards, coordination, infrastructure, and documentation in
+order to facilitate the distribution of climate model output
+\autocite{Eyring2016}.
+
+A further main aspect is to provide a set of standardized experiments for
+global climate model simulations. To participate in the latest phase of
+\ac{CMIP}, \acs{CMIP}6, climate models need to run a \emph{historical}
+simulation of the period \range{1850}{2014} and the so-called \ac{DECK}
+experiments, which include a pre-industrial control run (\emph{piControl}), a
+historical Atmospheric \ac{MIP} simulation (\emph{amip}), a simulation forced
+with an abrupt quadrupling of \ac{CO2} (\emph{abrupt-4xCO2}) and a simulation
+forced with a \pct{1} per year increase of the atmospheric \ac{CO2}
+concentration (\emph{1pctCO2}) \autocite{Eyring2016}. This is shown in the
+center of \cref{fig:02:cmip6}, which illustrates the experimental design of
+\acs{CMIP}6.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\figureWidth{}]{
+    ch02_scientific_background/figs/cmip6.pdf}
+  \caption{Schematic of the experiment design of Phase 6 of the \acl{CMIP}
+    (\acs{CMIP}6). The center of the circle illustrates the four \acs{DECK}
+    (\acl{DECK}) experiments and the \acs{CMIP}6 historical simulation. The
+    circular sectors show additional science themes that can be explored
+    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}. Taken from
+    \textcite{Simpkins2017}.}
+  \label{fig:02:cmip6}
+\end{figure}
+
+To increase diversity and answer more scientific questions, \acs{CMIP}6 models
+can participate in the so-called \acs{CMIP}6-Endorsed \acp{MIP}, of which
+\acs{CMIP}6 offers 21 (see circular sectors in \cref{fig:02:cmip6}). Some
+\acp{MIP} offer additional experiments to explore specific aspects of the Earth
+system, like the \ac{C4MIP} which focuses on the carbon cycle
+\autocite{Jones2016} or the \ac{CFMIP} which focuses on the evaluation of cloud
+feedbacks \autocite{Webb2017}. Other \acp{MIP} allow the assessment of future
+climate change. An example is the \ac{ScenarioMIP}, which provides common
+experiments that simulate different possible futures \autocite{ONeill2016}.
+These experiments are based on the so-called \acp{SSP}, a set of alternative
+pathways of future societal development \autocite{ONeill2013, ONeill2017}. For
+each experiment, a set of emissions and land use changes is calculated from the
+\acp{SSP} \autocite{Riahi2017} which are then used to force the global climate
+models. For \ac{ScenarioMIP}, five different \acp{SSP} are considered, ranging
+from \ac{SSP}1 (sustainability) to \ac{SSP}5 (fossil-fuel development). Each
+\ac{SSP} is combined with a climate outcome (measured as radiative forcing in
+the year 2100) based on a particular forcing pathway that \acp{IAM} have shown
+to be feasible. For example, \ac{SSP}5-8.5 represents a scenario based on a
+fossil-fuel development with a radiative forcing of $8.5 \unit{W m^{-2}}$ in
+2100 while \ac{SSP}1-2.6 represents a sustainable future with a radiative
+forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two other main scenarios
+(called \enquote{Tier 1} experiments in \ac{ScenarioMIP}) are the \ac{SSP}2-4.5
+and \ac{SSP}3-7.0 scenarios. In contrast to the \ac{ScenarioMIP} experiments,
+the corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012}, the so-called
+\ac{RCP}, only used the radiative forcing in 2100 as only dimension to describe
+the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5, \ac{RCP}2.6, \etc{}).
+
+In this thesis, climate model data from the two most recent \ac{CMIP}
+generations is used, \acs{CMIP}5 and \acs{CMIP}6. More detailed information
+about the specific variables and experiments analyzed is given in the
+corresponding chapters.
+
+\subsection{Sources of Uncertainties in Climate Model Projections}
+\label{subsec:02:source_of_uncertainties}
+
+\section{The Global Carbon Cycle}
+\label{sec:02:carbon_cycle}
+
+I like \ac{GPP}!
+
+\section{Techniques to reduce Uncertainties in Climate Model Projections}
 \label{sec:02:techniques}

### 2020-11-27 end_acronyms/acronyms.tex
Added further text and figures to ch.2
Added section 2.2

@@
 
+\begingroup
+\setstretch{0.5}
 \begin{acronym}
   \acro{AOGCM}{Atmosphere-Ocean General Circulation Model}
+  \acro{AR}{Assessment Report}
+  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
+  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
   \acro{CMIP}{Coupled Model Intercomparison Project}
-  \acro{CMIP3}{Phase 3 of the Coupled Model Intercomparison Project}
-  \acro{CMIP5}{Phase 5 of the Coupled Model Intercomparison Project}
-  \acro{CMIP6}{Phase 6 of the Coupled Model Intercomparison Project}
+  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
+  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
   \acro{DLR}{%
     Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
-  \acro{ECS}{Effective Climate Sensitivity}
+  \acro{ECS}{Equilibrium Climate Sensitivity}
   \acro{ESM}{Earth System Model}
   \acro{ESMValTool}{Earth System Model Evaluation Tool}
-  \acro{EqCS}{Equilibrium Climate Sensitivity}
+  \acro{GHG}{Greenhouse Gas}
   \acro{GPP}{Gross Primary Productivity}
+  \acro{IAM}{Integrated Assessment Models}
   \acro{IPA}{%
     Institut für Physik der Atmosphäre (Institute of Atmospheric Physics)}
+  \acro{IPCC}{Intergovernmental Panel on Climate Change}
   \acro{MIP}{Model Intercomparison Project}
   \acro{ML}{Machine Learning}
+  \acro{RCP}{Representative Concentration Pathway}
+  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
+  \acro{SSP}{Shared Socioeconomic Pathway}
   \acro{TCR}{Transient Climate Response}
+  \acro{WCRP}{World Climate Research Programme}
+  \acro{WGCM}{Working Group on Coupled Modelling}
 \end{acronym}
+\endgroup

### 2020-12-01 ch02_scientific_background/scientific_background.tex
Added first part of carbon cycle section

@@
 developing \acp{ESM} all around the world in order to obtain comparable output
-and to facilitate analysis. For this reason, the \ac{WGCM} of the \ac{WCRP}
-initiated the \ac{CMIP} in 1995, with the objective to \enquote{better
-understand past, present and future climate changes arising from natural,
-unforced variability or in response to changes in radiative forcing in a
-multi-model context} \autocite{WCRP2020}. One major element of \ac{CMIP} is to
-establish common standards, coordination, infrastructure, and documentation in
-order to facilitate the distribution of climate model output
+and to facilitate model analysis. For this reason, the \ac{WGCM} of the
+\ac{WCRP} initiated the \ac{CMIP} in 1995, with the objective to
+\enquote{better understand past, present and future climate changes arising
+  from natural, unforced variability or in response to changes in radiative
+  forcing in a multi-model context} \autocite{WCRP2020}. One major element of
+\ac{CMIP} is to establish common standards, coordination, infrastructure, and
+documentation in order to facilitate the distribution of climate model output
 \autocite{Eyring2016}.
@@
 the corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012}, the so-called
-\ac{RCP}, only used the radiative forcing in 2100 as only dimension to describe
-the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5, \ac{RCP}2.6, \etc{}).
+\acp{RCP}, only used the radiative forcing in 2100 as only dimension to
+describe the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5, \ac{RCP}2.6,
+\etc{}).
 
@@
 
+- anthropogenic forcing vs. natural forcing (volcanoes, solar cycle, etc.)
+
+- basic physics of greenhouse gases (vibration modes, etc.)
+
+- historical vs. future: in historical (future), uncertainty in forcing low
+(high)
+
+- definition of feedbacks, examples
+
+\section{Climate Sensitivity}
+\label{sec:02:climate_sensitivity}
+
+- Physical Feedbacks (ECS)
+
+- Equations for feedbacks + decomposition
+
+- ocean heat uptake (TCR)
+
+- different between effective climate sensitivity and equilibrium version
+
+- equations for ECS and TCR
+
 \section{The Global Carbon Cycle}
@@
 
-I like \ac{GPP}!
+Since one study presented in this thesis \autocite[see
+\cref{ch:06:paper_gpp}]{Schlund2020} aims to reduce uncertainties in carbon
+cycle-related processes, this chapter introduces the scientific background of
+the global carbon cycle.
+
+\subsection{Overview}
+\label{subsec:02:carbon_cycle_overview}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\figureWidth{}]{
+    ch02_scientific_background/figs/carbon_cycle_schematic.png}
+  \caption{Simplified schematic of the global carbon cycle including the
+    typical turnover time scales for carbon transfers through the major
+    reservoirs (atmosphere, land surface and ocean). Taken from
+    \textcite{Ciais2013}.}
+  \label{fig:02:carbon_cycle_schematic}
+\end{figure}
+
+A schematic overview of the global carbon cycle is shown in
+\cref{fig:02:carbon_cycle_schematic}. To quantify the carbon cycle, common
+units are \ac{ppm} for the atmospheric trace gas concentrations (dry-air mole
+fraction) and \ac{GtC} or $\si{\GtCyr{}}$ for the reservoirs masses or exchange
+fluxes, respectively. The carbon exchange processes between the different
+carbon reservoirs run on a wide range of time scales. Conceptually, one can
+distinguish between two domains of the global carbon cycle: a slow and a fast
+domain. The slow domain with turnover times (reservoir mass of carbon divided
+by exchange flux) of more than 10000 years consists of the large carbon stores
+in rocks and sediments which are connected to the rapid domain of the carbon
+cycle through volcanic emissions of \ac{CO2}, chemical weathering, erosion and
+sediment formation on the sea floor. These natural exchange fluxes between the
+slow and the fast domain are comparatively small ($< 0.3 \unit{\GtCyr{}}$) and
+can be assumed as approximately constant in time over the last few centuries
+\autocite{Ciais2013}.
+
+The fast domain of the global carbon cycle consists of three main carbon
+reservoirs: the atmosphere, the terrestrial biosphere and the ocean. In the
+atmosphere, carbon is mainly stored in trace gases, with \ac{CO2} as the major
+component with a current (2019) concentration of about $410 \unit{ppm}$
+\autocite{Friedlingstein2019}. Additional contributors to the atmospheric
+carbon content are the trace gas \ac{CH4}, the trace gas \ac{CO}, hydrocarbons,
+black carbon aerosols and organic compounds \autocite{Ciais2013}. Carbon in the
+terrestrial biosphere is mainly stored as organic compounds, with about
+$\range{450}{650} \unit{GtC}$ in the living vegetation biomass,
+$\range{1500}{2400} \unit{GtC}$ in dead organic matter in litter and soils and
+about $1700 \unit{GtC}$ in permafrost soils \autocite{Ciais2013}. The main
+component of the oceanic carbon reservoir is dissolved inorganic carbon
+(carbonic acid, bicarbonate ions and carbonate ions) with about $38000
+\unit{GtC}$. Further carbon is stored as dissolved organic carbon (about $700
+\unit{GtC}$), in surface sediments (about $1750 \unit{GtC}$) and in marine
+biota (about $3 \unit{GtC}$, predominantly phytoplankton and other
+microorganisms) \autocite{Ciais2013, Friedlingstein2019}.
+
+In the fast domain of the global carbon cycle, reservoir turnover times range
+from seconds to millennia. In contrast to the slow domain, the carbon exchange
+fluxes within the fast domain of the carbon cycle are much higher. One major
+group of exchange processes in the fast domain connects the atmosphere and the
+terrestrial biosphere. \ac{CO2} is removed from the atmosphere by plant
+photosynthesis with about $120 \unit{\GtCyr{}}$ \autocite{Ciais2013}. This
+process is also known as \ac{GPP}. The carbon fixed into plants can be released
+back into the atmosphere by autotrophic (plant) and heterotrophic (soil
+microbial and animal) respiration and additional disturbance processes like
+fires \autocite{Ciais2013}. Since the land \ac{CO2} uptake by photosynthesis
+occurs only during the growing season, whereas respiration occurs nearly all
+year, the larger amount of vegetation in the Northern hemisphere (due to the
+larger land mass) gives rise to a seasonal cycle of the atmospheric \ac{CO2}
+concentration \autocite{Keeling1995}. This seasonal cycle reflects the phase of
+the global carbon cycle and shows a maximum of the atmospheric \ac{CO2}
+concentration in the Northern hemisphere winter (net \ac{CO2} flux into
+atmosphere due to respiration) and a minimum during the Northern hemisphere
+summer (net \ac{CO2} flux into the land due to photosynthesis). Another major
+carbon exchange process connects the atmosphere and the ocean. Atmospheric
+\ac{CO2} is exchanged with the surface ocean through gas exchange, which is
+driven by the partial \ac{CO2} pressure difference between the air and the sea
+\autocite{Ciais2013}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\largeFigureWidth{}]{
+    ch02_scientific_background/figs/carbon_cycle_perturbation.png}
+  \caption{Schematic representation of the overall perturbation of the global
+    carbon cycle caused by anthropogenic activities, averaged globally for the
+    decade \range{2009}{2018}. Arrows represent carbon exchange fluxes; circles
+    carbon reservoirs. More details are given in the legend of this figure.
+    Taken from \textcite{Friedlingstein2019}.}
+  \label{fig:02:carbon_cycle_perturbation}
+\end{figure}
 
@@
 \label{sec:02:techniques}
+
+- Process-based weighting (Knutti + MDER)
+
+- emergent constraints
+
+- Wenzel et al., Nature (2016) as example (refer to carbon cycle chapter)
+
+- Review of ECS emergent constraints is given in chapter on Schlund et al., ESD
+(2020)

### 2020-12-02 ch02_scientific_background/scientific_background.tex
Started section on antrhopogenic influence on carbon cycle

@@
 
+\subsection{Anthropogenic Perturbations}
+\label{subsec:02:carbon_cycle_perturbations}
+
+Before the Industrial Era, the global carbon cycle was roughly in equilibrium.
+This can be inferred from ice core measurements, which show an almost constant
+atmospheric \ac{CO2} concentration over the last several thousand years before
+the Industrial Revolution in the \nth{19} century \autocite{Ciais2013}. Since
+the beginning of the Industrial Era, humanity is constantly emitting
+carbon-based \acp{GHG} (\eg{} \ac{CO2} and \ac{CH4}) into the atmosphere.
+Especially the atmospheric \ac{CO2} concentration has substantially increased,
+which has already been shown by Charles D. Keeling in 1976 by his continuous
+\ac{CO2} measurements at Mauna Loa, Hawaii that started in 1958
+\autocite{Keeling1976}. From 1958, the atmospheric \ac{CO2} concentration at
+Mauna Loa has steadily increased by about $100 \unit{ppm}$ to $410 \unit{ppm}$
+in the year 2019 \autocite{Keeling2005}. In addition to the monotonic increase,
+the so-called \enquote{Keeling Curve} is further superimposed with the seasonal
+\ac{CO2} cycle, which gives rise to local maxima of the atmospheric \ac{CO2}
+concentration in the Northern hemisphere winter and local minima in the
+Northern hemisphere summer \autocite[see
+\cref{subsec:02:carbon_cycle_overview}]{Keeling1995}. Due to its location in
+the middle of the Pacific Ocean, the Mauna Loa Observatory offers perfect
+conditions for \ac{CO2} measurements by being far away from big population
+centers. Moreover, its elevation of more than $3000 \unit{m}$ provides access
+to the well-mixed air of the Pacific Ocean in high altitudes, which prevents
+any interference from the vegetation present on the Hawaiian Islands.
+
+Apart from warming the Earth by altering its radiation budget, the emitted
+\ac{CO2} directly influences the carbon exchange fluxes of the global carbon
+cycle. which introduced net carbon fluxes between the different carbon
+reservoirs into the system, which left this steady state in shows ...
+
 \begin{figure}[t]

### 2020-12-03 ch02_scientific_background/scientific_background.tex
Finished section on carbon cycle

@@
 
-A possible way out is given by numerical climate models, which offer the
-possibility to simulate the Earth's climate on a computer. The first numerical
-climate models came up in the 1960s and were based on weather prediction models
-\autocite{Flato2011}. Early models from the 1970s simulated only the physical
-components of the climate system: atmosphere, land surface, ocean and sea ice
-(see \cref{fig:02:esms_historical_evolution}). The basis of these so-called
-\acp{AOGCM} \autocite{Flato2013} is the numerical solving of the differential
-equations describing the exchange of energy and matter between these physical
-components.
-
 \begin{figure}[t]
   \centering
-  \includegraphics[width=\figureWidth{}]{
+  \includegraphics[width=\FigureWidth{}]{
     ch02_scientific_background/figs/esms_historical_evolution.pdf}
@@
 \end{figure}
+
+A possible way out is given by numerical climate models, which offer the
+possibility to simulate the Earth's climate on a computer. The first numerical
+climate models came up in the 1960s and were based on weather prediction models
+\autocite{Flato2011}. Early models from the 1970s simulated only the physical
+components of the climate system: atmosphere, land surface, ocean and sea ice
+(see \cref{fig:02:esms_historical_evolution}). The basis of these so-called
+\acp{AOGCM} \autocite{Flato2013} is the numerical solving of the differential
+equations describing the exchange of energy and matter between these physical
+components.
 
@@
 atmosphere \autocite{Friedlingstein2019}. The residual part is absorbed by the
-two main carbon sinks of the planet, the terrestrial biosphere and the ocean.
-Since only atmospheric \ac{CO2} can act as \ac{GHG} by introducing an
+two other main carbon sinks of the planet, the terrestrial biosphere and the
+ocean. Since only atmospheric \ac{CO2} can act as \ac{GHG} by introducing an
 additional radiative forcing to the Earth System leading to increasing surface
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/cmip6.pdf}
+  \caption{Schematic of the experiment design of Phase 6 of the \acl{CMIP}
+    (\acs{CMIP}6). The center of the circle illustrates the four \acs{DECK}
+    (\acl{DECK}) experiments and the \acs{CMIP}6 historical simulation. The
+    circular sectors show additional science themes that can be explored
+    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}. Taken from
+    \textcite{Simpkins2017}.}
+  \label{fig:02:cmip6}
+\end{figure}
+
 A further main aspect is to provide a set of standardized experiments for
@@
 \acs{CMIP}6.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\figureWidth{}]{
-    ch02_scientific_background/figs/cmip6.pdf}
-  \caption{Schematic of the experiment design of Phase 6 of the \acl{CMIP}
-    (\acs{CMIP}6). The center of the circle illustrates the four \acs{DECK}
-    (\acl{DECK}) experiments and the \acs{CMIP}6 historical simulation. The
-    circular sectors show additional science themes that can be explored
-    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}. Taken from
-    \textcite{Simpkins2017}.}
-  \label{fig:02:cmip6}
-\end{figure}
 
@@
 forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two other main scenarios
-(called \enquote{Tier 1} experiments in \ac{ScenarioMIP}) are the \ac{SSP}2-4.5
+(called \emph{Tier 1} experiments in \ac{ScenarioMIP}) are the \ac{SSP}2-4.5
 and \ac{SSP}3-7.0 scenarios. In contrast to the \ac{ScenarioMIP} experiments,
@@
 
-Since one study presented in this thesis \autocite[see
-\cref{ch:06:paper_gpp}]{Schlund2020} aims to reduce uncertainties in carbon
-cycle-related processes, this chapter introduces the scientific background of
-the global carbon cycle.
+Since one study presented in this thesis aims to reduce uncertainties in carbon
+cycle-related processes \commentcite{Schlund2020}{see \cref{ch:06:paper_gpp}},
+this chapter introduces the scientific background of the global carbon cycle.
 
@@
   \centering
-  \includegraphics[width=\figureWidth{}]{
+  \includegraphics[width=\FigureWidth{}]{
     ch02_scientific_background/figs/carbon_cycle_schematic.png}
@@
 
-Before the Industrial Era, the global carbon cycle was roughly in equilibrium.
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/keeling_curve.pdf}
+  \caption{The Keeling Curve: monthly-mean atmospheric \acs{CO2} concentration
+    at the Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
+    \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
+    \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
+    concentration is superimposed with a seasonal oscillation caused by the
+    seasonal \acs{CO2} cycle (see \cref{subsec:02:carbon_cycle_overview}).}
+  \label{fig:02:keeling_curve}
+\end{figure}
+
+Before the Industrial Era, the global carbon cycle was roughly in a dynamic
+equilibrium, which means that exchange fluxes balanced each other and the
+amount of carbon in the different reservoirs did neither increase nor decrease.
 This can be inferred from ice core measurements, which show an almost constant
@@
 \ac{CO2} measurements at Mauna Loa, Hawaii that started in 1958
-\autocite{Keeling1976}. From 1958, the atmospheric \ac{CO2} concentration at
-Mauna Loa has steadily increased by about $100 \unit{ppm}$ to $410 \unit{ppm}$
-in the year 2019 \autocite{Keeling2005}. In addition to the monotonic increase,
-the so-called \enquote{Keeling Curve} is further superimposed with the seasonal
-\ac{CO2} cycle, which gives rise to local maxima of the atmospheric \ac{CO2}
-concentration in the Northern hemisphere winter and local minima in the
-Northern hemisphere summer \autocite[see
-\cref{subsec:02:carbon_cycle_overview}]{Keeling1995}. Due to its location in
-the middle of the Pacific Ocean, the Mauna Loa Observatory offers perfect
-conditions for \ac{CO2} measurements by being far away from big population
-centers. Moreover, its elevation of more than $3000 \unit{m}$ provides access
-to the well-mixed air of the Pacific Ocean in high altitudes, which prevents
-any interference from the vegetation present on the Hawaiian Islands.
-
-Apart from warming the Earth by altering its radiation budget, the emitted
-\ac{CO2} directly influences the carbon exchange fluxes of the global carbon
-cycle. which introduced net carbon fluxes between the different carbon
-reservoirs into the system, which left this steady state in shows ...
+\commentcite{Keeling1976}{see \cref{fig:02:keeling_curve}}. From 1958, the
+atmospheric \ac{CO2} concentration at Mauna Loa has steadily increased by about
+$100 \unit{ppm}$ to $410 \unit{ppm}$ in the year 2019 \autocite{Keeling2005}.
+In addition to the steady increase, the so-called \emph{Keeling Curve} is
+further superimposed with the seasonal \ac{CO2} cycle, which gives rise to
+local maxima of the atmospheric \ac{CO2} concentration in the Northern
+hemisphere winter and local minima in the Northern hemisphere summer
+\commentcite{Keeling1995}{see \cref{subsec:02:carbon_cycle_overview}}. Due to
+its location in the middle of the Pacific Ocean, the Mauna Loa Observatory
+offers perfect conditions for \ac{CO2} measurements by being far away from big
+population centers. Moreover, its elevation of more than $3000 \unit{m}$
+provides access to the well-mixed air of the Pacific Ocean in high altitudes,
+which prevents any interference from the vegetation present on the Hawaiian
+Islands.
 
@@
   \centering
-  \includegraphics[width=\largeFigureWidth{}]{
+  \includegraphics[width=\LargeFigureWidth{}]{
     ch02_scientific_background/figs/carbon_cycle_perturbation.png}
@@
 \end{figure}
+
+Apart from warming the Earth by altering its radiation budget, the
+anthropogenically emitted \ac{CO2} directly influences the carbon exchange
+fluxes of the global carbon cycle. Due to the excessive carbon in the
+atmosphere, there is now a net carbon flux from the atmosphere into the land
+and ocean reservoirs (see \cref{fig:02:carbon_cycle_perturbation}). Thus, the
+carbon cycle is not in a steady state anymore. In the decade
+\range{2009}{2018}, anthropogenic activities caused net carbon fluxes of $3.2
+\unit{\GtCyr{}}$ from the atmosphere into the terrestrial biosphere due to
+increased plant photosynthesis and $2.5 \unit{\GtCyr{}}$ from the atmosphere
+into the ocean due to increase dissolution of \ac{CO2} into the sea
+\autocite{Friedlingstein2019}. In the same time, the amount of carbon in the
+atmosphere reservoir increased with a rate of $4.9 \unit{\GtCyr{}}$, indicating
+that only about half of the anthropogenic \ac{CO2} emissions in the last decade
+remained in the atmosphere \autocite{Friedlingstein2019} where they can act as
+\ac{GHG}.
+
+Thus, this removal of \ac{CO2} from the atmosphere actively slows down global
+warming. However, whether this benefit will persist in the future remains
+unclear, which is primarily linked to two feedback processes connecting the
+climate system and the global carbon cycle: the \emph{concentration-carbon
+  feedback} and the \emph{climate-carbon feedback}
+\autocite{Friedlingstein2006, Gregory2009, Collins2013}. For the terrestrial
+biosphere, the concentration-carbon feedback is connected to the \emph{\ac{CO2}
+  fertilization effect} \autocite{Walker2020}, that causes an increase of
+photosynthesis rates when the atmospheric \ac{CO2} concentration increases,
+which in turns removes \ac{CO2} from the atmosphere, forming a negative
+feedback. For the ocean, the concentration-carbon feedback is negative as well.
+In this case, an elevated atmospheric \ac{CO2} concentration causes an
+increased dissolution of \ac{CO2} into the sea, which increases the ocean
+carbon uptake. On the other hand, the climate-carbon feedback is thought to be
+positive for both the terrestrial biosphere and the ocean
+\autocite{Gregory2009}. In the first case, temperature and precipitation
+changes due to anthropogenic activities decrease the land carbon uptake because
+of increased temperature and water stress on photosynthesis and higher
+ecosystem respiration costs, which accelerates global warming due to more
+\ac{CO2} that remains in the atmosphere. For the ocean, increased temperatures
+lead to a reduction of vertical transport in the ocean resulting from increased
+stability and reduced solubility of \ac{CO2} in the sea, which reduces the
+ocean carbon uptake and enhances climate change \autocite{Gregory2009}.
 

### 2020-12-04 ch01_introduction/introduction.tex
Finished section of ESMs

@@
 \label{ch:01:introduction}
+
+\section{Motivation}
+\label{ch:01:motivation}
+
+- Why is climate change relevant for the society?
+
+- Cite latest warming reports, also other variables (droughts, extreme events,
+etc.)
+
+- basic physics of greenhouse gases (vibration modes, etc.)
+
+- well-mixed greenhouse gases (water vapor, CO2, CH4) + sources
+
+- long-lived vs. short-lived GHGs?
+
+- aerosols + cooling effect
+
+- natural vs. anthropogenic effects (D \& A) -> clear evidence that climate
+change over 21st century is caused by human influence
+(https://globalwarmingindex.org/, Haustein et al., 2017)
 

### 2020-12-04 ch02_scientific_background/scientific_background.tex
Finished section of ESMs

@@
 
+A possible way out is given by numerical climate models, which offer the
+possibility to simulate the Earth's climate on a computer. The first numerical
+climate models came up in the 1960s and were based on weather prediction models
+\autocite{Flato2011}. Early models from the 1970s simulated only the physical
+components of the climate system: atmosphere, land surface, ocean and sea ice
+(see \cref{fig:02:esms_historical_evolution}). The basis of these so-called
+\acp{AOGCM} \autocite{Flato2013} is the numerical solving of the differential
+equations describing the exchange of energy and matter between these physical
+components.
+
 \begin{figure}[t]
@@
 \end{figure}
-
-A possible way out is given by numerical climate models, which offer the
-possibility to simulate the Earth's climate on a computer. The first numerical
-climate models came up in the 1960s and were based on weather prediction models
-\autocite{Flato2011}. Early models from the 1970s simulated only the physical
-components of the climate system: atmosphere, land surface, ocean and sea ice
-(see \cref{fig:02:esms_historical_evolution}). The basis of these so-called
-\acp{AOGCM} \autocite{Flato2013} is the numerical solving of the differential
-equations describing the exchange of energy and matter between these physical
-components.
 
@@
 
+A further main aspect is to provide a set of standardized experiments for
+global climate model simulations. To participate in the latest phase of
+\ac{CMIP}, \acs{CMIP}6, climate models need to run a \emph{historical}
+simulation of the period \range{1850}{2014} and the so-called \ac{DECK}
+experiments, which include a pre-industrial control run (\emph{piControl}), a
+historical Atmospheric \ac{MIP} simulation (\emph{amip}), a simulation forced
+with an abrupt quadrupling of \ac{CO2} (\emph{abrupt-4xCO2}) and a simulation
+forced with a \pct{1} per year increase of the atmospheric \ac{CO2}
+concentration (\emph{1pctCO2}) \autocite{Eyring2016}. This is shown in the
+center of \cref{fig:02:cmip6}, which illustrates the experimental design of
+\acs{CMIP}6.
+
 \begin{figure}[t]
@@
 \end{figure}
-
-A further main aspect is to provide a set of standardized experiments for
-global climate model simulations. To participate in the latest phase of
-\ac{CMIP}, \acs{CMIP}6, climate models need to run a \emph{historical}
-simulation of the period \range{1850}{2014} and the so-called \ac{DECK}
-experiments, which include a pre-industrial control run (\emph{piControl}), a
-historical Atmospheric \ac{MIP} simulation (\emph{amip}), a simulation forced
-with an abrupt quadrupling of \ac{CO2} (\emph{abrupt-4xCO2}) and a simulation
-forced with a \pct{1} per year increase of the atmospheric \ac{CO2}
-concentration (\emph{1pctCO2}) \autocite{Eyring2016}. This is shown in the
-center of \cref{fig:02:cmip6}, which illustrates the experimental design of
-\acs{CMIP}6.
 
@@
 
-- anthropogenic forcing vs. natural forcing (volcanoes, solar cycle, etc.)
+Simulations from climate model ensembles of \ac{CMIP} allow us to assess the
+impact of future climate change in a consistent and transparent way. Especially
+the \ac{ScenarioMIP} experiments can give valuable insights into the upcoming
+development of the Earth system by providing \emph{projections} of the future
+climate. In contrast to climate predictions, climate projections run over
+multiple decades and depend upon the future scenario considered, which are
+based on assumptions that may or may not turn out to be correct. On the
+contrary, climate predictions are attempts to predict the actual evolution of
+the climate on much shorter time scales from seasons to years. Similar to any
+other scientific experiment, climate model projections suffer from associated
+uncertainties. There are three major sources of climate model projections we
+can distinguish: natural variability, climate response uncertainty and emission
+uncertainty \autocite{Hawkins2009, Hawkins2010}.
+\Cref{fig:02:sources_of_uncertainty} shows these three sources for the
+projected global mean surface temperature anomaly over the \nth{21} century.
 
-- basic physics of greenhouse gases (vibration modes, etc.)
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch02_scientific_background/figs/sources_of_uncertainty.pdf}
+  \caption{Schematic illustration of the relative importance of different
+    sources of uncertainties in climate model projections and their evolution
+    in time. (a) Time series of the anomaly of the decadal and global mean
+    surface temperature relative to the period \range{1961}{1980}. The black
+    line shows the historical observations with estimates of uncertainty from
+    climate models (gray). The remaining colors show different sources of
+    uncertainty in future climate projections: Natural variability
+    (orange), climate response uncertainty (blue) and emission uncertainty
+    (green) \autocite{Hawkins2009, Hawkins2010}. Climate response uncertainty
+    can (b) increase in newer generations of climate models when a new process
+    is discovered to be relevant or (c) decrease with additional model
+    improvements and observational constraints. Taken from
+    \textcite{Cubasch2013}.}
+  \label{fig:02:sources_of_uncertainty}
+\end{figure}
 
-- historical vs. future: in historical (future), uncertainty in forcing low
-(high)
+\emph{Natural variability} is connected to the chaotic nature of the Earth
+system that arises from complex interactions between the ocean, atmosphere,
+land, biosphere and cryosphere \autocite{Cubasch2013}. It constitutes a
+fundamental limit of how precisely we can project the future climate since it
+is inherent in the Earth system and cannot be eliminated by more knowledge and
+more advanced climate models. Natural variability is more relevant on regional
+and local scales than on continental or global scales. Further contributions to
+natural variability on longer time scales come from phenomena like the
+\ac{ENSO} or the \ac{NAO} and from externally (and thus explainable) events
+like volcanic eruptions and variations in the solar activity. Natural
+variability can be seen as the \emph{noise} in the climate record as opposed to
+the anthropogenic \emph{signal} \autocite{Cubasch2013}. As illustrated by
+\cref{fig:02:sources_of_uncertainty}, the uncertainty associated with natural
+variability is constant over time.
 
-- definition of feedbacks, examples
+The second source of uncertainty in climate model projections is \emph{emission
+  uncertainty}. This arises from the different possible trajectories in terms
+of future forcing (\acp{GHG}, aerosols, land use changes, \etc{}) humankind
+might take. Examples for these are the \ac{SSP}-based experiments given by
+\ac{ScenarioMIP} that include a variety of different scenarios from a
+sustainable future to a full fossil fuel-based development (see
+\cref{subsec:02:cmip}). A possible approach to quantify emission uncertainty is
+to assess the climate impact of these different trajectories. Since the
+emission uncertainty strongly depends on the future development of the human
+society, it cannot be reduced by improving climate models. In contrast to
+natural variability, the emission uncertainty increases over time in climate
+projections, since estimating forcings for the near future is easier than for
+the far future.
 
-\section{Climate Sensitivity}
-\label{sec:02:climate_sensitivity}
+Finally, the third source of uncertainty in climate model projections is the
+\emph{climate response uncertainty}, which comes from our imperfect knowledge
+of how the climate system will respond to future anthropogenic forcings. Due to
+the complexity of the Earth system, the future climate could develop in many
+different ways that are all consistent with our current knowledge and models
+\autocite{Cubasch2013}. In the context of climate model ensembles, the climate
+response uncertainty is often also called \emph{model uncertainty} and reflects
+the different responses of the different climate models to a given forcing.
+Even though all climate models are built on the same physical principles, they
+differ in terms of spatial resolution, processes included and parametrizations
+of unresolved processes. The latter can be thought of as different
+approximations that are necessary to represent processes that take place on
+scales smaller than the common size of grid boxes in modern global climate
+models (about $\ang{1} \times \ang{1}$ horizontally), which is for example the
+case for some cloud processes and vegetation processes.
+
+These differences in the climate models also give rise to different intensities
+of \emph{climate feedbacks} (or even their presence/absence) in the models. A
+climate feedback is a mechanism that either amplifies (\emph{positive
+  feedback}) or weakens (\emph{negative feedback}) the effect of an external
+forcing. An example of a positive feedback is the water vapor feedback, where
+the increased surface temperature (caused by anthropogenic forcing) leads to an
+enhanced evaporation of water which enhances the amount of water vapor in the
+atmosphere. Since water vapor itself is a powerful \ac{GHG}, this amplifies the
+effect of the anthropogenic forcing by further increasing the surface
+temperatures \autocite{Cubasch2013}. Further examples and a mathematical
+description of feedbacks are given in
+\cref{sec:02:feedbacks_and_climate_sensitivity}.
+
+As sciences evolves, representations of already included processes can be
+improved in climate models. Moreover, new geophysical and biogeochemical
+processes can be added to them. On the one hand, this can increase the climate
+response uncertainty when a new process is discovered to be relevant
+\commentcite{Cubasch2013}{see \cref{fig:02:sources_of_uncertainty}b}. However,
+such an increase corresponds to a previously unmeasured uncertainty. An example
+for this has recently happened in \acs{CMIP}6: most likely due to changes in
+the cloud representation of the models the spread in the projected global mean
+surface air temperature caused by a doubling of the atmospheric \ac{CO2}
+concentration has substantially increased in \ac{CMIP}6 compared to older
+\ac{CMIP} generations \autocite{Zelinka2020}. On the other hand, the climate
+response uncertainty can decrease with additional model improvements and better
+understanding of the Earth system (see \cref{fig:02:sources_of_uncertainty}c).
+Moreover, it can also be reduced by observational constraints, which is the
+main topic of this thesis.
+
+\section{Feedback Processes and Climate Sensitivity}
+\label{sec:02:feedbacks_and_climate_sensitivity}
+
+- Overview over feedbacks (Cubasch et al., 2013, Figure 1.2)
+
+- Mathematical description of feedbacks, geometric series
+
+- Equations for feedbacks + decomposition
 
 - Physical Feedbacks (ECS)
-
-- Equations for feedbacks + decomposition
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\FigureWidth{}]{
-    ch02_scientific_background/figs/carbon_cycle_schematic.png}
-  \caption{Simplified schematic of the global carbon cycle including the
-    typical turnover time scales for carbon transfers through the major
-    reservoirs (atmosphere, land surface and ocean). Taken from
-    \textcite{Ciais2013}.}
-  \label{fig:02:carbon_cycle_schematic}
-\end{figure}
-
 A schematic overview of the global carbon cycle is shown in
@@
 \autocite{Ciais2013}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/carbon_cycle_schematic.png}
+  \caption{Simplified schematic of the global carbon cycle including the
+    typical turnover time scales for carbon transfers through the major
+    reservoirs (atmosphere, land surface and ocean). Taken from
+    \textcite{Ciais2013}.}
+  \label{fig:02:carbon_cycle_schematic}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\FigureWidth{}]{
-    ch02_scientific_background/figs/keeling_curve.pdf}
-  \caption{The Keeling Curve: monthly-mean atmospheric \acs{CO2} concentration
-    at the Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
-    \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
-    \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
-    concentration is superimposed with a seasonal oscillation caused by the
-    seasonal \acs{CO2} cycle (see \cref{subsec:02:carbon_cycle_overview}).}
-  \label{fig:02:keeling_curve}
-\end{figure}
-
 Before the Industrial Era, the global carbon cycle was roughly in a dynamic
@@
   \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
-    ch02_scientific_background/figs/carbon_cycle_perturbation.png}
-  \caption{Schematic representation of the overall perturbation of the global
-    carbon cycle caused by anthropogenic activities, averaged globally for the
-    decade \range{2009}{2018}. Arrows represent carbon exchange fluxes; circles
-    carbon reservoirs. More details are given in the legend of this figure.
-    Taken from \textcite{Friedlingstein2019}.}
-  \label{fig:02:carbon_cycle_perturbation}
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/keeling_curve.pdf}
+  \caption{The Keeling Curve: monthly-mean atmospheric \acs{CO2} concentration
+    at the Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
+    \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
+    \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
+    concentration is superimposed with a seasonal oscillation caused by the
+    seasonal \acs{CO2} cycle (see \cref{subsec:02:carbon_cycle_overview}).}
+  \label{fig:02:keeling_curve}
 \end{figure}
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch02_scientific_background/figs/carbon_cycle_perturbation.png}
+  \caption{Schematic representation of the overall perturbation of the global
+    carbon cycle caused by anthropogenic activities, averaged globally for the
+    decade \range{2009}{2018}. Arrows represent carbon exchange fluxes; circles
+    carbon reservoirs. More details are given in the legend of this figure.
+    Taken from \textcite{Friedlingstein2019}.}
+  \label{fig:02:carbon_cycle_perturbation}
+\end{figure}
+
 Thus, this removal of \ac{CO2} from the atmosphere actively slows down global
@@
 unclear, which is primarily linked to two feedback processes connecting the
-climate system and the global carbon cycle: the \emph{concentration-carbon
-  feedback} and the \emph{climate-carbon feedback}
+physical climate system and the global carbon cycle: the
+\emph{concentration-carbon feedback} and the \emph{climate-carbon feedback}
 \autocite{Friedlingstein2006, Gregory2009, Collins2013}. For the terrestrial
@@
 
+- Problems with multi-model analysis (search for refs in Schlund et al., ESD,
+2020) -> models are not independent, multi-model ensemble is not "real"
+statistical sample
+
 - Process-based weighting (Knutti + MDER)

### 2020-12-07 ch02_scientific_background/scientific_background.tex
Added climate feedback figure and adapted margins

@@
   \centering
-  \includegraphics[width=\FigureWidth{}]{
+  \includegraphics[width=\LargeFigureWidth{}]{
     ch02_scientific_background/figs/cmip6.pdf}
@@
 system, like the \ac{C4MIP} which focuses on the carbon cycle
-\autocite{Jones2016} or the \ac{CFMIP} which focuses on the evaluation of cloud
-feedbacks \autocite{Webb2017}. Other \acp{MIP} allow the assessment of future
-climate change. An example is the \ac{ScenarioMIP}, which provides common
-experiments that simulate different possible futures \autocite{ONeill2016}.
-These experiments are based on the so-called \acp{SSP}, a set of alternative
-pathways of future societal development \autocite{ONeill2013, ONeill2017}. For
-each experiment, a set of emissions and land use changes is calculated from the
+\autocite{Jones2016} or the \ac{AerChemMIP} which focuses on aerosol chemistry
+\autocite{Collins2017}. Other \acp{MIP} allow the assessment of future climate
+change. An example is the \ac{ScenarioMIP}, which provides common experiments
+that simulate different possible futures \autocite{ONeill2016}. These
+experiments are based on the so-called \acp{SSP}, a set of alternative pathways
+of future societal development \autocite{ONeill2013, ONeill2017}. For each
+experiment, a set of emissions and land use changes is calculated from the
 \acp{SSP} \autocite{Riahi2017} which are then used to force the global climate
@@
 description of feedbacks are given in
-\cref{sec:02:feedbacks_and_climate_sensitivity}.
+\cref{subsec:02:climate_feedbacks}.
 
@@
 
-\section{Feedback Processes and Climate Sensitivity}
-\label{sec:02:feedbacks_and_climate_sensitivity}
+\section{Climate Sensitivity}
+\label{sec:02:climate_sensitivity}
+
+Climate sensitivity refers to the change in the \ac{GSAT} that results from a
+change in the radiative forcing. In other words, it describes how sensitive the
+climate system is to an external forcing. The source of this forcing might
+either be natural (changes in the solar activity, volcanic eruptions, \etc{})
+or anthropogenic (emissions of \acp{GHG}, land use changes, \etc{}).
+
+\subsection{Climate Feedbacks}
+\label{subsec:02:climate_feedbacks}
+
+As already described in \cref{subsec:02:source_of_uncertainties}, the effects
+of an external forcing acting on the climate system can additionally be
+amplified or weakened by climate feedbacks. Thus, feedback processes play a
+crucial role in the evaluation of climate sensitivity.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch02_scientific_background/figs/feedbacks_overview.pdf}
+  \caption{Climate feedbacks and corresponding time scales. \enquote{+} refers
+    to positive feedbacks, which amplify the effect of the external forcing
+    (\eg{} the water vapor feedback). \enquote{-} refers to negative feedbacks,
+    which weaken the effect of the external forcing (\eg{} the longwave
+    radiation feedback). \enquote{+/-} refers to feedbacks which might be
+    either positive or negative (\eg{} the cloud feedback). The smaller box
+    highlights the large differences in time scales for the various feedbacks.
+    Taken from \textcite{Cubasch2013}.}
+  \label{fig:02:feedbacks_overview}
+\end{figure}
 
 - Overview over feedbacks (Cubasch et al., 2013, Figure 1.2)
+
+- Timescales of feedbacks
+
+- Longwave radiation feedback = Planck feedback
 

### 2020-12-08 ch02_scientific_background/scientific_background.tex
Added overview of climate feedbacks

@@
 climate feedback is a mechanism that either amplifies (\emph{positive
-  feedback}) or weakens (\emph{negative feedback}) the effect of an external
-forcing. An example of a positive feedback is the water vapor feedback, where
-the increased surface temperature (caused by anthropogenic forcing) leads to an
-enhanced evaporation of water which enhances the amount of water vapor in the
-atmosphere. Since water vapor itself is a powerful \ac{GHG}, this amplifies the
-effect of the anthropogenic forcing by further increasing the surface
-temperatures \autocite{Cubasch2013}. Further examples and a mathematical
-description of feedbacks are given in
+  feedback}) or diminishes (\emph{negative feedback}) the effect of an external
+forcing. An example of a strong positive feedback is the water vapor feedback,
+in which the increased surface temperature (caused by anthropogenic forcing)
+leads to an enhanced evaporation of water which enhances the amount of water
+vapor in the atmosphere. Since water vapor itself is a powerful \ac{GHG}, this
+amplifies the effect of the anthropogenic forcing by further increasing the
+surface temperatures \autocite{Cubasch2013}. Further examples and a
+mathematical description of feedbacks are given in
 \cref{subsec:02:climate_feedbacks}.
@@
 of an external forcing acting on the climate system can additionally be
-amplified or weakened by climate feedbacks. Thus, feedback processes play a
+amplified or diminished by climate feedbacks. Thus, feedback processes play a
 crucial role in the evaluation of climate sensitivity.
+\Cref{fig:02:feedbacks_overview} shows an overview of important feedbacks in
+the Earth system with their corresponding time scales on which they operate.
 
@@
     (\eg{} the water vapor feedback). \enquote{-} refers to negative feedbacks,
-    which weaken the effect of the external forcing (\eg{} the longwave
+    which diminish the effect of the external forcing (\eg{} the longwave
     radiation feedback). \enquote{+/-} refers to feedbacks which might be
@@
 
-- Overview over feedbacks (Cubasch et al., 2013, Figure 1.2)
+An example for a positive feedback is the already mentioned \emph{water vapor
+  feedback}. Being the primary \ac{GHG} in the Earth's atmosphere, water vapor
+is the largest contributor to the natural greenhouse effect. Since its amount
+in the atmosphere is mainly controlled by the air temperature and anthropogenic
+emissions of water vapor are negligible, the influence of water vapor on the
+climate system is described as a feedback mechanism and not as an external
+forcing \autocite{Myhre2013}. Basis of this feedback is the enhanced
+evaporation of water with increasing air temperatures. Each degree of warming
+allows the atmosphere to retain about \pct{7} more water vapor
+\autocite{Myhre2013}, which closes the positive feedback loop by further
+increasing air temperatures through the greenhouse effect. With a typical
+residence time of water vapor in the atmosphere of several days, the water
+vapor feedback operates on relatively short time scales. As the largest
+positive feedback in the Earth system \autocite{Soden2006}, the water vapor
+feedback amplifies any initial forcing (\eg{} caused by anthropogenic \ac{CO2}
+emissions) by a typical factor between two and three, rendering water vapor a
+fundamental agent of climate change \autocite{Myhre2013}. An example for a
+positive feedback that operates on longer time scales (several years) is the
+\emph{snow/ice albedo feedback}, in which the surface albedo decreases as
+highly reflective ice and snow surfaces melt with global warming, exposing the
+darker and more absorbing surfaces below \autocite{Cubasch2013}.
 
-- Timescales of feedbacks
+In contrast to positive feedbacks, negative feedbacks diminish the effect of an
+external forcing. An example for this is the \emph{blackbody feedback} (also
+known as \emph{Planck feedback} or \emph{longwave radiation feedback}), which
+is the strongest negative feedback in the Earth system \autocite{Cubasch2013}.
+It is based on the thermal electromagnetic radiation that any object with a
+non-zero temperature emits (the so-called \emph{blackbody radiation}). Since
+the power of this radiation strongly depends on the temperature of the object,
+higher surface temperatures of Earth increase the outgoing longwave radiation
+flux from the surface which reduces the effect of the external forcing and
+cools the planet.
 
-- Longwave radiation feedback = Planck feedback
+For some domains of the Earth system, feedbacks can be positive and/or
+negative, since a variety of different mechanisms is involved. An example for
+this is the cloud feedback. Changes in clouds induced by climate change can
+cause both longwave (greenhouse warming) and shortwave (reflective cooling)
+effects, which both need to be considered for the overall cloud feedback
+\autocite{Boucher2013}. Relevant cloud properties that may change as a response
+to an external forcing and that may alter the Earth's radiative budget are
+cloud cover, cloud optical thickness, cloud altitude and the geographical
+distribution of clouds. Examples for robust cloud feedback processes are the
+rise of high clouds in a warming climate which traps longwave radiation and
+enhances global warming and the reduction in mid- and low-level cloud cover
+which diminishes the reflection of incoming solar radiation and also increases
+the surface warming \autocite{Boucher2013}. In global climate model ensembles,
+the overall cloud feedback shows a large range with positive and negative
+values, but tends to be slightly positive on average \autocite{Soden2006,
+  Dufresne2008, Vial2013, Zelinka2020}. This large uncertainty in the cloud
+feedback is a major reason for uncertainties in the climate sensitivity of
+climate models \autocite{Boucher2013, Flato2013}.
 
-- Mathematical description of feedbacks, geometric series
+Further examples of feedbacks with positive and negative contributions are
+biogeochemical feedbacks. Negative contributions come from increased \ac{CO2}
+fluxes into the land and ocean carbon reservoirs due to increased
+photosynthesis rates and \ac{CO2} dissolution in the sea, which decrease the
+atmospheric \ac{CO2} content and diminish global warming. An example for a
+positive contribution is the decreased solubility of \ac{CO2} in water in a
+warmer climate, which reduces the atmosphere-ocean \ac{CO2} flux and enhances
+climate change. More details on this are given in
+\cref{subsec:02:carbon_cycle_perturbations}.
 
-- Equations for feedbacks + decomposition
+\subsection{Mathematical Description of Feedbacks}
+\label{subsec:02:mathematics_of_feedbacks}
+
+- Mathematical description of feedbacks
+
+- Equations for feedbacks + decomposition (reference for feedbacks: Planck, Roe
+2009)
 
@@
 
-\section{Techniques to reduce Uncertainties in Climate Model Projections}
+\section{Techniques to Reduce Uncertainties in Climate Model Projections}
 \label{sec:02:techniques}

### 2020-12-09 ch02_scientific_background/scientific_background.tex
Added subsection on mathematical framework for feedbacks

@@
     ch02_scientific_background/figs/sources_of_uncertainty.pdf}
-  \caption{Schematic illustration of the relative importance of different
-    sources of uncertainties in climate model projections and their evolution
-    in time. (a) Time series of the anomaly of the decadal and global mean
-    surface temperature relative to the period \range{1961}{1980}. The black
-    line shows the historical observations with estimates of uncertainty from
+  \caption{Schematic illustrating the importance of different sources of
+    uncertainties in climate model projections and their evolution in time. (a)
+    Time series of the anomaly of the decadal and global mean surface
+    temperature relative to the period \range{1961}{1980}. The black line
+    shows the historical observations with estimates of uncertainty from
     climate models (gray). The remaining colors show different sources of
-    uncertainty in future climate projections: Natural variability
-    (orange), climate response uncertainty (blue) and emission uncertainty
-    (green) \autocite{Hawkins2009, Hawkins2010}. Climate response uncertainty
-    can (b) increase in newer generations of climate models when a new process
-    is discovered to be relevant or (c) decrease with additional model
+    uncertainty in future climate projections: Natural variability (orange),
+    climate response uncertainty (blue) and emission uncertainty (green)
+    \autocite{Hawkins2009, Hawkins2010}. Climate response uncertainty can (b)
+    increase in newer generations of climate models when a new process is
+    discovered to be relevant or (c) decrease with additional model
     improvements and observational constraints. Taken from
@@
 surface temperatures \autocite{Cubasch2013}. Further examples and a
-mathematical description of feedbacks are given in
+mathematical framework for the analysis of feedbacks are given in
 \cref{subsec:02:climate_feedbacks}.
@@
 for this has recently happened in \acs{CMIP}6: most likely due to changes in
-the cloud representation of the models the spread in the projected global mean
-surface air temperature caused by a doubling of the atmospheric \ac{CO2}
-concentration has substantially increased in \ac{CMIP}6 compared to older
-\ac{CMIP} generations \autocite{Zelinka2020}. On the other hand, the climate
-response uncertainty can decrease with additional model improvements and better
-understanding of the Earth system (see \cref{fig:02:sources_of_uncertainty}c).
-Moreover, it can also be reduced by observational constraints, which is the
-main topic of this thesis.
+the cloud representation of the models the spread in the projected \ac{GSAT}
+caused by a doubling of the atmospheric \ac{CO2} concentration has
+substantially increased in \ac{CMIP}6 compared to older \ac{CMIP} generations
+\autocite{Zelinka2020}. On the other hand, the climate response uncertainty can
+decrease with additional model improvements and better understanding of the
+Earth system (see \cref{fig:02:sources_of_uncertainty}c). Moreover, it can also
+be reduced by observational constraints, which is the main topic of this
+thesis.
 
@@
 fluxes into the land and ocean carbon reservoirs due to increased
-photosynthesis rates and \ac{CO2} dissolution in the sea, which decrease the
-atmospheric \ac{CO2} content and diminish global warming. An example for a
-positive contribution is the decreased solubility of \ac{CO2} in water in a
-warmer climate, which reduces the atmosphere-ocean \ac{CO2} flux and enhances
-climate change. More details on this are given in
+photosynthesis rates and \ac{CO2} dissolution in the sea, respectively, which
+decrease the atmospheric \ac{CO2} content and diminish global warming. An
+example for a positive contribution is the decreased solubility of \ac{CO2} in
+water in a warmer climate, which reduces the atmosphere-ocean \ac{CO2} flux and
+enhances climate change. More details on this are given in
 \cref{subsec:02:carbon_cycle_perturbations}.
 
-\subsection{Mathematical Description of Feedbacks}
-\label{subsec:02:mathematics_of_feedbacks}
+\subsection{Mathematical Framework for Feedbacks Analysis}
+\label{subsec:02:mathematical_framework_feedbacks}
 
-- Mathematical description of feedbacks
+The foundation for a basic mathematical framework for the analysis of climate
+feedbacks is a simple energy balance model \autocite{Gregory2009, Roe2009}.
+Anthropogenic activities in the Earth system like the emissions of \acp{GHG} or
+aerosols introduce an external forcing to the climate system, which is
+quantified with a radiative forcing $F$ measured in $\si{W m^{-2}}$. To restore
+a stable state, the climate system opposes this forcing with a climate response
+$R$, leading to a net energy flux of
+\begin{equation}
+  N = F + R
+  \label{eq:02:N}
+\end{equation}
+into the system. Positive values of $N$, $F$ and $R$ indicate incoming fluxes;
+usually $F > 0$ and $R < 0$. On long time scales (multiple years), the net
+incoming radiative flux at the \ac{TOA} and the net heat flux into the ocean
+are basically equal definitions of $N$ since nearly all of the Earth's heat
+capacity resides in the ocean \autocite{Gregory2009}. While $N \neq 0$, the
+climate system evolves; when $N = 0$ a new steady state has been reached.
 
-- Equations for feedbacks + decomposition (reference for feedbacks: Planck, Roe
-2009)
+To quantify the effects of different feedbacks, a reference system with a basic
+response needs to be defined, which is a crucial aspect of feedback analysis
+\autocite{Roe2009}. Usually, the idealization of a blackbody Earth (without an
+atmosphere) is used for that: In equilibrium, the incoming solar irradiance is
+balanced with an outgoing thermal irradiance $J_0$ that solely depends on the
+global mean surface temperature $T_0$ following the Stefan–Boltzmann law
+\begin{equation}
+  J_0 = -\sigma T_0^4.
+  \label{eq:02:stefan_boltzmann_law_eq}
+\end{equation}
+$\sigma \approx 5.67 \unit{W m^{-2} K^{-4}}$ is the Stefan–Boltzmann constant.
+To answer an external forcing $F$, the climate system reacts with a response
+$R$ expressed by a change in the global mean surface temperature $\Delta T$:
+\begin{equation}
+  J_0 + R = -\sigma \left( T_0 + \Delta T \right)^4
+  \label{eq:02:stefan_boltzmann_law_non_eq}
+\end{equation}
+Since the temperature change caused by an anthropogenic forcing is much smaller
+than the equilibrium temperature $\Delta T \ll T_0 \approx 255 \unit{K}$, a
+simple \nth{1}-order Taylor expansion can be used to linearize the blackbody
+response:
+\begin{equation}
+  -\sigma \left( T_0 + \Delta T \right)^4 \approx J_0 - 4 \sigma T_0^3 \cdot
+  \Delta T
+  \label{eq:02:stefan_boltzmann_law_lin}
+\end{equation}
+Thus, by comparing
+\cref{eq:02:stefan_boltzmann_law_non_eq,eq:02:stefan_boltzmann_law_lin} the
+climate response $R$ can be expressed as
+\begin{equation}
+  R = -4 \sigma T_0^3 \cdot \Delta T \coloneqq \lambda_\text{BB} \cdot \Delta T
+  \label{eq:02:linear_blackbody_response}
+\end{equation}
+with the blackbody feedback parameter $\lambda_\text{BB} \approx -3.8 \unit{W
+  m^{-2} K^{-1}}$. Results from climate models and analyses of observations
+confirm this linear relationship between $R$ and $\Delta T$
+\autocite{Gregory2004}. However, the value of this linear constant $\lambda$,
+the \emph{climate feedback parameter}, is found to be considerably larger than
+the blackbody response ($\lambda \approx -1.0 \unit{W m^{-2} K^{-1}}$),
+indicating that additional processes affect the Earth's radiative balance: the
+climate feedbacks \autocite{Flato2013, Gregory2009}. Since climate models
+suggest that the radiative effects of these additional feedbacks are also
+proportional to $\Delta T$ \autocite{Gregory2008a},
+\cref{eq:02:linear_blackbody_response} can be adapted to
+\begin{equation}
+  R = \lambda \cdot \Delta T = \left( \lambda_\text{BB} + \lambda_\text{WV} +
+  \lambda_\text{Albedo} + \lambda_\text{Cloud} + \ldots \right) \cdot \Delta T.
+  \label{eq:02:linear_response}
+\end{equation}
+$\lambda_\text{WV}$ refers to the water vapor feedback, $\lambda_\text{Albedo}$
+to the snow/ice albedo feedback and $\lambda_\text{Cloud}$ to the cloud
+feedback. Thus, the overall climate feedback parameter $\lambda$ can be written
+as the sum of the individual feedback parameters $\lambda_i$:
+\begin{equation}
+  \lambda = \sum_i \lambda_i
+  \label{eq:02:lambda_as_sum_of_lambdas}
+\end{equation}
+Positive values of $\lambda_i$ indicate positive feedbacks (\eg{} the water
+vapor feedback) and negative values indicate negative feedbacks (\eg{} the
+blackbody feedback). This equation assumes that the individual radiative
+responses from the different feedbacks are independent, which is a reasonable
+first-order approximation but not entirely true \autocite{Soden2008}.
+
+\subsection{Equilibrium and \acl{ECS}}
+\label{subsec:02:ecs}
 

### 2020-12-10 ch02_scientific_background/scientific_background.tex
Added subsection on ECS

@@
 natural variability, the emission uncertainty increases over time in climate
-projections, since estimating forcings for the near future is easier than for
+projections, since estimating forcings for the near future is simpler than for
 the far future.
@@
 \emph{climate response uncertainty}, which comes from our imperfect knowledge
-of how the climate system will respond to future anthropogenic forcings. Due to
-the complexity of the Earth system, the future climate could develop in many
+of how the climate system will respond to anthropogenic forcings. Due to the
+complexity of the Earth system, the future climate could develop in many
 different ways that are all consistent with our current knowledge and models
@@
 models (about $\ang{1} \times \ang{1}$ horizontally), which is for example the
-case for some cloud processes and vegetation processes.
+case for cloud processes and vegetation processes.
 
@@
 in which the increased surface temperature (caused by anthropogenic forcing)
-leads to an enhanced evaporation of water which enhances the amount of water
+leads to an enhanced evaporation of water which increases the amount of water
 vapor in the atmosphere. Since water vapor itself is a powerful \ac{GHG}, this
@@
 
-Climate sensitivity refers to the change in the \ac{GSAT} that results from a
+Climate sensitivity refers to the change in \ac{GSAT} that results from a
 change in the radiative forcing. In other words, it describes how sensitive the
@@
 the power of this radiation strongly depends on the temperature of the object,
-higher surface temperatures of Earth increase the outgoing longwave radiation
-flux from the surface which reduces the effect of the external forcing and
-cools the planet.
+higher surface temperatures of the Earth increase the outgoing longwave
+radiation flux from the surface which reduces the effect of the external
+forcing and cools the planet.
 
@@
 
-The foundation for a basic mathematical framework for the analysis of climate
-feedbacks is a simple energy balance model \autocite{Gregory2009, Roe2009}.
-Anthropogenic activities in the Earth system like the emissions of \acp{GHG} or
-aerosols introduce an external forcing to the climate system, which is
-quantified with a radiative forcing $F$ measured in $\si{W m^{-2}}$. To restore
-a stable state, the climate system opposes this forcing with a climate response
-$R$, leading to a net energy flux of
+The foundation of a basic mathematical framework for climate feedback analysis
+is a simple energy balance model \autocite{Gregory2009, Roe2009}. Anthropogenic
+activities in the Earth system like the emissions of \acp{GHG} or aerosols
+introduce an external forcing to the climate system, which is quantified with a
+radiative forcing $F$ measured in $\si{W m^{-2}}$. To restore a stable state,
+the climate system opposes this forcing with a climate response $R$, leading to
+a net energy flux of
 \begin{equation}
@@
 response needs to be defined, which is a crucial aspect of feedback analysis
-\autocite{Roe2009}. Usually, the idealization of a blackbody Earth (without an
-atmosphere) is used for that: In equilibrium, the incoming solar irradiance is
+\autocite{Roe2009}. Usually, the idealization of a blackbody Earth without an
+atmosphere is used for that: In equilibrium, the incoming solar irradiance is
 balanced with an outgoing thermal irradiance $J_0$ that solely depends on the
@@
 than the equilibrium temperature $\Delta T \ll T_0 \approx 255 \unit{K}$, a
-simple \nth{1}-order Taylor expansion can be used to linearize the blackbody
+simple first-order Taylor expansion can be used to linearize the blackbody
 response:
@@
 proportional to $\Delta T$ \autocite{Gregory2008a},
-\cref{eq:02:linear_blackbody_response} can be adapted to
+\cref{eq:02:linear_blackbody_response} can be adjusted to
 \begin{equation}
@@
 
-- Physical Feedbacks (ECS)
+An important metric describing climate sensitivity is the Equilibrium Climate
+Sensitivity. It is defined as the change in \ac{GSAT} after an instantaneous
+doubling of the atmospheric \ac{CO2} concentration from pre-industrial
+conditions once the climate system reaches radiative equilibrium
+\autocite{Bindoff2013}. Being already used in one of the first assessments of
+the anthropogenic climate change, the \emph{Charney Report} from 1979
+\autocite{Charney1979}, the Equilibrium Climate Sensitivity is one of the
+oldest metrics describing climate change. However, in practice this traditional
+definition of is not useful. Running fully-coupled \acp{ESM} into equilibrium
+is computationally expensive as it would require thousands of model years
+\autocite{Rugenstein2020}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/gregory_regression.pdf}
+  \caption{Global and annual mean net \acf{TOA} radiation $N$ \vs{} global and
+    annual mean near-surface air temperature $\Delta T$ for 150 years of a
+    simulation with an abrupt quadrupling of the atmospheric \ac{CO2}
+    concentration (4x\acs{CO2}) for the \acs{CMIP}5 multi-model mean (blue
+    squares). To account for energy leakage and model drift, a linear fit of
+    the corresponding pre-industrial control run is subtracted from the
+    4x\acs{CO2} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope of
+    the linear regression (black line) corresponds to the climate feedback
+    parameter $\lambda$ and the y-intercept corresponds to the radiative
+    forcing $F_\text{4x}$. These can be used to calculate the \ac{ECS} with
+    the Gregory regression method following to \cref{eq:02:ecs}
+    \autocite{Gregory2004}.}
+  \label{fig:02:gregory_regression}
+\end{figure}
+
+For this reason, the Equilibrium Climate Sensitivity is commonly approximated
+with the \acf{ECS}, which can be derived from only 150 model years of a
+simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
+(4x\acs{CO2}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
+is the simple energy balance model introduced in
+\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
+equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
+\begin{equation}
+  \Delta T = -\frac{F}{\lambda}.
+  \label{eq:02:delta_t_eq}
+\end{equation}
+Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
+from the external forcing $F$ and the climate feedback parameter $\lambda$. The
+steady state values for $F$ and $\lambda$ can be estimated from a 4x\acs{CO2}
+simulation that is not in equilibrium by extrapolation with a linear regression
+\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
+  regression}, the global and annual mean net \ac{TOA} radiation $N$ \vs{} the
+change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
+4x\acs{CO2} run are plotted (see \cref{fig:02:gregory_regression}). To account
+for energy leakage and remove any model drift present in the control climate, a
+linear fit of the corresponding pre-industrial control run is subtracted from
+the 4x\acs{CO2} simulation beforehand \autocite{Andrews2012}. Since the
+combination of \cref{eq:02:N,eq:02:linear_response} yields
+\begin{equation}
+  N = F + \lambda \cdot \Delta T,
+  \label{eq:02:N_vs_delta_t}
+\end{equation}
+$F$ is now given by the y-intercept of this linear regression
+($F_\text{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
+\begin{equation}
+  \text{ECS} = - \frac{F_\text{4x}}{2 \lambda}.
+  \label{eq:02:ecs}
+\end{equation}
+The factor of $2$ in the denominator accounts for the fact the traditional
+Equilibrium Climate Sensitivity is defined for an abrupt \ac{CO2} doubling,
+whereas here a simulation with an abrupt quadrupling is considered.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/gregory_regression_different_years.png}
+  \caption{As \cref{fig:02:gregory_regression}, but for different time periods
+    considered in the Gregory regression resulting in different values for the
+    \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
+    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
+    different time periods of a long running simulation (see legend). The
+    initial years are simulated many times for different initial conditions.
+    Light gray colors correspond to the \acs{CMIP}5 ensemble (150 years each).
+    Taken from \textcite{Knutti2017}.}
+  \label{fig:02:gregory_regression_different_years}
+\end{figure}
+
+Although commonly used in the literature, the \ac{ECS} is known to be only an
+approximation of the true Equilibrium Climate Sensitivity. One major reason for
+this is the state and time dependence of the global feedbacks
+\autocite{Knutti2015, Knutti2017}. As
+\cref{fig:02:gregory_regression_different_years} shows, the slope in the
+Gregory regression is not constant, but rather decreases over time when a long
+running 4x\acs{CO2} simulation with more than 1000 model years is considered.
+As a result, the climate feedback parameter $\lambda$ decreases over time,
+resulting in a higher \ac{ECS}. Major reasons for this are temperature
+dependencies of the feedbacks, atmospheric and oceanic adjustments over time,
+changing warming patterns over time, non-additive feedbacks and dependencies on
+the type and magnitude of the external forcings \autocite{Knutti2017}. All in
+all, this demonstrates the limits of the linear feedback framework introduced
+in \cref{subsec:02:mathematical_framework_feedbacks} that is not capable of
+describing non-linear effects. A second major reason for the discrepancies
+between the Equilibrium Climate Sensitivity and \ac{ECS} is the use of a
+4x\acs{CO2} instead of a 2x\acs{CO2} simulation. The factor of $2$ in the
+denominator of \cref{eq:02:ecs} only partly compensates this difference since
+the radiative forcing logarithmically depends on the atmospheric \ac{CO2}
+concentration \autocite{Huang2014}.
+
+However, despite these deficiencies the \ac{ECS} is still a practical estimate
+of the Equilibrium Climate Sensitivity. With the help of climate models,
+\textcite{Sherwood2020} showed that the \ac{ECS} is only about \pct{6} lower
+than the best estimate of the true equilibrium warming obtained from
+integrating climate models until a new steady state is reached. Nevertheless,
+for \acs{CMIP}6 long running simulations from the \ac{LongRunMIP}
+\autocite{Rugenstein2019} can be a promising way forward to estimate the true
+Equilibrium Climate Sensitivity for \acp{ESM}.
+
+\subsection{\acl{TCR}}
+\label{subsec:02:tcr}
+
+A further metric describing the climate sensitivity is the \ac{TCR}. In
+contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
+the climate system but describes the transient response of an evolving climate.
+\ac{TCR} is defined as the change in the annual mean \ac{GSAT} at the time of
+\ac{CO2} doubling in a simulation with a \pct{1} per year increase of the
+atmospheric \ac{CO2} concentration (1\%\acs{CO2}) averaged over a 20 year
+period centered at the time of the doubling (years \range{61}{80} when the
+first year is indexed with 1) \autocite{Bindoff2013}. To account for model
+drift, the annual mean changes in \ac{GSAT} are calculated relative to a
+corresponding pre-industrial control simulation smoothed with a linear fit that
+considers 140 model years (length of the 1\%\acs{CO2} simulation).
 
@@
 
-- different between effective climate sensitivity and equilibrium version
+- residual heat -> 90\% go into ocean (Knutti et al. 2017)
 
-- equations for ECS and TCR
+- equations for TCR
 

### 2020-12-11 ch02_scientific_background/scientific_background.tex
Finished section on climate sensitivity

@@
     ch02_scientific_background/figs/gregory_regression.pdf}
-  \caption{Global and annual mean net \acf{TOA} radiation $N$ \vs{} global and
-    annual mean near-surface air temperature $\Delta T$ for 150 years of a
-    simulation with an abrupt quadrupling of the atmospheric \ac{CO2}
-    concentration (4x\acs{CO2}) for the \acs{CMIP}5 multi-model mean (blue
-    squares). To account for energy leakage and model drift, a linear fit of
-    the corresponding pre-industrial control run is subtracted from the
-    4x\acs{CO2} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope of
-    the linear regression (black line) corresponds to the climate feedback
+  \caption{Global and annual mean net \acf{TOA} radiation $N$ \vs{} the change
+    in global and annual mean near-surface air temperature $\Delta T$ for 150
+    years of a simulation with an abrupt quadrupling of the atmospheric
+    \ac{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 multi-model mean
+    (blue circles). To account for energy leakage and model drift, a linear
+    fit of the corresponding pre-industrial control run is subtracted from the
+    \nxcotwo{4} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope
+    of the linear regression (black line) corresponds to the climate feedback
     parameter $\lambda$ and the y-intercept corresponds to the radiative
-    forcing $F_\text{4x}$. These can be used to calculate the \ac{ECS} with
-    the Gregory regression method following to \cref{eq:02:ecs}
-    \autocite{Gregory2004}.}
+    forcing $F_\text{4x}$. These can be used to calculate the \acf{ECS} with
+    the Gregory regression method according to \cref{eq:02:ecs}
+    \autocite{Gregory2004}. \acs{ECS} is equivalently given by the x-intercept
+    of the linear regression line divided by $2$. Here, $\text{ECS} = 3.14
+    \unit{K}$.}
   \label{fig:02:gregory_regression}
@@
 simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
-(4x\acs{CO2}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
+(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
 is the simple energy balance model introduced in
@@
 from the external forcing $F$ and the climate feedback parameter $\lambda$. The
-steady state values for $F$ and $\lambda$ can be estimated from a 4x\acs{CO2}
+steady state values for $F$ and $\lambda$ can be estimated from a \nxcotwo{4}
 simulation that is not in equilibrium by extrapolation with a linear regression
@@
 change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
-4x\acs{CO2} run are plotted (see \cref{fig:02:gregory_regression}). To account
+\nxcotwo{4} run are plotted (see \cref{fig:02:gregory_regression}). To account
 for energy leakage and remove any model drift present in the control climate, a
 linear fit of the corresponding pre-industrial control run is subtracted from
-the 4x\acs{CO2} simulation beforehand \autocite{Andrews2012}. Since the
+the \nxcotwo{4} simulation beforehand \autocite{Andrews2012}. Since the
 combination of \cref{eq:02:N,eq:02:linear_response} yields
@@
 \begin{equation}
-  \text{ECS} = - \frac{F_\text{4x}}{2 \lambda}.
+  \text{ECS} = -\frac{F_\text{4x}}{2 \lambda}.
   \label{eq:02:ecs}
@@
 Gregory regression is not constant, but rather decreases over time when a long
-running 4x\acs{CO2} simulation with more than 1000 model years is considered.
+running \nxcotwo{4} simulation with more than 1000 model years is considered.
 As a result, the climate feedback parameter $\lambda$ decreases over time,
@@
 between the Equilibrium Climate Sensitivity and \ac{ECS} is the use of a
-4x\acs{CO2} instead of a 2x\acs{CO2} simulation. The factor of $2$ in the
+\nxcotwo{4} instead of a \nxcotwo{2} simulation. The factor of $2$ in the
 denominator of \cref{eq:02:ecs} only partly compensates this difference since
@@
 contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
-the climate system but describes the transient response of an evolving climate.
-\ac{TCR} is defined as the change in the annual mean \ac{GSAT} at the time of
-\ac{CO2} doubling in a simulation with a \pct{1} per year increase of the
-atmospheric \ac{CO2} concentration (1\%\acs{CO2}) averaged over a 20 year
-period centered at the time of the doubling (years \range{61}{80} when the
-first year is indexed with 1) \autocite{Bindoff2013}. To account for model
-drift, the annual mean changes in \ac{GSAT} are calculated relative to a
-corresponding pre-industrial control simulation smoothed with a linear fit that
-considers 140 model years (length of the 1\%\acs{CO2} simulation).
+the Earth system but describes the transient response of an evolving climate.
+Following \textcite{Bindoff2013}, \ac{TCR} is defined as the change in the
+\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a \pct{1} per
+year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}). For
+this, the annual mean \acp{GSAT} are averaged over a 20 year period centered at
+the time of the \ac{CO2} doubling (years \range{61}{80} when the first year is
+indexed with 1). To account for model drift, the annual mean changes in
+\ac{GSAT} are calculated relative to a corresponding pre-industrial control
+simulation smoothed with a linear fit that considers 140 model years (length of
+the \onepctcotwo{} simulation). An illustration of that calculation is shown in
+\cref{fig:02:tcr}.
 
-- ocean heat uptake (TCR)
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/tcr_calculation.pdf}
+  \caption{Global and annual mean near-surface air temperature change $\Delta
+    T$ for 140 years of a simulation with a \pct{1} per year increase of the
+    atmospheric \ac{CO2} concentration (\onepctcotwo{}) for the \acf{CESM} of
+    the \acf{NCAR} (blue circles). The temperature change is calculated
+    relative to a corresponding pre-industrial control simulation smoothed
+    with a linear fit over all 140 years. The \acf{TCR} is defined as $\Delta
+    T$ at the time of \acs{CO2} doubling averaged over a 20 year period
+    (illustrated by the vertical dashed lines). Here, $\text{TCR} = 2.35
+    \unit{K}$ (horizontal dashed line).}
+  \label{fig:02:tcr}
+\end{figure}
 
-- residual heat -> 90\% go into ocean (Knutti et al. 2017)
+Similar to \ac{ECS}, \ac{TCR} can also be defined in terms of an external
+forcing and climate feedbacks. For this, the energy balance
+\cref{eq:02:N_vs_delta_t} can be slightly adjusted. Since over \pct{90} of the
+excess energy introduced into the system by the radiative forcing $F$ is taken
+up by the ocean due to its large heat capacity, $N$ can be taken equal to the
+global ocean heat uptake \autocite{Knutti2017}. In experiments with a steadily
+increasing radiative forcing, which is the case for the \onepctcotwo{}
+simulation, this ocean heat uptake can be approximated with
+\begin{equation}
+  N = \kappa \cdot \Delta T,
+  \label{eq:02:ocean_heat_uptake}
+\end{equation}
+where $\kappa$ is the ocean heat uptake efficiency \autocite{Gregory2008}.
+$\kappa > 0$, since there is a net energy flux into the climate system ($N >
+0$) due to the external forcing $F$. This approximation becomes less accurate
+as the deeper ocean warms up and cannot be applied to simulations with a steady
+state climate change ($N \to 0$) where the external forcings tends stabilize
+\autocite{Gregory2009}. By applying the definition of \ac{TCR} (transient
+\ac{GSAT} change $\Delta T$ at the time of \ac{CO2} doubling) and combining
+\cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be estimated as
+\begin{equation}
+  \text{TCR} = \frac{F_\text{2x}}{\kappa - \lambda},
+  \label{eq:02:tcr}
+\end{equation}
+where $F_\text{2x}$ is the radiative forcing induced by a doubling of the
+atmospheric \ac{CO2} concentration.
 
-- equations for TCR
+This equation can be used to derive a relationship between \ac{TCR} and
+\ac{ECS}. Writing $\text{ECS} = -F_\text{2x} / \lambda$ and assuming path
+independence of the forcing (\ie{} the resulting radiative forcing from the
+\onepctcotwo{} and \nxcotwo{2} runs are comparable) gives
+\begin{equation}
+  \text{TCR} = \frac{F_\text{2x} \cdot \text{ECS}}{F_\text{2x} + \kappa \cdot
+    \text{ECS}},
+  \label{eq:02:tcr_vs_ecs}
+\end{equation}
+which demonstrates the non-linear connection between \ac{TCR} and \ac{ECS} \autocite{Gregory2008, Nijsse2020}.
 

### 2020-12-11 ch01_introduction/introduction.tex
Added climate feedback figure and adapted margins
Added overview of climate feedbacks
Started section on techniques for reducing uncertainties in climate model projections

@@
 
+- definition of radiative forcing
+
 - well-mixed greenhouse gases (water vapor, CO2, CH4) + sources
@@
 
-\section{Structure of the thesis}
+\section{Structure of the Thesis}
 \label{sec:01:structure}
@@
 \cref{ch:05:paper_ecs} describes the assessment of emergent constraints (a
-technique to reduce uncertainties in climate model projections, see
-\vref{sec:02:techniques}) on the \ac{ECS} for these \acp{ESM}. The contents of
-this chapter are published in \emph{Earth System Dynamics}
+technique to reduce uncertainties in multi-model climate projections, see
+\vref{subsec:02:emergent_constraints}) on the \ac{ECS} for these \acp{ESM}. The
+contents of this chapter are published in \emph{Earth System Dynamics}
 \autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
-reduce climate model uncertainties based on \ac{ML}. As an example, the method
-is applied to the photosynthesis rate at the end of the \nth{21} century, which
-is already published in the \emph{Journal of Geophysical Research:
-  Biogeosciences} \autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook}
-provides a summary of the results of this thesis and gives an outlook of
-possible future works.
+reduce uncertainties in multi-model projections of the future climate based on
+\ac{ML}. As an example, the method is applied to the photosynthesis rate at the
+end of the \nth{21} century, which is already published in the \emph{Journal of
+  Geophysical Research: Biogeosciences} \autocite{Schlund2020}. Finally,
+\cref{ch:07:summary_outlook} provides a summary of the results of this thesis
+and gives an outlook of possible future works.

### 2020-12-11 ch02_scientific_background/scientific_background.tex
Started section on techniques for reducing uncertainties in climate model projections

@@
 \cref{eq:02:N_vs_delta_t} can be slightly adjusted. Since over \pct{90} of the
-excess energy introduced into the system by the radiative forcing $F$ is taken
-up by the ocean due to its large heat capacity, $N$ can be taken equal to the
-global ocean heat uptake \autocite{Knutti2017}. In experiments with a steadily
-increasing radiative forcing, which is the case for the \onepctcotwo{}
+excess energy introduced into the climate system by the radiative forcing $F$
+is taken up by the ocean due to its large heat capacity, $N$ can be taken equal
+to the global ocean heat uptake \autocite{Knutti2017}. In experiments with a
+steadily increasing radiative forcing, which is the case for the \onepctcotwo{}
 simulation, this ocean heat uptake can be approximated with
@@
 where $\kappa$ is the ocean heat uptake efficiency \autocite{Gregory2008}.
-$\kappa > 0$, since there is a net energy flux into the climate system ($N >
-0$) due to the external forcing $F$. This approximation becomes less accurate
-as the deeper ocean warms up and cannot be applied to simulations with a steady
-state climate change ($N \to 0$) where the external forcings tends stabilize
-\autocite{Gregory2009}. By applying the definition of \ac{TCR} (transient
-\ac{GSAT} change $\Delta T$ at the time of \ac{CO2} doubling) and combining
-\cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be estimated as
+Since there is a net energy flux into the climate system ($N > 0$) due to the
+external forcing $F$, $\kappa$ is positive. This approximation becomes less
+accurate as the deeper ocean warms up and cannot be applied to simulations with
+a steady state climate change ($N \to 0$) where the external forcing tends
+stabilize \autocite{Gregory2009}. By applying the definition of \ac{TCR}
+(transient \ac{GSAT} change $\Delta T$ at the time of \ac{CO2} doubling) and
+combining \cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be
+estimated as
 \begin{equation}
@@
 \end{equation}
-which demonstrates the non-linear connection between \ac{TCR} and \ac{ECS} \autocite{Gregory2008, Nijsse2020}.
+which demonstrates the non-linear connection between \ac{TCR} and \ac{ECS}
+\autocite{Gregory2008, Nijsse2020}. Since $\lambda < 0$ and $\kappa > 0$,
+\cref{eq:02:tcr} implies that the equilibrium response \ac{ECS} is (as
+expected) larger than transient response \ac{TCR}, \ie{} $\text{ECS} >
+\text{TCR}$.
 
@@
 
-\section{Techniques to Reduce Uncertainties in Climate Model Projections}
-\label{sec:02:techniques}
+\section{Techniques to Reduce Uncertainties in Multi-Model Climate Projections}
+\label{sec:02:techniques_to_reduce_uncertainties}
+
+As shown in \cref{subsec:02:source_of_uncertainties}, projections of the future
+climate are always associated with uncertainties. In the context of this
+thesis, the most relevant source for these is the climate response uncertainty.
+It originates from the different responses of the different climate models in a
+multi-model ensemble to a given forcing and affects all multi-model projections
+of the future climate (\eg{} the climate sensitivity or the future development
+of the global carbon cycle). This section introduces three state-of-the-art
+techniques to reduce uncertainties in multi-model climate model projections
+with observations. The presented methods form the baseline for the work
+presented in this thesis.
+
+\subsection{Emergent Constraints}
+\label{subsec:02:emergent_constraints}
+
+- equations for emergent constraints, see Schlund ESD, following Cox 2013
+
+- Flat prior, see Schlund ESD
+
+- Wenzel et al., Nature (2016) as example (refer to carbon cycle chapter)
+
+- Review of ECS emergent constraints is given in chapter on Schlund et al., ESD
+(2020)
 
@@
 
-- Process-based weighting (Knutti + MDER)
+\subsection{Performance- and Interdependence-based Weighting of Climate Models}
+\label{subsec:02:model_weighting}
 
-- emergent constraints
+- Knutti, Sanderson
 
-- Wenzel et al., Nature (2016) as example (refer to carbon cycle chapter)
+- recent studies: Brunner (T), Paul Young (Ozone), Lorenz?
 
-- Review of ECS emergent constraints is given in chapter on Schlund et al., ESD
-(2020)
+- Lorenz (2018): reasoning for non-equal weighting, Collins 2013 as bad example
+
+- Basic assumption: Models good in past -> good in future (PERFORMANCE)
+
+- Models share code (INTERDEPENDENCE)
+
+- Review by Abramowitz
+
+- Theory: Knutt, p.1910
+
+\subsection{\acl{MDER}}
+\label{subsec:02:mder}
+
+- some equations for MDER
+
+- Karpechko, Wenzel, Senftleben
+
+- basic assumption (in contrast to weighting), see Schlund JGR 2020

### 2020-12-13 ch02_scientific_background/scientific_background.tex
Added section description for section on reducing uncertainties

@@
 
-\section{Techniques to Reduce Uncertainties in Multi-Model Climate Projections}
+\section{Techniques to Reduce Uncertainties in Multi-Model Climate Projections
+  with Observations}
 \label{sec:02:techniques_to_reduce_uncertainties}
@@
 climate are always associated with uncertainties. In the context of this
-thesis, the most relevant source for these is the climate response uncertainty.
-It originates from the different responses of the different climate models in a
-multi-model ensemble to a given forcing and affects all multi-model projections
-of the future climate (\eg{} the climate sensitivity or the future development
-of the global carbon cycle). This section introduces three state-of-the-art
-techniques to reduce uncertainties in multi-model climate model projections
-with observations. The presented methods form the baseline for the work
-presented in this thesis.
+thesis, the most relevant source of uncertainty is the climate response
+uncertainty. It originates from our imperfect knowledge on how the climate
+system will respond to external forcing. In a multi-model ensemble (\eg{} from
+\ac{CMIP}), this is expressed as the different responses of the different
+climate models to a given forcing. A common approach to distill information
+about a projected quantity from multi-model ensembles is to treat the
+unweighted \ac{MMM} and the multi-model range as best estimate and uncertainty
+measure of this quantity \autocite{Collins2013}. This \emph{one model one vote}
+or \emph{model democracy} approach basically assumes that all climate models
+are independent, equally plausible, distributed around reality and that the
+projected multi-model range is representative for the uncertainty in the
+projected quantity \autocite{Knutti2017a}. However, since the \ac{CMIP}
+ensembles, sometimes referred to as \emph{ensembles of opportunity}, have not
+been designed to represent a true statistical sample of the reality composed of
+independent climate models \autocite{Tebaldi2007}, these assumptions do not
+hold in practice. The main reasons for this are that different climate models
+(even for different modeling institutions) share parts of their code
+\autocite{Knutti2013}, that models do not equally well represent the observed
+past and present-day climate \autocite{Knutti2013} and that models might suffer
+from common structural limitations like missing processes
+\autocite{Knutti2017a}.
+
+Thus, more sophisticated techniques are necessary to evaluate multi-model
+climate projections. This section introduces three state-of-the-art methods to
+assess multi-model projections and reduce associated uncertainties with
+observations. These techniques form the baseline of the work presented in the following chapters.
 
@@
 
+From Knutti (2017): A way out, at least partially, is to reframe the
+competition of “which model is the best” to “which models are adequate for
+predicting X” [Parker, 2009]
+
 - some equations for MDER

### 2020-12-14 ch02_scientific_background/scientific_background.tex
Added section on emergent constraints

@@
 2100 while \ac{SSP}1-2.6 represents a sustainable future with a radiative
-forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two other main scenarios
-(called \emph{Tier 1} experiments in \ac{ScenarioMIP}) are the \ac{SSP}2-4.5
-and \ac{SSP}3-7.0 scenarios. In contrast to the \ac{ScenarioMIP} experiments,
-the corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012}, the so-called
-\acp{RCP}, only used the radiative forcing in 2100 as only dimension to
-describe the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5, \ac{RCP}2.6,
-\etc{}).
+forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two scenarios in the
+main category of \ac{ScenarioMIP}, the \emph{Tier 1} experiments, are the
+\ac{SSP}2-4.5 and \ac{SSP}3-7.0 scenarios. In contrast to the \ac{ScenarioMIP}
+experiments, the corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012},
+the so-called \acp{RCP}, only used the radiative forcing in 2100 as only
+dimension to describe the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5,
+\ac{RCP}2.6, \etc{}).
 
@@
 \begin{equation}
-  J_0 + R = -\sigma \left( T_0 + \Delta T \right)^4
+  J_0 + R = -\sigma \left( T_0 + \Delta T \right)^4.
   \label{eq:02:stefan_boltzmann_law_non_eq}
@@
   -\sigma \left( T_0 + \Delta T \right)^4 \approx J_0 - 4 \sigma T_0^3 \cdot
-  \Delta T
+  \Delta T.
   \label{eq:02:stefan_boltzmann_law_lin}
@@
 \begin{equation}
-  R = -4 \sigma T_0^3 \cdot \Delta T \coloneqq \lambda_\text{BB} \cdot \Delta T
+  R = -4 \sigma T_0^3 \cdot \Delta T \coloneq \lambda_\text{BB} \cdot \Delta T
   \label{eq:02:linear_blackbody_response}
@@
 \begin{equation}
-  \lambda = \sum_i \lambda_i
+  \lambda = \sum_i \lambda_i.
   \label{eq:02:lambda_as_sum_of_lambdas}
@@
 
-\section{Techniques to Reduce Uncertainties in Multi-Model Climate Projections
+\section{Reducing Uncertainties in Multi-Model Climate Projections
   with Observations}
-\label{sec:02:techniques_to_reduce_uncertainties}
+\label{sec:02:reducing_uncertainties}
 
@@
 assess multi-model projections and reduce associated uncertainties with
-observations. These techniques form the baseline of the work presented in the following chapters.
+observations. These techniques form the baseline of the work presented in the
+following chapters.
 
@@
 
-- equations for emergent constraints, see Schlund ESD, following Cox 2013
+As indicated at the beginning of this section, one main issue of multi-model
+ensembles is that not all participating climate models are equally plausible.
+Usually, this is quantified with some kind of measure of the models'
+\emph{performance}, \ie{} their agreement with observations of the real climate
+system. However, this model performance can only be evaluated against
+observations of the past and present-day climate, which does not necessarily
+provide insights into the quality of model projections of the future climate.
 
-- Flat prior, see Schlund ESD
+The \emph{emergent constraint} approach tackles this problem by
+\enquote{identifying robust, physically interpretable relationships between
+  Earth system feedback behaviors on short, well-observed time scales and on
+  time scales that span the twenty-first century and beyond}
+\autocite{Eyring2019}. An illustration of the concept of emergent constraints
+is shown in \cref{fig:02:emergent_constraint}. Each emergent constraint
+requires two key components: an \emph{emergent relationship} and a
+corresponding observation of the real world \autocite{Eyring2019}. The emergent
+relationship (red line in \cref{fig:02:emergent_constraint}) is a robust and
+physically interpretable inter-model relationship between a target variable $y$
+related to the future climate and an observable $x$ of the past or present-day
+climate. Basis for the relationship is output of the different climate models
+of a multi-model ensemble (blue circles in \cref{fig:02:emergent_constraint}).
+Using an observation of $x$, this emergent relationship can then be used to
+derive a emergent constraint on $y$ (gray shaded area in
+\cref{fig:02:emergent_constraint}) that considers uncertainties in the emergent
+relationship itself (red shaded area in \cref{fig:02:emergent_constraint}) and
+uncertainties in the observation (blue shaded area in
+\cref{fig:02:emergent_constraint}).
 
-- Wenzel et al., Nature (2016) as example (refer to carbon cycle chapter)
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/emergent_constraint.pdf}
+  \caption{Schematic illustration of the emergent constraint approach. Basis of
+    every emergent constraint is a robust and physically interpretable emergent
+    relationship (red line) between a target variable $y$ (\eg{} Earth system
+    sensitivity or projection of future climate change) and an observable $x$
+    (\eg{} past or present-day trend or variation) for the climate models of a
+    multi-model ensemble (blue circles). With an observation of $x$,
+    uncertainties in $y$ in the multi-model ensemble, illustrated by the yellow
+    \acf{PDF}, can be reduced (gray \acs{PDF}). Uncertainties in the target
+    variable $y$ arise from two sources: uncertainties in the observation (blue
+    shaded area) and uncertainties in the emergent relationship (red shaded
+    area). Taken and modified from \textcite{Eyring2019}.}
+  \label{fig:02:emergent_constraint}
+\end{figure}
 
-- Review of ECS emergent constraints is given in chapter on Schlund et al., ESD
-(2020)
+One possible mathematical framework for the evaluation of emergent constraints
+is based on linear regression and Gaussian probability densities
+\autocite{Cox2013, Cox2018}. Let $x_m$ be the observable predictor variable for
+climate model $m$ and $y_m$ the corresponding target variable. To find the
+linear emergent relationship for the climate model ensemble $\left\{ \left(
+x_m, y_m \right) \right\}$, a linear regression model
+\begin{equation}
+  \hat{y}(x) = a + b x
+  \label{eq:02:linear_regression_y}
+\end{equation}
+for the predicted target variable $\hat{y}$ with intercept $a$ and slope $b$ is
+used. Fitting this regression line with ordinary least squares includes
+minimizing the standard error $s$ of the estimate
+\begin{equation}
+  s^2 = \frac{1}{M - 2} \sum_{m=1}^M \left( y_m - \hat{y}_m \right)^2,
+  \label{eq:02:sse}
+\end{equation}
+where $\hat{y}_m \coloneq \hat{y}(x_m)$ is the predicted target variable for
+climate model $m$ and $M$ is the total number of climate models. The
+uncertainty of the emergent relationship for a value $x$ that has not been used
+to fit the regression line is given by the standard prediction error
+$\sigma_{\hat{y}}(x)$:
+\begin{equation}
+  \sigma_{\hat{y}}^2(x) = s^2 \left[ 1 + \frac{1}{M} + \frac{\left( x - \bar{x}
+  \right)^2}{\sum_{m=1}^M \left( x_m - \bar{x} \right)^2} \right].
+  \label{eq:02:spe}
+\end{equation}
+Here, $\bar{x}$ indicates the arithmetic mean of $x$ over all climate models.
+Assuming Gaussian errors and a mean of $\hat{y}(x)$ (\ie{} the best estimate of
+the target variable $y$ is given by the regression line), \cref{eq:02:spe} can
+be used to define a conditional \acf{PDF} for predicting a target variable of
+$y$ given $x$:
+\begin{equation}
+  P(y|x) = \frac{1}{\sqrt{2 \pi \sigma_{\hat{y}}^2(x)}} \exp \left[
+  -\frac{\left( y - \hat{y}(x) \right)^2}{2 \sigma_{\hat{y}}^2(x)} \right].
+  \label{eq:02:pdf_y_given_x}
+\end{equation}
+This distribution describes the uncertainty in the emergent relationship itself
+introduced by the imperfect alignment of the climate model data (red shaded
+area in \cref{fig:02:emergent_constraint}). Its maximum is given by the
+emergent relationship itself (red line in \cref{fig:02:emergent_constraint}).
+The conditional \ac{PDF} can be interpreted as the posterior distribution of
+the regression model based on the climate model output but constrained on the
+observable $x$. However, the observed value of $x$, called $x_0$, also has
+uncertainties associated with it (blue shaded are in
+\cref{fig:02:emergent_constraint}). Assuming again a Gaussian distribution, the
+observational \ac{PDF} for observing $x_0$ given the true value $x$ can be
+written as
+\begin{equation}
+  P(x_0|x) = \frac{1}{\sqrt{2 \pi \sigma_x^2}} \exp \left[ -\frac{\left( x_0 -
+  x \right)^2}{2 \sigma_x^2} \right],
+  \label{eq:02:pdf_obs}
+\end{equation}
+where $\sigma_x$ is the standard deviation of the observation around the true
+value. Assuming an imperfect uniform prior $P(x) \propto 1$ with cut-offs at
+$-\infty$ and $+\infty$ and using Bayes' theorem implies $P(x_0|x) = P(x|x_0)$.
+In a final step, this can be used to calculate the posterior \ac{PDF} for the
+constrained prediction of the target variable $y$ given the observation $x_0$
+(gray \ac{PDF} in \cref{fig:02:emergent_constraint}) with numerical
+integration:
+\begin{equation}
+  P(y|x_0) = \int_{-\infty}^{+\infty} P(y|x) \, P(x|x_0) \dd{}x.
+  \label{eq:02:pdf_y_given_x0}
+\end{equation}
+Posterior estimates of the target variable are influenced by the way the
+statistical inference has been performed. Alternative methods that can be used
+include Bayesian frameworks \autocite{Renoult2020}, information theoretic
+approaches based on the Kullback–Leibler divergence between the models'
+\acp{PDF} of $x$ and the observational \ac{PDF} \autocite{Brient2016} and
+linear regression models based on hierarchical Bayesian models
+\autocite{Nijsse2020}. However, no consensus has yet been found for this
+statistical inference \autocite{Brient2020}.
 
-- Problems with multi-model analysis (search for refs in Schlund et al., ESD,
-2020) -> models are not independent, multi-model ensemble is not "real"
-statistical sample
+One limitation of the presented approach is the assumption that the individual
+data points from the different climate models are independent. As already noted
+in the beginning of \cref{sec:02:reducing_uncertainties}, this is not the case
+for typical climate models ensembles, as some modeling groups provide output
+for multiple climate models and some climate models from different modeling
+institutions share components and code \autocite{Knutti2013}. The duplicated
+code in the different climate models leads to an overestimation of the sample
+size and may result in spurious correlations \autocite{Sanderson2015}. Possible
+approaches to tackle this problem are presented in
+\cref{subsec:02:model_weighting} and include a weighting of the climate models
+based on their degree of interdependence \autocite{Knutti2017a, Sanderson2015,
+  Sanderson2017}. A further limitation of this approach is the use of an
+ordinary least squares linear regression model. This is not always appropriate,
+for example when a non-linear emergent relationship is expected
+\autocite{Nijsse2020} or when additional physical considerations further
+constrain the regression model, \eg{} by demanding a zero intercept ($a = 0$)
+\autocite{Annan2020, JimenezdelaCuesta2019}. Moreover, using only a single
+observational dataset to estimate $x_0$ and $\sigma_x$ when different datasets
+are available might lead to an underestimation of the observational
+uncertainty, as different observational datasets might lead to different
+emergent constraints.
+
+A crucial aspect for every emergent constraint is a verifiable physical process
+explaining the correlation between $x$ and $y$ \autocite{Hall2019}. Only if the
+underlying emergent relationship can be derived from a robust and plausible
+physical mechanism, an emergent constraint can be considered credible. The
+reason for this are spurious relationships: Due to the large number of possible
+observables provided by modern \acp{ESM} and the comparatively small number of
+climate models, spurious relationships are possible just by chance
+\autocite{Caldwell2014}. Furthermore, out-of-sample tests offer an important
+tool to evaluate the credibility of emergent constraints \autocite{Hall2019}.
+These ensure that the existence of an emergent relationship is not limited to a
+certain climate model ensemble and might indicate that the relationship is also
+valid for the true climate system. Testing emergent constraints in different
+\ac{CMIP} generations offers a straightforward setup for out-of-sample testing
+\autocite{Caldwell2018}, which is discussed in more detail in
+\cref{ch:05:paper_ecs}, where eleven emergent constraints on \ac{ECS} are
+evaluated on the new \acs{CMIP}6 ensemble.
+
+In the last two decades, many emergent constraints on various aspects of the
+Earth system have been published. Early studies tackled the hydrological cycle
+\autocite{Allen2002} and the snow-albedo feedback \autocite{Hall2006}. Over the
+years, the climate sensitivity expressed by the \ac{ECS} has been a prominent
+target variable. Since cloud feedbacks are a major source of uncertainty for
+climate sensitivity, a variety of papers focus on constraining \ac{ECS} with
+cloud-related processes \autocite{Brient2015, Brient2016, Fasullo2012,
+  Lipat2017, Sherwood2014, Su2014, Tian2015, Volodin2008, Qu2013, Zhai2015},
+which are discussed in detail in
+\cref{sec:05:comparison_of_emergent_constraints}. More recent studies aim to
+constrain \ac{ECS} with the historical temperature variability
+\autocite{Cox2018} or the historical warming trend
+\autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}. Emergent
+constraints are not only limited to physical processes, but can also be applied
+to other domains, like the global carbon cycle. For example, \textcite{Cox2013}
+and \textcite{Wenzel2014} provide an emergent constraint on the terrestrial
+climate-carbon feedback, \textcite{Wenzel2016} and \textcite{Winkler2019} on
+the terrestrial concentration-carbon feedback. An extensive discussion on the
+emergent constraint by \textcite{Wenzel2016} is given in \cref{sec:06:step_1}.
+A further example of an emergent constraints on the carbon cycle focus on
+primary production in the tropical oceans \autocite{Kwiatkowski2017}.
 
@@
 
-- Theory: Knutt, p.1910
+- Theory: Knutti, p.1910
 
@@
 \label{subsec:02:mder}
+
+I do not like \ac{MDER}!
 
@@
 - basic assumption (in contrast to weighting), see Schlund JGR 2020
+
+- Difference to standard model weighting: weaker assumption (see Karpechko
+2013)
+
+- Disadvantages: no interdependence, no errors in OBS

### 2020-12-15 ch02_scientific_background/scientific_background.tex
Added subsection on Knutti weighting

@@
 about a projected quantity from multi-model ensembles is to treat the
-unweighted \ac{MMM} and the multi-model range as best estimate and uncertainty
-measure of this quantity \autocite{Collins2013}. This \emph{one model one vote}
-or \emph{model democracy} approach basically assumes that all climate models
-are independent, equally plausible, distributed around reality and that the
-projected multi-model range is representative for the uncertainty in the
-projected quantity \autocite{Knutti2017a}. However, since the \ac{CMIP}
-ensembles, sometimes referred to as \emph{ensembles of opportunity}, have not
-been designed to represent a true statistical sample of the reality composed of
-independent climate models \autocite{Tebaldi2007}, these assumptions do not
-hold in practice. The main reasons for this are that different climate models
-(even for different modeling institutions) share parts of their code
-\autocite{Knutti2013}, that models do not equally well represent the observed
-past and present-day climate \autocite{Knutti2013} and that models might suffer
-from common structural limitations like missing processes
-\autocite{Knutti2017a}.
+arithmetic \ac{MMM} and the multi-model range as best estimate and uncertainty
+measure of this quantity \autocite{Collins2013}. This \emph{model democracy}
+approach basically assumes that all climate models are independent, equally
+plausible, distributed around reality and that the projected multi-model range
+is representative for the uncertainty in the projected quantity
+\autocite{Knutti2017a}. However, since the \ac{CMIP} ensembles, sometimes
+referred to as \emph{ensembles of opportunity}, have not been designed to
+represent a true statistical sample of the reality composed of independent
+climate models \autocite{Tebaldi2007}, these assumptions do not hold in
+practice. The main reasons for this are that different climate models (even for
+different modeling institutions) share parts of their code
+\autocite{Abramowitz2019, Knutti2013}, that models do not equally well
+represent the observed past and present-day climate \autocite{Gleckler2008,
+  Knutti2013} and that models might suffer from common structural limitations
+like missing processes \autocite{Knutti2017a}.
 
@@
 emergent constraint by \textcite{Wenzel2016} is given in \cref{sec:06:step_1}.
-A further example of an emergent constraints on the carbon cycle focus on
+A further example of an emergent constraints on the carbon cycle focuses on
 primary production in the tropical oceans \autocite{Kwiatkowski2017}.
@@
 
-- Knutti, Sanderson
-
-- recent studies: Brunner (T), Paul Young (Ozone), Lorenz?
-
-- Lorenz (2018): reasoning for non-equal weighting, Collins 2013 as bad example
-
-- Basic assumption: Models good in past -> good in future (PERFORMANCE)
-
-- Models share code (INTERDEPENDENCE)
-
-- Review by Abramowitz
-
-- Theory: Knutti, p.1910
+An alternative approach to emergent constraints are climate model weighting
+schemes. Their basic idea is to abandon model democracy by replacing the
+arithmetic mean used to calculate the \acp{MMM} by a weighted mean of the form
+\begin{equation}
+  y = \sum_{m=1}^{M} w_m y_m
+  \label{eq:02:weighted_mean}
+\end{equation}
+with normalized weights $w_m$. Similar to the notation introduced in the
+previous section, $y$ is a target variable (\eg{} a projection of the future
+climate) and $m$ indexes the $M$ different climate models. To address two major
+issues of model democracy (different climate models are not equally plausible
+and not independent; see beginning of \cref{sec:02:reducing_uncertainties}),
+\textcite{Knutti2017a} propose a weighting scheme based on climate model
+performance and interdependence with the following weights:
+\begin{equation}
+  w_m \propto \frac{\exp \left( -\frac{D_m^2}{\sigma_D^2} \right)}{1 + \sum_{n
+    \neq m}^{M} \exp \left( -\frac{S_{mn}^2}{\sigma_S^2} \right)}.
+  \label{eq:02:knutti_weights}
+\end{equation}
+The metric $D_m$ describes the distance between climate model $m$ and
+observations (= model performance) and the metric $S_{mn}$ describes the
+distance between climate model $m$ and $n$ (= model interdependence).
+$\sigma_D$ and $\sigma_S$ are constants that determine the individual strength
+of the performance and interdependence weighting, respectively.
+
+A commonly used distance metric to measure model performance $D_m$ and model
+interdependence $S_{mn}$ is the \ac{RMSE}, but others are possible
+\autocite{Knutti2017a}. The metrics are evaluated on a set of past or
+present-day diagnostics and variables, whose choice is crucial for the
+weighting scheme. A helpful strategy for this is to focus on addressing the
+question \enquote{which climate model is adequate for predicting the target
+  variable $y$?} instead of trying to answer the question \enquote{which
+  climate model is the best?} \autocite{Parker2009}. Thus, diagnostics and
+variables are chosen that are relevant for the projection of the target
+variable \autocite{Knutti2017a}. In practice, this choice is either based on
+expert judgment about relevant processes, on emergent relationships (see
+\cref{subsec:02:emergent_constraints}) or on multivariate regression models
+\commentcite{Karpechko2013, Sanderson2015a}{see \cref{subsec:02:mder}}. It
+might also be beneficial to use different diagnostics for the calculations of
+the performance and interdependence metrics \autocite{Merrifield2020} and/or to
+remove selected diagnostics based on their mutual correlation
+\autocite{Lorenz2018}.
+
+The constants $\sigma_D$ and $\sigma_S$ determine how strongly the climate
+models' performance and interdependence are weighted \autocite{Knutti2017a}.
+Small values of the performance parameter $\sigma_D$ lead to an aggressive
+weighting with only a few climate models receiving a majority of the weight,
+while large values of $\sigma_D$ result in an equal weighting. For the
+interdependence parameter $\sigma_S$, this is slightly different: Here, small
+(all climate models are independent) and large (all climate models are
+dependent) values lead to an almost equal weighting. Thus, an optimal choice
+for $\sigma_D$ and $\sigma_S$ is crucial. A useful tool to estimate these
+optimal parameters is the \emph{leave-one-model-out \ac{CV}} approach, which is
+also known as \emph{pseudo-reality}, \emph{model-as-truth} or \emph{perfect
+  model} setup \autocite{Elia2002, Karpechko2013}. For this, a single climate
+model is removed from the multi-model ensemble and treated as observation
+(\emph{pseudo-observation}). Then, a weighted \ac{MMM} with weights computed
+from the updated model ensemble is calculated, which gives a prediction for the
+\enquote{true} climate model. This allows a simple quantitative assessment of
+the weighting scheme by calculating the \ac{RMSE} between the prediction and
+the known ground truth of the pseudo-observation. The whole process is repeated
+for every climate model of the ensemble to get a statistical distribution of
+\acp{RMSE}. Finally, different \ac{RMSE} distributions calculated from
+different parameters $\sigma_D$ and $\sigma_S$ can be assessed using specific
+criteria to find optimal values for $\sigma_D$ and $\sigma_S$
+\autocite{Knutti2017a}. Furthermore, the leave-one-model-out \ac{CV} approach
+can be used to evaluate different climate model weightings schemes (including
+the unweighted \ac{MMM}) and compare them against each other.
+
+Due to its flexibility, the climate model weighting scheme given by
+\cref{eq:02:knutti_weights} has already been used for various target variables:
+Arctic sea ice \autocite{Knutti2017a}, Antarctic ozone concentrations
+\autocite{Amos2020}, North American maximum temperature \autocite{Lorenz2018},
+European temperature and precipitation \autocite{Brunner2019, Merrifield2020}
+and global warming over the \nth{21} century \autocite{Brunner2020,
+  Liang2020a}.
 

### 2020-12-16 ch02_scientific_background/scientific_background.tex
Started section on MDER

@@
   R = \lambda \cdot \Delta T = \left( \lambda_\text{BB} + \lambda_\text{WV} +
-  \lambda_\text{Albedo} + \lambda_\text{Cloud} + \ldots \right) \cdot \Delta T.
+  \lambda_\text{Albedo} + \lambda_\text{Cloud} + \dots \right) \cdot \Delta T.
   \label{eq:02:linear_response}
@@
 \begin{equation}
-  \hat{y}(x) = a + b x
+  \hat{y}(x) = b_0 + b_1 x
   \label{eq:02:linear_regression_y}
 \end{equation}
-for the predicted target variable $\hat{y}$ with intercept $a$ and slope $b$ is
-used. Fitting this regression line with ordinary least squares includes
-minimizing the standard error $s$ of the estimate
+for the predicted target variable $\hat{y}$ with intercept $b_0$ and slope
+$b_1$ is used. Fitting this regression line with ordinary least squares
+includes minimizing the standard error $s$ of the estimate
 \begin{equation}
@@
 \autocite{Nijsse2020} or when additional physical considerations further
-constrain the regression model, \eg{} by demanding a zero intercept ($a = 0$)
+constrain the regression model, \eg{} by demanding a zero intercept ($b_0 = 0$)
 \autocite{Annan2020, JimenezdelaCuesta2019}. Moreover, using only a single
@@
 constraints are not only limited to physical processes, but can also be applied
-to other domains, like the global carbon cycle. For example, \textcite{Cox2013}
-and \textcite{Wenzel2014} provide an emergent constraint on the terrestrial
-climate-carbon feedback, \textcite{Wenzel2016} and \textcite{Winkler2019} on
-the terrestrial concentration-carbon feedback. An extensive discussion on the
-emergent constraint by \textcite{Wenzel2016} is given in \cref{sec:06:step_1}.
-A further example of an emergent constraints on the carbon cycle focuses on
-primary production in the tropical oceans \autocite{Kwiatkowski2017}.
+to other domains, like the global carbon cycle \autocite{Cox2013,
+  Kwiatkowski2017, Wenzel2014, Wenzel2016, Winkler2019}. An extensive
+discussion on the emergent constraint by \textcite{Wenzel2016}, which focuses
+on the concentration-carbon feedback, is given in \cref{sec:06:step_1}.
 
@@
 
-An alternative approach to emergent constraints are climate model weighting
-schemes. Their basic idea is to abandon model democracy by replacing the
-arithmetic mean used to calculate the \acp{MMM} by a weighted mean of the form
+A further technique to reduce uncertainties in climate model projections with
+observations are model weighting schemes. Their basic idea is to abandon model
+democracy by replacing the arithmetic mean used to calculate the \acp{MMM} by a
+weighted mean of the form
 \begin{equation}
@@
 
-Due to its flexibility, the climate model weighting scheme given by
-\cref{eq:02:knutti_weights} has already been used for various target variables:
+The definition of the weights according to \cref{eq:02:knutti_weights} is based
+on reasonable and comprehensible principles. However, the exact form of the
+equation is purely subjective. Moreover, the additional freedom in choosing a
+suitable metric and optimal values for the parameters $\sigma_D$ and $\sigma_S$
+adds another level of subjectivity to the weighting scheme, which can partly be
+addressed with the introduced leave-one-model-out \ac{CV} setup. Nevertheless,
+due to its flexibility, the climate model weighting scheme of
+\textcite{Knutti2017a} has already been used for various target variables:
 Arctic sea ice \autocite{Knutti2017a}, Antarctic ozone concentrations
@@
 
-I do not like \ac{MDER}!
+An alternative approach to weighting climate models based on performance and
+interdependence metrics is a technique called \emph{\ac{MDER}}. Similar to all
+methods presented in \cref{sec:02:reducing_uncertainties}, it can be used to
+reduce uncertainties in climate model projections with observations. The basis
+of \ac{MDER} is a set of $K$ predictor diagnostics $\left\{ x^{(k)} \right\}$
+with $k \in \left\{ 1, 2, \dots, K \right\}$ which are relevant for the
+projection of the target variable $y$. The reasoning for this choice of the
+diagnostics is similar to the one presented in the previous section: Weighting
+schemes should address the question \enquote{which climate model is adequate
+  for predicting the target variable?} and not \enquote{which climate model is
+  the best?}.
+
+The concept of \ac{MDER} is mathematically similar to the concept of emergent constraints.
+In a first step, an inter-model relationship between the target variable $y$ and the process-relevant diagnostics $\left\{ x^{(k)} \right\}$ is used to fit a multivariate linear regression model
+\begin{equation}
+  \bm{X}
+\end{equation}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/mder_schematic.pdf}
+  \caption{Schematic illustration of the \acf{MDER} approach. First,
+    inter-model relationships between a target variable $y$ and multiple
+    process-based predictors $x^{(k)}$ (here: two predictors $x^{(1)}$ and
+    $x^{(2)}$) are used to fit a multivariate linear regression model (gray
+    plane). Second, observations of the predictors (horizontal thick black
+    lines) are fed into the regression model to calculate climate models
+    weights $w_m$ following \cref{eq:02:mder_weights}. The black circle
+    indicates the best estimate for the target variable $y$ given by the
+    observed values of $x^{(1)}$ and $x^{(2)}$. Each of the remaining colored
+    circles represents a single climate model of the multi-model ensemble. The
+    vertical dashed lines are auxiliary lines to visualize the distance
+    between the climate model data and the linear regression surface.}
+  \label{fig:02:mder}
+\end{figure}
+
+\begin{equation}
+  w_m = xD
+  \label{eq:02:mder_weights}
+\end{equation}
 
@@
 - Difference to standard model weighting: weaker assumption (see Karpechko
-2013)
+2013): The assumption that the regression between the present-day diagnostic
+and a future-climatic variable— derived from climate models—holds also for true
+cli- mate might seem to be weak at first sight, especially since it explicitly
+requires imperfect models away from observed climate to span the desired
+relationship. How- ever, a much weaker assumption is done in traditional
+model-weighting approaches: a model that simulates present-day climate better
+than another model neces- sarily better simulates future climate. Whereas
+traditional approaches do not show this relationship, ensemble re- gression
+establishes exactly such a relationship across a broad model ensemble.
+
+- Techniques for diagnostic selection: hypothesis testing, forward iteration,
+backwards elimination
 

### 2020-12-17 ch02_scientific_background/scientific_background.tex
Finished section on MDER

@@
   forcing in a multi-model context} \autocite{WCRP2020}. One major element of
-\ac{CMIP} is to establish common standards, coordination, infrastructure, and
+\ac{CMIP} is to establish common standards, coordination, infrastructure and
 documentation in order to facilitate the distribution of climate model output
@@
   R = \lambda \cdot \Delta T = \left( \lambda_\text{BB} + \lambda_\text{WV} +
-  \lambda_\text{Albedo} + \lambda_\text{Cloud} + \dots \right) \cdot \Delta T.
+  \lambda_\text{Albedo} + \lambda_\text{Cloud} + \ldots \right) \cdot \Delta T.
   \label{eq:02:linear_response}
@@
 \begin{equation}
-  \hat{y}(x) = b_0 + b_1 x
+  \hat{y}(x) = \hat{b}_0 + \hat{b}_1 x
   \label{eq:02:linear_regression_y}
 \end{equation}
-for the predicted target variable $\hat{y}$ with intercept $b_0$ and slope
-$b_1$ is used. Fitting this regression line with ordinary least squares
+for the predicted target variable $\hat{y}$ with estimated intercept
+$\hat{b}_0$ and slope $\hat{b}_1$ is used (see \cref{eq:02:linear_coefficients}
+for details on this). Fitting this regression line with ordinary least squares
 includes minimizing the standard error $s$ of the estimate
@@
 \autocite{Nijsse2020} or when additional physical considerations further
-constrain the regression model, \eg{} by demanding a zero intercept ($b_0 = 0$)
-\autocite{Annan2020, JimenezdelaCuesta2019}. Moreover, using only a single
-observational dataset to estimate $x_0$ and $\sigma_x$ when different datasets
-are available might lead to an underestimation of the observational
+constrain the regression model, \eg{} by demanding a zero intercept ($\hat{b}_0
+= 0$) \autocite{Annan2020, JimenezdelaCuesta2019}. Moreover, using only a
+single observational dataset to estimate $x_0$ and $\sigma_x$ when different
+datasets are available might lead to an underestimation of the observational
 uncertainty, as different observational datasets might lead to different
@@
 
-An alternative approach to weighting climate models based on performance and
-interdependence metrics is a technique called \emph{\ac{MDER}}. Similar to all
-methods presented in \cref{sec:02:reducing_uncertainties}, it can be used to
-reduce uncertainties in climate model projections with observations. The basis
-of \ac{MDER} is a set of $K$ predictor diagnostics $\left\{ x^{(k)} \right\}$
-with $k \in \left\{ 1, 2, \dots, K \right\}$ which are relevant for the
-projection of the target variable $y$. The reasoning for this choice of the
-diagnostics is similar to the one presented in the previous section: Weighting
-schemes should address the question \enquote{which climate model is adequate
-  for predicting the target variable?} and not \enquote{which climate model is
-  the best?}.
-
-The concept of \ac{MDER} is mathematically similar to the concept of emergent constraints.
-In a first step, an inter-model relationship between the target variable $y$ and the process-relevant diagnostics $\left\{ x^{(k)} \right\}$ is used to fit a multivariate linear regression model
-\begin{equation}
-  \bm{X}
-\end{equation}
+An alternative climate model weighting scheme is the \emph{\ac{MDER}} approach
+\autocite{Karpechko2013}. Similar to all methods presented in
+\cref{sec:02:reducing_uncertainties}, it can be used to reduce uncertainties in
+climate model projections with observations. The basis of \ac{MDER} is a set of
+$K$ predictor diagnostics $\left\{ x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right\}$
+which are relevant for the projection of the target variable $y$. The reasoning
+for this choice of the diagnostics is similar to the one presented in the
+previous section: Weighting schemes should address the question \enquote{which
+  climate model is adequate for predicting the target variable?} and not
+\enquote{which climate model is the best?}.
 
@@
     ch02_scientific_background/figs/mder_schematic.pdf}
-  \caption{Schematic illustration of the \acf{MDER} approach. First,
-    inter-model relationships between a target variable $y$ and multiple
-    process-based predictors $x^{(k)}$ (here: two predictors $x^{(1)}$ and
-    $x^{(2)}$) are used to fit a multivariate linear regression model (gray
-    plane). Second, observations of the predictors (horizontal thick black
-    lines) are fed into the regression model to calculate climate models
-    weights $w_m$ following \cref{eq:02:mder_weights}. The black circle
-    indicates the best estimate for the target variable $y$ given by the
-    observed values of $x^{(1)}$ and $x^{(2)}$. Each of the remaining colored
-    circles represents a single climate model of the multi-model ensemble. The
-    vertical dashed lines are auxiliary lines to visualize the distance
-    between the climate model data and the linear regression surface.}
+  \caption{Schematic illustration of the \acf{MDER} approach
+    \autocite{Karpechko2013}. First, inter-model relationships between a target
+    variable $y$ and multiple process-based predictors $x^{(k)}$ (here: two
+    predictors $x^{(1)}$ and $x^{(2)}$) are used to fit a multivariate linear
+    regression model (gray surface). Second, observations of the predictors
+    (horizontal black lines) are fed into the regression model to calculate an
+    observation-based best estimate of the target variable $\hat{y}_0$
+    following \cref{eq:02:mder_best_estimate}. The black circle indicates the
+    best estimate for the target variable $y$ given by the observed values of
+    $x^{(1)}$ and $x^{(2)}$. Each of the remaining colored circles represents
+    a single climate model of the multi-model ensemble. The vertical dashed
+    lines visualize the distance between the climate model data and the linear
+    regression surface and represent the noise terms $\bm{\varepsilon}$ in
+    \cref{eq:02:multivariate_linear_model}.}
   \label{fig:02:mder}
@@
 
-\begin{equation}
-  w_m = xD
+The concept of \ac{MDER} is mathematically similar to the concept of emergent
+constraints. In a first step, an inter-model relationship between the target
+variable and the process-relevant diagnostics is used to fit a multivariate
+linear regression model. Let $\bm{y} = \left( y_1, y_2, \ldots, y_M \right)^T
+\in \mathbb{R}^M$ be the vector of target variables of the $M$ climate models
+($T$ denotes the transpose) and $\bm{X} \in \mathbb{R}^{M \times (K + 1)}$ the
+design matrix representing the predictors:
+\begin{equation}
+  \bm{X} =
+  \begin{pmatrix}
+    1 & x_1^{(1)} & x_1^{(2)} & \cdots & x_1^{(K)} \\
+    1 & x_2^{(1)} & x_2^{(2)} & \cdots & x_2^{(K)} \\
+    \vdots & \vdots & \vdots & \ddots & \vdots \\
+    1 & x_M^{(1)} & x_M^{(2)} & \cdots & x_M^{(K)} \\
+  \end{pmatrix}.
+  \label{eq:02:design_matrix}
+\end{equation}
+The entry $x_m^{(k)}$ of this matrix refers to the diagnostic variable of
+diagnostic $k$ and climate model $m$. With this notation, the linear
+inter-model relationship can be written as
+\begin{equation}
+  \bm{y} = \bm{X} \bm{b} + \bm{\varepsilon},
+  \label{eq:02:multivariate_linear_model}
+\end{equation}
+where $\bm{b} = \left( b_0, b_1, \ldots, b_K \right)^T \in \mathbb{R}^{(K +
+  1)}$ is the vector of linear coefficients (with intercept $b_0$) and
+$\bm{\varepsilon} = \left( \varepsilon_1, \varepsilon_2, \ldots, \varepsilon_M
+\right)^T \in \mathbb{R}^M$ a vector of independent random variables
+representing the noise in the target variable. \Cref{fig:02:mder} shows a
+schematic that illustrates this linear relationship (gray surface) for the
+different climate models (colored circles) for two diagnostics ($K = 2$). Using
+ordinary least squares regression, the estimated linear coefficients
+$\hat{\bm{b}}$ are given by
+\begin{equation}
+  \hat{\bm{b}} = \left( \bm{X}^T \bm{X} \right)^{-1} \bm{X}^T \bm{y},
+  \label{eq:02:linear_coefficients}
+\end{equation}
+where the exponent $-1$ denotes the inverse matrix. Since this definition works
+for any number of diagnostics $K$, it can also be used to calculate the linear
+coefficients $\hat{b}_0$ (intercept) and $\hat{b}_1$ (slope) that define
+emergent relationships (see \cref{eq:02:linear_regression_y}).
+
+In a second step, observed data of the process-based diagnostics $\bm{x}_0 =
+\left( 1, x_0^{(1)}, x_0^{(2)}, \ldots, x_0^{(K)} \right)^T \in \mathbb{R}^{(K
+  + 1)}$ is fed into the multivariate linear regression model to get an
+observation-based prediction of the target variable $\hat{y}_0$
+\autocite{Karpechko2013}:
+\begin{equation}
+  \hat{y}_0 = \bm{x}_0^T \hat{\bm{b}}
+  \label{eq:02:mder_best_estimate}
+\end{equation}
+This is mathematically similar to the calculation of the best estimate target
+variable $y$ for emergent constraints. In \cref{fig:02:mder}, the observations
+of the predictors $\bm{x}_0$ are illustrated with horizontal black lines and
+the best estimate $\hat{y}_0$ is shown as a black circle. By combining
+\cref{eq:02:linear_coefficients,eq:02:mder_best_estimate} and comparing this to
+the definition of weighted means in \cref{eq:02:weighted_mean}, climate model
+weights $\bm{w} = \left( w_1, w_2, \ldots, w_M \right)^T \in \mathbb{R}^M$ can
+be defined by
+\begin{equation}
+  \bm{w} = \left[ \bm{x}_0^T \left( \bm{X}^T \bm{X} \right)^{-1} \bm{X}^T
+  \right]^T,
   \label{eq:02:mder_weights}
 \end{equation}
-
-From Knutti (2017): A way out, at least partially, is to reframe the
-competition of “which model is the best” to “which models are adequate for
-predicting X” [Parker, 2009]
-
-- some equations for MDER
-
-- Karpechko, Wenzel, Senftleben
-
-- basic assumption (in contrast to weighting), see Schlund JGR 2020
-
-- Difference to standard model weighting: weaker assumption (see Karpechko
-2013): The assumption that the regression between the present-day diagnostic
-and a future-climatic variable— derived from climate models—holds also for true
-cli- mate might seem to be weak at first sight, especially since it explicitly
-requires imperfect models away from observed climate to span the desired
-relationship. How- ever, a much weaker assumption is done in traditional
-model-weighting approaches: a model that simulates present-day climate better
-than another model neces- sarily better simulates future climate. Whereas
-traditional approaches do not show this relationship, ensemble re- gression
-establishes exactly such a relationship across a broad model ensemble.
-
-- Techniques for diagnostic selection: hypothesis testing, forward iteration,
-backwards elimination
-
-- Disadvantages: no interdependence, no errors in OBS
+which can be used to calculate the weighted target variable by
+\begin{equation}
+  \hat{y}_0 = \hat{\bm{w}}^T \bm{y}.
+  \label{eq:02:weighted_mean_vector}
+\end{equation}
+
+A crucial aspect for the success of the \ac{MDER} approach is the choice of the
+process-relevant diagnostics. In addition to the pre-selection based on expert
+judgment, an additional selection based on statistical criteria is necessary
+for two reasons: First, predictors which only show a weak correlation with the
+target variable should not be included in the regression model since they
+introduce additional noise and might lead to overconfident results. Second,
+multicollinearity (\ie{} mutually correlated predictors) should be avoided
+since this reduces the robustness of the linear regression. A common technique
+to deal with these problems is a stepwise feature selection algorithm based on
+statistical tests of the correlations between the involved variables
+\autocite{Karpechko2013}.
+
+The basic assumption of the \ac{MDER} algorithm is that the inter-model
+relationship between the process-based predictors and the target variable also
+holds for the true climate. This may seem weak at first glance, especially
+since it explicitly requires climate models that deviate from the observed
+climate to span the desired relationship (similar to emergent constraints).
+However, a much weaker assumption is made traditional in weighting approaches
+which assume that climate models that are better in simulating the past or
+present‐day climate are necessarily better in simulating the future climate. In
+contrast to these other weighting approaches, \ac{MDER} explicitly establishes
+the relationship between past/present and future within the climate model
+ensemble \autocite{Karpechko2013}.
+
+Drawbacks of the \ac{MDER} approach are the missing consideration of errors in
+the observational data, the limitation to linear relationships between the
+process-relevant diagnostics and the target variable and the limitation to a
+single data point per climate model. Despite these, \ac{MDER} has been
+successfully used to constrain uncertainties in Antarctic total ozone
+projections \autocite{Karpechko2013}, in the projected change of the austral
+jet position \autocite{Wenzel2016a} and in projections of the Arctic sea ice
+extent \autocite{Senftleben2020}.

### 2020-12-18 ch02_scientific_background/scientific_background.tex
Finished chapter 2 and solved formatting issues with underfull vboxes

@@
     relative to a corresponding pre-industrial control simulation smoothed
-    with a linear fit over all 140 years. The \acf{TCR} is defined as $\Delta
-    T$ at the time of \acs{CO2} doubling averaged over a 20 year period
-    (illustrated by the vertical dashed lines). Here, $\text{TCR} = 2.35
-    \unit{K}$ (horizontal dashed line).}
+    with a linear fit over all 140 years. The \acf{TCR} is defined as the
+    temperature change $\Delta T$ at the time of \acs{CO2} doubling averaged
+    over a 20 year period (illustrated by the vertical dashed lines). Here,
+    $\text{TCR} = 2.35 \unit{K}$ (horizontal dashed line).}
   \label{fig:02:tcr}
@@
 climate model $m$ and $y_m$ the corresponding target variable. To find the
-linear emergent relationship for the climate model ensemble $\left\{ \left(
-x_m, y_m \right) \right\}$, a linear regression model
+linear emergent relationship for a climate model ensemble with $M$ climate
+models and data $\left\{ \left( x_m, y_m \right) \mid m \in I_M \right\}$ with index set $I_M = \left\{ 1, 2, \ldots, M\right\}$, a linear
+regression model
 \begin{equation}
@@
 \begin{equation}
-  P(y|x) = \frac{1}{\sqrt{2 \pi \sigma_{\hat{y}}^2(x)}} \exp \left[
+  \cond{y}{x} = \frac{1}{\sqrt{2 \pi \sigma_{\hat{y}}^2(x)}} \exp \left[
     -\frac{\left( y - \hat{y}(x) \right)^2}{2 \sigma_{\hat{y}}^2(x)} \right].
@@
 \begin{equation}
-  P(x_0|x) = \frac{1}{\sqrt{2 \pi \sigma_x^2}} \exp \left[ -\frac{\left( x_0 -
-    x \right)^2}{2 \sigma_x^2} \right],
+  \cond{x_0}{x} = \frac{1}{\sqrt{2 \pi \sigma_x^2}} \exp \left[ -\frac{\left(
+    x_0 - x \right)^2}{2 \sigma_x^2} \right],
   \label{eq:02:pdf_obs}
@@
 value. Assuming an imperfect uniform prior $P(x) \propto 1$ with cut-offs at
-$-\infty$ and $+\infty$ and using Bayes' theorem implies $P(x_0|x) = P(x|x_0)$.
-In a final step, this can be used to calculate the posterior \ac{PDF} for the
-constrained prediction of the target variable $y$ given the observation $x_0$
-(gray \ac{PDF} in \cref{fig:02:emergent_constraint}) with numerical
-integration:
-\begin{equation}
-  P(y|x_0) = \int_{-\infty}^{+\infty} P(y|x) \, P(x|x_0) \dd{}x.
+$-\infty$ and $+\infty$ and using Bayes' theorem implies $\cond{x_0}{x} =
+\cond{x}{x_0}$. In a final step, this can be used to calculate the posterior
+\ac{PDF} for the constrained prediction of the target variable $y$ given the
+observation $x_0$ (gray \ac{PDF} in \cref{fig:02:emergent_constraint}) with
+numerical integration:
+\begin{equation}
+  \cond{y}{x_0} = \int_{-\infty}^{+\infty} \cond{y}{x} \, \cond{x}{x_0} \dd{}x.
   \label{eq:02:pdf_y_given_x0}
@@
 
-One limitation of the presented approach is the assumption that the individual
+A convenient metric to quantify the skill of an emergent relationship is the
+coefficient of determination $R^2$ of its underlying statistical model. In the
+presented framework which is based on univariate ordinary least squares
+regression, $R^2$ is given by the squared Pearson correlation coefficient $r$
+evaluated on the climate model ensemble data $\left\{ \left( x_m, y_m \right)
+\mid m \in I_M \right\}$, \ie{} $R^2 = r^2$. A further quantity describing the
+skill of an emergent relationship is its statistical significance. In the
+introduced framework, a two-sided $t$-test can be used to determine how likely
+the correlation found between the target variable $y$ and the predictor $x$
+would be to appear by chance. The null hypothesis for this test is that the
+predictor and the target variable are not linearly correlated, \ie{} that the
+true underlying Pearson correlation coefficient of the population is zero. If
+this null hypothesis is true, the probability distribution of the variable
+\begin{equation}
+  t = \frac{r \sqrt{M - 2}}{\sqrt{1 - r^2}}
+  \label{eq:02:t}
+\end{equation}
+is a Student's $t$-distribution with $M - 2$ degrees of freedom. The
+statistical significance can then be measured with the $p$-value of this
+two-sided $t$-test, which describes the probability of obtaining an absolute
+sample Pearson correlation coefficient greater than $\abs*{r}$ if the null
+hypothesis is true. Smaller $p$-values indicate higher a higher statistical
+significance and vice versa.
+
+One limitation of the presented framework is the assumption that the individual
 data points from the different climate models are independent. As already noted

### 2020-12-21 ch01_introduction/introduction.tex
Added section on emergent constraints
Added introduction of chapter on ESMValTool

--- 2020-12-04 ch01_introduction/introduction.tex
@@
 \section{Motivation}
-\label{ch:01:motivation}
+\label{sec:01:motivation}
 
@@
 
-- natural vs. anthropogenic effects (D \& A) -> clear evidence that climate
+- natural vs. anthropogenic effects (D\&A) -> clear evidence that climate
 change over 21st century is caused by human influence
@@
 
-\section{Structure of the thesis}
+- remove description of emergent constraints from section 1.3, since this will
+probably be included in this section.
+
+- answer question: what is climate sensitivity? What is \ac{ECS}? What is
+\ac{TCR}? What is \ac{CMIP}?
+
+- One sentence on \ac{ML}.
+
+\section{Research Questions}
+\label{sec:01:research_questions}
+
+This section will present the \textbf{AWESOME} research questions!!!
+
+\begin{enumerate}
+  \item Did climate sensitivity change in the latest generation of climate
+  models from \acs{CMIP}6 and can we see progress compared to older climate
+  model generations?
+
+  \item Can we use emergent constraints established in \acs{CMIP}3 and
+  \acs{CMIP}5 to reduce uncertainties in climate sensitivity in the \acs{CMIP}6
+  ensemble?
+
+  \item Can we use \ac{ML}-based approaches to constrain gridded projections of
+  the future climate with observations?
+\end{enumerate}
+
+\section{Structure of the Thesis}
 \label{sec:01:structure}
 
-Parts of this thesis are published in multiple peer-reviewed publications (two
-first-author studies and six co-author studies). If applicable, this is clearly
-stated at the beginning of each chapter. \Cref{ch:02:scientific_background}
-introduces the scientific background for this thesis. This includes relevant
-literature that is used as a baseline for this thesis. \Cref{ch:03:esmvaltool}
-gives an overview over the contributions made to the \ac{ESMValTool}, an
-open-source software for the analysis of \acp{ESM}. These contributions helped
-improving the routine evaluation of \acp{ESM} which is useful for the whole
-scientific community and lead to co-authorship in four peer-reviewed studies
-\autocite{Righi2020, Eyring2020, Lauer2020, Weigel2020}.
-\Cref{ch:04:assessment_climate_metrics} covers the assessment of
-policy-relevant climate metrics like the \ac{ECS} and the \ac{TCR} in the
-latest generation of \acp{ESM}. This work is already published in two
-scientific publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and
-\ac{TCR} are considerably higher in this new climate model generation,
-\cref{ch:05:paper_ecs} describes the assessment of emergent constraints (a
-technique to reduce uncertainties in climate model projections, see
-\vref{sec:02:techniques}) on the \ac{ECS} for these \acp{ESM}. The contents of
-this chapter are published in \emph{Earth System Dynamics}
-\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
-reduce climate model uncertainties based on \ac{ML}. As an example, the method
-is applied to the photosynthesis rate at the end of the \nth{21} century, which
-is already published in the \emph{Journal of Geophysical Research:
-  Biogeosciences} \autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook}
-provides a summary of the results of this thesis and gives an outlook of
-possible future works.
+Parts of this thesis are already published in multiple peer-reviewed
+publications (two first-author studies and six co-author studies). Unless
+stated otherwise, results, discussions and conclusions presented from these
+studies originate from the author of this thesis. A detailed list of
+corresponding references is given at the beginning of the relevant chapters.
+
+This thesis is structured as follows: \Cref{ch:02:scientific_background}
+introduces the scientific background. This includes relevant literature that is
+used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
+over the contributions made to the \ac{ESMValTool}, an open-source software for
+the analysis of \acp{ESM}. These contributions helped improving the routine
+evaluation of \acp{ESM} which is useful for the whole scientific community and
+lead to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
+  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:assessment_climate_metrics}
+covers the assessment of policy-relevant climate metrics like the \ac{ECS} and
+the \ac{TCR} in the latest generation of \acp{ESM}. This work is already
+published in two scientific publications \autocite{Bock2020, Meehl2020}. Since
+the \ac{ECS} and \ac{TCR} are considerably higher in this new climate model
+generation, \cref{ch:05:paper_ecs} describes the assessment of emergent
+constraints (a technique to reduce uncertainties in multi-model climate
+projections, see \cref{subsec:02:emergent_constraints}) on the \ac{ECS} for
+these \acp{ESM}. The contents of this chapter are published in \emph{Earth
+  System Dynamics} \autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a
+new method to reduce uncertainties in gridded multi-model projections of the
+future climate with observations based on \ac{ML}. As an example, the method is
+applied to the photosynthesis rate at the end of the \nth{21} century, which is
+already published in the \emph{Journal of Geophysical Research: Biogeosciences}
+\autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook} provides a
+summary of the results of this thesis and gives an outlook of possible future
+works.

### 2020-12-21 ch02_scientific_background/scientific_background.tex
Added introduction of chapter on ESMValTool

@@
 concepts of climate model simulations and associated uncertainties are
-introduced. Next, the fundamental biogeochemical processes of the global carbon
-cycle and important metrics describing climate change are presented. Finally,
-state-of-the art techniques used reduce uncertainties in projections of the
-future climate are shown. These methods form the basis for the new techniques
-developed in this thesis.
+introduced. Next, important metrics describing climate sensitivity and
+fundamental biogeochemical processes of the global carbon cycle are presented.
+Finally, state-of-the art techniques that can be used to reduce uncertainties
+in projections of the future climate with observations are shown. These methods
+form the basis for new techniques developed in this thesis.
 

### 2020-12-21 ch03_esmvaltool/esmvaltool.tex
Initialized chapters
Added first figure
Added further text and figures to ch.2
Added section 2.2
Started section on techniques for reducing uncertainties in climate model projections
Added introduction of chapter on ESMValTool

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                                 ESMValTool                                  %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Improving Routine Climate Model Evaluation}
+\label{ch:03:esmvaltool}
 
-The \ac{ECS} is really cool. I like it very much!
+In order to answer the research questions of this thesis posed in
+\cref{sec:01:research_questions}, a reliable tool to read, process and evaluate
+output from climate model simulations is necessary. A valuable software that
+fits these criteria is the \ac{ESMValTool}. The \ac{ESMValTool} is an
+open-source community diagnostics and performance metrics tool for the routine
+evaluation of \ac{ESM} output, which particularly facilitates the analysis of
+\ac{CMIP} models (\url{https://www.esmvaltool.org}). For this reason, all
+analyses presented in
+\cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp} of this
+thesis were implemented into the \ac{ESMValTool}. Apart from that, further
+substantial contributions to the code base of the \ac{ESMValTool} have been
+made, which lead to co-authorship in the four peer-reviewed studies that
+provide its scientific documentation \autocite{Eyring2020, Lauer2020,
+  Righi2020, Weigel2020}. After a brief overview over the \ac{ESMValTool}'s
+structure, this chapter presents all these contributions that are not
+documented in other chapters of this thesis.
+
+\section{The \acf{ESMValTool}}
+\label{sec:03:esmvaltool}
+
+- overview ESMValTool
+
+- CMOR
+
+- Changes in ESMValCore + table
+
+- Added recipes and diagnostics for ESMValTool + table
+
+* Anav13
+
+* MDER
+
+* TCR
+
+* ECS
+
+* Schlund20jgr
+
+* Schlund20ESD
+
+* Flato13
+
+- total number of changed lines of code

### 2020-12-21 ch01_introduction/introduction.tex
Optimized introduction on chapter 3

@@
 the analysis of \acp{ESM}. These contributions helped improving the routine
-evaluation of \acp{ESM} which is useful for the whole scientific community and
+evaluation of \acp{ESM} which is useful for the entire scientific community and
 lead to co-authorship in four peer-reviewed studies \autocite{Eyring2020,

### 2020-12-21 ch03_esmvaltool/esmvaltool.tex
Optimized introduction on chapter 3

@@
 In order to answer the research questions of this thesis posed in
-\cref{sec:01:research_questions}, a reliable tool to read, process and evaluate
-output from climate model simulations is necessary. A valuable software that
-fits these criteria is the \ac{ESMValTool}. The \ac{ESMValTool} is an
-open-source community diagnostics and performance metrics tool for the routine
-evaluation of \ac{ESM} output, which particularly facilitates the analysis of
-\ac{CMIP} models (\url{https://www.esmvaltool.org}). For this reason, all
-analyses presented in
+\cref{sec:01:research_questions}, a reliable and efficient tool to read,
+process and evaluate output from climate model simulations is necessary. A
+valuable software that fits these criteria is the \ac{ESMValTool}. The
+\ac{ESMValTool} is an open-source community diagnostics and performance metrics
+tool for the routine evaluation of \ac{ESM} output, which notably facilitates
+the analysis of \ac{CMIP} models (\url{https://www.esmvaltool.org}). For this
+reason, all analyses presented in
 \cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp} of this
@@
 substantial contributions to the code base of the \ac{ESMValTool} have been
-made, which lead to co-authorship in the four peer-reviewed studies that
-provide its scientific documentation \autocite{Eyring2020, Lauer2020,
-  Righi2020, Weigel2020}. After a brief overview over the \ac{ESMValTool}'s
-structure, this chapter presents all these contributions that are not
-documented in other chapters of this thesis.
+made that improve the routine evaluation of climate models, which is beneficial
+for the entire scientific community. Moreover, this lead to co-authorship in
+the scientific documentation of the \ac{ESMValTool}, which is published in four
+peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
+After a brief overview over the \ac{ESMValTool}'s structure, this chapter
+presents all these contributions that are not documented in other chapters of
+this thesis.
 

### 2020-12-22 ch03_esmvaltool/esmvaltool.tex
Added section on ESMValTool

@@
 \cref{sec:01:research_questions}, a reliable and efficient tool to read,
-process and evaluate output from climate model simulations is necessary. A
-valuable software that fits these criteria is the \ac{ESMValTool}. The
+process and evaluate climate model output and observational data is necessary.
+A valuable software that fits these criteria is the \ac{ESMValTool}. The
 \ac{ESMValTool} is an open-source community diagnostics and performance metrics
@@
 
-- overview ESMValTool
+As shown in \cref{subsec:02:climate_modeling}, climate models have been
+continuously improved and extended over the last decades from the relatively
+simple \acp{AOGCM} to the complex state-of-the-art \acp{ESM} that include an
+immense number of variables and processes. In \acs{CMIP}6, more modeling
+institutes provide data for more versions of these complex models running
+simulations for more experiments. Consequently, the data volume of the entire
+\acs{CMIP}6 archive is expected to reach up to $80 \unit{PB}$
+\autocite{Balaji2018}, which is a vast increase in comparison to the $2
+\unit{PB}$ of \acs{CMIP}5. The increasing complexity and data volume of the
+climate models pose a major challenge for the evaluation and analysis of the
+model output. To address this big data challenge and support the scientific
+community, the \ac{ESMValTool} has been developed to provide an
+\enquote{open-source, standardized, community-based software package for the
+  systematic, efficient, and well-documented analysis of \ac{ESM} results}
+\autocite{Righi2020}. Moreover, the \ac{ESMValTool} allows a routine comparison
+of single or multiple climate models against predecessor versions and/or
+observations.
 
-- CMOR
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch03_esmvaltool/figs/esmvaltool_schematic.pdf}
+  \caption{Schematic representation of the \acf{ESMValTool}. First, input data
+    is preprocessed by the \acf{ESMValCore}. Second, the preprocessed data is
+    read by diagnostic scripts, which create the final output (\eg{} plots).
+    The user can specify the desired input datasets, preprocessing operations
+    and diagnostic scripts in the recipe, which is a configuration file that
+    controls the main workflow of the \acs{ESMValTool}. Taken and modified from
+    \textcite{Righi2020}.}
+  \label{fig:03:esmvaltool_schematic}
+\end{figure}
 
-- Changes in ESMValCore + table
+Since its first release in 2016 \autocite{Eyring2016a}, the \ac{ESMValTool} has
+been greatly extended and improved. A schematic of the current structure of the
+\ac{ESMValTool} is illustrated in \cref{fig:03:esmvaltool_schematic}. The first
+major element of this workflow is an extensive preprocessing of the input data.
+For this, the Python-based \acf{ESMValCore} provide a set of operators
+(\emph{preprocessors}) that are applied to the input data. There are two
+classes of preprocessors: non-optional and optional ones. Examples for
+non-optional preprocessors are the \ac{I/O} operations \emph{load} (loading
+input data from climate models and observations) and \emph{save} (saving the
+preprocessed data), which utilize the \ac{netCDF} format (a self-describing,
+machine-independent binary file format used for the storage of array-based
+scientific data). Further non-optional preprocessors include \emph{checks} and
+\emph{fixes} of the input data, which test whether the input data adheres to
+standards given by the \ac{CMOR} format and fixes the data if necessary. The
+\ac{CMOR} format ensures that the output from the many different modeling
+institutes within a generation of \ac{CMIP} follows common standards. The
+optional preprocessors include commonly used operations on spatiotemporal
+datasets, like spatial and temporal \emph{subsetting}, the calculation of
+\emph{spatial and temporal statistics}, horizontal and vertical
+\emph{interpolation}, land/sea/ice \emph{masking}, \emph{unit conversion}, or
+the calculation of \emph{multi-model statistics}. A further example is the
+\emph{variable derivation}, which can be used to derive non-\ac{CMOR} variables
+from the input data. The aim of the preprocessor is to facilitate the routine
+evaluation of climate models by providing a set of commonly used data
+operations. To reduce computation times, \ac{ESMValCore} allows parallel
+preprocessing of different datasets.
+
+The second main element of the \ac{ESMValTool} workflow is the calculation of
+\emph{diagnostics}. This task is performed by the diagnostic scripts, which can
+be written in multiple programming languages. Currently, the languages Python,
+NCL, R and Julia are supported. The diagnostic scripts, which contain the code
+that runs the actual scientific evaluation, read the preprocessed datasets from
+\ac{ESMValCore} and use these to create the final output of the tool. Apart
+from \ac{netCDF} files and plots, this includes log files with provenance
+information that ensure the reproducibility and transparency of the results.
+The whole workflow of the \ac{ESMValTool} is controlled with the \emph{recipe},
+which is a configuration file in which the user can specify the desired input
+datasets, preprocessing operations and diagnostic scripts. Implementing a new
+scientific assessment into the \ac{ESMValTool} usually includes writing a new
+recipe and one or more diagnostic scripts. A third major element of the
+\ac{ESMValTool} is the \emph{\ac{CMOR}ization} (\ie{} the reformatting of data
+so it adheres to \ac{CMOR} standards) of raw observational datasets (not shown
+in \cref{fig:03:esmvaltool_schematic}). This ensures that the \ac{ESMValTool}
+can process arbitrary observational datasets, which can for example be used to
+assess the skill of climate model simulations. More details on the structure of
+the \ac{ESMValTool} are given in its extensive documentation
+(\url{https://esmvaltool.readthedocs.io}).
+
+To ensure a rapid evaluation of the \acs{CMIP}6 models, the \ac{ESMValTool} is
+fully integrated into the infrastructure of the \ac{ESGF}, which provides the
+\ac{CMIP} model output for the general public \autocite{Eyring2016b}. As soon
+as new model data is published on the \ac{ESGF} servers, it can be accessed
+with the \ac{ESMValTool} and analyzed. This instantaneous evaluation of the
+\ac{CMIP} models is urgently needed since there is a growing dependency on
+\ac{CMIP} products by a broad research community and by national and
+international climate assessments. For this reason, the \ac{ESMValTool} is for
+example used to evaluate climate model output and observational data in several
+chapters of the upcoming Sixth \ac{AR} of the \ac{IPCC}.
+
+The \ac{ESMValTool} is developed open-source on GitHub
+(\url{https://github.com/ESMValGroup}). It is released under the Apache
+License, version 2.0. The source code of the latest released version of the
+\ac{ESMValCore} package, which includes the core functionalities of the
+\ac{ESMValTool}, is publicly available at Zenodo \autocite{Andela2020a}.
+Similarly, the source code of the latest released version of the
+\ac{ESMValTool} package, which includes the recipe and diagnostic scripts that
+can for example be used to reproduce scientific assessments, is also available
+at Zenodo \autocite{Andela2020}.
+
+\section{Contributions to \acs{ESMValCore}}
+\label{sec:03:contributions_to_esmvalcore}
+
+- total number of changed lines of code
+
+- Changes in \ac{ESMValCore} + table
+
+\section{Contributions to \acs{ESMValTool}}
+\label{sec:03:contributions_to_esmvaltool}
+
+- total number of changed lines of code
 
@@
 * Flato13
-
-- total number of changed lines of code

### 2020-12-22 end_acronyms/acronyms.tex
Added first part of carbon cycle section
Finished section of ESMs
Added climate feedback figure and adapted margins
Added subsection on mathematical framework for feedbacks
Added subsection on ECS
Started section on techniques for reducing uncertainties in climate model projections
Added section description for section on reducing uncertainties
Added section on emergent constraints
Added subsection on Knutti weighting
Added section on ESMValTool

@@
 \begin{acronym}
+  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
   \acro{AOGCM}{Atmosphere-Ocean General Circulation Model}
@@
   \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
-  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
+  \acro{CESM}{Community Earth System Model}
   \acro{CMIP}{Coupled Model Intercomparison Project}
+  \acro{CH4}[CH\textsubscript{4}]{methane}
+  \acro{CMOR}{Climate Model Output Rewriter}
+  \acro{CO}{carbon monoxide}
   \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
+  \acro{CV}{cross-validation}
   \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
@@
     Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
-  \acro{ECS}{Equilibrium Climate Sensitivity}
+  \acro{ECS}{Effective Climate Sensitivity}
+  \acro{ENSO}{El Niño-Southern Oscillation}
+  \acro{ESGF}{Earth System Grid Federation}
   \acro{ESM}{Earth System Model}
+  \acro{ESMValCore}{core functionalities of the ESMValTool}
   \acro{ESMValTool}{Earth System Model Evaluation Tool}
@@
   \acro{GPP}{Gross Primary Productivity}
+  \acro{GSAT}{global mean near-surface air temperature}
+  \acro{GtC}{gigatonnes of carbon}
+  \acro{I/O}{Input/Output}
   \acro{IAM}{Integrated Assessment Models}
@@
   \acro{IPCC}{Intergovernmental Panel on Climate Change}
+  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
+  \acro{MDER}{Multiple Diagnostic Ensemble Regression}
   \acro{MIP}{Model Intercomparison Project}
   \acro{ML}{Machine Learning}
+  \acro{MMM}{multi-model mean}
+  \acro{NAO}{North Atlantic Oscillation}
+  \acro{NCAR}{National Center for Atmospheric Research}
+  \acro{netCDF}{Network Common Data Form}
+  \acro{PDF}{probability density function}
+  \acro{ppm}{parts per million}
   \acro{RCP}{Representative Concentration Pathway}
+  \acro{RMSE}{root-mean-square error}
   \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
@@
   \acro{TCR}{Transient Climate Response}
+  \acro{TOA}{top of the atmosphere}
   \acro{WCRP}{World Climate Research Programme}

### 2021-01-04 ch02_scientific_background/scientific_background.tex
Added section on ESMValTool
Added subsection on calculation of CRE feedback parameters

@@
 documentation in order to facilitate the distribution of climate model output
-\autocite{Eyring2016}.
+\autocite{Eyring2016, Juckes2020}.
 
@@
 
+\subsection{Cloud-Related Feedback Parameters}
+\label{subsec:02:cloud_feedback_parameters}
+
+In addition to the calculation of the external forcing $F$, the overall climate
+feedback parameter $\lambda$ and the \ac{ECS}, the Gregory regression can also
+be used to estimate cloud-related feedback parameters. For this, the net
+\ac{TOA} radiation $N$ on the y-axis in \cref{fig:02:gregory_regression} is
+replaced with the \ac{CRE}, which is defined as the difference between the
+all-sky (\ie{} with clouds if present) net \ac{TOA} radiation and the clear-sky
+(\ie{} clouds artificially removed) net \ac{TOA} radiation
+\autocite{Andrews2012}. This can be done for the shortwave $N_\text{SWCRE}$ and
+longwave $N_\text{LWCRE}$ components separately, but also for the combined
+effect $N_\text{CRE} = N_\text{SWCRE} + N_\text{LWCRE}$. The slopes in the
+corresponding Gregory regressions are the so-called \emph{\ac{CRE} feedback
+  parameters} $\lambda_\text{SWCRE}$, $\lambda_\text{LWCRE}$ and
+$\lambda_\text{CRE}$, which quantify the change in \ac{CRE} as a response to
+increasing \acp{GSAT}.
+
 \subsection{\acl{TCR}}
@@
   \centering
-  \includegraphics[width=\FigureWidth{}]{
+  \includegraphics[width=\LargeFigureWidth{}]{
     ch02_scientific_background/figs/keeling_curve.pdf}

### 2021-01-04 ch03_esmvaltool/esmvaltool.tex
Added tables that summarizes changes made to the ESMValCore

@@
 
+\begin{table}[t]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Type & Name & Description \\
+    \midrule
+    Derived variable & \emph{co2s} & Atmospheric \ac{CO2} concentration at
+    surface \\
+    \cmidrule(l){2-3}
+    & \emph{et} & Evapotranspiration \\
+    \cmidrule(l){2-3}
+    & \emph{uajet} & Position of austral jet stream \\
+    \midrule
+    Preprocessor & Amplitude & Amplitude of periodic phenomena (\eg{} diurnal
+    or seasonal cycles) \\
+    \cmidrule(l){2-3}
+    & Land/sea fraction weighting & Weighting of fields based on
+    the land or sea fraction of the respective grid cells \\
+    \cmidrule(l){2-3}
+    & Trend & Calculation of linear trends (slope of ordinary least squares
+    regression) \\
+    \midrule
+    Miscellaneous & Concatenate experiments & Added option to concatenate
+    multiple experiments (\eg{} historical simulation with future scenarios) \\
+    \cmidrule(l){2-3}
+    & Expanded variable derivation & Simplified, cleaned and extended variable
+    derivation \\
+    \bottomrule
+  \end{tabulary}
+  \caption{Summary of features added to the \acf{ESMValCore}.}
+  \label{tab:03:changes_to_esmvalcore}
+\end{table}
+
 \section{Contributions to \acs{ESMValTool}}

### 2021-01-05 ch03_esmvaltool/esmvaltool.tex
Added contributions to ESMValCore to chapter 03

--- 2020-12-22 ch03_esmvaltool/esmvaltool.tex
@@
 thesis were implemented into the \ac{ESMValTool}. Apart from that, further
-substantial contributions to the code base of the \ac{ESMValTool} have been
-made that improve the routine evaluation of climate models, which is beneficial
-for the entire scientific community. Moreover, this lead to co-authorship in
-the scientific documentation of the \ac{ESMValTool}, which is published in four
-peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
-After a brief overview over the \ac{ESMValTool}'s structure, this chapter
-presents all these contributions that are not documented in other chapters of
-this thesis.
+substantial changes and additions to the code base of the \ac{ESMValTool} have
+been contributed that improve the routine evaluation of climate models, which
+is beneficial for the entire scientific community. This lead to co-authorship
+in the scientific documentation of the \ac{ESMValTool}, which is published in
+four peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020,
+  Weigel2020}. After a brief overview over the \ac{ESMValTool}'s structure,
+this chapter presents all these contributions that are not documented in other
+chapters of this thesis.
 
@@
 
-- total number of changed lines of code
+As of December 2020, $40408$ lines of code have been added and $21050$ lines of
+code have been removed from to the GitHub repository of \ac{ESMValCore} by the
+author of this thesis. Apart from general improvements of the code base, these
+changes and additions mainly include new preprocessor functions that can be
+applied to the input data and derivation scripts for new non-\ac{CMOR}
+variables. \Cref{tab:03:changes_to_esmvalcore} shows a summary of these main
+contributions, which are partly published in the scientific documentation of
+\ac{ESMValCore} \autocite{Righi2020}.
 
-- Changes in \ac{ESMValCore} + table
+\begin{table}[t]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Type & Name & Description \\
+    \midrule
+    Preprocessor & \emph{amplitude} & Amplitude of periodic phenomena (\eg{}
+    diurnal or seasonal cycles) \\
+    \cmidrule(l){2-3}
+    & \emph{land/sea fraction weighting} & Weighting of fields based on
+    the land or sea fraction of the respective grid cells \\
+    \cmidrule(l){2-3}
+    & \emph{trend} & Linear trend (slope of ordinary least squares regression)
+    \\
+    \midrule
+    Derived variable & \emph{asr} & Absorbed shortwave radiation \\
+    \cmidrule(l){2-3}
+    & \emph{co2s} & Atmospheric \ac{CO2} concentration at
+    surface \\
+    \cmidrule(l){2-3}
+    & \emph{et} & Evapotranspiration \\
+    \cmidrule(l){2-3}
+    & \emph{rlntcs} & Clear-sky net \acf{TOA} longwave radiation \\
+    \cmidrule(l){2-3}
+    & \emph{rsntcs} & Clear-sky net \ac{TOA} shortwave radiation \\
+    \cmidrule(l){2-3}
+    & \emph{uajet} & Position of austral jet stream \\
+    \bottomrule
+  \end{tabulary}
+  \caption{Summary of changes and features contributed to the
+    \acf{ESMValCore}.}
+  \label{tab:03:changes_to_esmvalcore}
+\end{table}
+
+In total, three new preprocessor functions have been added: \emph{amplitude},
+\emph{land/sea fraction weighting} and \emph{trend}. The \emph{amplitude}
+preprocessor calculates the peak-to-peak amplitude of periodic phenomena along
+an arbitrary coordinate. Usually, this coordinate is a temporal dimension. A
+common application of this preprocessor is the calculation of a variable's
+diurnal or seasonal cycle amplitude. The \emph{land/sea fraction weighting}
+preprocessor weights fields with the land or sea fraction of the respective
+grid cells. For example, this is necessary for the spatial integration of
+flux-related variables which are reported in units of \enquote{per square meter
+  of land/sea} and not in \enquote{per square meter of grid cell}. After the
+weighting, the grid cell areas can be used to integrate the flux-related
+variable over a desired region. The \emph{trend} preprocessor calculates the
+linear trend of a variable along an arbitrary coordinate. The linear trend is
+defined as the slope of an ordinary least squares linear regression of the
+variable against the selected coordinate. For example, this can be used to
+calculate the temporal trend of the \ac{GSAT} over the \nth{20} century.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[t]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/co2s_map.pdf}
+    \caption{}
+    \label{fig:03:co2s:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch03_esmvaltool/figs/co2s_mlo_cycle.pdf}
+    \caption{}
+    \label{fig:03:co2s:b}
+  \end{subfigure}
+  \caption{(a) \acs{CMIP}6 \acf{MMM} of the global atmospheric \acs{CO2}
+    concentration at surface (\emph{co2s}) averaged over the months \acf{JJA}
+    of the year 2014 in the emission-driven historical simulation. The
+    \acs{CMIP}6 \ac{MMM} includes the four climate models that provide all
+    data needed for the calculation of \emph{co2s} (see legend in (b)). (b)
+    Monthly-mean \emph{co2s} at Mauna Loa, Hawaii ($19.5 \unit{\degree N}$,
+    $155.6 \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2014.
+    The thick black line shows observations from the Mauna Loa Observatory
+    \autocite{Keeling2005}; the remaining colored lines show emission-driven
+    historical simulations from individual \acs{CMIP}6 models.}
+  \label{fig:03:co2s}
+\end{figure}
+
+Apart from new preprocessor functions, six derivation scripts for the
+non-\ac{CMOR} variables \emph{asr}, \emph{co2s}, \emph{et}, \emph{rlntcs},
+\emph{rsntcs} and \emph{uajet} have been added. The absorbed shortwave
+radiation (\emph{asr}) is defined as the difference in the incoming \ac{TOA}
+shortwave radiation and the outgoing \ac{TOA} shortwave radiation. The
+atmospheric \ac{CO2} concentration at surface (\emph{co2s}) can be calculated
+from the pressure level-dependent atmospheric \ac{CO2} concentration and the
+surface air pressure using interpolation. An example illustrating \emph{co2s}
+is given in \cref{fig:03:co2s}. \Cref{fig:03:co2s:a} shows the \acs{CMIP}6
+\ac{MMM} of the global \emph{co2s} for the months \acf{JJA} of the year 2014 in
+the emission-driven historical simulation. As expected, the map shows high
+atmospheric \ac{CO2} concentrations over large metropolitan areas with high
+\ac{CO2} emissions (\eg{} over North America, the Arabian Peninsula and East
+Asia). Moreover, regions with high photosynthetic activity (\eg{} the boreal
+forests in the Northern hemisphere summer and the tropical rainforests in South
+America) exhibit smaller \ac{CO2} concentrations. \Cref{fig:03:co2s:b} shows
+the monthly-mean \emph{co2s} at Mauna Loa, Hawaii from 1958 to 2014 for
+observations performed at the Mauna Loa Observatory \autocite{Keeling2005}
+(thick black line) and emission-driven historical simulations from four
+\acs{CMIP}6 models (remaining colored lines). Apart from the model CNRM-ESM2-1,
+the simulated atmospheric \ac{CO2} concentrations show the expected seasonal
+cycle present in the Keeling Curve (see \cref{fig:02:keeling_curve}). All
+climate models correctly simulate the increase in \emph{co2s} over the years.
+However, their are some differences in the absolute values with the model
+BCC-CSM2-MR showing the largest deviations.
+
+The evapotranspiration (\emph{et}) defined as the sum of evaporation and plant
+transpiration can be calculated from the surface latent heat flux and the
+latent heat vaporization constant. The longwave/shortwave clear-sky net
+\acf{TOA} radiations (\emph{rlntcs}/\emph{rsntcs}) is defined as the difference
+between the longwave/shortwave incoming \ac{TOA} radiation assuming clear-sky
+and the longwave/shortwave outgoing \ac{TOA} radiation assuming clear-sky.
+Finally, the position of the austral jet stream (\emph{uajet}) is given by the
+latitude with maximum zonal-mean eastward wind speed at $850 \unit{hPa}$ on the
+Southern hemisphere between $80 \unit{\degree S}$ and $30 \unit{\degree S}$.
 

### 2021-01-06 ch02_scientific_background/scientific_background.tex
Cleaned code

@@
 properties of the climate system. For example, due to the global carbon cycle,
-only about \pct{50} of the emitted \ac{CO2} by humankind remains in the
+only about $50 \unit{\%}$ of the emitted \ac{CO2} by humankind remains in the
 atmosphere \autocite{Friedlingstein2019}. The residual part is absorbed by the
@@
 with an abrupt quadrupling of \ac{CO2} (\emph{abrupt-4xCO2}) and a simulation
-forced with a \pct{1} per year increase of the atmospheric \ac{CO2}
+forced with a $1 \unit{\%}$ per year increase of the atmospheric \ac{CO2}
 concentration (\emph{1pctCO2}) \autocite{Eyring2016}. This is shown in the
@@
 evaporation of water with increasing air temperatures. Each degree of warming
-allows the atmosphere to retain about \pct{7} more water vapor
+allows the atmosphere to retain about $7 \unit{\%}$ more water vapor
 \autocite{Myhre2013}, which closes the positive feedback loop by further
@@
 feedback amplifies any initial forcing (\eg{} caused by anthropogenic \ac{CO2}
-emissions) by a typical factor between two and three, rendering water vapor a
+emissions) by a typical factor between $2$ and $3$, rendering water vapor a
 fundamental agent of climate change \autocite{Myhre2013}. An example for a
@@
     of the linear regression (black line) corresponds to the climate feedback
-    parameter $\lambda$ and the y-intercept corresponds to the radiative
+    parameter $\lambda$ and the $y$-intercept corresponds to the radiative
     forcing $F_\text{4x}$. These can be used to calculate the \acf{ECS} with
     the Gregory regression method according to \cref{eq:02:ecs}
-    \autocite{Gregory2004}. \acs{ECS} is equivalently given by the x-intercept
-    of the linear regression line divided by $2$. Here, $\text{ECS} = 3.14
-    \unit{K}$.}
+    \autocite{Gregory2004}. \acs{ECS} is equivalently given by the
+    \xintercept{} of the linear regression line divided by $2$. Here,
+    $\text{ECS} = 3.14 \unit{K}$.}
   \label{fig:02:gregory_regression}
@@
 \end{equation}
-$F$ is now given by the y-intercept of this linear regression
-($F_\text{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
+$F$ is now given by the \yintercept{} of this linear regression ($F_\text{4x}$)
+and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
 \begin{equation}
@@
 of the Equilibrium Climate Sensitivity. With the help of climate models,
-\textcite{Sherwood2020} showed that the \ac{ECS} is only about \pct{6} lower
-than the best estimate of the true equilibrium warming obtained from
+\textcite{Sherwood2020} showed that the \ac{ECS} is only about $6 \unit{\%}$
+lower than the best estimate of the true equilibrium warming obtained from
 integrating climate models until a new steady state is reached. Nevertheless,
@@
 be used to estimate cloud-related feedback parameters. For this, the net
-\ac{TOA} radiation $N$ on the y-axis in \cref{fig:02:gregory_regression} is
+\ac{TOA} radiation $N$ on the \yaxis{} in \cref{fig:02:gregory_regression} is
 replaced with the \ac{CRE}, which is defined as the difference between the
@@
 Following \textcite{Bindoff2013}, \ac{TCR} is defined as the change in the
-\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a \pct{1} per
-year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}). For
-this, the annual mean \acp{GSAT} are averaged over a 20 year period centered at
-the time of the \ac{CO2} doubling (years \range{61}{80} when the first year is
-indexed with 1). To account for model drift, the annual mean changes in
-\ac{GSAT} are calculated relative to a corresponding pre-industrial control
-simulation smoothed with a linear fit that considers 140 model years (length of
-the \onepctcotwo{} simulation). An illustration of that calculation is shown in
-\cref{fig:02:tcr}.
+\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a $1 \unit{\%}$
+per year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}).
+For this, the annual mean \acp{GSAT} are averaged over a 20 year period
+centered at the time of the \ac{CO2} doubling (years \range{61}{80} when the
+first year is indexed with $1$). To account for model drift, the annual mean
+changes in \ac{GSAT} are calculated relative to a corresponding pre-industrial
+control simulation smoothed with a linear fit that considers 140 model years
+(length of the \onepctcotwo{} simulation). An illustration of that calculation
+is shown in \cref{fig:02:tcr}.
 
@@
   \caption{Global and annual mean near-surface air temperature change $\Delta
-    T$ for 140 years of a simulation with a \pct{1} per year increase of the
-    atmospheric \ac{CO2} concentration (\onepctcotwo{}) for the \acf{CESM} of
-    the \acf{NCAR} (blue circles). The temperature change is calculated
+    T$ for 140 years of a simulation with a $1 \unit{\%}$ per year increase of
+    the atmospheric \ac{CO2} concentration (\onepctcotwo{}) for the \acf{CESM}
+    of the \acf{NCAR} (blue circles). The temperature change is calculated
     relative to a corresponding pre-industrial control simulation smoothed
@@
 forcing and climate feedbacks. For this, the energy balance
-\cref{eq:02:N_vs_delta_t} can be slightly adjusted. Since over \pct{90} of the
-excess energy introduced into the climate system by the radiative forcing $F$
-is taken up by the ocean due to its large heat capacity, $N$ can be taken equal
-to the global ocean heat uptake \autocite{Knutti2017}. In experiments with a
-steadily increasing radiative forcing, which is the case for the \onepctcotwo{}
-simulation, this ocean heat uptake can be approximated with
+\cref{eq:02:N_vs_delta_t} can be slightly adjusted. Since over $90 \unit{\%}$
+of the excess energy introduced into the climate system by the radiative
+forcing $F$ is taken up by the ocean due to its large heat capacity, $N$ can be
+taken equal to the global ocean heat uptake \autocite{Knutti2017}. In
+experiments with a steadily increasing radiative forcing, which is the case for
+the \onepctcotwo{} simulation, this ocean heat uptake can be approximated with
 \begin{equation}

### 2021-01-06 ch03_esmvaltool/esmvaltool.tex
Cleaned code

@@
 example used to evaluate climate model output and observational data in several
-chapters of the upcoming Sixth \ac{AR} of the \ac{IPCC}.
+chapters of the upcoming \ac{AR6} of the \ac{IPCC}.
 
@@
 
-As of December 2020, $40408$ lines of code have been added and $21050$ lines of
+As of December 2020, 40408 lines of code have been added and 21050 lines of
 code have been removed from to the GitHub repository of \ac{ESMValCore} by the

### 2021-01-06 02_abstract/abstract_english.tex
Adapted file permissions
Added figures for ESMValTool contributions (chapter 3)

--- 2020-11-13 ch1_introduction/introduction.tex
@@
 %=============================================================================%
-%                              01 - Introduction                              %
+%                             Abstract (English)                              %
 %=============================================================================%
 
-\chapter{Introduction}
+\chapter*{Abstract (English version)}
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Hendrerit gravida rutrum quisque non tellus orci ac auctor augue.
-Eros donec ac odio tempor orci dapibus.
-Ultrices neque ornare aenean euismod elementum nisi.
-Vivamus arcu felis bibendum ut tristique et egestas quis ipsum.
-Elit ut aliquam purus sit amet luctus.
-Et malesuada fames ac turpis egestas sed.
-Egestas dui id ornare arcu odio ut sem nulla pharetra.
-Porta lorem mollis aliquam ut porttitor leo a diam.
-Ultricies tristique nulla aliquet enim.
-Urna nunc id cursus metus aliquam.
-Faucibus a pellentesque sit amet porttitor.
-Tincidunt ornare massa eget egestas purus.
-Venenatis tellus in metus vulputate eu scelerisque felis imperdiet.
-Posuere urna nec tincidunt praesent.
-Risus pretium quam vulputate dignissim suspendisse in.
-
-Aliquam id diam maecenas ultricies mi eget.
-Mauris pharetra et ultrices neque ornare aenean euismod.
-Vitae suscipit tellus mauris a diam maecenas.
-Massa massa ultricies mi quis hendrerit.
-Tincidunt eget nullam non nisi est sit amet.
-Vitae purus faucibus ornare suspendisse sed nisi.
-Enim sit amet venenatis urna.
-Malesuada fames ac turpis egestas.
-Viverra nam libero justo laoreet sit amet.
-Ornare massa eget egestas purus viverra accumsan in nisl nisi.
-Ac feugiat sed lectus vestibulum mattis.
-
-\autocite{Schlund2020}
-\autocite{Schlund2020a}
+TBA.

### 2021-01-06 ch02_scientific_background/scientific_background.tex
Added figures for ESMValTool contributions (chapter 3)

@@
     ch02_scientific_background/figs/gregory_regression.pdf}
-  \caption{Global and annual mean net \acf{TOA} radiation $N$ \vs{} the change
+  \caption{Global and annual mean net \acf{TOA} radiation $N$ versus the change
     in global and annual mean near-surface air temperature $\Delta T$ for 150
     years of a simulation with an abrupt quadrupling of the atmospheric
-    \ac{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 multi-model mean
+    \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 multi-model mean
     (blue circles). To account for energy leakage and model drift, a linear
@@
 \autocite{Gregory2004}. For this method, the so-called \emph{Gregory
-  regression}, the global and annual mean net \ac{TOA} radiation $N$ \vs{} the
+  regression}, the global and annual mean net \ac{TOA} radiation $N$ versus the
 change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
@@
     T$ for 140 years of a simulation with a $1 \unit{\%}$ per year increase of
-    the atmospheric \ac{CO2} concentration (\onepctcotwo{}) for the \acf{CESM}
+    the atmospheric \acs{CO2} concentration (\onepctcotwo{}) for the \acf{CESM}
     of the \acf{NCAR} (blue circles). The temperature change is calculated

### 2021-01-06 ch03_esmvaltool/esmvaltool.tex
Added figures for ESMValTool contributions (chapter 3)

@@
     of the year 2014 in the emission-driven historical simulation. The
-    \acs{CMIP}6 \ac{MMM} includes the four climate models that provide all
+    \acs{CMIP}6 \acs{MMM} includes the four climate models that provide all
     data needed for the calculation of \emph{co2s} (see legend in (b)). (b)
@@
 
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[t]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_psi.pdf}
+    \caption{}
+    \label{fig:03:cox18:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_ec.pdf}
+    \caption{}
+    \label{fig:03:cox18:b}
+  \end{subfigure}
+  \caption{Emergent constraint on the \acf{ECS} from global temperature
+    variability \autocite{Cox2018}. Magenta colors indicate high sensitivity
+    climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$) and green colors
+    indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}
+    K^{-1}}$). Both panels were created with the \acs{ESMValTool} and are
+    similar to \textcite{Cox2018}. (a) Temperature variability metric $\Psi$
+    versus time for historical simulations from \acs{CMIP}5 models (colored
+    lines) and observations from the HadCRUT4 dataset (black circles)
+    \autocite{Morice2012}. (b) Emergent relationship between \acs{ECS} and the
+    temperature variability metric $\Psi$. The black dot-dashed line shows the
+    linear regression across the climate model ensemble with its associated
+    uncertainties indicated by the black dashed lines (standard prediction
+    error; see \cref{eq:02:spe}). Blue lines show the observational constraint
+    from HadCRUT4 with its best estimate (dot-dashed line) and standard
+    deviation (dashed lines). The letters represent individual \acs{CMIP}5
+    models (see \textcite{Cox2018} for details).}
+  \label{fig:03:cox18}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[t]{0.34\columnwidth}
+    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942a.pdf}
+    \caption{}
+    \label{fig:03:flato13:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{0.34\columnwidth}
+    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942b.pdf}
+    \caption{}
+    \label{fig:03:flato13:b}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{0.24\columnwidth}
+    \raisebox{7.5mm}{\includegraphics[width=\columnwidth]{
+      ch03_esmvaltool/figs/flato13_942_legend.pdf}}
+  \end{subfigure}
+  \caption{Relationships between climate metrics within the \acs{CMIP}5 climate
+    model ensemble. Both panels were created with the \acs{ESMValTool} and are
+    similar to \textcite{Flato2013}. (a) \Acf{GSAT} versus the \acf{ECS} for
+    \acs{CMIP}5 models for the period \range{1961}{1990} (larger symbols) and
+    for pre-industrial control runs (smaller symbols). (b) \acf{TCR} versus
+    \acs{ECS} for \acs{CMIP}5 models. The black line shows a linear fit
+    between \acs{TCR} and \acs{ECS}.}
+\end{figure}
+
 - total number of changed lines of code

### 2021-01-07 ch03_esmvaltool/esmvaltool.tex
Finished chapter 3

@@
   \end{tabulary}
-  \caption{Summary of changes and features contributed to the
-    \acf{ESMValCore}.}
+  \caption{Summary of new preprocessor functions and variable derivation
+    scripts contributed to the \acf{ESMValCore} by the author of this thesis.}
   \label{tab:03:changes_to_esmvalcore}
@@
 climate models correctly simulate the increase in \emph{co2s} over the years.
-However, their are some differences in the absolute values with the model
+However, there are some differences in the absolute values with the model
 BCC-CSM2-MR showing the largest deviations.
@@
 \label{sec:03:contributions_to_esmvaltool}
+
+In addition to the analyses presented in
+\cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp},
+several more recipes and \ac{CMOR}ization scripts for observational data have
+been added to the main \ac{ESMValTool} repository on GitHub. Moreover, multiple
+minor changes and additions have been implemented to improve the code base
+and/or to fix bugs. In total, the author of this study has added 88668 lines of
+code and removed 47816 lines of code from the \ac{ESMValTool} repository (as of
+December 2020). Parts of these implementations are already published in the
+scientific documentation of the \ac{ESMValTool}, which covers large-scale
+diagnostics \autocite{Eyring2020}, diagnostics for emergent constraints and
+future projections \autocite{Lauer2020} and diagnostics for extreme events and
+regional evaluation \autocite{Weigel2020}.
+
+The implemented \ac{CMOR}ization scripts (also referred to as
+\emph{\ac{CMOR}izers}) allow the processing of arbitrary observational data
+with the \ac{ESMValTool}. For this, the raw observational datasets are
+reformatted in such way that they adhere to the \ac{CMOR} standards. Since the
+\ac{CMOR}ization scripts are publicly available like any other code in the
+\ac{ESMValTool} and \ac{ESMValCore} repositories, the entire scientific
+community can use them to \ac{CMOR}ize the observational data. A complete list
+of all eleven newly implemented \ac{CMOR}izers including the main reference for
+the corresponding observational datasets is shown in
+\cref{tab:03:changes_to_esmvaltool}.
+
+Due to an exhaustive overhaul of the \ac{ESMValTool} version 2, older recipes
+from version 1 \autocite{Eyring2016a} cannot be used directly in the latest
+release. Porting recipes from the old version to the current version requires
+substantial changes on the recipes themselves, but also on the corresponding
+diagnostic scripts. As of December 2020, the author of this thesis has ported
+two recipes originally written by other authors to the new version of
+\ac{ESMValTool}: \emph{recipe\_anav13jclim.yml} and
+\emph{recipe\_wenzel16jclim.yml}. \emph{recipe\_anav13jclim.yml} reproduces the
+analysis of \textcite{Anav2013}, who evaluate carbon cycle-related variables
+for climate models of the \acs{CMIP}5 ensemble. This includes climatologies,
+trends and variabilities of important quantities of the terrestrial and oceanic
+carbon cycle like carbon fluxes, carbon reservoir sizes and vegetation
+distributions. \emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis
+performed by \textcite{Wenzel2016a} to constrain future changes in the position
+of the austral jet stream. Due to its flexible code, the corresponding
+diagnostic scripts can be used to apply the \ac{MDER} method to arbitrary
+target variables and predictors.
+
+\begin{table}[t]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Type & Name & Main reference \\
+    \midrule
+    \ac{CMOR}izer & CRU & \autocite{Harris2014} \\
+    & CT2019 & \autocite{Jacobson2020} \\
+    & GCP & \autocite{Friedlingstein2019} \\
+    & HWSD & \autocite{Wieder2014} \\
+    & JMA-TRANSCOM & \autocite{Maki2010} \\
+    & LAI3g & \autocite{Zhu2013} \\
+    & LandFlux-EVAL & \autocite{Mueller2013} \\
+    & MLS-Aura & \autocite{Read2015} \\
+    & MTE & \autocite{Jung2011} \\
+    & NDP & \autocite{Gibbs2006} \\
+    & Scripps-\ac{CO2} & \autocite{Keeling2005} \\
+    \midrule
+    Recipe & \emph{bock20jgr} (figures 8 and 10) & \autocite{Bock2020} \\
+    & \emph{cox18nature} & \autocite{Cox2018} \\
+    & \emph{ecs} & \autocite{Gregory2004} \\
+    & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
+    & \emph{schlund20esd} & \autocite{Schlund2020a} \\
+    & \emph{schlund20jgr\_gpp\_abs\_rcp85} & \autocite{Schlund2020} \\
+    & \emph{schlund20jgr\_gpp\_change\_1pct} & \autocite{Schlund2020} \\
+    & \emph{schlund20jgr\_gpp\_change\_rcp85} & \autocite{Schlund2020} \\
+    & \emph{tcr} & \autocite{Gregory2008} \\
+    \bottomrule
+  \end{tabulary}
+  \caption{Summary of new \acs{CMOR}ization scripts for observational data
+    and recipes contributed to the \acf{ESMValTool} by the author of this
+    thesis. The corresponding file names of the recipes in the
+    \acs{ESMValTool} repository are given by
+    \emph{recipe\_(name\_in\_table).yml}.}
+  \label{tab:03:changes_to_esmvaltool}
+\end{table}
+
+In addition to the ported recipes, several new recipes and diagnostic scripts
+have been contributed to the public \ac{ESMValTool} repository (see
+\cref{tab:03:changes_to_esmvaltool} for a summary). Some of these include the
+analyses performed in other chapters of this thesis, namely
+\emph{recipe\_bock20jgr.yml} (\cref{ch:04:assessment_climate_metrics}),
+\emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and
+\emph{recipe\_schlund20jgr\_*.yml} (\cref{ch:06:paper_gpp}).
+\emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml} contain diagnostic scripts
+that are able to calculate the climate metrics \ac{ECS} and \ac{TCR} for
+arbitrary model output (see \cref{sec:02:climate_sensitivity}). Corresponding
+plots are shown in \cref{fig:02:gregory_regression,fig:02:tcr}.
+\emph{recipe\_cox18nature.yml} reproduces the analysis of \textcite{Cox2018},
+who introduce an emergent constraint in \ac{ECS} based on a global temperature
+variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
+years for \acs{CMIP}5 models and observations form HadCRUT4
+\autocite{Morice2012}. Since high sensitivity climate models exhibit higher
+values of $\Psi$ than low sensitivity models, an emergent relationship between
+\ac{ECS} and $\Psi$ can be established (\cref{fig:03:cox18:b}). With
+observations of $\Psi$, \ac{ECS} can eventually be constrained to $\left( 2.8
+\pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). More details on
+this including the calculation of $\Psi$ are given in \cref{subsec:05:cox}.
 
@@
     indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}
-    K^{-1}}$). Both panels were created with the \acs{ESMValTool} and are
-    similar to \textcite{Cox2018}. (a) Temperature variability metric $\Psi$
-    versus time for historical simulations from \acs{CMIP}5 models (colored
-    lines) and observations from the HadCRUT4 dataset (black circles)
-    \autocite{Morice2012}. (b) Emergent relationship between \acs{ECS} and the
-    temperature variability metric $\Psi$. The black dot-dashed line shows the
-    linear regression across the climate model ensemble with its associated
-    uncertainties indicated by the black dashed lines (standard prediction
-    error; see \cref{eq:02:spe}). Blue lines show the observational constraint
-    from HadCRUT4 with its best estimate (dot-dashed line) and standard
-    deviation (dashed lines). The letters represent individual \acs{CMIP}5
-    models (see \textcite{Cox2018} for details).}
+      K^{-1}}$). Both panels were created with the \acs{ESMValTool} using
+    \emph{recipe\_cox18nature.yml} and are similar to \textcite{Cox2018}. (a)
+    Temperature variability metric $\Psi$ versus time for historical
+    simulations from \acs{CMIP}5 models (colored lines) and observations from
+    the HadCRUT4 dataset (black circles) \autocite{Morice2012}. (b) Emergent
+    relationship between \acs{ECS} and the temperature variability metric
+    $\Psi$. The black dot-dashed line shows the linear regression across the
+    climate model ensemble with its associated uncertainties indicated by the
+    black dashed lines (standard prediction error; see \cref{eq:02:spe}). Blue
+    lines show the observational constraint from HadCRUT4 with its best
+    estimate (dot-dashed line) and standard deviation (dashed lines). The
+    letters represent individual \acs{CMIP}5 models (see \textcite{Cox2018}
+    for details).}
   \label{fig:03:cox18}
 \end{figure}
+
+An important class of recipes in the \ac{ESMValTool} covers international climate assessments.
+One such example is \emph{recipe\_flato13ipcc.yml}, which reproduces large parts of the climate model evaluation performed in chapter 9 of the \ac{IPCC}'s \ac{AR5} \autocite{Flato2013}.
+For this recipe, the two-panel figure 9.42 has been added, which shows the relationship between the historical and pre-industrial \ac{GSAT} and \ac{ECS} (\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS} (\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
+\Cref{fig:03:flato13:a} illustrates that there is no clear connection between \ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high temperatures in the historical or pre-industrial period do no necessarily exhibit a high equilibrium warming in the future or vice versa.
+On the other hand, \ac{TCR} and \ac{ECS} are in fact well-correlated within the \acs{CMIP}5 ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}.
+However, this figure only shows a linear regression fit between the two variables and not the expected reciprocal relation.
 
@@
   \caption{Relationships between climate metrics within the \acs{CMIP}5 climate
-    model ensemble. Both panels were created with the \acs{ESMValTool} and are
-    similar to \textcite{Flato2013}. (a) \Acf{GSAT} versus the \acf{ECS} for
-    \acs{CMIP}5 models for the period \range{1961}{1990} (larger symbols) and
-    for pre-industrial control runs (smaller symbols). (b) \acf{TCR} versus
-    \acs{ECS} for \acs{CMIP}5 models. The black line shows a linear fit
-    between \acs{TCR} and \acs{ECS}.}
+    model ensemble. Both panels were created with the \acs{ESMValTool} using
+    \emph{recipe\_flato13ipcc.yml} and are similar to \textcite{Flato2013}. (a)
+    \Acf{GSAT} versus the \acf{ECS} for \acs{CMIP}5 models for the period
+    \range{1961}{1990} (larger symbols) and for pre-industrial control runs
+    (smaller symbols). (b) \acf{TCR} versus \acs{ECS} for \acs{CMIP}5 models.
+    The black line shows a linear fit between \acs{TCR} and \acs{ECS}.}
+  \label{fig:03:flato13}
 \end{figure}
-
-- total number of changed lines of code
-
-- Added recipes and diagnostics for ESMValTool + table
-
-* Anav13
-
-* MDER
-
-* TCR
-
-* ECS
-
-* Schlund20jgr
-
-* Schlund20ESD
-
-* Flato13

### 2021-01-08 ch01_introduction/introduction.tex
Cleaned code
Added figures for ESMValTool contributions (chapter 3)
Cleaned tex files and started chapter 04

@@
 - natural vs. anthropogenic effects (D\&A) -> clear evidence that climate
-change over 21st century is caused by human influence
-(https://globalwarmingindex.org/, Haustein et al., 2017)
+change over 21st century is caused by human influence. The following website:
+is relevant: (https://globalwarmingindex.org/, Haustein et al., 2017)
 
@@
 
+- Mention IPCC assessments, older ones and upcoming AR6
+
+- "climate model" is used as generic term which include \acp{AOGCM} and
+\acp{ESM} (see \cref{subsec:02:numerical_climate_modeling}) (subcategories).
+Since most modern climate models participating in CMIP are \acp{ESM} (or at
+least have \ac{ESM} versions), the terms "climate model" and "ESM" are used
+interchangeably in this thesis.
+
 \section{Research Questions}
@@
 \begin{enumerate}
-  \item Did climate sensitivity change in the latest generation of climate
-  models from \acs{CMIP}6 and can we see progress compared to older climate
+  \item Did climate sensitivity change in the latest generation of \acp{ESM} from \acs{CMIP}6 and can we see progress compared to older climate
   model generations?
@@
   Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:assessment_climate_metrics}
-covers the assessment of policy-relevant climate metrics like the \ac{ECS} and
-the \ac{TCR} in the latest generation of \acp{ESM}. This work is already
-published in two scientific publications \autocite{Bock2020, Meehl2020}. Since
-the \ac{ECS} and \ac{TCR} are considerably higher in this new climate model
-generation, \cref{ch:05:paper_ecs} describes the assessment of emergent
-constraints (a technique to reduce uncertainties in multi-model climate
-projections, see \cref{subsec:02:emergent_constraints}) on the \ac{ECS} for
-these \acp{ESM}. The contents of this chapter are published in \emph{Earth
+covers the assessment of climate sensitivity metrics like the \ac{ECS} or
+\ac{TCR} in the latest generation of \acp{ESM} (\acs{CMIP}6). This work is
+already published in two scientific publications \autocite{Bock2020,
+  Meehl2020}. Since the \ac{ECS} and \ac{TCR} are considerably higher in this
+new climate model generation, \cref{ch:05:paper_ecs} describes the assessment
+of emergent constraints (a technique to reduce uncertainties in multi-model
+climate projections, see \cref{subsec:02:emergent_constraints}) on the \ac{ECS}
+for these \acp{ESM}. The contents of this chapter are published in \emph{Earth
   System Dynamics} \autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a

### 2021-01-08 ch03_esmvaltool/esmvaltool.tex
Cleaned tex files and started chapter 04

@@
 
-As shown in \cref{subsec:02:climate_modeling}, climate models have been
-continuously improved and extended over the last decades from the relatively
-simple \acp{AOGCM} to the complex state-of-the-art \acp{ESM} that include an
-immense number of variables and processes. In \acs{CMIP}6, more modeling
-institutes provide data for more versions of these complex models running
-simulations for more experiments. Consequently, the data volume of the entire
-\acs{CMIP}6 archive is expected to reach up to $80 \unit{PB}$
+As outlined in \cref{subsec:02:numerical_climate_modeling}, climate models have
+been continuously improved and extended over the last decades from the
+relatively simple \acp{AOGCM} to the complex state-of-the-art \acp{ESM} that
+include an immense number of variables and processes. In \acs{CMIP}6, more
+modeling institutes provide data for more versions of these complex models
+running simulations for more experiments. Consequently, the data volume of the
+entire \acs{CMIP}6 archive is expected to reach up to $80 \unit{PB}$
 \autocite{Balaji2018}, which is a vast increase in comparison to the $2
@@
 cycle present in the Keeling Curve (see \cref{fig:02:keeling_curve}). All
-climate models correctly simulate the increase in \emph{co2s} over the years.
+\acp{ESM} correctly simulate the increase in \emph{co2s} over the years.
 However, there are some differences in the absolute values with the model
@@
 analysis of \textcite{Anav2013}, who evaluate carbon cycle-related variables
-for climate models of the \acs{CMIP}5 ensemble. This includes climatologies,
-trends and variabilities of important quantities of the terrestrial and oceanic
-carbon cycle like carbon fluxes, carbon reservoir sizes and vegetation
-distributions. \emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis
-performed by \textcite{Wenzel2016a} to constrain future changes in the position
-of the austral jet stream. Due to its flexible code, the corresponding
-diagnostic scripts can be used to apply the \ac{MDER} method to arbitrary
-target variables and predictors.
+for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends
+and variabilities of important quantities of the terrestrial and oceanic carbon
+cycle like carbon fluxes, carbon reservoir sizes and vegetation distributions.
+\emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis performed by
+\textcite{Wenzel2016a} to constrain future changes in the position of the
+austral jet stream. Due to its flexible code, the corresponding diagnostic
+scripts can be used to apply the \ac{MDER} method to arbitrary target variables
+and predictors.
 

### 2021-01-08 ch04_assessment_climate_metrics/assessment_climate_metrics.tex
Initialized chapters
Wrote first page of introduction
Added first figure
Added section 2.2
Added overview of climate feedbacks
Added figures for ESMValTool contributions (chapter 3)
Cleaned tex files and started chapter 04

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                        Assessment of climate metrics                        %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Assessment of Climate Sensitivity in \acs{CMIP}6}
+\label{ch:04:assessment_climate_metrics}
 
-The \ac{ECS} is really cool. I like it very much!
+In order to reduce uncertainties in multi-model climate projections, a first
+important step is the assessment of the desired target variables in the
+corresponding climate model ensemble. This is particularly relevant when a new
+generation of \acp{ESM} has been published, which has happened in 2020 with the
+\acs{CMIP}6 ensemble.

### 2021-01-11 ch03_esmvaltool/esmvaltool.tex
Cleaned tex files and added text to chapter 4

@@
 reason, all analyses presented in
-\cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp} of this
+\cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs,ch:06:paper_gpp} of this
 thesis were implemented into the \ac{ESMValTool}. Apart from that, further
@@
 
-As of December 2020, 40408 lines of code have been added and 21050 lines of
-code have been removed from to the GitHub repository of \ac{ESMValCore} by the
+As of \TheMonth{}, 40408 lines of code have been added and 21050 lines of code
+have been removed from to the GitHub repository of \ac{ESMValCore} by the
 author of this thesis. Apart from general improvements of the code base, these
@@
 In addition to the analyses presented in
-\cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp},
-several more recipes and \ac{CMOR}ization scripts for observational data have
-been added to the main \ac{ESMValTool} repository on GitHub. Moreover, multiple
+\cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs,ch:06:paper_gpp}, several
+more recipes and \ac{CMOR}ization scripts for observational data have been
+added to the main \ac{ESMValTool} repository on GitHub. Moreover, multiple
 minor changes and additions have been implemented to improve the code base
@@
 code and removed 47816 lines of code from the \ac{ESMValTool} repository (as of
-December 2020). Parts of these implementations are already published in the
+\TheMonth{}). Parts of these implementations are already published in the
 scientific documentation of the \ac{ESMValTool}, which covers large-scale
@@
 substantial changes on the recipes themselves, but also on the corresponding
-diagnostic scripts. As of December 2020, the author of this thesis has ported
-two recipes originally written by other authors to the new version of
+diagnostic scripts. As of \TheMonth{}, the author of this thesis has ported two
+recipes originally written by other authors to the new version of
 \ac{ESMValTool}: \emph{recipe\_anav13jclim.yml} and
@@
 analyses performed in other chapters of this thesis, namely
-\emph{recipe\_bock20jgr.yml} (\cref{ch:04:assessment_climate_metrics}),
+\emph{recipe\_bock20jgr.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
 \emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and

### 2021-01-11 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Cleaned tex files and added text to chapter 4

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                        Papers ECS and TCR assessment                        %
%=============================================================================%

\chapter{Assessment of Climate Sensitivity in the \acs{CMIP}6 Ensemble}
\label{ch:04:papers_ecs_tcr_assessment}

In order to reduce uncertainties in multi-model climate projections, a first
important step is the assessment of the desired target variables in the
corresponding climate model ensemble. This is particularly relevant when a new
generation of \acp{ESM} is published that includes considerable modifications
compared to the respective predecessor model versions. Currently, we are in a
situation like this with the \acs{CMIP}6 ensemble, for which new data is still
released every day (as of \TheMonth{}). In light of the upcoming international
climate assessment of the \ac{IPCC} (\ac{AR6}), the evaluation of
policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6
models is crucial since they provide vital information about the future climate
of planet Earth. As the \ac{ESMValTool} is an excellent tool that allows for a
quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we assess and analyze
climate sensitivity (expressed by \ac{ECS} and \ac{TCR}) of the \acs{CMIP}6
models and compare it to corresponding results of predecessor model
generations. This work is already published in two scientific papers:
\textcite{Bock2020} and \textcite{Meehl2020}.

\section{Evaluation of \acs{ECS} and \acs{TCR} in \acs{CMIP}6}
\label{sec:04:evaluation_ecs_and_tcr}

### 2021-01-11 ch05_paper_ecs/paper_ecs.tex
Initialized chapters
Added first figure
Added further text and figures to ch.2
Added section 2.2
Added section on emergent constraints
Added figures for ESMValTool contributions (chapter 3)
Finished chapter 3
Cleaned tex files and added text to chapter 4

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                                  Paper ECS                                  %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Evaluation of Emergent Constraints on the \acl{ECS} in \acs{CMIP}6}
+\label{ch:05:paper_ecs}
 
-The \ac{ECS} is really cool. I like it very much!
+TBA.
+
+\section{Comparison of Emergent Constraints on \acs{ECS} for \acs{CMIP}5 and \acs{CMIP}6}
+\label{sec:05:comparison_of_emergent_constraints}
+
+\subsection{COX}
+\label{subsec:05:cox}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua. Sed arcu non odio euismod. Quis
+risus sed vulputate odio ut. Quis commodo odio aenean sed adipiscing diam donec
+adipiscing tristique. Habitant morbi tristique senectus et. Eleifend donec
+pretium vulputate sapien nec sagittis aliquam malesuada bibendum. Lobortis
+feugiat vivamus at augue. Sit amet commodo nulla facilisi nullam. Viverra vitae
+congue eu consequat ac felis donec. Tortor dignissim convallis aenean et tortor
+at risus. Condimentum id venenatis a condimentum. Orci phasellus egestas tellus
+rutrum tellus pellentesque eu tincidunt tortor. Neque volutpat ac tincidunt
+vitae semper quis. Etiam erat velit scelerisque in dictum non consectetur a.
+Pellentesque adipiscing commodo elit at imperdiet dui accumsan sit. Senectus et
+netus et malesuada. Eget gravida cum sociis natoque penatibus et. Purus ut
+faucibus pulvinar elementum. Nulla facilisi etiam dignissim diam quis enim
+lobortis scelerisque fermentum.
+
+Non diam phasellus vestibulum lorem sed risus. Vitae suscipit tellus mauris a
+diam maecenas sed enim. Sed odio morbi quis commodo odio aenean sed adipiscing
+diam. Mauris pharetra et ultrices neque ornare aenean euismod. Diam quis enim
+lobortis scelerisque. Nisi vitae suscipit tellus mauris a diam. Et netus et
+malesuada fames. Urna neque viverra justo nec ultrices dui. Maecenas accumsan
+lacus vel facilisis volutpat est velit egestas dui. Eget magna fermentum
+iaculis eu non diam phasellus vestibulum lorem. Volutpat consequat mauris nunc
+congue nisi vitae suscipit tellus.
+
+Ac tortor dignissim convallis aenean et tortor at risus. Consequat mauris nunc
+congue nisi vitae. Gravida neque convallis a cras semper auctor. Faucibus
+turpis in eu mi bibendum neque. Lorem mollis aliquam ut porttitor leo a diam.
+Facilisis gravida neque convallis a cras semper auctor neque vitae. Nunc
+aliquet bibendum enim facilisis gravida neque convallis a cras. Bibendum enim
+facilisis gravida neque convallis a cras semper auctor. Ut enim blandit
+volutpat maecenas volutpat blandit. Quis ipsum suspendisse ultrices gravida
+dictum fusce ut. Dictumst quisque sagittis purus sit amet volutpat consequat
+mauris nunc. Non arcu risus quis varius quam quisque id diam vel.
+
+Pellentesque sit amet porttitor eget. Enim nulla aliquet porttitor lacus. Enim
+eu turpis egestas pretium aenean pharetra magna ac placerat. In vitae turpis
+massa sed elementum tempus. Cras semper auctor neque vitae tempus quam
+pellentesque nec nam. Sollicitudin tempor id eu nisl. Nunc aliquet bibendum
+enim facilisis gravida neque. Imperdiet sed euismod nisi porta. Est ullamcorper
+eget nulla facilisi etiam dignissim diam quis. Cursus mattis molestie a iaculis
+at erat pellentesque adipiscing commodo. Elit sed vulputate mi sit amet mauris
+commodo quis. Quis risus sed vulputate odio. Orci a scelerisque purus semper
+eget. Scelerisque eu ultrices vitae auctor. Cum sociis natoque penatibus et
+magnis dis parturient. Id aliquet risus feugiat in ante metus. Pretium quam
+vulputate dignissim suspendisse in.
+
+Velit aliquet sagittis id consectetur purus ut faucibus pulvinar. Venenatis a
+condimentum vitae sapien pellentesque habitant morbi tristique. Ut etiam sit
+amet nisl purus in mollis. Etiam dignissim diam quis enim. Id leo in vitae
+turpis. Et magnis dis parturient montes nascetur ridiculus. Gravida rutrum
+quisque non tellus orci. Eu consequat ac felis donec et. Egestas purus viverra
+accumsan in nisl nisi scelerisque eu. Elit ullamcorper dignissim cras tincidunt
+lobortis feugiat vivamus at. Maecenas pharetra convallis posuere morbi leo urna
+molestie. Nunc id cursus metus aliquam eleifend mi. Mauris ultrices eros in
+cursus turpis massa. Lectus nulla at volutpat diam.
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua. Sed arcu non odio euismod. Quis
+risus sed vulputate odio ut. Quis commodo odio aenean sed adipiscing diam donec
+adipiscing tristique. Habitant morbi tristique senectus et. Eleifend donec
+pretium vulputate sapien nec sagittis aliquam malesuada bibendum. Lobortis
+feugiat vivamus at augue. Sit amet commodo nulla facilisi nullam. Viverra vitae
+congue eu consequat ac felis donec. Tortor dignissim convallis aenean et tortor
+at risus. Condimentum id venenatis a condimentum. Orci phasellus egestas tellus
+rutrum tellus pellentesque eu tincidunt tortor. Neque volutpat ac tincidunt
+vitae semper quis. Etiam erat velit scelerisque in dictum non consectetur a.
+Pellentesque adipiscing commodo elit at imperdiet dui accumsan sit. Senectus et
+netus et malesuada. Eget gravida cum sociis natoque penatibus et. Purus ut
+faucibus pulvinar elementum. Nulla facilisi etiam dignissim diam quis enim
+lobortis scelerisque fermentum.
+
+Non diam phasellus vestibulum lorem sed risus. Vitae suscipit tellus mauris a
+diam maecenas sed enim. Sed odio morbi quis commodo odio aenean sed adipiscing
+diam. Mauris pharetra et ultrices neque ornare aenean euismod. Diam quis enim
+lobortis scelerisque. Nisi vitae suscipit tellus mauris a diam. Et netus et
+malesuada fames. Urna neque viverra justo nec ultrices dui. Maecenas accumsan
+lacus vel facilisis volutpat est velit egestas dui. Eget magna fermentum
+iaculis eu non diam phasellus vestibulum lorem. Volutpat consequat mauris nunc
+congue nisi vitae suscipit tellus.
+
+Ac tortor dignissim convallis aenean et tortor at risus. Consequat mauris nunc
+congue nisi vitae. Gravida neque convallis a cras semper auctor. Faucibus
+turpis in eu mi bibendum neque. Lorem mollis aliquam ut porttitor leo a diam.
+Facilisis gravida neque convallis a cras semper auctor neque vitae. Nunc
+aliquet bibendum enim facilisis gravida neque convallis a cras. Bibendum enim
+facilisis gravida neque convallis a cras semper auctor. Ut enim blandit
+volutpat maecenas volutpat blandit. Quis ipsum suspendisse ultrices gravida
+dictum fusce ut. Dictumst quisque sagittis purus sit amet volutpat consequat
+mauris nunc. Non arcu risus quis varius quam quisque id diam vel.
+
+Pellentesque sit amet porttitor eget. Enim nulla aliquet porttitor lacus. Enim
+eu turpis egestas pretium aenean pharetra magna ac placerat. In vitae turpis
+massa sed elementum tempus. Cras semper auctor neque vitae tempus quam
+pellentesque nec nam. Sollicitudin tempor id eu nisl. Nunc aliquet bibendum
+enim facilisis gravida neque. Imperdiet sed euismod nisi porta. Est ullamcorper
+eget nulla facilisi etiam dignissim diam quis. Cursus mattis molestie a iaculis
+at erat pellentesque adipiscing commodo. Elit sed vulputate mi sit amet mauris
+commodo quis. Quis risus sed vulputate odio. Orci a scelerisque purus semper
+eget. Scelerisque eu ultrices vitae auctor. Cum sociis natoque penatibus et
+magnis dis parturient. Id aliquet risus feugiat in ante metus. Pretium quam
+vulputate dignissim suspendisse in.
+
+Velit aliquet sagittis id consectetur purus ut faucibus pulvinar. Venenatis a
+condimentum vitae sapien pellentesque habitant morbi tristique. Ut etiam sit
+amet nisl purus in mollis. Etiam dignissim diam quis enim. Id leo in vitae
+turpis. Et magnis dis parturient montes nascetur ridiculus. Gravida rutrum
+quisque non tellus orci. Eu consequat ac felis donec et. Egestas purus viverra
+accumsan in nisl nisi scelerisque eu. Elit ullamcorper dignissim cras tincidunt
+lobortis feugiat vivamus at. Maecenas pharetra convallis posuere morbi leo urna
+molestie. Nunc id cursus metus aliquam eleifend mi. Mauris ultrices eros in
+cursus turpis massa. Lectus nulla at volutpat diam.

### 2021-01-11 ch06_paper_gpp/paper_gpp.tex
Initialized chapters
Added first figure
Added further text and figures to ch.2
Added overview of climate feedbacks
Added section on emergent constraints
Added figures for ESMValTool contributions (chapter 3)
Cleaned tex files and added text to chapter 4

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                                  Paper GPP                                  %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Constraining Future \acl{GPP} with Machine Learning}
+\label{ch:06:paper_gpp}
 
-The \ac{ECS} is really cool. I like it very much!
+TBA.
+
+\section{Step 1: XXXXXXXXXXXXXXXXXX}
+\label{sec:06:step_1}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua. Sed arcu non odio euismod. Quis
+risus sed vulputate odio ut. Quis commodo odio aenean sed adipiscing diam donec
+adipiscing tristique. Habitant morbi tristique senectus et. Eleifend donec
+pretium vulputate sapien nec sagittis aliquam malesuada bibendum. Lobortis
+feugiat vivamus at augue. Sit amet commodo nulla facilisi nullam. Viverra vitae
+congue eu consequat ac felis donec. Tortor dignissim convallis aenean et tortor
+at risus. Condimentum id venenatis a condimentum. Orci phasellus egestas tellus
+rutrum tellus pellentesque eu tincidunt tortor. Neque volutpat ac tincidunt
+vitae semper quis. Etiam erat velit scelerisque in dictum non consectetur a.
+Pellentesque adipiscing commodo elit at imperdiet dui accumsan sit. Senectus et
+netus et malesuada. Eget gravida cum sociis natoque penatibus et. Purus ut
+faucibus pulvinar elementum. Nulla facilisi etiam dignissim diam quis enim
+lobortis scelerisque fermentum.
+
+Non diam phasellus vestibulum lorem sed risus. Vitae suscipit tellus mauris a
+diam maecenas sed enim. Sed odio morbi quis commodo odio aenean sed adipiscing
+diam. Mauris pharetra et ultrices neque ornare aenean euismod. Diam quis enim
+lobortis scelerisque. Nisi vitae suscipit tellus mauris a diam. Et netus et
+malesuada fames. Urna neque viverra justo nec ultrices dui. Maecenas accumsan
+lacus vel facilisis volutpat est velit egestas dui. Eget magna fermentum
+iaculis eu non diam phasellus vestibulum lorem. Volutpat consequat mauris nunc
+congue nisi vitae suscipit tellus.
+
+Ac tortor dignissim convallis aenean et tortor at risus. Consequat mauris nunc
+congue nisi vitae. Gravida neque convallis a cras semper auctor. Faucibus
+turpis in eu mi bibendum neque. Lorem mollis aliquam ut porttitor leo a diam.
+Facilisis gravida neque convallis a cras semper auctor neque vitae. Nunc
+aliquet bibendum enim facilisis gravida neque convallis a cras. Bibendum enim
+facilisis gravida neque convallis a cras semper auctor. Ut enim blandit
+volutpat maecenas volutpat blandit. Quis ipsum suspendisse ultrices gravida
+dictum fusce ut. Dictumst quisque sagittis purus sit amet volutpat consequat
+mauris nunc. Non arcu risus quis varius quam quisque id diam vel.
+
+Pellentesque sit amet porttitor eget. Enim nulla aliquet porttitor lacus. Enim
+eu turpis egestas pretium aenean pharetra magna ac placerat. In vitae turpis
+massa sed elementum tempus. Cras semper auctor neque vitae tempus quam
+pellentesque nec nam. Sollicitudin tempor id eu nisl. Nunc aliquet bibendum
+enim facilisis gravida neque. Imperdiet sed euismod nisi porta. Est ullamcorper
+eget nulla facilisi etiam dignissim diam quis. Cursus mattis molestie a iaculis
+at erat pellentesque adipiscing commodo. Elit sed vulputate mi sit amet mauris
+commodo quis. Quis risus sed vulputate odio. Orci a scelerisque purus semper
+eget. Scelerisque eu ultrices vitae auctor. Cum sociis natoque penatibus et
+magnis dis parturient. Id aliquet risus feugiat in ante metus. Pretium quam
+vulputate dignissim suspendisse in.
+
+Velit aliquet sagittis id consectetur purus ut faucibus pulvinar. Venenatis a
+condimentum vitae sapien pellentesque habitant morbi tristique. Ut etiam sit
+amet nisl purus in mollis. Etiam dignissim diam quis enim. Id leo in vitae
+turpis. Et magnis dis parturient montes nascetur ridiculus. Gravida rutrum
+quisque non tellus orci. Eu consequat ac felis donec et. Egestas purus viverra
+accumsan in nisl nisi scelerisque eu. Elit ullamcorper dignissim cras tincidunt
+lobortis feugiat vivamus at. Maecenas pharetra convallis posuere morbi leo urna
+molestie. Nunc id cursus metus aliquam eleifend mi. Mauris ultrices eros in
+cursus turpis massa. Lectus nulla at volutpat diam.
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
+incididunt ut labore et dolore magna aliqua. Sed arcu non odio euismod. Quis
+risus sed vulputate odio ut. Quis commodo odio aenean sed adipiscing diam donec
+adipiscing tristique. Habitant morbi tristique senectus et. Eleifend donec
+pretium vulputate sapien nec sagittis aliquam malesuada bibendum. Lobortis
+feugiat vivamus at augue. Sit amet commodo nulla facilisi nullam. Viverra vitae
+congue eu consequat ac felis donec. Tortor dignissim convallis aenean et tortor
+at risus. Condimentum id venenatis a condimentum. Orci phasellus egestas tellus
+rutrum tellus pellentesque eu tincidunt tortor. Neque volutpat ac tincidunt
+vitae semper quis. Etiam erat velit scelerisque in dictum non consectetur a.
+Pellentesque adipiscing commodo elit at imperdiet dui accumsan sit. Senectus et
+netus et malesuada. Eget gravida cum sociis natoque penatibus et. Purus ut
+faucibus pulvinar elementum. Nulla facilisi etiam dignissim diam quis enim
+lobortis scelerisque fermentum.
+
+Non diam phasellus vestibulum lorem sed risus. Vitae suscipit tellus mauris a
+diam maecenas sed enim. Sed odio morbi quis commodo odio aenean sed adipiscing
+diam. Mauris pharetra et ultrices neque ornare aenean euismod. Diam quis enim
+lobortis scelerisque. Nisi vitae suscipit tellus mauris a diam. Et netus et
+malesuada fames. Urna neque viverra justo nec ultrices dui. Maecenas accumsan
+lacus vel facilisis volutpat est velit egestas dui. Eget magna fermentum
+iaculis eu non diam phasellus vestibulum lorem. Volutpat consequat mauris nunc
+congue nisi vitae suscipit tellus.
+
+Ac tortor dignissim convallis aenean et tortor at risus. Consequat mauris nunc
+congue nisi vitae. Gravida neque convallis a cras semper auctor. Faucibus
+turpis in eu mi bibendum neque. Lorem mollis aliquam ut porttitor leo a diam.
+Facilisis gravida neque convallis a cras semper auctor neque vitae. Nunc
+aliquet bibendum enim facilisis gravida neque convallis a cras. Bibendum enim
+facilisis gravida neque convallis a cras semper auctor. Ut enim blandit
+volutpat maecenas volutpat blandit. Quis ipsum suspendisse ultrices gravida
+dictum fusce ut. Dictumst quisque sagittis purus sit amet volutpat consequat
+mauris nunc. Non arcu risus quis varius quam quisque id diam vel.
+
+Pellentesque sit amet porttitor eget. Enim nulla aliquet porttitor lacus. Enim
+eu turpis egestas pretium aenean pharetra magna ac placerat. In vitae turpis
+massa sed elementum tempus. Cras semper auctor neque vitae tempus quam
+pellentesque nec nam. Sollicitudin tempor id eu nisl. Nunc aliquet bibendum
+enim facilisis gravida neque. Imperdiet sed euismod nisi porta. Est ullamcorper
+eget nulla facilisi etiam dignissim diam quis. Cursus mattis molestie a iaculis
+at erat pellentesque adipiscing commodo. Elit sed vulputate mi sit amet mauris
+commodo quis. Quis risus sed vulputate odio. Orci a scelerisque purus semper
+eget. Scelerisque eu ultrices vitae auctor. Cum sociis natoque penatibus et
+magnis dis parturient. Id aliquet risus feugiat in ante metus. Pretium quam
+vulputate dignissim suspendisse in.
+
+Velit aliquet sagittis id consectetur purus ut faucibus pulvinar. Venenatis a
+condimentum vitae sapien pellentesque habitant morbi tristique. Ut etiam sit
+amet nisl purus in mollis. Etiam dignissim diam quis enim. Id leo in vitae
+turpis. Et magnis dis parturient montes nascetur ridiculus. Gravida rutrum
+quisque non tellus orci. Eu consequat ac felis donec et. Egestas purus viverra
+accumsan in nisl nisi scelerisque eu. Elit ullamcorper dignissim cras tincidunt
+lobortis feugiat vivamus at. Maecenas pharetra convallis posuere morbi leo urna
+molestie. Nunc id cursus metus aliquam eleifend mi. Mauris ultrices eros in
+cursus turpis massa. Lectus nulla at volutpat diam.

### 2021-01-12 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added ECS and TCR table
Added correct ECS and TCR tables for CMIP5 and 6

@@
 \label{sec:04:evaluation_ecs_and_tcr}
+
+\begin{table}[t]
+  \centering
+  \csvreader[ECSAndTCRTable]{
+    ch04_papers_ecs_tcr_assessment/data/CMIP5_ecs-tcr.csv}{}{
+    \dataset & \idx & \ecs & \tcr}
+  \caption{\acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
+    Details on the calculation on \acs{ECS} and \acs{TCR} are given in
+    \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively.}
+  \label{tab:04:ecs_tcr_cmip5}
+\end{table}
+
+\begin{table}[t]
+  \centering
+  \csvreader[ECSAndTCRTable]{
+    ch04_papers_ecs_tcr_assessment/data/CMIP6_ecs-tcr.csv}{}{
+    \dataset & \idx & \ecs & \tcr}
+  \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.}
+  \label{tab:04:ecs_tcr_cmip6}
+\end{table}

### 2021-01-13 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added code to create CMIP6 MMM Gregory plot

@@
 quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
-on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we assess and analyze
-climate sensitivity (expressed by \ac{ECS} and \ac{TCR}) of the \acs{CMIP}6
-models and compare it to corresponding results of predecessor model
+on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we use it to assess
+and analyze climate sensitivity (expressed by \ac{ECS} and \ac{TCR}) of the
+\acs{CMIP}6 models and compare it to corresponding results of predecessor model
 generations. This work is already published in two scientific papers:
@@
 \label{sec:04:evaluation_ecs_and_tcr}
+
+Following \cref{subsec:02:ecs}, \ac{ECS} is calculated with the Gregory
+regression method using \ac{GSAT} and \ac{TOA} net radiation data for 150 years
+of a \nxcotwo{4} simulation \autocite{Gregory2004}. In contrast to that, the
+transient response of the climate system \ac{TCR} is calculated from \ac{GSAT}
+data of a \onepctcotwo{} run \commentcite{Bindoff2013}{see
+  \cref{subsec:02:tcr}}.
 

### 2021-01-14 ch01_introduction/introduction.tex
Cleaned tex files and added text to chapter 4
Adapted margins

@@
 - natural vs. anthropogenic effects (D\&A) -> clear evidence that climate
-change over 21st century is caused by human influence. The following website:
-is relevant: (https://globalwarmingindex.org/, Haustein et al., 2017)
+change over 21st century is caused by human influence. Website:
+(https://globalwarmingindex.org/, Haustein et al., 2017)
 
@@
 lead to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
-  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:assessment_climate_metrics}
+  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:papers_ecs_tcr_assessment}
 covers the assessment of climate sensitivity metrics like the \ac{ECS} or

### 2021-01-14 ch02_scientific_background/scientific_background.tex
Cleaned tex files and started chapter 04
Cleaned tex files and added text to chapter 4
Adapted margins

@@
 \subsection{Numerical Climate Modeling}
-\label{subsec:02:climate_modeling}
+\label{subsec:02:numerical_climate_modeling}
 
@@
 
-\subsection{\acs{CMIP}}
+\subsection{The \acf{CMIP}}
 \label{subsec:02:cmip}
@@
 and to facilitate model analysis. For this reason, the \ac{WGCM} of the
-\ac{WCRP} initiated the \ac{CMIP} in 1995, with the objective to
+\ac{WCRP} initiated the \acf{CMIP} in 1995, with the objective to
 \enquote{better understand past, present and future climate changes arising
@@
   \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
+  \includegraphics[width=\FigureWidth{}]{
     ch02_scientific_background/figs/cmip6.pdf}
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
-    ch02_scientific_background/figs/feedbacks_overview.pdf}
-  \caption{Climate feedbacks and corresponding time scales. \enquote{+} refers
-    to positive feedbacks, which amplify the effect of the external forcing
-    (\eg{} the water vapor feedback). \enquote{-} refers to negative feedbacks,
-    which diminish the effect of the external forcing (\eg{} the longwave
-    radiation feedback). \enquote{+/-} refers to feedbacks which might be
-    either positive or negative (\eg{} the cloud feedback). The smaller box
-    highlights the large differences in time scales for the various feedbacks.
-    Taken from \textcite{Cubasch2013}.}
-  \label{fig:02:feedbacks_overview}
-\end{figure}
-
 An example for a positive feedback is the already mentioned \emph{water vapor
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch02_scientific_background/figs/feedbacks_overview.pdf}
+  \caption{Climate feedbacks and corresponding time scales. \enquote{+} refers
+    to positive feedbacks, which amplify the effect of the external forcing
+    (\eg{} the water vapor feedback). \enquote{-} refers to negative feedbacks,
+    which diminish the effect of the external forcing (\eg{} the longwave
+    radiation feedback). \enquote{+/-} refers to feedbacks which might be
+    either positive or negative (\eg{} the cloud feedback). The smaller box
+    highlights the large differences in time scales for the various feedbacks.
+    Taken from \textcite{Cubasch2013}.}
+  \label{fig:02:feedbacks_overview}
+\end{figure}
+
 In contrast to positive feedbacks, negative feedbacks diminish the effect of an
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\FigureWidth{}]{
-    ch02_scientific_background/figs/gregory_regression_different_years.png}
-  \caption{As \cref{fig:02:gregory_regression}, but for different time periods
-    considered in the Gregory regression resulting in different values for the
-    \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
-    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
-    different time periods of a long running simulation (see legend). The
-    initial years are simulated many times for different initial conditions.
-    Light gray colors correspond to the \acs{CMIP}5 ensemble (150 years each).
-    Taken from \textcite{Knutti2017}.}
-  \label{fig:02:gregory_regression_different_years}
-\end{figure}
-
 Although commonly used in the literature, the \ac{ECS} is known to be only an
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/gregory_regression_different_years.png}
+  \caption{As \cref{fig:02:gregory_regression}, but for different time periods
+    considered in the Gregory regression resulting in different values for the
+    \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
+    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
+    different time periods of a long running simulation (see legend). The
+    initial years are simulated many times for different initial conditions.
+    Light gray colors correspond to the \acs{CMIP}5 ensemble (150 years each).
+    Taken from \textcite{Knutti2017}.}
+  \label{fig:02:gregory_regression_different_years}
+\end{figure}
+
 However, despite these deficiencies the \ac{ECS} is still a practical estimate
@@
 \autocite{Ciais2013}.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\FigureWidth{}]{
-    ch02_scientific_background/figs/carbon_cycle_schematic.png}
-  \caption{Simplified schematic of the global carbon cycle including the
-    typical turnover time scales for carbon transfers through the major
-    reservoirs (atmosphere, land surface and ocean). Taken from
-    \textcite{Ciais2013}.}
-  \label{fig:02:carbon_cycle_schematic}
-\end{figure}
 
@@
 microorganisms) \autocite{Ciais2013, Friedlingstein2019}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/carbon_cycle_schematic.png}
+  \caption{Simplified schematic of the global carbon cycle including the
+    typical turnover time scales for carbon transfers through the major
+    reservoirs (atmosphere, land surface and ocean). Taken from
+    \textcite{Ciais2013}.}
+  \label{fig:02:carbon_cycle_schematic}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
+  \includegraphics[width=\FigureWidth{}]{
     ch02_scientific_background/figs/keeling_curve.pdf}
@@
 
+The concept of \ac{MDER} is mathematically similar to the concept of emergent
+constraints. In a first step, an inter-model relationship between the target
+variable and the process-relevant diagnostics is used to fit a multivariate
+linear regression model. Let $\bm{y} = \left( y_1, y_2, \ldots, y_M \right)^T
+\in \mathbb{R}^M$ be the vector of target variables of the $M$ climate models
+($T$ denotes the transpose) and $\bm{X} \in \mathbb{R}^{M \times (K + 1)}$ the
+design matrix representing the predictors:
+\begin{equation}
+  \bm{X} =
+  \begin{pmatrix}
+    1 & x_1^{(1)} & x_1^{(2)} & \cdots & x_1^{(K)} \\
+    1 & x_2^{(1)} & x_2^{(2)} & \cdots & x_2^{(K)} \\
+    \vdots & \vdots & \vdots & \ddots & \vdots \\
+    1 & x_M^{(1)} & x_M^{(2)} & \cdots & x_M^{(K)} \\
+  \end{pmatrix}.
+  \label{eq:02:design_matrix}
+\end{equation}
+The entry $x_m^{(k)}$ of this matrix refers to the diagnostic variable of
+diagnostic $k$ and climate model $m$. With this notation, the linear
+inter-model relationship can be written as
+\begin{equation}
+  \bm{y} = \bm{X} \bm{b} + \bm{\varepsilon},
+  \label{eq:02:multivariate_linear_model}
+\end{equation}
+where $\bm{b} = \left( b_0, b_1, \ldots, b_K \right)^T \in \mathbb{R}^{(K +
+  1)}$ is the vector of linear coefficients (with intercept $b_0$) and
+$\bm{\varepsilon} = \left( \varepsilon_1, \varepsilon_2, \ldots, \varepsilon_M
+\right)^T \in \mathbb{R}^M$ a vector of independent random variables
+representing the noise in the target variable. \Cref{fig:02:mder} shows a
+schematic that illustrates this linear relationship (gray surface) for the
+different climate models (colored circles) for two diagnostics ($K = 2$). Using
+ordinary least squares regression, the estimated linear coefficients
+$\hat{\bm{b}}$ are given by
+\begin{equation}
+  \hat{\bm{b}} = \left( \bm{X}^T \bm{X} \right)^{-1} \bm{X}^T \bm{y},
+  \label{eq:02:linear_coefficients}
+\end{equation}
+where the exponent \enquote{$-1$} denotes the inverse matrix. Since this
+definition works for any number of diagnostics $K$, it can also be used to
+calculate the linear coefficients $\hat{b}_0$ (intercept) and $\hat{b}_1$
+(slope) that define emergent relationships (see
+\cref{eq:02:linear_regression_y}).
+
 \begin{figure}[t]
@@
 
-The concept of \ac{MDER} is mathematically similar to the concept of emergent
-constraints. In a first step, an inter-model relationship between the target
-variable and the process-relevant diagnostics is used to fit a multivariate
-linear regression model. Let $\bm{y} = \left( y_1, y_2, \ldots, y_M \right)^T
-\in \mathbb{R}^M$ be the vector of target variables of the $M$ climate models
-($T$ denotes the transpose) and $\bm{X} \in \mathbb{R}^{M \times (K + 1)}$ the
-design matrix representing the predictors:
-\begin{equation}
-  \bm{X} =
-  \begin{pmatrix}
-    1 & x_1^{(1)} & x_1^{(2)} & \cdots & x_1^{(K)} \\
-    1 & x_2^{(1)} & x_2^{(2)} & \cdots & x_2^{(K)} \\
-    \vdots & \vdots & \vdots & \ddots & \vdots \\
-    1 & x_M^{(1)} & x_M^{(2)} & \cdots & x_M^{(K)} \\
-  \end{pmatrix}.
-  \label{eq:02:design_matrix}
-\end{equation}
-The entry $x_m^{(k)}$ of this matrix refers to the diagnostic variable of
-diagnostic $k$ and climate model $m$. With this notation, the linear
-inter-model relationship can be written as
-\begin{equation}
-  \bm{y} = \bm{X} \bm{b} + \bm{\varepsilon},
-  \label{eq:02:multivariate_linear_model}
-\end{equation}
-where $\bm{b} = \left( b_0, b_1, \ldots, b_K \right)^T \in \mathbb{R}^{(K +
-  1)}$ is the vector of linear coefficients (with intercept $b_0$) and
-$\bm{\varepsilon} = \left( \varepsilon_1, \varepsilon_2, \ldots, \varepsilon_M
-\right)^T \in \mathbb{R}^M$ a vector of independent random variables
-representing the noise in the target variable. \Cref{fig:02:mder} shows a
-schematic that illustrates this linear relationship (gray surface) for the
-different climate models (colored circles) for two diagnostics ($K = 2$). Using
-ordinary least squares regression, the estimated linear coefficients
-$\hat{\bm{b}}$ are given by
-\begin{equation}
-  \hat{\bm{b}} = \left( \bm{X}^T \bm{X} \right)^{-1} \bm{X}^T \bm{y},
-  \label{eq:02:linear_coefficients}
-\end{equation}
-where the exponent $-1$ denotes the inverse matrix. Since this definition works
-for any number of diagnostics $K$, it can also be used to calculate the linear
-coefficients $\hat{b}_0$ (intercept) and $\hat{b}_1$ (slope) that define
-emergent relationships (see \cref{eq:02:linear_regression_y}).
-
-In a second step, observed data of the process-based diagnostics $\bm{x}_0 =
+In the second step of the algorithm, observed data of the process-based diagnostics $\bm{x}_0 =
 \left( 1, x_0^{(1)}, x_0^{(2)}, \ldots, x_0^{(K)} \right)^T \in \mathbb{R}^{(K

### 2021-01-14 ch03_esmvaltool/esmvaltool.tex
Adapted margins

@@
 Since its first release in 2016 \autocite{Eyring2016a}, the \ac{ESMValTool} has
-been greatly extended and improved. A schematic of the current structure of the
-\ac{ESMValTool} is illustrated in \cref{fig:03:esmvaltool_schematic}. The first
-major element of this workflow is an extensive preprocessing of the input data.
-For this, the Python-based \acf{ESMValCore} provide a set of operators
-(\emph{preprocessors}) that are applied to the input data. There are two
-classes of preprocessors: non-optional and optional ones. Examples for
-non-optional preprocessors are the \ac{I/O} operations \emph{load} (loading
-input data from climate models and observations) and \emph{save} (saving the
-preprocessed data), which utilize the \ac{netCDF} format (a self-describing,
-machine-independent binary file format used for the storage of array-based
-scientific data). Further non-optional preprocessors include \emph{checks} and
-\emph{fixes} of the input data, which test whether the input data adheres to
-standards given by the \ac{CMOR} format and fixes the data if necessary. The
-\ac{CMOR} format ensures that the output from the many different modeling
-institutes within a generation of \ac{CMIP} follows common standards. The
-optional preprocessors include commonly used operations on spatiotemporal
-datasets, like spatial and temporal \emph{subsetting}, the calculation of
-\emph{spatial and temporal statistics}, horizontal and vertical
-\emph{interpolation}, land/sea/ice \emph{masking}, \emph{unit conversion}, or
-the calculation of \emph{multi-model statistics}. A further example is the
-\emph{variable derivation}, which can be used to derive non-\ac{CMOR} variables
-from the input data. The aim of the preprocessor is to facilitate the routine
-evaluation of climate models by providing a set of commonly used data
-operations. To reduce computation times, \ac{ESMValCore} allows parallel
-preprocessing of different datasets.
+been greatly extended and improved. A schematic representation of the current
+structure of the \ac{ESMValTool} is illustrated in
+\cref{fig:03:esmvaltool_schematic}. The first major element of this workflow is
+an extensive preprocessing of the input data. For this, the Python-based
+\acf{ESMValCore} provide a set of operators (\emph{preprocessors}) that are
+applied to the input data. There are two classes of preprocessors: non-optional
+and optional ones. Examples for non-optional preprocessors are the \ac{I/O}
+operations \emph{load} (loading input data from climate models and
+observations) and \emph{save} (saving the preprocessed data), which utilize the
+\ac{netCDF} format (a self-describing, machine-independent binary file format
+used for the storage of array-based scientific data). Further non-optional
+preprocessors include \emph{checks} and \emph{fixes} of the input data, which
+test whether the input data adheres to standards given by the \ac{CMOR} format
+and fixes the data if necessary. The \ac{CMOR} format ensures that the output
+from the many different modeling institutes within a generation of \ac{CMIP}
+follows common standards. The optional preprocessors include commonly used
+operations on spatiotemporal datasets, like spatial and temporal
+\emph{subsetting}, the calculation of \emph{spatial and temporal statistics},
+horizontal and vertical \emph{interpolation}, land/sea/ice \emph{masking},
+\emph{unit conversion}, or the calculation of \emph{multi-model statistics}. A
+further example is the \emph{variable derivation}, which can be used to derive
+non-\ac{CMOR} variables from the input data. The aim of the preprocessor is to
+facilitate the routine evaluation of climate models by providing a set of
+commonly used data operations. To reduce computation times, \ac{ESMValCore}
+allows parallel preprocessing of different datasets. However, since this leads
+to a higher memory usage, a compromise has to be found in practice to optimize
+the use of time and memory resources.
 
@@
 
-The \ac{ESMValTool} is developed open-source on GitHub
+The \ac{ESMValTool} is developed open-source on GitHub, a web service that
+provides hosting for version control with \emph{git}
 (\url{https://github.com/ESMValGroup}). It is released under the Apache
@@
     Preprocessor & \emph{amplitude} & Amplitude of periodic phenomena (\eg{}
-    diurnal or seasonal cycles) \\
+    cycles) \\
     \cmidrule(l){2-3}
@@
     \cmidrule(l){2-3}
-    & \emph{trend} & Linear trend (slope of ordinary least squares regression)
-    \\
+    & \emph{trend} & Linear trend (slope of linear regression) \\
     \midrule
@@
 preprocessor calculates the peak-to-peak amplitude of periodic phenomena along
-an arbitrary coordinate. Usually, this coordinate is a temporal dimension. A
-common application of this preprocessor is the calculation of a variable's
-diurnal or seasonal cycle amplitude. The \emph{land/sea fraction weighting}
-preprocessor weights fields with the land or sea fraction of the respective
-grid cells. For example, this is necessary for the spatial integration of
-flux-related variables which are reported in units of \enquote{per square meter
-  of land/sea} and not in \enquote{per square meter of grid cell}. After the
-weighting, the grid cell areas can be used to integrate the flux-related
-variable over a desired region. The \emph{trend} preprocessor calculates the
-linear trend of a variable along an arbitrary coordinate. The linear trend is
-defined as the slope of an ordinary least squares linear regression of the
-variable against the selected coordinate. For example, this can be used to
-calculate the temporal trend of the \ac{GSAT} over the \nth{20} century.
+an arbitrary coordinate, which is usually time. A common application of this
+preprocessor is the calculation of a variable's diurnal or seasonal cycle
+amplitude. The \emph{land/sea fraction weighting} preprocessor weights fields
+with the land or sea fraction of the respective grid cells. For example, this
+is necessary for the spatial integration of flux-related variables which are
+reported in units of \enquote{per square meter of land/sea} and not in
+\enquote{per square meter of grid cell}. After the weighting, the grid cell
+areas can be used to integrate the flux-related variable over a desired region.
+The \emph{trend} preprocessor calculates the linear trend of a variable along
+an arbitrary coordinate. The linear trend is defined as the slope of an
+ordinary least squares linear regression of the variable against the selected
+coordinate. For example, this can be used to calculate the temporal trend of
+the \ac{GSAT} over the \nth{20} century.
 
@@
   \end{subfigure}
-  \caption{(a) \acs{CMIP}6 \acf{MMM} of the global atmospheric \acs{CO2}
-    concentration at surface (\emph{co2s}) averaged over the months \acf{JJA}
-    of the year 2014 in the emission-driven historical simulation. The
-    \acs{CMIP}6 \acs{MMM} includes the four climate models that provide all
-    data needed for the calculation of \emph{co2s} (see legend in (b)). (b)
-    Monthly-mean \emph{co2s} at Mauna Loa, Hawaii ($19.5 \unit{\degree N}$,
-    $155.6 \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2014.
-    The thick black line shows observations from the Mauna Loa Observatory
-    \autocite{Keeling2005}; the remaining colored lines show emission-driven
-    historical simulations from individual \acs{CMIP}6 models.}
+  \caption{(a) \acs{CMIP}6 \acf{MMM} of the atmospheric \acs{CO2} concentration
+    at surface (\emph{co2s}) averaged over the months \acf{JJA} of the year
+    2014 in the emission-driven historical simulation. The \acs{CMIP}6
+    \acs{MMM} includes the four climate models that provide all data needed
+    for the calculation of \emph{co2s} (see legend in (b)). (b) Monthly-mean
+    \emph{co2s} at Mauna Loa, Hawaii ($19.5 \unit{\degree N}$, $155.6
+    \unit{\degree W}$) from 1958 to 2014. The black line shows observations
+    from the Mauna Loa Observatory \autocite{Keeling2005}; the remaining lines
+    show emission-driven historical simulations from individual \acs{CMIP}6
+    models.}
   \label{fig:03:co2s}
@@
 more recipes and \ac{CMOR}ization scripts for observational data have been
-added to the main \ac{ESMValTool} repository on GitHub. Moreover, multiple
-minor changes and additions have been implemented to improve the code base
-and/or to fix bugs. In total, the author of this study has added 88668 lines of
-code and removed 47816 lines of code from the \ac{ESMValTool} repository (as of
+added to the main \ac{ESMValTool} repository on GitHub. A summary of these is
+given in \cref{tab:03:changes_to_esmvaltool}. Moreover, multiple minor changes
+and additions have been implemented to improve the code base and/or to fix
+bugs. In total, the author of this study has added 88668 lines of code and
+removed 47816 lines of code from the \ac{ESMValTool} repository (as of
 \TheMonth{}). Parts of these implementations are already published in the
@@
 
-The implemented \ac{CMOR}ization scripts (also referred to as
-\emph{\ac{CMOR}izers}) allow the processing of arbitrary observational data
-with the \ac{ESMValTool}. For this, the raw observational datasets are
-reformatted in such way that they adhere to the \ac{CMOR} standards. Since the
-\ac{CMOR}ization scripts are publicly available like any other code in the
-\ac{ESMValTool} and \ac{ESMValCore} repositories, the entire scientific
-community can use them to \ac{CMOR}ize the observational data. A complete list
-of all eleven newly implemented \ac{CMOR}izers including the main reference for
-the corresponding observational datasets is shown in
-\cref{tab:03:changes_to_esmvaltool}.
+\begin{table}[t]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Type & Name & Main reference \\
+    \midrule
+    Recipe & \emph{bock20jgr} (figures 8 and 10) & \autocite{Bock2020} \\
+    & \emph{cox18nature} & \autocite{Cox2018} \\
+    & \emph{ecs} & \autocite{Gregory2004} \\
+    & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
+    & \emph{schlund20esd} & \autocite{Schlund2020a} \\
+    & \emph{schlund20jgr\_gpp\_abs\_rcp85} & \autocite{Schlund2020} \\
+    & \emph{schlund20jgr\_gpp\_change\_1pct} & \autocite{Schlund2020} \\
+    & \emph{schlund20jgr\_gpp\_change\_rcp85} & \autocite{Schlund2020} \\
+    & \emph{tcr} & \autocite{Gregory2008} \\
+    \midrule
+    \ac{CMOR}izer & CRU & \autocite{Harris2014} \\
+    & CT2019 & \autocite{Jacobson2020} \\
+    & GCP & \autocite{Friedlingstein2019} \\
+    & HWSD & \autocite{Wieder2014} \\
+    & JMA-TRANSCOM & \autocite{Maki2010} \\
+    & LAI3g & \autocite{Zhu2013} \\
+    & LandFlux-EVAL & \autocite{Mueller2013} \\
+    & MLS-Aura & \autocite{Read2015} \\
+    & MTE & \autocite{Jung2011} \\
+    & NDP & \autocite{Gibbs2006} \\
+    & Scripps-\ac{CO2} & \autocite{Keeling2005} \\
+    \bottomrule
+  \end{tabulary}
+  \caption{Summary of new recipes and \acs{CMOR}ization scripts for
+    observational data contributed to the \acf{ESMValTool} by the author of
+    this thesis. The corresponding file names of the recipes in the
+    \acs{ESMValTool} repository are given by
+    \emph{recipe\_(name\_in\_table).yml}.}
+  \label{tab:03:changes_to_esmvaltool}
+\end{table}
 
@@
 and predictors.
-
-\begin{table}[t]
-  \centering
-  \begin{tabulary}{\columnwidth}{L L L}
-    \toprule
-    Type & Name & Main reference \\
-    \midrule
-    \ac{CMOR}izer & CRU & \autocite{Harris2014} \\
-    & CT2019 & \autocite{Jacobson2020} \\
-    & GCP & \autocite{Friedlingstein2019} \\
-    & HWSD & \autocite{Wieder2014} \\
-    & JMA-TRANSCOM & \autocite{Maki2010} \\
-    & LAI3g & \autocite{Zhu2013} \\
-    & LandFlux-EVAL & \autocite{Mueller2013} \\
-    & MLS-Aura & \autocite{Read2015} \\
-    & MTE & \autocite{Jung2011} \\
-    & NDP & \autocite{Gibbs2006} \\
-    & Scripps-\ac{CO2} & \autocite{Keeling2005} \\
-    \midrule
-    Recipe & \emph{bock20jgr} (figures 8 and 10) & \autocite{Bock2020} \\
-    & \emph{cox18nature} & \autocite{Cox2018} \\
-    & \emph{ecs} & \autocite{Gregory2004} \\
-    & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
-    & \emph{schlund20esd} & \autocite{Schlund2020a} \\
-    & \emph{schlund20jgr\_gpp\_abs\_rcp85} & \autocite{Schlund2020} \\
-    & \emph{schlund20jgr\_gpp\_change\_1pct} & \autocite{Schlund2020} \\
-    & \emph{schlund20jgr\_gpp\_change\_rcp85} & \autocite{Schlund2020} \\
-    & \emph{tcr} & \autocite{Gregory2008} \\
-    \bottomrule
-  \end{tabulary}
-  \caption{Summary of new \acs{CMOR}ization scripts for observational data
-    and recipes contributed to the \acf{ESMValTool} by the author of this
-    thesis. The corresponding file names of the recipes in the
-    \acs{ESMValTool} repository are given by
-    \emph{recipe\_(name\_in\_table).yml}.}
-  \label{tab:03:changes_to_esmvaltool}
-\end{table}
-
-In addition to the ported recipes, several new recipes and diagnostic scripts
-have been contributed to the public \ac{ESMValTool} repository (see
-\cref{tab:03:changes_to_esmvaltool} for a summary). Some of these include the
-analyses performed in other chapters of this thesis, namely
-\emph{recipe\_bock20jgr.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
-\emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and
-\emph{recipe\_schlund20jgr\_*.yml} (\cref{ch:06:paper_gpp}).
-\emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml} contain diagnostic scripts
-that are able to calculate the climate metrics \ac{ECS} and \ac{TCR} for
-arbitrary model output (see \cref{sec:02:climate_sensitivity}). Corresponding
-plots are shown in \cref{fig:02:gregory_regression,fig:02:tcr}.
-\emph{recipe\_cox18nature.yml} reproduces the analysis of \textcite{Cox2018},
-who introduce an emergent constraint in \ac{ECS} based on a global temperature
-variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
-years for \acs{CMIP}5 models and observations form HadCRUT4
-\autocite{Morice2012}. Since high sensitivity climate models exhibit higher
-values of $\Psi$ than low sensitivity models, an emergent relationship between
-\ac{ECS} and $\Psi$ can be established (\cref{fig:03:cox18:b}). With
-observations of $\Psi$, \ac{ECS} can eventually be constrained to $\left( 2.8
-\pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). More details on
-this including the calculation of $\Psi$ are given in \cref{subsec:05:cox}.
 
@@
 \end{figure}
+
+In addition to the ported recipes, several new recipes and diagnostic scripts
+have been contributed to the public \ac{ESMValTool} repository (see
+\cref{tab:03:changes_to_esmvaltool} for a summary). Some of these include the
+analyses performed in other chapters of this thesis, namely
+\emph{recipe\_bock20jgr.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
+\emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and
+\emph{recipe\_schlund20jgr\_*.yml} (\cref{ch:06:paper_gpp}).
+\emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml} contain diagnostic scripts
+that are able to calculate the climate metrics \ac{ECS} and \ac{TCR} for
+arbitrary model output (see \cref{sec:02:climate_sensitivity}). Corresponding
+plots are shown in \cref{fig:02:gregory_regression,fig:02:tcr}.
+\emph{recipe\_cox18nature.yml} reproduces the analysis of \textcite{Cox2018},
+who introduce an emergent constraint in \ac{ECS} based on a global temperature
+variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
+years for \acs{CMIP}5 models and observations form HadCRUT4
+\autocite{Morice2012}. Since high sensitivity climate models exhibit higher
+values of $\Psi$ than low sensitivity models, an emergent relationship between
+\ac{ECS} and $\Psi$ can be established (\cref{fig:03:cox18:b}). With
+observations of $\Psi$, \ac{ECS} can eventually be constrained to $\left( 2.8
+\pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). More details on
+this including the calculation of $\Psi$ are given in \cref{subsec:05:cox}.
 
@@
 \end{figure}
+
+The implemented \ac{CMOR}ization scripts (also referred to as
+\emph{\ac{CMOR}izers}) allow the processing of arbitrary observational data
+with the \ac{ESMValTool}. For this, the raw observational datasets are
+reformatted in such way that they adhere to the \ac{CMOR} standards. Since the
+\ac{CMOR}ization scripts are publicly available like any other code in the
+\ac{ESMValTool} and \ac{ESMValCore} repositories, the entire scientific
+community can use them to \ac{CMOR}ize the observational data. A complete list
+of all eleven newly implemented \ac{CMOR}izers including the main reference for
+the corresponding observational datasets is shown in
+\cref{tab:03:changes_to_esmvaltool}.

### 2021-01-15 ch03_esmvaltool/esmvaltool.tex
Added text and figures to section 4.1

@@
 example used to evaluate climate model output and observational data in several
-chapters of the upcoming \ac{AR6} of the \ac{IPCC}.
+chapters of the upcoming Sixth \acl{AR} (\acs{AR}6) of the \ac{IPCC}.
 
@@
 
-\begin{table}[t]
+\begin{table}[!b]
   \centering
@@
 
-An important class of recipes in the \ac{ESMValTool} covers international climate assessments.
-One such example is \emph{recipe\_flato13ipcc.yml}, which reproduces large parts of the climate model evaluation performed in chapter 9 of the \ac{IPCC}'s \ac{AR5} \autocite{Flato2013}.
-For this recipe, the two-panel figure 9.42 has been added, which shows the relationship between the historical and pre-industrial \ac{GSAT} and \ac{ECS} (\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS} (\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
-\Cref{fig:03:flato13:a} illustrates that there is no clear connection between \ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high temperatures in the historical or pre-industrial period do no necessarily exhibit a high equilibrium warming in the future or vice versa.
-On the other hand, \ac{TCR} and \ac{ECS} are in fact well-correlated within the \acs{CMIP}5 ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}.
-However, this figure only shows a linear regression fit between the two variables and not the expected reciprocal relation.
+An important class of recipes in the \ac{ESMValTool} covers international
+climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which
+reproduces large parts of the climate model evaluation performed in chapter 9
+of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
+recipe, the two-panel figure 9.42 has been added, which shows the relationship
+between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
+(\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS}
+(\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
+\Cref{fig:03:flato13:a} illustrates that there is no clear connection between
+\ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high
+temperatures in the historical or pre-industrial period do no necessarily
+exhibit a high equilibrium warming in the future or vice versa. On the other
+hand, \ac{TCR} and \ac{ECS} are in fact well-correlated within the \acs{CMIP}5
+ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this
+figure only shows a linear regression fit between the two variables and not the
+expected reciprocal relation.
 

### 2021-01-15 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added text and figures to section 4.1

@@
 released every day (as of \TheMonth{}). In light of the upcoming international
-climate assessment of the \ac{IPCC} (\ac{AR6}), the evaluation of
+climate assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of
 policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6
@@
 generations. This work is already published in two scientific papers:
-\textcite{Bock2020} and \textcite{Meehl2020}.
+\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
+author of this thesis contributed their figure 8 (a bar chart showing the
+\ac{ECS} for the three climate model generations \acs{CMIP}3, \acs{CMIP}5 and
+\acs{CMIP}6), their figure 10 (map plots showing cloud-related feedback
+parameters for the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
+\cref{fig:04:cloud_feedback_parameters}) and discussion about \ac{ECS} of the
+\acs{CMIP}6 models. For \textcite{Meehl2020}, the author of this thesis
+contributed all figures and tables of the study and details on the calculations
+of \ac{ECS} and \ac{TCR}.
 
-\section{Evaluation of \acs{ECS} and \acs{TCR} in \acs{CMIP}6}
+\section{Evaluation of \acs{ECS} and \acs{TCR} over the Years}
 \label{sec:04:evaluation_ecs_and_tcr}
@@
 transient response of the climate system \ac{TCR} is calculated from \ac{GSAT}
-data of a \onepctcotwo{} run \commentcite{Bindoff2013}{see
-  \cref{subsec:02:tcr}}.
+data at the time of \ac{CO2} doubling in a \onepctcotwo{} run
+\commentcite{Bindoff2013}{see \cref{subsec:02:tcr}}. Using these definitions,
+we assessed \ac{ECS} and \ac{TCR} for all \acs{CMIP}5 and \acs{CMIP}6 models
+where the necessary data is available (as of \TheMonth{}) using the
+\ac{ESMValTool}. \Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an
+overview over the results for the \acs{CMIP}5 and \acs{CMIP}6 models,
+respectively.
 
-\begin{table}[t]
+\begin{table}[!t]
   \centering
@@
 
-\begin{table}[t]
+\begin{table}[p]
   \centering
@@
 \end{table}
+
+The first striking feature of these two tables is the increased \ac{MMM} of
+\ac{ECS} and \ac{TCR} in \acs{CMIP}6. For \ac{ECS}, the \acs{CMIP}6 \ac{MMM} is
+about $16 \unit{\%}$ ($0.51 \unit{K}$) higher than the corresponding
+\acs{CMIP}5 \ac{MMM}. For \ac{TCR}, the relative difference between the two
+model ensembles is notably smaller with about $10 \unit{\%}$ ($0.19 \unit{K}$).
+The spread in the multi-model ensembles (expressed as the multi-model standard
+deviation) shows an even larger increase in the \acs{CMIP}6 ensemble: For
+\ac{ECS}, the relative difference is about $49 \unit{\%}$ ($0.72 \unit{K}$ in
+\acs{CMIP}5 to $1.07 \unit{K}$ in \acs{CMIP}6) and for \ac{TCR}, the relative
+difference is about $17 \unit{\%}$ ($0.36 \unit{K}$ in \acs{CMIP}5 to $0.42
+\unit{K}$ in \acs{CMIP}6). The main reason for the increased \ac{MMM} and
+spread in the \acs{CMIP}6 ensemble is the existence of several models with very
+high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exists a
+number of models with very low values of \ac{ECS}. Therefore, the \ac{CMIP}6
+model ranges of \ac{ECS} and \ac{TCR} are well outside the corresponding
+assessed ranges given by the latest published \acl{AR} of the \ac{IPCC} from
+2013 \autocite{Stocker2013} with $\range{1.8}{5.6} \unit{K}$ (\acs{AR}5:
+$\range{1.5}{4.5} \unit{K}$) and  $\range{1.3}{3.0} \unit{K}$ (\acs{AR}5:
+$\range{1.0}{2.5} \unit{K}$) for \ac{ECS} and \ac{TCR}, respectively.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch04_papers_ecs_tcr_assessment/figs/historical_ecs_tcr_ranges.pdf}
+  \caption{Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red
+    bars) over the years in the Charney report from 1979 \autocite{Charney1979}
+    and the subsequent \acp{AR} of the \acf{IPCC}. Orange and green bars
+    represent the modeled ranges of \acs{ECS} and \acs{TCR}, respectively, from
+    the different climate model ensembles of the \acf{CMIP} with the
+    corresponding \acf{MMM} illustrated with the horizontal black lines (the
+    green dot represents output from a single climate model). The numbers
+    correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
+    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
+    \textcite{Meehl2020}.}
+  \label{fig:04:historical_ecs_tcr_ranges}
+\end{figure}
+
+To illustrate this and put it into historical context,
+\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
+\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
+and the different \acp{AR} of the \ac{IPCC} in combination with the
+corresponding modeled ranges from the different \ac{CMIP} generations. Since
+the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
+\unit{K}$ has almost remained unchanged during the last 40 years (with the
+exception of \acs{AR}4, where the lower bound was temporarily increased to $2.0
+\unit{K}$). For \acs{AR}2, \acs{AR}4 and \acs{AR}5, the corresponding climate
+model generations from \acs{CMIP}1, \acs{CMIP}5 and \acs{CMIP}6 more or less
+agree with this assessed range. In contrast to that, some of the early climate
+models used in \acs{AR}1 and the \acs{CMIP}2 models used in \acs{AR}3 exhibit
+values of \acs{ECS} above $4.5 \unit{K}$, resulting in a upper model range of
+about $0.6 \unit{K}$ higher than the assessed range. However, these deviations
+from the assessed range are negligible to the ones present in the \acs{CMIP}6
+ensemble, which have become evident over the course of 2020. As shown in
+\cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper bounds are
+more extreme than in any previous climate model generation, resulting in a
+\acs{CMIP}6 model range of $\range{1.8}{5.6} \unit{K}$ (as of \TheMonth{}).
+This increase in range is particularly relevant for the upper bound, which is
+about $1.1 \unit{K}$ higher than the assessed upper bound of $4.5 \unit{K}$.
+Moreover, not only a small fraction of models exceeds this upper bounds: out of
+the 42 analyzed \ac{CMIP}6 models, a third (14 models) has an \ac{ECS} above
+$4.5 \unit{K}$. Because of the massive political and societal relevance of
+\ac{ECS}, its apparent increase in the \ac{CMIP}6 ensemble is currently one of
+the most important questions for the entire climate modeling community. A list
+of possible reasons for this increased climate sensitivity is given in
+\cref{sec:04:possible_reasons_high_ecs_cmip6}.
+
+\section{Possible Reasons for High Climate Sensitivity in \acs{CMIP}6}
+\label{sec:04:possible_reasons_high_ecs_cmip6}

### 2021-01-18 ch02_scientific_background/scientific_background.tex
Finished section 4.1 and added figures for section 4.2

@@
     years of a simulation with an abrupt quadrupling of the atmospheric
-    \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 multi-model mean
-    (blue circles). To account for energy leakage and model drift, a linear
-    fit of the corresponding pre-industrial control run is subtracted from the
+    \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 \acl{MMM} (blue
+    circles). To account for energy leakage and model drift, a linear fit of
+    the corresponding pre-industrial control run is subtracted from the
     \nxcotwo{4} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope
@@
     ch02_scientific_background/figs/gregory_regression_different_years.png}
-  \caption{As \cref{fig:02:gregory_regression}, but for different time periods
-    considered in the Gregory regression resulting in different values for the
-    \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
-    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
-    different time periods of a long running simulation (see legend). The
+  \caption{As in \cref{fig:02:gregory_regression} but for different time
+    periods considered in the Gregory regression resulting in different values
+    for the \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray,
+    blue, red and orange colors correspond to the \acf{CESM} of the \acf{NCAR}
+    for different time periods of a long running simulation (see legend). The
     initial years are simulated many times for different initial conditions.

### 2021-01-18 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Finished section 4.1 and added figures for section 4.2

@@
 
-\section{Evaluation of \acs{ECS} and \acs{TCR} over the Years}
+\section{Evaluation of \acs{ECS} and \acs{TCR} in \acs{CMIP}6}
 \label{sec:04:evaluation_ecs_and_tcr}
@@
 regression method using \ac{GSAT} and \ac{TOA} net radiation data for 150 years
-of a \nxcotwo{4} simulation \autocite{Gregory2004}. In contrast to that, the
-transient response of the climate system \ac{TCR} is calculated from \ac{GSAT}
-data at the time of \ac{CO2} doubling in a \onepctcotwo{} run
-\commentcite{Bindoff2013}{see \cref{subsec:02:tcr}}. Using these definitions,
-we assessed \ac{ECS} and \ac{TCR} for all \acs{CMIP}5 and \acs{CMIP}6 models
-where the necessary data is available (as of \TheMonth{}) using the
-\ac{ESMValTool}. \Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an
-overview over the results for the \acs{CMIP}5 and \acs{CMIP}6 models,
-respectively.
+of a \nxcotwo{4} simulation \autocite{Gregory2004}. This calculation is
+illustrated in \cref{fig:04:gregory_regression_cmip6_mmm} for the \acs{CMIP}6
+\ac{MMM}, which yields an \ac{ECS} of $3.74 \unit{K}$ when all 150 years of the
+run are used. However, similar to the \acs{CMIP}5 ensemble, the exact value of
+\ac{ECS} depends on the years considered in the Gregory regression (see
+\cref{fig:02:gregory_regression_different_years}). Using only the first 20
+years of the simulation gives a significantly lower \ac{ECS} of $3.31 \unit{K}$
+than using only the last 130 years of the simulation, which gives an \ac{ECS}
+of $4.05 \unit{K}$. As thoroughly describes in \cref{subsec:02:ecs}, the reason
+for this is the state and time dependence of the climate feedback parameter,
+which is given by the slope of the Gregory regression line. Due to non-linear
+effects in the feedbacks, this slope changes over time, resulting in lower
+(higher) values of \ac{ECS} when only early (late) years of the simulation are
+considered.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch04_papers_ecs_tcr_assessment/figs/gregory_regression_cmip6_mmm.pdf}
+  \caption{Gregory regression for the \acs{CMIP}6 \acf{MMM} following
+    \textcite{Gregory2004}: Global and annual mean net \acf{TOA} radiation $N$
+    versus the change in global and annual mean near-surface air temperature
+    $\Delta T$ for 150 years of a simulation with an abrupt quadrupling of the
+    atmospheric \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}6
+    \acs{MMM} (circles). To account for energy leakage and model drift, a
+    linear fit of the corresponding pre-industrial control run is subtracted
+    from the \nxcotwo{4} simulation. A linear regression can be used to
+    calculate the \acf{ECS} according to \cref{eq:02:ecs} with the radiative
+    forcing $F_\text{4x}$ (\yintercept) and the climate feedback parameter
+    $\lambda$ (slope) for all 150 years (black line), which results in
+    $\text{ECS} = 3.74 \unit{K}$. Due to non-linear effects, the exact value
+    of \acs{ECS} depends on the years considered for the Gregory regression
+    (see \cref{subsec:02:ecs} and
+    \cref{fig:02:gregory_regression_different_years}): Using only the first 20
+    years of the simulation (blue circles and line) yields a considerably
+    smaller \acs{ECS} than using only the last 130 years of the simulation
+    (orange circles and line). Updated from \textcite{Meehl2020}.}
+  \label{fig:04:gregory_regression_cmip6_mmm}
+\end{figure}
+
+In contrast to that, the transient response of the climate system \ac{TCR} is
+calculated from \ac{GSAT} data at the time of \ac{CO2} doubling in a
+\onepctcotwo{} run \commentcite{Bindoff2013}{see \cref{subsec:02:tcr}}. Using
+these two definitions and the \ac{ESMValTool}, we assess \ac{ECS} and \ac{TCR}
+for all \acs{CMIP}5 and \acs{CMIP}6 models where the necessary temperature and
+radiation data is available (as of \TheMonth{}).
+\Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an overview over the
+results for the \acs{CMIP}5 and \acs{CMIP}6 models, respectively.
 
@@
   \caption{\acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
-    Details on the calculation on \acs{ECS} and \acs{TCR} are given in
-    \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively.}
+    Details on the calculation of \acs{ECS} and \acs{TCR} are given in
+    \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively. The \acf{MMM}
+    is calculated from the Gregory regression method using the \acs{MMM} net
+    \acf{TOA} radiation and the \acs{MMM} change in \acf{GSAT} similar to
+    \cref{fig:04:gregory_regression_cmip6_mmm}. The multi-model standard
+    deviation is given by the sample standard deviation of \acs{ECS} evaluated
+    over all climate models (using the normalization $1 / M$, where $M$ is the
+    number of models).}
   \label{tab:04:ecs_tcr_cmip5}
@@
 model ranges of \ac{ECS} and \ac{TCR} are well outside the corresponding
-assessed ranges given by the latest published \acl{AR} of the \ac{IPCC} from
-2013 \autocite{Stocker2013} with $\range{1.8}{5.6} \unit{K}$ (\acs{AR}5:
-$\range{1.5}{4.5} \unit{K}$) and  $\range{1.3}{3.0} \unit{K}$ (\acs{AR}5:
-$\range{1.0}{2.5} \unit{K}$) for \ac{ECS} and \ac{TCR}, respectively.
+assessed ranges given by the latest published \ac{AR} of the \ac{IPCC} from
+2013 \autocite{Stocker2013} with $\range{1.8}{5.6} \unit{K}$ for \ac{ECS}
+(\acs{AR}5: $\range{1.5}{4.5} \unit{K}$) and  $\range{1.3}{3.0} \unit{K}$
+(\acs{AR}5: $\range{1.0}{2.5} \unit{K}$) for \ac{TCR}.
 
@@
     bars) over the years in the Charney report from 1979 \autocite{Charney1979}
-    and the subsequent \acp{AR} of the \acf{IPCC}. Orange and green bars
-    represent the modeled ranges of \acs{ECS} and \acs{TCR}, respectively, from
-    the different climate model ensembles of the \acf{CMIP} with the
-    corresponding \acf{MMM} illustrated with the horizontal black lines (the
-    green dot represents output from a single climate model). The numbers
-    correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
+    and the subsequent \acp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
+      Kattenberg1996, Albritton2001, Solomon2007, Stocker2013}. Orange and
+    green bars represent the modeled ranges of \acs{ECS} and \acs{TCR},
+    respectively, from the different climate model ensembles of the \acf{CMIP}
+    with the corresponding \acf{MMM} illustrated with the horizontal black
+    lines (the green dot represents output from a single climate model). The
+    numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
     \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
@@
 the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
-\unit{K}$ has almost remained unchanged during the last 40 years (with the
-exception of \acs{AR}4, where the lower bound was temporarily increased to $2.0
-\unit{K}$). For \acs{AR}2, \acs{AR}4 and \acs{AR}5, the corresponding climate
-model generations from \acs{CMIP}1, \acs{CMIP}5 and \acs{CMIP}6 more or less
-agree with this assessed range. In contrast to that, some of the early climate
-models used in \acs{AR}1 and the \acs{CMIP}2 models used in \acs{AR}3 exhibit
-values of \acs{ECS} above $4.5 \unit{K}$, resulting in a upper model range of
-about $0.6 \unit{K}$ higher than the assessed range. However, these deviations
-from the assessed range are negligible to the ones present in the \acs{CMIP}6
+\unit{K}$ has almost remained unchanged during the last 40 years
+\autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
+  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
+temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
+\acs{AR}4 and \acs{AR}5, the corresponding climate model generations from
+\acs{CMIP}1, \acs{CMIP}3 and \acs{CMIP}5 more or less agree with this assessed
+range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
+some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
+used in \acs{AR}3 exhibit values of \acs{ECS} above $4.5 \unit{K}$, resulting
+in a upper model range of about $0.6 \unit{K}$ higher than the assessed range
+\autocite{Mitchell1990, Cubasch2001}. However, these deviations from the
+assessed range are insignificant to the ones present in the \acs{CMIP}6
 ensemble, which have become evident over the course of 2020. As shown in
@@
 more extreme than in any previous climate model generation, resulting in a
-\acs{CMIP}6 model range of $\range{1.8}{5.6} \unit{K}$ (as of \TheMonth{}).
-This increase in range is particularly relevant for the upper bound, which is
-about $1.1 \unit{K}$ higher than the assessed upper bound of $4.5 \unit{K}$.
-Moreover, not only a small fraction of models exceeds this upper bounds: out of
-the 42 analyzed \ac{CMIP}6 models, a third (14 models) has an \ac{ECS} above
-$4.5 \unit{K}$. Because of the massive political and societal relevance of
-\ac{ECS}, its apparent increase in the \ac{CMIP}6 ensemble is currently one of
-the most important questions for the entire climate modeling community. A list
-of possible reasons for this increased climate sensitivity is given in
+\acs{CMIP}6  \ac{ECS} model range of $\range{1.8}{5.6} \unit{K}$ (as of
+\TheMonth{}). This increase in range is particularly relevant for the upper
+bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
+$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that
+exceeds this upper bound: out of the 42 analyzed \ac{CMIP}6 models, a third (14
+models) has an \ac{ECS} above $4.5 \unit{K}$. Because of the massive political
+and societal relevance of \ac{ECS}, its apparent increase in the \ac{CMIP}6
+ensemble is currently one of the most important questions for the entire
+climate modeling community. A list of possible reasons for it is given in
 \cref{sec:04:possible_reasons_high_ecs_cmip6}.
+
+In contrast to \ac{ECS}, \ac{TCR} has only been evaluated since \acs{AR}1 in
+1990, in which it is estimated with $2.3 \unit{K}$ using a single climate model
+\commentcite{Bretherton1990}{see single green circle in
+  \cref{fig:04:historical_ecs_tcr_ranges}}. Over the years, the modeled range
+of \ac{TCR} has decreased to about $\range{1.1}{2.6} \unit{K}$ in the
+\acs{CMIP}5 ensemble \autocite{Flato2013}. Similarly, the assessed range of
+\ac{TCR}, which was first quoted in \acs{AR}4 with $\range{1.0}{3.0} \unit{K}$
+\autocite{Solomon2007}, has been reduced to $\range{1.0}{2.5} \unit{K}$ in
+\acs{AR}5 \autocite{Stocker2013}. While the corresponding climate model ranges
+from \acs{CMIP}3 and \acs{CMIP}5 agree with these assessed ranges
+\autocite{Randall2007, Flato2013}, the upper bound of the \acs{CMIP}6 model
+range ($\range{1.3}{3.0} \unit{K}$) is well above the assessed upper bound of
+$2.5 \unit{K}$ given by \acs{AR}5. However, due to the correlation of \ac{TCR}
+and \ac{ECS} given by \cref{eq:02:tcr_vs_ecs}, this result is not surprising in
+light of the large \ac{ECS} range combined with the many high sensitivity
+models of the \acs{CMIP}6 ensemble.
 

### 2021-01-19 ch03_esmvaltool/esmvaltool.tex
Finished chapter 4

@@
 Finally, the position of the austral jet stream (\emph{uajet}) is given by the
-latitude with maximum zonal-mean eastward wind speed at $850 \unit{hPa}$ on the
+latitude with maximum zonal mean eastward wind speed at $850 \unit{hPa}$ on the
 Southern hemisphere between $80 \unit{\degree S}$ and $30 \unit{\degree S}$.

### 2021-01-19 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Finished chapter 4

@@
   \caption{Gregory regression for the \acs{CMIP}6 \acf{MMM} following
-    \textcite{Gregory2004}: Global and annual mean net \acf{TOA} radiation $N$
+    \textcite{Gregory2004}: Global and annual mean net \acl{TOA} radiation $N$
     versus the change in global and annual mean near-surface air temperature
@@
 exceeds this upper bound: out of the 42 analyzed \ac{CMIP}6 models, a third (14
-models) has an \ac{ECS} above $4.5 \unit{K}$. Because of the massive political
-and societal relevance of \ac{ECS}, its apparent increase in the \ac{CMIP}6
-ensemble is currently one of the most important questions for the entire
-climate modeling community. A list of possible reasons for it is given in
-\cref{sec:04:possible_reasons_high_ecs_cmip6}.
+models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
+several modeling centers independently confirmed high \ac{ECS} values in their
+models \autocite{Andrews2019, Gettelman2019, Wyser2020}. A list of possible
+reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble is
+given in \cref{sec:04:possible_reasons_high_ecs_cmip6}.
 
@@
 \label{sec:04:possible_reasons_high_ecs_cmip6}
+
+Because of the massive political and societal relevance of \ac{ECS}, its
+apparent increase in the \ac{CMIP}6 ensemble is currently one of the most
+important questions for the entire climate modeling community. For this reason,
+this section discusses possible reasons for the increased climate sensitivity
+in \acs{CMIP}6.
+
+Various improvements of the underlying physical, chemical and biological
+processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
+represent the coupled Earth system in more detail. Since many of these
+processes directly influence the models' \ac{ECS} \autocite{Forster2020},
+determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
+highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
+possible reason is the consideration of aerosol-cloud interactions. New
+prognostic aerosol schemes added to some \acs{CMIP}6 models that include
+aerosol-cloud interactions could have produced overly large negative radiative
+forcing, which then required a stronger model response to \ac{GHG} forcing in
+order to reproduce the observed historical temperature trend. In fact,
+\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
+forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
+present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
+have also been found for previous climate model generations
+\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
+over time, these relations cannot be used to tune \ac{ECS} based on the models'
+responses to aerosols over the different periods of the \nth{20} century
+\autocite{Dittus2020}.
+
+Another possible reason for the high climate sensitivity in the \acs{CMIP}6
+ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
+discussed in \cref{subsec:02:climate_feedbacks}, uncertainties in cloud
+feedbacks are a major source of uncertainty of \ac{ECS} in modern \acp{ESM}
+\autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
+to cloud feedbacks immediately impact the climate models' climate sensitivity.
+The particular feedback mechanism that is thought to be relevant for the high
+\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
+Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5). As
+the climate warms, the predominantly ice clouds over the Southern Ocean in
+these climate models become liquid clouds \autocite{McCoy2015}. As a result,
+the cloud will get more reflective to the incoming solar radiation since a
+cloud consisting of smaller liquid droplets reflects more sunlight than a cloud
+consisting of larger ice crystals (for a fixed water content). In addition,
+predominantly liquid clouds tend to precipitate less than mixed-phase clouds
+(composed of liquid water and ice), leading to more liquid water that stays in
+the cloud. Overall, this forms a strong negative shortwave cloud feedback over
+the Southern Ocean in climate models from previous \ac{CMIP} generations.
+
+However, it is well-known that the low-level mixed-phase clouds over the
+Southern Ocean in these earlier climate models are biased towards too large
+amounts of ice crystals and too little amounts of supercooled liquid water
+\autocite{BodasSalcedo2016, McCoy2015, McCoy2016, Tan2016}. For this reason,
+the microphysical representation of mixed-phase clouds has been substantially
+improved in many \acs{CMIP}6 models \autocite{BodasSalcedo2019, Gettelman2019,
+  Mulcahy2020, Williams2020}, leading for example to a better agreement of the
+simulated \ac{TOA} shortwave \ac{CRE} with corresponding observations
+\autocite{Bock2020}. As a consequence, the strong negative cloud feedback
+described above that results from a cloud phase change from ice clouds in the
+present-day to liquid clouds in the future is removed (or significantly
+reduced) in those \acs{CMIP}6 models that simulate predominantly liquid clouds
+over the present-day Southern Ocean. Moreover, this reduction of the negative
+shortwave cloud feedback increases the global net cloud feedback
+\autocite{BodasSalcedo2019, Tan2016} and eventually the \ac{ECS} of the
+affected \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019}.
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_netcre.pdf}
+    \caption{}
+    \label{fig:04:cloud_feedback_parameters:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_netcre-cmip5_netcre.pdf}
+    \caption{}
+    \label{fig:04:cloud_feedback_parameters:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_swcre.pdf}
+    \caption{}
+    \label{fig:04:cloud_feedback_parameters:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_swcre-cmip5_swcre.pdf}
+    \caption{}
+    \label{fig:04:cloud_feedback_parameters:d}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_lwcre.pdf}
+    \caption{}
+    \label{fig:04:cloud_feedback_parameters:e}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_lwcre-cmip5_lwcre.pdf}
+    \caption{}
+    \label{fig:04:cloud_feedback_parameters:f}
+  \end{subfigure}
+  \caption{Geographical distributions of the net (a), shortwave (c) and
+    longwave (e) \acf{CRE} feedback parameters for the \acs{CMIP}6 \acf{MMM}
+    and the corresponding differences to the \acs{CMIP}5 \acs{MMM}, again for
+    the net (b), shortwave (d) and longwave (f) components. The \acs{CRE}
+    feedback parameters for each grid cell are calculated with a Gregory
+    regression of the grid cell's \acs{CRE} versus the change in the
+    \acl{GSAT} (see \cref{subsec:02:cloud_feedback_parameters}). Updated from
+    \textcite{Bock2020}.}
+  \label{fig:04:cloud_feedback_parameters}
+\end{figure}
+
+To further confirm this hypothesis, we show the global distributions of
+relevant cloud feedback parameters (expressed by the net, shortwave and
+longwave \ac{CRE} feedback parameters; see
+\cref{subsec:02:cloud_feedback_parameters}) for the \acs{CMIP}6 \ac{MMM} and
+its difference to the \acs{CMIP}5 \ac{MMM} in
+\cref{fig:04:cloud_feedback_parameters}. First of all, the comparison of
+\cref{fig:04:cloud_feedback_parameters:a,fig:04:cloud_feedback_parameters:c,%
+  fig:04:cloud_feedback_parameters:e} demonstrates that the shortwave component
+dominates the net \ac{CRE} feedback parameter over large swaths of the globe
+in the \acs{CMIP}6 \ac{MMM}. The sign change at approximately $60
+\unit{\degree S}$ in the shortwave \ac{CRE} feedback parameter (see
+\cref{fig:04:cloud_feedback_parameters:b}) shows where the climate models are
+switching from simulating almost purely liquid clouds (north of $60
+\unit{\degree S}$) to clouds with an increasing ice component (south of $60
+\unit{\degree S}$) in their pre-industrial and present-day experiments. The
+increasing fraction of ice in the simulated mixed-phase clouds when moving
+south supports a negative shortwave cloud feedback on warming in these areas.
+\Cref{fig:04:cloud_feedback_parameters:b,fig:04:cloud_feedback_parameters:d}
+show a higher shortwave and net \ac{CRE} feedback parameter over the Southern
+Ocean in \acs{CMIP}6 compared to \acs{CMIP}5. This is consistent to the
+hypothesis presented in the paragraph above, which claims that the reduced
+negative cloud phase change feedback over the Southern Ocean in \acs{CMIP}6
+leads to an increased net cloud feedback parameter in comparison to older
+\ac{CMIP} generations. This result is further supported by
+\textcite{Zelinka2020}, who find a larger positive zonal mean net cloud
+feedback parameter in \acs{CMIP}6 compared to \acs{CMIP}5 due to a more
+positive (reduced negative) low‐level cloud feedback, primarily in the
+extratropics. Moreover, \textcite{Zelinka2020} show that this mainly results
+from an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds
+for the pre-industrial and present‐day periods, leading to the aforementioned
+reduction in the negative cloud phase change feedback on warming and
+eventually to higher \ac{ECS} values.

### 2021-01-20 ch02_scientific_background/scientific_background.tex
Added correct line breaks

@@
 linear emergent relationship for a climate model ensemble with $M$ climate
-models and data $\left\{ \left( x_m, y_m \right) \mid m \in I_M \right\}$ with index set $I_M = \left\{ 1, 2, \ldots, M\right\}$, a linear
+models and data $\left\{ \left( x_m, y_m \right) \mid m \in I_M \right\}$ with
+index set $I_M = \left\{ 1, 2, \ldots, M\right\}$, a linear
 regression model
@@
 
-In the second step of the algorithm, observed data of the process-based diagnostics $\bm{x}_0 =
-\left( 1, x_0^{(1)}, x_0^{(2)}, \ldots, x_0^{(K)} \right)^T \in \mathbb{R}^{(K
-  + 1)}$ is fed into the multivariate linear regression model to get an
-observation-based prediction of the target variable $\hat{y}_0$
+In the second step of the algorithm, observed data of the process-based
+diagnostics $\bm{x}_0 = \left( 1, x_0^{(1)}, x_0^{(2)}, \ldots, x_0^{(K)}
+\right)^T \in \mathbb{R}^{(K + 1)}$ is fed into the multivariate linear
+regression model to get an observation-based prediction of the target variable
+$\hat{y}_0$
 \autocite{Karpechko2013}:

### 2021-01-21 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added schematic illustrating the cloud phase change feedback

@@
     bars) over the years in the Charney report from 1979 \autocite{Charney1979}
-    and the subsequent \acp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
+    and the subsequent \acfp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
       Kattenberg1996, Albritton2001, Solomon2007, Stocker2013}. Orange and
@@
 some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
-used in \acs{AR}3 exhibit values of \acs{ECS} above $4.5 \unit{K}$, resulting
-in a upper model range of about $0.6 \unit{K}$ higher than the assessed range
-\autocite{Mitchell1990, Cubasch2001}. However, these deviations from the
-assessed range are insignificant to the ones present in the \acs{CMIP}6
-ensemble, which have become evident over the course of 2020. As shown in
-\cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper bounds are
-more extreme than in any previous climate model generation, resulting in a
-\acs{CMIP}6  \ac{ECS} model range of $\range{1.8}{5.6} \unit{K}$ (as of
+used in \acs{AR}3 exhibit values of \acs{ECS} well above $4.5 \unit{K}$,
+resulting in a upper model range of about $0.6 \unit{K}$ higher than the
+assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
+from the assessed range are insignificant to the ones present in the
+\acs{CMIP}6 ensemble, which have become evident over the course of 2020. As
+shown in \cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper
+bound are more extreme than in any previous climate model generation, resulting
+in a \acs{CMIP}6  \ac{ECS} model range of $\range{1.8}{5.6} \unit{K}$ (as of
 \TheMonth{}). This increase in range is particularly relevant for the upper
@@
 models \autocite{Andrews2019, Gettelman2019, Wyser2020}. A list of possible
-reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble is
-given in \cref{sec:04:possible_reasons_high_ecs_cmip6}.
+reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble
+including aerosol-cloud interactions and changes in the shortwave cloud
+feedback over the Southern Ocean is given in
+\cref{sec:04:possible_reasons_high_ecs_cmip6}.
 
 In contrast to \ac{ECS}, \ac{TCR} has only been evaluated since \acs{AR}1 in
-1990, in which it is estimated with $2.3 \unit{K}$ using a single climate model
-\commentcite{Bretherton1990}{see single green circle in
+1990, in which it is estimated with $2.3 \unit{K}$ using only a single climate
+model \commentcite{Bretherton1990}{see single green circle in
   \cref{fig:04:historical_ecs_tcr_ranges}}. Over the years, the modeled range
-of \ac{TCR} has decreased to about $\range{1.1}{2.6} \unit{K}$ in the
-\acs{CMIP}5 ensemble \autocite{Flato2013}. Similarly, the assessed range of
-\ac{TCR}, which was first quoted in \acs{AR}4 with $\range{1.0}{3.0} \unit{K}$
-\autocite{Solomon2007}, has been reduced to $\range{1.0}{2.5} \unit{K}$ in
-\acs{AR}5 \autocite{Stocker2013}. While the corresponding climate model ranges
-from \acs{CMIP}3 and \acs{CMIP}5 agree with these assessed ranges
-\autocite{Randall2007, Flato2013}, the upper bound of the \acs{CMIP}6 model
-range ($\range{1.3}{3.0} \unit{K}$) is well above the assessed upper bound of
-$2.5 \unit{K}$ given by \acs{AR}5. However, due to the correlation of \ac{TCR}
-and \ac{ECS} given by \cref{eq:02:tcr_vs_ecs}, this result is not surprising in
-light of the large \ac{ECS} range combined with the many high sensitivity
-models of the \acs{CMIP}6 ensemble.
+of \ac{TCR} has decreased from about $\range{1.3}{3.8} \unit{K}$ in the
+\acs{CMIP}1 models \autocite{Kattenberg1996} to about $\range{1.1}{2.6}
+\unit{K}$ in the \acs{CMIP}5 ensemble \autocite{Flato2013}. Similarly, the
+assessed range of \ac{TCR}, which was first quoted in \acs{AR}4 with
+$\range{1.0}{3.0} \unit{K}$ \autocite{Solomon2007}, has been reduced to
+$\range{1.0}{2.5} \unit{K}$ in the subsequent \acs{AR}5 \autocite{Stocker2013}.
+While the corresponding climate model ranges from \acs{CMIP}3 and \acs{CMIP}5
+agree with these assessed ranges \autocite{Randall2007, Flato2013}, the upper
+bound of the \acs{CMIP}6 model range ($\range{1.3}{3.0} \unit{K}$) is well
+above the assessed upper bound of $2.5 \unit{K}$ given by \acs{AR}5. However,
+unlike \ac{ECS}, also the lower bound of the \acs{CMIP}6 model range increased
+compared to the \acs{CMIP}5 ensemble, leading to a similar \ac{TCR} model
+spread in both climate model generations. Due to the correlation of \ac{TCR}
+and \ac{ECS} given by \cref{eq:02:tcr_vs_ecs}, the increase of \ac{TCR} in many
+\acs{CMIP}6 models is not surprising in light of the many high \ac{ECS} models
+that are present the \acs{CMIP}6 ensemble.
 
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch04_papers_ecs_tcr_assessment/figs/cloud_phase_change_feedback.pdf}
+  \caption{Schematic illustration of the strong negative shortwave cloud phase
+    change feedback over the Southern Ocean, which is present in earlier
+    \acs{CMIP} model generations like \acs{CMIP}5 (green arrows). Due to
+    substantial improvements in the microphysical representation of mixed-phase
+    clouds in many \acs{CMIP}6 models, the predominantly ice clouds over the
+    Southern Ocean in the present-day climate of previous climate model
+    generations have been replaced with clouds that predominantly consist of
+    (supercooled) liquid clouds \autocite{BodasSalcedo2019, Gettelman2019,
+      Mulcahy2020, Williams2020}. Thus, in these models a cloud phase change
+    due to global warming is no longer possible, which removes (or
+    significantly reduces) the strong negative cloud phase change feedback
+    (orange arrows). This leads to an increase in the global net cloud
+    feedback \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS} in the
+    corresponding \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019},
+    which is a possible explanation for the high \acs{ECS} values found in
+    many \acs{CMIP}6 models.}
+  \label{fig:04:cloud_phase_change_feedback}
+\end{figure}
+
 Another possible reason for the high climate sensitivity in the \acs{CMIP}6
@@
 \ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
-Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5). As
-the climate warms, the predominantly ice clouds over the Southern Ocean in
-these climate models become liquid clouds \autocite{McCoy2015}. As a result,
-the cloud will get more reflective to the incoming solar radiation since a
-cloud consisting of smaller liquid droplets reflects more sunlight than a cloud
-consisting of larger ice crystals (for a fixed water content). In addition,
-predominantly liquid clouds tend to precipitate less than mixed-phase clouds
-(composed of liquid water and ice), leading to more liquid water that stays in
-the cloud. Overall, this forms a strong negative shortwave cloud feedback over
-the Southern Ocean in climate models from previous \ac{CMIP} generations.
+Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5).
+This so-called \emph{cloud phase change feedback} is illustrated in
+\cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate warms,
+the predominantly ice clouds over the Southern Ocean in these climate models
+become liquid clouds \autocite{McCoy2015}. As a result, the cloud will get more
+reflective to the incoming solar radiation since a cloud consisting of smaller
+liquid droplets reflects more sunlight than a cloud consisting of larger ice
+crystals (for a fixed water content), which has a cooling effect on the climate
+system. In addition, predominantly liquid clouds tend to precipitate less than
+mixed-phase clouds (composed of liquid water and ice). This leads to a higher
+liquid water content of the cloud and a further amplification of the cooling
+effect. Overall, this forms a strong negative shortwave cloud feedback over the
+Southern Ocean in climate models from previous \ac{CMIP} generations.
 
@@
 reduced) in those \acs{CMIP}6 models that simulate predominantly liquid clouds
-over the present-day Southern Ocean. Moreover, this reduction of the negative
-shortwave cloud feedback increases the global net cloud feedback
-\autocite{BodasSalcedo2019, Tan2016} and eventually the \ac{ECS} of the
-affected \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019}.
+over the present-day Southern Ocean. This is illustrated in
+\cref{fig:04:cloud_phase_change_feedback} (orange arrows). Moreover, this
+reduction of the negative shortwave cloud phase change feedback increases the
+global net cloud feedback \autocite{BodasSalcedo2019, Tan2016} and eventually
+the \ac{ECS} of the affected \acs{CMIP}6 models \autocite{Andrews2019,
+  Gettelman2019}, offering a possible explanation for the high \ac{ECS} values
+in some \acs{CMIP}6 models.
 
-\begin{figure}[p]
+\begin{figure}[!t]
   \centering
@@
 positive (reduced negative) low‐level cloud feedback, primarily in the
-extratropics. Moreover, \textcite{Zelinka2020} show that this mainly results
-from an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds
-for the pre-industrial and present‐day periods, leading to the aforementioned
+extratropics. In comparison to earlier \acs{CMIP} generations, the
+\acs{CMIP}6 models exhibit weaker increases in the extratropical low‐level
+cloud cover and the corresponding liquid water content as a result of global
+warming. Moreover, \textcite{Zelinka2020} show that this mainly results from
+an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds for
+the pre-industrial and present‐day periods, leading to the aforementioned
 reduction in the negative cloud phase change feedback on warming and

### 2021-01-21 ch02_scientific_background/scientific_background.tex
Added additional section to chapter 04

@@
 concepts of climate model simulations and associated uncertainties are
-introduced. Next, important metrics describing climate sensitivity and
-fundamental biogeochemical processes of the global carbon cycle are presented.
+introduced (\cref{sec:02:esms}). Next, important metrics describing climate
+sensitivity (\cref{sec:02:climate_sensitivity}) and fundamental biogeochemical
+processes of the global carbon cycle (\cref{sec:02:esms}) are presented.
 Finally, state-of-the art techniques that can be used to reduce uncertainties
-in projections of the future climate with observations are shown. These methods
-form the basis for new techniques developed in this thesis.
+in projections of the future climate with observations are shown
+(\cref{sec:02:reducing_uncertainties}). These methods form the basis for new
+techniques developed in this thesis.
 

### 2021-01-21 ch03_esmvaltool/esmvaltool.tex
Added additional section to chapter 04

@@
 four peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020,
-  Weigel2020}. After a brief overview over the \ac{ESMValTool}'s structure,
-this chapter presents all these contributions that are not documented in other
-chapters of this thesis.
+  Weigel2020}. After a brief overview of the \ac{ESMValTool}'s structure
+(\cref{sec:03:esmvaltool}), this chapter presents all these contributions that
+are not documented in other chapters of this thesis
+(\cref{sec:03:contributions_to_esmvalcore,sec:03:contributions_to_esmvaltool}).
 

### 2021-01-21 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added additional section to chapter 04

@@
 on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we use it to assess
-and analyze climate sensitivity (expressed by \ac{ECS} and \ac{TCR}) of the
-\acs{CMIP}6 models and compare it to corresponding results of predecessor model
-generations. This work is already published in two scientific papers:
+and analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to
+corresponding results of predecessor model generations and international
+climate assessments. This work is already published in two scientific papers:
 \textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
@@
 \acs{CMIP}6 models. For \textcite{Meehl2020}, the author of this thesis
-contributed all figures and tables of the study and details on the calculations
-of \ac{ECS} and \ac{TCR}.
+contributed all figures and tables of the study. This chapter first presents
+the evaluation of climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6
+ensembles (\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding
+\ac{ECS} and \ac{TCR} values into historical context
+(\cref{sec:04:historical_context}). Finally, possible reasons for the apparent
+increase of climate sensitivity in the \acs{CMIP}6 models are briefly discussed
+(\cref{sec:04:possible_reasons_high_ecs_cmip6}).
 
-\section{Evaluation of \acs{ECS} and \acs{TCR} in \acs{CMIP}6}
+\section{Evaluation of \acs{ECS} and \acs{TCR} in \acs{CMIP}5 and \acs{CMIP}6}
 \label{sec:04:evaluation_ecs_and_tcr}
@@
 (higher) values of \ac{ECS} when only early (late) years of the simulation are
-considered.
+considered. In contrast to that, the transient response of the climate system
+\ac{TCR} is calculated from \ac{GSAT} data at the time of \ac{CO2} doubling in
+a \onepctcotwo{} run \commentcite{Bindoff2013}{see \cref{subsec:02:tcr}}. Using
+these two definitions and the \ac{ESMValTool}, we assess \ac{ECS} and \ac{TCR}
+for all \acs{CMIP}5 and \acs{CMIP}6 models where the necessary temperature and
+radiation data is available (as of \TheMonth{}).
+\Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an overview over the
+results for the \acs{CMIP}5 and \acs{CMIP}6 models, respectively.
 
@@
 \end{figure}
-
-In contrast to that, the transient response of the climate system \ac{TCR} is
-calculated from \ac{GSAT} data at the time of \ac{CO2} doubling in a
-\onepctcotwo{} run \commentcite{Bindoff2013}{see \cref{subsec:02:tcr}}. Using
-these two definitions and the \ac{ESMValTool}, we assess \ac{ECS} and \ac{TCR}
-for all \acs{CMIP}5 and \acs{CMIP}6 models where the necessary temperature and
-radiation data is available (as of \TheMonth{}).
-\Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an overview over the
-results for the \acs{CMIP}5 and \acs{CMIP}6 models, respectively.
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
-    ch04_papers_ecs_tcr_assessment/figs/historical_ecs_tcr_ranges.pdf}
-  \caption{Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red
-    bars) over the years in the Charney report from 1979 \autocite{Charney1979}
-    and the subsequent \acfp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
-      Kattenberg1996, Albritton2001, Solomon2007, Stocker2013}. Orange and
-    green bars represent the modeled ranges of \acs{ECS} and \acs{TCR},
-    respectively, from the different climate model ensembles of the \acf{CMIP}
-    with the corresponding \acf{MMM} illustrated with the horizontal black
-    lines (the green dot represents output from a single climate model). The
-    numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
-    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
-    \textcite{Meehl2020}.}
-  \label{fig:04:historical_ecs_tcr_ranges}
-\end{figure}
+\section{Comparison to Previous \acs{CMIP} Generations and International
+  Climate Assessments}
+\label{sec:04:historical_context}
 
-To illustrate this and put it into historical context,
-\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
-\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
-and the different \acp{AR} of the \ac{IPCC} in combination with the
-corresponding modeled ranges from the different \ac{CMIP} generations. Since
-the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
-\unit{K}$ has almost remained unchanged during the last 40 years
-\autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
+To illustrate the results of the previous section and put them into historical
+context, \cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS}
+and \ac{TCR} ranges over the years from the Charney report
+\autocite{Charney1979} and the different \acp{AR} of the \ac{IPCC} in
+combination with the corresponding modeled ranges from the different \ac{CMIP}
+generations. Since the Charney report in 1979, the assessed range of \ac{ECS}
+of $\range{1.5}{4.5} \unit{K}$ has almost remained unchanged during the last 40
+years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
   Stocker2013} with the exception of \acs{AR}4, where the lower bound was
@@
 \cref{sec:04:possible_reasons_high_ecs_cmip6}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch04_papers_ecs_tcr_assessment/figs/historical_ecs_tcr_ranges.pdf}
+  \caption{Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red
+    bars) over the years in the Charney report from 1979 \autocite{Charney1979}
+    and the subsequent \acfp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
+      Kattenberg1996, Albritton2001, Solomon2007, Stocker2013}. Orange and
+    green bars represent the modeled ranges of \acs{ECS} and \acs{TCR},
+    respectively, from the different climate model ensembles of the \acf{CMIP}
+    with the corresponding \acf{MMM} illustrated with the horizontal black
+    lines (the green dot represents output from a single climate model). The
+    numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
+    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
+    \textcite{Meehl2020}.}
+  \label{fig:04:historical_ecs_tcr_ranges}
+\end{figure}
 
@@
 
+Another possible reason for the high climate sensitivity in the \acs{CMIP}6
+ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
+discussed in \cref{subsec:02:climate_feedbacks}, uncertainties in cloud
+feedbacks are a major source of uncertainty of \ac{ECS} in modern \acp{ESM}
+\autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
+to cloud feedbacks immediately impact the climate models' climate sensitivity.
+The particular feedback mechanism that is thought to be relevant for the high
+\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
+Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5).
+This so-called \emph{cloud phase change feedback} is illustrated in
+\cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate warms,
+the predominantly ice clouds over the Southern Ocean in these climate models
+become liquid clouds \autocite{McCoy2015}. As a result, the cloud will get more
+reflective to the incoming solar radiation since a cloud consisting of smaller
+liquid droplets reflects more sunlight than a cloud consisting of larger ice
+crystals (for a fixed water content), which has a cooling effect on the climate
+system. In addition, predominantly liquid clouds tend to precipitate less than
+mixed-phase clouds (composed of liquid water and ice). This leads to a higher
+liquid water content of the cloud and a further amplification of the cooling
+effect. Overall, this forms a strong negative shortwave cloud feedback over the
+Southern Ocean in climate models from previous \ac{CMIP} generations.
+
 \begin{figure}[t]
@@
 
-Another possible reason for the high climate sensitivity in the \acs{CMIP}6
-ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
-discussed in \cref{subsec:02:climate_feedbacks}, uncertainties in cloud
-feedbacks are a major source of uncertainty of \ac{ECS} in modern \acp{ESM}
-\autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
-to cloud feedbacks immediately impact the climate models' climate sensitivity.
-The particular feedback mechanism that is thought to be relevant for the high
-\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
-Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5).
-This so-called \emph{cloud phase change feedback} is illustrated in
-\cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate warms,
-the predominantly ice clouds over the Southern Ocean in these climate models
-become liquid clouds \autocite{McCoy2015}. As a result, the cloud will get more
-reflective to the incoming solar radiation since a cloud consisting of smaller
-liquid droplets reflects more sunlight than a cloud consisting of larger ice
-crystals (for a fixed water content), which has a cooling effect on the climate
-system. In addition, predominantly liquid clouds tend to precipitate less than
-mixed-phase clouds (composed of liquid water and ice). This leads to a higher
-liquid water content of the cloud and a further amplification of the cooling
-effect. Overall, this forms a strong negative shortwave cloud feedback over the
-Southern Ocean in climate models from previous \ac{CMIP} generations.
-
 However, it is well-known that the low-level mixed-phase clouds over the
@@
 
-\begin{figure}[!t]
+\begin{figure}[p]
   \centering

### 2021-01-21 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added introduction to chapter 05

@@
 (\cref{sec:04:historical_context}). Finally, possible reasons for the apparent
-increase of climate sensitivity in the \acs{CMIP}6 models are briefly discussed
-(\cref{sec:04:possible_reasons_high_ecs_cmip6}).
+increase of climate sensitivity in the \acs{CMIP}6 models are discussed in
+detail (\cref{sec:04:possible_reasons_high_ecs_cmip6}).
 

### 2021-01-21 ch05_paper_ecs/paper_ecs.tex
Added correct line breaks
Added additional section to chapter 04
Added introduction to chapter 05

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                                  Paper ECS                                  %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Evaluation of Emergent Constraints on the \acl{ECS} in \acs{CMIP}6}
+\label{ch:05:paper_ecs}
 
-The \ac{ECS} is really cool. I like it very much!
+The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
+previous chapter begs the question about the interpretation and credibility of
+these results. Since the \acs{CMIP}6 models provide crucial input for the
+upcoming international climate assessment of the \ac{IPCC} (\acs{AR}6), this is
+not only important for the scientific community, but also directly affects
+society and policymakers. An important question in this context could be the
+following: Is it reasonable to adapt the assessed \ac{ECS} range from \acs{AR}5
+($\range{1.5}{4.5} \unit{K}$) in \acs{AR}6 based on the \acs{CMIP}6 model
+range? An argument for this is that due to the many new and improved processes
+included to the models, the \acs{CMIP}6 ensemble is arguably more realistic
+than its predecessor generations. However, even the \acs{CMIP}6 models are far
+from fully simulating all aspects and processes of the vastly complex Earth
+system. For this reason, the apparent increase in climate sensitivity in the
+\acs{CMIP}6 models might be wrongly caused by missing processes that cancel out
+the effects of the newly included processes, leading to a biased \ac{ECS} even
+though the models are more realistic. Thus, it is vital to take other sources
+of information into account before answering the question raised above.
+
+One of the most important sources of information for this purpose is
+undoubtedly observational data. As presented in
+\cref{subsec:02:emergent_constraints}, emergent constraints are a beneficial
+tool to assess the accuracy of climate model projections and reduce associated
+uncertainties with appropriate observations of the Earth system. For \ac{ECS},
+a large variety of different emergent constraints has been developed over the
+years, which we now evaluate on the new \acs{CMIP}6 ensemble. In total we
+assess eleven emergent constraints (selected by their availability in the
+\ac{ESMValTool}), out of which the most are related to cloud feedbacks, since
+these constitute the most important source of uncertainty for \ac{ECS}
+\commentcite{Boucher2013, Flato2013}{see also
+  \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
+constraints have been derived on the \acs{CMIP}3 or \ac{CMIP}5 ensemble, the
+\ac{CMIP}6 models offer an exciting possibility to perform out-of-sample
+testing using the previously unseen data. We quantify the skill of the emergent
+constraints with three metrics: the coefficient of determination $R^2$ of the
+linear relationship, the statistical significance using a two-sided $t$-test
+based on the Pearson correlation coefficient $r$ (see \cref{eq:02:t}) and the
+constrained range of \ac{ECS} based on the constrained \ac{PDF} (see
+\cref{eq:02:pdf_y_given_x0}). This analysis has already been published in
+\textcite{Schlund2020a}. In this chapter, we first introduce the eleven
+emergent constraints on \acs{ECS} and evaluate them on \acs{CMIP}5 and
+\acs{CMIP}6 data (\cref{sec:05:comparison_of_emergent_constraints}). After
+that, we provide a discussion of the results and draw important conclusions
+from them (\cref{sec:05:discussion_and_conclusions}).
+
+\section{Comparison of Emergent Constraints on \acs{ECS} for \acs{CMIP}5 and
+  \acs{CMIP}6}
+\label{sec:05:comparison_of_emergent_constraints}
+
+\subsection{COX}
+\label{subsec:05:cox}
+
+\section{Discussion and Conclusions}
+\label{sec:05:discussion_and_conclusions}
+
+- mention limitation of emergent constraints of chapter 2 again
+
+- mention why significance
+
+- it is not possible to constrain ECS with the emergent constraints

### 2021-01-22 ch03_esmvaltool/esmvaltool.tex
Updated several figures

@@
   \centering
-  \begin{subfigure}[t]{\SubfigureWidth{}}
-    \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/co2s_map.pdf}
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \raisebox{2.5mm}{\includegraphics[width=\columnwidth]{
+      ch03_esmvaltool/figs/co2s_map.pdf}}
     \caption{}
@@
   ~
-  \begin{subfigure}[t]{\SubfigureWidth{}}
+  \begin{subfigure}[b]{\SubfigureWidth{}}
     \includegraphics[width=\columnwidth]{
@@
   \centering
-  \begin{subfigure}[t]{\SubfigureWidth{}}
+  \begin{subfigure}[b]{\SubfigureWidth{}}
     \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_psi.pdf}
@@
   ~
-  \begin{subfigure}[t]{\SubfigureWidth{}}
+  \begin{subfigure}[b]{\SubfigureWidth{}}
     \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/cox18_ec.pdf}
@@
   \centering
-  \begin{subfigure}[t]{0.34\columnwidth}
+  \begin{subfigure}[b]{0.35\columnwidth}
     \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942a.pdf}
@@
   ~
-  \begin{subfigure}[t]{0.34\columnwidth}
+  \begin{subfigure}[b]{0.35\columnwidth}
     \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942b.pdf}
@@
   ~
-  \begin{subfigure}[t]{0.24\columnwidth}
-    \raisebox{7.5mm}{\includegraphics[width=\columnwidth]{
+  \begin{subfigure}[b]{0.25\columnwidth}
+    \raisebox{12.5mm}{\includegraphics[width=\columnwidth]{
       ch03_esmvaltool/figs/flato13_942_legend.pdf}}

### 2021-01-22 ch05_paper_ecs/paper_ecs.tex
Added figures for chapter 05

@@
 
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_BRI.pdf}
+    \caption{}
+    \label{fig:05:bri_cox_lip:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_BRI.pdf}
+    \caption{}
+    \label{fig:05:bri_cox_lip:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_COX.pdf}
+    \caption{}
+    \label{fig:05:bri_cox_lip:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+  ch05_paper_ecs/figs/target_distribution_training_data_COX.pdf}
+    \caption{}
+    \label{fig:05:bri_cox_lip:d}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_LIP.pdf}
+    \caption{}
+    \label{fig:05:bri_cox_lip:e}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+  ch05_paper_ecs/figs/target_distribution_training_data_LIP.pdf}
+    \caption{}
+    \label{fig:05:bri_cox_lip:f}
+  \end{subfigure}
+  \caption{xD. Updated from
+    \textcite{Schlund2020a}.}
+  \label{fig:05:bri_cox_lip}
+\end{figure}
+
 \section{Discussion and Conclusions}

### 2021-01-27 ch02_scientific_background/scientific_background.tex
Updated GCP 2019 paper to GCP 2020 paper

@@
 only about $50 \unit{\%}$ of the emitted \ac{CO2} by humankind remains in the
-atmosphere \autocite{Friedlingstein2019}. The residual part is absorbed by the
+atmosphere \autocite{Friedlingstein2020}. The residual part is absorbed by the
 two other main carbon sinks of the planet, the terrestrial biosphere and the
@@
 component with a current (2019) concentration of about $410 \unit{ppm}$
-\autocite{Friedlingstein2019}. Additional contributors to the atmospheric
+\autocite{Friedlingstein2020}. Additional contributors to the atmospheric
 carbon content are the trace gas \ac{CH4}, the trace gas \ac{CO}, hydrocarbons,
@@
 biota (about $3 \unit{GtC}$, predominantly phytoplankton and other
-microorganisms) \autocite{Ciais2013, Friedlingstein2019}.
+microorganisms) \autocite{Ciais2013, Friedlingstein2020}.
 
@@
 carbon cycle is not in a steady state anymore. In the decade
-\range{2009}{2018}, anthropogenic activities caused net carbon fluxes of $3.2
+\range{2010}{2019}, anthropogenic activities caused net carbon fluxes of $3.4
 \unit{\GtCyr{}}$ from the atmosphere into the terrestrial biosphere due to
@@
 into the ocean due to increase dissolution of \ac{CO2} into the sea
-\autocite{Friedlingstein2019}. In the same time, the amount of carbon in the
-atmosphere reservoir increased with a rate of $4.9 \unit{\GtCyr{}}$, indicating
+\autocite{Friedlingstein2020}. In the same time, the amount of carbon in the
+atmosphere reservoir increased with a rate of $5.1 \unit{\GtCyr{}}$, indicating
 that only about half of the anthropogenic \ac{CO2} emissions in the last decade
-remained in the atmosphere \autocite{Friedlingstein2019} where they can act as
+remained in the atmosphere \autocite{Friedlingstein2020} where they can act as
 \ac{GHG}.
@@
     carbon cycle caused by anthropogenic activities, averaged globally for the
-    decade \range{2009}{2018}. Arrows represent carbon exchange fluxes; circles
+    decade \range{2010}{2019}. Arrows represent carbon exchange fluxes; circles
     carbon reservoirs. More details are given in the legend of this figure.
-    Taken from \textcite{Friedlingstein2019}.}
+    Taken from \textcite{Friedlingstein2020}.}
   \label{fig:02:carbon_cycle_perturbation}

### 2021-01-27 ch03_esmvaltool/esmvaltool.tex
Updated GCP 2019 paper to GCP 2020 paper

@@
     & CT2019 & \autocite{Jacobson2020} \\
-    & GCP & \autocite{Friedlingstein2019} \\
+    & GCP & \autocite{Friedlingstein2020} \\
     & HWSD & \autocite{Wieder2014} \\

### 2021-01-27 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Updated several figures
Adapted copyright statements

@@
     smaller \acs{ECS} than using only the last 130 years of the simulation
-    (orange circles and line). Updated from \textcite{Meehl2020}.}
+    (orange circles and line). \AdaptedFrom{Meehl2020}.}
   \label{fig:04:gregory_regression_cmip6_mmm}
@@
     over all climate models (using the normalization $1 / M$, where $M$ is the
-    number of models).}
+    number of models). \AdaptedFrom{Meehl2020}.}
   \label{tab:04:ecs_tcr_cmip5}
@@
     \dataset & \idx & \ecs & \tcr}
-  \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.}
+  \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.
+    \AdaptedFrom{Meehl2020}.}
   \label{tab:04:ecs_tcr_cmip6}
@@
     numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
-    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
-    \textcite{Meehl2020}.}
+    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}).
+    \AdaptedFrom{Meehl2020}.}
   \label{fig:04:historical_ecs_tcr_ranges}
@@
   \centering
-  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+  \begin{subfigure}[b]{\SubfigureWidth{}}
     \includegraphics[width=\columnwidth]{
@@
   ~
-  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+  \begin{subfigure}[b]{\SubfigureWidth{}}
     \includegraphics[width=\columnwidth]{
@@
   \\
-  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
+  \begin{subfigure}[b]{\SubfigureWidth{}}
     \includegraphics[width=\columnwidth]{
@@
     regression of the grid cell's \acs{CRE} versus the change in the
-    \acl{GSAT} (see \cref{subsec:02:cloud_feedback_parameters}). Updated from
-    \textcite{Bock2020}.}
+    \acl{GSAT} (see \cref{subsec:02:cloud_feedback_parameters}).
+    \AdaptedFrom{Bock2020}.}
   \label{fig:04:cloud_feedback_parameters}

### 2021-01-28 ch01_introduction/introduction.tex
Added correct line breaks
Upadted author contributions

@@
 \begin{enumerate}
-  \item Did climate sensitivity change in the latest generation of \acp{ESM} from \acs{CMIP}6 and can we see progress compared to older climate
-  model generations?
+  \item Did climate sensitivity change in the latest generation of \acp{ESM}
+  from \acs{CMIP}6 and can we see progress compared to older climate model
+  generations?
 
@@
 Parts of this thesis are already published in multiple peer-reviewed
-publications (two first-author studies and six co-author studies). Unless
-stated otherwise, results, discussions and conclusions presented from these
-studies originate from the author of this thesis. A detailed list of
-corresponding references is given at the beginning of the relevant chapters.
+publications (two first-author studies and six co-author studies). Wherever
+material from these studies is presented in this thesis, the pronoun
+\enquote{we} is used to increase readability by avoiding the passive voice and
+to acknowledge all involved contributors. However, unless stated otherwise, all
+contents from these publications (text, figures and tables) shown in this
+thesis originate from the author of this thesis. A detailed list of
+contributions to these studies (including their full reference) is given in the
+corresponding chapters.
 

### 2021-01-28 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Upadted author contributions

@@
 author of this thesis contributed their figure 8 (a bar chart showing the
-\ac{ECS} for the three climate model generations \acs{CMIP}3, \acs{CMIP}5 and
-\acs{CMIP}6), their figure 10 (map plots showing cloud-related feedback
-parameters for the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
-\cref{fig:04:cloud_feedback_parameters}) and discussion about \ac{ECS} of the
-\acs{CMIP}6 models. For \textcite{Meehl2020}, the author of this thesis
-contributed all figures and tables of the study. This chapter first presents
-the evaluation of climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6
-ensembles (\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding
-\ac{ECS} and \ac{TCR} values into historical context
-(\cref{sec:04:historical_context}). Finally, possible reasons for the apparent
-increase of climate sensitivity in the \acs{CMIP}6 models are discussed in
-detail (\cref{sec:04:possible_reasons_high_ecs_cmip6}).
+\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5 and \acs{CMIP}6 models), their
+figure 10 (map plots showing cloud-related feedback parameters for the
+\acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
+\cref{fig:04:cloud_feedback_parameters}), code to create these figures with the
+\ac{ESMValTool} and text for the manuscript (in particular a section about the
+evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author of this thesis
+contributed all figures and tables, code to produce these with the
+\ac{ESMValTool} and text for the manuscript. This chapter first presents the
+evaluation of climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6 ensembles
+(\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding \ac{ECS} and
+\ac{TCR} values into historical context (\cref{sec:04:historical_context}).
+Finally, possible reasons for the apparent increase of climate sensitivity in
+the \acs{CMIP}6 models are discussed in detail
+(\cref{sec:04:possible_reasons_high_ecs_cmip6}).
 

### 2021-01-28 ch05_paper_ecs/paper_ecs.tex
Adapted copyright statements
Upadted author contributions

@@
 \cref{eq:02:pdf_y_given_x0}). This analysis has already been published in
-\textcite{Schlund2020a}. In this chapter, we first introduce the eleven
-emergent constraints on \acs{ECS} and evaluate them on \acs{CMIP}5 and
-\acs{CMIP}6 data (\cref{sec:05:comparison_of_emergent_constraints}). After
-that, we provide a discussion of the results and draw important conclusions
-from them (\cref{sec:05:discussion_and_conclusions}).
+\textcite{Schlund2020a}. For this paper, the author of this thesis lead the
+writing and the analysis and contributed the code to reproduce this analysis
+including all figures and tables with the \ac{ESMValTool}. In this chapter, we
+first introduce the eleven emergent constraints on \acs{ECS} and evaluate them
+on \acs{CMIP}5 and \acs{CMIP}6 data
+(\cref{sec:05:comparison_of_emergent_constraints}). After that, we provide a
+discussion of the results and draw important conclusions from them
+(\cref{sec:05:discussion_and_conclusions}).
 
@@
   \end{subfigure}
-  \caption{xD. Updated from
-    \textcite{Schlund2020a}.}
+  \caption{xD. \AdaptedFrom{Schlund2020a}.}
   \label{fig:05:bri_cox_lip}

### 2021-01-28 ch03_esmvaltool/esmvaltool.tex
Adapted copyright statements
Added recipe for Meehl et al. (2020) to chapter 03

@@
     and diagnostic scripts in the recipe, which is a configuration file that
-    controls the main workflow of the \acs{ESMValTool}. Taken and modified from
-    \textcite{Righi2020}.}
+    controls the main workflow of the \acs{ESMValTool}.
+    \AdaptedFrom{Righi2020}.}
   \label{fig:03:esmvaltool_schematic}
@@
     & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
+    & \emph{meehl20sciadv} & \autocite{Meehl2020} \\
     & \emph{schlund20esd} & \autocite{Schlund2020a} \\
@@
 \emph{recipe\_bock20jgr.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
+\emph{recipe\_meehl20sciadv.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
 \emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and

### 2021-02-01 ch01_introduction/introduction.tex
Adapted key science questions

@@
 
-\section{Research Questions}
+\section{Key Science Questions}
 \label{sec:01:research_questions}
 
-This section will present the \textbf{AWESOME} research questions!!!
+This section will present the \textbf{AWESOME} key science questions!!!
 
 \begin{enumerate}
-  \item Did climate sensitivity change in the latest generation of \acp{ESM}
-  from \acs{CMIP}6 and can we see progress compared to older climate model
-  generations?
+  \item What is the range of climate sensitivity in the latest generation of
+  \acp{ESM} compared to previous multi-model ensembles, and do we understand
+  the processes that determine this uncertainty range?
 
-  \item Can we use emergent constraints established in \acs{CMIP}3 and
-  \acs{CMIP}5 to reduce uncertainties in climate sensitivity in the \acs{CMIP}6
-  ensemble?
+  \item Can uncertainties in climate sensitivity be reduced with observations
+  using the emergent constraint approach?
 
-  \item Can we use \ac{ML}-based approaches to constrain gridded projections of
-  the future climate with observations?
+  \item Can uncertainties in multi-dimensional (gridded) climate projections be
+  reduced with \ac{ML} techniques and observations?
 \end{enumerate}

### 2021-02-04 ch05_paper_ecs/paper_ecs.tex
Added further tables for chapter 05 and text for chapter 06

@@
 including all figures and tables with the \ac{ESMValTool}. In this chapter, we
-first introduce the eleven emergent constraints on \acs{ECS} and evaluate them
-on \acs{CMIP}5 and \acs{CMIP}6 data
-(\cref{sec:05:comparison_of_emergent_constraints}). After that, we provide a
-discussion of the results and draw important conclusions from them
-(\cref{sec:05:discussion_and_conclusions}).
+first introduce the methods and data (climate model output and observations)
+used in this study (\cref{sec:05:methods_and_data}). After that, we present the
+eleven emergent constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5
+and \acs{CMIP}6 ensemble (\cref{sec:05:comparison_of_emergent_constraints}).
+Finally, we provide a discussion (\cref{sec:05:discussion}) and a summary of
+the results (\cref{sec:05:summary}).
+
+\section{Methods and Data}
+\label{sec:05:methods_and_data}
+
+In this study we use the output from climate models participating in
+\acs{CMIP}5 and \acs{CMIP}6, shown in \cref{sec:app:cmip_models_in_chapter_5}
+in \cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models}, respectively.
 
@@
 
-\section{Discussion and Conclusions}
-\label{sec:05:discussion_and_conclusions}
+\section{Discussion}
+\label{sec:05:discussion}
 
-- mention limitation of emergent constraints of chapter 2 again
-
-- mention why significance
-
-- it is not possible to constrain ECS with the emergent constraints
+\section{Summary}
+\label{sec:05:summary}

### 2021-02-04 ch06_paper_gpp/paper_gpp.tex
Added further tables for chapter 05 and text for chapter 06

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                                  Paper GPP                                  %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Constraining Future \acl{GPP} with Machine Learning}
+\label{ch:06:paper_gpp}
 
-The \ac{ECS} is really cool. I like it very much!
+- Single-process emergent constraints have disadvantages
+
+To overcome these, we explore an alternative approach that expands the MDER
+method with an ML algorithm. In contrast to emergent constraints, this approach
+considers (1) multiple process-based predictors, (2) multi-dimensional
+(gridded) target variables and (3) non-linear relationships between the
+predictors and the target variable. The target variable analyzed is GPP. As
+described in detail in section XXX, the carbon cycle slows down global warming
+by absorbing 30 \% CO2, thus accurately quantifying the carbon sink is crucial.
+GPP is most largest flux of the terrestrial carbon cycle. Due to feedbacks
+(section XXX, NAME OF FEEDBACKS), multi-model projectionf of future GPP is
+uncertain, especially in CMIP5 RCP 8.5 at the end of the century. Thus, the aim
+of the works presented in this chapter is to reduce uncertainties in
+multi-model projections of future GPP in the RCP 8.5 scenario at the end of the
+21st century with a new ML-based weighting technique and observations. The new
+two‐step approach utilizes aspects of emergent constraints and the MDER
+technique in combination with a supervised machine learning algorithm. In the
+first step, we apply an existing emergent... (see SCHLUND ET AL., 2020).
+
+\section{Step 1: XXXXXXXXXXXXXXXXXX}
+\label{sec:06:step_1}
+
+TBA.

### 2021-02-04 end_appendix/appendix.tex
Wrote first page of introduction
Added first figure
Finished section on carbon cycle
Added tables that summarizes changes made to the ESMValCore
Added figures for ESMValTool contributions (chapter 3)
Added further tables for chapter 05 and text for chapter 06

--- 2020-11-14 end_appendix/appendix.tex
@@
 
+\begingroup
+
+\crefalias{section}{appendix}
+\numberwithin{equation}{section}
+\numberwithin{figure}{section}
+\numberwithin{table}{section}
 \renewcommand{\thesection}{\Alph{section}}
 
-\section{TBA}
+\section{Overview of \acs{CMIP} Models used in
+  \texorpdfstring{\Cref{ch:05:paper_ecs}}{Chapter \ref{ch:05:paper_ecs}}}
+\label{sec:app:cmip_models_in_chapter_5}
 
-TBA.
+\Cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models} give an overview over the
+\acs{CMIP}5 and \acs{CMIP}6 models, respectively, which are used for the
+evaluation of emergent constraints on \ac{ECS} \autocite{Schlund2020a} as
+presented in \cref{ch:05:paper_ecs}.
 
-\section{TBA}
+\begin{table}[!b]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Model & Index & Main reference \\
+    \midrule
+    ACCESS1-0 & 1 & \textcite{Dix2013} \\
+    ACCESS1-3 & 2 & \textcite{Dix2013} \\
+    BNU-ESM & 3 & \textcite{Ji2014} \\
+    CCSM4 & 4 & \textcite{Gent2011, Meehl2012} \\
+    CNRM-CM5 & 5 & \textcite{Voldoire2013} \\
+    CNRM-CM5-2 & 6 & \textcite{Voldoire2013} \\
+    CSIRO-Mk3-6-0 & 7 & \textcite{Rotstayn2012} \\
+    CanESM2 & 8 & \textcite{Arora2011} \\
+    FGOALS-g2 & 9 & \textcite{Li2013} \\
+    GFDL-CM3 & 10 & \textcite{Donner2011} \\
+    GFDL-ESM2G & 11 & \textcite{Dunne2012} \\
+    GFDL-ESM2M & 12 & \textcite{Dunne2012} \\
+    GISS-E2-H & 13 & \textcite{Schmidt2006} \\
+    GISS-E2-R & 14 & \textcite{Schmidt2006} \\
+    HadGEM2-ES & 15 & \textcite{Collins2011} \\
+    IPSL-CM5A-LR & 16 & \textcite{Dufresne2013} \\
+    IPSL-CM5A-MR & 17 & \textcite{Dufresne2013} \\
+    IPSL-CM5B-LR & 18 & \textcite{Dufresne2013} \\
+    MIROC-ESM & 19 & \textcite{Watanabe2011} \\
+    MIROC5 & 20 & \textcite{Watanabe2010} \\
+    MPI-ESM-LR & 21 & \textcite{Giorgetta2013} \\
+    MPI-ESM-MR & 22 & \textcite{Giorgetta2013} \\
+    MPI-ESM-P & 23 & \textcite{Giorgetta2013} \\
+    MRI-CGCM3 & 24 & \textcite{Yukimoto2012} \\
+    NorESM1-M & 25 & \textcite{Bentsen2013, Iversen2013} \\
+    bcc-csm1-1 & 26 & \textcite{Wu2014} \\
+    bcc-csm1-1-m & 27 & \textcite{Wu2014} \\
+    inmcm4 & 28 & \textcite{Volodin2010} \\
+    \bottomrule
+  \end{tabulary}
+  \caption{List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs} alongside
+    the main reference and the index used in the corresponding figures.
+    \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:cmip5_models}
+\end{table}
 
-TBA.
+\begin{table}[p]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Model & Index & Main reference \\
+    \midrule
+    ACCESS-CM2 & 29 & \textcite{Bi2013} \\
+    ACCESS-ESM1-5 & 30 & \textcite{Law2017, Ziehn2017} \\
+    AWI-CM-1-1-MR & 31 & \textcite{Rackow2018, Sidorenko2015} \\
+    BCC-CSM2-MR & 32 & \textcite{Wu2019} \\
+    BCC-ESM1 & 33 & \textcite{Wu2019} \\
+    CAMS-CSM1-0 & 34 & \textcite{Rong2018} \\
+    CAS-ESM2-0 & 35 & \textcite{Wang2020} \\
+    CESM2 & 36 & \textcite{Danabasoglu2020} \\
+    CESM2-FV2 & 37 & \textcite{Danabasoglu2020} \\
+    CESM2-WACCM & 38 & \textcite{Danabasoglu2020, Gettelman2019a} \\
+    CESM2-WACCM-FV2 & 39 & \textcite{Danabasoglu2020, Gettelman2019a} \\
+    CMCC-CM2-SR5 & 40 & \textcite{Cherchi2019} \\
+    CNRM-CM6-1 & 41 & \textcite{Voldoire2019} \\
+    CNRM-CM6-1-HR & 42 & \textcite{Voldoire2019} \\
+    CNRM-ESM2-1 & 43 & \textcite{Seferian2019} \\
+    CanESM5 & 44 & \textcite{Swart2019} \\
+    E3SM-1-0 & 45 & \textcite{Golaz2019} \\
+    EC-Earth3-Veg & 46 & \textcite{Wyser2020} \\
+    FGOALS-f3-L & 47 & \textcite{Guo2020, He2019, He2020} \\
+    FGOALS-g3 & 48 & \textcite{Li2020} \\
+    GISS-E2-1-G & 49 & \textcite{Rind2020} \\
+    GISS-E2-1-H & 50 & \textcite{Rind2020} \\
+    HadGEM3-GC31-LL & 51 & \textcite{Kuhlbrodt2018} \\
+    HadGEM3-GC31-MM & 52 & \textcite{Williams2018} \\
+    INM-CM4-8 & 53 & \textcite{Volodin2017,Volodin2017a} \\
+    INM-CM5-0 & 54 & \textcite{Volodin2017,Volodin2017a} \\
+    IPSL-CM6A-LR & 55 & \textcite{Boucher2020} \\
+    KACE-1-0-G & 56 & \textcite{Lee2020} \\
+    MCM-UA-1-0 & 57 & \textcite{Delworth2002} \\
+    MIROC-ES2L & 58 & \textcite{Hajima2020} \\
+    MIROC6 & 59 & \textcite{Tatebe2019} \\
+    MPI-ESM-1-2-HAM & 60 & \textcite{Mauritsen2019} \\
+    MPI-ESM1-2-HR & 61 & \textcite{Muller2018} \\
+    MPI-ESM1-2-LR & 62 & \textcite{Mauritsen2019} \\
+    MRI-ESM2-0 & 63 & \textcite{Yukimoto2019} \\
+    NESM3 & 64 & \textcite{Cao2018} \\
+    NorCPM1 & 65 & \textcite{Counillon2016} \\
+    NorESM2-LM & 66 & \textcite{Seland2020} \\
+    NorESM2-MM & 67 & \textcite{Seland2020} \\
+    SAM0-UNICON & 68 & \textcite{Park2019} \\
+    TaiESM1 & 69 & \textcite{Lee2020a} \\
+    UKESM1-0-LL & 70 & \textcite{Sellar2019} \\
+    \bottomrule
+  \end{tabulary}
+  \caption{As in \cref{tab:app:a:cmip5_models}, but for the \acs{CMIP}6 models
+    used in \cref{ch:05:paper_ecs}. \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:cmip6_models}
+\end{table}
+
+\endgroup

### 2021-02-08 ch02_scientific_background/scientific_background.tex
Adapted copyright statements
Added tables for chapter 05

@@
     Eventually, more and more components (aerosols, carbon cycle, \etc{}) were
-    added to the coupled system, forming the modern \acfp{ESM}. Taken from
-    \textcite{Cubasch2013}.}
+    added to the coupled system, forming the modern \acfp{ESM}.
+    \IPCCReproducedFrom{Cubasch2013}{their figure 1.13}.}
   \label{fig:02:esms_historical_evolution}
@@
     circular sectors show additional science themes that can be explored
-    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}. Taken from
-    \textcite{Simpkins2017}.}
+    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}.
+    \SpringerAdaptedFrom{Simpkins2017}.}
   \label{fig:02:cmip6}
@@
     discovered to be relevant or (c) decrease with additional model
-    improvements and observational constraints. Taken from
-    \textcite{Cubasch2013}.}
+    improvements and observational constraints.
+    \IPCCReproducedFrom{Cubasch2013}{their FAQ 1.1, figure 1}.}
   \label{fig:02:sources_of_uncertainty}
@@
 
-\begin{figure}[t]
+\begin{figure}[!t]
   \centering
@@
     highlights the large differences in time scales for the various feedbacks.
-    Taken from \textcite{Cubasch2013}.}
+    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.}
   \label{fig:02:feedbacks_overview}
@@
     Light gray colors correspond to the \acs{CMIP}5 ensemble (150 years each).
-    Taken from \textcite{Knutti2017}.}
+    \SpringerAdaptedFrom{Knutti2017}.}
   \label{fig:02:gregory_regression_different_years}
@@
     typical turnover time scales for carbon transfers through the major
-    reservoirs (atmosphere, land surface and ocean). Taken from
-    \textcite{Ciais2013}.}
+    reservoirs (atmosphere, land surface and ocean).
+    \IPCCReproducedFrom{Ciais2013}{their FAQ 6.2, figure 1}.}
   \label{fig:02:carbon_cycle_schematic}
@@
 occurs only during the growing season, whereas respiration occurs nearly all
-year, the larger amount of vegetation in the Northern hemisphere (due to the
+year, the larger amount of vegetation in the Northern Hemisphere (due to the
 larger land mass) gives rise to a seasonal cycle of the atmospheric \ac{CO2}
@@
 the global carbon cycle and shows a maximum of the atmospheric \ac{CO2}
-concentration in the Northern hemisphere winter (net \ac{CO2} flux into
-atmosphere due to respiration) and a minimum during the Northern hemisphere
+concentration in the Northern Hemisphere winter (net \ac{CO2} flux into
+atmosphere due to respiration) and a minimum during the Northern Hemisphere
 summer (net \ac{CO2} flux into the land due to photosynthesis). Another major
@@
 local maxima of the atmospheric \ac{CO2} concentration in the Northern
-hemisphere winter and local minima in the Northern hemisphere summer
+Hemisphere winter and local minima in the Northern Hemisphere summer
 \commentcite{Keeling1995}{see \cref{subsec:02:carbon_cycle_overview}}. Due to
@@
     carbon reservoirs. More details are given in the legend of this figure.
-    Taken from \textcite{Friedlingstein2020}.}
+    \AdaptedFrom{Friedlingstein2020}.}
   \label{fig:02:carbon_cycle_perturbation}
@@
     shaded area) and uncertainties in the emergent relationship (red shaded
-    area). Taken and modified from \textcite{Eyring2019}.}
+    area). \SpringerAdaptedFrom{Eyring2019}.}
   \label{fig:02:emergent_constraint}

### 2021-02-08 ch03_esmvaltool/esmvaltool.tex
Adapted key science questions
Added tables for chapter 05

@@
 
-In order to answer the research questions of this thesis posed in
+In order to answer the key science questions of this thesis posed in
 \cref{sec:01:research_questions}, a reliable and efficient tool to read,
@@
 Asia). Moreover, regions with high photosynthetic activity (\eg{} the boreal
-forests in the Northern hemisphere summer and the tropical rainforests in South
+forests in the Northern Hemisphere summer and the tropical rainforests in South
 America) exhibit smaller \ac{CO2} concentrations. \Cref{fig:03:co2s:b} shows
@@
 latitude with maximum zonal mean eastward wind speed at $850 \unit{hPa}$ on the
-Southern hemisphere between $80 \unit{\degree S}$ and $30 \unit{\degree S}$.
+Southern Hemisphere between $80 \unit{\degree S}$ and $30 \unit{\degree S}$.
 

### 2021-02-08 ch05_paper_ecs/paper_ecs.tex
Adapted headings in ch. 05
Added tables for chapter 05

@@
 following: Is it reasonable to adapt the assessed \ac{ECS} range from \acs{AR}5
-($\range{1.5}{4.5} \unit{K}$) in \acs{AR}6 based on the \acs{CMIP}6 model
-range? An argument for this is that due to the many new and improved processes
-included to the models, the \acs{CMIP}6 ensemble is arguably more realistic
-than its predecessor generations. However, even the \acs{CMIP}6 models are far
-from fully simulating all aspects and processes of the vastly complex Earth
-system. For this reason, the apparent increase in climate sensitivity in the
-\acs{CMIP}6 models might be wrongly caused by missing processes that cancel out
-the effects of the newly included processes, leading to a biased \ac{ECS} even
-though the models are more realistic. Thus, it is vital to take other sources
-of information into account before answering the question raised above.
+($\range{1.5}{4.5} \unit{K}$) in \acs{AR}6 based on the \acs{CMIP}6 model range
+of $\range{1.8}{5.6} \unit{K}$? An argument for this is that due to the many
+new and improved processes included to the models, the \acs{CMIP}6 ensemble is
+arguably more realistic than its predecessor generations. However, even the
+\acs{CMIP}6 models are far from fully simulating all aspects and processes of
+the full vastly complex Earth system. This could introduce new biases which
+have not been present in earlier \ac{CMIP} models. For example, the apparent
+increase in climate sensitivity in the \acs{CMIP}6 models might be wrongly
+caused by missing processes that cancel out the effects of the newly included
+processes. Thus, it is vital to take other sources of information into account
+before answering the question raised above.
 
@@
 undoubtedly observational data. As presented in
-\cref{subsec:02:emergent_constraints}, emergent constraints are a beneficial
-tool to assess the accuracy of climate model projections and reduce associated
-uncertainties with appropriate observations of the Earth system. For \ac{ECS},
-a large variety of different emergent constraints has been developed over the
-years, which we now evaluate on the new \acs{CMIP}6 ensemble. In total we
-assess eleven emergent constraints (selected by their availability in the
-\ac{ESMValTool}), out of which the most are related to cloud feedbacks, since
-these constitute the most important source of uncertainty for \ac{ECS}
-\commentcite{Boucher2013, Flato2013}{see also
-  \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
-constraints have been derived on the \acs{CMIP}3 or \ac{CMIP}5 ensemble, the
-\ac{CMIP}6 models offer an exciting possibility to perform out-of-sample
-testing using the previously unseen data. We quantify the skill of the emergent
-constraints with three metrics: the coefficient of determination $R^2$ of the
-linear relationship, the statistical significance using a two-sided $t$-test
-based on the Pearson correlation coefficient $r$ (see \cref{eq:02:t}) and the
-constrained range of \ac{ECS} based on the constrained \ac{PDF} (see
-\cref{eq:02:pdf_y_given_x0}). This analysis has already been published in
-\textcite{Schlund2020a}. For this paper, the author of this thesis lead the
-writing and the analysis and contributed the code to reproduce this analysis
-including all figures and tables with the \ac{ESMValTool}. In this chapter, we
-first introduce the methods and data (climate model output and observations)
-used in this study (\cref{sec:05:methods_and_data}). After that, we present the
-eleven emergent constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5
-and \acs{CMIP}6 ensemble (\cref{sec:05:comparison_of_emergent_constraints}).
-Finally, we provide a discussion (\cref{sec:05:discussion}) and a summary of
-the results (\cref{sec:05:summary}).
+\cref{subsec:02:emergent_constraints}, appropriate observations of the Earth
+system can be utilized within the emergent constraints framework to assess the
+accuracy of climate model projections and reduce associated uncertainties. For
+\ac{ECS}, a large variety of different emergent constraints has been developed
+over the years, which we now evaluate on the \acs{CMIP}6 ensemble in order to
+assess whether they still hold for these new climate models. In total we assess
+eleven emergent constraints on \ac{ECS}, out of which the most are related to
+cloud feedbacks since these constitute the most important source of uncertainty
+for \ac{ECS} \commentcite{Boucher2013, Flato2013}{see also
+  \cref{subsec:02:climate_feedbacks}}. A complete list of these emergent
+constraints, which are selected by their availability in the \ac{ESMValTool},
+is given in \cref{tab:05:overview_emergent_constraints}. Since all of the
+evaluated emergent constraints have been derived on the \acs{CMIP}3 or
+\ac{CMIP}5 ensemble, the \ac{CMIP}6 models offer an exciting possibility to
+perform out-of-sample testing using the previously unseen data.
 
-\section{Methods and Data}
-\label{sec:05:methods_and_data}
+\begin{table}[!t]
+  \centering
+  \hbadness=3500
+  \begin{tabular}{p{0.07\columnwidth} p{0.24\columnwidth} p{0.29\columnwidth}
+      p{0.29\columnwidth}}
+    \toprule
+    Label & Reference & Description of \xaxis{} & Variables and corresponding
+    observations \\
+    \midrule
+    BRI & \textcite{Brient2016} & Sensitivity of shortwave cloud albedo
+    to changes in \acf{SST} [$\si{\pctK{}}$] & \makecell{\tabitem\emph{hur}
+      (ERA-Interim) \\ \tabitem \emph{rsdt} (CERES-EBAF) \\ \tabitem
+      \emph{rsut} (CERES-EBAF) \\ \tabitem \emph{rsutcs} (CERES-EBAF) \\
+      \tabitem \emph{ts} (HadISST)} \\
+    \midrule
+    COX & \textcite{Cox2018} & $\Psi$ (temperature variability metric)
+    [$\si{K}$] &
+    \tabitem \emph{tas} (HadCRUT4) \\
+    \midrule
+    LIP & \textcite{Lipat2017} & Extent of Southern Hemisphere Hadley cell
+    [$\si{\degree}$] & \tabitem \emph{va} (ERA-Interim) \\
+    \midrule
+    SHD & \textcite{Sherwood2014} & $D$-index (large-scale lower-tropospheric
+    mixing) [$\si{1}$] & \tabitem \emph{wap} (ERA-Interim) \\
+    \midrule
+    SHL & \textcite{Sherwood2014} & \acs{LTMI} (\acl{LTMI}) [$\si{1}$] &
+    \makecell{\tabitem \emph{hur} (ERA-Interim) \\ \tabitem \emph{ta}
+      (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
+    \midrule
+    SHS & \textcite{Sherwood2014} & $S$-index (small-scale lower-tropospheric
+    mixing) [$\si{1}$] & \makecell{\tabitem \emph{hur} (ERA-Interim) \\
+      \tabitem \emph{ta} (ERA-Interim) \\ \tabitem \emph{wap} (ERA-Interim)} \\
+    \midrule
+    SU & \textcite{Su2014} & Error in vertical profile of relative humidity
+    [$\si{1}$] & \tabitem \emph{hur} (AIRS, MLS-Aura) \\
+    \midrule
+    TIH & \textcite{Tian2015} & Tropical mid-tropospheric humidity asymmetry
+    index from AIRS [$\si{\%}$] & \tabitem \emph{hus} (AIRS) \\
+    \midrule
+    TII & \textcite{Tian2015} & Southern ITCZ index from GPCP
+    [$\si{\mmday{}}$] & \tabitem \emph{pr} (GPCP) \\
+    \midrule
+    VOL* & \textcite{Volodin2008} & Difference between tropical and southern
+    midlatitudinal cloud fraction [$\si{\%}$] & \tabitem \emph{clt} (ISCCP
+    D-2) \\
+    \midrule
+    ZHA* & \textcite{Zhai2015} & Seasonal response of marine boundary layer
+    cloud fraction to changes in \acs{SST} [$\si{\pctK{}}$] &
+    \makecell{\tabitem \emph{cl} (CloudSat/CALIPSO) \\ \tabitem \emph{tos}
+      (AMSRE SST) \\ \tabitem \emph{wap} (ERA-Interim)} \\
+    \bottomrule
+  \end{tabular}
+  \caption{Overview of the eleven emergent constraints on the \acf{ECS}
+    evaluated in this chapter. Detailed descriptions of the variables used to
+    calculate the \xaxis{} values of the emergent constraints and the
+    references for the corresponding observations (fourth column) are given in
+    \cref{tab:app:a:overview_variables,tab:app:a:observations}, respectively.
+    For emergent constraints marked with an asterisk (*), the observational
+    value of the original publication is used to calculate the observational
+    constraint. For all others, the observational value is computed with the
+    \acs{ESMValTool}. \AdaptedFrom{Schlund2020a}.}
+  \label{tab:05:overview_emergent_constraints}
+\end{table}
+
+This analysis is already published in \textcite{Schlund2020a}. For this paper,
+the author of this thesis lead the writing and the analysis and implemented the
+code to reproduce this analysis with all figures and tables using the
+\ac{ESMValTool}. In this chapter, we first introduce the data (climate model
+output and observations) and methods used in this study
+(\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
+constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
+ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
+provide a discussion (\cref{sec:05:discussion}) and a summary of the results
+(\cref{sec:05:summary}).
+
+\section{Data and Methods}
+\label{sec:05:data_and_methods}
 
@@
 in \cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models}, respectively.
+
+We quantify the skill of the emergent constraints with three metrics: the
+coefficient of determination $R^2$ of the linear relationship, the statistical
+significance using a two-sided $t$-test based on the Pearson correlation
+coefficient $r$ (see \cref{eq:02:t}) and the constrained range of \ac{ECS}
+based on the constrained \ac{PDF} (see \cref{eq:02:pdf_y_given_x0}).
 
@@
 
+\subsection{SHL}
+\label{subsec:05:shl}
+
+This is the \ac{LTMI}!!
+
 \section{Discussion}

### 2021-02-08 end_appendix/appendix.tex
Added tables for chapter 05

@@
 
-\section{Overview of \acs{CMIP} Models used in
+\section{Supplementary Materials for
   \texorpdfstring{\Cref{ch:05:paper_ecs}}{Chapter \ref{ch:05:paper_ecs}}}
@@
 
-\Cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models} give an overview over the
-\acs{CMIP}5 and \acs{CMIP}6 models, respectively, which are used for the
-evaluation of emergent constraints on \ac{ECS} \autocite{Schlund2020a} as
-presented in \cref{ch:05:paper_ecs}.
+This section contains supplementary tables for \cref{ch:05:paper_ecs}
+(\nameref*{ch:05:paper_ecs}; see \cite{Schlund2020a}).
 
 \begin{table}[!b]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L}
+    \toprule
+    Variable short name & Description \\
+    \midrule
+    \emph{cl} & Cloud area fraction \\
+    \emph{clt} & Total cloud area fraction \\
+    \emph{hur} & Relative humidity \\
+    \emph{hus} & Specific humidity \\
+    \emph{pr} & Precipitation \\
+    \emph{rsdt} & \Acf{TOA} incoming shortwave radiation \\
+    \emph{rsut} & \acs{TOA} outgoing shortwave radiation \\
+    \emph{rsutcs} & Clear-sky \acs{TOA} outgoing shortwave radiation \\
+    \emph{ta} & Air temperature \\
+    \emph{tas} & Surface air temperature \\
+    \emph{tos} & Sea surface temperature \\
+    \emph{ts} & Surface temperature \\
+    \emph{va} & Northward wind speed \\
+    \emph{wap} & Vertical velocity \\
+    \bottomrule
+  \end{tabulary}
+  \caption{Overview of the variables used in \cref{ch:05:paper_ecs}
+    (\nameref*{ch:05:paper_ecs}). More details are given in
+    \cref{tab:05:overview_emergent_constraints}, which lists the corresponding
+    variable short names for each emergent constraint.}
+  \label{tab:app:a:overview_variables}
+\end{table}
+
+\begin{table}[!b]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Observational dataset & Corresponding variables & Reference \\
+    \midrule
+    AIRS & \emph{hur}, \emph{hus} & \textcite{Aumann2003} \\
+    AMSRE SST & \emph{tos} & \textcite{AMSRE2011} \\
+    CERES-EBAF & \emph{rsdt}, \emph{rsut}, \emph{rsutcs} & \textcite{Loeb2018}
+    \\
+    CloudSat/CALIPSO & \emph{cl} & \textcite{Mace2009} \\
+    ERA-Interim & \emph{hur}, \emph{ta}, \emph{va}, \emph{wap} &
+    \textcite{Dee2011} \\
+    GPCP & \emph{pr} & \textcite{Adler2003} \\
+    HadCRUT4 & \emph{tas} & \textcite{Morice2012} \\
+    HadISST & \emph{ts} & \textcite{Rayner2003} \\
+    ISCCP D-2 & \emph{clt} & \textcite{Rossow1991} \\
+    MLS-Aura & \emph{hur} & \textcite{Beer2006} \\
+    \bottomrule
+  \end{tabulary}
+  \caption{References for all observational datasets used in
+    \cref{ch:05:paper_ecs} (\nameref*{ch:05:paper_ecs}).}
+  \label{tab:app:a:observations}
+\end{table}
+
+\begin{table}[p]
   \centering
@@
   \end{tabulary}
-  \caption{List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs} alongside
-    the main reference and the index used in the corresponding figures.
-    \AdaptedFrom{Schlund2020a}.}
+  \caption{List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
+    (\nameref*{ch:05:paper_ecs}) alongside the main reference and the index
+    used in the corresponding figures. \AdaptedFrom{Schlund2020a}.}
   \label{tab:app:a:cmip5_models}
@@
   \end{tabulary}
-  \caption{As in \cref{tab:app:a:cmip5_models}, but for the \acs{CMIP}6 models
-    used in \cref{ch:05:paper_ecs}. \AdaptedFrom{Schlund2020a}.}
+  \caption{As in \cref{tab:app:a:cmip5_models}, but for the \acs{CMIP}6 models.
+    \AdaptedFrom{Schlund2020a}.}
   \label{tab:app:a:cmip6_models}
@@
 
+\begin{table}[p]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    A & B & C \\
+    \midrule
+    111111111111111111111111111111111111111111111 & 2 & 3 \\
+    \bottomrule
+  \end{tabulary}
+  \caption{\acs{CMIP}5. \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:emergent_constraints_cmip5}
+\end{table}
+
+\begin{table}[p]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    A & B & C \\
+    \midrule
+    111111111111111111111111111111111111111111111 & 2 & 3 \\
+    \bottomrule
+  \end{tabulary}
+  \caption{\acs{CMIP}6. \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:emergent_constraints_cmip6}
+\end{table}
+
 \endgroup

### 2021-02-09 ch02_scientific_background/scientific_background.tex
Added further tables to chapter 05

@@
 models. For \ac{ScenarioMIP}, five different \acp{SSP} are considered, ranging
-from \ac{SSP}1 (sustainability) to \ac{SSP}5 (fossil-fuel development). Each
+from \acs{SSP}1 (sustainability) to \acs{SSP}5 (fossil-fuel development). Each
 \ac{SSP} is combined with a climate outcome (measured as radiative forcing in
 the year 2100) based on a particular forcing pathway that \acp{IAM} have shown
-to be feasible. For example, \ac{SSP}5-8.5 represents a scenario based on a
+to be feasible. For example, \acs{SSP}5-8.5 represents a scenario based on a
 fossil-fuel development with a radiative forcing of $8.5 \unit{W m^{-2}}$ in
-2100 while \ac{SSP}1-2.6 represents a sustainable future with a radiative
+2100 while \acs{SSP}1-2.6 represents a sustainable future with a radiative
 forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two scenarios in the
 main category of \ac{ScenarioMIP}, the \emph{Tier 1} experiments, are the
-\ac{SSP}2-4.5 and \ac{SSP}3-7.0 scenarios. In contrast to the \ac{ScenarioMIP}
-experiments, the corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012},
-the so-called \acp{RCP}, only used the radiative forcing in 2100 as only
-dimension to describe the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5,
-\ac{RCP}2.6, \etc{}).
+\acs{SSP}2-4.5 and \acs{SSP}3-7.0 scenarios. In contrast to the
+\ac{ScenarioMIP} experiments, the corresponding \acs{CMIP}5 counterparts
+\autocite{Taylor2012}, the so-called \acp{RCP}, only used the radiative forcing
+in 2100 as only dimension to describe the possible futures (\eg{} \acs{RCP}8.5,
+\acs{RCP}4.5, \acs{RCP}2.6, \etc{}).
 
@@
 caused by a doubling of the atmospheric \ac{CO2} concentration has
-substantially increased in \ac{CMIP}6 compared to older \ac{CMIP} generations
+substantially increased in \acs{CMIP}6 compared to older \ac{CMIP} generations
 \autocite{Zelinka2020}. On the other hand, the climate response uncertainty can
@@
 \begin{equation}
-  R = -4 \sigma T_0^3 \cdot \Delta T \coloneq \lambda_\text{BB} \cdot \Delta T
+  R = -4 \sigma T_0^3 \cdot \Delta T \coloneq \lambda_\mathtxt{BB} \cdot
+  \Delta T
   \label{eq:02:linear_blackbody_response}
 \end{equation}
-with the blackbody feedback parameter $\lambda_\text{BB} \approx -3.8 \unit{W
-  m^{-2} K^{-1}}$. Results from climate models and analyses of observations
-confirm this linear relationship between $R$ and $\Delta T$
+with the blackbody feedback parameter $\lambda_\mathtxt{BB} \approx -3.8
+\unit{W m^{-2} K^{-1}}$. Results from climate models and analyses of
+observations confirm this linear relationship between $R$ and $\Delta T$
 \autocite{Gregory2004}. However, the value of this linear constant $\lambda$,
@@
 \begin{equation}
-  R = \lambda \cdot \Delta T = \left( \lambda_\text{BB} + \lambda_\text{WV} +
-  \lambda_\text{Albedo} + \lambda_\text{Cloud} + \ldots \right) \cdot \Delta T.
+  R = \lambda \cdot \Delta T = \left( \lambda_\mathtxt{BB} +
+  \lambda_\mathtxt{WV} + \lambda_\mathtxt{Albedo} + \lambda_\mathtxt{Cloud} +
+  \ldots \right) \cdot \Delta T.
   \label{eq:02:linear_response}
 \end{equation}
-$\lambda_\text{WV}$ refers to the water vapor feedback, $\lambda_\text{Albedo}$
-to the snow/ice albedo feedback and $\lambda_\text{Cloud}$ to the cloud
-feedback. Thus, the overall climate feedback parameter $\lambda$ can be written
-as the sum of the individual feedback parameters $\lambda_i$:
+$\lambda_\mathtxt{WV}$ refers to the water vapor feedback,
+$\lambda_\mathtxt{Albedo}$ to the snow/ice albedo feedback and
+$\lambda_\mathtxt{Cloud}$ to the cloud feedback. Thus, the overall climate
+feedback parameter $\lambda$ can be written as the sum of the individual
+feedback parameters $\lambda_i$:
 \begin{equation}
@@
     parameter $\lambda$ and the $y$-intercept corresponds to the radiative
-    forcing $F_\text{4x}$. These can be used to calculate the \acf{ECS} with
+    forcing $F_\mathtxt{4x}$. These can be used to calculate the \acf{ECS} with
     the Gregory regression method according to \cref{eq:02:ecs}
@@
     \xintercept{} of the linear regression line divided by $2$. Here,
-    $\text{ECS} = 3.14 \unit{K}$.}
+    $\mathtxt{ECS} = 3.14 \unit{K}$.}
   \label{fig:02:gregory_regression}
@@
 \end{equation}
-$F$ is now given by the \yintercept{} of this linear regression ($F_\text{4x}$)
-and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
-\begin{equation}
-  \text{ECS} = -\frac{F_\text{4x}}{2 \lambda}.
+$F$ is now given by the \yintercept{} of this linear regression
+($F_\mathtxt{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
+\begin{equation}
+  \mathtxt{ECS} = -\frac{F_\mathtxt{4x}}{2 \lambda}.
   \label{eq:02:ecs}
@@
 (\ie{} clouds artificially removed) net \ac{TOA} radiation
-\autocite{Andrews2012}. This can be done for the shortwave $N_\text{SWCRE}$ and
-longwave $N_\text{LWCRE}$ components separately, but also for the combined
-effect $N_\text{CRE} = N_\text{SWCRE} + N_\text{LWCRE}$. The slopes in the
-corresponding Gregory regressions are the so-called \emph{\ac{CRE} feedback
-  parameters} $\lambda_\text{SWCRE}$, $\lambda_\text{LWCRE}$ and
-$\lambda_\text{CRE}$, which quantify the change in \ac{CRE} as a response to
-increasing \acp{GSAT}.
+\autocite{Andrews2012}. This can be done for the shortwave $N_\mathtxt{SWCRE}$
+and longwave $N_\mathtxt{LWCRE}$ components separately, but also for the
+combined effect $N_\mathtxt{CRE} = N_\mathtxt{SWCRE} + N_\mathtxt{LWCRE}$. The
+slopes in the corresponding Gregory regressions are the so-called
+\emph{\ac{CRE} feedback parameters} $\lambda_\mathtxt{SWCRE}$,
+$\lambda_\mathtxt{LWCRE}$ and $\lambda_\mathtxt{CRE}$, which quantify the
+change in \ac{CRE} as a response to increasing \acp{GSAT}.
 
@@
     over a 20 year period (illustrated by the vertical dashed lines). Here,
-    $\text{TCR} = 2.35 \unit{K}$ (horizontal dashed line).}
+    $\mathtxt{TCR} = 2.35 \unit{K}$ (horizontal dashed line).}
   \label{fig:02:tcr}
@@
 \begin{equation}
-  \text{TCR} = \frac{F_\text{2x}}{\kappa - \lambda},
+  \mathtxt{TCR} = \frac{F_\mathtxt{2x}}{\kappa - \lambda},
   \label{eq:02:tcr}
 \end{equation}
-where $F_\text{2x}$ is the radiative forcing induced by a doubling of the
+where $F_\mathtxt{2x}$ is the radiative forcing induced by a doubling of the
 atmospheric \ac{CO2} concentration.
@@
 This equation can be used to derive a relationship between \ac{TCR} and
-\ac{ECS}. Writing $\text{ECS} = -F_\text{2x} / \lambda$ and assuming path
+\ac{ECS}. Writing $\mathtxt{ECS} = -F_\mathtxt{2x} / \lambda$ and assuming path
 independence of the forcing (\ie{} the resulting radiative forcing from the
@@
 \begin{equation}
-  \text{TCR} = \frac{F_\text{2x} \cdot \text{ECS}}{F_\text{2x} + \kappa \cdot
-    \text{ECS}},
+  \mathtxt{TCR} = \frac{F_\mathtxt{2x} \cdot \mathtxt{ECS}}{F_\mathtxt{2x} +
+    \kappa \cdot \mathtxt{ECS}},
   \label{eq:02:tcr_vs_ecs}
@@
 \cref{eq:02:tcr} implies that the equilibrium response \ac{ECS} is (as
-expected) larger than transient response \ac{TCR}, \ie{} $\text{ECS} >
-\text{TCR}$.
+expected) larger than transient response \ac{TCR}, \ie{} $\mathtxt{ECS} >
+\mathtxt{TCR}$.
 

### 2021-02-09 ch03_esmvaltool/esmvaltool.tex
Added further tables to chapter 05

@@
   \centering
-  \begin{tabulary}{\columnwidth}{L L L}
+  \begin{tabulary}{\columnwidth}{L >{\em}L L}
     \toprule
-    Type & Name & Description \\
+    Type & \textnormal{Name} & Description \\
     \midrule
-    Preprocessor & \emph{amplitude} & Amplitude of periodic phenomena (\eg{}
-    cycles) \\
+    Preprocessor & amplitude & Amplitude of periodic phenomena (\eg{} cycles)
+    \\
     \cmidrule(l){2-3}
-    & \emph{land/sea fraction weighting} & Weighting of fields based on
-    the land or sea fraction of the respective grid cells \\
+    & land/sea fraction weighting & Weighting of fields based on the land or
+    sea fraction of the respective grid cells \\
     \cmidrule(l){2-3}
-    & \emph{trend} & Linear trend (slope of linear regression) \\
+    & trend & Linear trend (slope of linear regression) \\
     \midrule
-    Derived variable & \emph{asr} & Absorbed shortwave radiation \\
+    Derived variable & asr & Absorbed shortwave radiation \\
     \cmidrule(l){2-3}
-    & \emph{co2s} & Atmospheric \ac{CO2} concentration at
+    & co2s & Atmospheric \ac{CO2} concentration at
     surface \\
     \cmidrule(l){2-3}
-    & \emph{et} & Evapotranspiration \\
+    & et & Evapotranspiration \\
     \cmidrule(l){2-3}
-    & \emph{rlntcs} & Clear-sky net \acf{TOA} longwave radiation \\
+    & rlntcs & Clear-sky net \acf{TOA} longwave radiation \\
     \cmidrule(l){2-3}
-    & \emph{rsntcs} & Clear-sky net \ac{TOA} shortwave radiation \\
+    & rsntcs & Clear-sky net \ac{TOA} shortwave radiation \\
     \cmidrule(l){2-3}
-    & \emph{uajet} & Position of austral jet stream \\
+    & uajet & Position of austral jet stream \\
     \bottomrule

### 2021-02-09 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added further tables to chapter 05

@@
     calculate the \acf{ECS} according to \cref{eq:02:ecs} with the radiative
-    forcing $F_\text{4x}$ (\yintercept) and the climate feedback parameter
+    forcing $F_\mathtxt{4x}$ (\yintercept) and the climate feedback parameter
     $\lambda$ (slope) for all 150 years (black line), which results in
-    $\text{ECS} = 3.74 \unit{K}$. Due to non-linear effects, the exact value
-    of \acs{ECS} depends on the years considered for the Gregory regression
-    (see \cref{subsec:02:ecs} and
+    $\mathtxt{ECS} = 3.74 \unit{K}$. Due to non-linear effects, the exact
+    value of \acs{ECS} depends on the years considered for the Gregory
+    regression (see \cref{subsec:02:ecs} and
     \cref{fig:02:gregory_regression_different_years}): Using only the first 20
@@
   \csvreader[ECSAndTCRTable]{
-    ch04_papers_ecs_tcr_assessment/data/CMIP5_ecs-tcr.csv}{}{
-    \dataset & \idx & \ecs & \tcr}
+    ch04_papers_ecs_tcr_assessment/data/cmip5_ecs-tcr.csv}{}{
+    \dataset & $\idx$ & $\ecs$ & $\tcr$}
   \caption{\acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
@@
   \csvreader[ECSAndTCRTable]{
-    ch04_papers_ecs_tcr_assessment/data/CMIP6_ecs-tcr.csv}{}{
-    \dataset & \idx & \ecs & \tcr}
+    ch04_papers_ecs_tcr_assessment/data/cmip6_ecs-tcr.csv}{}{
+    \dataset & $\idx$ & $\ecs$ & $\tcr$}
   \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.
@@
 high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exists a
-number of models with very low values of \ac{ECS}. Therefore, the \ac{CMIP}6
+number of models with very low values of \ac{ECS}. Therefore, the \acs{CMIP}6
 model ranges of \ac{ECS} and \ac{TCR} are well outside the corresponding
@@
 $4.5 \unit{K}$. Moreover, it is not only a small fraction of models that
-exceeds this upper bound: out of the 42 analyzed \ac{CMIP}6 models, a third (14
-models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
+exceeds this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third
+(14 models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
 several modeling centers independently confirmed high \ac{ECS} values in their
@@
     with the corresponding \acf{MMM} illustrated with the horizontal black
-    lines (the green dot represents output from a single climate model). The
+    lines (the green circle represents output from a single climate model). The
     numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
@@
 Because of the massive political and societal relevance of \ac{ECS}, its
-apparent increase in the \ac{CMIP}6 ensemble is currently one of the most
+apparent increase in the \acs{CMIP}6 ensemble is currently one of the most
 important questions for the entire climate modeling community. For this reason,

### 2021-02-09 ch05_paper_ecs/paper_ecs.tex
Added further tables to chapter 05

@@
 assess whether they still hold for these new climate models. In total we assess
-eleven emergent constraints on \ac{ECS}, out of which the most are related to
-cloud feedbacks since these constitute the most important source of uncertainty
-for \ac{ECS} \commentcite{Boucher2013, Flato2013}{see also
-  \cref{subsec:02:climate_feedbacks}}. A complete list of these emergent
-constraints, which are selected by their availability in the \ac{ESMValTool},
-is given in \cref{tab:05:overview_emergent_constraints}. Since all of the
-evaluated emergent constraints have been derived on the \acs{CMIP}3 or
-\ac{CMIP}5 ensemble, the \ac{CMIP}6 models offer an exciting possibility to
-perform out-of-sample testing using the previously unseen data.
+eleven emergent constraints on \ac{ECS} (see
+\cref{tab:05:overview_emergent_constraints}), out of which the most are related
+to cloud feedbacks since these constitute the most important source of
+uncertainty for \ac{ECS} \commentcite{Boucher2013, Flato2013}{see also
+  \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
+constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, the
+\acs{CMIP}6 models offer an exciting possibility to perform out-of-sample
+testing using the previously unseen data.
 
-\begin{table}[!t]
+This analysis is already published in \textcite{Schlund2020a}. For this paper,
+the author of this thesis lead the writing and the analysis and implemented the
+code to reproduce this analysis with all figures and tables using the
+\ac{ESMValTool}. In this chapter, we first introduce the data (climate model
+output and observations) and methods used in this study
+(\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
+constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
+ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
+provide a discussion (\cref{sec:05:discussion}) and a summary of the results
+(\cref{sec:05:summary}).
+
+\begin{table}[p]
   \centering
@@
 
-This analysis is already published in \textcite{Schlund2020a}. For this paper,
-the author of this thesis lead the writing and the analysis and implemented the
-code to reproduce this analysis with all figures and tables using the
-\ac{ESMValTool}. In this chapter, we first introduce the data (climate model
-output and observations) and methods used in this study
-(\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
-constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
-ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
-provide a discussion (\cref{sec:05:discussion}) and a summary of the results
-(\cref{sec:05:summary}).
-
 \section{Data and Methods}
@@
 In this study we use the output from climate models participating in
-\acs{CMIP}5 and \acs{CMIP}6, shown in \cref{sec:app:cmip_models_in_chapter_5}
-in \cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models}, respectively.
-
-We quantify the skill of the emergent constraints with three metrics: the
-coefficient of determination $R^2$ of the linear relationship, the statistical
-significance using a two-sided $t$-test based on the Pearson correlation
-coefficient $r$ (see \cref{eq:02:t}) and the constrained range of \ac{ECS}
-based on the constrained \ac{PDF} (see \cref{eq:02:pdf_y_given_x0}).
+\acs{CMIP}5 and \acs{CMIP}6 as shown in
+\cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models}, respectively. In
+addition, for each emergent constraint at least one observational dataset is
+used to calculate the observational constraint. A complete list of these
+datasets is given in \cref{tab:05:overview_emergent_constraints} and in
+\cref{tab:app:a:observations}. Following other similar studies, we calculate
+\ac{ECS} with the Gregory regression method \autocite{Gregory2004}, which is
+described in detail in \cref{subsec:02:ecs}. An overview of the eleven emergent
+constraints on \ac{ECS} analyzed in this study including the variables required
+for their calculations is given in \cref{tab:05:overview_emergent_constraints}
+and \cref{sec:05:comparison_of_emergent_constraints}. We chose these particular
+emergent constraints since these were already implemented in the
+\ac{ESMValTool} at the time of writing this study, which greatly facilitated
+this analysis. For all emergent constraints, we use the historical simulations
+of \acs{CMIP}5 and \acs{CMIP}6 in order to ensure maximum agreement with the
+observational data. If necessary, the historical simulation of \acs{CMIP}5 is
+extended after its final year 2005 with data from the \acs{RCP}8.5 scenario
+\autocite{Riahi2011}. Note that we only use data through 2014, during which
+time all \ac{RCP} scenarios behave similarly and the choice of the scenario is
+not expected to affect the results considerably. Such an extension is not
+needed for \acs{CMIP}6 models as their historical simulations cover a longer
+time period until 2014. We quantify the skill of the emergent constraints with
+three metrics: the coefficient of determination $R^2$ of the linear
+relationship, the statistical significance using a two-sided $t$-test based on
+the Pearson correlation coefficient $r$ (see \cref{eq:02:t}) and the
+constrained range of \ac{ECS} based on the constrained \ac{PDF} (see
+\cref{eq:02:pdf_y_given_x0}).
 
@@
   \centering
-  \begin{subfigure}[b]{\SubfigureWidth{}}
+  \begin{subfigure}[b]{\SmallSubfigureWidth{}}
     \includegraphics[width=\columnwidth]{
@@
   ~
-  \begin{subfigure}[b]{\SubfigureWidth{}}
+  \begin{subfigure}[b]{\SmallSubfigureWidth{}}
     \includegraphics[width=\columnwidth]{
@@
   \\
-  \begin{subfigure}[b]{\SubfigureWidth{}}
+  \begin{subfigure}[b]{\SmallSubfigureWidth{}}
     \includegraphics[width=\columnwidth]{
@@
   \end{subfigure}
-  \caption{xD. \AdaptedFrom{Schlund2020a}.}
+  \caption{Emergent constraints BRI, COX and LIP applied to the \acs{CMIP}5
+    (blue) and \acs{CMIP}6 ensemble (orange). (a), (c), (e) Emergent
+    relationships (solid lines) for the \acs{CMIP} models (numbers; see
+    \cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models} for details) with
+    their standard prediction errors (blue and orange shaded areas; see
+    \cref{eq:02:spe}). The vertical dashed line corresponds to the
+    observational reference with its standard error (gray shaded area). The
+    horizontal dashed lines show the best estimates of the constrained
+    \acf{ECS} and the colored points mark the \aclp{MMM} for \acs{CMIP}5
+    (blue) and \acs{CMIP}6 (orange). (b), (d), (f) \Aclp{PDF} for \acs{ECS} of
+    the emergent constraints (solid lines; see \cref{eq:02:pdf_y_given_x0})
+    and the unconstrained model ensembles (histograms). Due to the
+    availability of data, the histograms may differ for the different emergent
+    constraints. \AdaptedFrom{Schlund2020a}.}
   \label{fig:05:bri_cox_lip}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_SHD.pdf}
+    \caption{}
+    \label{fig:05:shd_shl_shs:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_SHD.pdf}
+    \caption{}
+    \label{fig:05:shd_shl_shs:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_SHL.pdf}
+    \caption{}
+    \label{fig:05:shd_shl_shs:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_SHL.pdf}
+    \caption{}
+    \label{fig:05:shd_shl_shs:d}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_SHS.pdf}
+    \caption{}
+    \label{fig:05:shd_shl_shs:e}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_SHS.pdf}
+    \caption{}
+    \label{fig:05:shd_shl_shs:f}
+  \end{subfigure}
+  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints
+    SHD, SHL and SHS. \AdaptedFrom{Schlund2020a}.}
+  \label{fig:05:shd_shl_shs}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_SU.pdf}
+    \caption{}
+    \label{fig:05:su_tih_tii:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_SU.pdf}
+    \caption{}
+    \label{fig:05:su_tih_tii:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_TIH.pdf}
+    \caption{}
+    \label{fig:05:su_tih_tii:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_TIH.pdf}
+    \caption{}
+    \label{fig:05:su_tih_tii:d}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_TII.pdf}
+    \caption{}
+    \label{fig:05:su_tih_tii:e}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_TII.pdf}
+    \caption{}
+    \label{fig:05:su_tih_tii:f}
+  \end{subfigure}
+  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints
+    SU, TIH and TII. \AdaptedFrom{Schlund2020a}.}
+  \label{fig:05:su_tih_tii}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_VOL.pdf}
+    \caption{}
+    \label{fig:05:vol_zha:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_VOL.pdf}
+    \caption{}
+    \label{fig:05:vol_zha:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_ZHA.pdf}
+    \caption{}
+    \label{fig:05:vol_zha:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch05_paper_ecs/figs/target_distribution_training_data_ZHA.pdf}
+    \caption{}
+    \label{fig:05:vol_zha:d}
+  \end{subfigure}
+  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints
+    VOL and ZHA. \AdaptedFrom{Schlund2020a}.}
+  \label{fig:05:vol_zha}
 \end{figure}
@@
 
+\begin{center}
+  \newcommand{\testiii}{YOTHISISATEST!}
+  hi
+  \testiii{}
+\end{center}
+
 \section{Discussion}

### 2021-02-09 end_appendix/appendix.tex
Added further tables to chapter 05

@@
   \end{tabulary}
-  \caption{As in \cref{tab:app:a:cmip5_models}, but for the \acs{CMIP}6 models.
+  \caption{As in \cref{tab:app:a:cmip5_models} but for the \acs{CMIP}6 models.
     \AdaptedFrom{Schlund2020a}.}
@@
   \centering
-  \begin{tabulary}{\columnwidth}{L L L}
-    \toprule
-    A & B & C \\
-    \midrule
-    111111111111111111111111111111111111111111111 & 2 & 3 \\
-    \bottomrule
-  \end{tabulary}
-  \caption{\acs{CMIP}5. \AdaptedFrom{Schlund2020a}.}
-  \label{tab:app:a:emergent_constraints_cmip5}
+  \csvreader[EmergentConstraintsPart1Table]{
+    ch05_paper_ecs/data/cmip5_emergent_constraints.csv}{}{
+    \dataset & $\idx$ & $\ECS$ & $\BRI$ & $\COX$ & $\LIP$ & $\SHD$ & $\SHL$ &
+    $\SHS$}
+  \caption{All \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
+    (\nameref*{ch:05:paper_ecs}) including their \acf{ECS} and \xaxis{} values
+    for the emergent constraints BRI, COX, LIP, SHD, SHL and SHS. More details
+    on the emergent constraints are given in
+    \cref{tab:05:overview_emergent_constraints}. The specified index
+    corresponds to the index used in the associated figures.
+    \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:cmip5_emergent_constraints_part1}
 \end{table}
@@
   \centering
-  \begin{tabulary}{\columnwidth}{L L L}
-    \toprule
-    A & B & C \\
-    \midrule
-    111111111111111111111111111111111111111111111 & 2 & 3 \\
-    \bottomrule
-  \end{tabulary}
-  \caption{\acs{CMIP}6. \AdaptedFrom{Schlund2020a}.}
-  \label{tab:app:a:emergent_constraints_cmip6}
+  \csvreader[EmergentConstraintsPart2Table]{
+    ch05_paper_ecs/data/cmip5_emergent_constraints.csv}{}{
+    \dataset & $\idx$ & $\ECS$ & $\SU$ & $\TIH$ & $\TII$ & $\VOL$ & $\ZHA$}
+  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
+    emergent constraints SU, TIH, TII, VOL and ZHA.
+    \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:cmip5_emergent_constraints_part2}
+\end{table}
+
+\begin{table}[p]
+  \centering
+  \csvreader[EmergentConstraintsPart1Table]{
+    ch05_paper_ecs/data/cmip6_emergent_constraints.csv}{}{
+    \dataset & $\idx$ & $\ECS$ & $\BRI$ & $\COX$ & $\LIP$ & $\SHD$ & $\SHL$ &
+    $\SHS$}
+  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
+    \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:cmip6_emergent_constraints_part1}
+\end{table}
+
+\begin{table}[p]
+  \centering
+  \csvreader[EmergentConstraintsPart2Table]{
+    ch05_paper_ecs/data/cmip6_emergent_constraints.csv}{}{
+    \dataset & $\idx$ & $\ECS$ & $\SU$ & $\TIH$ & $\TII$ & $\VOL$ & $\ZHA$}
+  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part2} but for the
+  \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.}
+  \label{tab:app:a:cmip6_emergent_constraints_part2}
 \end{table}

### 2021-02-10 ch02_scientific_background/scientific_background.tex
Added final table to chapter 05

@@
 immediately impacts the global carbon cycle by inserting additional carbon into
-the system (for details see \cref{sec:02:carbon_cycle}). Further examples
+the system (see \cref{sec:02:carbon_cycle} for details). Further examples
 include land use changes like the deforestation of tropical rainforests, which
@@
 $\hat{b}_0$ and slope $\hat{b}_1$ is used (see \cref{eq:02:linear_coefficients}
-for details on this). Fitting this regression line with ordinary least squares
-includes minimizing the standard error $s$ of the estimate
+for details). Fitting this regression line with ordinary least squares includes
+minimizing the standard error $s$ of the estimate
 \begin{equation}

### 2021-02-10 ch05_paper_ecs/paper_ecs.tex
Added final table to chapter 05

@@
 
-This is the \ac{LTMI}!!
+I like \ac{LTMI}!!
 
-\begin{center}
-  \newcommand{\testiii}{YOTHISISATEST!}
-  hi
-  \testiii{}
-\end{center}
+\begin{table}[t]
+  \centering
+  \begin{tabular}{p{0.06\columnwidth} p{0.34\columnwidth} y{0.12} y{0.12}
+      y{0.1} y{0.1}}
+    \toprule
+    Label & \ecreshead{\acs{ECS}}{(original publication)} &
+    \ecreshead{\acs{ECS} [$\si{K}$]}{(\acs{CMIP}5)} & \ecreshead{\acs{ECS}
+      [$\si{K}$]}{(\acs{CMIP}6)} & \ecreshead{$p$}{(\acs{CMIP}5)} &
+    \ecreshead{$p$}{(\acs{CMIP}6)} \\
+    \midrule
+    BRI & most likely $4.0 \unit{K}$, $< 2.30 \unit{K}$ very unlikely ($90
+    \unit{\%}$ confidence) & 3.72
+    \pm 0.59 & 4.32 \pm 1.07 & 0.0005 & 0.0355 \\
+    COX & $\pmrangeunit{2.8}{0.6}{K}$ & 3.03 \pm 0.73 & 3.71 \pm 1.09 & 0.0032
+    & 0.5415 \\
+    LIP & no best estimate given & 2.97 \pm 0.75 & 3.75 \pm 1.11 & 0.0228 &
+    0.6791 \\
+    SHD & none --- see SHL & 3.65 \pm 0.64 & 3.77 \pm 1.06 & 0.0037 & 0.2805 \\
+    SHL & most likely $4 \unit{K}$, lower limit $3 \unit{K}$ & 3.42 \pm 0.65 &
+    3.67 \pm 1.06 & 0.0002 & 0.0138 \\
+    SHS & none --- see SHL & 3.07 \pm 0.73 & 3.48 \pm 1.07 & 0.0647 & 0.0396 \\
+    SU & most likely $4 \unit{K}$, lower limit $3 \unit{K}$ & 3.30 \pm 0.88 &
+    3.77 \pm 1.35 & 0.1676 & 0.1935 \\
+    TIH & most likely $4.0 \unit{K}$ & 3.88 \pm 0.75 & 4.15 \pm 1.10 & 0.0089
+    & 0.1348 \\
+    TII & most likely $4.0 \unit{K}$ & 3.87 \pm 0.67 & 3.84 \pm 1.09 & 0.0013
+    & 0.8236 \\
+    VOL & $\pmrangeunit{3.6}{0.4}{K}$ (\stddev{}) & 3.74 \pm 0.64 & 4.21 \pm
+    1.04 & 0.0004 & 0.0056 \\
+    ZHA & $\pmrangeunit{3.90}{0.45}{K}$ (\stddev{}) & 3.35 \pm 0.74 & 3.79 \pm
+    0.67 & 0.2567 & < 0.0001 \\
+    \bottomrule
+  \end{tabular}
+  \caption{Overview of the constrained \acf{ECS} ranges and $p$-values for all
+    eleven analyzed emergent constraints. If not further specified, the
+    uncertainty ranges correspond to the $66 \unit{\%}$ confidence intervals
+    ($\range{17}{83} \unit{\%}$). For \acs{CMIP}5 and \acs{CMIP}6, these are
+    evaluated from the \acl{PDF} given by \cref{eq:02:pdf_y_given_x0} (see also
+    the right columns of \cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,%
+      fig:05:su_tih_tii,fig:05:vol_zha}). Note that even though \acs{CMIP}5
+    models were used for some constraints in the original publications, the
+    constrained ranges in the second and third column might differ due to the
+    use of a different subset of climate models (in this study, we use output
+    from all \acs{CMIP} models that is publicly available; see
+    \cref{tab:app:a:cmip5_emergent_constraints_part1,%
+      tab:app:a:cmip5_emergent_constraints_part2,%
+      tab:app:a:cmip6_emergent_constraints_part1,%
+      tab:app:a:cmip6_emergent_constraints_part2} for details). The $p$-values
+    describing the significance of the emergent relationships are defined as
+    the probability to obtain an absolute correlation coefficient $\abs*{r}$ or
+    higher under the null hypothesis that the true underlying correlation
+    coefficient between the predictor and \acs{ECS} is zero. Smaller $p$-values
+    point to higher significance and vice versa (see
+    \cref{subsec:02:emergent_constraints} for details).
+    \AdaptedFrom{Schlund2020a}.}
+  \label{tab:05:overview_results}
+\end{table}
 

### 2021-02-10 ch02_scientific_background/scientific_background.tex
Added first parts of section 5.2

@@
 balanced with an outgoing thermal irradiance $J_0$ that solely depends on the
-global mean surface temperature $T_0$ following the Stefan–Boltzmann law
+global mean surface temperature $T_0$ following the Stefan-Boltzmann law
 \begin{equation}
@@
 \end{equation}
-$\sigma \approx 5.67 \unit{W m^{-2} K^{-4}}$ is the Stefan–Boltzmann constant.
+$\sigma \approx 5.67 \unit{W m^{-2} K^{-4}}$ is the Stefan-Boltzmann constant.
 To answer an external forcing $F$, the climate system reacts with a response
@@
 include Bayesian frameworks \autocite{Renoult2020}, information theoretic
-approaches based on the Kullback–Leibler divergence between the models'
+approaches based on the Kullback-Leibler divergence between the models'
 \acp{PDF} of $x$ and the observational \ac{PDF} \autocite{Brient2016} and

### 2021-02-10 ch05_paper_ecs/paper_ecs.tex
Added first parts of section 5.2

@@
     \midrule
-    ZHA* & \textcite{Zhai2015} & Seasonal response of marine boundary layer
-    cloud fraction to changes in \acs{SST} [$\si{\pctK{}}$] &
+    ZHA* & \textcite{Zhai2015} & Seasonal response of \acl{MBLC} cloud
+    fraction to changes in \acs{SST} [$\si{\pctK{}}$] &
     \makecell{\tabitem \emph{cl} (CloudSat/CALIPSO) \\ \tabitem \emph{tos}
@@
 
-\subsection{COX}
-\label{subsec:05:cox}
+In this section we describe and discuss the eleven emergent constraints on
+\ac{ECS} summarized in \cref{tab:05:overview_emergent_constraints} using
+\acs{CMIP}5 and \acs{CMIP}6 data
+(\cref{subsec:05:bri,subsec:05:cox,subsec:05:lip,subsec:05:shd,subsec:05:shs,%
+  subsec:05:shl,subsec:05:su,subsec:05:tih,subsec:05:tii,subsec:05:vol,%
+  subsec:05:zha}) and provide a best estimate for \ac{ECS} and statistical
+significance of the eleven emergent constraints in
+\cref{subsec:05:emergent_constraints_summary}. While most of these emergent
+constraints have been derived using data from the \acs{CMIP}5 and/or
+\acs{CMIP}3 ensembles, to our knowledge none of them has been evaluated on
+the \acs{CMIP}6 ensemble so far. The results for the individual emergent
+constraints described in the following are shown in
+\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}.
+The left columns in these figures show the emergent relationships, including
+the uncertainty of the linear regressions (blue and orange shaded areas; see
+\cref{eq:02:spe}) and the uncertainty in the observations (gray shaded area;
+see \cref{eq:02:pdf_obs}). The right columns show the \acp{PDF} of \ac{ECS} in
+the original model ensemble (histogram) and the constrained distribution given
+by the emergent constraints (blue and orange line; see
+\cref{eq:02:pdf_y_given_x0}). \Cref{tab:05:overview_results} shows the
+corresponding $66 \unit{\%}$ confidence intervals (\ie{} the $\range{17}{83}
+\unit{\%}$ intervals) of \ac{ECS} derived from the \acp{PDF} given by
+\cref{eq:02:pdf_y_given_x0} and the $p$-values used to assess the significance
+of the emergent relationships.
+
+\subsection{Sensitivity of Shortwave Cloud Albedo to changes in \acl{SST}
+  (BRI)}
+\label{subsec:05:bri}
+
+In this emergent constraint proposed by \textcite{Brient2016}, \ac{ECS} is
+correlated with the \ac{TLC} albedo, \ie{} using the covariance of clouds with
+changes in \acp{SST}. Differences in the \ac{TLC} albedo account for more than
+half of the variance of \ac{ECS} in the \acs{CMIP}5 ensemble. Following
+\textcite{Brient2016}, \ac{TLC} regions are defined as grid points that are in
+the driest quartile of $500 \unit{hPa}$ relative humidity of all grid cells
+over the ocean between $30 \unit{\degree S}$ and $30 \unit{\degree N}$. The
+\ac{TLC} albedo is obtained by calculating the ratio of \ac{TOA} shortwave
+cloud radiative forcing and solar insolation averaged over the \ac{TLC} region.
+The regression coefficients of de-seasonalized variations of \ac{TLC} shortwave
+albedo and \ac{SST} (in $\si{\pctK{}}$) are then used as an emergent constraint
+for \ac{ECS}. Here, we use observational data from HadISST for \ac{SST}
+\autocite{Rayner2003}, ERA-Interim for $500 \unit{hPa}$ relative humidity
+\autocite{Dee2011} and CERES-EBAF \autocite{Loeb2018} for the \ac{TOA}
+radiative fluxes over the time period \range{2001}{2005}. In the original
+publication, \textcite{Brient2016} use similar observation-based datasets with
+the exception of \ac{SST}, where they take ERSST data \autocite{Smith2003} as
+reference instead. Our analysis yields a $66 \unit{\%}$ confidence range for
+\ac{ECS} of $\pmrangeunit{3.72}{0.59}{K}$ for \acs{CMIP}5 ($R^2 = 0.38$) and
+$\pmrangeunit{4.32}{1.07}{K}$ for \acs{CMIP}6, with much lower $R^2 = 0.12$.
+The original publication states a best estimate of $4.0 \unit{K}$ with a very
+low likelihood of values below $2.3 \unit{K}$ ($90 \unit{\%}$ confidence). The
+statistical significance of the emergent relationship drops from $p = 0.0005$
+for \acs{CMIP}5 to $p = 0.0355$ for \acs{CMIP}6.
 
@@
 
+\subsection{Temperature Variability Metric (COX)}
+\label{subsec:05:cox}
+
+The emergent constraint on \ac{ECS} proposed by \textcite{Cox2018} uses a
+temperature variability metric $\Psi$ that is based on the interannual
+variation of \ac{GSAT} calculated from its variance (in time) and $1$-year lag
+autocorrelation. In contrast to the majority of emergent constraints that focus
+on cloud-related processes, this constraint is based on the
+fluctuation-dissipation theorem, which relates the long-term response of the
+climate system to an external forcing (\ac{ECS}) and short-term variations of
+the climate system (climate variability). This arguably places the constraint
+on a more solid theoretical foundation, although several questions have been
+raised on the robustness of the results to choices made in the analysis
+\autocite{Brown2018, PoChedley2018, Rypdal2018}. For example,
+\textcite{Annan2020} has shown that the assumed linear relationship between
+$\Psi$ and \ac{ECS} does not hold when adding a deep ocean to the model. As
+observational data, here we use the HadCRUT4 dataset \autocite{Morice2012} over
+the time period \range{1880}{2014}. Under the COX constraint we assess a $66
+\unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.03}{0.73}{K}$ for \acs{CMIP}5
+($R^2 = 0.31$) and $\pmrangeunit{3.71}{1.09}{K}$ for \acs{CMIP}6 ($R^2 =
+0.01$). \textcite{Cox2018} derive a $66 \unit{\%}$ range of
+$\pmrangeunit{2.8}{0.6}{K}$ from a different subset of \acs{CMIP}5 models but
+the same observations. When moving from \acs{CMIP}5 to \acs{CMIP}6, the
+significance of the emergent relation drops massively from $p = 0.0032$ to $p =
+0.5415$, respectively.
+
+\subsection{Extent of Southern Hemisphere Hadley Cell (LIP)}
+\label{subsec:05:lip}
+
+The results of \textcite{Lipat2017} show that the multi-year average extent of
+the Hadley cell correlates with \acs{ECS} in \acs{CMIP}5 models. The Hadley
+cell edge is defined as the latitude of the first two grid cells from the
+Equator going south where the zonal average $500 \unit{hPa}$ mass stream
+function calculated from \ac{DJF} means of the meridional wind field changes
+sign from negative to positive. \textcite{Lipat2017} explain this correlation
+by tying it to the observed correlation of the interannual variability in
+midlatitudinal clouds and their radiative effects with the poleward extent of
+the Hadley cell. For the calculation of the emergent constraint, we use
+reanalysis data from ERA-Interim \autocite{Dee2011} for the meridional wind
+speed over the time period \range{1980}{2005}. Our application of this emergent
+constraint gives \ac{ECS} $66 \unit{\%}$ ranges of
+$\pmrangeunit{2.97}{0.75}{K}$ for \acs{CMIP}5 ($R^2 = 0.18$) and
+$\pmrangeunit{3.75}{1.11}{K}$ for \acs{CMIP}6 ($R^2 < 0.01$). The original
+publication does not specify an \ac{ECS} range. For \acs{CMIP}6, the emergent
+constraint shows a much lower statistical significance ($p = 0.6791$) than for
+\acs{CMIP}5 ($p = 0.0228$).
+
+\subsection{Large-Scale Lower-Tropospheric Mixing (SHD)}
+\label{subsec:05:shd}
+
+\textcite{Sherwood2014} propose that the degree of mixing in the lower
+troposphere determines the response of boundary layer clouds and humidity to
+climate warming, as the associated moisture transport would increase rapidly in
+a warmer atmosphere due to the Clausius-Clapeyron relationship. The large-scale
+component $D$ of this mixing is defined as the ratio of shallow to deep
+overturning. $D$ is calculated from the vertical velocities averaged over two
+height regions: $850$ and $700 \unit{hPa}$ for shallow overturning and $600$,
+$500$ and $400 \unit{hPa}$ for deep overturning. Both quantities are averaged
+over parts of the tropical ocean region away from the regions of highest
+\ac{SST} and strongest mid-level ascent, specifically the region $\range{30
+  \unit{\degree S}}{30 \unit{\degree N}}$, $\range{160 \unit{\degree W}}{30
+  \unit{\degree E}}$, wherever air is ascending at low levels. As
+observation-based data, we use vertical velocities from ERA-Interim
+\autocite{Dee2011} over the time period \range{1989}{1998} similar to the
+original publication. We derive \ac{ECS} $66 \unit{\%}$ confidence ranges of
+$\pmrangeunit{3.65}{0.64}{K}$ for \acs{CMIP}5 ($R^2 = 0.28$) and
+$\pmrangeunit{3.77}{1.06}{K}$ for \acs{CMIP}6 ($R^2 = 0.03$).
+\textcite{Sherwood2014} do not give a best estimate for \ac{ECS} based on the
+large-scale component of mixing $D$ or its small-scale counterpart $S$ (see
+\cref{subsec:05:shs}) but instead for the sum of $D + S$ only (see
+\cref{subsec:05:shl}). The regression shows a much lower significance for
+\acs{CMIP}6 ($p = 0.2805$) than for \acs{CMIP}5 ($p = 0.0037$).
+
 \begin{figure}[p]
@@
 \end{figure}
+
+\subsection{Small-Scale Lower-Tropospheric Mixing (SHS)}
+\label{subsec:05:shs}
+
+\subsection{Lower-Tropospheric Mixing Index (SHL)}
+\label{subsec:05:shl}
+
+I like the \ac{LTMI}!!
+
+\subsection{Error in Vertical Profile of Relative Humidity (SU)}
+\label{subsec:05:su}
+
+\subsection{Tropical Mid-Tropospheric Humidity Asymmetry Index (TIH)}
+\label{subsec:05:tih}
+
+\subsection{Southern \acs{ITCZ} index from (TII)}
+\label{subsec:05:tii}
+
+I like the \ac{ITCZ}!!
+
+\subsection{Difference Between Tropical and Southern Midlatitudinal Cloud
+  Fraction (VOL)}
+\label{subsec:05:vol}
+
+\subsection{Response of Marine Boundary Layer Cloud Fraction to changes in
+  \acl{SST} (ZHA)}
+\label{subsec:05:zha}
+
+I like the \ac{MBLC}!!
+
+\subsection{Constrained \acs{ECS} ranges and statistical significance of the
+  eleven emergent constraints}
+\label{subsec:05:emergent_constraints_summary}
 
@@
 
-\subsection{SHL}
-\label{subsec:05:shl}
-
-I like \ac{LTMI}!!
-
 \begin{table}[t]

### 2021-02-11 ch03_esmvaltool/esmvaltool.tex
Added final table to chapter 05
Finished section 5.2

@@
 \cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs,ch:06:paper_gpp} of this
-thesis were implemented into the \ac{ESMValTool}. Apart from that, further
+thesis have been implemented into the \ac{ESMValTool}. Apart from that, further
 substantial changes and additions to the code base of the \ac{ESMValTool} have
@@
     \toprule
-    Type & \textnormal{Name} & Description \\
+    Type & \ignorecolumntype{l}{Name} & Description \\
     \midrule
@@
     indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}
-      K^{-1}}$). Both panels were created with the \acs{ESMValTool} using
+      K^{-1}}$). Both panels have been created with the \acs{ESMValTool} using
     \emph{recipe\_cox18nature.yml} and are similar to \textcite{Cox2018}. (a)
@@
     relationship between \acs{ECS} and the temperature variability metric
-    $\Psi$. The black dot-dashed line shows the linear regression across the
+    $\Psi$. The dot-dashed black line shows the linear regression across the
     climate model ensemble with its associated uncertainties indicated by the
-    black dashed lines (standard prediction error; see \cref{eq:02:spe}). Blue
+    dashed black lines (standard prediction error; see \cref{eq:02:spe}). Blue
     lines show the observational constraint from HadCRUT4 with its best
@@
 exhibit a high equilibrium warming in the future or vice versa. On the other
-hand, \ac{TCR} and \ac{ECS} are in fact well-correlated within the \acs{CMIP}5
+hand, \ac{TCR} and \ac{ECS} are in fact well correlated within the \acs{CMIP}5
 ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this
@@
   \caption{Relationships between climate metrics within the \acs{CMIP}5 climate
-    model ensemble. Both panels were created with the \acs{ESMValTool} using
-    \emph{recipe\_flato13ipcc.yml} and are similar to \textcite{Flato2013}. (a)
-    \Acf{GSAT} versus the \acf{ECS} for \acs{CMIP}5 models for the period
-    \range{1961}{1990} (larger symbols) and for pre-industrial control runs
-    (smaller symbols). (b) \acf{TCR} versus \acs{ECS} for \acs{CMIP}5 models.
-    The black line shows a linear fit between \acs{TCR} and \acs{ECS}.}
+    model ensemble. Both panels have been created with the \acs{ESMValTool}
+    using \emph{recipe\_flato13ipcc.yml} and are similar to
+    \textcite{Flato2013}. (a) \Acf{GSAT} versus the \acf{ECS} for \acs{CMIP}5
+    models for the period \range{1961}{1990} (larger symbols) and for
+    pre-industrial control runs (smaller symbols). (b) \acf{TCR} versus
+    \acs{ECS} for \acs{CMIP}5 models. The black line shows a linear fit
+    between \acs{TCR} and \acs{ECS}.}
   \label{fig:03:flato13}

### 2021-02-11 ch05_paper_ecs/paper_ecs.tex
Finished section 5.2

@@
     \midrule
-    TII & \textcite{Tian2015} & Southern ITCZ index from GPCP
+    TII & \textcite{Tian2015} & Southern \acs{ITCZ} index from GPCP
     [$\si{\mmday{}}$] & \tabitem \emph{pr} (GPCP) \\
@@
     fraction to changes in \acs{SST} [$\si{\pctK{}}$] &
-    \makecell{\tabitem \emph{cl} (CloudSat/CALIPSO) \\ \tabitem \emph{tos}
-      (AMSRE SST) \\ \tabitem \emph{wap} (ERA-Interim)} \\
+    \makecell{\tabitem \emph{cl} (Cloudsat/CALIPSO) \\ \tabitem \emph{tos}
+      (AMSRE \ac{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
     \bottomrule
@@
 and \cref{sec:05:comparison_of_emergent_constraints}. We chose these particular
-emergent constraints since these were already implemented in the
+emergent constraints since these had already been implemented in the
 \ac{ESMValTool} at the time of writing this study, which greatly facilitated
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_BRI.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_bri.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_BRI.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_bri.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_COX.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_cox.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_COX.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_cox.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_LIP.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_lip.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_LIP.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_lip.pdf}
     \caption{}
@@
     horizontal dashed lines show the best estimates of the constrained
-    \acf{ECS} and the colored points mark the \aclp{MMM} for \acs{CMIP}5
+    \acf{ECS} and the colored circles mark the \aclp{MMM} for \acs{CMIP}5
     (blue) and \acs{CMIP}6 (orange). (b), (d), (f) \Aclp{PDF} for \acs{ECS} of
@@
 
-\begin{figure}[p]
+\begin{figure}[!t]
   \centering
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_SHD.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_shd.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_SHD.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_shd.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_SHL.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_shl.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_SHL.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_shl.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_SHS.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_shs.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_SHS.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_shs.pdf}
     \caption{}
@@
 
+The small-scale mixing $S$ \autocite{Sherwood2014} is calculated from the
+differences in relative humidity and temperature between $700$ and $850
+\unit{hPa}$. The differences are averaged over all grid cells within the upper
+quartile of the annual mean $500 \unit{hPa}$ ascent rate (within ascending
+regions) in the tropics. The tropics are defined as the region between $30
+\unit{\degree S}$ and $30 \unit{\degree N}$. In the \ac{CFMIP} models
+\autocite{Webb2017}, for which convective tendencies are available, upward
+moisture transport by parameterized convection is shown to increase more
+rapidly with warming for higher values of $S$. We use reanalysis data from
+ERA-Interim \autocite{Dee2011} for temperature and relative humidity to
+calculate the observational constraint (\range{1989}{1998}). Our analysis shows
+a $66 \unit{\%}$ range of \ac{ECS} of $\pmrangeunit{3.07}{0.73}{K}$ for
+\acs{CMIP}5 ($R^2 = 0.13$) and $\pmrangeunit{3.48}{1.07}{K}$ for \acs{CMIP}6
+($R^2 = 0.12$). The correlation of $S$ and \acs{ECS} shows a slightly higher
+significance in the \acs{CMIP}6 ensemble ($p = 0.0396$) than in the \acs{CMIP}5
+ensemble ($p = 0.0647$). The SHS constraint is one of the two analyzed emergent
+constraints (ZHA being the other exception) that shows a higher statistical
+significance for the \acs{CMIP}6 than for the \acs{CMIP}5 ensemble.
+
 \subsection{Lower-Tropospheric Mixing Index (SHL)}
@@
 
-I like the \ac{LTMI}!!
+The \ac{LTMI} formulated by \textcite{Sherwood2014} is defined as the sum of
+the small-scale mixing $S$ (see \cref{subsec:05:shs}) and the large-scale
+mixing $D$ (see \cref{subsec:05:shd}), which are supposed to capture
+complementary components of the total mixing phenomenon.
+\textcite{Sherwood2014} argue that the increase in dehydration depends on
+initial mixing linking it to cloud feedbacks and thus also to \ac{ECS}. For
+this constraint, we derive an \ac{ECS} $66 \unit{\%}$ confidence range of
+$\pmrangeunit{3.42}{0.65}{K}$ for \acs{CMIP}5 ($R^2 = 0.41$) and
+$\pmrangeunit{3.67}{1.06}{K}$ for \acs{CMIP}6 ($R^2 = 0.16$).
+\textcite{Sherwood2014} give a best estimate of about $4 \unit{K}$ with a lower
+limit of $3 \unit{K}$. Similar to both other constraints by
+\textcite{Sherwood2014}, SHD and SHS, the statistical significance of the SHL
+emergent relation decreases in \acs{CMIP}6 ($p = 0.0138$) compared to
+\acs{CMIP}5 ($p = 0.0002$).
 
@@
 
-\subsection{Tropical Mid-Tropospheric Humidity Asymmetry Index (TIH)}
-\label{subsec:05:tih}
-
-\subsection{Southern \acs{ITCZ} index from (TII)}
-\label{subsec:05:tii}
-
-I like the \ac{ITCZ}!!
-
-\subsection{Difference Between Tropical and Southern Midlatitudinal Cloud
-  Fraction (VOL)}
-\label{subsec:05:vol}
-
-\subsection{Response of Marine Boundary Layer Cloud Fraction to changes in
-  \acl{SST} (ZHA)}
-\label{subsec:05:zha}
-
-I like the \ac{MBLC}!!
-
-\subsection{Constrained \acs{ECS} ranges and statistical significance of the
-  eleven emergent constraints}
-\label{subsec:05:emergent_constraints_summary}
-
-\begin{figure}[p]
+\begin{figure}[!t]
   \centering
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_SU.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_su.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_SU.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_su.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_TIH.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_tih.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_TIH.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_tih.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_TII.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_tii.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_TII.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_tii.pdf}
     \caption{}
@@
 
-\begin{figure}[p]
+Another emergent constraint on \ac{ECS} that targets uncertainties in cloud
+feedbacks is proposed by \textcite{Su2014}. They show that changes in the
+Hadley circulation are physically connected to changes in tropical clouds and
+thus \ac{ECS}. Consequently, the inter-model spread in the change of the Hadley
+circulation in an ensemble of climate models is well correlated with the
+corresponding changes in the \ac{TOA} \ac{CRE}. Moreover, \textcite{Su2014}
+find a correlation between a model's \ac{ECS} and its ability to represent the
+present-day Hadley circulation. The latter is calculated from the tropical
+($\range{45 \unit{\degree S}}{40 \unit{\degree N}}$) zonal-mean vertical
+profiles of relative humidity from the surface to $100 \unit{hPa}$. These
+profiles are then used to define the \xaxis{} of the SU constraint by
+calculating a performance metric based on the slope of the linear regression
+between a climate model's relative humidity profile and the corresponding
+observational reference. Similarly to the original publication, we use humidity
+observations from AIRS \autocite{Aumann2003} for pressure levels greater than
+$300 \unit{hPa}$ and MLS-Aura data \autocite{Beer2006} for pressure levels of
+less than $300 \unit{hPa}$. Our analysis yields a constrained $66 \unit{\%}$
+range of \ac{ECS} of $\pmrangeunit{3.30}{0.88}{K}$ for \acs{CMIP}5 ($R^2 =
+0.08$) and $\pmrangeunit{3.77}{1.35}{K}$ for \acs{CMIP}6 ($R^2 = 0.05$). The
+original publication gives a best estimate of $4 \unit{K}$ with a lower limit
+of $3 \unit{K}$. \Cref{fig:05:su_tih_tii:a} shows that in addition to the low
+$R^2$ values, the emergent relationship shows different slopes for \acs{CMIP}5
+and \acs{CMIP}6. For \acs{CMIP}5, the expected positive correlation is found,
+while for \acs{CMIP}6, a negative correlation is found. This suggests that the
+constraint is not working (anymore) when applied to the \acs{CMIP}6 data.
+Consequently, the SU constraint shows a weaker statistical significance in
+\acs{CMIP}6 ($p = 0.1935$) than in \acs{CMIP}5 ($p = 0.1676$). The SU
+constraint is related to an emergent constraint on \ac{ECS} proposed by
+\textcite{Fasullo2012}, who correlate \range{May}{August} zonal-mean relative
+humidity against \ac{ECS}. In contrast to \textcite{Su2014}, they do not use
+the entire tropics, but identify two distinct regions with the largest
+correlation.
+
+\subsection{Tropical Mid-Tropospheric Humidity Asymmetry Index (TIH)}
+\label{subsec:05:tih}
+
+\textcite{Tian2015} finds a link between mid-tropospheric humidity over the
+tropical Pacific and simulated moisture, precipitation, clouds, large-scale
+circulation, and thus \ac{ECS} in \acs{CMIP}3 and \acs{CMIP}5 models. The study
+explains this link with the similarity of mid-tropospheric humidity and
+precipitation patterns as both are related to the \ac{ITCZ}. The proposed
+tropical mid-tropospheric humidity asymmetry index to constrain \ac{ECS} is
+defined as relative bias (in $\si{\%}$) in simulated annual mean $500
+\unit{hPa}$ specific humidity averaged over the Southern Hemisphere tropical
+Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{80
+  \unit{\degree W}}{120 \unit{\degree E}}$) minus the bias averaged over the
+Northern Hemisphere tropical Pacific ($\range{0 \unit{\degree N}}{20
+  \unit{\degree N}}$, $\range{80 \unit{\degree W}}{120 \unit{\degree E}}$) when
+compared with observations. Similar to the SU constraint, the index proposed by
+\textcite{Tian2015} seems to be related to the emergent constraint by
+\textcite{Fasullo2012}, who find correlations between relative humidity of the
+middle and upper troposphere and \ac{ECS}. Here, we use humidity observations
+from AIRS \autocite{Aumann2003} over the time period \range{2003}{2005} as the
+reference dataset. We assess a $66 \unit{\%}$ \ac{ECS} range of
+$\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5 ($R^2 = 0.24$) and
+$\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 = 0.06$).
+\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
+significance of the emergent relationship drops massively from $p = 0.0089$ in
+\acs{CMIP}5 to $p = 0.1348$ in \acs{CMIP}6.
+
+\subsection{Southern \acs{ITCZ} index from (TII)}
+\label{subsec:05:tii}
+
+In addition to the humidity index, \textcite{Tian2015} proposes an emergent
+constraint on \ac{ECS} based on the southern \ac{ITCZ} index
+\autocite{Bellucci2010, Hirota2011}. This index is defined as the
+climatological annual mean precipitation bias averaged over the southeastern
+Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{150
+  \unit{\degree W}}{100 \unit{\degree W}}$). The southern \ac{ITCZ} index is
+calculated in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ},
+a common problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds
+a link between double-\ac{ITCZ} bias and simulated moisture, precipitation,
+clouds, and large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He
+argues that this could explain the link between the double-\acs{ITCZ} bias and
+\ac{ECS}. As reference data, we use observed precipitation data for the years
+\range{1986}{2005} from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66
+\unit{\%}$ confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5
+($R^2 = 0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 <
+0.01$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
+emergent relationship shows a much lower statistical significance in
+\acs{CMIP}6 ($p = 0.8236$) than in \acs{CMIP}5 ($p = 0.0013$).
+
+\subsection{Difference Between Tropical and Southern Midlatitudinal Cloud
+  Fraction (VOL)}
+\label{subsec:05:vol}
+
+The study by \textcite{Volodin2008} aims at constraining \ac{ECS} based on the
+geographical distribution of clouds in climate models. Since this early
+emergent constraint has originally been trained on \acs{CMIP}3 models, both
+\acs{CMIP}5 and \acs{CMIP}6 are out-of-sample tests for it.
+\textcite{Volodin2008} shows that high \ac{ECS} models tend to simulate a
+higher total cloud cover over the southern midlatitudes and a lower total cloud
+cover over the tropics (relative to the \ac{MMM}). This can be used to
+establish an emergent relationship between the \ac{ECS} and the difference in
+tropical total cloud cover ($\range{28 \unit{\degree S}}{28 \unit{\degree N}}$)
+and the southern midlatitudinal total cloud cover ($\range{56 \unit{\degree
+    S}}{36 \unit{\degree S}}$). Analogous to the original study, we use the
+ISCCP-D2 data \autocite{Rossow1991} as observational reference. For the VOL
+constraint, we calculate a constrained $66 \unit{\%}$ range of \ac{ECS} of
+$\pmrangeunit{3.74}{0.64}{K}$ for \acs{CMIP}5 ($R^2 = 0.38$) and
+$\pmrangeunit{4.21}{1.04}{K}$ for \acs{CMIP}6 ($R^2 = 0.18$), whereas the
+original publication gives a range of $\pmrangeunit{3.6}{0.4}{K}$ (standard
+deviation) for a climate model ensemble of \acs{CMIP}3 models. The emergent
+constraint by \textcite{Volodin2008} shows a lower significance in the
+\acs{CMIP}6 ensemble ($p = 0.0056$) than in the \acs{CMIP}5 ensemble ($p =
+0.0004$).
+
+\begin{figure}[t]
   \centering
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_VOL.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_vol.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_VOL.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_vol.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/scatterplot_merged_training_data_ZHA.pdf}
+      ch05_paper_ecs/figs/scatterplot_merged_training_data_zha.pdf}
     \caption{}
@@
     \includegraphics[width=\columnwidth]{
-      ch05_paper_ecs/figs/target_distribution_training_data_ZHA.pdf}
+      ch05_paper_ecs/figs/target_distribution_training_data_zha.pdf}
     \caption{}
@@
 
-\begin{table}[t]
+\subsection{Response of Marine Boundary Layer Cloud Fraction to changes in
+  \acl{SST} (ZHA)}
+\label{subsec:05:zha}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{ch05_paper_ecs/figs/zha.pdf}
+  \caption{Emergent relationship ZHA \autocite{Zhai2015} for different subsets
+    of \acs{CMIP}5 models. Blue circles show the 15 \acs{CMIP}5 models used in
+    the original publication (except for CESM1-CAM5). The solid blue line and
+    blue shaded area show the emergent relationships evaluated on these models
+    including the uncertainty range. In our study, we added eleven more
+    \acs{CMIP}5 models (red circles). The corresponding emergent relationship
+    that considers all available \acs{CMIP}5 models is shown in red colors.
+    This relationship shows a considerably lower coefficient of determination
+    ($R^2$) and higher $p$-value than the relationship using the original
+    subset of \acs{CMIP}5 models. The vertical dashed line and shaded area
+    correspond to the observational reference and the horizontal dashed lines
+    show the corresponding constraints on the \acf{ECS} using this observation.
+    \AdaptedFrom{Schlund2020a}.}
+  \label{fig:05:zha}
+\end{figure}
+
+\textcite{Zhai2015} focus on the variations of \acp{MBLC}, which largely
+contribute to the shortwave cloud feedback and thus to the uncertainty in
+modeled \ac{ECS}. Their central quantity is the response of the \ac{MBLC}
+fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
+for both hemispheres ($\range{20 \unit{\degree}}{40 \unit{\degree}}$). On short
+(seasonal) and long (centennial under a forcing) timescales, this quantity is
+well correlated with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5
+models. Together with observations of cloud fraction from Cloudsat/CALIPSO
+\autocite{Mace2009}, \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011}, and
+vertical velocity from ERA-Interim \autocite{Dee2011}, the seasonal response of
+\ac{MBLC} fraction to changes in \ac{SST} forms an emergent constraint on
+\ac{ECS}. We assess a $66 \unit{\%}$ \ac{ECS} range of
+$\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5 ($R^2 = 0.05$) and
+$\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 = 0.62$). In their original
+publication, \textcite{Zhai2015} find an \ac{ECS} range of
+$\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination of
+\acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance, the
+results of the ZHA constraints are somewhat surprising: although \acs{CMIP}5
+data (in combination with \acs{CMIP}3 data) has successfully been used in the
+original publication, our approach finds that the statistical significance of
+the emergent relationship is much higher in the unseen \acs{CMIP}6 ensemble ($p
+< 0.0001$) than in the previously available \acs{CMIP}5 ensemble ($p =
+0.2567$). The ZHA constraint is the only emergent constraint analyzed here that
+shows this extreme behavior (only one other constraint, SHS, shows a slightly
+higher significance in \acs{CMIP}6; all other constraints show lower
+significances in \acs{CMIP}6). The reason for the erratic skill in \acs{CMIP}5
+is the set of climate models used. For our analysis, we use eleven additional
+\acs{CMIP}5 models that have not been used in the original publication
+(ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m, CCSM4, GFDL-ESM2G, GFDL-ESM2M,
+IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and MPI-ESM-P). Due to a lack of
+publicly available data, the model CESM1-CAM5 that is used in the original
+publication is not included in our analysis. The effect of choosing different
+subsets of \acs{CMIP}5 models on the emergent relationship is illustrated in
+\cref{fig:05:zha}. Using the original \acs{CMIP}5 models from the original
+publication gives a considerably higher correlation ($R^2 = 0.38$) than using
+all available \acs{CMIP}5 models ($R^2 = 0.05$). This result shows a strong
+dependency of this emergent constraint on the subset of climate models used.
+Nonetheless, the performance on \acs{CMIP}6 models is, surprisingly, the best
+of all the constraints, and much better than on either subset of \acs{CMIP}5
+models.
+
+\subsection{Constrained \acs{ECS} ranges and statistical significance of the
+  eleven emergent constraints}
+\label{subsec:05:emergent_constraints_summary}
+
+\begin{table}[!b]
   \centering
@@
       fig:05:su_tih_tii,fig:05:vol_zha}). Note that even though \acs{CMIP}5
-    models were used for some constraints in the original publications, the
-    constrained ranges in the second and third column might differ due to the
-    use of a different subset of climate models (in this study, we use output
-    from all \acs{CMIP} models that is publicly available; see
+    models have been used for some constraints in the original publications,
+    the constrained ranges in the second and third column might differ due to
+    the use of a different subset of climate models (in this study, we use
+    output from all \acs{CMIP} models that is publicly available; see
     \cref{tab:app:a:cmip5_emergent_constraints_part1,%
@@
     describing the significance of the emergent relationships are defined as
-    the probability to obtain an absolute correlation coefficient $\abs*{r}$ or
-    higher under the null hypothesis that the true underlying correlation
-    coefficient between the predictor and \acs{ECS} is zero. Smaller $p$-values
-    point to higher significance and vice versa (see
+    the probability to obtain an absolute correlation coefficient $\abs*{r}$
+    or higher under the null hypothesis that the true underlying correlation
+    coefficient between the predictor and \acs{ECS} is zero. Smaller
+    $p$-values point to higher significance and vice versa (see
     \cref{subsec:02:emergent_constraints} for details).
@@
 \end{table}
+
+In most cases, the emergent relationships (left columns of
+\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha})
+show the same sign of the slope (as expected from the theory) for \acs{CMIP}5
+and \acs{CMIP}6, with the SU constraint being the only exception. However, the
+coefficient of determination ($R^2$) is lower for \acs{CMIP}6 compared to
+\acs{CMIP}5 for all one constraint: ZHA. The probability distributions of the
+constrained \ac{ECS} that we obtain (right columns of
+\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha})
+give similar results: except for the ZHA constraint, the constraint on the
+\acs{CMIP}6 ensemble is weaker, \ie{} the constrained \acp{PDF} derived from
+the \acs{CMIP}6 ensemble are broader than their respective \acs{CMIP}5
+counterparts. As shown in \cref{tab:05:overview_results} and
+\cref{fig:05:overview_results}, for \acs{CMIP}5, the range of the best (maximum
+likelihood) estimates for \acs{ECS} is $2.97 \unit{K}$ to $3.88 \unit{K}$,
+while the corresponding \acs{CMIP}6 best estimates are higher for almost every
+tested emergent constraint (TII being the only exception), resulting in a range
+of best estimates of $3.48 \unit{K}$ to $4.32 \unit{K}$. Using the arithmetic
+mean of all analyzed emergent constraints, this results in a mean increase of
+the \ac{ECS} best estimate of $12 \unit{\%}$ in \acs{CMIP}6 compared to
+\acs{CMIP}5. Similarly, the size of the $66 \unit{\%}$ \ac{ECS} ranges
+($\range{17}{83} \unit{\%}$ confidence) shows values of $1.16 \unit{K}$ to
+$1.75 \unit{K}$ in \acs{CMIP}5 and $1.32 \unit{K}$ to $2.70 \unit{K}$ in
+\acs{CMIP}6, resulting in an increase of $51 \unit{\%}$ averaged over all
+emergent constraints.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch05_paper_ecs/figs/constrained_ranges_summary.pdf}
+  \caption{Overview of the constrained \acf{ECS} ranges given by the eleven
+    analyzed emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6
+    (orange) ensemble. The bars correspond to the $66 \unit{\%}$ confidence
+    intervals ($\range{17}{83} \unit{\%}$) calculated from the constrained
+    \acl{PDF} of \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical
+    black lines illustrate the corresponding best estimates of the
+    distributions. For almost all emergent constraints, the constrained best
+    estimate and range of \ac{ECS} is higher in the \acs{CMIP}6 ensemble.}
+  \label{fig:05:overview_results}
+\end{figure}
+
+In summary, the $R^2$ of the emergent relationships and the constrained range
+of \ac{ECS} each depend strongly on the climate model ensemble used, even
+though a physical explanation is given for each emergent constraint that is
+thought to be valid for every climate model ensemble. The same behavior is
+found for the statistical significance of the emergent relationships using the
+null hypothesis that there is no correlation between the predictor and
+\ac{ECS}. Except for the ZHA and SHS constraints, every emergent relationship
+investigated shows a lower statistical significance (\ie{} a higher $p$- value)
+in the \acs{CMIP}6 ensemble than in the \acs{CMIP}5 ensemble. If a conventional
+significance test of $p < 0.05$ was applied, eight of the eleven constraints
+would pass this test on \acs{CMIP}5 model data but only five (BRI, SHL, SHS,
+VOL, and ZHA) would pass this test on \acs{CMIP}6. This is still much better
+than would be expected purely by chance. Hence, there is still skill in at
+least a few of the constraints, but it is much lower than was suggested in
+nearly all of the initial studies.
 

### 2021-02-11 end_appendix/appendix.tex
Finished section 5.2

@@
     AIRS & \emph{hur}, \emph{hus} & \textcite{Aumann2003} \\
-    AMSRE SST & \emph{tos} & \textcite{AMSRE2011} \\
+    AMSRE \ac{SST} & \emph{tos} & \textcite{AMSRE2011} \\
     CERES-EBAF & \emph{rsdt}, \emph{rsut}, \emph{rsutcs} & \textcite{Loeb2018}
     \\
-    CloudSat/CALIPSO & \emph{cl} & \textcite{Mace2009} \\
+    Cloudsat/CALIPSO & \emph{cl} & \textcite{Mace2009} \\
     ERA-Interim & \emph{hur}, \emph{ta}, \emph{va}, \emph{wap} &

### 2021-02-11 ch03_esmvaltool/esmvaltool.tex
Finished chapter 05

@@
   \caption{Emergent constraint on the \acf{ECS} from global temperature
-    variability \autocite{Cox2018}. Magenta colors indicate high sensitivity
+    variability \autocite{Cox2018}. Magenta colors indicate high-sensitivity
     climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$) and green colors
@@
 years for \acs{CMIP}5 models and observations form HadCRUT4
-\autocite{Morice2012}. Since high sensitivity climate models exhibit higher
+\autocite{Morice2012}. Since high-sensitivity climate models exhibit higher
 values of $\Psi$ than low sensitivity models, an emergent relationship between

### 2021-02-11 ch05_paper_ecs/paper_ecs.tex
Finished chapter 05

@@
 for both hemispheres ($\range{20 \unit{\degree}}{40 \unit{\degree}}$). On short
-(seasonal) and long (centennial under a forcing) timescales, this quantity is
+(seasonal) and long (centennial under a forcing) time scales, this quantity is
 well correlated with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5
 models. Together with observations of cloud fraction from Cloudsat/CALIPSO
-\autocite{Mace2009}, \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011}, and
+\autocite{Mace2009}, \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and
 vertical velocity from ERA-Interim \autocite{Dee2011}, the seasonal response of
@@
 Nonetheless, the performance on \acs{CMIP}6 models is, surprisingly, the best
-of all the constraints, and much better than on either subset of \acs{CMIP}5
+of all the constraints and much better than on either subset of \acs{CMIP}5
 models.
@@
 would pass this test on \acs{CMIP}5 model data but only five (BRI, SHL, SHS,
-VOL, and ZHA) would pass this test on \acs{CMIP}6. This is still much better
+VOL and ZHA) would pass this test on \acs{CMIP}6. This is still much better
 than would be expected purely by chance. Hence, there is still skill in at
-least a few of the constraints, but it is much lower than was suggested in
-nearly all of the initial studies.
+least a few of the constraints, but it is much lower than suggested in nearly
+all of the initial studies.
 
@@
 
+As shown in the previous sections, most emergent relationships show smaller
+coefficients of determination when evaluated on the new \acs{CMIP}6 ensemble
+compared to the \acs{CMIP}5 ensemble. In this section, we discuss possible
+reasons for, and implications of, these differences. As reported by
+\textcite{Caldwell2014}, the large amount of data provided by modern \acp{ESM}
+can generate spurious correlations of variables between past climate and
+\ac{ECS} just by chance, especially when only a small number of climate models
+is considered. This would cause the performance of the emergent constraint to
+be reduced on out-of-sample data (like the new \acs{CMIP}6 ensemble), since the
+emergent relationship appeared just by chance and not because of a physically
+based mechanism.
+
+A further reason for the weaker emergent relationships in \acs{CMIP}6 may be
+the increased complexity of the participating \acp{ESM}. Each emergent
+constraint approach is based on the assumption that a single observable process
+or physical aspect in the current climate dominates the uncertainty in
+\ac{ECS}. Some emergent constraints such as ZHA and BRI relate changes in cloud
+properties (ZHA: low-level cloud fraction; BRI: cloud reflectivity) on seasonal
+or interannual time scales driven by changes in \ac{SST} to \ac{ECS}. This
+means that it has to be implicitly assumed that the observable changes in these
+properties on seasonal or interannual time scales are related to those
+occurring as a result of climate forcing in a way successfully captured by the
+\acp{ESM}. While this assumption seems to make sense, we do not know whether
+the \acp{ESM} cover all relevant processes of the real Earth system. For
+example, it may be possible that there exist processes that are unimportant in
+the \acp{ESM} (and hence are not captured by the emergent constraints) but are
+actually important in reality. This lack of relevant processes may lead to an
+overconfident constraint. Thus, the more complex \acp{ESM} of the \acs{CMIP}6
+ensemble are more likely to capture relevant processes of the true climate,
+which leads to weaker emergent relationships. On the other hand, emergent
+constraints on the less complex \acs{CMIP}3 and \acs{CMIP}5 ensemble may be
+overconfident.
+
+For \acs{CMIP}6 models, \textcite{Zelinka2020} show that cloud feedbacks and
+thus \ac{ECS} in high-sensitivity models are to some extent associated with
+changes in clouds over the Southern Ocean, while in \acs{CMIP}3 and \acs{CMIP}5
+the uncertainty in cloud feedbacks is dominated by clouds in the subtropical
+subsidence regions. One might speculate that a possible reason for this might
+be an improved simulation of clouds over the Southern Ocean in some models
+\autocite{BodasSalcedo2019, Gettelman2019}, as shown for some pre-\acs{CMIP}6
+model versions evaluated by \textcite{Lauer2018}. The findings of
+\textcite{Zelinka2020} can also at least partly explain the larger inter-model
+spread in climate sensitivity due to a greater diversity of cloud feedbacks,
+which also results in a weaker emergent constraint compared with \acs{CMIP}5
+models, as most of them constrained low-cloud feedbacks. They find that on
+average, the shortwave low-cloud feedback is larger in \acs{CMIP}6 than in
+\acs{CMIP}5, which they primarily relate to changes in the representation of
+clouds. As a possible explanation, \textcite{Zelinka2020} give an increase in
+mean-state supercooled liquid water (\ie{} increase in the cloud water liquid
+fraction) in mixed-phase clouds resulting in less pronounced increases in
+low-level cloud cover and water content with warmer \acp{SST} particularly in
+midlatitudes.
+
+Our findings suggest that the process-oriented emergent constraints (\ie{} all
+of the emergent constraints investigated here except COX) are only successful
+in constraining \ac{ECS} as long as the uncertainty in \ac{ECS} is dominated by
+the same process or feedback. In the \acs{CMIP}5 ensemble, cloud feedback is
+the main contributor to the spread in \ac{ECS} with low-level clouds in
+tropical subsidence regions dominating the spread in cloud feedback
+\autocite{Ceppi2017}. If any other process or feedback is biased (or missing)
+in the ensemble as a whole, then these process-oriented emergent constraints
+will be biased in their estimates of \ac{ECS}. The appearance of diverse new
+feedback processes in \acs{CMIP}6 could explain the reduced skill when applied
+to \acs{CMIP}6 data, and a tendency for these to be positive would explain the
+upward shift in the model \ac{ECS} distribution that is not captured by the
+\acs{CMIP}5-trained constraints. Process-oriented emergent constraints are
+therefore perhaps best thought of as constraints on the processes that they
+target, rather than constraints on \ac{ECS}.
+
+Emergent constraints that use global temperature change as a way of
+constraining \ac{ECS} could in principle overcome this problem. If one feedback
+is biased in an ensemble the constraint might still work as both global
+temperature change and \ac{ECS} might similarly reflect the sum of all
+feedbacks. Emergent constraints of this kind include the tropical temperature
+during the Last Glacial Maximum \autocite{Hargreaves2012}, tropical temperature
+anomalies during the mid-Pliocene Warm Period \autocite{Hargreaves2016} and
+post-1970s warming \autocite{JimenezdelaCuesta2019}. This seems to be supported
+by the findings of \textcite{Tokarska2020}, who test an emergent constraint for
+the \ac{TCR} based on recent global warming trends on the \acs{CMIP}5 and
+\acs{CMIP}6 ensembles with similar results for both model ensembles.
+Nevertheless, these temperature-based estimates are sensitive to assumptions
+about forcings and unforced decadal temperature variations, which could also be
+incorrect, as could model-predicted relationships between feedbacks on short
+and long time scales that are implicit in most such measures.
+
 \section{Summary}
 \label{sec:05:summary}
+
+This paper assesses eleven different emergent constraints on \ac{ECS}, of which
+most are directly or indirectly related to cloud feedbacks, by applying them to
+results from \acp{ESM} contributing to \acs{CMIP}5 and \acs{CMIP}6. Of
+particular interest are the results from \acs{CMIP}6, since all analyzed
+emergent constraints have been published prior to the availability of
+\acs{CMIP}6 data. In summary, the best estimate of \ac{ECS} averaged over all
+emergent constraints increases by $12 \unit{\%}$ when moving from relationships
+trained on \acs{CMIP}5 to those trained on \acs{CMIP}6. Some increase is
+predicted by almost every constraint we analyze (with TII as the only
+exception) and can be at least partly explained by the increased multi-model
+mean \ac{ECS} of \acs{CMIP}6, which has not been accompanied by systematic
+changes in the constraint variables that could explain this increase, leading
+to regression fits with higher intercept values at observed constraint values.
+This is also illustrated by the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM} in the
+left columns of
+\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
+(colored circles), in which the connecting line between the \acs{CMIP}5 and
+\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
+for all emergent constraints.
+
+Our results also show that, except for ZHA, all emergent relationships are
+weaker (in terms of the coefficient of determination $R^2$) in \acs{CMIP}6
+compared to \acs{CMIP}5, which means that the corresponding emergent
+relationships are able to explain less of the \ac{ECS} variation simulated by
+the newer \acs{CMIP}6 models than by those of \acs{CMIP}5. This is also
+demonstrated by the statistical significance, which is lower in \acs{CMIP}6
+than in \acs{CMIP}5 for all emergent constraint except for ZHA and SHS. As
+described in \cref{subsec:02:emergent_constraints}, our test for statistical
+significance uses the null hypothesis that there is no correlation between the
+predictor variable of the emergent constraint and \ac{ECS}. Further evidence of
+the decreased performance of the emergent constraints in \acs{CMIP}6 is given
+by the size of the constrained \acs{ECS} ranges, which widens by $51 \unit{\%}$
+in \acs{CMIP}6 compared to \acs{CMIP}5 on average. Moreover, our more detailed
+analysis of the ZHA constraint (see \cref{fig:05:zha}) shows that this emergent
+constraint is very sensitive to outliers and the subset of the climate model
+ensemble used to fit the emergent relationship. Such a behavior might not be
+unique to the ZHA constraint but could apply to other emergent constraints as
+well. This in turn suggests that the number of climate models commonly used for
+emergent constraints might be too low, leading to non-robust relationships.
+
+Our analysis makes a number of simplifying assumptions common to other studies,
+such as model independence, discussed in detail in
+\cref{subsec:02:ecs,subsec:02:emergent_constraints}. These assumptions affect
+the significance of emergent relationships and the \acp{PDF} of \ac{ECS} based
+on a constraint. However, they do not affect our main conclusions here, which
+concern the change in performance on \acs{CMIP}6 relative to \acs{CMIP}5 and
+the implications for robustness and future use of emergent constraints.
+
+\ac{ECS} is the product of the complex interactions of the many components and
+feedbacks. Thus, constraining \ac{ECS} with a single physical process might
+overly simplify this problem. Such single process-oriented emergent constraints
+therefore do not seem to be helpful in constraining \ac{ECS} but should
+probably rather be thought of as constraints for the process or feedback they
+are actually targeting (if that can be clearly identified). With increasing
+computational resources available to climate science, more and more detailed
+process interactions can be taken into account in a modern \ac{ESM}. In
+contrast, the predecessor versions \acs{CMIP}3 and \acs{CMIP}5 are less complex
+with simpler atmospheric process representations, so constraining uncertainties
+of a single dominant process may allow for an apparently more successful
+constraint of \ac{ECS} than would be achieved in more complex models. As a
+conclusion, we argue that to constrain \ac{ECS} in a more robust way, it might
+be beneficial to apply multivariate approaches that are able to consider
+multiple (different) relevant physical processes and feedbacks at once and thus
+are able to get a broader picture of the complex reality. A possible approach
+for this is given by \textcite{Bretherton2020}, who combine the information
+from multiple emergent constraints on \ac{ECS} using a multivariate Gaussian
+and multiple linear regression. For the \acs{CMIP}3 and \acs{CMIP}5 ensembles,
+they find an increased best estimate relative to the unweighted ensemble mean
+similar to the participating individual emergent constraints, but with lower
+uncertainty range. Moreover, new machine learning techniques are a promising
+avenue forward for such multivariate approaches and for constraining
+uncertainties in multi-model projections \commentcite{Schlund2020}{see
+  \cref{ch:06:paper_gpp}} with the aim of further improving climate modeling
+and analysis \autocite{Reichstein2019}.

### 2021-02-12 ch05_paper_ecs/paper_ecs.tex
Optimized introduction of chapter 05

@@
 not only important for the scientific community, but also directly affects
-society and policymakers. An important question in this context could be the
-following: Is it reasonable to adapt the assessed \ac{ECS} range from \acs{AR}5
-($\range{1.5}{4.5} \unit{K}$) in \acs{AR}6 based on the \acs{CMIP}6 model range
-of $\range{1.8}{5.6} \unit{K}$? An argument for this is that due to the many
-new and improved processes included to the models, the \acs{CMIP}6 ensemble is
-arguably more realistic than its predecessor generations. However, even the
-\acs{CMIP}6 models are far from fully simulating all aspects and processes of
-the full vastly complex Earth system. This could introduce new biases which
+society and policymakers. Due to the many new and improved processes included
+to the models, the \acs{CMIP}6 ensemble is arguably more realistic than its
+predecessor generations. However, does this automatically make results derived
+from \acs{CMIP}6 models more trustworthy than those derived from earlier
+\ac{CMIP} models? In general, the answer to this question is \enquote{no}. Even
+the \acs{CMIP}6 models are far from fully simulating all aspects and processes
+of the full vastly complex Earth system, which could introduce new biases which
 have not been present in earlier \ac{CMIP} models. For example, the apparent
@@
 caused by missing processes that cancel out the effects of the newly included
-processes. Thus, it is vital to take other sources of information into account
-before answering the question raised above.
+processes. Thus, to assess the credibility of \ac{ESM} results it is vital to
+take other sources of information into account.
 
@@
 \cref{subsec:02:emergent_constraints}, appropriate observations of the Earth
-system can be utilized within the emergent constraints framework to assess the
-accuracy of climate model projections and reduce associated uncertainties. For
-\ac{ECS}, a large variety of different emergent constraints has been developed
-over the years, which we now evaluate on the \acs{CMIP}6 ensemble in order to
-assess whether they still hold for these new climate models. In total we assess
-eleven emergent constraints on \ac{ECS} (see
-\cref{tab:05:overview_emergent_constraints}), out of which the most are related
-to cloud feedbacks since these constitute the most important source of
-uncertainty for \ac{ECS} \commentcite{Boucher2013, Flato2013}{see also
-  \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
-constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, the
-\acs{CMIP}6 models offer an exciting possibility to perform out-of-sample
-testing using the previously unseen data.
+system can be utilized within the so-called \emph{emergent constraints}
+framework to assess the accuracy of climate model projections and reduce
+associated uncertainties. For \ac{ECS}, a large variety of different emergent
+constraints has been developed over the years, which we now evaluate on the
+\acs{CMIP}6 ensemble in order to assess whether they still hold for these new
+climate models, \ie{} whether they have skill to reduce uncertainties in
+\ac{ECS} in the \acs{CMIP}6 ensemble. In total we assess eleven emergent
+constraints on \ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out
+of which the most are related to cloud feedbacks since these constitute the
+most important source of uncertainty for \ac{ECS} \commentcite{Boucher2013,
+  Flato2013}{see also \cref{subsec:02:climate_feedbacks}}. Since all of the
+evaluated emergent constraints have been derived on the \acs{CMIP}3 or
+\acs{CMIP}5 ensemble, the \acs{CMIP}6 models offer an exciting possibility to
+perform out-of-sample testing using the previously unseen data.
 
-This analysis is already published in \textcite{Schlund2020a}. For this paper,
-the author of this thesis lead the writing and the analysis and implemented the
-code to reproduce this analysis with all figures and tables using the
-\ac{ESMValTool}. In this chapter, we first introduce the data (climate model
-output and observations) and methods used in this study
-(\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
-constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
-ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
-provide a discussion (\cref{sec:05:discussion}) and a summary of the results
-(\cref{sec:05:summary}).
-
-\begin{table}[p]
+\begin{table}[!t]
   \centering
@@
 \end{table}
+
+This analysis is already published in \textcite{Schlund2020a}. For this paper,
+the author of this thesis lead the writing and the analysis and implemented the
+code to reproduce this analysis with all figures and tables using the
+\ac{ESMValTool}. In this chapter, we first introduce the data (climate model
+output and observations) and methods used in this study
+(\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
+constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
+ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
+provide a discussion (\cref{sec:05:discussion}) and a summary of the results
+(\cref{sec:05:summary}).
 

### 2021-02-12 ch01_introduction/introduction.tex
Added introduction of chapter 06 and fixed some acronyms

@@
   System Dynamics} \autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a
-new method to reduce uncertainties in gridded multi-model projections of the
-future climate with observations based on \ac{ML}. As an example, the method is
-applied to the photosynthesis rate at the end of the \nth{21} century, which is
-already published in the \emph{Journal of Geophysical Research: Biogeosciences}
-\autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook} provides a
-summary of the results of this thesis and gives an outlook of possible future
-works.
+new method to reduce uncertainties in multi-dimensional (gridded) multi-model
+projections of the future climate with observations based on \ac{ML}. As an
+example, the method is applied to the photosynthesis rate at the end of the
+\nth{21} century, which is already published in the \emph{Journal of
+  Geophysical Research: Biogeosciences} \autocite{Schlund2020}. Finally,
+\cref{ch:07:summary_outlook} provides a summary of the results of this thesis
+and gives an outlook of possible future works.

### 2021-02-12 ch02_scientific_background/scientific_background.tex
Added introduction of chapter 06 and fixed some acronyms

@@
 
-\section{\aclp{ESM}: Simulations and Analysis}
+\section{Earth System Models: Simulations and Analysis}
 \label{sec:02:esms}
@@
 
-\subsection{Equilibrium and \acl{ECS}}
+\subsection{Equilibrium and Effective Climate Sensitivity}
 \label{subsec:02:ecs}
 
-An important metric describing climate sensitivity is the Equilibrium Climate
-Sensitivity. It is defined as the change in \ac{GSAT} after an instantaneous
+An important metric describing climate sensitivity is the equilibrium climate
+sensitivity. It is defined as the change in \ac{GSAT} after an instantaneous
 doubling of the atmospheric \ac{CO2} concentration from pre-industrial
@@
 the anthropogenic climate change, the \emph{Charney Report} from 1979
-\autocite{Charney1979}, the Equilibrium Climate Sensitivity is one of the
+\autocite{Charney1979}, the equilibrium climate sensitivity is one of the
 oldest metrics describing climate change. However, in practice this traditional
@@
 
-For this reason, the Equilibrium Climate Sensitivity is commonly approximated
+For this reason, the equilibrium climate sensitivity is commonly approximated
 with the \acf{ECS}, which can be derived from only 150 model years of a
@@
 The factor of $2$ in the denominator accounts for the fact the traditional
-Equilibrium Climate Sensitivity is defined for an abrupt \ac{CO2} doubling,
+equilibrium climate sensitivity is defined for an abrupt \ac{CO2} doubling,
 whereas here a simulation with an abrupt quadrupling is considered.
@@
 Although commonly used in the literature, the \ac{ECS} is known to be only an
-approximation of the true Equilibrium Climate Sensitivity. One major reason for
+approximation of the true equilibrium climate sensitivity. One major reason for
 this is the state and time dependence of the global feedbacks
@@
 describing non-linear effects. A second major reason for the discrepancies
-between the Equilibrium Climate Sensitivity and \ac{ECS} is the use of a
+between the equilibrium climate sensitivity and \ac{ECS} is the use of a
 \nxcotwo{4} instead of a \nxcotwo{2} simulation. The factor of $2$ in the
@@
 However, despite these deficiencies the \ac{ECS} is still a practical estimate
-of the Equilibrium Climate Sensitivity. With the help of climate models,
+of the equilibrium climate sensitivity. With the help of climate models,
 \textcite{Sherwood2020} showed that the \ac{ECS} is only about $6 \unit{\%}$
@@
 \autocite{Rugenstein2019} can be a promising way forward to estimate the true
-Equilibrium Climate Sensitivity for \acp{ESM}.
+equilibrium climate sensitivity for \acp{ESM}.
 
@@
 
-\subsection{\acl{TCR}}
+\subsection{Transient Climate Response}
 \label{subsec:02:tcr}
@@
 discussion on the emergent constraint by \textcite{Wenzel2016}, which focuses
-on the concentration-carbon feedback, is given in \cref{sec:06:step_1}.
+on the concentration-carbon feedback, is given in \cref{subsec:06:step_1}.
 
@@
 
-\subsection{\acl{MDER}}
+\subsection{Multiple Diagnostic Ensemble Regression}
 \label{subsec:02:mder}

### 2021-02-12 ch03_esmvaltool/esmvaltool.tex
Added introduction of chapter 06 and fixed some acronyms

@@
 applied to the input data. There are two classes of preprocessors: non-optional
-and optional ones. Examples for non-optional preprocessors are the \ac{I/O}
+and optional ones. Examples for non-optional preprocessors are the input/output
 operations \emph{load} (loading input data from climate models and

### 2021-02-12 ch05_paper_ecs/paper_ecs.tex
Added introduction of chapter 06 and fixed some acronyms

@@
 
-\chapter{Evaluation of Emergent Constraints on the \acl{ECS} in \acs{CMIP}6}
+\chapter{Evaluation of Emergent Constraints on the Effective Climate
+  Sensitivity in \acs{CMIP}6}
 \label{ch:05:paper_ecs}
@@
 
-This analysis is already published in \textcite{Schlund2020a}. For this paper,
-the author of this thesis lead the writing and the analysis and implemented the
-code to reproduce this analysis with all figures and tables using the
+This work is already published in \textcite{Schlund2020a}. For this paper, the
+author of this thesis lead the writing and the analysis and implemented the
+code to reproduce this study with all figures and tables using the
 \ac{ESMValTool}. In this chapter, we first introduce the data (climate model
@@
 feedbacks. Thus, constraining \ac{ECS} with a single physical process might
-overly simplify this problem. Such single process-oriented emergent constraints
+overly simplify this problem. Such single-process-oriented emergent constraints
 therefore do not seem to be helpful in constraining \ac{ECS} but should
@@
 similar to the participating individual emergent constraints, but with lower
-uncertainty range. Moreover, new machine learning techniques are a promising
-avenue forward for such multivariate approaches and for constraining
-uncertainties in multi-model projections \commentcite{Schlund2020}{see
-  \cref{ch:06:paper_gpp}} with the aim of further improving climate modeling
-and analysis \autocite{Reichstein2019}.
+uncertainty range. Moreover, new \ac{ML} techniques are a promising avenue
+forward for such multivariate approaches and for constraining uncertainties in
+multi-model projections \commentcite{Schlund2020}{see \cref{ch:06:paper_gpp}}
+with the aim of further improving climate modeling and analysis
+\autocite{Reichstein2019}.

### 2021-02-12 ch06_paper_gpp/paper_gpp.tex
Added introduction of chapter 06 and fixed some acronyms

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                                  Paper GPP                                  %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Constraining Future Gross Primary Production with Machine Learning}
+\label{ch:06:paper_gpp}
 
-The \ac{ECS} is really cool. I like it very much!
+As argued in detail in the previous chapter, single-process-oriented emergent
+constraints do not seem to be beneficial in constraining climate projections
+from the vastly complex modern-day \acp{ESM} since they might overly simplify
+the problem. To overcome this issue, we explore an alternative approach that
+expands the \acf{MDER} method with a supervised \acf{ML} algorithm. In contrast
+to emergent constraints, this approach considers (1) multiple process-based
+predictors, (2) multi-dimensional (gridded) target variables and (3) non-linear
+relationships between the predictors and the target variable. Since the new
+technique relies on a large number of data points in order to train the \ac{ML}
+algorithm, \ac{ECS} as a single scalar metric for each climate model is not an
+appropriate target variable. As an alternative, we analyze projections of
+future \acf{GPP} in this study, which constitutes the largest flux of the
+terrestrial carbon uptake. As described in detail in
+\cref{subsec:02:carbon_cycle_perturbations}, the global carbon cycle actively
+slows down global warming by removing \ac{CO2} from the atmosphere. In
+particular, the terrestrial biosphere absorbs about $30 \unit{\%}$ of the total
+anthropogenic \ac{CO2} emission \autocite{Friedlingstein2020}. Thus, accurately
+quantifying the future evolution of the terrestrial carbon sink is crucial for
+reliable climate change projections.
+
+Due to complex feedback processes connected to the global carbon cycle
+(concentration-carbon feedback and climate-carbon feedback; see
+\cref{subsec:02:carbon_cycle_perturbations}), the response of the terrestrial
+carbon cycle to changes in the atmospheric \ac{CO2} concentration and climate
+is a major source of uncertainty in climate projections \autocite{Bodman2013,
+  Booth2012, Collins2013}. This applies for example to multi-model projections
+of future \ac{GPP} of the \acs{CMIP}5 ensemble. Thus, the aim of this study is
+to reduce uncertainties in multi-model projections of future \ac{GPP} in the
+\acs{CMIP}5 \acs{RCP}8.5 scenario at the end of the \nth{21} century with a new
+\ac{ML}-based weighting technique and appropriate observations. The new
+two‐step approach utilizes aspects of emergent constraints and the \ac{MDER}
+technique in combination with a \ac{ML} algorithm. In the first step, we apply
+an existing emergent constraint on \ac{CO2} fertilization \autocite{Wenzel2016}
+to constrain the \acp{ESM}' responses to rising atmospheric \ac{CO2}
+concentration using observations of the increase of the \ac{CO2} seasonal cycle
+amplitude at Cape Kumukahi, Hawaii \autocite{Keeling2005}. In a second step, we
+introduce a supervised \ac{ML} algorithm based on boosting trees
+\autocite{Friedman2001} to learn an empirical spatial relationship that links
+grid‐wise future \ac{GPP} to historical processes relevant to its simulation
+under present‐day conditions. In combination with observational products of the
+predictors, that relationship can be used to further constrain uncertainties in
+the projected spatial maps of \ac{GPP} at the end of the \nth{21} century in
+the \acs{RCP}8.5 scenario. We examine both constraining the absolute \ac{GPP}
+and the fractional change in \ac{GPP} as two independent approaches and target
+variables. Unlike univariate linear regression used in the \ac{MDER} algorithm,
+the proposed \ac{GBRT} algorithm is able to handle multiple predictors and
+copes with non-linearities in the data. \ac{GBRT} is a well‐known and
+successful tool used for interpolation, classification and prediction in other
+fields of data science and engineering \autocite{Death2007, Elith2008}. In the
+context of climate science, \ac{GBRT} was recently applied to identify the key
+drivers of spatial variations of the ratio of plant transpiration to total
+terrestrial evapotranspiration in \acp{ESM} \autocite{Lian2018}.
+
+This work is already published in \textcite{Schlund2020}. For this paper, the
+author of this thesis lead the writing and the analysis and implemented the
+code to reproduce this study with all figures and tables using the
+\ac{ESMValTool}. \Cref{sec:06:data_and_methods} of this chapter provides an
+overview of the data and methods used in this paper. The results are presented
+in \cref{sec:06:results} and \cref{sec:06:summary_and_discussion} closes with a
+summary and discussion.
+
+\section{Data and Methods}
+\label{sec:06:data_and_methods}
+
+TBA.
+
+\subsection{Constraining the \texorpdfstring{\acs{CO2}}{CO2} Fertilization
+  Effect (Step 1)}
+\label{subsec:06:step_1}
+
+TBA.
+
+\section{Results}
+\label{sec:06:results}
+
+TBA.
+
+\section{Summary and Discussion}
+\label{sec:06:summary_and_discussion}
+
+TBA.

### 2021-02-15 ch02_scientific_background/scientific_background.tex
Added first figures for chapter 6

@@
 the weighting scheme by calculating the \ac{RMSE} between the prediction and
-the known ground truth of the pseudo-observation. The whole process is repeated
-for every climate model of the ensemble to get a statistical distribution of
-\acp{RMSE}. Finally, different \ac{RMSE} distributions calculated from
-different parameters $\sigma_D$ and $\sigma_S$ can be assessed using specific
-criteria to find optimal values for $\sigma_D$ and $\sigma_S$
-\autocite{Knutti2017a}. Furthermore, the leave-one-model-out \ac{CV} approach
-can be used to evaluate different climate model weightings schemes (including
-the unweighted \ac{MMM}) and compare them against each other.
+the known ground truth of the pseudo-observation, which is also known as
+\ac{RMSEP}. The whole process is repeated for every climate model of the
+ensemble to get a statistical distribution of \acp{RMSEP}. Finally, different
+\ac{RMSEP} distributions calculated from different parameters $\sigma_D$ and
+$\sigma_S$ can be assessed using specific criteria to find optimal values for
+$\sigma_D$ and $\sigma_S$ \autocite{Knutti2017a}. Furthermore, the
+leave-one-model-out \ac{CV} approach can be used to evaluate different climate
+model weightings schemes (including the unweighted \ac{MMM}) and compare them
+against each other.
 
@@
 which assume that climate models that are better in simulating the past or
-present‐day climate are necessarily better in simulating the future climate. In
+present-day climate are necessarily better in simulating the future climate. In
 contrast to these other weighting approaches, \ac{MDER} explicitly establishes

### 2021-02-15 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added first figures for chapter 6

@@
 feedback parameter in \acs{CMIP}6 compared to \acs{CMIP}5 due to a more
-positive (reduced negative) low‐level cloud feedback, primarily in the
+positive (reduced negative) low-level cloud feedback, primarily in the
 extratropics. In comparison to earlier \acs{CMIP} generations, the
-\acs{CMIP}6 models exhibit weaker increases in the extratropical low‐level
+\acs{CMIP}6 models exhibit weaker increases in the extratropical low-level
 cloud cover and the corresponding liquid water content as a result of global
@@
 an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds for
-the pre-industrial and present‐day periods, leading to the aforementioned
+the pre-industrial and present-day periods, leading to the aforementioned
 reduction in the negative cloud phase change feedback on warming and

### 2021-02-15 ch06_paper_gpp/paper_gpp.tex
Added first figures for chapter 6

@@
 \ac{ML}-based weighting technique and appropriate observations. The new
-two‐step approach utilizes aspects of emergent constraints and the \ac{MDER}
+two-step approach utilizes aspects of emergent constraints and the \ac{MDER}
 technique in combination with a \ac{ML} algorithm. In the first step, we apply
@@
 \autocite{Friedman2001} to learn an empirical spatial relationship that links
-grid‐wise future \ac{GPP} to historical processes relevant to its simulation
-under present‐day conditions. In combination with observational products of the
+grid-wise future \ac{GPP} to historical processes relevant to its simulation
+under present-day conditions. In combination with observational products of the
 predictors, that relationship can be used to further constrain uncertainties in
@@
 the proposed \ac{GBRT} algorithm is able to handle multiple predictors and
-copes with non-linearities in the data. \ac{GBRT} is a well‐known and
+copes with non-linearities in the data. \ac{GBRT} is a well-known and
 successful tool used for interpolation, classification and prediction in other
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch06_paper_gpp/figs/schematic_steps.pdf}
+  \caption{Schematic illustration of our two-step approach. In step 1 an
+    emergent constraint by \textcite{Wenzel2016} is used to constrain the
+    global mean fractional change in \acf{GPP} over the 21st century.
+    Moreover, this constraint is used to rescale two different gridded target
+    variables: absolute \acs{GPP} at the end of the \nth{21} century and
+    fractional \acs{GPP} change over the \nth{21} century. In step 2, a
+    \acl{ML} model is used to constrain these two target variables (step 2a:
+    absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
+    independent approaches. \AdaptedFrom{Schlund2020a}.}
+  \label{fig:06:schematic_steps}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=\LargerFigureWidth{}]{
+    ch06_paper_gpp/figs/mlr_concept.pdf}
+  \caption{Schematic illustration of our \acf{ML} approach to constrain
+    projected absolute \acf{GPP} in step 2a. (a) In the training phase of the
+    algorithm, the model is fitted to the training data interpolating the
+    empirical (non-linear) relationship between two process-oriented
+    diagnostics of the past climate $\left\{ x^{(1)}, x^{(2)} \right\}$ and
+    (rescaled) future \acs{GPP} (gray surface). The points show the training
+    points for the supervised \acs{ML} algorithm, each of them representing a
+    single grid cell/month of a single climate model (the different colors
+    correspond to different climate models). (b) In the prediction phase,
+    observation-based values of the diagnostic (black points) are fed into the
+    trained \acs{ML} model to constrain \acs{GPP} for every grid cell/month to
+    a value which best agrees with the observations. (c) For an independent
+    validation of our method, we use an out-of-sample testing setup based on a
+    leave-one-model-out \acl{CV} approach (see
+    \cref{subsec:02:model_weighting} for details). The schematic illustration
+    of step 2b differs only in the target variable used (fractional \acs{GPP}
+    change instead of absolute \acs{GPP}). \AdaptedFrom{Schlund2020a}.}
+  \label{fig:06:mlr_concept}
+\end{figure}
+
 \subsection{Constraining the \texorpdfstring{\acs{CO2}}{CO2} Fertilization

### 2021-02-16 ch06_paper_gpp/paper_gpp.tex
Added further code and figures to chapter 06

@@
     absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
-    independent approaches. \AdaptedFrom{Schlund2020a}.}
+    independent approaches. \AdaptedFrom{Schlund2020}.}
   \label{fig:06:schematic_steps}
@@
     of step 2b differs only in the target variable used (fractional \acs{GPP}
-    change instead of absolute \acs{GPP}). \AdaptedFrom{Schlund2020a}.}
+    change instead of absolute \acs{GPP}). \AdaptedFrom{Schlund2020}.}
   \label{fig:06:mlr_concept}
@@
 
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \raisebox{5.5mm}{\includegraphics[width=\columnwidth]{
+      ch06_paper_gpp/figs/step2a_rmse.pdf}}
+    \caption{}
+    \label{fig:06:step2a_ml_inference:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{
+      ch06_paper_gpp/figs/step2a_global_feature_importance.pdf}
+    \caption{}
+    \label{fig:06:step2a_ml_inference:b}
+  \end{subfigure}
+  \caption{(a) Box plot of the \acf{RMSEP} distributions for six different
+    statistical models used to predict future absolute \acf{GPP} in step 2a
+    using a leave-one-model-out \acl{CV} approach. The distribution for each
+    statistical model contains seven elements (black points; one for each
+    climate model used as truth) and is represented in the following way: The
+    lower and upper limit of the blue boxes correspond to the $25 \unit{\%}$
+    and $75 \unit{\%}$ quantiles, respectively. The central line in the box
+    shows the median and the black \enquote{x} shows the mean of the
+    distribution. The whiskers outside the box represent the range of the
+    distribution. Compared to the \acs{CMIP}5 \acf{MMM} and its corresponding
+    rescaled version (r\acs{MMM}), the prediction uncertainty measured by the
+    mean \acs{RMSEP} is significantly reduced by up to $48 \unit{\%}$ and $39
+    \unit{\%}$, respectively, when using other statistical models. Moreover,
+    the non-linear \acf{GBRT} models can slightly reduce the mean \acs{RMSEP}
+    compared to the linear \acf{LASSO} models by about $2 \unit{\%}$ for
+    \acs{GBRT}-1D (using historical \acs{GPP} as single predictor) and $3
+    \unit{\%}$ for the full \acs{GBRT} model (using all predictors). (b)
+    Relative global feature importance for the different predictors used in
+    the \acs{GBRT} model to predict future absolute \acs{GPP} (step 2a). The
+    red bars correspond to positive Pearson correlation coefficients between
+    all predictors and the target variable. Due to its strong positive linear
+    relationship with the future \acs{GPP}, the historical \acs{GPP} is by far
+    the most important predictor in the model. A local feature importance map
+    using \acf{LIME} is not shown here because \acs{GPP} is the dominant
+    predictor for all grid cells. \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:step2a_ml_inference}
+\end{figure}
+
 \section{Summary and Discussion}

### 2021-02-16 end_acronyms/acronyms.tex
Added subsection on calculation of CRE feedback parameters
Added contributions to ESMValCore to chapter 03
Cleaned code
Added figures for ESMValTool contributions (chapter 3)
Finished chapter 3
Cleaned tex files and added text to chapter 4
Added text and figures to section 4.1
Added tables for chapter 05
Added first parts of section 5.2
Finished section 5.2
Added introduction of chapter 06 and fixed some acronyms
Added first figures for chapter 6
Added further code and figures to chapter 06

@@
   \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
-  \acro{AOGCM}{Atmosphere-Ocean General Circulation Model}
+  \acro{AOGCM}{atmosphere-ocean general circulation model}
   \acro{AR}{Assessment Report}
@@
   \acro{CESM}{Community Earth System Model}
+  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
   \acro{CMIP}{Coupled Model Intercomparison Project}
@@
   \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
+  \acro{CRE}{cloud radiative effect}
   \acro{CV}{cross-validation}
   \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
+  \acro{DJF}{December-January-February}
   \acro{DLR}{%
     Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
-  \acro{ECS}{Effective Climate Sensitivity}
+  \acro{ECS}{effective climate sensitivity}
   \acro{ENSO}{El Niño-Southern Oscillation}
   \acro{ESGF}{Earth System Grid Federation}
-  \acro{ESM}{Earth System Model}
+  \acro{ESM}{Earth system model}
   \acro{ESMValCore}{core functionalities of the ESMValTool}
   \acro{ESMValTool}{Earth System Model Evaluation Tool}
-  \acro{GHG}{Greenhouse Gas}
-  \acro{GPP}{Gross Primary Productivity}
+  \acro{GBRT}{gradient boosted regression tree}
+  \acro{GHG}{greenhouse gas}
+  \acro{GPP}{gross primary production}
   \acro{GSAT}{global mean near-surface air temperature}
   \acro{GtC}{gigatonnes of carbon}
-  \acro{I/O}{Input/Output}
-  \acro{IAM}{Integrated Assessment Models}
+  \acro{IAM}{integrated assessment model}
   \acro{IPA}{%
@@
   \acro{IPCC}{Intergovernmental Panel on Climate Change}
+  \acro{ITCZ}{Intertropical Convergence Zone}
+  \acro{JJA}{June-July-August}
+  \acro{LASSO}{least absolute shrinkage and selection operator}
+  \acro{LIME}{local interpretable model-agnostic explanations}
   \acro{LongRunMIP}{Long Run Model Intercomparison Project}
-  \acro{MDER}{Multiple Diagnostic Ensemble Regression}
+  \acro{LTMI}{lower-tropospheric mixing index}
+  \acro{MBLC}{marine boundary layer cloud}
+  \acro{MDER}{multiple diagnostic ensemble regression}
   \acro{MIP}{Model Intercomparison Project}
-  \acro{ML}{Machine Learning}
+  \acro{ML}{machine learning}
   \acro{MMM}{multi-model mean}
@@
   \acro{RMSE}{root-mean-square error}
+  \acro{RMSEP}{root-mean-square error of prediction}
   \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
   \acro{SSP}{Shared Socioeconomic Pathway}
-  \acro{TCR}{Transient Climate Response}
+  \acro{SST}{sea surface temperature}
+  \acro{TCR}{transient climate response}
+  \acro{TLC}{tropical low-level cloud}
   \acro{TOA}{top of the atmosphere}

### 2021-02-16 ch06_paper_gpp/paper_gpp.tex
Renamed figures for chapter 06

@@
   \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
-    ch06_paper_gpp/figs/schematic_steps.pdf}
+  \includegraphics[width=\LargeFigureWidth{}]{ch06_paper_gpp/figs/1.pdf}
   \caption{Schematic illustration of our two-step approach. In step 1 an
@@
   \centering
-  \includegraphics[width=\LargerFigureWidth{}]{
-    ch06_paper_gpp/figs/mlr_concept.pdf}
+  \includegraphics[width=\LargerFigureWidth{}]{ch06_paper_gpp/figs/2.pdf}
   \caption{Schematic illustration of our \acf{ML} approach to constrain
@@
     \raisebox{5.5mm}{\includegraphics[width=\columnwidth]{
-      ch06_paper_gpp/figs/step2a_rmse.pdf}}
+      ch06_paper_gpp/figs/4a.pdf}}
     \caption{}
@@
   \begin{subfigure}[b]{\SubfigureWidth{}}
-    \includegraphics[width=\columnwidth]{
-      ch06_paper_gpp/figs/step2a_global_feature_importance.pdf}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/4b.pdf}
     \caption{}

### 2021-02-17 ch02_scientific_background/scientific_background.tex
Cleaned text and added table to chapter 06

@@
 models. For \ac{ScenarioMIP}, five different \acp{SSP} are considered, ranging
-from \acs{SSP}1 (sustainability) to \acs{SSP}5 (fossil-fuel development). Each
-\ac{SSP} is combined with a climate outcome (measured as radiative forcing in
-the year 2100) based on a particular forcing pathway that \acp{IAM} have shown
-to be feasible. For example, \acs{SSP}5-8.5 represents a scenario based on a
-fossil-fuel development with a radiative forcing of $8.5 \unit{W m^{-2}}$ in
-2100 while \acs{SSP}1-2.6 represents a sustainable future with a radiative
-forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two scenarios in the
-main category of \ac{ScenarioMIP}, the \emph{Tier 1} experiments, are the
-\acs{SSP}2-4.5 and \acs{SSP}3-7.0 scenarios. In contrast to the
-\ac{ScenarioMIP} experiments, the corresponding \acs{CMIP}5 counterparts
+from \acs{SSP}1 (sustainability) to \acs{SSP}5 (fossil fuel--based
+development). Each \ac{SSP} is combined with a climate outcome (measured as
+radiative forcing in the year 2100) based on a particular forcing pathway that
+\acp{IAM} have shown to be feasible. For example, \acs{SSP}5-8.5 represents a
+scenario based on a fossil fuel--based development with a radiative forcing of
+$8.5 \unit{W m^{-2}}$ in 2100 while \acs{SSP}1-2.6 represents a sustainable
+future with a radiative forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The
+two scenarios in the main category of \ac{ScenarioMIP}, the \emph{Tier 1}
+experiments, are the \acs{SSP}2-4.5 and \acs{SSP}3-7.0 scenarios. In contrast
+to the \ac{ScenarioMIP} experiments, the corresponding \acs{CMIP}5 counterparts
 \autocite{Taylor2012}, the so-called \acp{RCP}, only used the radiative forcing
@@
 \ac{ScenarioMIP} that include a variety of different scenarios from a
-sustainable future to a full fossil fuel-based development (see
+sustainable future to a full fossil fuel--based development (see
 \cref{subsec:02:cmip}). A possible approach to quantify emission uncertainty is
@@
 Since one study presented in this thesis aims to reduce uncertainties in carbon
-cycle-related processes \commentcite{Schlund2020}{see \cref{ch:06:paper_gpp}},
+cycle--related processes \commentcite{Schlund2020}{see \cref{ch:06:paper_gpp}},
 this chapter introduces the scientific background of the global carbon cycle.
@@
     ch02_scientific_background/figs/keeling_curve.pdf}
-  \caption{The Keeling Curve: monthly-mean atmospheric \acs{CO2} concentration
+  \caption{The Keeling Curve: monthly mean atmospheric \acs{CO2} concentration
     at the Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6

### 2021-02-17 ch03_esmvaltool/esmvaltool.tex
Cleaned text and added table to chapter 06

@@
     \acs{MMM} includes the four climate models that provide all data needed
-    for the calculation of \emph{co2s} (see legend in (b)). (b) Monthly-mean
+    for the calculation of \emph{co2s} (see legend in (b)). (b) Monthly mean
     \emph{co2s} at Mauna Loa, Hawaii ($19.5 \unit{\degree N}$, $155.6
@@
 atmospheric \ac{CO2} concentration at surface (\emph{co2s}) can be calculated
-from the pressure level-dependent atmospheric \ac{CO2} concentration and the
+from the pressure level--dependent atmospheric \ac{CO2} concentration and the
 surface air pressure using interpolation. An example illustrating \emph{co2s}
@@
 America) exhibit smaller \ac{CO2} concentrations. \Cref{fig:03:co2s:b} shows
-the monthly-mean \emph{co2s} at Mauna Loa, Hawaii from 1958 to 2014 for
+the monthly mean \emph{co2s} at Mauna Loa, Hawaii from 1958 to 2014 for
 observations performed at the Mauna Loa Observatory \autocite{Keeling2005}
@@
 \emph{recipe\_wenzel16jclim.yml}. \emph{recipe\_anav13jclim.yml} reproduces the
-analysis of \textcite{Anav2013}, who evaluate carbon cycle-related variables
+analysis of \textcite{Anav2013}, who evaluate carbon cycle--related variables
 for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends

### 2021-02-17 ch05_paper_ecs/paper_ecs.tex
Cleaned text and added table to chapter 06

@@
 present-day Hadley circulation. The latter is calculated from the tropical
-($\range{45 \unit{\degree S}}{40 \unit{\degree N}}$) zonal-mean vertical
+($\range{45 \unit{\degree S}}{40 \unit{\degree N}}$) zonal mean vertical
 profiles of relative humidity from the surface to $100 \unit{hPa}$. These
@@
 constraint is related to an emergent constraint on \ac{ECS} proposed by
-\textcite{Fasullo2012}, who correlate \range{May}{August} zonal-mean relative
+\textcite{Fasullo2012}, who correlate \range{May}{August} zonal mean relative
 humidity against \ac{ECS}. In contrast to \textcite{Su2014}, they do not use

### 2021-02-17 ch06_paper_gpp/paper_gpp.tex
Cleaned text and added table to chapter 06

@@
 
+\begin{table}[t]
+  \centering
+  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
+      p{0.3\columnwidth}}
+    \toprule
+    Predictor & \makecell{Observation‐driven \\ data} & Used time range &
+    Physical connection to \acs{GPP} \\
+    \midrule
+    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
+    \range{1991}{2000} & -- \\
+    \midrule
+    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
+    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
+    grid cell. The photosynthesis rate is highly dependent on the number of
+    leaves (and vegetation in general). \\
+    \midrule
+    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
+    \range{1901}{2005} & Water is essential for the chemical processes of
+    photosynthesis. \\
+    \midrule
+    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
+    \range{1979}{2005} & Solar radiation is essential for the chemical
+    processes of photosynthesis. \\
+    \midrule
+    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
+    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
+    (incoming solar radiation). \\
+    \bottomrule
+  \end{tabular}
+  \caption{Process-oriented diagnostics (\enquote{predictors} or
+    \enquote{features}) used in the \acf{GBRT} model to predict the target
+    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
+    variables are monthly climatologies of the specified time ranges in the
+    historical climate. For step 2b (target variable: fractional \acs{GPP}
+    change), the temporal mean over the specified time ranges is calculated for
+    all variables. \AdaptedFrom{Schlund2020}.}
+  \label{tab:06:predictors}
+\end{table}
+
 \subsection{Constraining the \texorpdfstring{\acs{CO2}}{CO2} Fertilization

### 2021-02-17 end_appendix/appendix.tex
Cleaned text and added table to chapter 06

--- 2021-02-09 end_appendix/appendix.tex
@@
 
-This section contains supplementary tables for \cref{ch:05:paper_ecs}
-(\nameref*{ch:05:paper_ecs}; see \cite{Schlund2020a}).
+\vspace{\fill}
 
-\begin{table}[!b]
+\begin{table}[!h]
   \centering
-  \begin{tabulary}{\columnwidth}{L L}
+  \begin{tabulary}{\columnwidth}{>{\em}L L}
     \toprule
-    Variable short name & Description \\
+    \ignorecolumntype{l}{Variable short name} & Description \\
     \midrule
-    \emph{cl} & Cloud area fraction \\
-    \emph{clt} & Total cloud area fraction \\
-    \emph{hur} & Relative humidity \\
-    \emph{hus} & Specific humidity \\
-    \emph{pr} & Precipitation \\
-    \emph{rsdt} & \Acf{TOA} incoming shortwave radiation \\
-    \emph{rsut} & \acs{TOA} outgoing shortwave radiation \\
-    \emph{rsutcs} & Clear-sky \acs{TOA} outgoing shortwave radiation \\
-    \emph{ta} & Air temperature \\
-    \emph{tas} & Surface air temperature \\
-    \emph{tos} & Sea surface temperature \\
-    \emph{ts} & Surface temperature \\
-    \emph{va} & Northward wind speed \\
-    \emph{wap} & Vertical velocity \\
+    cl & Cloud area fraction \\
+    clt & Total cloud area fraction \\
+    hur & Relative humidity \\
+    hus & Specific humidity \\
+    pr & Precipitation \\
+    rsdt & \Acf{TOA} incoming shortwave radiation \\
+    rsut & \acs{TOA} outgoing shortwave radiation \\
+    rsutcs & Clear-sky \acs{TOA} outgoing shortwave radiation \\
+    ta & Air temperature \\
+    tas & Surface air temperature \\
+    tos & Sea surface temperature \\
+    ts & Surface temperature \\
+    va & Northward wind speed \\
+    wap & Vertical velocity \\
     \bottomrule
@@
     (\nameref*{ch:05:paper_ecs}). More details are given in
-    \cref{tab:05:overview_emergent_constraints}, which lists the corresponding
-    variable short names for each emergent constraint.}
+    \cref{tab:05:overview_emergent_constraints}, which lists the variables
+    used for each emergent constraint.}
   \label{tab:app:a:overview_variables}
@@
 
+\vspace{\fill}
+
 \begin{table}[!b]
   \centering
-  \begin{tabulary}{\columnwidth}{L L L}
+  \begin{tabulary}{\columnwidth}{L >{\em}L L}
     \toprule
-    Observational dataset & Corresponding variables & Reference \\
+    Observational dataset & \ignorecolumntype{l}{Corresponding variables} &
+    Reference \\
     \midrule
-    AIRS & \emph{hur}, \emph{hus} & \textcite{Aumann2003} \\
-    AMSRE SST & \emph{tos} & \textcite{AMSRE2011} \\
-    CERES-EBAF & \emph{rsdt}, \emph{rsut}, \emph{rsutcs} & \textcite{Loeb2018}
-    \\
-    CloudSat/CALIPSO & \emph{cl} & \textcite{Mace2009} \\
-    ERA-Interim & \emph{hur}, \emph{ta}, \emph{va}, \emph{wap} &
-    \textcite{Dee2011} \\
-    GPCP & \emph{pr} & \textcite{Adler2003} \\
-    HadCRUT4 & \emph{tas} & \textcite{Morice2012} \\
-    HadISST & \emph{ts} & \textcite{Rayner2003} \\
-    ISCCP D-2 & \emph{clt} & \textcite{Rossow1991} \\
-    MLS-Aura & \emph{hur} & \textcite{Beer2006} \\
+    AIRS & hur, hus & \textcite{Aumann2003} \\
+    AMSRE \ac{SST} & tos & \textcite{AMSRE2011} \\
+    CERES-EBAF & rsdt, rsut, rsutcs & \textcite{Loeb2018} \\
+    Cloudsat/CALIPSO & cl & \textcite{Mace2009} \\
+    ERA-Interim & hur, ta, va, wap & \textcite{Dee2011} \\
+    GPCP & pr & \textcite{Adler2003} \\
+    HadCRUT4 & tas & \textcite{Morice2012} \\
+    HadISST & ts & \textcite{Rayner2003} \\
+    ISCCP D-2 & clt & \textcite{Rossow1991} \\
+    MLS-Aura & hur & \textcite{Beer2006} \\
     \bottomrule
@@
   \caption{References for all observational datasets used in
-    \cref{ch:05:paper_ecs} (\nameref*{ch:05:paper_ecs}).}
+    \cref{ch:05:paper_ecs} (\nameref*{ch:05:paper_ecs}). More details are
+    given in \cref{tab:05:overview_emergent_constraints}, which lists the
+    observational datasets used for each emergent constraint.}
   \label{tab:app:a:observations}
@@
   \centering
-  \begin{tabulary}{\columnwidth}{L L L}
+  \begin{tabulary}{\columnwidth}{L >{$}L<{$} L}
     \toprule
-    Model & Index & Main reference \\
+    Model & \ignorecolumntype{l}{Index} & Main reference \\
     \midrule

### 2021-02-17 ch05_paper_ecs/paper_ecs.tex
Added figures of chapter 06 to tex file

@@
 \ac{ECS}. Except for the ZHA and SHS constraints, every emergent relationship
-investigated shows a lower statistical significance (\ie{} a higher $p$- value)
+investigated shows a lower statistical significance (\ie{} a higher $p$-value)
 in the \acs{CMIP}6 ensemble than in the \acs{CMIP}5 ensemble. If a conventional

### 2021-02-17 ch06_paper_gpp/paper_gpp.tex
Added figures of chapter 06 to tex file

@@
 concentration using observations of the increase of the \ac{CO2} seasonal cycle
-amplitude at Cape Kumukahi, Hawaii \autocite{Keeling2005}. In a second step, we
-introduce a supervised \ac{ML} algorithm based on boosting trees
-\autocite{Friedman2001} to learn an empirical spatial relationship that links
-grid-wise future \ac{GPP} to historical processes relevant to its simulation
-under present-day conditions. In combination with observational products of the
-predictors, that relationship can be used to further constrain uncertainties in
-the projected spatial maps of \ac{GPP} at the end of the \nth{21} century in
-the \acs{RCP}8.5 scenario. We examine both constraining the absolute \ac{GPP}
-and the fractional change in \ac{GPP} as two independent approaches and target
-variables. Unlike univariate linear regression used in the \ac{MDER} algorithm,
-the proposed \ac{GBRT} algorithm is able to handle multiple predictors and
-copes with non-linearities in the data. \ac{GBRT} is a well-known and
-successful tool used for interpolation, classification and prediction in other
-fields of data science and engineering \autocite{Death2007, Elith2008}. In the
-context of climate science, \ac{GBRT} was recently applied to identify the key
-drivers of spatial variations of the ratio of plant transpiration to total
-terrestrial evapotranspiration in \acp{ESM} \autocite{Lian2018}.
+amplitude at \ac{KUM} \autocite{Keeling2005}. In a second step, we introduce a
+supervised \ac{ML} algorithm based on boosting trees \autocite{Friedman2001} to
+learn an empirical spatial relationship that links grid-wise future \ac{GPP} to
+historical processes relevant to its simulation under present-day conditions.
+In combination with observational products of the predictors, that relationship
+can be used to further constrain uncertainties in the projected spatial maps of
+\ac{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario. We
+examine both constraining the absolute \ac{GPP} and the fractional change in
+\ac{GPP} as two independent approaches and target variables. Unlike univariate
+linear regression used in the \ac{MDER} algorithm, the proposed \ac{GBRT}
+algorithm is able to handle multiple predictors and copes with non-linearities
+in the data. \ac{GBRT} is a well-known and successful tool used for
+interpolation, classification and prediction in other fields of data science
+and engineering \autocite{Death2007, Elith2008}. In the context of climate
+science, \ac{GBRT} was recently applied to identify the key drivers of spatial
+variations of the ratio of plant transpiration to total terrestrial
+evapotranspiration in \acp{ESM} \autocite{Lian2018}.
 
@@
     absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
-    independent approaches. \AdaptedFrom{Schlund2020}.}
+    independent approaches. \AdaptedFrom{Schlund2020}. \cref{eq:06:xxx}}
   \label{fig:06:schematic_steps}
@@
   \centering
+  \begin{subfigure}[b]{0.37\columnwidth}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3a.pdf}
+    \caption{}
+    \label{fig:06:step1:a}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.37\columnwidth}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3b.pdf}
+    \caption{}
+    \label{fig:06:step1:b}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.24\columnwidth}
+    \raisebox{12.9mm}{\includegraphics[width=\columnwidth]{
+      ch06_paper_gpp/figs/3c.pdf}}
+  \end{subfigure}
+  \caption{Emergent relationship between the global mean fractional change in
+    \acf{GPP} and the sensitivity of the \acs{CO2} seasonal cycle amplitude to
+    rising atmospheric \acs{CO2} concentrations observed at \acf{KUM}. Colored
+    points refer to \acs{CMIP}5 models, the orange line and shaded area to the
+    linear  regression fit and its corresponding standard prediction error and
+    the dashed black lines to the observational constraint. (a) Similar to
+    \textcite{Wenzel2016}: The global mean fractional \acs{GPP} change after
+    \acs{CO2} doubling in a \acf{1BGC} defined as 10-year mean of the
+    \acs{1BGC} run centered at the time of \acs{CO2} doubling relative to the
+    10-year mean of pre-industrial control conditions at the beginning of the
+    \acs{1BGC} run. The sensitivity of the \acs{CO2} amplitude for the
+    \acs{CMIP}5 models is calculated from the years \range{1860}{2005}, the
+    corresponding observational value from the years \range{1979}{2019}. The
+    constrained global mean \acs{GPP} change after \acs{CO2} doubling in the
+    \acs{1BGC} run is $\pmrangeunit{30}{9}{\%}$. (b) Fractional global mean
+    \acs{GPP} change over the \nth{21} century calculated from the 10-year
+    mean \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
+    the emission-driven fully coupled \acs{RCP}8.5 simulations and the 10-year
+    mean \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in
+    the emission‐driven fully-coupled historical run (see \cref{eq:06:xxx}).
+    In contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
+    amplitude is calculated from the years \range{1979}{2019} for climate
+    models and observations for better comparability (the historical
+    \acs{CMIP}5 simulations are extended with the \acs{RCP}8.5 simulations for
+    the years \range{2006}{2019}). The constrained global mean \acs{GPP}
+    change over the \nth{21} century is $\pmrangeunit{39}{7}{\%}$.
+    \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:step1}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
   \begin{subfigure}[b]{\SubfigureWidth{}}
@@
 
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5a.pdf}
+    \caption{}
+    \label{fig:06:step2a_results:a}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5b.pdf}
+    \caption{}
+    \label{fig:06:step2a_results:b}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5c.pdf}
+    \caption{}
+    \label{fig:06:step2a_results:c}
+  \end{subfigure}
+  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the absolute
+    \acf{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
+    \acs{RCP}8.5 scenario using \cref{eq:06:xxx} (r\acs{MMM}) and its
+    unweighted version (\acs{MMM}). The plot shows an almost constant value
+    over the whole globe with a mean of $0.92$, which corresponds to the ratio
+    of the constrained global mean \acs{GPP} change over the \nth{21} century
+    from step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
+    \acs{GPP} change ($43 \unit{\%}$). All values close to $0$ for the dataset
+    in the denominator have been masked to avoid divisions by $0$. (b) Bias
+    between r\acs{MMM} and our \acf{GBRT} prediction for the end of the
+    \nth{21} century. This corresponds to step 2a of our approach. (c) Bias
+    between the modeled \acs{GPP} in the \acs{CMIP}5 \acs{MMM} of the
+    historical simulation and the FLUXNET-MTE observation‐based estimate of
+    \acs{GPP} \autocite{Jung2011} averaged between 1991 and 2000. Over large
+    swaths of the globe, the \acs{CMIP}5 ensemble overestimates \acs{GPP} (red
+    color). Panels (b) and (c) show similar bias patterns (pattern correlation
+    of $R^2 = 0.88$). Thus, the \acs{GBRT} prediction in step 2a is able to
+    correct the historical bias of the \acs{CMIP}5 ensemble relative to the
+    FLUXNET‐MTE product. \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:step2a_results}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6a.pdf}
+    \caption{}
+    \label{fig:06:step2a_results_biases:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6b.pdf}
+    \caption{}
+    \label{fig:06:step2a_results_biases:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6c.pdf}
+    \caption{}
+    \label{fig:06:step2a_results_biases:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/6d.pdf}
+    \caption{}
+    \label{fig:06:step2a_results_biases:d}
+  \end{subfigure}
+  \caption{Difference (a) and ratio (b) of the both biases shown in
+    \cref{fig:06:step2a_results:b,fig:06:step2a_results:c}. For panel (b), all
+    values close to $0$ for the dataset in the denominator have been masked to
+    avoid divisions by $0$. Both panels show that our approach is only to
+    first-order a bias correction, in which case both plots would only show
+    constant values. (c) Comparison of the \acf{GBRT} versus \acs{GBRT}-1D
+    projections of future \acf{GPP}. This panel indicates a clear difference
+    between the full \acs{GBRT} model using all predictors and the \acs{GBRT}
+    model using only historical \acs{GPP} as single predictor. (d) Comparison
+    of the \acs{GBRT} versus the \acf{LASSO} projections of future \acs{GPP}.
+    This panel shows that there is a clear difference between using the
+    non-linear \acs{GBRT} model and the linear \acs{LASSO} model. The results
+    of the \acs{LASSO}-1D model are not shown here because they are very
+    similar to the results of the \acs{LASSO} model.
+    \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:step2a_results_biases}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \raisebox{5mm}{\includegraphics[width=\columnwidth]{
+      ch06_paper_gpp/figs/7a.pdf}}
+    \caption{}
+    \label{fig:06:step2b_ml_inference:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/7b.pdf}
+    \caption{}
+    \label{fig:06:step2b_ml_inference:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \raisebox{12.5mm}{\includegraphics[width=\columnwidth]{
+      ch06_paper_gpp/figs/7c.pdf}}
+    \caption{}
+    \label{fig:06:step2b_ml_inference:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/7d.pdf}
+    \caption{}
+    \label{fig:06:step2b_ml_inference:d}
+  \end{subfigure}
+  \caption{(a) Box plot of the \acf{RMSEP} distributions for four different
+    statistical models used to predict the fractional change in \acf{GPP} over
+    the \nth{21} century (step 2b) using the leave‐one‐model‐out \acl{CV}
+    approach. A detailed description of the representation of these
+    distributions is given in \cref{fig:06:step2a_ml_inference:a}. The
+    \acf{GBRT} algorithm shows the minimal mean and median \acs{RMSEP}.
+    Compared to the \acs{CMIP}5 \acf{MMM}, its corresponding rescaled version
+    (r\acs{MMM}), and the linear \acf{LASSO} model, the mean \acs{RMSEP} of the
+    \acs{GBRT} model is reduced by more than $16 \unit{\%}$, $9 \unit{\%}$, and
+    $3 \unit{\%}$, respectively. (b) Relative global feature importance for
+    the different features used in the \acs{LASSO} model to predict the
+    fractional \acs{GPP} change (step 2b). The feature importance for the
+    \acs{LASSO} model is given by the normalized linear coefficients of the
+    model. (c) Local feature importance for the \acs{GBRT} model used to
+    predict fractional \acs{GPP} change (step 2b) calculated using the
+    \acf{LIME} approach \autocite{Ribeiro2016} for the three dominant features
+    \acf{T}, \acf{LAI} and \acs{GPP}. The relative influence of these three
+    features (ignoring all other features) is color-coded according to the
+    triangle in the lower left corner. Over large parts of the globe, \acs{T}
+    is the dominant feature. (d) Relative global feature importance for the
+    different features used in the \acs{GBRT} model to predict the fractional
+    \acs{GPP} change (step 2b). The blue bars for the two global feature
+    importance plots (b) and (d) correspond to negative Pearson correlation
+    coefficients between all predictors and the target variable.
+    \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:step2b_ml_inference}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8a.pdf}
+    \caption{}
+    \label{fig:06:step2b_results:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8b.pdf}
+    \caption{}
+    \label{fig:06:step2b_results:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8c.pdf}
+    \caption{}
+    \label{fig:06:step2b_results:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8d.pdf}
+    \caption{}
+    \label{fig:06:step2b_results:d}
+  \end{subfigure}
+  \caption{Fractional change in \acf{GPP} over the \nth{21} century (2100
+    versus 2000) in the \acs{RCP}8.5 scenario for different statistical models
+    (step 2b): (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change,
+    (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:xxx}, (c)
+    linear \acf{LASSO} model, and (d) \acf{GBRT} model. The geographical
+    patterns from the different statistical models are very similar and show a
+    higher \acs{GPP} increase in high latitudes and a lower \acs{GPP} closer
+    to the equator. \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:step2b_results}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/9a.pdf}
+    \caption{}
+    \label{fig:06:step2b_results_biases:a}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/9b.pdf}
+    \caption{}
+    \label{fig:06:step2b_results_biases:b}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/9c.pdf}
+    \caption{}
+    \label{fig:06:step2b_results_biases:c}
+  \end{subfigure}
+  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the
+    fractional change in \acf{GPP} over the \nth{21} century using
+    \cref{eq:06:xxx} (r\acs{MMM}; see \cref{fig:06:step2b_results:b}) and its
+    unweighted version (\acs{MMM}; see \cref{fig:06:step2b_results:a}). The
+    plot shows an almost constant value over the whole globe with a mean of
+    $0.91$, which corresponds to the ratio of the constrained global mean
+    \acs{GPP} change over the \nth{21} century from step 1 ($39 \unit{\%}$)
+    and the \acs{CMIP}5 ensemble mean global mean GPP change ($43 \unit{\%}$).
+    All values close to $0$ for the dataset in the denominator have been
+    masked to avoid divisions by $0$. Absolute (b) and relative (c)
+    differences between the fractional \acs{GPP} change over the \nth{21}
+    century predicted by the \acf{GBRT} model (see
+    \cref{fig:06:step2b_results:d}) and r\acs{MMM} (see
+    \cref{fig:06:step2b_results:b}). For panel (c), all values close to $0$ for
+    the dataset in the denominator have been masked to avoid divisions by $0$.
+    Both plots show a good agreement over large parts of the globe
+    (corresponding to values of $0$). There are large absolute differences in
+    the Sahara region and central Asia. The largest relative differences
+    (except for the Sahara and Arabian Peninsula region) appear over South
+    America, South Africa, the west coast of Africa, the Middle East, parts of
+    Australia and western parts of the United States.
+    \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:step2b_results_biases}
+\end{figure}
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10a.pdf}
+    \caption{}
+    \label{fig:06:comparison_step2a_step2b:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10b.pdf}
+    \caption{}
+    \label{fig:06:comparison_step2a_step2b:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10c.pdf}
+    \caption{}
+    \label{fig:06:comparison_step2a_step2b:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/10d.pdf}
+    \caption{}
+    \label{fig:06:comparison_step2a_step2b:d}
+  \end{subfigure}
+  \caption{Top row: Absolute future \acf{GPP} at the end of the \nth{21}
+    century (\range{2091}{2100}) calculated using step 2a (a) and step 2b (b).
+    Both approaches give similar results with global averages of $169
+    \unit{\GtCyr{}}$ (a) and $175 \unit{\GtCyr{}}$ (b), which are both
+    consistent with the global result of step 1 of
+    $\pmrangeunit{171}{12}{\GtCyr{}}$. The pattern correlation between both
+    approaches is $R^2 = 0.97$. Bottom row: Absolute (c) and relative (d)
+    differences between panels (a) and (b). \AdaptedFrom{Schlund2020}.}
+  \label{fig:06:comparison_step2a_step2b}
+\end{figure}
+
 \section{Summary and Discussion}

### 2021-02-17 end_acronyms/acronyms.tex
Cleaned text and added table to chapter 06
Added figures of chapter 06 to tex file

@@
 \begin{acronym}
+  \acro{1BGC}[1\%BGC]{biogeochemically-coupled simulation where the %
+    atmospheric CO\textsubscript{2} concentration increases by $1 \unit{\%}$ %
+    per year}
   \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
@@
   \acro{DJF}{December-January-February}
-  \acro{DLR}{%
-    Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace Center)}
+  \acro{DLR}{Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace %
+    Center)}
   \acro{ECS}{effective climate sensitivity}
@@
   \acro{IAM}{integrated assessment model}
-  \acro{IPA}{%
-    Institut für Physik der Atmosphäre (Institute of Atmospheric Physics)}
+  \acro{IPA}{Institut für Physik der Atmosphäre (Institute of Atmospheric %
+    Physics)}
   \acro{IPCC}{Intergovernmental Panel on Climate Change}
@@
   \acro{JJA}{June-July-August}
+  \acro{KUM}{Cape Kumukahi, Hawaii}
+  \acro{LAI}{leaf area index}
   \acro{LASSO}{least absolute shrinkage and selection operator}
@@
   \acro{ppm}{parts per million}
+  \acro{PR}{precipitation}
   \acro{RCP}{Representative Concentration Pathway}
@@
   \acro{RMSEP}{root-mean-square error of prediction}
+  \acro{RSDS}{downwelling solar radiation at surface}
   \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
@@
   \acro{SST}{sea surface temperature}
+  \acro{T}{near-surface air temperature}
   \acro{TCR}{transient climate response}

### 2021-02-17 ch05_paper_ecs/paper_ecs.tex
Added figures of appendix B to tex file

@@
     relationships (solid lines) for the \acs{CMIP} models (numbers; see
-    \cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models} for details) with
-    their standard prediction errors (blue and orange shaded areas; see
+    \cref{tab:app:a:cmip5_emergent_constraints_part1,%
+      tab:app:a:cmip6_emergent_constraints_part1} for details) with their
+    standard prediction errors (blue and orange shaded areas; see
     \cref{eq:02:spe}). The vertical dashed line corresponds to the
@@
   \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints
-    SHD, SHL and SHS. \AdaptedFrom{Schlund2020a}.}
+    SHD, SHL and SHS (see \cref{tab:app:a:cmip5_emergent_constraints_part1,%
+      tab:app:a:cmip6_emergent_constraints_part1} for details).
+    \AdaptedFrom{Schlund2020a}.}
   \label{fig:05:shd_shl_shs}
@@
   \end{subfigure}
-  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints
-    SU, TIH and TII. \AdaptedFrom{Schlund2020a}.}
+  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU,
+    TIH and TII (see \cref{tab:app:a:cmip5_emergent_constraints_part2,%
+      tab:app:a:cmip6_emergent_constraints_part2} for details).
+    \AdaptedFrom{Schlund2020a}.}
   \label{fig:05:su_tih_tii}
@@
   \end{subfigure}
-  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints
-    VOL and ZHA. \AdaptedFrom{Schlund2020a}.}
+  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints VOL
+    and ZHA (see \cref{tab:app:a:cmip5_emergent_constraints_part2,%
+      tab:app:a:cmip6_emergent_constraints_part2} for details).
+    \AdaptedFrom{Schlund2020a}.}
   \label{fig:05:vol_zha}

### 2021-02-17 end_appendix/appendix.tex
Added figures of appendix B to tex file

@@
   \texorpdfstring{\Cref{ch:05:paper_ecs}}{Chapter \ref{ch:05:paper_ecs}}}
-\label{sec:app:cmip_models_in_chapter_5}
+\label{sec:app:si_for_paper_ecs}
 
@@
 
+\section{Supplementary Materials for
+  \texorpdfstring{\Cref{ch:06:paper_gpp}}{Chapter \ref{ch:06:paper_gpp}}}
+\label{sec:app:si_for_paper_gpp}
+
+\begin{table}[p]
+  \centering
+  \begin{tabulary}{\columnwidth}{L L L}
+    \toprule
+    Climate model & Land model & Main reference \\
+    \midrule
+    CESM1-BGC & CLM4 & \textcite{Gent2011} \\
+    CanESM2 & CLASS2.7 + CTEM1 & \textcite{Arora2011} \\
+    GFDL-ESM2M & LM3 & \textcite{Dunne2012} \\
+    HadGEM2-ES & JULES + TRIFFID & \textcite{Collins2011} \\
+    MIROC-ESM & MATSIRO + SEIB-DGVM & \textcite{Watanabe2011} \\
+    MPI-ESM-LR & JSBACH + BETHY & \textcite{Giorgetta2013} \\
+    NorESM1-ME & CLM4 & \textcite{Iversen2013} \\
+    \bottomrule
+  \end{tabulary}
+  \caption{List of the seven \acs{CMIP}5 models used in \cref{ch:06:paper_gpp}
+    (\nameref*{ch:06:paper_gpp}) alongside the main reference. More details are
+    given by \textcite{Anav2013}. We chose all \acs{CMIP}5 models which provide
+    all necessary variables (\emph{co2}, \emph{gpp}, \emph{lai}, \emph{pr},
+    \emph{rsds} and \emph{tas}) for all used experiments (\emph{esmHistorical},
+    \emph{esmrcp85} and \emph{esmFixClim1}). For all models, we only use the
+    first ensemble member available. \AdaptedFrom{Schlund2020}.}
+  \label{tab:app:b:cmip5_models}
+\end{table}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{0.39\columnwidth}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s1a.pdf}
+    \caption{}
+    \label{fig:app:b:co2:a}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.39\columnwidth}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s1b.pdf}
+    \caption{}
+    \label{fig:app:b:co2:b}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.2\columnwidth}
+    \raisebox{14.5mm}{\includegraphics[width=\columnwidth]{
+      ch06_paper_gpp/figs/s1c.pdf}}
+  \end{subfigure}
+  \caption{(a) Monthly mean atmospheric \acs{CO2} concentration at \acf{KUM}
+    from 1979 to 2019. The thin colored lines show the individual \acs{CMIP}5
+    models (emission-driven historical simulations for the years
+    \range{1979}{2005} and emission-driven \acs{RCP}8.5 simulations for the
+    years \range{2006}{2019}; the latter is not available for HadGEM2-ES). The
+    thick black line shows the observations. For the \acs{CMIP}5 models, the
+    grid cell closest to \acs{KUM} is considered. The curves show an increase
+    of the atmospheric \acs{CO2} concentration superimposed by a pronounced
+    seasonal cycle (see \cref{subsec:02:carbon_cycle_perturbations}). (b)
+    Annual amplitude of the seasonal cycle of \acs{CO2} (defined as the
+    difference between the maximum and the minimum monthly mean atmospheric
+    \acs{CO2} concentration for each year) against the annual mean atmospheric
+    \acs{CO2} concentration at \acs{KUM}. Colored points show the \acs{CMIP}5
+    models (similar time ranges as in (a)) and thick black points show the
+    observations. The lines show the corresponding linear regression fits for
+    each dataset. The slopes of these linear fits define the sensitivity of
+    the seasonal \acs{CO2} cycle amplitude to atmospheric \acs{CO2}
+    concentrations, which is used as predictor for the emergent constraint
+    step of our approach (step 1). \AdaptedFrom{Schlund2020}.}
+  \label{fig:app:b:co2}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s2a.pdf}
+    \caption{}
+    \label{fig:app:b:residuals:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s2b.pdf}
+    \caption{}
+    \label{fig:app:b:residuals:b}
+  \end{subfigure}
+  \caption{Distribution of the residuals of the \acf{GBRT} model for the two
+    different target variables used in step 2a (absolute \acf{GPP} at the end
+    of the \nth{21} century) and step 2b (fractional \acs{GPP} change over the
+    \nth{21} century). The distributions are derived by kernel density
+    estimation using training (blue) and test (green) data. The plots show
+    approximately unbiased distributions for the training and the test
+    datasets, which are very similar to each other. This indicates that the
+    \acl{ML} model does not overfit the data. \AdaptedFrom{Schlund2020}.}
+  \label{fig:app:b:residuals}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s3a.pdf}
+    \caption{}
+    \label{fig:app:b:cmip5_hist_mte:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s3b.pdf}
+    \caption{}
+    \label{fig:app:b:cmip5_hist_mte:b}
+  \end{subfigure}
+  \caption{Geographical distributions of the historical \acf{GPP} averaged
+    between 1991 and 2000. (a) C\acs{MIP}5 \acf{MMM}. (b) FLUXNET-MTE product
+    \autocite{Jung2011}. \AdaptedFrom{Schlund2020}.}
+  \label{fig:app:b:cmip5_hist_mte}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4a.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4b.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results:b}
+  \end{subfigure}
+  \\
+    \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4c.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4d.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results:d}
+  \end{subfigure}
+  \\
+    \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4e.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results:e}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s4f.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results:f}
+  \end{subfigure}
+  \caption{Geographical distributions of the absolute \acf{GPP} at the end of
+    the \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for different
+    statistical models. (a) \acs{CMIP}5 \acf{MMM}. (b) Rescaled \acs{CMIP}5
+    \acs{MMM} using \cref{eq:06:xxx}. (c) \Acf{LASSO} model using only the
+    historical \acs{GPP} as single predictor. (d) \Acf{GBRT} model using only
+    the historical \acs{GPP} as single predictor. (e) \acs{LASSO} model using
+    all predictors. (f) \acs{GBRT} model using all predictors.
+    \AdaptedFrom{Schlund2020}.}
+  \label{fig:app:b:step2a_results}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5a.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results_errors:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5b.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results_errors:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5c.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results_errors:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5d.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results_errors:d}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5e.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results_errors:e}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s5f.pdf}
+    \caption{}
+    \label{fig:app:b:step2a_results_errors:f}
+  \end{subfigure}
+  \caption{Geographical distributions of the \acfp{SPE} of the absolute
+    \acf{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario
+    (step 2a) for different statistical models. Details on the calculation of
+    the \acs{SPE} are given in \cref{subsec:app:b:xxx}. (a) \acs{CMIP}5
+    \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:xxx}. (c)
+    \Acf{LASSO} model using only the historical \acs{GPP} as single predictor.
+    (d) \Acf{GBRT} model using only the historical \acs{GPP} as single
+    predictor. (e) \acs{LASSO} model using all predictors. (f) \acs{GBRT}
+    model using all predictors. The \acs{SPE} is minimal for the \acs{GBRT}
+    model using all predictors. \AdaptedFrom{Schlund2020}.}
+  \label{fig:app:b:step2a_results_errors}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s6a.pdf}
+    \caption{}
+    \label{fig:app:b:step2b_results_errors:a}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s6b.pdf}
+    \caption{}
+    \label{fig:app:b:step2b_results_errors:b}
+  \end{subfigure}
+  \\
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s6c.pdf}
+    \caption{}
+    \label{fig:app:b:step2b_results_errors:c}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{\SubfigureWidth{}}
+    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/s6d.pdf}
+    \caption{}
+    \label{fig:app:b:step2b_results_errors:d}
+  \end{subfigure}
+  \caption{Geographical distributions of the \acfp{SPE} of the fractional
+    change in \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario
+    (step 2b) for different statistical models. Details on the calculation of
+    the \acs{SPE} are given in \cref{subsec:app:b:xxx}. (a) \acs{CMIP}5
+    \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:xxx}. (c)
+    \Acf{LASSO} model. (d) \Acf{GBRT} model. The \acs{SPE} is minimal for the
+    \acs{GBRT} model. \AdaptedFrom{Schlund2020}.}
+  \label{fig:app:b:step2b_results_errors}
+\end{figure}
+
 \endgroup

### 2021-02-17 ch03_esmvaltool/esmvaltool.tex
Added figures of chapter 06 to tex file
Used acronym for SPE everywhere

@@
     climate model ensemble with its associated uncertainties indicated by the
-    dashed black lines (standard prediction error; see \cref{eq:02:spe}). Blue
-    lines show the observational constraint from HadCRUT4 with its best
-    estimate (dot-dashed line) and standard deviation (dashed lines). The
-    letters represent individual \acs{CMIP}5 models (see \textcite{Cox2018}
-    for details).}
+    dashed black lines (\acl{SPE}; see \cref{eq:02:spe}). Blue lines show the
+    observational constraint from HadCRUT4 with its best estimate (dot-dashed
+    line) and standard deviation (dashed lines). The letters represent
+    individual \acs{CMIP}5 models (see \textcite{Cox2018} for details).}
   \label{fig:03:cox18}
@@
   \centering
-  \begin{subfigure}[b]{0.35\columnwidth}
+  \begin{subfigure}[b]{0.36\columnwidth}
     \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942a.pdf}
@@
   \end{subfigure}
-  ~
-  \begin{subfigure}[b]{0.35\columnwidth}
+  \begin{subfigure}[b]{0.36\columnwidth}
     \includegraphics[width=\columnwidth]{ch03_esmvaltool/figs/flato13_942b.pdf}
@@
   \end{subfigure}
-  ~
-  \begin{subfigure}[b]{0.25\columnwidth}
+  \begin{subfigure}[b]{0.26\columnwidth}
     \raisebox{12.5mm}{\includegraphics[width=\columnwidth]{

### 2021-02-17 ch05_paper_ecs/paper_ecs.tex
Used acronym for SPE everywhere

@@
       tab:app:a:cmip6_emergent_constraints_part1} for details) with their
-    standard prediction errors (blue and orange shaded areas; see
-    \cref{eq:02:spe}). The vertical dashed line corresponds to the
-    observational reference with its standard error (gray shaded area). The
-    horizontal dashed lines show the best estimates of the constrained
-    \acf{ECS} and the colored circles mark the \aclp{MMM} for \acs{CMIP}5
-    (blue) and \acs{CMIP}6 (orange). (b), (d), (f) \Aclp{PDF} for \acs{ECS} of
-    the emergent constraints (solid lines; see \cref{eq:02:pdf_y_given_x0})
-    and the unconstrained model ensembles (histograms). Due to the
-    availability of data, the histograms may differ for the different emergent
-    constraints. \AdaptedFrom{Schlund2020a}.}
+    \aclp{SPE} (blue and orange shaded areas; see \cref{eq:02:spe}). The
+    vertical dashed line corresponds to the observational reference with its
+    standard error (gray shaded area). The horizontal dashed lines show the
+    best estimates of the constrained \acf{ECS} and the colored circles mark
+    the \aclp{MMM} for \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange). (b), (d),
+    (f) \Aclp{PDF} for \acs{ECS} of the emergent constraints (solid lines; see
+    \cref{eq:02:pdf_y_given_x0}) and the unconstrained model ensembles
+    (histograms). Due to the availability of data, the histograms may differ
+    for the different emergent constraints. \AdaptedFrom{Schlund2020a}.}
   \label{fig:05:bri_cox_lip}

### 2021-02-17 ch06_paper_gpp/paper_gpp.tex
Used acronym for SPE everywhere

@@
     points refer to \acs{CMIP}5 models, the orange line and shaded area to the
-    linear  regression fit and its corresponding standard prediction error and
-    the dashed black lines to the observational constraint. (a) Similar to
+    linear regression fit and its corresponding \acl{SPE} and the dashed black
+    lines to the observational constraint. (a) Similar to
     \textcite{Wenzel2016}: The global mean fractional \acs{GPP} change after

### 2021-02-18 ch06_paper_gpp/paper_gpp.tex
Added text to chapter 06

@@
 \ac{GPP} as two independent approaches and target variables. Unlike univariate
-linear regression used in the \ac{MDER} algorithm, the proposed \ac{GBRT}
-algorithm is able to handle multiple predictors and copes with non-linearities
-in the data. \ac{GBRT} is a well-known and successful tool used for
-interpolation, classification and prediction in other fields of data science
-and engineering \autocite{Death2007, Elith2008}. In the context of climate
-science, \ac{GBRT} was recently applied to identify the key drivers of spatial
-variations of the ratio of plant transpiration to total terrestrial
+linear regression used in the \ac{MDER} algorithm, the proposed
+\emph{\ac{GBRT}} algorithm is able to handle multiple predictors and copes with
+non-linearities in the data. \ac{GBRT} is a well-known and successful tool used
+for interpolation, classification and prediction in other fields of data
+science and engineering \autocite{Death2007, Elith2008}. In the context of
+climate science, \ac{GBRT} was recently applied to identify the key drivers of
+spatial variations of the ratio of plant transpiration to total terrestrial
 evapotranspiration in \acp{ESM} \autocite{Lian2018}.
@@
 
-TBA.
+Here we briefly describe the data and methods used in this study.
+\Cref{subsec:06:step_1} reviews the method of a previously published emergent
+constraint on \ac{CO2} fertilization from \textcite{Wenzel2016} that we apply
+here to our study to rescale the climate model output (step 1). This includes a
+discussion of the caveats that are associated with this emergent constraint
+that attributes changes in the observed increase in the \ac{CO2} seasonal cycle
+amplitude at \ac{KUM} entirely to the increase in atmospheric \ac{CO2}
+concentration, yet other studies provide indications that part of this increase
+is due to climate and land use changes \autocite{Bastos2019, Forkel2016,
+  Piao2018, Zhao2016}. In \cref{subsec:06:step_2} the \ac{ML} technique used in
+step 2 to constrain gridded \ac{GPP} projections is presented.
+\Cref{fig:06:schematic_steps} shows an overview of our two-step approach.
+\Cref{subsec:06:uncertainty_and_skill} describes how uncertainty and predictive
+skill are quantified. A complete list of all \acs{CMIP}5 models used in this
+study is given in \cref{tab:app:b:cmip5_models}. To apply the emergent
+constraint in step 1, we need emission-driven \acs{CMIP}5 simulations. The
+number of \acp{ESM} included is therefore seven, as in \textcite{Wenzel2016}.
+More details on the methods and the experimental setup are given in
+\cref{sec:app:si_for_paper_gpp}.
 
@@
     absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
-    independent approaches. \AdaptedFrom{Schlund2020}. \cref{eq:06:xxx}}
+    independent approaches. \AdaptedFrom{Schlund2020}.}
   \label{fig:06:schematic_steps}
 \end{figure}
+
+\subsection{Constraining the \texorpdfstring{\acs{CO2}}{CO2} Fertilization
+  Effect (Step 1)}
+\label{subsec:06:step_1}
+
+In this section, we present the observational emergent constraint on the
+\ac{CO2} fertilization effect by \textcite{Wenzel2016} which we use to globally
+rescale the models for their individual bias in the sensitivity of \ac{GPP} to
+an increase in atmospheric \ac{CO2} concentration. Our objective here is to
+present a general methodology and suggest further improvements upon this
+constraint that could be used in the future. \textcite{Wenzel2016} show a
+correlation between the increase in the \ac{CO2} seasonal cycle amplitude
+simulated by models and the projected \ac{CO2} fertilization. When combined
+with the observed trends in the \ac{CO2} seasonal cycle amplitude, the
+fractional \ac{GPP} change at the time of \ac{CO2} doubling relative to
+pre-industrial conditions in a \ac{1BGC} is constrained. The predictor in this
+study is the sensitivity of the \ac{CO2} seasonal cycle amplitude to rising
+atmospheric \ac{CO2} concentrations, defined as the slope of the linear
+regression between the \ac{CO2} seasonal cycle amplitude and the annual mean
+atmospheric \ac{CO2} concentrations (see \cref{fig:app:b:co2} for details). The
+emergent constraint is physically motivated by the hypothesis that increasing
+terrestrial \ac{GPP} is a main driver for the observed changes in the \ac{CO2}
+seasonal cycle amplitude \autocite{Gray2014, Keeling1996, Zhao2014}. This
+hypothesis is based on the fact that the seasonal cycle in atmospheric \ac{CO2}
+concentration originates from photosynthesis and decomposition processes: in
+summer, photosynthesis removes \ac{CO2} from the atmosphere; in winter,
+additional \ac{CO2} is added to the atmosphere due to the decomposition of
+organic matter \autocite{Keeling1995}. Thus, \ac{CO2} fertilization leads to an
+increase of the \acs{CO2} seasonal cycle amplitude due to increased \acs{CO2}
+removal from the atmosphere. We note that this emergent constraint implicitly
+assumes that there is no temperature-driven increase in respiration over time.
+Therefore, this approach most likely overestimates the true sensitivity of
+\ac{GPP} to an increase in the atmospheric \ac{CO2} concentration.
+
+However, we note that this emergent constraint is not undisputed. In other
+studies, the observed change in the \ac{CO2} seasonal cycle amplitude is
+attributed to other factors. Using a dynamic global vegetation model,
+\textcite{Forkel2016} find that the increase in the \ac{CO2} seasonal cycle
+amplitude above $40 \unit{\degree N}$ is mainly driven by the response of
+plants to global warming instead of \ac{CO2}. Similarly, \textcite{Piao2018}
+show that elevated atmospheric \acs{CO2} concentration and climate change
+equally contribute to the increase in the \ac{CO2} seasonal cycle in northern
+high latitudes with additional smaller contributions from air-sea carbon fluxes
+and land use changes. For their analysis, \textcite{Piao2018} use \ac{CO2}
+records from 26 stations and nine terrestrial ecosystem models. An evaluation
+of two atmospheric inversions and eleven land surface models shows contrasting
+effects of \ac{CO2} fertilization (positive) and global warming (negative) on
+the \ac{CO2} seasonal cycle amplitude increase in northern high latitudes with
+negligible contributions from land use changes \autocite{Bastos2019}. A further
+study by \textcite{Zhao2016} finds that for seven out of nine dynamic
+vegetation models, the \ac{CO2} fertilization effect is the strongest driver
+for the observed changes in the \ac{CO2} seasonal cycle amplitude, while the
+two remaining models equally attribute this to \ac{CO2} fertilization, climate
+change and land use changes. Moreover, \textcite{Winkler2019a} discuss
+limitations of carbon cycle emergent constraints related to the calculation of
+the predictor from observation-based data and the inter-model relationship. As
+a case study, they adopt a further emergent constraint by
+\textcite{Winkler2019} that uses the response of the \ac{LAI} to ambient
+\ac{CO2} instead of the \ac{CO2} seasonal cycle amplitude sensitivity to
+constrain the \ac{GPP} increase in the northern high latitudes. Furthermore, it
+has been shown that the \acs{CMIP}5 models \autocite{Graven2013} and the TRENDY
+dynamic vegetation models \autocite{Thomas2016} greatly underestimate long-term
+changes in the \ac{CO2} seasonal cycle which questions their ability to
+attribute changes in the \ac{CO2} seasonal cycle to different contributors.
+Nevertheless, we emphasize that we here present a generic methodology and use
+the results of an existing emergent constraint. Our proposed framework is
+universal and could be based on other emergent constraints in the future.
+
+In contrast to \textcite{Wenzel2016}, instead of idealized simulations we apply
+the presented emergent constraint on the emission-driven \acs{RCP}8.5
+simulation \autocite{Riahi2011} and examine whether it still holds for this
+simulation where forcings other than \ac{CO2} also change. Moreover, we
+consider the global mean \ac{GPP} change instead of the \ac{GPP} change in the
+northern extratropics as in \textcite{Wenzel2016}. Due to the strong annual
+mixing of \ac{CO2} in the atmosphere, the change in the \ac{CO2} seasonal cycle
+amplitude at \ac{KUM} is representative for the whole globe. Thus, we apply
+this emergent constraint to the global mean \ac{GPP} change. However, we note
+that this approach might introduce errors since the \ac{CO2} fertilization
+effect could be different for tropical and extratropical ecosystems due to
+nutrient limitations and the temperature dependence of Rubisco kinetics
+\autocite{CraftsBrandner2000} in the tropics. Nevertheless, we argue that we
+can use the constraint if the emergent relationship still holds within our
+considered framework (the \acs{RCP}8.5 scenario simulated by the \acs{CMIP}5
+models).
+
+Therefore, our target variable in step 1 is the global mean fractional
+\acs{GPP} change over the \nth{21} century $f$ in the \acs{RCP}8.5 scenario,
+which is calculated from the 10-year mean \ac{GPP} at the end of the \nth{21}
+century in the emission‐driven fully-coupled \acs{RCP}8.5 simulations
+(\range{2091}{2100}) and the 10-year mean \ac{GPP} at the end of the \nth{20}
+century (\range{1991}{2000}) in the emission-driven historical simulation:
+\begin{equation}
+  f = \frac{\mathtxt{GPP (\range{2091}{2100})}}{
+    \mathtxt{GPP (\range{1991}{2000})}} - 1.
+  \label{eq:06:f}
+\end{equation}
+This variable is given as a percentage, where negative values correspond to a
+decrease in \ac{GPP}, positive values to an increase in \ac{GPP} and a value of
+$0 \unit{\%}$ to no \ac{GPP} change. The \ac{CO2} amplitude sensitivity is
+calculated for the years \range{1979}{2019}, which is the full available time
+range for the observed \ac{CO2} record at \ac{KUM} \autocite{Keeling2005}.
+Similar \textcite{Wenzel2016}, we extract the observed \ac{CO2} amplitude from
+smoothed atmospheric \ac{CO2} concentrations (using a stiff cubic spline
+function plus a four-harmonics functions with linear gain) as provided by
+\textcite{Keeling2005}. For the \acs{CMIP}5 models, the respective grid cell
+closest to \ac{KUM} ($19.5 \unit{\degree N}$, $154.8 \unit{\degree W}$) is
+extracted. Additionally, the emission-driven historical simulations are
+extended with the emission-driven \acs{RCP}8.5 simulations for the years 2006
+to 2019. More details including an illustration of this calculation are shown
+in \cref{fig:app:b:co2}.
+
+Since the global constraint above cannot be used directly as a predictor in the
+\ac{ML} approach which operates on gridded values (see
+\cref{subsec:06:step_2}), we use it as first step to globally rescale the
+\ac{ESM} output. This important first step allows us to constrain the large
+uncertainty in the global \ac{CO2} fertilization effect of the \acs{CMIP}5
+models. In this study, two different gridded target variables are examined (see
+\cref{fig:06:schematic_steps}): the monthly climatology of future absolute
+\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
+\acs{RCP}8.5 scenario (see step 2a) and the fractional \acs{GPP} change over
+the \nth{21} century as defined in \cref{eq:06:f} (see step 2b). The global
+rescaling of the gridded target variables works as follows: Let $f_m$ be the
+global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
+$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change
+and $y_{mi}$ the target variable of climate model $m$. Then, we define the
+rescaled target variable $y'_{mi}$ by
+\begin{equation}
+  y'_{mi} = y_{mi} \cdot \frac{f'}{f_m},
+  \label{eq:06:y}
+\end{equation}
+where the index $i$ counts over all grid cells (and months when absolute
+\ac{GPP} is used as target variable in step 2a) and $m$ refers to the \ac{ESM}
+($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
+grid cells/months are rescaled by a factor that is determined by the ratio
+between the models' global mean \ac{GPP} change and the constrained global mean
+\ac{GPP} change derived from the \textcite{Wenzel2016} emergent constraint
+using the observed change in the \ac{CO2} cycle amplitude. Models that simulate
+a too high global mean \ac{GPP} change are nudged toward smaller values and
+vice versa. The rescaled target variables are then fed into the \ac{ML} model
+(step 2) presented in the following section. We note that this global rescaling
+assumes that the relative biases in the fractional \ac{GPP} change of the
+individual \acp{ESM} are constant over the whole globe.
 
@@
   \end{tabular}
-  \caption{Process-oriented diagnostics (\enquote{predictors} or
-    \enquote{features}) used in the \acf{GBRT} model to predict the target
-    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
-    variables are monthly climatologies of the specified time ranges in the
-    historical climate. For step 2b (target variable: fractional \acs{GPP}
-    change), the temporal mean over the specified time ranges is calculated for
-    all variables. \AdaptedFrom{Schlund2020}.}
+  \caption{Process-oriented diagnostics (\emph{predictors} or \emph{features})
+    used in the \acf{GBRT} model to predict the target variables. For step 2a
+    (target variable: absolute \acs{GPP}), all listed variables are monthly
+    climatologies of the specified time ranges in the historical climate. For
+    step 2b (target variable: fractional \acs{GPP} change), the temporal mean
+    over the specified time ranges is calculated for all variables.
+    \AdaptedFrom{Schlund2020}.}
   \label{tab:06:predictors}
@@
 
-\subsection{Constraining the \texorpdfstring{\acs{CO2}}{CO2} Fertilization
-  Effect (Step 1)}
-\label{subsec:06:step_1}
+\subsection{Gradient Boosted Regression Tree (\acs{GBRT}) Algorithm (Step 2)}
+\label{subsec:06:step_2}
+
+For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
+function $F$ relating the target variable $y$ to a set of $K$ predictors
+$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (\emph{features} or
+\emph{covariates}) such that the predicted target variable is given by $\hat{y}
+= F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001, Friedman2002}. Its
+basic elements are decision trees, which use binary splits of the input data to
+create decision rules. Due to their simple nature, decision trees are easy to
+interpret but cannot be used to create satisfying predictions for very complex
+datasets. This issue can be overcome by a technique called \emph{boosting}
+\autocite{Freund1996}. Boosting iteratively improves the performance of
+\emph{weak learners} (in our case decision trees) in the following way: At
+first, a single decision tree is fitted to a randomly selected subsample of the
+training data $\left\{ \left( \bm{x}_i, y_i \right) \mid i \in I \right\}$ with
+an arbitrary index set $I$ (using a predefined loss function, \eg{} the mean
+square error). Then, further decision trees are linearly added to the
+regression function $F$ (scaled by a regularization parameter $\nu \ll 1$
+called \emph{learning rate}) and the model is fitted again in the same manner.
+This ensures that the prediction for poorly modeled training points in the
+beginning will gradually improve during the training. The whole procedure is
+repeated until the loss function does not improve anymore, as evaluated on an
+independent validation data set. Relevant hyperparameters of this algorithm
+(maximum number of decision trees, maximum depth of the decision trees,
+learning rate and the subsample ratio) are optimized by $5$-fold \ac{CV} using
+the training data. More details on the \ac{GBRT} algorithm are given in
+\cref{subsec:app:b:gbrt}.
+
+\subsection{Assessment of Uncertainty and Predictive Skill}
+\label{subsec:06:uncertainty_and_skill}
 
@@
     mean \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in
-    the emission‐driven fully-coupled historical run (see \cref{eq:06:xxx}).
-    In contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
+    the emission‐driven fully-coupled historical run (see \cref{eq:06:f}). In
+    contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
     amplitude is calculated from the years \range{1979}{2019} for climate
@@
     \acf{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
-    \acs{RCP}8.5 scenario using \cref{eq:06:xxx} (r\acs{MMM}) and its
-    unweighted version (\acs{MMM}). The plot shows an almost constant value
-    over the whole globe with a mean of $0.92$, which corresponds to the ratio
-    of the constrained global mean \acs{GPP} change over the \nth{21} century
-    from step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
+    \acs{RCP}8.5 scenario using \cref{eq:06:y} (r\acs{MMM}) and its unweighted
+    version (\acs{MMM}). The plot shows an almost constant value over the
+    whole globe with a mean of $0.92$, which corresponds to the ratio of the
+    constrained global mean \acs{GPP} change over the \nth{21} century from
+    step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
     \acs{GPP} change ($43 \unit{\%}$). All values close to $0$ for the dataset
@@
     (step 2b): (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change,
-    (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:xxx}, (c)
+    (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:y}, (c)
     linear \acf{LASSO} model, and (d) \acf{GBRT} model. The geographical
@@
     fractional change in \acf{GPP} over the \nth{21} century using
-    \cref{eq:06:xxx} (r\acs{MMM}; see \cref{fig:06:step2b_results:b}) and its
+    \cref{eq:06:y} (r\acs{MMM}; see \cref{fig:06:step2b_results:b}) and its
     unweighted version (\acs{MMM}; see \cref{fig:06:step2b_results:a}). The

### 2021-02-18 end_appendix/appendix.tex
Added text to chapter 06

@@
 \crefalias{section}{appendix}
+\crefalias{subsection}{appendix}
 \numberwithin{equation}{section}
@@
 
+\subsection{Data Preprocessing}
+\label{subsec:app:b:data_preprocessing}
+
+The raw monthly mean output of every participating climate model and
+observation-driven dataset is regridded to a $\ang{2} \times \ang{2}$ grid and
+masked with a common mask to remove all oceans and Antarctica. For step 2a
+(target variable: absolute \acf{GPP} at the end of the \nth{21} century),
+monthly climatologies are calculated for every dataset by averaging over all
+available years for every month. For step 2b (target variable: fractional
+\ac{GPP} change over the \nth{21} century), temporal means are calculated for
+every dataset by averaging over the full time dimension. In addition, values
+greater than $300 \unit{\%}$ in the target variable in step 2b (fractional
+\ac{GPP} change over the \nth{21} century) are masked to avoid numerical
+inconsistencies caused by the division of small numbers in the derivation of
+the target variable. In the next step, the multi-dimensional data is flattened
+and all the training data from the different climate models is stacked into a
+single large training array. Finally, to account for the varying magnitudes of
+the different predictors, all of them are linearly scaled by their respective
+means and standard deviations so that they have a mean of zero and unit
+variance. In total, 237852 (16503) training data points, 79284 (5501) hold-out
+test data points, and 46344 (3727) points for the prediction are used in the
+machine learning model for step 2a (step 2b).
+
+\subsection{Gradient Boosted Regression Tree (\acs{GBRT}) Algorithm}
+\label{subsec:app:b:gbrt}
+
+The basic elements of the \ac{GBRT} algorithm are decision trees. These models
+create decision rules based on binary splits to predict a target variable $y$
+from a set of predictors $\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)}
+\right)$ ($K$ is the number of predictors). These predictors do not need to be
+of the same type: \ac{GBRT} allows the simultaneous input of numerical and
+categorical variables, which is a great advantage for our use case. There is no
+need to encode the categorical predictors in any way. Due to their simple
+nature, \acf{ML} models based on decision trees are easy to interpret and
+explain but cannot be used to create satisfying predictions for complex
+datasets. This issue can be overcome by a technique called \emph{boosting}.
+Boosting improves the performance of \emph{weak learners} (in our case decision
+trees) by combining a large number of them  \autocite{Freund1996}. The
+regression function used to predict $\hat{y} = F(\bm{x})$ can be written as a
+linear combination of simple decision trees $h(\bm{x})$
+\begin{equation}
+  F(\bm{x}) = \sum_{j=0}^{H} \beta_j h(\bm{x}, \bm{\alpha}_j),
+  \label{eq:app:b:f_x_sum}
+\end{equation}
+where $H$ is the total number of decision trees, $\beta_j$ the expansion
+coefficient for tree $j$ and $\bm{\alpha}_j$ the hyperparameters for tree $j$.
+Using all $N$ training data points $\left\{ \left( \bm{x}_i, y_i \right) \mid m
+\in I_N \right\}$ with index set $I_N = \left\{ 1, 2, \ldots, N \right\}$
+(\enquote{classic} gradient boosting), the expansion coefficients and
+parameters are jointly fitted by minimizing a loss function $L \left( y,
+F(\bm{x}) \right)$ in a forward iteration:
+\begin{equation}
+  \left( \beta_j, \bm{\alpha}_j \right) = \underset{\beta,
+    \bm{\alpha}}{\arg\min} \sum_{i=1}^{N} L \left( y_i, F_{j-1}(\bm{x}_i) +
+  \beta h(\bm{x}_i, \bm{\alpha}) \right).
+  \label{eq:app:b:beta_alpha}
+\end{equation}
+In practice, this iteration step only uses a randomly selected subsample of the
+training data (drawn without replacement), \ie{} the sum does not cover all $N$
+training points. Starting with an initial guess $F_0(\bm{x})$, the model is
+recursively built by
+\begin{equation}
+  F_j(\bm{x}) = F_{j-1}(\bm{x}) + \beta_j h(\bm{x}, \bm{\alpha}_j).
+  \label{eq:app:b:f_x_recursion}
+\end{equation}
+The minimization procedure of the loss function (mean square error with
+additional sample weights determined by the grid cell areas) is called
+\emph{stochastic gradient boosting} and is explained in detail by
+\textcite{Friedman2001, Friedman2002}. Fitting the \ac{GBRT} model involves
+building the decision trees by splitting the data at points with maximum
+information gain. Boosting those simple trees greatly improves the overall
+predictive power of the \ac{ML} algorithm: Poorly modeled training points in
+the early stages of the algorithm will gradually improve throughout the
+training process.
+
+A crucial criterion for the successful application of any \ac{GBRT} algorithm
+is the choice of several hyperparameters. The three main control parameters of
+the learning procedure are the total number of decision trees $H$, the
+complexity of the individual trees (for example measured by the maximum tree
+depth) and the learning rate $\nu \ll 1$. The latter parameter is used for
+regularization and dramatically reduces the risk of overfitting by scaling down
+the contribution of each added weak learner \autocite{Death2007, Elith2008,
+  Friedman2001}. A common way to optimize the algorithm is $n$-fold \acf{CV}
+\autocite{Bishop2006}: The data is randomly divided into a training and a
+validation dataset and the \acs{GBRT} model is fitted on the training data
+only. After that, the performance of this model can be evaluated on the
+validation dataset by a suitable metric (\eg{} the mean square error). This
+process is repeated $n$ times so that every input point is part of the
+validation set at least once. The optimal hyperparameters are the set of
+hyperparameters with optimal performance on the validation datasets
+\autocite{Elith2008}.
+
 \begin{table}[p]
@@
     statistical models. (a) \acs{CMIP}5 \acf{MMM}. (b) Rescaled \acs{CMIP}5
-    \acs{MMM} using \cref{eq:06:xxx}. (c) \Acf{LASSO} model using only the
+    \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO} model using only the
     historical \acs{GPP} as single predictor. (d) \Acf{GBRT} model using only
@@
     the \acs{SPE} are given in \cref{subsec:app:b:xxx}. (a) \acs{CMIP}5
-    \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:xxx}. (c)
+    \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c)
     \Acf{LASSO} model using only the historical \acs{GPP} as single predictor.
@@
     the \acs{SPE} are given in \cref{subsec:app:b:xxx}. (a) \acs{CMIP}5
-    \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:xxx}. (c)
+    \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c)
     \Acf{LASSO} model. (d) \Acf{GBRT} model. The \acs{SPE} is minimal for the

### 2021-02-18 ch02_scientific_background/scientific_background.tex
Used acronym for SPE everywhere
Added text to chapter 06
Added appendix B

@@
   \caption{Historical evolution of coupled climate models over the last 45
-    years. In early days, these models were so-called \acfp{AOGCM} and only
+    years. In early days, these models were so-called \acfp{AOGCM}, which only
     included three components: the atmosphere, the land surface and the ocean.
@@
     climate response uncertainty (blue) and emission uncertainty (green)
-    \autocite{Hawkins2009, Hawkins2010}. Climate response uncertainty can (b)
-    increase in newer generations of climate models when a new process is
-    discovered to be relevant or (c) decrease with additional model
+    \autocite{Hawkins2009, Hawkins2010}. Climate response uncertainty can
+    increase (b) in newer generations of climate models when a new process is
+    discovered to be relevant or decrease (c) with additional model
     improvements and observational constraints.
@@
     of the linear regression (black line) corresponds to the climate feedback
-    parameter $\lambda$ and the $y$-intercept corresponds to the radiative
+    parameter $\lambda$, and the $y$-intercept corresponds to the radiative
     forcing $F_\mathtxt{4x}$. These can be used to calculate the \acf{ECS} with
@@
 models and data $\left\{ \left( x_m, y_m \right) \mid m \in I_M \right\}$ with
-index set $I_M = \left\{ 1, 2, \ldots, M\right\}$, a linear
+index set $I_M = \left\{ 1, 2, \ldots, M \right\}$, a linear
 regression model
@@
 uncertainty of the emergent relationship for a value $x$ that has not been used
-to fit the regression line is given by the standard prediction error
-$\sigma_{\hat{y}}(x)$:
+to fit the regression line is given by the \ac{SPE} $\sigma_{\hat{y}}(x)$:
 \begin{equation}
@@
     lines visualize the distance between the climate model data and the linear
-    regression surface and represent the noise terms $\bm{\varepsilon}$ in
+    regression surface, which represents the noise term $\bm{\varepsilon}$ in
     \cref{eq:02:multivariate_linear_model}.}

### 2021-02-18 ch03_esmvaltool/esmvaltool.tex
Added appendix B

@@
     variability \autocite{Cox2018}. Magenta colors indicate high-sensitivity
-    climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$) and green colors
+    climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$), and green colors
     indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}

### 2021-02-18 ch05_paper_ecs/paper_ecs.tex
Added appendix B

@@
     standard error (gray shaded area). The horizontal dashed lines show the
-    best estimates of the constrained \acf{ECS} and the colored circles mark
+    best estimates of the constrained \acf{ECS}, and the colored circles mark
     the \aclp{MMM} for \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange). (b), (d),
@@
     subset of \acs{CMIP}5 models. The vertical dashed line and shaded area
-    correspond to the observational reference and the horizontal dashed lines
+    correspond to the observational reference, and the horizontal dashed lines
     show the corresponding constraints on the \acf{ECS} using this observation.

### 2021-02-18 ch06_paper_gpp/paper_gpp.tex
Added appendix B

@@
     points refer to \acs{CMIP}5 models, the orange line and shaded area to the
-    linear regression fit and its corresponding \acl{SPE} and the dashed black
+    linear regression fit and its corresponding \acl{SPE}, and the dashed black
     lines to the observational constraint. (a) Similar to
@@
     (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:y}, (c)
-    linear \acf{LASSO} model, and (d) \acf{GBRT} model. The geographical
+    linear \acf{LASSO} model and (d) \acf{GBRT} model. The geographical
     patterns from the different statistical models are very similar and show a

### 2021-02-18 end_appendix/appendix.tex
Added appendix B

@@
 coefficient for tree $j$ and $\bm{\alpha}_j$ the hyperparameters for tree $j$.
-Using all $N$ training data points $\left\{ \left( \bm{x}_i, y_i \right) \mid m
+Using all $N$ training data points $\left\{ \left( \bm{x}_i, y_i \right) \mid i
 \in I_N \right\}$ with index set $I_N = \left\{ 1, 2, \ldots, N \right\}$
-(\enquote{classic} gradient boosting), the expansion coefficients and
+(\enquote{classic} gradient boosting), the expansion coefficients and tree
 parameters are jointly fitted by minimizing a loss function $L \left( y,
@@
 \autocite{Elith2008}.
+
+\subsection{Evaluation of Prediction Uncertainty}
+\label{subsec:app:b:uncertainty}
+
+We estimate the \acf{SPE} of the \ac{GBRT} model itself as the \ac{RMSE} of the
+predicted target variable $\hat{y}$ and the corresponding true values $y$ of a
+hold-out test dataset, the so-called \emph{\acf{RMSEP}} \autocite{Bishop2006},
+which is assumed to be constant for all prediction input points:
+\begin{equation}
+  \sigma_\mathtxt{GBRT} = \mathtxt{RMSE}(\hat{y}, y).
+  \label{eq:app:b:sigma_gbrt}
+\end{equation}
+For this, we randomly select $25 \unit{\%}$ of the input data prior to
+training, so that this part of the data neither enters the training of the
+\ac{GBRT} model nor the hyperparameter optimization process. Moreover, the test
+dataset allows an assessment of the prediction residuals, which is useful to
+detect overfitting (see \cref{fig:app:b:residuals}).
+
+A second source of uncertainty is the error in the rescaling of the target
+variable in step 1 of our approach. Analogous to \cref{eq:06:y}, we estimate
+this error as
+\begin{equation}
+  \sigma_{i, \mathtxt{RESC}} = \bar{y}_i \cdot \frac{\sigma_{f'}}{\bar{f}}
+  \label{eq:app:b:sigma_rescaling}
+\end{equation}
+for each prediction input point $i$ ($i$ runs over all grid cells and months).
+$\bar{y}_i$ is the \acs{CMIP}5 \acf{MMM} of the target variable at prediction
+input point $i$ (step 2a: absolute \ac{GPP} at the end of the \nth{21} century;
+step 2b: fractional \ac{GPP} change over the \nth{21} century), $\sigma_{f'}$
+the standard error in the global mean fractional \acs{GPP} change given by the
+emergent constraint from step 1 and $\bar{f}$ the \acs{CMIP}5 \ac{MMM} global
+mean fractional \ac{GPP} change over the \nth{21} century.
+
+The final source of uncertainty is the error of the prediction input data
+$\sigma_{ik}$ ($i$ corresponds to the prediction input point and $k$ to the
+predictor). These are only available for the FLUXNET-MTE product
+\autocite{Jung2011}. To account for this, we use the \acf{LIME} technique
+\commentcite{Ribeiro2016}{see \cref{subsec:06:step_2}} to build a local linear
+model for every sample point, which yields the linear coefficients $b_{ik}$.
+Using error propagation for all predictors and assuming independence of all
+individual errors, the \ac{SPE} due to observational uncertainty $\sigma_{i,
+  \mathtxt{OBS}}$ then be calculated as
+\begin{equation}
+  \sigma_{i, \mathtxt{OBS}} = \sum_{k=1}^{K} b_{ik}^2 \sigma_{ik}^2.
+  \label{eq:app:b:sigma_obs}
+\end{equation}
+The total \ac{SPE} at a prediction input point $i$ is the sum of the squared
+errors presented above (assuming all of them are independent):
+\begin{equation}
+  \sigma_i^2 = \sigma_\mathtxt{GBRT}^2 + \sigma_{i, \mathtxt{RESC}}^2 +
+  \sigma_{i, \mathtxt{OBS}}^2.
+  \label{eq:app:b:sigma_sum}
+\end{equation}
+The specified error ranges for the \ac{MMM} approaches are calculated in a
+similar way: The constant \ac{SPE} per grid cell is estimated by the mean
+\ac{RMSEP} given in the leave-one-model-out \ac{CV} experiment (see
+\cref{subsec:02:model_weighting}). For the plain \ac{MMM}, this is the only
+source of uncertainty. For the rescaled \ac{MMM}, the total error can be
+calculated similarly to \cref{eq:app:b:sigma_sum} without the last term
+(observational uncertainty).
+
+\subsection{Evaluation of Residuals}
+\label{subsec:app:b:residuals}
+
+A convenient way to gain information about statistical models is to analyze the
+residuals $\varepsilon_i$, which are defined as the difference between the true
+value of the target variable $y_i$ and the predicted value $\hat{y}_i$ at a
+sample point $i$ with known ground truth:
+\begin{equation}
+  \varepsilon_i = y_i - \hat{y}_i.
+  \label{eq:app:b:residuals}
+\end{equation}
+A common way to visualize the residuals is to plot their \acfp{PDF} (see
+\cref{fig:app:b:residuals}). The two panels (for steps 2a and 2b) show that the
+\ac{ML} model is not overfitting in both cases: The distributions of the
+training data and the independent hold-out data are very similar. Moreover, the
+\acp{PDF} do not show significant biases, as the residuals are approximately
+unbiased (zero mean) for the training and the test dataset. This justifies the
+use of the \ac{RMSEP} to estimate the \ac{SPE}, since for unbiased residuals
+the \ac{RMSEP} is equal to the standard deviation of the residuals.
 
@@
     \acs{CO2} concentration at \acs{KUM}. Colored points show the \acs{CMIP}5
-    models (similar time ranges as in (a)) and thick black points show the
+    models (similar time ranges as in (a)), and thick black points show the
     observations. The lines show the corresponding linear regression fits for
@@
     (step 2a) for different statistical models. Details on the calculation of
-    the \acs{SPE} are given in \cref{subsec:app:b:xxx}. (a) \acs{CMIP}5
+    the \acs{SPE} are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5
     \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c)
@@
     (step 2b) for different statistical models. Details on the calculation of
-    the \acs{SPE} are given in \cref{subsec:app:b:xxx}. (a) \acs{CMIP}5
+    the \acs{SPE} are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5
     \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c)

### 2021-02-19 ch01_introduction/introduction.tex
Finished section 6.1

--- 2021-02-01 ch01_introduction/introduction.tex
@@
 \section{Key Science Questions}
-\label{sec:01:research_questions}
+\label{sec:01:key_science_questions}
 
-This section will present the \textbf{AWESOME} key science questions!!!
+This \namecref{sec:01:key_science_questions} will present the \textbf{AWESOME}
+key science questions!!!
 
@@
 contributions to these studies (including their full reference) is given in the
-corresponding chapters.
+corresponding \namecrefs{ch:03:esmvaltool}.
 
@@
 climate projections, see \cref{subsec:02:emergent_constraints}) on the \ac{ECS}
-for these \acp{ESM}. The contents of this chapter are published in \emph{Earth
-  System Dynamics} \autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a
-new method to reduce uncertainties in gridded multi-model projections of the
-future climate with observations based on \ac{ML}. As an example, the method is
-applied to the photosynthesis rate at the end of the \nth{21} century, which is
-already published in the \emph{Journal of Geophysical Research: Biogeosciences}
+for these \acp{ESM}. The contents of this \namecref{ch:05:paper_ecs} are
+published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.
+\Cref{ch:06:paper_gpp} focuses on a new method to reduce uncertainties in
+multi-dimensional (gridded) multi-model projections of the future climate with
+observations based on \ac{ML}. As an example, the method is applied to the
+photosynthesis rate at the end of the \nth{21} century, which is already
+published in the \emph{Journal of Geophysical Research: Biogeosciences}
 \autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook} provides a

### 2021-02-19 ch02_scientific_background/scientific_background.tex
Finished section 6.1

@@
 
-This chapter introduces the scientific background of this thesis. First, basic
-concepts of climate model simulations and associated uncertainties are
-introduced (\cref{sec:02:esms}). Next, important metrics describing climate
-sensitivity (\cref{sec:02:climate_sensitivity}) and fundamental biogeochemical
-processes of the global carbon cycle (\cref{sec:02:esms}) are presented.
-Finally, state-of-the art techniques that can be used to reduce uncertainties
-in projections of the future climate with observations are shown
+This \namecref{ch:02:scientific_background} introduces the scientific
+background of this thesis. First, basic concepts of climate model simulations
+and associated uncertainties are introduced (\cref{sec:02:esms}). Next,
+important metrics describing climate sensitivity
+(\cref{sec:02:climate_sensitivity}) and fundamental biogeochemical processes of
+the global carbon cycle (\cref{sec:02:esms}) are presented. Finally,
+state-of-the art techniques that can be used to reduce uncertainties in
+projections of the future climate with observations are shown
 (\cref{sec:02:reducing_uncertainties}). These methods form the basis for new
@@
 about the specific variables and experiments analyzed is given in the
-corresponding chapters.
+corresponding \namecrefs{ch:04:papers_ecs_tcr_assessment}.
 
@@
 like volcanic eruptions and variations in the solar activity. Natural
-variability can be seen as the \emph{noise} in the climate record as opposed to
-the anthropogenic \emph{signal} \autocite{Cubasch2013}. As illustrated by
+variability can be seen as the \enquote{noise} in the climate record as opposed
+to the anthropogenic \enquote{signal} \autocite{Cubasch2013}. As illustrated by
 \cref{fig:02:sources_of_uncertainty}, the uncertainty associated with natural
@@
 response uncertainty when a new process is discovered to be relevant
-\commentcite{Cubasch2013}{see \cref{fig:02:sources_of_uncertainty}b}. However,
-such an increase corresponds to a previously unmeasured uncertainty. An example
-for this has recently happened in \acs{CMIP}6: most likely due to changes in
-the cloud representation of the models the spread in the projected \ac{GSAT}
-caused by a doubling of the atmospheric \ac{CO2} concentration has
-substantially increased in \acs{CMIP}6 compared to older \ac{CMIP} generations
+\commentcite{Cubasch2013}{see
+  \cref{fig:02:sources_of_uncertainty}{\color{BrownRed} b}}. However, such an
+increase corresponds to a previously unmeasured uncertainty. An example for
+this has recently happened in \acs{CMIP}6: most likely due to changes in the
+cloud representation of the models the spread in the projected \ac{GSAT} caused
+by a doubling of the atmospheric \ac{CO2} concentration has substantially
+increased in \acs{CMIP}6 compared to older \ac{CMIP} generations
 \autocite{Zelinka2020}. On the other hand, the climate response uncertainty can
 decrease with additional model improvements and better understanding of the
-Earth system (see \cref{fig:02:sources_of_uncertainty}c). Moreover, it can also
-be reduced by observational constraints, which is the main topic of this
-thesis.
+Earth system (see \cref{fig:02:sources_of_uncertainty}{\color{BrownRed} c}).
+Moreover, it can also be reduced by observational constraints, which is the
+main topic of this thesis.
 
@@
 cycle--related processes \commentcite{Schlund2020}{see \cref{ch:06:paper_gpp}},
-this chapter introduces the scientific background of the global carbon cycle.
+this \namecref{sec:02:carbon_cycle} introduces the scientific background of the
+global carbon cycle.
 
@@
 Thus, more sophisticated techniques are necessary to evaluate multi-model
-climate projections. This section introduces three state-of-the-art methods to
-assess multi-model projections and reduce associated uncertainties with
-observations. These techniques form the baseline of the work presented in the
-following chapters.
+climate projections. This \namecref{sec:02:reducing_uncertainties} introduces
+three state-of-the-art methods to assess multi-model projections and reduce
+associated uncertainties with observations. These techniques form the baseline
+of the work presented in the following \namecrefs{ch:05:paper_ecs}.
 
@@
 
-As indicated at the beginning of this section, one main issue of multi-model
-ensembles is that not all participating climate models are equally plausible.
-Usually, this is quantified with some kind of measure of the models'
-\emph{performance}, \ie{} their agreement with observations of the real climate
-system. However, this model performance can only be evaluated against
-observations of the past and present-day climate, which does not necessarily
-provide insights into the quality of model projections of the future climate.
+As indicated at the beginning of this \namecref{sec:02:reducing_uncertainties},
+one main issue of multi-model ensembles is that not all participating climate
+models are equally plausible. Usually, this is quantified with some kind of
+measure of the models' \emph{performance}, \ie{} their agreement with
+observations of the real climate system. However, this model performance can
+only be evaluated against observations of the past and present-day climate,
+which does not necessarily provide insights into the quality of model
+projections of the future climate.
 
@@
 with normalized weights $w_m$. Similar to the notation introduced in the
-previous section, $y$ is a target variable (\eg{} a projection of the future
-climate) and $m$ indexes the $M$ different climate models. To address two major
-issues of model democracy (different climate models are not equally plausible
-and not independent; see beginning of \cref{sec:02:reducing_uncertainties}),
-\textcite{Knutti2017a} propose a weighting scheme based on climate model
-performance and interdependence with the following weights:
+previous \namecref{subsec:02:emergent_constraints}, $y$ is a target variable
+(\eg{} a projection of the future climate) and $m$ indexes the $M$ different
+climate models. To address two major issues of model democracy (different
+climate models are not equally plausible and not independent; see beginning of
+\cref{sec:02:reducing_uncertainties}), \textcite{Knutti2017a} propose a
+weighting scheme based on climate model performance and interdependence with
+the following weights:
 \begin{equation}
@@
 the known ground truth of the pseudo-observation, which is also known as
-\ac{RMSEP}. The whole process is repeated for every climate model of the
+\emph{\ac{RMSEP}}. The whole process is repeated for every climate model of the
 ensemble to get a statistical distribution of \acp{RMSEP}. Finally, different
@@
 for this choice of the diagnostics is similar to the one presented in the
-previous section: Weighting schemes should address the question \enquote{which
-  climate model is adequate for predicting the target variable?} and not
-\enquote{which climate model is the best?}.
+previous \namecref{subsec:02:model_weighting}: Weighting schemes should address
+the question \enquote{which climate model is adequate for predicting the target
+  variable?} and not \enquote{which climate model is the best?}.
 

### 2021-02-19 ch03_esmvaltool/esmvaltool.tex
Finished section 6.1

@@
 In order to answer the key science questions of this thesis posed in
-\cref{sec:01:research_questions}, a reliable and efficient tool to read,
+\cref{sec:01:key_science_questions}, a reliable and efficient tool to read,
 process and evaluate climate model output and observational data is necessary.
@@
   Weigel2020}. After a brief overview of the \ac{ESMValTool}'s structure
-(\cref{sec:03:esmvaltool}), this chapter presents all these contributions that
-are not documented in other chapters of this thesis
+(\cref{sec:03:esmvaltool}), this \namecref{ch:03:esmvaltool} presents all these
+contributions that are not documented in other
+\namecrefs{ch:04:papers_ecs_tcr_assessment} of this thesis
 (\cref{sec:03:contributions_to_esmvalcore,sec:03:contributions_to_esmvaltool}).
@@
 \cref{tab:03:changes_to_esmvaltool} for a summary). Some of these include the
-analyses performed in other chapters of this thesis, namely
-\emph{recipe\_bock20jgr.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
-\emph{recipe\_meehl20sciadv.yml} (\cref{ch:04:papers_ecs_tcr_assessment}),
-\emph{recipe\_schlund20esd.yml} (\cref{ch:05:paper_ecs}) and
-\emph{recipe\_schlund20jgr\_*.yml} (\cref{ch:06:paper_gpp}).
-\emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml} contain diagnostic scripts
-that are able to calculate the climate metrics \ac{ECS} and \ac{TCR} for
-arbitrary model output (see \cref{sec:02:climate_sensitivity}). Corresponding
-plots are shown in \cref{fig:02:gregory_regression,fig:02:tcr}.
-\emph{recipe\_cox18nature.yml} reproduces the analysis of \textcite{Cox2018},
-who introduce an emergent constraint in \ac{ECS} based on a global temperature
-variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
-years for \acs{CMIP}5 models and observations form HadCRUT4
-\autocite{Morice2012}. Since high-sensitivity climate models exhibit higher
-values of $\Psi$ than low sensitivity models, an emergent relationship between
-\ac{ECS} and $\Psi$ can be established (\cref{fig:03:cox18:b}). With
-observations of $\Psi$, \ac{ECS} can eventually be constrained to $\left( 2.8
-\pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). More details on
-this including the calculation of $\Psi$ are given in \cref{subsec:05:cox}.
+analyses performed in other \namecrefs{ch:04:papers_ecs_tcr_assessment} of this
+thesis, namely \emph{recipe\_bock20jgr.yml}
+(\cref{ch:04:papers_ecs_tcr_assessment}), \emph{recipe\_meehl20sciadv.yml}
+(\cref{ch:04:papers_ecs_tcr_assessment}), \emph{recipe\_schlund20esd.yml}
+(\cref{ch:05:paper_ecs}) and \emph{recipe\_schlund20jgr\_*.yml}
+(\cref{ch:06:paper_gpp}). \emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml}
+contain diagnostic scripts that are able to calculate the climate metrics
+\ac{ECS} and \ac{TCR} for arbitrary model output (see
+\cref{sec:02:climate_sensitivity}). Corresponding plots are shown in
+\cref{fig:02:gregory_regression,fig:02:tcr}. \emph{recipe\_cox18nature.yml}
+reproduces the analysis of \textcite{Cox2018}, who introduce an emergent
+constraint in \ac{ECS} based on a global temperature variability metric $\Psi$.
+\Cref{fig:03:cox18:a} shows $\Psi$ over the past 85 years for \acs{CMIP}5
+models and observations form HadCRUT4 \autocite{Morice2012}. Since
+high-sensitivity climate models exhibit higher values of $\Psi$ than low
+sensitivity models, an emergent relationship between \ac{ECS} and $\Psi$ can be
+established (\cref{fig:03:cox18:b}). With observations of $\Psi$, \ac{ECS} can
+eventually be constrained to $\left( 2.8 \pm 0.6 \right) \unit{K}$ ($66
+\unit{\%}$ confidence range). More details on this including the calculation of
+$\Psi$ are given in \cref{subsec:05:cox}.
 

### 2021-02-19 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added figures of appendix B to tex file
Finished section 6.1

@@
 contributed all figures and tables, code to produce these with the
-\ac{ESMValTool} and text for the manuscript. This chapter first presents the
-evaluation of climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6 ensembles
+\ac{ESMValTool} and text for the manuscript. This
+\namecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
+climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6 ensembles
 (\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding \ac{ECS} and
@@
     \dataset & $\idx$ & $\ecs$ & $\tcr$}
-  \caption{\acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
+  \caption{\Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
     Details on the calculation of \acs{ECS} and \acs{TCR} are given in
@@
     over all climate models (using the normalization $1 / M$, where $M$ is the
-    number of models). \AdaptedFrom{Meehl2020}.}
+    number of models). Corresponding references for each model are given in
+    \cref{tab:app:a:cmip5_models}. \AdaptedFrom{Meehl2020}.}
   \label{tab:04:ecs_tcr_cmip5}
@@
   \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.
-    \AdaptedFrom{Meehl2020}.}
+    Corresponding references for each model are given in
+    \cref{tab:app:a:cmip6_models}. \AdaptedFrom{Meehl2020}.}
   \label{tab:04:ecs_tcr_cmip6}
@@
 
-To illustrate the results of the previous section and put them into historical
-context, \cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS}
-and \ac{TCR} ranges over the years from the Charney report
-\autocite{Charney1979} and the different \acp{AR} of the \ac{IPCC} in
-combination with the corresponding modeled ranges from the different \ac{CMIP}
-generations. Since the Charney report in 1979, the assessed range of \ac{ECS}
-of $\range{1.5}{4.5} \unit{K}$ has almost remained unchanged during the last 40
-years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
+To illustrate the results of the previous
+\namecref{sec:04:evaluation_ecs_and_tcr} and put them into historical context,
+\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
+\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
+and the different \acp{AR} of the \ac{IPCC} in combination with the
+corresponding modeled ranges from the different \ac{CMIP} generations. Since
+the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
+\unit{K}$ has almost remained unchanged during the last 40 years
+\autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
   Stocker2013} with the exception of \acs{AR}4, where the lower bound was
@@
 important questions for the entire climate modeling community. For this reason,
-this section discusses possible reasons for the increased climate sensitivity
-in \acs{CMIP}6.
+this \namecref{sec:04:possible_reasons_high_ecs_cmip6} discusses possible
+reasons for the increased climate sensitivity in \acs{CMIP}6.
 

### 2021-02-19 ch05_paper_ecs/paper_ecs.tex
Finished section 6.1

@@
 The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
-previous chapter begs the question about the interpretation and credibility of
-these results. Since the \acs{CMIP}6 models provide crucial input for the
-upcoming international climate assessment of the \ac{IPCC} (\acs{AR}6), this is
-not only important for the scientific community, but also directly affects
-society and policymakers. Due to the many new and improved processes included
-to the models, the \acs{CMIP}6 ensemble is arguably more realistic than its
-predecessor generations. However, does this automatically make results derived
-from \acs{CMIP}6 models more trustworthy than those derived from earlier
-\ac{CMIP} models? In general, the answer to this question is \enquote{no}. Even
-the \acs{CMIP}6 models are far from fully simulating all aspects and processes
-of the full vastly complex Earth system, which could introduce new biases which
-have not been present in earlier \ac{CMIP} models. For example, the apparent
-increase in climate sensitivity in the \acs{CMIP}6 models might be wrongly
-caused by missing processes that cancel out the effects of the newly included
-processes. Thus, to assess the credibility of \ac{ESM} results it is vital to
-take other sources of information into account.
+previous \namecref{ch:04:papers_ecs_tcr_assessment} begs the question about the
+interpretation and credibility of these results. Since the \acs{CMIP}6 models
+provide crucial input for the upcoming international climate assessment of the
+\ac{IPCC} (\acs{AR}6), this is not only important for the scientific community,
+but also directly affects society and policymakers. Due to the many new and
+improved processes included to the models, the \acs{CMIP}6 ensemble is arguably
+more realistic than its predecessor generations. However, does this
+automatically make results derived from \acs{CMIP}6 models more trustworthy
+than those derived from earlier \ac{CMIP} models? In general, the answer to
+this question is \enquote{no}. Even the \acs{CMIP}6 models are far from fully
+simulating all aspects and processes of the full vastly complex Earth system,
+which could introduce new biases which have not been present in earlier
+\ac{CMIP} models. For example, the apparent increase in climate sensitivity in
+the \acs{CMIP}6 models might be wrongly caused by missing processes that cancel
+out the effects of the newly included processes. Thus, to assess the
+credibility of \ac{ESM} results it is vital to take other sources of
+information into account.
 
@@
   \caption{Overview of the eleven emergent constraints on the \acf{ECS}
-    evaluated in this chapter. Detailed descriptions of the variables used to
-    calculate the \xaxis{} values of the emergent constraints and the
-    references for the corresponding observations (fourth column) are given in
+    evaluated in this \namecref{ch:05:paper_ecs}. Detailed descriptions of the
+    variables used to calculate the \xaxis{} values of the emergent
+    constraints and the references for the corresponding observations (fourth
+    column) are given in
     \cref{tab:app:a:overview_variables,tab:app:a:observations}, respectively.
@@
 code to reproduce this study with all figures and tables using the
-\ac{ESMValTool}. In this chapter, we first introduce the data (climate model
-output and observations) and methods used in this study
+\ac{ESMValTool}. In this \namecref{ch:05:paper_ecs}, we first introduce the
+data (climate model output and observations) and methods used in this study
 (\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
@@
 
-In this section we describe and discuss the eleven emergent constraints on
-\ac{ECS} summarized in \cref{tab:05:overview_emergent_constraints} using
-\acs{CMIP}5 and \acs{CMIP}6 data
+In this \namecref{sec:05:comparison_of_emergent_constraints} we describe and
+discuss the eleven emergent constraints on \ac{ECS} summarized in
+\cref{tab:05:overview_emergent_constraints} using \acs{CMIP}5 and \acs{CMIP}6
+data
 (\cref{subsec:05:bri,subsec:05:cox,subsec:05:lip,subsec:05:shd,subsec:05:shs,%
@@
 
-As shown in the previous sections, most emergent relationships show smaller
-coefficients of determination when evaluated on the new \acs{CMIP}6 ensemble
-compared to the \acs{CMIP}5 ensemble. In this section, we discuss possible
-reasons for, and implications of, these differences. As reported by
-\textcite{Caldwell2014}, the large amount of data provided by modern \acp{ESM}
-can generate spurious correlations of variables between past climate and
-\ac{ECS} just by chance, especially when only a small number of climate models
-is considered. This would cause the performance of the emergent constraint to
-be reduced on out-of-sample data (like the new \acs{CMIP}6 ensemble), since the
-emergent relationship appeared just by chance and not because of a physically
-based mechanism.
+As shown in the previous \namecrefs{sec:05:comparison_of_emergent_constraints},
+most emergent relationships show smaller coefficients of determination when
+evaluated on the new \acs{CMIP}6 ensemble compared to the \acs{CMIP}5 ensemble.
+In this \namecref{sec:05:discussion}, we discuss possible reasons for, and
+implications of, these differences. As reported by \textcite{Caldwell2014}, the
+large amount of data provided by modern \acp{ESM} can generate spurious
+correlations of variables between past climate and \ac{ECS} just by chance,
+especially when only a small number of climate models is considered. This would
+cause the performance of the emergent constraint to be reduced on out-of-sample
+data (like the new \acs{CMIP}6 ensemble), since the emergent relationship
+appeared just by chance and not because of a physically based mechanism.
 

### 2021-02-19 ch06_paper_gpp/paper_gpp.tex
Finished section 6.1

@@
 
-As argued in detail in the previous chapter, single-process-oriented emergent
-constraints do not seem to be beneficial in constraining climate projections
-from the vastly complex modern-day \acp{ESM} since they might overly simplify
-the problem. To overcome this issue, we explore an alternative approach that
-expands the \acf{MDER} method with a supervised \acf{ML} algorithm. In contrast
-to emergent constraints, this approach considers (1) multiple process-based
+As argued in detail in the previous \namecref{ch:05:paper_ecs},
+single-process-oriented emergent constraints do not seem to be beneficial in
+constraining climate projections from the vastly complex modern-day \acp{ESM}
+since they might overly simplify the problem. To overcome this issue, we
+explore an alternative approach that expands the \acf{MDER} method (see
+\cref{subsec:02:mder}) with a supervised \acf{ML} algorithm. In contrast to
+emergent constraints, this approach considers (1) multiple process-based
 predictors, (2) multi-dimensional (gridded) target variables and (3) non-linear
@@
 code to reproduce this study with all figures and tables using the
-\ac{ESMValTool}. \Cref{sec:06:data_and_methods} of this chapter provides an
-overview of the data and methods used in this paper. The results are presented
-in \cref{sec:06:results} and \cref{sec:06:summary_and_discussion} closes with a
-summary and discussion.
+\ac{ESMValTool}. \Cref{sec:06:data_and_methods} of this
+\namecref{ch:06:paper_gpp} provides an overview of the data and methods used in
+this paper. The results are presented in \cref{sec:06:results} and
+\cref{sec:06:summary_and_discussion} closes with a summary and discussion.
 
@@
 
-In this section, we present the observational emergent constraint on the
-\ac{CO2} fertilization effect by \textcite{Wenzel2016} which we use to globally
-rescale the models for their individual bias in the sensitivity of \ac{GPP} to
-an increase in atmospheric \ac{CO2} concentration. Our objective here is to
-present a general methodology and suggest further improvements upon this
-constraint that could be used in the future. \textcite{Wenzel2016} show a
-correlation between the increase in the \ac{CO2} seasonal cycle amplitude
-simulated by models and the projected \ac{CO2} fertilization. When combined
-with the observed trends in the \ac{CO2} seasonal cycle amplitude, the
-fractional \ac{GPP} change at the time of \ac{CO2} doubling relative to
-pre-industrial conditions in a \ac{1BGC} is constrained. The predictor in this
-study is the sensitivity of the \ac{CO2} seasonal cycle amplitude to rising
-atmospheric \ac{CO2} concentrations, defined as the slope of the linear
-regression between the \ac{CO2} seasonal cycle amplitude and the annual mean
-atmospheric \ac{CO2} concentrations (see \cref{fig:app:b:co2} for details). The
-emergent constraint is physically motivated by the hypothesis that increasing
-terrestrial \ac{GPP} is a main driver for the observed changes in the \ac{CO2}
-seasonal cycle amplitude \autocite{Gray2014, Keeling1996, Zhao2014}. This
-hypothesis is based on the fact that the seasonal cycle in atmospheric \ac{CO2}
-concentration originates from photosynthesis and decomposition processes: in
-summer, photosynthesis removes \ac{CO2} from the atmosphere; in winter,
-additional \ac{CO2} is added to the atmosphere due to the decomposition of
-organic matter \autocite{Keeling1995}. Thus, \ac{CO2} fertilization leads to an
-increase of the \acs{CO2} seasonal cycle amplitude due to increased \acs{CO2}
-removal from the atmosphere. We note that this emergent constraint implicitly
-assumes that there is no temperature-driven increase in respiration over time.
-Therefore, this approach most likely overestimates the true sensitivity of
-\ac{GPP} to an increase in the atmospheric \ac{CO2} concentration.
+In this \namecref{subsec:06:step_1}, we present the observational emergent
+constraint on the \ac{CO2} fertilization effect by \textcite{Wenzel2016} which
+we use to globally rescale the models for their individual bias in the
+sensitivity of \ac{GPP} to an increase in atmospheric \ac{CO2} concentration.
+Our objective here is to present a general methodology and suggest further
+improvements upon this constraint that could be used in the future.
+\textcite{Wenzel2016} show a correlation between the increase in the \ac{CO2}
+seasonal cycle amplitude simulated by models and the projected \ac{CO2}
+fertilization. When combined with the observed trends in the \ac{CO2} seasonal
+cycle amplitude, the fractional \ac{GPP} change at the time of \ac{CO2}
+doubling relative to pre-industrial conditions in a \ac{1BGC} is constrained.
+The predictor in this study is the sensitivity of the \ac{CO2} seasonal cycle
+amplitude to rising atmospheric \ac{CO2} concentrations, defined as the slope
+of the linear regression between the \ac{CO2} seasonal cycle amplitude and the
+annual mean atmospheric \ac{CO2} concentrations (see \cref{fig:app:b:co2} for
+details). The emergent constraint is physically motivated by the hypothesis
+that increasing terrestrial \ac{GPP} is a main driver for the observed changes
+in the \ac{CO2} seasonal cycle amplitude \autocite{Gray2014, Keeling1996,
+  Zhao2014}. This hypothesis is based on the fact that the seasonal cycle in
+atmospheric \ac{CO2} concentration originates from photosynthesis and
+decomposition processes: in summer, photosynthesis removes \ac{CO2} from the
+atmosphere; in winter, additional \ac{CO2} is added to the atmosphere due to
+the decomposition of organic matter \autocite{Keeling1995}. Thus, \ac{CO2}
+fertilization leads to an increase of the \acs{CO2} seasonal cycle amplitude
+due to increased \acs{CO2} removal from the atmosphere. We note that this
+emergent constraint implicitly assumes that there is no temperature-driven
+increase in respiration over time. Therefore, this approach most likely
+overestimates the true sensitivity of \ac{GPP} to an increase in the
+atmospheric \ac{CO2} concentration.
 
@@
 vice versa. The rescaled target variables are then fed into the \ac{ML} model
-(step 2) presented in the following section. We note that this global rescaling
-assumes that the relative biases in the fractional \ac{GPP} change of the
-individual \acp{ESM} are constant over the whole globe.
+(step 2) presented in the following \namecref{subsec:06:step_2}. We note that
+this global rescaling assumes that the relative biases in the fractional
+\ac{GPP} change of the individual \acp{ESM} are constant over the whole globe.
+
+\subsection{Gradient Boosted Regression Tree (\acs{GBRT}) Algorithm (Step 2)}
+\label{subsec:06:step_2}
+
+For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
+function $F$ relating the target variable $y$ to a set of $K$ predictors
+$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (also called
+\emph{features} or \emph{covariates}) such that the predicted target variable
+is given by $\hat{y} = F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001,
+  Friedman2002}. Its basic elements are decision trees, which use binary splits
+of the input data to create decision rules. Due to their simple nature,
+decision trees are easy to interpret but cannot be used to create satisfying
+predictions for very complex datasets. This issue can be overcome by a
+technique called \emph{boosting} \autocite{Freund1996}. Boosting iteratively
+improves the performance of \emph{weak learners} (in our case decision trees)
+in the following way: At first, a single decision tree is fitted to a randomly
+selected subsample of the training data $\left\{ \left( \bm{x}_i, y_i \right)
+\mid i \in I \right\}$ with an arbitrary index set $I$ (using a predefined loss
+function, \eg{} the mean square error). Then, further decision trees are
+linearly added to the regression function $F$ (scaled by a regularization
+parameter $\nu \ll 1$ called \emph{learning rate}) and the model is fitted
+again in the same manner. This ensures that the prediction for poorly modeled
+training points in the beginning will gradually improve during the training.
+The whole procedure is repeated until the loss function does not improve
+anymore, as evaluated on an independent validation data set. Relevant
+hyperparameters of this algorithm (maximum number of decision trees, maximum
+depth of the decision trees, learning rate and the subsample ratio) are
+optimized by $5$-fold \acf{CV} using the training data. More details on the
+\ac{GBRT} algorithm are given in \cref{subsec:app:b:gbrt}.
+
+A convenient way to evaluate and interpret \ac{GBRT} models (or other
+statistical models in general) is to identify the most relevant predictors.
+Globally, this so-called \emph{feature importance} is closely related to the
+number of appearances of a predictor in the different decision trees that set
+up the \ac{GBRT} model \autocite{Friedman2001}. To evaluate the feature
+importance regionally (for the different grid cells), we use a model
+distillation technique called \emph{\ac{LIME}}, which approximates the
+\ac{GBRT} model with a local linear regression model for each grid cell
+\autocite{Ribeiro2016}. The corresponding linear coefficients $b_k$ for each
+predictor $x^{(k)}$ constitute a good proxy to the model's Jacobian
+\begin{equation}
+  b_k \approx \frac{\partial y}{\partial x^{(k)}},
+  \label{eq:06:jacobian}
+\end{equation}
+which is difficult to compute explicitly for this particular \ac{ML} model.
+Thus, features with high (low) absolute $b_k$ have a high (low) impact on the
+local prediction of the target variable $y$.
 
@@
   \end{tabular}
-  \caption{Process-oriented diagnostics (\emph{predictors} or \emph{features})
-    used in the \acf{GBRT} model to predict the target variables. For step 2a
-    (target variable: absolute \acs{GPP}), all listed variables are monthly
-    climatologies of the specified time ranges in the historical climate. For
-    step 2b (target variable: fractional \acs{GPP} change), the temporal mean
-    over the specified time ranges is calculated for all variables.
-    \AdaptedFrom{Schlund2020}.}
+  \caption{Process-oriented diagnostics (also called \emph{predictors} or
+    \emph{features}) used in the \acf{GBRT} model to predict the target
+    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
+    variables are monthly climatologies of the specified time ranges in the
+    historical climate. For step 2b (target variable: fractional \acs{GPP}
+    change), the temporal mean over the specified time ranges is calculated
+    for all variables. \AdaptedFrom{Schlund2020}.}
   \label{tab:06:predictors}
@@
 
-\subsection{Gradient Boosted Regression Tree (\acs{GBRT}) Algorithm (Step 2)}
-\label{subsec:06:step_2}
+As already described in the previous \namecref{subsec:06:step_1}, we analyze
+two different target variables in this study: the monthly climatology of
+absolute \ac{GPP} at the end of the \nth{21} century (step 2a) and the
+fractional \ac{GPP} change over the \nth{21} century (step 2b). A schematic
+representation of step 2a is shown in \cref{fig:06:mlr_concept}. The target
+variable $y$ for the \ac{ML} model is the rescaled monthly climatology of
+future GPP at the end of the \nth{21} century (\range{2091}{2100}) in the
+\acs{RCP}8.5 scenario given by \cref{eq:06:y}. We relate this target variable
+to several process-oriented diagnostics $\bm{x}$ of the past climate
+(historical simulation) listed in \cref{tab:06:predictors}. A single training
+point $\left( \bm{x}_i, y_i \right)$ for the supervised \ac{ML} algorithm
+corresponds to the set of the aforementioned diagnostics evaluated on a single
+grid cell/month of a single climate model (see colored points in
+\namecrefs{fig:06:mlr_concept} \labelcref{fig:06:mlr_concept}{\color{BrownRed}
+  a} and \labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). Step 2b differs
+from step 2a only in the target variable used. Both steps are independent
+approaches and are evaluated separately.
 
-For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
-function $F$ relating the target variable $y$ to a set of $K$ predictors
-$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (\emph{features} or
-\emph{covariates}) such that the predicted target variable is given by $\hat{y}
-= F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001, Friedman2002}. Its
-basic elements are decision trees, which use binary splits of the input data to
-create decision rules. Due to their simple nature, decision trees are easy to
-interpret but cannot be used to create satisfying predictions for very complex
-datasets. This issue can be overcome by a technique called \emph{boosting}
-\autocite{Freund1996}. Boosting iteratively improves the performance of
-\emph{weak learners} (in our case decision trees) in the following way: At
-first, a single decision tree is fitted to a randomly selected subsample of the
-training data $\left\{ \left( \bm{x}_i, y_i \right) \mid i \in I \right\}$ with
-an arbitrary index set $I$ (using a predefined loss function, \eg{} the mean
-square error). Then, further decision trees are linearly added to the
-regression function $F$ (scaled by a regularization parameter $\nu \ll 1$
-called \emph{learning rate}) and the model is fitted again in the same manner.
-This ensures that the prediction for poorly modeled training points in the
-beginning will gradually improve during the training. The whole procedure is
-repeated until the loss function does not improve anymore, as evaluated on an
-independent validation data set. Relevant hyperparameters of this algorithm
-(maximum number of decision trees, maximum depth of the decision trees,
-learning rate and the subsample ratio) are optimized by $5$-fold \ac{CV} using
-the training data. More details on the \ac{GBRT} algorithm are given in
-\cref{subsec:app:b:gbrt}.
+The main part of steps 2a and 2b each consist of two phases. In the first phase
+(training of the algorithm), we build a statistical model based on the
+empirical relationship between observable process-oriented variables of the
+present-day climate and future \ac{GPP} in the \acs{CMIP}5 ensemble (see gray
+surface in \namecrefs{fig:06:mlr_concept}
+\labelcref{fig:06:mlr_concept}{\color{BrownRed} a} and
+\labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). This has a flavor similar
+to the \ac{MDER} algorithm \commentcite{Karpechko2013, Senftleben2020,
+  Wenzel2016a}{see \cref{subsec:02:mder}}. However, in contrast to the original
+\ac{MDER} algorithm, our multivariate \ac{GBRT} approach is able to handle
+multiple predictors with non-linear dependencies. Moreover, instead of
+considering a single globally or regionally averaged value per climate model,
+our approach uses spatiotemporally gridded data for every \ac{ESM} in the
+ensemble. This dramatically increases the available number of training data
+(even if they are not statistically independent) and uniquely enables the
+statistical model to identify and exploit regional processes. Furthermore, the
+large training database allows us to perform extensive out-of-sample testing,
+which is crucial to ensure the \ac{ML} model's robustness.
+
+In the second phase (predicting with the trained \ac{GBRT} model), we use
+observational data products to constrain the \acs{RCP}8.5 \acs{CMIP}5
+projection of \ac{GPP}. To do this, observational data for all predictors
+$\bm{x}$ of the historical climate (for every grid cell/month) are fed into the
+trained \ac{GBRT} model in order to get a spatially constrained projection of
+\ac{GPP} (see black points in \cref{fig:06:mlr_concept}{\color{BrownRed} b}).
+Similar to the \ac{MDER} algorithm, our approach assumes that there exists an
+inter-model relationship between the predictors and the target variable which
+also holds for the true climate. This assumption may seem weak at first glance,
+but compared to the assumption of traditional weighting approaches (models that
+are better in simulating the present-day climate are also better in simulating
+the future climate and vice versa), the \ac{MDER} assumption is more reasonable
+and allows correcting models instead of just weighing
+\commentcite{Karpechko2013}{see \cref{subsec:02:mder}}. Following other
+\ac{MDER} studies \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}, we
+also select diagnostics as predictor variables $\bm{x}$ that are related to
+\ac{GPP}-relevant processes (see \cref{tab:06:predictors} for detailed physical
+explanations of the links between the predictors and \ac{GPP}).
+
+The raw \acs{CMIP}5 and observation-driven datasets are preprocessed before
+entering the \ac{ML} algorithm (see \cref{subsec:app:b:data_preprocessing}). In
+particular, increases greater than $300 \unit{\%}$ in the fractional \ac{GPP}
+change that is used as target variable in step 2b are removed. These
+unrealistically high values (making up about $5.9 \unit{\%}$ of the data) occur
+in places with very small absolute \ac{GPP} values in the historical simulation
+that are used in the denominator of the derivation of the fractional \ac{GPP}
+change (see \cref{eq:06:f}). In addition, we randomly select $25 \unit{\%}$ of
+the training data and use it as hold-out test data. This dataset is excluded
+from the training and can therefore be used for an independent validation of
+our statistical model.
 
@@
 
-TBA.
+The estimation of the \acf{SPE} for the \ac{GBRT} models needs to consider
+three sources of uncertainty: the uncertainty of the \ac{ML} model itself, the
+uncertainty in the rescaling of the target variable (see \cref{eq:06:y}) and
+the uncertainty in the input data, that is, errors in the observation-based
+products. First, the \ac{SPE} of the \ac{GBRT} model itself (\ie{} uncertainty
+in the position of the gray surface in \namecrefs{fig:06:mlr_concept}
+\labelcref{fig:06:mlr_concept}{\color{BrownRed} a} and
+\labelcref{fig:06:mlr_concept}{\color{BrownRed} b}) is estimated as the
+\emph{\acf{RMSEP}}, that is, the \ac{RMSE} between the predicted and true
+values of the independent hold-out test dataset \autocite{Bishop2006}. This
+approach is justified when the distribution of residuals is unbiased (see
+\cref{subsec:app:b:residuals}), in which case the \ac{RMSEP} equals the
+standard deviation of the residuals (which we interpret as error).
+\Cref{fig:app:b:residuals} shows that this is approximately correct.
+
+The \ac{SPE} of the \ac{GBRT} model can be understood as combined error due to
+the internal variability of the climate system, the different model responses
+within the climate model ensemble to forcing and the incomplete description of
+the climate system with the limited number of variables used as input features.
+Second, the uncertainty in the rescaling is simply assessed as the uncertainty
+given by the emergent constraint from step 1. Finally, the uncertainty in the
+observation-based data (\ie{} uncertainty in the position of the black points
+in \cref{fig:06:mlr_concept}{\color{BrownRed} b}) is assessed by error
+propagation. The necessary sensitivities of the target variable to the various
+predictors for each grid cell/month are given by the linear coefficients $b_k$
+given by the \ac{LIME} approach introduced in \cref{subsec:06:step_2}. All
+sources of uncertainty are assumed independent and are combined by adding the
+squared errors, yielding a single estimate of the \ac{SPE} for each grid
+cell/month. Further details on the calculation of errors are given in
+\cref{subsec:app:b:uncertainty}.
+
+The predictive power of the \ac{GBRT} algorithm and its robustness is assessed
+using a \emph{leave-one-model-out \ac{CV}} approach \commentcite{Elia2002,
+  Karpechko2013}{see \cref{fig:06:mlr_concept}{\color{BrownRed} c} and
+  \cref{subsec:02:model_weighting}}. For this, a single climate model is removed
+from the multi-model ensemble and considered as being the true climate. Our
+statistical model is then fitted on all remaining climate models and a
+prediction for the \enquote{true} model is created. This allows a simple
+evaluation of the predictive power by calculating the \ac{RMSEP}. The whole
+process is repeated for every climate model of the ensemble in order to get a
+distribution of \acp{RMSEP}. Since this approach is not limited to \ac{GBRT},
+it allows a simple comparison of the predictive skill of different statistical
+models. We compare the \ac{GBRT} model to the simple \acs{CMIP}5 \acf{MMM} and
+a \ac{LASSO} model \autocite{Tibshirani1996}. The \ac{LASSO} model is an
+extended linear regression model that regularizes the least squares solution
+with an L$1$ penalty term. In contrast to the ordinary least squares
+regression, the \ac{LASSO} model promotes sparse solutions (\ie{} only a few
+non-zero coefficients/weights), which acts as an intrinsic feature selection
+and provides an effective way to combat overfitting induced by collinearity.
 

### 2021-02-19 ch02_scientific_background/scientific_background.tex
Added sections 6.2.1 and 6.2.2

@@
 terrestrial biosphere is mainly stored as organic compounds, with about
-$\range{450}{650} \unit{GtC}$ in the living vegetation biomass,
-$\range{1500}{2400} \unit{GtC}$ in dead organic matter in litter and soils and
+$\rangeunit{450}{650}{GtC}$ in the living vegetation biomass,
+$\rangeunit{1500}{2400}{GtC}$ in dead organic matter in litter and soils and
 about $1700 \unit{GtC}$ in permafrost soils \autocite{Ciais2013}. The main

### 2021-02-19 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added sections 6.2.1 and 6.2.2

@@
 assessed ranges given by the latest published \ac{AR} of the \ac{IPCC} from
-2013 \autocite{Stocker2013} with $\range{1.8}{5.6} \unit{K}$ for \ac{ECS}
-(\acs{AR}5: $\range{1.5}{4.5} \unit{K}$) and  $\range{1.3}{3.0} \unit{K}$
-(\acs{AR}5: $\range{1.0}{2.5} \unit{K}$) for \ac{TCR}.
+2013 \autocite{Stocker2013} with $\rangeunit{1.8}{5.6}{K}$ for \ac{ECS}
+(\acs{AR}5: $\rangeunit{1.5}{4.5}{K}$) and $\rangeunit{1.3}{3.0}{K}$
+(\acs{AR}5: $\rangeunit{1.0}{2.5}{K}$) for \ac{TCR}.
 
@@
 corresponding modeled ranges from the different \ac{CMIP} generations. Since
-the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
-\unit{K}$ has almost remained unchanged during the last 40 years
-\autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
+the Charney report in 1979, the assessed range of \ac{ECS} of
+$\rangeunit{1.5}{4.5}{K}$ has almost remained unchanged during the last 40
+years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
   Stocker2013} with the exception of \acs{AR}4, where the lower bound was
@@
 bound are more extreme than in any previous climate model generation, resulting
-in a \acs{CMIP}6  \ac{ECS} model range of $\range{1.8}{5.6} \unit{K}$ (as of
+in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$ (as of
 \TheMonth{}). This increase in range is particularly relevant for the upper
@@
   \cref{fig:04:historical_ecs_tcr_ranges}}. Over the years, the modeled range
-of \ac{TCR} has decreased from about $\range{1.3}{3.8} \unit{K}$ in the
-\acs{CMIP}1 models \autocite{Kattenberg1996} to about $\range{1.1}{2.6}
-\unit{K}$ in the \acs{CMIP}5 ensemble \autocite{Flato2013}. Similarly, the
-assessed range of \ac{TCR}, which was first quoted in \acs{AR}4 with
-$\range{1.0}{3.0} \unit{K}$ \autocite{Solomon2007}, has been reduced to
-$\range{1.0}{2.5} \unit{K}$ in the subsequent \acs{AR}5 \autocite{Stocker2013}.
-While the corresponding climate model ranges from \acs{CMIP}3 and \acs{CMIP}5
-agree with these assessed ranges \autocite{Randall2007, Flato2013}, the upper
-bound of the \acs{CMIP}6 model range ($\range{1.3}{3.0} \unit{K}$) is well
-above the assessed upper bound of $2.5 \unit{K}$ given by \acs{AR}5. However,
-unlike \ac{ECS}, also the lower bound of the \acs{CMIP}6 model range increased
-compared to the \acs{CMIP}5 ensemble, leading to a similar \ac{TCR} model
-spread in both climate model generations. Due to the correlation of \ac{TCR}
-and \ac{ECS} given by \cref{eq:02:tcr_vs_ecs}, the increase of \ac{TCR} in many
-\acs{CMIP}6 models is not surprising in light of the many high \ac{ECS} models
-that are present the \acs{CMIP}6 ensemble.
+of \ac{TCR} has decreased from about $\rangeunit{1.3}{3.8}{K}$ in the
+\acs{CMIP}1 models \autocite{Kattenberg1996} to about $\rangeunit{1.1}{2.6}{K}$
+in the \acs{CMIP}5 ensemble \autocite{Flato2013}. Similarly, the assessed range
+of \ac{TCR}, which was first quoted in \acs{AR}4 with $\rangeunit{1.0}{3.0}{K}$
+\autocite{Solomon2007}, has been reduced to $\rangeunit{1.0}{2.5}{K}$ in the
+subsequent \acs{AR}5 \autocite{Stocker2013}. While the corresponding climate
+model ranges from \acs{CMIP}3 and \acs{CMIP}5 agree with these assessed ranges
+\autocite{Randall2007, Flato2013}, the upper bound of the \acs{CMIP}6 model
+range ($\rangeunit{1.3}{3.0}{K}$) is well above the assessed upper bound of
+$2.5 \unit{K}$ given by \acs{AR}5. However, unlike \ac{ECS}, also the lower
+bound of the \acs{CMIP}6 model range increased compared to the \acs{CMIP}5
+ensemble, leading to a similar \ac{TCR} model spread in both climate model
+generations. Due to the correlation of \ac{TCR} and \ac{ECS} given by
+\cref{eq:02:tcr_vs_ecs}, the increase of \ac{TCR} in many \acs{CMIP}6 models is
+not surprising in light of the many high \ac{ECS} models that are present the
+\acs{CMIP}6 ensemble.
 

### 2021-02-19 ch05_paper_ecs/paper_ecs.tex
Added sections 6.2.1 and 6.2.2

@@
 \cref{eq:02:pdf_y_given_x0}). \Cref{tab:05:overview_results} shows the
-corresponding $66 \unit{\%}$ confidence intervals (\ie{} the $\range{17}{83}
-\unit{\%}$ intervals) of \ac{ECS} derived from the \acp{PDF} given by
-\cref{eq:02:pdf_y_given_x0} and the $p$-values used to assess the significance
-of the emergent relationships.
+corresponding $66 \unit{\%}$ confidence intervals (\ie{} the
+$\rangeunit{17}{83}{\%}$ intervals) of \ac{ECS} derived from the \acp{PDF}
+given by \cref{eq:02:pdf_y_given_x0} and the $p$-values used to assess the
+significance of the emergent relationships.
 
@@
   \unit{\degree W}}{120 \unit{\degree E}}$) minus the bias averaged over the
-Northern Hemisphere tropical Pacific ($\range{0 \unit{\degree N}}{20
-  \unit{\degree N}}$, $\range{80 \unit{\degree W}}{120 \unit{\degree E}}$) when
-compared with observations. Similar to the SU constraint, the index proposed by
+Northern Hemisphere tropical Pacific ($\rangeunit{0}{20}{\degree N}$,
+$\range{80 \unit{\degree W}}{120 \unit{\degree E}}$) when compared with
+observations. Similar to the SU constraint, the index proposed by
 \textcite{Tian2015} seems to be related to the emergent constraint by
@@
 climatological annual mean precipitation bias averaged over the southeastern
-Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{150
-  \unit{\degree W}}{100 \unit{\degree W}}$). The southern \ac{ITCZ} index is
-calculated in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ},
-a common problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds
-a link between double-\ac{ITCZ} bias and simulated moisture, precipitation,
-clouds, and large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He
-argues that this could explain the link between the double-\acs{ITCZ} bias and
+Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$,
+$\rangeunit{150}{100}{\degree W}$). The southern \ac{ITCZ} index is calculated
+in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ}, a common
+problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds a link
+between double-\ac{ITCZ} bias and simulated moisture, precipitation, clouds,
+and large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He argues
+that this could explain the link between the double-\acs{ITCZ} bias and
 \ac{ECS}. As reference data, we use observed precipitation data for the years
@@
 tropical total cloud cover ($\range{28 \unit{\degree S}}{28 \unit{\degree N}}$)
-and the southern midlatitudinal total cloud cover ($\range{56 \unit{\degree
-    S}}{36 \unit{\degree S}}$). Analogous to the original study, we use the
-ISCCP-D2 data \autocite{Rossow1991} as observational reference. For the VOL
-constraint, we calculate a constrained $66 \unit{\%}$ range of \ac{ECS} of
+and the southern midlatitudinal total cloud cover ($\rangeunit{56}{36}{\degree
+  S}$). Analogous to the original study, we use the ISCCP-D2 data
+\autocite{Rossow1991} as observational reference. For the VOL constraint, we
+calculate a constrained $66 \unit{\%}$ range of \ac{ECS} of
 $\pmrangeunit{3.74}{0.64}{K}$ for \acs{CMIP}5 ($R^2 = 0.38$) and
@@
 fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
-for both hemispheres ($\range{20 \unit{\degree}}{40 \unit{\degree}}$). On short
-(seasonal) and long (centennial under a forcing) time scales, this quantity is
-well correlated with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5
-models. Together with observations of cloud fraction from Cloudsat/CALIPSO
-\autocite{Mace2009}, \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and
-vertical velocity from ERA-Interim \autocite{Dee2011}, the seasonal response of
-\ac{MBLC} fraction to changes in \ac{SST} forms an emergent constraint on
-\ac{ECS}. We assess a $66 \unit{\%}$ \ac{ECS} range of
-$\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5 ($R^2 = 0.05$) and
-$\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 = 0.62$). In their original
-publication, \textcite{Zhai2015} find an \ac{ECS} range of
-$\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination of
-\acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance, the
-results of the ZHA constraints are somewhat surprising: although \acs{CMIP}5
-data (in combination with \acs{CMIP}3 data) has successfully been used in the
-original publication, our approach finds that the statistical significance of
-the emergent relationship is much higher in the unseen \acs{CMIP}6 ensemble ($p
-< 0.0001$) than in the previously available \acs{CMIP}5 ensemble ($p =
-0.2567$). The ZHA constraint is the only emergent constraint analyzed here that
-shows this extreme behavior (only one other constraint, SHS, shows a slightly
-higher significance in \acs{CMIP}6; all other constraints show lower
-significances in \acs{CMIP}6). The reason for the erratic skill in \acs{CMIP}5
-is the set of climate models used. For our analysis, we use eleven additional
-\acs{CMIP}5 models that have not been used in the original publication
-(ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m, CCSM4, GFDL-ESM2G, GFDL-ESM2M,
-IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and MPI-ESM-P). Due to a lack of
-publicly available data, the model CESM1-CAM5 that is used in the original
-publication is not included in our analysis. The effect of choosing different
-subsets of \acs{CMIP}5 models on the emergent relationship is illustrated in
-\cref{fig:05:zha}. Using the original \acs{CMIP}5 models from the original
-publication gives a considerably higher correlation ($R^2 = 0.38$) than using
-all available \acs{CMIP}5 models ($R^2 = 0.05$). This result shows a strong
-dependency of this emergent constraint on the subset of climate models used.
-Nonetheless, the performance on \acs{CMIP}6 models is, surprisingly, the best
-of all the constraints and much better than on either subset of \acs{CMIP}5
-models.
+for both hemispheres ($\rangeunit{20}{40}{\degree}$). On short (seasonal) and
+long (centennial under a forcing) time scales, this quantity is well correlated
+with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5 models. Together
+with observations of cloud fraction from Cloudsat/CALIPSO \autocite{Mace2009},
+\ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and vertical velocity from
+ERA-Interim \autocite{Dee2011}, the seasonal response of \ac{MBLC} fraction to
+changes in \ac{SST} forms an emergent constraint on \ac{ECS}. We assess a $66
+\unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5
+($R^2 = 0.05$) and $\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 =
+0.62$). In their original publication, \textcite{Zhai2015} find an \ac{ECS}
+range of $\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination
+of \acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance,
+the results of the ZHA constraints are somewhat surprising: although
+\acs{CMIP}5 data (in combination with \acs{CMIP}3 data) has successfully been
+used in the original publication, our approach finds that the statistical
+significance of the emergent relationship is much higher in the unseen
+\acs{CMIP}6 ensemble ($p < 0.0001$) than in the previously available
+\acs{CMIP}5 ensemble ($p = 0.2567$). The ZHA constraint is the only emergent
+constraint analyzed here that shows this extreme behavior (only one other
+constraint, SHS, shows a slightly higher significance in \acs{CMIP}6; all other
+constraints show lower significances in \acs{CMIP}6). The reason for the
+erratic skill in \acs{CMIP}5 is the set of climate models used. For our
+analysis, we use eleven additional \acs{CMIP}5 models that have not been used
+in the original publication (ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m,
+CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and
+MPI-ESM-P). Due to a lack of publicly available data, the model CESM1-CAM5 that
+is used in the original publication is not included in our analysis. The effect
+of choosing different subsets of \acs{CMIP}5 models on the emergent
+relationship is illustrated in \cref{fig:05:zha}. Using the original
+\acs{CMIP}5 models from the original publication gives a considerably higher
+correlation ($R^2 = 0.38$) than using all available \acs{CMIP}5 models ($R^2 =
+0.05$). This result shows a strong dependency of this emergent constraint on
+the subset of climate models used. Nonetheless, the performance on \acs{CMIP}6
+models is, surprisingly, the best of all the constraints and much better than
+on either subset of \acs{CMIP}5 models.
 
@@
     uncertainty ranges correspond to the $66 \unit{\%}$ confidence intervals
-    ($\range{17}{83} \unit{\%}$). For \acs{CMIP}5 and \acs{CMIP}6, these are
+    ($\rangeunit{17}{83}{\%}$). For \acs{CMIP}5 and \acs{CMIP}6, these are
     evaluated from the \acl{PDF} given by \cref{eq:02:pdf_y_given_x0} (see also
@@
 \acs{CMIP}5. Similarly, the size of the $66 \unit{\%}$ \ac{ECS} ranges
-($\range{17}{83} \unit{\%}$ confidence) shows values of $1.16 \unit{K}$ to
-$1.75 \unit{K}$ in \acs{CMIP}5 and $1.32 \unit{K}$ to $2.70 \unit{K}$ in
-\acs{CMIP}6, resulting in an increase of $51 \unit{\%}$ averaged over all
-emergent constraints.
+($\rangeunit{17}{83}{\%}$ confidence) shows values of $1.16 \unit{K}$ to $1.75
+\unit{K}$ in \acs{CMIP}5 and $1.32 \unit{K}$ to $2.70 \unit{K}$ in \acs{CMIP}6,
+resulting in an increase of $51 \unit{\%}$ averaged over all emergent
+constraints.
 
@@
     (orange) ensemble. The bars correspond to the $66 \unit{\%}$ confidence
-    intervals ($\range{17}{83} \unit{\%}$) calculated from the constrained
+    intervals ($\rangeunit{17}{83}{\%}$) calculated from the constrained
     \acl{PDF} of \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical

### 2021-02-19 ch06_paper_gpp/paper_gpp.tex
Added sections 6.2.1 and 6.2.2

@@
 three sources of uncertainty: the uncertainty of the \ac{ML} model itself, the
-uncertainty in the rescaling of the target variable (see \cref{eq:06:y}) and
-the uncertainty in the input data, that is, errors in the observation-based
-products. First, the \ac{SPE} of the \ac{GBRT} model itself (\ie{} uncertainty
-in the position of the gray surface in \namecrefs{fig:06:mlr_concept}
+uncertainty in the rescaling of the target variable and the uncertainty in the
+input data, that is, errors in the observation-based products. First, the
+\ac{SPE} of the \ac{GBRT} model itself (\ie{} uncertainty in the position of
+the gray surface in \namecrefs{fig:06:mlr_concept}
 \labelcref{fig:06:mlr_concept}{\color{BrownRed} a} and
@@
   Karpechko2013}{see \cref{fig:06:mlr_concept}{\color{BrownRed} c} and
-  \cref{subsec:02:model_weighting}}. For this, a single climate model is removed
-from the multi-model ensemble and considered as being the true climate. Our
-statistical model is then fitted on all remaining climate models and a
+  \cref{subsec:02:model_weighting}}. For this, a single climate model is
+removed from the multi-model ensemble and considered as being the true climate.
+Our statistical model is then fitted on all remaining climate models and a
 prediction for the \enquote{true} model is created. This allows a simple
@@
 
-TBA.
+\subsection{Constraining Global Mean \acs{GPP} Projections (Step 1)}
+\label{subsec:06:results_step1}
+
+In \textcite{Wenzel2016}, the observed atmospheric \ac{CO2} concentration at
+\ac{KUM} \autocite{Keeling2005} is used to constrain the \ac{GPP} change in the
+\ac{1BGC} run resulting from a doubling of atmospheric \ac{CO2} concentrations
+to $\pmrangeunit{32}{9}{\%}$ for extratropical ecosystems
+($\rangeunit{30}{90}{\degree N}$). As mentioned in \cref{subsec:06:step_1}, we
+argue that due to the strong annual mixing of \ac{CO2} in the atmosphere, this
+emergent constraint can also be applied to the global mean \ac{GPP} change if
+the emergent relationship holds. In order to test this, we first use the
+idealized \acs{CMIP}5 simulations that were analyzed by \textcite{Wenzel2016}
+to calculate the emergent constraint (\cref{fig:06:step1:a}). The relationship
+is statistically significant at a $5 \unit{\%}$ significance level ($R^2 =
+0.79$, $p = 0.007$) and predicts a constrained global mean \acs{GPP} change of
+$\pmrangeunit{30}{9}{\%}$, which is consistent with the findings of
+\textcite{Wenzel2016}. This result gives us confidence that the emergent
+relationship holds for the global mean \ac{GPP}. Details on the mathematical
+derivation of the constrained range are given in
+\cref{subsec:02:emergent_constraints}.
 
@@
     mean \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
-    the emission-driven fully coupled \acs{RCP}8.5 simulations and the 10-year
+    the emission-driven fully-coupled \acs{RCP}8.5 simulations and the 10-year
     mean \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in
@@
 
-\begin{figure}[t]
+In addition to applying it to the idealized simulations, we apply the emergent
+constraint to the \acs{RCP}8.5 scenario. In the \ac{1BGC} simulation, the
+\ac{CO2} fertilization effect is the only driver of future \ac{GPP} change
+since the carbon cycle components of the \acp{ESM} only see the increasing
+\ac{CO2} concentration. However, since the magnitude of the
+concentration-carbon feedback is believed to be four times the size of the
+climate-carbon feedback \autocite{Gregory2009}, the \ac{CO2} fertilization
+effect is still the dominant driver of \ac{GPP} change in the fully-coupled
+simulations \autocite{Huntzinger2017}. This is also supported by
+\textcite{Wenzel2016}, who show that the \ac{GPP} increase due to rising
+atmospheric \ac{CO2} concentrations in the fully-coupled historical simulations
+is proportional to the same quantity in the \ac{1BGC} run. Additional lines of
+evidence suggest that the modeled \ac{CO2} sensitivity of photosynthesis is the
+main source of uncertainty of future \ac{GPP} \autocite{Arora2013, Haverd2020,
+  Rogers2017}. Thus, the sensitivity of the \ac{CO2} amplitude can be used to
+approximately constrain global mean \ac{GPP} change over the \nth{21} century
+in \acs{RCP}8.5, which is shown in \cref{fig:06:step1:b}. Assuming a
+significance level of $5 \unit{\%}$, the emergent relationship is statistically
+significant ($R^2 = 0.65$, $p = 0.029$).
+
+Using the emergent relationship, the global mean \ac{GPP} change over the
+\nth{21} century can be constrained to $\pmrangeunit{39}{7}{\%}$ (standard
+error), which is at the lower end of the original \acs{CMIP}5 range
+($\rangeunit{31}{57}{\%}$). Moreover, the best estimate $39 \unit{\%}$ is also
+slightly lower than the \acs{CMIP}5 \ac{MMM} of $43 \unit{\%}$. The observed
+\ac{GPP} averaged over the years \range{1991}{2000} from the FLUXNET-MTE
+product is $\pmrangeunit{123}{6}{\GtCyr{}}$ \autocite{Jung2011}. Assuming
+independent errors and using Gaussian error propagation, the constrained
+fractional change corresponds to a global mean \ac{GPP} of
+$\pmrangeunit{171}{12}{\GtCyr{}}$ (standard error) at the end of the \nth{21}
+century (\range{2091}{2100}) in the \acs{RCP}8.5 scenario. In contrast to that,
+the unconstrained\acs{CMIP}5 ensemble has a model range of
+$\rangeunit{156}{247}{\GtCyr{}}$ for future \ac{GPP} in the \acs{RCP}8.5
+scenario (\range{2091}{2100}). The resulting global mean \ac{GPP} change is
+used to rescale the gridded output of the different climate models with
+\cref{eq:06:y}. The following two \namecrefs{subsec:06:results_step2a} show the
+results for the rescaled absolute future \ac{GPP} (step 2a) and the rescaled
+fractional \ac{GPP} change (step 2b) further constrained by the \ac{GBRT}
+model.
+
+\subsection{Constraining Gridded Absolute \acs{GPP} Projections (Step 2a)}
+\label{subsec:06:results_step2a}
+
+In the second step, the objective is to constrain the spatial distribution of
+the projected \ac{GPP}. We use the grid-wise monthly mean climatology of
+absolute \ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) as
+target variable, rescaled using \cref{eq:06:y} to reduce the large
+uncertainties in the \acs{CO2} fertilization effect in the models. In the first
+part, we evaluate the GBRT model using the leave-one-model-out \ac{CV} approach
+and compare it to other statistical approaches and the unweighted \ac{MMM}. In
+the second part, we use observation-based products to predict the target
+variable.
+
+\subsubsection{Prediction Error in Leave‐One‐Model‐Out \acs{CV} Approach and
+  Feature Importance}
+\label{subsubsec:06:results_step2a_ml_inference}
+
+To get a detailed insight into the performance of our \ac{GBRT} model, we
+compare it to five other statistical models: the unweighted \acs{CMIP}5
+\ac{MMM} of future \ac{GPP} (\acs{MMM}), its rescaled version using
+\cref{eq:06:y} (r\acs{MMM}), a linear \ac{LASSO} regression model using all
+features as defined in \cref{tab:06:predictors} (\ac{LASSO}), a
+single-predictor \ac{LASSO} model (\acs{LASSO}-1D) and a single-predictor
+\ac{GBRT} model (\acs{GBRT}-1D). Both single-predictor models use the
+historical \ac{GPP} as only feature. The predictive power (in terms of
+\ac{RMSEP}) of each statistical model is assessed using the leave-one-model-out
+\ac{CV} approach (see \cref{subsec:06:uncertainty_and_skill}). This allows us
+to create an \ac{RMSEP} distribution for each statistical approach, where the
+different points of the distributions refer to different training/prediction
+dataset combinations generated by the leave-one-model-out \ac{CV} approach.
+Thus, each distribution consists of seven points, one for each climate model.
+\Cref{fig:06:step2a_ml_inference:a} shows the \ac{RMSEP} distributions for the
+six different statistical models.
+
+\begin{figure}[!t]
   \centering
@@
 
-\begin{figure}[t]
+In terms of raw predictive power, the simple \ac{MMM} is outperformed by every
+other model. Its prediction uncertainty expressed as mean \ac{RMSEP} can be
+reduced by more than $48 \unit{\%}$ by using other statistical models. However,
+this is not surprising: In contrast to the other statistical models, the simple
+\acs{RCP}8.5 \ac{MMM} does not take further evidence in form of observations of
+the historical climate into account. Step 1 of our algorithm (r\acs{MMM}) can
+reduce this mean \ac{RMSEP} by $15 \unit{\%}$ due to the simple rescaling of
+the gridded climate model output with the global mean \ac{GPP} constraint on
+the \ac{CO2} fertilization effect. However, since there is a considerably large
+\ac{GPP} range in the individual climate models themselves, this reduction is
+rather small for the gridded values. A far larger reduction of the \ac{RMSEP}
+can be achieved by using the regression models \acs{LASSO}‐1D, \ac{LASSO},
+\acs{GBRT}‐1D and \ac{GBRT}. All of them share similar \ac{RMSEP}
+distributions, which can be explained as follows: The historical \ac{GPP} is
+strongly (near linearly) correlated to the rescaled end-of-century \ac{GPP}
+(with pattern correlation of $R^2 = 0.83$ for the whole multi-model ensemble).
+Because of that, all regression models are able to considerably reduce the
+\acp{RMSEP} compared to the \acs{MMM} and r\acs{MMM}. This also shows that the
+\ac{GBRT} models successfully learn the linear connection between past and
+future \ac{GPP}. Moreover, the non-linear \ac{GBRT} models can slightly reduce
+the mean \ac{RMSEP} even further compared to the linear \ac{LASSO} models by
+about $2 \unit{\%}$ for \acs{GBRT}‐1D and $3 \unit{\%}$ for the full \ac{GBRT}
+model. This leads to the conclusion that in addition to the strong linear
+relationship between past and future \ac{GPP}, there are small non-linear
+relations between the predictors and the target variable that can be used to
+further reduce the \ac{RMSEP}. Since the full \ac{GBRT} model with access to
+all predictors shows the minimal mean \ac{RMSEP}, we argue that it is
+beneficial for our approach to use the multivariate non-linear \ac{GBRT} model
+instead of a linear model because only the \ac{GBRT} algorithm is able to make
+use of all predictors and exploit more complex relationships.
+
+The influence of the different predictors can be further analyzed by evaluating
+the global feature importance by using the whole training dataset and the
+already trained model. This is shown in \cref{fig:06:step2a_ml_inference:b}.
+The past \ac{GPP} with its strong positive linear correlation to future
+\ac{GPP} is the most relevant predictor of the trained model with a relative
+importance of approximately $95 \unit{\%}$. All other features show values of
+less than $2 \unit{\%}$. The global feature importance determines the expected
+impact of a feature in predictions of the target variable. In our \ac{GBRT}
+model, the prediction input data of historical \ac{GPP} mainly determines the
+\ac{ML} model's prediction of the future \ac{GPP} in the \acs{RCP}8.5 scenario
+at the end of the \nth{21} century. However, we emphasize that this is only
+valid for our specific setup (\ie{} which variables are used, which algorithm
+is used, which climate models are considered, \etc{}). For another problem
+setup, the global feature importance could change significantly. Since the
+historical \ac{GPP} is the most important predictor for all grid cells, we do
+not show the plot of the local feature importance using \ac{LIME}. We can get
+further insights into our \ac{GBRT} model by analyzing the residuals of the
+prediction using an independent test dataset. The resulting plots show that the
+algorithm is not overfitting the training data and that the prediction errors
+on the whole are approximately unbiased (see \cref{fig:app:b:residuals:a}).
+
+\subsubsection{Observation-Based \acs{GBRT} Prediction of Absolute \acs{GPP}}
+\label{subsubsec:06:results_step2a_obs}
+
+In this \namecref{subsubsec:06:results_step2a_obs}, we use the different
+statistical models to predict absolute \ac{GPP} at the end of the \nth{21}
+century. For this, we feed observation-based data (see
+\cref{tab:06:predictors}) into the regression models \ac{GBRT}, \acs{GBRT}‐1D,
+\ac{LASSO} and \acs{LASSO}‐1D.
+
+The result of step 1 on the gridded data is illustrated in
+\cref{fig:06:step2a_results:a}, which shows the ratio between r\acs{MMM} and
+\acs{MMM}. This ratio is almost constant over the whole globe, which is not
+surprising due to the global nature of the rescaling (see \cref{eq:06:y}). The
+use of the global emergent constraint predicts a slightly lower \ac{GPP}
+increase over the \nth{21} century ($39 \unit{\%}$) than the unweighted
+\acs{CMIP}5 ensemble mean ($43 \unit{\%}$). The corresponding ratio $39
+\unit{\%} / 43 \unit{\%} \approx 0.91$ is approximately equal to the mean value
+of r\acs{MMM}/\acs{MMM} ($0.92$). Consequently, the global estimate of
+r\acs{MMM} gives a lower \ac{GPP} at the end of the century ($179
+\unit{\GtCyr{}}$) as \acs{MMM} ($198 \unit{\GtCyr{}}$), which can be
+interpreted as a correction of the \acp{ESM}' overestimation of future \ac{GPP}
+to changes in the atmospheric \ac{CO2} concentration.
+
+\begin{figure}[!t]
   \centering
@@
     avoid divisions by $0$. Both panels show that our approach is only to
-    first-order a bias correction, in which case both plots would only show
+    first order a bias correction, in which case both plots would only show
     constant values. (c) Comparison of the \acf{GBRT} versus \acs{GBRT}-1D
@@
 \end{figure}
+
+The spatial result of our \ac{GBRT} model in step 2a can be visualized by
+comparing r\acs{MMM} and the output of the \acs{GBRT} model.
+\Cref{fig:06:step2a_results:b} shows the bias of the rescaled \acs{CMIP}5
+\acs{RCP}8.5 \ac{MMM} (r\acs{MMM}) compared to our \acs{GBRT} predicted
+\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}), while
+\Cref{fig:06:step2a_results:c} shows the bias of the historical \acs{CMIP}5
+\ac{MMM} compared to the observational FLUXNET-MTE product \autocite{Jung2011}
+averaged for the period \range{1991}{2000}. As reported by \textcite{Anav2013},
+the historical \acs{CMIP}5 ensemble mean overestimates \ac{GPP} in most
+regions, leading to a global mean of $138 \unit{\GtCyr{}}$, which is larger
+than the $123 \unit{\GtCyr{}}$ estimate for the FLUXNET-MTE product. Regions
+where \ac{GPP} is largely overestimated are the western parts of South America,
+central and southern Africa, and East Asia. On the contrary, \ac{GPP} is
+underestimated in small areas of Central America and northern parts of South
+America. The bias patterns in
+\cref{fig:06:step2a_results:b,fig:06:step2a_results:c} are very similar
+(pattern correlation of $R^2 = 0.88$), which means that the \ac{GBRT} algorithm
+detects regional biases in the historical simulation and corrects them in its
+future predictions. This is also illustrated in
+\cref{fig:06:step2a_results_biases:b}, in which many values are close to $1$,
+while only a few negative values exist. However, we emphasize that our approach
+is only to first order a bias correction (\ie{} subtracting the historical bias
+from the \acs{RCP}8.5 \ac{MMM}). There are clear differences between
+\cref{fig:06:step2a_results:b,fig:06:step2a_results:c}, which are also
+illustrated in
+\cref{fig:06:step2a_results_biases:a,fig:06:step2a_results_biases:b}. A simple
+bias correction would only show constant values for the whole globe, whereas
+our approach applies different corrections (in sign and magnitude) for
+different regions. On the global domain, our approach predicts a \ac{GPP} at
+the end of the century of $169 \unit{\GtCyr{}}$, which is consistent with the
+global constraint from step 1 of $\pmrangeunit{171}{12}{\GtCyr{}}$. In summary,
+our approach first corrects the \acs{CMIP}5 models response to \ac{CO2} (step
+1) and second corrects the historical bias of the \acp{ESM} relative to
+observations (step 2a).
 
@@
 \end{figure}
+
+Further illustrations of these results including the uncertainties can be found
+in \cref{fig:app:b:cmip5_hist_mte,fig:app:b:step2a_results,%
+  fig:app:b:step2a_results_errors}. Analogous to the results from the
+leave-one-model-out \ac{CV} approach, the uncertainties on grid cell level
+are significantly reduced for the \ac{GBRT} prediction compared to the
+\ac{MMM} results. These errors consider all three sources of uncertainty that
+we present in \cref{subsec:06:uncertainty_and_skill} (error in the \ac{GBRT}
+model, error in the rescaling and error in the observational-driven
+products). In contrast to that, the error derived from the
+leave-one-model-out \ac{CV} approach illustrated in
+\cref{fig:06:step2a_ml_inference:a} only shows the error in the statistical
+models themselves. We note that local relative errors can be very large
+(especially in regions with low absolute \ac{GPP}) due to the uncertainty
+calculation method: Regardless of the absolute value, the error of each grid
+cell is at least $535 \unit{gC \, m^{-2} yr^{-1}}$ (estimated \ac{SPE} of the
+\ac{GBRT} model using the \ac{RMSEP}), which is added to the propagated
+errors of the observation-driven datasets and the error derived from the
+rescaling using \cref{eq:app:b:sigma_rescaling}. The \ac{GBRT} model error is
+the dominant source of uncertainty. Since the true covariance structure of
+these global fields is unknown, a global (or at least regional) aggregation
+of these uncertainties is not possible.
+
+To further compare our \ac{GBRT} approach to other statistical models, we also
+fed the observation-driven data into the \acs{GBRT}‐1D, \ac{LASSO}, and
+\acs{LASSO}‐1D models to get a constrained projection of the \acs{RCP}8.5
+\ac{GPP}. \Cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}
+show the difference of the other models to the full \ac{GBRT} approach. Both
+panels show values close to $0$ for most of the globe (\ie{} the differences
+between the \ac{GBRT} and \acs{GBRT}‐1D/\ac{LASSO} models are small),
+indicating that the bias-correcting property of our approach is also present in
+the other regression models that use only the historical \ac{GPP} as single
+predictors. Thus, we conclude that the bias correction of our \ac{GBRT} model
+originates from the observation of the historical \ac{GPP}, which is also
+supported by \cref{fig:06:step2a_ml_inference:b}, which shows historical
+\ac{GPP} as by far most important feature. On the contrary, second-order
+corrections originate from the observations in the remaining predictors. These
+second-order corrections are also visible as the non-zero values in
+\Cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}. This
+indicates that using the non-linear \ac{GBRT} model with all predictors
+improves the final result. Moreover, the similar patterns in both panels
+demonstrate that the \acs{GBRT}‐1D model emulates the strong linear relation of
+the historical \ac{GPP} to future \ac{GPP} and performs equally well as the
+linear models. The global estimates of $167 \unit{\GtCyr{}}$ (\acs{GBRT}‐1D),
+$163 \unit{\GtCyr{}}$ (\ac{LASSO}), and $162 \unit{\GtCyr{}}$ (\acs{LASSO}‐1D)
+are all consistent with the global constraint of
+$\pmrangeunit{171}{12}{\GtCyr{}}$.
 

### 2021-02-22 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Finished section 6.2

@@
 
-However, it is well-known that the low-level mixed-phase clouds over the
+However, it is well known that the low-level mixed-phase clouds over the
 Southern Ocean in these earlier climate models are biased towards too large

### 2021-02-22 ch06_paper_gpp/paper_gpp.tex
Finished section 6.2

@@
 compare it to five other statistical models: the unweighted \acs{CMIP}5
-\ac{MMM} of future \ac{GPP} (\acs{MMM}), its rescaled version using
-\cref{eq:06:y} (r\acs{MMM}), a linear \ac{LASSO} regression model using all
-features as defined in \cref{tab:06:predictors} (\ac{LASSO}), a
-single-predictor \ac{LASSO} model (\acs{LASSO}-1D) and a single-predictor
-\ac{GBRT} model (\acs{GBRT}-1D). Both single-predictor models use the
-historical \ac{GPP} as only feature. The predictive power (in terms of
-\ac{RMSEP}) of each statistical model is assessed using the leave-one-model-out
-\ac{CV} approach (see \cref{subsec:06:uncertainty_and_skill}). This allows us
-to create an \ac{RMSEP} distribution for each statistical approach, where the
-different points of the distributions refer to different training/prediction
-dataset combinations generated by the leave-one-model-out \ac{CV} approach.
-Thus, each distribution consists of seven points, one for each climate model.
+\ac{MMM} of future \ac{GPP}, its rescaled version using \cref{eq:06:y}
+(r\acs{MMM}), a linear \ac{LASSO} regression model using all features as
+defined in \cref{tab:06:predictors}, a single-predictor \ac{LASSO} model
+(\acs{LASSO}-1D) and a single-predictor \ac{GBRT} model (\acs{GBRT}-1D). Both
+single-predictor models use the historical \ac{GPP} as only feature. The
+predictive power (in terms of \ac{RMSEP}) of each statistical model is assessed
+using the leave-one-model-out \ac{CV} approach (see
+\cref{subsec:06:uncertainty_and_skill}). This allows us to create an \ac{RMSEP}
+distribution for each statistical approach, where the different points of the
+distributions refer to different training/prediction dataset combinations
+generated by the leave-one-model-out \ac{CV} approach. Thus, each distribution
+consists of seven points, one for each climate model.
 \Cref{fig:06:step2a_ml_inference:a} shows the \ac{RMSEP} distributions for the
@@
 Because of that, all regression models are able to considerably reduce the
-\acp{RMSEP} compared to the \acs{MMM} and r\acs{MMM}. This also shows that the
+\acp{RMSEP} compared to the \ac{MMM} and r\acs{MMM}. This also shows that the
 \ac{GBRT} models successfully learn the linear connection between past and
@@
 \cref{fig:06:step2a_results:a}, which shows the ratio between r\acs{MMM} and
-\acs{MMM}. This ratio is almost constant over the whole globe, which is not
+\ac{MMM}. This ratio is almost constant over the whole globe, which is not
 surprising due to the global nature of the rescaling (see \cref{eq:06:y}). The
@@
 r\acs{MMM} gives a lower \ac{GPP} at the end of the century ($179
-\unit{\GtCyr{}}$) as \acs{MMM} ($198 \unit{\GtCyr{}}$), which can be
+\unit{\GtCyr{}}$) as \ac{MMM} ($198 \unit{\GtCyr{}}$), which can be
 interpreted as a correction of the \acp{ESM}' overestimation of future \ac{GPP}
@@
 
+Further illustrations of these results including the uncertainties can be found
+in \cref{fig:app:b:cmip5_hist_mte,fig:app:b:step2a_results,%
+  fig:app:b:step2a_results_errors}. Analogous to the results from the
+leave-one-model-out \ac{CV} approach, the uncertainties on grid cell level
+are significantly reduced for the \ac{GBRT} prediction compared to the
+\ac{MMM} results. These errors consider all three sources of uncertainty that
+we present in \cref{subsec:06:uncertainty_and_skill} (error in the \ac{GBRT}
+model, error in the rescaling and error in the observational-driven
+products). In contrast to that, the error derived from the
+leave-one-model-out \ac{CV} approach illustrated in
+\cref{fig:06:step2a_ml_inference:a} only shows the error in the statistical
+models themselves. We note that local relative errors can be very large
+(especially in regions with low absolute \ac{GPP}) due to the uncertainty
+calculation method: Regardless of the absolute value, the error of each grid
+cell is at least $535 \unit{gC \, m^{-2} yr^{-1}}$ (estimated \ac{SPE} of the
+\ac{GBRT} model using the \ac{RMSEP}), which is added to the propagated
+errors of the observation-driven datasets and the error derived from the
+rescaling using \cref{eq:app:b:sigma_rescaling}. The \ac{GBRT} model error is
+the dominant source of uncertainty. Since the true covariance structure of
+these global fields is unknown, a global (or at least regional) aggregation
+of these uncertainties is not possible.
+
+To further compare our \ac{GBRT} approach to other statistical models, we also
+fed the observation-driven data into the \acs{GBRT}‐1D, \ac{LASSO}, and
+\acs{LASSO}‐1D models to get a constrained projection of the \acs{RCP}8.5
+\ac{GPP}. \Cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}
+show the difference of the other models to the full \ac{GBRT} approach. Both
+panels show values close to $0$ for most of the globe (\ie{} the differences
+between the \ac{GBRT} and \acs{GBRT}‐1D/\ac{LASSO} models are small),
+indicating that the bias-correcting property of our approach is also present in
+the other regression models that use only the historical \ac{GPP} as single
+predictors. Thus, we conclude that the bias correction of our \ac{GBRT} model
+originates from the observation of the historical \ac{GPP}, which is also
+supported by \cref{fig:06:step2a_ml_inference:b}, which shows historical
+\ac{GPP} as by far most important feature. On the contrary, second-order
+corrections originate from the observations in the remaining predictors. These
+second-order corrections are also visible as the non-zero values in
+\Cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}. This
+indicates that using the non-linear \ac{GBRT} model with all predictors
+improves the final result. Moreover, the similar patterns in both panels
+demonstrate that the \acs{GBRT}‐1D model emulates the strong linear relation of
+the historical \ac{GPP} to future \ac{GPP} and performs equally well as the
+linear models. The global estimates of $167 \unit{\GtCyr{}}$ (\acs{GBRT}‐1D),
+$163 \unit{\GtCyr{}}$ (\ac{LASSO}), and $162 \unit{\GtCyr{}}$ (\acs{LASSO}‐1D)
+are all consistent with the global constraint of
+$\pmrangeunit{171}{12}{\GtCyr{}}$.
+
+\subsection{Constraining Gridded Fractional \acs{GPP} Change Projections (Step
+  2b)}
+\label{subsec:06:results_step2b}
+
+In step 2b, we constrain the gridded fractional \ac{GPP} change over the
+\nth{21} century (2100 versus 2000) in the emission-driven \acs{RCP}8.5
+scenario (rescaled using \cref{eq:06:y}). This step is independent from step 2a
+and only differs in the target variable (\ac{GPP} change instead of absolute
+\ac{GPP}). The remaining setup including the predictors, the \ac{GBRT} model
+itself and the datasets are similar.
+
+\subsubsection{Prediction Error in Leave‐One‐Model‐Out \acs{CV} Approach and
+  Feature Importance}
+\label{subsubsec:06:results_step2b_ml_inference}
+
+\Cref{fig:06:step2b_ml_inference:a} shows the \ac{RMSEP} distributions for four
+different statistical models: the simple \acs{CMIP}5 \ac{MMM} of the fractional
+\ac{GPP} change, its rescaled version using \cref{eq:06:y} (r\acs{MMM}), a
+linear \ac{LASSO} regression model and the \ac{GBRT} model. Similar to step 2a,
+the \ac{GBRT} approach shows the smallest mean \ac{RMSEP} values of all
+statistical models. Compared to \ac{MMM} and r\acs{MMM} the mean is reduced by
+$16 \unit{\%}$ and $9 \unit{\%}$, respectively. Moreover, the non-linear
+\ac{GBRT} model also outperforms the linear \ac{LASSO} model (mean reduced by
+$3 \unit{\%}$). In contrast to step 2a, there is not a single predictor which
+heavily dominates the feature importance of the \ac{GBRT} model. Thus, using
+single-predictor models (\acs{GBRT}‐1D and \acs{LASSO}‐1D) is not necessary in
+step 2b. \Cref{fig:06:step2b_ml_inference:b,fig:06:step2b_ml_inference:d} show
+the global feature importance for the \ac{LASSO} and \ac{GBRT} model,
+respectively. Both models agree that the \acf{T} and \ac{LAI} are the two
+dominant predictors with a relative importance of more than $65 \unit{\%}$. The
+\ac{LASSO} model shows a relative importance of less than $10 \unit{\%}$ for
+each of the remaining features, while the \ac{GBRT} model exhibits a
+considerably high relative importance for \ac{GPP} with almost $20 \unit{\%}$.
+All predictors are negatively correlated to the target variable (as given by
+the sign of the linear Pearson correlation coefficient). That means cold areas
+with a small \ac{LAI} in the historical climate show a larger \ac{GPP} change
+over the \nth{21} century and vice versa. Indeed, the fertilization effect is
+expected to be stronger in lower \ac{LAI} regions. To get more detailed
+insights on this, observation-driven data is fed into the \ac{GBRT} model to
+obtain a constrained fractional \ac{GPP} change over the \nth{21} century.
+
 \begin{figure}[p]
@@
 
-Further illustrations of these results including the uncertainties can be found
-in \cref{fig:app:b:cmip5_hist_mte,fig:app:b:step2a_results,%
-  fig:app:b:step2a_results_errors}. Analogous to the results from the
-leave-one-model-out \ac{CV} approach, the uncertainties on grid cell level
-are significantly reduced for the \ac{GBRT} prediction compared to the
-\ac{MMM} results. These errors consider all three sources of uncertainty that
-we present in \cref{subsec:06:uncertainty_and_skill} (error in the \ac{GBRT}
-model, error in the rescaling and error in the observational-driven
-products). In contrast to that, the error derived from the
-leave-one-model-out \ac{CV} approach illustrated in
-\cref{fig:06:step2a_ml_inference:a} only shows the error in the statistical
-models themselves. We note that local relative errors can be very large
-(especially in regions with low absolute \ac{GPP}) due to the uncertainty
-calculation method: Regardless of the absolute value, the error of each grid
-cell is at least $535 \unit{gC \, m^{-2} yr^{-1}}$ (estimated \ac{SPE} of the
-\ac{GBRT} model using the \ac{RMSEP}), which is added to the propagated
-errors of the observation-driven datasets and the error derived from the
-rescaling using \cref{eq:app:b:sigma_rescaling}. The \ac{GBRT} model error is
-the dominant source of uncertainty. Since the true covariance structure of
-these global fields is unknown, a global (or at least regional) aggregation
-of these uncertainties is not possible.
-
-To further compare our \ac{GBRT} approach to other statistical models, we also
-fed the observation-driven data into the \acs{GBRT}‐1D, \ac{LASSO}, and
-\acs{LASSO}‐1D models to get a constrained projection of the \acs{RCP}8.5
-\ac{GPP}. \Cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}
-show the difference of the other models to the full \ac{GBRT} approach. Both
-panels show values close to $0$ for most of the globe (\ie{} the differences
-between the \ac{GBRT} and \acs{GBRT}‐1D/\ac{LASSO} models are small),
-indicating that the bias-correcting property of our approach is also present in
-the other regression models that use only the historical \ac{GPP} as single
-predictors. Thus, we conclude that the bias correction of our \ac{GBRT} model
-originates from the observation of the historical \ac{GPP}, which is also
-supported by \cref{fig:06:step2a_ml_inference:b}, which shows historical
-\ac{GPP} as by far most important feature. On the contrary, second-order
-corrections originate from the observations in the remaining predictors. These
-second-order corrections are also visible as the non-zero values in
-\Cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}. This
-indicates that using the non-linear \ac{GBRT} model with all predictors
-improves the final result. Moreover, the similar patterns in both panels
-demonstrate that the \acs{GBRT}‐1D model emulates the strong linear relation of
-the historical \ac{GPP} to future \ac{GPP} and performs equally well as the
-linear models. The global estimates of $167 \unit{\GtCyr{}}$ (\acs{GBRT}‐1D),
-$163 \unit{\GtCyr{}}$ (\ac{LASSO}), and $162 \unit{\GtCyr{}}$ (\acs{LASSO}‐1D)
-are all consistent with the global constraint of
-$\pmrangeunit{171}{12}{\GtCyr{}}$.
+\subsubsection{Observation-Based \acs{GBRT} Prediction of Relative \acs{GPP}}
+\label{subsubsec:06:results_step2b_obs}
+
+The global distribution of the fractional \ac{GPP} change for the different
+statistical models is shown in \cref{fig:06:step2b_results}. All panels show a
+\ac{GPP} increase over the \nth{21} century in the \acs{RCP}8.5 scenario for
+almost all regions of the globe. Compared to the unweighted \acs{CMIP}5
+\ac{MMM}, its rescaled version (r\acs{MMM}) shows a slightly lower \ac{GPP}
+increase, which is consistent with the emergent constraint approach from step 1
+that shows a lower global mean \ac{GPP} increase over the \nth{21} century ($39
+\unit{\%}$) than the global \acs{CMIP}5 ensemble mean ($43 \unit{\%}$). This is
+also illustrated in \cref{fig:06:step2b_results_biases:a}, which shows an
+almost constant pattern over the whole globe with a mean of $0.91$ that is
+consistent with the ratio $39 \unit{\%} / 43 \unit{\%} \approx 0.91$. Similar
+to step 2a, this can be interpreted as a global correction of the \acp{ESM}'
+response of future \ac{GPP} to changes in the atmospheric \ac{CO2}
+concentration.
 
@@
 
+All four patterns of \cref{fig:06:step2b_results} show very similar
+geographical patterns. Obvious exceptions are the Sahara desert and the Arabian
+Peninsula, which show a noisy behavior for the \aclp{MMM} (\ac{MMM} and
+r\acs{MMM}) and a small \ac{GPP} increase for the two regression models
+(\ac{LASSO} and \ac{GBRT}). This noise-like pattern occurs due to numerical
+inconsistencies produced by very small values in the absolute historical and
+future \ac{GPP} in the climate models in this region that are used to derive
+the fractional \ac{GPP} change following \cref{eq:06:f}. However, due to the
+small impact on the global mean \ac{GPP}, this region is negligible in our
+analysis. The fractional \ac{GPP} change is not uniformly distributed over the
+globe. Instead, there is a pronounced latitudinal dependency in its
+geographical pattern: In high latitudes, the projected \ac{GPP} change is
+larger than in regions closer to the equator. This effect is particularly
+strong for the northern high latitudes and consistent with the results of
+\textcite{Wenzel2016}, who find an increased \ac{GPP} change in northern
+high-latitude ecosystems compared to the northern extratropics. A possible
+reason for this is the extension of the growing season in high latitudes caused
+by climate change connected to a \emph{greening}, which has already been
+observed in the past climate through satellite measurements
+\autocite{Lucht2002, Myneni1997, Zhang2020} and associated with the increase in
+the \ac{CO2} seasonal cycle amplitude \autocite{Forkel2016}. This is consistent
+with the feature map in \cref{fig:06:step2b_ml_inference:c}, which shows an
+increased relative importance of the \ac{LAI} (a quantity directly related to
+the greening) in the high-latitude regions with high \ac{GPP} changes. On the
+contrary, in the tropical ecosystems the growing season already covers the
+whole year. Thus, a further climate change--induced extension is not possible,
+leading to a smaller fractional \ac{GPP} change in regions closer to the
+equator. For most parts of the remaining globe, \ac{T} is the dominant
+predictor, with some areas in northern Africa, the Middle East, India and
+Australia showing an increased relative importance of the historical \ac{GPP}.
+Overall, the negative correlation of the three most important features (\ac{T},
+\ac{LAI}, and historical \ac{GPP}) to the target variable is well reflected in
+the global pattern of the fractional \ac{GPP} change: The high latitudes with
+lower \ac{T}, lower \ac{LAI} and lower historical \ac{GPP} show higher \ac{GPP}
+changes, whereas regions closer to the equator with higher \ac{T}, higher
+\ac{LAI} and higher historical \ac{GPP} show smaller \ac{GPP} changes. At this
+point we want to emphasize that it is not possible to infer global or other
+large-scale averages of the fractional \ac{GPP} change from the geographical
+distributions in \cref{fig:06:step2b_results} since division operations and
+averaging operations do not commute: in general, $\sum_i \left( a_i / b_i
+\right) \neq \left( \sum_i a_i \right) / \left( \sum_i b_i \right)$. Thus, a
+direct comparison to other studies that give locally or globally aggregated
+results is not possible.
+
+To get further insights in the \ac{GBRT} algorithm, we compare it directly to
+the rescaled \acs{CMIP}5 \ac{MMM} (r\acs{MMM}), which is illustrated in
+\cref{fig:06:step2b_results_biases:b,fig:06:step2b_results_biases:c}. The
+mismatch between \ac{GBRT} and r\acs{MMM} can be interpreted as additional
+information that is added to the target variable by the observations in the
+predictors in step 2b. Overall, the absolute difference
+(\cref{fig:06:step2b_results_biases:b}) is small over the whole globe,
+indicating that the r\acs{MMM} and \ac{GBRT} agree in the general pattern of
+the fractional \ac{GPP} change. The most striking deviation in this panel is
+again the Sahara and Arabian Peninsula region, which shows the already
+discussed large absolute differences. Apart from that, there is a large patch
+in central Asia that shows high negative differences, that is, the \ac{GBRT}
+model predicts a smaller \ac{GPP} change in this area. The relative difference
+(\cref{fig:06:step2b_results_biases:c}) indicates many regions where the
+\ac{GBRT} model predicts an increased \ac{GPP} change compared to r\acs{MMM},
+in particular South America, South Africa, the west coast of Africa, the Middle
+East, parts of Australia and western parts of the United States. However, all
+these correspond to low absolute \ac{GPP} changes.
+\Cref{fig:app:b:step2b_results_errors} shows the corresponding gridded
+uncertainties for all four statistical models that also cover the uncertainty
+in the emergent constraint and the observational uncertainty (in contrast to
+\cref{fig:06:step2b_ml_inference:a}, which only covers the uncertainty in the
+statistical models). Similar to step 2a, the local standard errors are large,
+even for the \ac{GBRT} model (that performs better than any other statistical
+model) which shows errors of at least $43.6 \unit{\%}$ for every grid cell
+caused by the uncertainty in the statistical model itself.
+\Cref{fig:app:b:residuals:b} illustrates the residuals for the training and
+independent test datasets and shows that the \ac{GBRT} model has symmetrical
+residuals and is not overfitting.
+
+\subsection{Comparison of the Absolute and Relative \acs{GBRT} Prediction}
+\label{subsec:06:comparison_step2a_step2b}
+
+In a last step, we use the FLUXNET‐MTE observational product to constrain the
+absolute \ac{GPP} increase at the end of the \nth{21} century
+(\range{2091}{2100}) from the fractional \ac{GPP} change given by the \ac{GBRT}
+model in step 2b. This can be directly compared to the output of step 2a and
+serves as a sanity check between steps 2a and 2b. As shown by
+\cref{fig:06:comparison_step2a_step2b}, both approaches show similar
+geographical patterns (pattern correlation: $R^2 = 0.97$). The difference
+between the two approaches is in the same order of magnitude as the difference
+between the different statistical models used in step 2a (see
+\cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}) and
+considerably smaller than the difference between the \ac{GBRT} and r\acs{MMM}
+approaches (see \cref{fig:06:step2a_results:b}). Moreover, the globally
+aggregated results ($169 \unit{\GtCyr{}}$ for step 2a and $175 \unit{\GtCyr{}}$
+for step 2b) are both compatible with the global value of
+$\pmrangeunit{171}{12}{\GtCyr{}}$ given by the global emergent constraint in
+step 1.
+
 \begin{figure}[t]

### 2021-02-22 ch02_scientific_background/scientific_background.tex
Finished chapter 06

@@
 relationship (red line in \cref{fig:02:emergent_constraint}) is a robust and
-physically interpretable inter-model relationship between a target variable $y$
+physically-interpretable inter-model relationship between a target variable $y$
 related to the future climate and an observable $x$ of the past or present-day
@@
   \caption{Schematic illustration of the emergent constraint approach. Basis of
-    every emergent constraint is a robust and physically interpretable emergent
+    every emergent constraint is a robust and physically-interpretable emergent
     relationship (red line) between a target variable $y$ (\eg{} Earth system

### 2021-02-22 ch05_paper_ecs/paper_ecs.tex
Finished chapter 06

@@
 data (like the new \acs{CMIP}6 ensemble), since the emergent relationship
-appeared just by chance and not because of a physically based mechanism.
+appeared just by chance and not because of a physically-based mechanism.
 

### 2021-02-22 ch06_paper_gpp/paper_gpp.tex
Finished chapter 06

@@
 \Cref{fig:06:step2a_results:b} shows the bias of the rescaled \acs{CMIP}5
-\acs{RCP}8.5 \ac{MMM} (r\acs{MMM}) compared to our \acs{GBRT} predicted
+\acs{RCP}8.5 \ac{MMM} (r\acs{MMM}) compared to our \acs{GBRT}-predicted
 \ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}), while
@@
 
-TBA.
+In this paper, we develop a two-step approach to constrain the projected
+\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
+\acs{RCP}8.5 scenario. In the first step, we constrain the global mean \ac{GPP}
+to $\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint
+approach which we assume can also be applied to the global mean \ac{GPP} at the
+end of the \nth{21} century in the \acs{RCP}8.5 scenario (see discussion in
+\cref{subsec:06:step_1}). This step corresponds to the correction of the
+\acp{ESM}' biases in the response of future \ac{GPP} to rising atmospheric
+\ac{CO2} concentration, which is the main source of uncertainty in future
+\ac{GPP} projections \autocite{Arora2013, Haverd2020, Rogers2017}. As already
+discussed in \cref{subsec:06:step_1}, due to other than \ac{CO2} fertilization
+drivers of the increase in the \ac{CO2} seasonal cycle amplitude
+\autocite{Bastos2019, Forkel2016, Piao2018, Zhao2016} this emergent constraint
+is not undisputed and can be replaced with another emergent constraint in
+future studies. In the second step, a \ac{ML} approach is used to further
+constrain the gridded \ac{GPP} based on process-based present-day predictors.
+We consider two target variables: first (step 2a) the gridded monthly
+climatologies of absolute \ac{GPP} (\range{2091}{2100}) and second (step 2b)
+the gridded fractional \ac{GPP} change over the \nth{21} century (years
+\range{2091}{2100} versus years \range{1991}{2000}; see \cref{eq:06:f}). Both
+approaches give consistent results (see
+\cref{fig:06:comparison_step2a_step2b}). The latter quantity shows an increased
+\ac{GPP} change in the high latitudes compared to regions closer to the
+equator, which might be attributed to an additional greening trend in the high
+latitudes as supported by the feature map in
+\cref{fig:06:step2b_ml_inference:c} that shows \ac{LAI} as dominant predictor
+in this region. However, in the absence of a robust physical mechanism this
+connection cannot be proven. For both target variables, the \ac{GBRT} algorithm
+is superior to all other considered statistical models in terms of prediction
+uncertainty evaluated in a leave-one-model-out \ac{CV} approach among different
+statistical models. Compared to the unweighted \acs{CMIP}5 \ac{MMM} the mean
+prediction error is reduced by $48 \unit{\%}$ (step 2a) and $16 \unit{\%}$
+(step 2b); compared to a linear \ac{LASSO} regression model the mean prediction
+error is reduced by approximately $3 \unit{\%}$ for both cases. However, local
+standard errors are still large for both target variables, even for the
+\ac{GBRT} model. Due to the unknown covariance structure a global aggregation
+of these errors is not possible. Step 2a mainly corrects the bias of the
+simulated absolute historical \ac{GPP} relative to observations
+\autocite{Anav2013}. Consequently, the historical \ac{GPP} is by far the most
+important predictor for future absolute \ac{GPP} (regions with high \ac{GPP} in
+the past are likely to have high \ac{GPP} in the future and vice versa).
+However, as we show in \cref{fig:06:step2a_results_biases}, the \ac{GBRT} model
+expands this bias correction by taking more predictors than the historical
+\ac{GPP} into account. A similar result is provided by step 2b:
+\cref{fig:06:step2b_results_biases:b,fig:06:step2b_results_biases:c} show the
+impact of the additional predictors as difference between the \acs{CMIP}5
+\ac{MMM} and the \ac{GBRT} prediction of the \ac{GPP} change. In this case, the
+target variable (fractional \ac{GPP} change) is already normalized with the
+historical \ac{GPP}. Therefore, \ac{T} (second most important feature in step
+2a) is now the most important predictor instead of historical \ac{GPP} (see
+\cref{fig:06:step2b_ml_inference:d}). As shown in
+\cref{fig:06:step2b_ml_inference:c}, the different features have also different
+dominant regional impacts; but again, since there is no distinct physical
+mechanism relating the predictors to the target variable this map requires
+further analysis and needs to be treated with care.
+
+Our \ac{GBRT} approach (step 2) is mathematically similar to \ac{MDER}
+\autocite{Karpechko2013, Senftleben2020, Wenzel2016a}: We establish a
+relationship between process-oriented, physically-relevant diagnostics and
+future projections and then utilize this to predict today's observed conditions
+into the future. We emphasize that the exact nature of this relationship is
+strongly dependent on the climate model ensemble considered. Although derived
+empirically, we argue that this relationship can still be used to extract
+further information from the climate model ensemble for two reasons: First, we
+consider process-oriented variables which are physically linked to \ac{GPP},
+and second, we ensure its statistical robustness. The latter is achieved by
+considering gridded climatological data instead of global or regional means.
+The number of points establishing the relation is dramatically increased from
+about 50 points for classical emergent constraints to 237852 points (step 2a)
+and 16503 (step 2b) in our approach. We validate the relationship by extensive
+out-of-sample testing using a leave-one-model-out \ac{CV} approach and randomly
+excluding parts of the data prior to training. Furthermore, the large number of
+points in combination with the non-linear \ac{GBRT} model allows us to exploit
+more information than classical emergent constraints and include non-linear
+multivariate relations in the statistical model, including the information
+encoded in spatial variation. The prediction phase of our approach (feeding
+observation-based data into the trained \ac{GBRT} model) can be interpreted as
+an implicit performance weighting: The \ac{GBRT} model creates predictions
+based on today's conditions (in form of variables which are physically
+connected to \ac{GPP}). However, unlike other performance weighting techniques
+\autocite{Knutti2017a, Sanderson2017}, we do not assign constant weights to the
+different climate models. On the contrary, since the algorithm works on grid
+cell level, it is able to adapt to regional climate characteristics and thus
+applies an implicit \emph{local} and \emph{climate-dependent} weighting.
+Moreover, the algorithm adds another level of weighting by assigning a relative
+relevance score to the different predictors. We emphasize that this is only an
+implicit weighting, since it is not possible to extract specific values for the
+individual weights due to the complex structure of the \ac{GBRT} model.
+
+Yet, like emergent constraints and performance weighting techniques, our
+\ac{GBRT} approach in step 2 hinges on the assumption that the climate models
+reflect the real world. This is certainly only partially true, since there are
+processes like \ac{CO2} fertilization, land use and land cover changes,
+nutrient cycles and limitations \autocite{Du2020,Fleischer2019}, disturbances,
+and induced vegetation dynamics which may strongly alter the future
+trajectories of \ac{GPP} without this being fully encoded in today's
+observational products. Moreover, all \ac{CO2}-related effects are only
+implicitly incorporated in our approach by the target period we are using and
+the global rescaling using the emergent constraint in step 1. Thus, our
+reduction of uncertainty only applies to the specific setup we are considering
+here: the future \ac{GPP} in the \acs{CMIP}5 \acs{RCP}8.5 scenario at the end
+of the \nth{21} century. It has to be viewed as a nominal reduction, while the
+real uncertainty remains unknown and is likely larger than the nominal one.
+These issues could be tested in the future by including other (even offline)
+models, which account for the aforementioned processes. A true validation will
+only be possible once we experience the changed condition or could be based on
+paleo data. In summary, our approach does not address the coupled system
+directly but hinges on the climate predictions of the individual \acp{ESM} and
+hence is not able to implicitly correct respective biases of the \acs{CMIP}5
+ensemble.
+
+The presented approach based on \ac{ML} is not limited to projections of the
+future carbon cycle, which is used as an illustrative example. Indeed, since
+its only prerequisites are the availability of gridded climate model data and
+gridded observation-based data, the \ac{GBRT} algorithm can be applied to any
+variable of interest if physically-relevant diagnostics that influence the
+target variable are known. This opens a wide range of possibilities for
+constraining uncertainties in projected variable scenarios. In addition, maps
+of the local relative feature importance as shown in
+\cref{fig:06:step2b_ml_inference:c} can be used to reveal connections between
+different Earth system variables that are currently unknown. With the proposed
+data-driven approach, a possibly extended application could be to constrain
+projections of not just one but several variables simultaneously. Moreover, the
+concept of our method is not limited to \ac{GBRT} but can be used with any
+other (\ac{ML}) regression algorithm.

### 2021-02-22 ch07_summary_outlook/summary_outlook.tex
Initialized chapters
Added first figure
Added figures for ESMValTool contributions (chapter 3)
Added further tables for chapter 05 and text for chapter 06
Finished chapter 06

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                             Summary and Outlook                             %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Summary and Outlook}
+\label{ch:07:summary_outlook}
 
-The \ac{ECS} is really cool. I like it very much!
+Summary:
+
+- Here we present methods to constrain uncertainties in multi-model projections
+of the future climate.
+
+- ESMValTool is great tool to evaluate multi-model output (thus also
+projections)
+
+- ESMValTool is expanded and also used to create analyses presented in chapters
+4, 5 and 6
+
+- We find high multi-model means and high ranges of ECS and TCR in new CMIP6
+ensemble (reasons, specific numbers)
+
+- State-of-the-art methods (emergent constraints) cannot reduce uncertainties
+in CMIP6 ECS projections (single-process, more complex models, etc.)
+
+- Add numbers for chapter 5
+
+- Novel ML-based approach: Cannot be used on ECS (scalar values)
+
+- Add numbers for chapter 6
+
+Outlook:
+
+- Why GBRT not used for ECS?
+
+- GBRT used for temperature anomaly based on other studies (emergent
+constraints, historical T trend, weighting following Brunner et al.)

### 2021-02-24 ch01_introduction/introduction.tex
Restyled title page and renamed final chapter

@@
 
+\begingroup
+\crefalias{enumi}{question}
 \begin{enumerate}
-  \item What is the range of climate sensitivity in the latest generation of
-  \acp{ESM} compared to previous multi-model ensembles, and do we understand
-  the processes that determine this uncertainty range?
+  \item \label{enum:01:question_1} What is the range of climate sensitivity in
+  the latest generation of \acp{ESM} compared to previous multi-model
+  ensembles, and do we understand the processes that determine this
+  uncertainty range?
 
-  \item Can uncertainties in climate sensitivity be reduced with observations
-  using the emergent constraint approach?
+  \item \label{enum:01:question_2} Can uncertainties in climate sensitivity be
+  reduced with observations using the emergent constraint approach?
 
-  \item Can uncertainties in multi-dimensional (gridded) climate projections be
-  reduced with \ac{ML} techniques and observations?
+  \item \label{enum:01:question_3} Can uncertainties in multi-dimensional
+  (gridded) climate projections be reduced with \ac{ML} techniques and
+  observations?
 \end{enumerate}
+\endgroup
 
@@
 published in the \emph{Journal of Geophysical Research: Biogeosciences}
-\autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook} provides a
-summary of the results of this thesis and gives an outlook of possible future
-works.
+\autocite{Schlund2020}. Finally, \cref{ch:07:conclusion} provides a summary of
+the results of this thesis and gives an outlook of possible future works.

### 2021-02-25 ch01_introduction/introduction.tex
Added overall summary

@@
 This \namecref{sec:01:key_science_questions} will present the \textbf{AWESOME}
-key science questions!!!
+\namecrefs{enum:01:question_1}!!
 
@@
 \begin{enumerate}
-  \item \label{enum:01:question_1} What is the range of climate sensitivity in
-  the latest generation of \acp{ESM} compared to previous multi-model
-  ensembles, and do we understand the processes that determine this
-  uncertainty range?
-
-  \item \label{enum:01:question_2} Can uncertainties in climate sensitivity be
-  reduced with observations using the emergent constraint approach?
-
-  \item \label{enum:01:question_3} Can uncertainties in multi-dimensional
-  (gridded) climate projections be reduced with \ac{ML} techniques and
-  observations?
+  \item \label{enum:01:question_1} \KeyScienceQuestionOne{}
+  \item \label{enum:01:question_2} \KeyScienceQuestionTwo{}
+  \item \label{enum:01:question_3} \KeyScienceQuestionThree{}
 \end{enumerate}

### 2021-02-25 ch06_paper_gpp/paper_gpp.tex
Added overall summary

@@
 relationship between process-oriented, physically-relevant diagnostics and
-future projections and then utilize this to predict today's observed conditions
+future projections and then utilize this to project today's observed conditions
 into the future. We emphasize that the exact nature of this relationship is

### 2021-02-25 ch07_conclusion/conclusion.tex
Restyled title page and renamed final chapter
Added overall summary

--- 2020-11-14 ch01_introduction/introduction.tex
@@
 %=============================================================================%
-%                                Introduction                                 %
+%                                 Conclusion                                  %
 %=============================================================================%
 
-\chapter{Introduction}
-\label{ch:introduction}
+\chapter{Conclusion}
+\label{ch:07:conclusion}
 
-The \ac{ECS} is really cool. I like it very much!
+\section{Overall Summary}
+\label{sec:07:overall_summary}
+
+The analysis of future climate projections from numerical climate model
+simulations is of paramount importance for an accurate quantification of the
+future climate change. Since this involves metrics of public interest like the
+allowable fossil fuel emissions to meet particular warming targets, for example
+the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015},
+this scientific question is not only relevant to climate scientists but to
+policymakers and the whole human society. In the light of the large spread of
+climate sensitivity in the most recent generation of climate models from
+\acs{CMIP}6 \autocite{Meehl2020}, a careful statistical evaluation and
+refinement of the output of multi-model climate projections is as relevant as
+ever. This thesis quantifies the associated uncertainties, presents the
+evaluation of established methods to reduce these uncertainties in the new
+climate model ensemble and finally introduces a novel alternative technique
+based on supervised \ac{ML}.
+
+To answer the \namecrefs{enum:01:question_1} posed in
+\cref{sec:01:key_science_questions} and to ensure a consistent evaluation of
+the participating climate model ensembles in this thesis, the \ac{ESMValTool}
+is used, an open-source community diagnostics and performance metrics tool for
+the routine evaluation of \acp{ESM}. All analyses shown in this thesis are
+performed with the \ac{ESMValTool}. Apart from that, further substantial
+changes and additions to the code base of the tool have been implemented (see
+\cref{ch:03:esmvaltool}), which lead to co-authorship in the scientific
+documentation of the \ac{ESMValTool} \autocite{Eyring2020, Lauer2020,
+  Righi2020, Weigel2020}. Since the open-source tool is freely available for
+everyone, these improvements are beneficial for the entire scientific
+community.
+
+In a first analysis, we evaluate the climate sensitivity metrics \ac{ECS} and
+\ac{TCR} for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
+which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
+published in \textcite{Bock2020, Meehl2020}, directly addresses
+\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS},
+we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which is higher
+on the upper and lower end than any model range from previous \ac{CMIP}
+generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of
+\ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus
+$3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
+\acs{CMIP}6 models range is considerably larger than in the assessed range of
+$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
+\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
+\unit{K}$ is exceeded by about a third of the \acs{CMIP}6 models, with many
+models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
+range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
+\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
+\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
+climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
+representation of mixed-phase clouds above the Southern Ocean. For one thing,
+this change leads to an improved shortwave \ac{CRE} in these models when
+compared to observations. However, this change also substantially reduces the
+strong negative shortwave cloud feedback over the Southern Ocean that is
+present in previous \ac{CMIP} generations and that results from a cloud phase
+change from ice clouds in the present-day to liquid clouds in the future. In
+the affected \acs{CMIP}6 models, this cloud phase change is not possible
+anymore since the models predominantly simulate liquid clouds over the
+present-day Southern Ocean.
+
+In order to reduce this large range of \ac{ECS} in the latest generation of
+climate  models, we evaluate suitable emergent constraints on the \acs{CMIP}6
+ensemble. Emergent constraints use a physically-based inter-model relationship
+between an observable quantity of the Earth system and a target variable to
+reduce uncertainties in the target variable with observations. In total we
+assess eleven emergent constraints on \ac{ECS}, which are mostly related to
+cloud feedbacks since these constitute the most important source of uncertainty
+for \ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated
+emergent constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5
+ensemble, we can perform out-of-sample tests on the emergent constraints by
+assessing whether they still hold for the \acs{CMIP}6 models. In this study,
+which is shown in \cref{ch:05:paper_ecs} and already published in
+\textcite{Schlund2020a}, we find a substantial reduction of skill for the
+majority of emergent constraints when applied to the \acs{CMIP}6 ensemble in
+comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
+decrease of the coefficient of determination $R^2$ of the emergent relationship
+and a decrease of the statistical significance using the null hypothesis that
+there is no correlation between the predictor and \ac{ECS}. Moreover, the
+corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
+\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
+constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
+(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
+constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
+\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
+$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
+\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
+emergent constraints. Thus, \cref{enum:01:question_2}
+(\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{no}
+for the \acs{CMIP}6 ensemble. The increased best estimates and spreads
+resulting from the emergent constraints in \acs{CMIP}6 are likely related to
+the increased \ac{MMM} and multi-model spread of \ac{ECS} in \acs{CMIP}6. A
+possible reason for the reduced skill of the emergent constraints when applied
+to the \acs{CMIP}6 ensemble is the increased complexity of the \acs{CMIP}6
+models: A basic assumption for these single-process-oriented emergent
+constraints is that a single observable process dominates the uncertainty in
+\ac{ECS}, which might not be valid anymore due to an increased number of
+processes that are included in the \acs{CMIP}6 models.
+
+To overcome these issues of single-process-oriented emergent constraints,
+\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
+This work is already published in \textcite{Schlund2020}. Since the new
+technique relies on a large number of data points in order to train the \ac{ML}
+algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
+not an appropriate target variable. As an alternative, we study projections of
+the future \ac{GPP}, which is the largest flux of the terrestrial carbon uptake
+and slows down global warming by removing \ac{CO2} from the atmosphere. In the
+first step of our new two-step approach, we constrain the global mean \ac{GPP}
+at the end of the \nth{21} century in the \acs{CMIP}5 \acs{RCP}8.5 scenario to
+$\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
+\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models' GPP
+response to \ac{CO2}, which shows a range of $\rangeunit{156}{248}{\GtCyr{}}$
+in the raw multi-model ensemble. In the second step, a \ac{ML}-based climate
+model weighting approach is used to further constrain the gridded \ac{GPP}
+based on present-day predictors that are relevant for the simulation of
+\ac{GPP} in the \acp{ESM}. Our \ac{ML} approach is mathematically similar to
+the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a},
+but additionally considers multi-dimensional (gridded) target variables and
+non-linear relationships between the predictors and the target variable. We
+establish a relationship between process‐oriented predictors and future
+projections of \ac{GPP} and then utilize this to project today's observed
+conditions into the future. The prediction phase of our method can be
+interpreted as an implicit performance weighting. However, due to the complex
+structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
+extract specific values for the individual weights. We consider two target
+variables: the gridded monthly climatologies of absolute \ac{GPP}
+(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
+\nth{21} century (2100 versus 2000). The latter quantity shows an increased
+\acs{GPP} change in the high latitudes compared to regions closer to the
+equator. The results of both approaches are consistent with each other and with
+the global constraint of the first step. We validate our approach by comparing
+it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
+\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
+(\ac{LASSO}), we find a reduction of the resulting mean \ac{RMSEP} of up to $48
+\unit{\%}$ ($3 \unit{\%}$) when using our \ac{ML} approach. Moreover, the
+evaluation of the global and local feature importance allows us to get further
+insights into the \ac{ML} model. For the first target variable (absolute
+\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
+can be explained with a correction of the historical bias in \ac{GPP} by our
+approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
+features. This study directly addresses \cref{enum:01:question_3}
+(\KeyScienceQuestionThree{}), which we can answer with \enquote{yes} based on
+the results presented here.
+
+\section{Outlook}
+\label{sec:07:outlook}
+
+- Why GBRT not used for ECS?
+
+- GBRT used for temperature anomaly based on other studies (emergent
+constraints, historical T trend, weighting following Brunner et al.)
+
+- Missing feedback process relevant for high ECS (Grise et al. 2016)?
+
+- Causal model evaluation (Nowack et al., Runge et al.)?

### 2021-02-25 ch07_conclusion/conclusion.tex
Added formatting of key science questions on chapter 07

@@
 published in \textcite{Bock2020, Meehl2020}, directly addresses
-\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS},
-we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which is higher
-on the upper and lower end than any model range from previous \ac{CMIP}
-generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of
-\ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus
-$3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
-\acs{CMIP}6 models range is considerably larger than in the assessed range of
-$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
-\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
-\unit{K}$ is exceeded by about a third of the \acs{CMIP}6 models, with many
-models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
-range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
+\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
+\ac{ECS}, we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which
+is higher on the upper and lower end than any model range from previous
+\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
+\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
+\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
+upper bound of the \acs{CMIP}6 models range is considerably larger than in the
+assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
+\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
+bound of $4.5 \unit{K}$ is exceeded by about a third of the \acs{CMIP}6 models,
+with many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the
+model range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
 \acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
@@
 emergent constraints. Thus, \cref{enum:01:question_2}
-(\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{no}
-for the \acs{CMIP}6 ensemble. The increased best estimates and spreads
-resulting from the emergent constraints in \acs{CMIP}6 are likely related to
-the increased \ac{MMM} and multi-model spread of \ac{ECS} in \acs{CMIP}6. A
-possible reason for the reduced skill of the emergent constraints when applied
-to the \acs{CMIP}6 ensemble is the increased complexity of the \acs{CMIP}6
-models: A basic assumption for these single-process-oriented emergent
-constraints is that a single observable process dominates the uncertainty in
-\ac{ECS}, which might not be valid anymore due to an increased number of
-processes that are included in the \acs{CMIP}6 models.
+(\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to be answered with a
+\enquote{no} for the \acs{CMIP}6 ensemble. The increased best estimates and
+spreads resulting from the emergent constraints in \acs{CMIP}6 are likely
+related to the increased \ac{MMM} and multi-model spread of \ac{ECS} in
+\acs{CMIP}6. A possible reason for the reduced skill of the emergent
+constraints when applied to the \acs{CMIP}6 ensemble is the increased
+complexity of the \acs{CMIP}6 models: A basic assumption for these
+single-process-oriented emergent constraints is that a single observable
+process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
+due to an increased number of processes that are included in the \acs{CMIP}6
+models.
 
@@
 features. This study directly addresses \cref{enum:01:question_3}
-(\KeyScienceQuestionThree{}), which we can answer with \enquote{yes} based on
-the results presented here.
+(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can answer with
+\enquote{yes} based on the results presented here.
 

### 2021-02-26 ch02_scientific_background/scientific_background.tex
Finished chapter 07

@@
   \caption{Global and annual mean net \acf{TOA} radiation $N$ versus the change
-    in global and annual mean near-surface air temperature $\Delta T$ for 150
-    years of a simulation with an abrupt quadrupling of the atmospheric
-    \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 \acl{MMM} (blue
-    circles). To account for energy leakage and model drift, a linear fit of
-    the corresponding pre-industrial control run is subtracted from the
-    \nxcotwo{4} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope
-    of the linear regression (black line) corresponds to the climate feedback
-    parameter $\lambda$, and the $y$-intercept corresponds to the radiative
-    forcing $F_\mathtxt{4x}$. These can be used to calculate the \acf{ECS} with
-    the Gregory regression method according to \cref{eq:02:ecs}
-    \autocite{Gregory2004}. \acs{ECS} is equivalently given by the
-    \xintercept{} of the linear regression line divided by $2$. Here,
-    $\mathtxt{ECS} = 3.14 \unit{K}$.}
+    in global and annual mean \acl{T} $\Delta T$ for 150 years of a simulation
+    with an abrupt quadrupling of the atmospheric \acs{CO2} concentration
+    (\nxcotwo{4}) for the \acs{CMIP}5 \acl{MMM} (blue circles). To account for
+    energy leakage and model drift, a linear fit of the corresponding
+    pre-industrial control run is subtracted from the \nxcotwo{4} simulation.
+    As given by \cref{eq:02:N_vs_delta_t}, the slope of the linear regression
+    (black line) corresponds to the climate feedback parameter $\lambda$, and
+    the \yintercept{} corresponds to the radiative forcing $F_\mathtxt{4x}$.
+    These can be used to calculate the \acf{ECS} with the Gregory regression
+    method according to \cref{eq:02:ecs} \autocite{Gregory2004}. \acs{ECS} is
+    equivalently given by the \xintercept{} of the linear regression line
+    divided by $2$. Here, $\mathtxt{ECS} = 3.14 \unit{K}$.}
   \label{fig:02:gregory_regression}
@@
     ch02_scientific_background/figs/tcr_calculation.pdf}
-  \caption{Global and annual mean near-surface air temperature change $\Delta
-    T$ for 140 years of a simulation with a $1 \unit{\%}$ per year increase of
-    the atmospheric \acs{CO2} concentration (\onepctcotwo{}) for the \acf{CESM}
-    of the \acf{NCAR} (blue circles). The temperature change is calculated
-    relative to a corresponding pre-industrial control simulation smoothed
-    with a linear fit over all 140 years. The \acf{TCR} is defined as the
-    temperature change $\Delta T$ at the time of \acs{CO2} doubling averaged
-    over a 20 year period (illustrated by the vertical dashed lines). Here,
+  \caption{Global and annual mean \acl{T} change $\Delta T$ for 140 years of a
+    simulation with a $1 \unit{\%}$ per year increase of the atmospheric
+    \acs{CO2} concentration (\onepctcotwo{}) for the \acf{CESM} of the
+    \acf{NCAR} (blue circles). The temperature change is calculated relative
+    to a corresponding pre-industrial control simulation smoothed with a
+    linear fit over all 140 years. The \acf{TCR} is defined as the temperature
+    change $\Delta T$ at the time of \acs{CO2} doubling averaged over a
+    20-year period (illustrated by the vertical dashed lines). Here,
     $\mathtxt{TCR} = 2.35 \unit{K}$ (horizontal dashed line).}

### 2021-02-26 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Finished chapter 07

@@
     \textcite{Gregory2004}: Global and annual mean net \acl{TOA} radiation $N$
-    versus the change in global and annual mean near-surface air temperature
-    $\Delta T$ for 150 years of a simulation with an abrupt quadrupling of the
-    atmospheric \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}6
-    \acs{MMM} (circles). To account for energy leakage and model drift, a
-    linear fit of the corresponding pre-industrial control run is subtracted
-    from the \nxcotwo{4} simulation. A linear regression can be used to
-    calculate the \acf{ECS} according to \cref{eq:02:ecs} with the radiative
-    forcing $F_\mathtxt{4x}$ (\yintercept) and the climate feedback parameter
+    versus the change in global and annual mean \acl{T} $\Delta T$ for 150
+    years of a simulation with an abrupt quadrupling of the atmospheric
+    \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}6 \acs{MMM}
+    (circles). To account for energy leakage and model drift, a linear fit of
+    the corresponding pre-industrial control run is subtracted from the
+    \nxcotwo{4} simulation. A linear regression can be used to calculate the
+    \acf{ECS} according to \cref{eq:02:ecs} with the radiative forcing
+    $F_\mathtxt{4x}$ (\yintercept) and the climate feedback parameter
     $\lambda$ (slope) for all 150 years (black line), which results in

### 2021-02-26 ch06_paper_gpp/paper_gpp.tex
Finished chapter 07

@@
     emergent constraint by \textcite{Wenzel2016} is used to constrain the
-    global mean fractional change in \acf{GPP} over the 21st century.
+    global mean fractional change in \acf{GPP} over the \nth{21} century.
     Moreover, this constraint is used to rescale two different gridded target

### 2021-02-26 ch07_conclusion/conclusion.tex
Finished chapter 07

@@
 $\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
-\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models' GPP
-response to \ac{CO2}, which shows a range of $\rangeunit{156}{248}{\GtCyr{}}$
-in the raw multi-model ensemble. In the second step, a \ac{ML}-based climate
-model weighting approach is used to further constrain the gridded \ac{GPP}
-based on present-day predictors that are relevant for the simulation of
-\ac{GPP} in the \acp{ESM}. Our \ac{ML} approach is mathematically similar to
-the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a},
-but additionally considers multi-dimensional (gridded) target variables and
-non-linear relationships between the predictors and the target variable. We
-establish a relationship between process‐oriented predictors and future
-projections of \ac{GPP} and then utilize this to project today's observed
-conditions into the future. The prediction phase of our method can be
-interpreted as an implicit performance weighting. However, due to the complex
-structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
-extract specific values for the individual weights. We consider two target
+\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models'
+\ac{GPP} response to \ac{CO2}, which shows a range of
+$\rangeunit{156}{248}{\GtCyr{}}$ in the raw multi-model ensemble. In the second
+step, a \ac{ML}-based climate model weighting approach is used to further
+constrain the gridded \ac{GPP} based on present-day predictors that are
+relevant for the simulation of \ac{GPP} in the \acp{ESM}. Our \ac{ML} approach
+is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
+  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
+(gridded) target variables and non-linear relationships between the predictors
+and the target variable. We establish a relationship between process‐oriented
+predictors and future projections of \ac{GPP} and then utilize this to project
+today's observed conditions into the future. The prediction phase of our method
+can be interpreted as an implicit performance weighting. However, due to the
+complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
+to extract specific values for the individual weights. We consider two target
 variables: the gridded monthly climatologies of absolute \ac{GPP}
@@
 features. This study directly addresses \cref{enum:01:question_3}
-(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can answer with
-\enquote{yes} based on the results presented here.
+(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can confidently answer
+with \enquote{yes} based on the results we find.
 
@@
 
-- Why GBRT not used for ECS?
+Climate change is a global phenomenon affecting the entire human society. Due
+to the vastly complex nature of the Earth system, accurately quantifying future
+climate change is a highly non-trivial task. This thesis covers several
+important aspects of the climate system and metrics describing the changing
+climate, but it is by no means complete and can be extended in an arbitrary
+number of ways. This final \namecref{sec:07:outlook} presents several
+suggestions on possible ways forward.
 
-- GBRT used for temperature anomaly based on other studies (emergent
-constraints, historical T trend, weighting following Brunner et al.)
+\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for the
+increased climate sensitivity in the \acs{CMIP}6 models (changes in the
+aerosol-cloud interaction and changes in the shortwave cloud phase change
+feedback over the Southern Ocean). While these are likely explanations, there
+might be additional relevant aspects that have not been analyzed in detail yet.
+This could include additional feedback processes that are not present in older
+\ac{CMIP} generations or missing feedback processes that appear in previous
+\ac{CMIP} models but are absent in the \acs{CMIP}6 ensemble. Obvious candidates
+for these feedback processes are cloud-related feedbacks, which pose a major
+source of uncertainty in climate sensitivity in modern-day \acp{ESM}
+\autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
+can further help to gain a better understanding of the \acs{CMIP}6 models and
+potentially to reduce associated uncertainties in the entire multi-model
+ensemble.
 
-- Missing feedback process relevant for high ECS (Grise et al. 2016)?
+Due to the flexibility of our \ac{ML}-based climate model weighting approach
+presented in \cref{ch:06:paper_gpp}, one natural way forward is to apply this
+technique to other target variables. While the scalar metrics \ac{ECS} or
+\ac{TCR} with one value per climate model do not offer enough training points
+for this technique, a possible alternative could be the gridded \acl{T} over
+the \nth{21} century in the \ac{SSP} scenarios. A relevant study in this
+context is given by \textcite{Brunner2020}, who use performance and
+interdependence-based climate model weighting to constrain the \acl{T} over the
+\nth{21} century based on observable process-based predictors of today's
+climate and which could serve as a possible baseline for our new \ac{ML}-based
+weighting approach. An exciting and valuable addition to our method could be
+causal inference \autocite{Nowack2020, Runge2019}. In the current approach, the
+relationships between the different predictors and the target variable are
+based on statistical correlation alone. Moreover, the local feature importance
+maps do not reveal true causal relationships between the predictors and the
+target variable, but only show the relative weight that is given to a specific
+predictor at a specific location for the prediction of the target variable by
+the \ac{ML} model. By integrating causal networks into our approach, it might
+be possible to explore and utilize true causal connections instead.
 
-- Causal model evaluation (Nowack et al., Runge et al.)?
+The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
+ever increasing computational power and resources provided by modern
+supercomputers has the potential for enormous improvements in climate modeling
+and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
+weighting approach, further possible applications of \ac{AI} and \ac{ML} in
+climate science involve for example \ac{ML}-based parametrizations that are
+learned from high-resolution climate models \autocite{Rasp2018}, the
+\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
+  Mansfield2020} or the learning of entire \acp{ESM} from observational
+products \autocite{Geer2021}. Therefore, the foundation for innovative and
+groundbreaking research in climate science has been laid, which will hopefully
+help us in further understanding and mitigating one of the greatest challenges
+for humankind today: climate change.

### 2021-03-01 ch01_introduction/introduction.tex
Finished chapter 07
Finished chapter 01

@@
 
-- Why is climate change relevant for the society?
+Climate change is one of the greatest challenges for humankind today with
+far-reaching consequences for all human beings on planet Earth. The warming of
+the climate system is \enquote{unequivocal} and \enquote{many of the observed
+  changes are unprecedented over decades to millennia} \autocite{IPCC2014}. The
+changing climate increases the \enquote{likelihood of severe, pervasive and
+  irreversible impacts for people, species and ecosystems} with \enquote{mostly
+  negative impacts for biodiversity, ecosystem services and economic
+  development} and amplifies \enquote{risks for livelihoods and for food and
+  human security} \autocite{IPCC2014}. Potential drivers for climate change
+are all natural and anthropogenic substances and processes that may alter the
+Earth's energy budget. The human influence on the climate system is clear: Ever
+increasing emissions of \acp{GHG} since the end of the pre-industrial era
+largely driven by economic and population growth led to atmospheric
+concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O} that are
+\enquote{unprecedented in at least the last 800000 years} \autocite{IPCC2014}.
+The effects of these \ac{GHG} emissions and other anthropogenic drivers have
+been \enquote{detected throughout the climate system and are extremely likely
+  to have been the dominant cause of the observed warming since the
+  mid-\nth{20} century} \autocite{IPCC2014}. The greenhouse effect is based on
+the optical properties of the \acp{GHG}: While the atmosphere is mostly
+transparent for the incoming solar (shortwave) radiation, the outgoing
+infrared (longwave) radiation that is reflected from the planet's surface is
+partly absorbed by the corresponding \ac{GHG} molecules through an excitation
+of their corresponding vibrational modes and re-emitted isotropically. This
+process traps energy near the surface and leads to a warming of the Earth's
+surface and the lower atmosphere. The individual impact of the different
+drivers of climate change measured with the so-called \emph{radiative
+forcing}, which quantifies the change in energy fluxes caused by changes in
+these drivers relative to pre-industrial conditions \autocite{IPCC2013}. The
+total radiative forcing is positive (which corresponds to a warming) and its
+largest contribution is caused by the increase in the atmospheric
+concentration of \ac{CO2} due to fossil fuel emissions since the year 1750
+\autocite{IPCC2014}. Apart from their physical warming effect on the climate,
+carbon-based \acp{GHG} like \ac{CO2} also directly influence the global
+carbon cycle, an important biogeochemical cycle of the Earth. A crucial flux
+of the carbon cycle is the \emph{\ac{GPP}}, which describes the carbon uptake
+of the terrestrial biosphere due to photosynthesis. Since this land carbon
+uptake absorbs about $30 \unit{\%}$ of the anthropogenic \ac{CO2} emissions in
+today's climate \autocite{Friedlingstein2020}, this process substantially slows
+down global warming. Other important anthropogenic drivers of climate change
+are the emission of aerosols and land use/land cover changes. Apart from
+anthropogenic drivers, there are also natural processes which impact the
+climate system like changes in the solar activity or volcanic eruptions.
+However, there is clear evidence that these natural drives alone cannot explain
+the observed climate change \autocite{Haustein2017}.
 
-- Cite latest warming reports, also other variables (droughts, extreme events,
-etc.)
+To successfully mitigate the massive impacts of climate change, a first
+important step is the understanding of climate change and its accurate
+quantification. Extremely valuable tools in this context are climate models,
+which allow us to simulate the behavior of the climate system under arbitrary
+conditions without having to perform (ethnically questionable) experiments in
+the real world. All around the world climate research institutes provide a
+variety of different climate models. Many of them participate in the \ac{CMIP},
+which was initiated in 1995 by the \ac{WGCM} of the \ac{WCRP} to
+\enquote{better understand past, present and future climate changes arising
+  from natural, unforced variability or in response to changes in radiative
+  forcing in a multi-model context} \autocite{WCRP2020}. The \ac{CMIP} models
+provide crucial input for the international climate assessments given by the
+\acp{AR} of the \ac{IPCC}. For example, the latest generation of climate models
+from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
+  as \acs{CMIP}6} will be used in the upcoming Sixth \acl{AR} (\acs{AR}6) of
+the \ac{IPCC}, and their predecessor models from \acs{CMIP}5
+\autocite{Taylor2012} have been successfully evaluated in the Fifth \acl{AR}
+(\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
+allow the simulation of biological and chemical processes in addition to the
+dynamics of the physical components of the Earth system, are also known as
+\acp{ESM} and provide the most sophisticated simulations of the Earth's
+climate. In this thesis, the terms \enquote{climate model} and
+\enquote{\acl{ESM}} are used interchangeably since most modern models
+participating in \ac{CMIP} are \acp{ESM} or at least have \ac{ESM} versions.
 
-- basic physics of greenhouse gases (vibration modes, etc.)
-
-- definition of radiative forcing
-
-- well-mixed greenhouse gases (water vapor, CO2, CH4) + sources
-
-- long-lived vs. short-lived GHGs?
-
-- aerosols + cooling effect
-
-- natural vs. anthropogenic effects (D\&A) -> clear evidence that climate
-change over 21st century is caused by human influence. Website:
-(https://globalwarmingindex.org/, Haustein et al., 2017)
-
-- remove description of emergent constraints from section 1.3, since this will
-probably be included in this section.
-
-- answer question: what is climate sensitivity? What is \ac{ECS}? What is
-\ac{TCR}? What is \ac{CMIP}?
-
-- One sentence on \ac{ML}.
-
-- Mention IPCC assessments, older ones and upcoming AR6
-
-- "climate model" is used as generic term which include \acp{AOGCM} and
-\acp{ESM} (see \cref{subsec:02:numerical_climate_modeling}) (subcategories).
-Since most modern climate models participating in CMIP are \acp{ESM} (or at
-least have \ac{ESM} versions), the terms "climate model" and "ESM" are used
-interchangeably in this thesis.
+Simulations that extrapolate state of the climate system into the future are
+called \emph{projections} of the future climate. These include idealized
+simulations with only a prescribed change in the atmospheric \ac{CO2}
+concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
+increase of $1 \unit{\%}$ per year) as well as more realistic projections that
+consider different future scenarios (\eg{} a fossil fuel--based future or a
+scenario that is based on a sustainable development). In many variables that
+are relevant for climate change, multi-model projections from \ac{CMIP} show a
+large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
+policy-relevant example for this is the climate sensitivity, which refers to
+the change in the \ac{GSAT} that results from a change in the radiative
+forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
+equilibrium response of the climate system after a doubling of the atmospheric
+\ac{CO2} concentration and the \emph{\ac{TCR}}, which describes the transient
+response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
+have been assessed with high ranges of $\rangeunit{1.5}{4.5}{K}$ and
+$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
+\autocite{Stocker2013}. The corresponding inter-model ranges from the
+\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
+a careful statistical evaluation and further refinement of the output of
+multi-model climate projections is necessary in order to reduce associated
+uncertainties. State-of-the-art techniques for this are so-called
+\emph{emergent constraints}, which use a physically-based inter-model
+relationship between an observable quantity of the Earth system and a target
+variable to reduce uncertainties in the target variable with observations
+\autocite{Allen2002}. An alternative approach is climate model weighting based
+on performance (\ie{} the distance of the climate models relative to
+observational products) and interdependence (\ie{} the distance between the
+climate models among each other) \autocite{Knutti2017a}. These techniques form
+the baseline for the new analyses and results presented in this thesis.
 
@@
 
-This \namecref{sec:01:key_science_questions} will present the \textbf{AWESOME}
-\namecrefs{enum:01:question_1}!!
+The aim of this thesis is to reduce uncertainties in multi-model climate
+projections with observations by addressing the following three
+\namecrefs{enum:01:question_1}:
 
@@
 covers the assessment of climate sensitivity metrics like the \ac{ECS} or
-\ac{TCR} in the latest generation of \acp{ESM} (\acs{CMIP}6). This work is
+\ac{TCR} in the latest generation of \acp{ESM} from \acs{CMIP}6. This work is
 already published in two scientific publications \autocite{Bock2020,
@@
 new climate model generation, \cref{ch:05:paper_ecs} describes the assessment
-of emergent constraints (a technique to reduce uncertainties in multi-model
-climate projections, see \cref{subsec:02:emergent_constraints}) on the \ac{ECS}
-for these \acp{ESM}. The contents of this \namecref{ch:05:paper_ecs} are
-published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.
-\Cref{ch:06:paper_gpp} focuses on a new method to reduce uncertainties in
-multi-dimensional (gridded) multi-model projections of the future climate with
-observations based on \ac{ML}. As an example, the method is applied to the
-photosynthesis rate at the end of the \nth{21} century, which is already
-published in the \emph{Journal of Geophysical Research: Biogeosciences}
+of emergent constraints on the \ac{ECS} for these \acs{CMIP}6 models and
+compares these to results derived from \acs{CMIP}5 models. The contents of this
+\namecref{ch:05:paper_ecs} are published in \emph{Earth System Dynamics}
+\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
+reduce uncertainties in multi-dimensional (gridded) multi-model projections of
+the future climate with observations based on \ac{ML}. As an example, the
+method is applied to \ac{GPP} at the end of the \nth{21} century, which is
+already published in the \emph{Journal of Geophysical Research: Biogeosciences}
 \autocite{Schlund2020}. Finally, \cref{ch:07:conclusion} provides a summary of

### 2021-03-01 ch02_scientific_background/scientific_background.tex
Finished chapter 01

@@
 
-An important metric describing climate sensitivity is the equilibrium climate
-sensitivity. It is defined as the change in \ac{GSAT} after an instantaneous
-doubling of the atmospheric \ac{CO2} concentration from pre-industrial
-conditions once the climate system reaches radiative equilibrium
+An important metric describing climate sensitivity is the \emph{equilibrium
+  climate sensitivity}. It is defined as the change in \ac{GSAT} after an
+instantaneous doubling of the atmospheric \ac{CO2} concentration from
+pre-industrial conditions once the climate system reaches radiative equilibrium
 \autocite{Bindoff2013}. Being already used in one of the first assessments of
@@
 For this reason, the equilibrium climate sensitivity is commonly approximated
-with the \acf{ECS}, which can be derived from only 150 model years of a
+with the \emph{\acf{ECS}}, which can be derived from only 150 model years of a
 simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
@@
 
-A further metric describing the climate sensitivity is the \ac{TCR}. In
+A further metric describing the climate sensitivity is the \emph{\ac{TCR}}. In
 contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
@@
 photosynthesis with about $120 \unit{\GtCyr{}}$ \autocite{Ciais2013}. This
-process is also known as \ac{GPP}. The carbon fixed into plants can be released
-back into the atmosphere by autotrophic (plant) and heterotrophic (soil
-microbial and animal) respiration and additional disturbance processes like
-fires \autocite{Ciais2013}. Since the land \ac{CO2} uptake by photosynthesis
-occurs only during the growing season, whereas respiration occurs nearly all
-year, the larger amount of vegetation in the Northern Hemisphere (due to the
-larger land mass) gives rise to a seasonal cycle of the atmospheric \ac{CO2}
-concentration \autocite{Keeling1995}. This seasonal cycle reflects the phase of
-the global carbon cycle and shows a maximum of the atmospheric \ac{CO2}
-concentration in the Northern Hemisphere winter (net \ac{CO2} flux into
-atmosphere due to respiration) and a minimum during the Northern Hemisphere
-summer (net \ac{CO2} flux into the land due to photosynthesis). Another major
-carbon exchange process connects the atmosphere and the ocean. Atmospheric
-\ac{CO2} is exchanged with the surface ocean through gas exchange, which is
-driven by the partial \ac{CO2} pressure difference between the air and the sea
-\autocite{Ciais2013}.
+process is also known as \emph{\ac{GPP}}. The carbon fixed into plants can be
+released back into the atmosphere by autotrophic (plant) and heterotrophic
+(soil microbial and animal) respiration and additional disturbance processes
+like fires \autocite{Ciais2013}. Since the land \ac{CO2} uptake by
+photosynthesis occurs only during the growing season, whereas respiration
+occurs nearly all year, the larger amount of vegetation in the Northern
+Hemisphere (due to the larger land mass) gives rise to a seasonal cycle of the
+atmospheric \ac{CO2} concentration \autocite{Keeling1995}. This seasonal cycle
+reflects the phase of the global carbon cycle and shows a maximum of the
+atmospheric \ac{CO2} concentration in the Northern Hemisphere winter (net
+\ac{CO2} flux into atmosphere due to respiration) and a minimum during the
+Northern Hemisphere summer (net \ac{CO2} flux into the land due to
+photosynthesis). Another major carbon exchange process connects the atmosphere
+and the ocean. Atmospheric \ac{CO2} is exchanged with the surface ocean through
+gas exchange, which is driven by the partial \ac{CO2} pressure difference
+between the air and the sea \autocite{Ciais2013}.
 

### 2021-03-01 ch07_conclusion/conclusion.tex
Finished chapter 01

@@
 between an observable quantity of the Earth system and a target variable to
-reduce uncertainties in the target variable with observations. In total we
-assess eleven emergent constraints on \ac{ECS}, which are mostly related to
-cloud feedbacks since these constitute the most important source of uncertainty
-for \ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated
-emergent constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5
-ensemble, we can perform out-of-sample tests on the emergent constraints by
-assessing whether they still hold for the \acs{CMIP}6 models. In this study,
-which is shown in \cref{ch:05:paper_ecs} and already published in
-\textcite{Schlund2020a}, we find a substantial reduction of skill for the
-majority of emergent constraints when applied to the \acs{CMIP}6 ensemble in
-comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
-decrease of the coefficient of determination $R^2$ of the emergent relationship
-and a decrease of the statistical significance using the null hypothesis that
-there is no correlation between the predictor and \ac{ECS}. Moreover, the
-corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
-\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
-constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
-(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
-constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
-\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
-$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
-\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
-emergent constraints. Thus, \cref{enum:01:question_2}
-(\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to be answered with a
-\enquote{no} for the \acs{CMIP}6 ensemble. The increased best estimates and
-spreads resulting from the emergent constraints in \acs{CMIP}6 are likely
-related to the increased \ac{MMM} and multi-model spread of \ac{ECS} in
-\acs{CMIP}6. A possible reason for the reduced skill of the emergent
-constraints when applied to the \acs{CMIP}6 ensemble is the increased
+reduce uncertainties in the target variable with observations
+\autocite{Allen2002}. In total we assess eleven emergent constraints on
+\ac{ECS}, which are mostly related to cloud feedbacks since these constitute
+the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
+  Flato2013}. Since all of the evaluated emergent constraints have been derived
+on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, we can perform out-of-sample tests
+on the emergent constraints by assessing whether they still hold for the
+\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
+already published in \textcite{Schlund2020a}, we find a substantial reduction
+of skill for the majority of emergent constraints when applied to the
+\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
+skill is expressed as a decrease of the coefficient of determination $R^2$ of
+the emergent relationship and a decrease of the statistical significance using
+the null hypothesis that there is no correlation between the predictor and
+\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
+show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
+confidence) for almost all emergent constraints, resulting in values of
+$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
+$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
+an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
+the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
+\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
+increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
+\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
+be answered with a \enquote{no} for the \acs{CMIP}6 ensemble. The increased
+best estimates and spreads resulting from the emergent constraints in
+\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
+of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
+emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
 complexity of the \acs{CMIP}6 models: A basic assumption for these
@@
 products \autocite{Geer2021}. Therefore, the foundation for innovative and
-groundbreaking research in climate science has been laid, which will hopefully
-help us in further understanding and mitigating one of the greatest challenges
-for humankind today: climate change.
+groundbreaking research in climate science for the near future has been laid,
+which will hopefully help us in further understanding and fighting one of the
+greatest challenges for humankind today: climate change.

### 2021-03-02 02_abstract/abstract_english.tex
Added first part of English abstract and cleaned tex files

@@
 
-TBA.
+A precise quantification of climate change is crucial in order to assess
+optimal mitigation and adaptation strategies. Valuable tools in this context
+are \acp{ESM}, which are state-of-the-art climate models that allow numerical
+simulations of the complex physical, biological and chemical processes of the
+Earth system. Of particular relevance are projections of the future climate,
+which enable an assessment of future climate change. Due to the chaotic nature
+of the climate system, unknowns in future emission pathways and uncertainties
+in the climate models, projections of the future climate are associated with
+large uncertainties. In this thesis, we analyze future climate projections from
+\acp{ESM} participating in the \ac{CMIP} and reduce corresponding uncertainties
+with observations.

### 2021-03-02 ch01_introduction/introduction.tex
Added first part of English abstract and cleaned tex files

@@
 multi-model climate projections is necessary in order to reduce associated
-uncertainties. State-of-the-art techniques for this are so-called
-\emph{emergent constraints}, which use a physically-based inter-model
-relationship between an observable quantity of the Earth system and a target
-variable to reduce uncertainties in the target variable with observations
-\autocite{Allen2002}. An alternative approach is climate model weighting based
-on performance (\ie{} the distance of the climate models relative to
-observational products) and interdependence (\ie{} the distance between the
-climate models among each other) \autocite{Knutti2017a}. These techniques form
-the baseline for the new analyses and results presented in this thesis.
+uncertainties. A state-of-the-art technique for this is the \emph{emergent
+  constraints} method, which uses a physically-based inter-model relationship
+between an observable quantity of the Earth system and a target variable to
+reduce uncertainties in the target variable with observations
+\autocite{Allen2002}. An alternative approach is the weighting of climate
+models based on their performance (\ie{} the distance of one model to
+observational products) and interdependence (\ie{} the distance of one model to
+other climate models) \autocite{Knutti2017a}. These techniques form the
+baseline for the new analyses and results presented in this thesis, which
+partly utilize methods from a new emerging research field in climate sciences:
+\ac{AI} and \ac{ML}.
 

### 2021-03-02 ch02_scientific_background/scientific_background.tex
Added first part of English abstract and cleaned tex files

@@
 positive for both the terrestrial biosphere and the ocean
-\autocite{Gregory2009}. In the first case, temperature and precipitation
-changes due to anthropogenic activities decrease the land carbon uptake because
-of increased temperature and water stress on photosynthesis and higher
-ecosystem respiration costs, which accelerates global warming due to more
-\ac{CO2} that remains in the atmosphere. For the ocean, increased temperatures
-lead to a reduction of vertical transport in the ocean resulting from increased
-stability and reduced solubility of \ac{CO2} in the sea, which reduces the
-ocean carbon uptake and enhances climate change \autocite{Gregory2009}.
+\autocite{Gregory2009}. In the first case, temperature and \acl{PR} changes due
+to anthropogenic activities decrease the land carbon uptake because of
+increased temperature and water stress on photosynthesis and higher ecosystem
+respiration costs, which accelerates global warming due to more \ac{CO2} that
+remains in the atmosphere. For the ocean, increased temperatures lead to a
+reduction of vertical transport in the ocean resulting from increased stability
+and reduced solubility of \ac{CO2} in the sea, which reduces the ocean carbon
+uptake and enhances climate change \autocite{Gregory2009}.
 
@@
 \autocite{Amos2020}, North American maximum temperature \autocite{Lorenz2018},
-European temperature and precipitation \autocite{Brunner2019, Merrifield2020}
-and global warming over the \nth{21} century \autocite{Brunner2020,
-  Liang2020a}.
+European temperature and \acl{PR} \autocite{Brunner2019, Merrifield2020} and
+global warming over the \nth{21} century \autocite{Brunner2020, Liang2020a}.
 

### 2021-03-02 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added first part of English abstract and cleaned tex files

@@
 
-Various improvements of the underlying physical, chemical and biological
+Various improvements of the underlying physical, biological and chemical
 processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to

### 2021-03-02 ch05_paper_ecs/paper_ecs.tex
Added overall summary
Added first part of English abstract and cleaned tex files

@@
 \ac{IPCC} (\acs{AR}6), this is not only important for the scientific community,
-but also directly affects society and policymakers. Due to the many new and
+but also directly affects policymakers and society. Due to the many new and
 improved processes included to the models, the \acs{CMIP}6 ensemble is arguably
@@
 \textcite{Tian2015} finds a link between mid-tropospheric humidity over the
-tropical Pacific and simulated moisture, precipitation, clouds, large-scale
+tropical Pacific and simulated moisture, \acl{PR}, clouds, large-scale
 circulation, and thus \ac{ECS} in \acs{CMIP}3 and \acs{CMIP}5 models. The study
 explains this link with the similarity of mid-tropospheric humidity and
-precipitation patterns as both are related to the \ac{ITCZ}. The proposed
-tropical mid-tropospheric humidity asymmetry index to constrain \ac{ECS} is
-defined as relative bias (in $\si{\%}$) in simulated annual mean $500
-\unit{hPa}$ specific humidity averaged over the Southern Hemisphere tropical
-Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{80
-  \unit{\degree W}}{120 \unit{\degree E}}$) minus the bias averaged over the
-Northern Hemisphere tropical Pacific ($\rangeunit{0}{20}{\degree N}$,
-$\range{80 \unit{\degree W}}{120 \unit{\degree E}}$) when compared with
-observations. Similar to the SU constraint, the index proposed by
-\textcite{Tian2015} seems to be related to the emergent constraint by
-\textcite{Fasullo2012}, who find correlations between relative humidity of the
-middle and upper troposphere and \ac{ECS}. Here, we use humidity observations
-from AIRS \autocite{Aumann2003} over the time period \range{2003}{2005} as the
-reference dataset. We assess a $66 \unit{\%}$ \ac{ECS} range of
-$\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5 ($R^2 = 0.24$) and
-$\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 = 0.06$).
-\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
+\acl{PR} patterns as both are related to the \ac{ITCZ}. The proposed tropical
+mid-tropospheric humidity asymmetry index to constrain \ac{ECS} is defined as
+relative bias (in $\si{\%}$) in simulated annual mean $500 \unit{hPa}$ specific
+humidity averaged over the Southern Hemisphere tropical Pacific ($\range{30
+  \unit{\degree S}}{0 \unit{\degree N}}$, $\range{80 \unit{\degree W}}{120
+  \unit{\degree E}}$) minus the bias averaged over the Northern Hemisphere
+tropical Pacific ($\rangeunit{0}{20}{\degree N}$, $\range{80 \unit{\degree
+    W}}{120 \unit{\degree E}}$) when compared with observations. Similar to
+the SU constraint, the index proposed by \textcite{Tian2015} seems to be
+related to the emergent constraint by \textcite{Fasullo2012}, who find
+correlations between relative humidity of the middle and upper troposphere and
+\ac{ECS}. Here, we use humidity observations from AIRS \autocite{Aumann2003}
+over the time period \range{2003}{2005} as the reference dataset. We assess a
+$66 \unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5
+($R^2 = 0.24$) and $\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 =
+0.06$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
 significance of the emergent relationship drops massively from $p = 0.0089$ in
@@
 \autocite{Bellucci2010, Hirota2011}. This index is defined as the
-climatological annual mean precipitation bias averaged over the southeastern
-Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$,
+climatological annual mean \acl{PR} bias averaged over the southeastern Pacific
+($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$,
 $\rangeunit{150}{100}{\degree W}$). The southern \ac{ITCZ} index is calculated
@@
 problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds a link
-between double-\ac{ITCZ} bias and simulated moisture, precipitation, clouds,
-and large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He argues
-that this could explain the link between the double-\acs{ITCZ} bias and
-\ac{ECS}. As reference data, we use observed precipitation data for the years
-\range{1986}{2005} from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66
-\unit{\%}$ confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5
-($R^2 = 0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 <
-0.01$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
-emergent relationship shows a much lower statistical significance in
-\acs{CMIP}6 ($p = 0.8236$) than in \acs{CMIP}5 ($p = 0.0013$).
+between double-\ac{ITCZ} bias and simulated moisture, \acl{PR}, clouds, and
+large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He argues that
+this could explain the link between the double-\acs{ITCZ} bias and \ac{ECS}. As
+reference data, we use observed \acl{PR} data for the years \range{1986}{2005}
+from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66 \unit{\%}$
+confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5 ($R^2 =
+0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 < 0.01$).
+\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The emergent
+relationship shows a much lower statistical significance in \acs{CMIP}6 ($p =
+0.8236$) than in \acs{CMIP}5 ($p = 0.0013$).
 

### 2021-03-02 end_acronyms/acronyms.tex
Added figures of appendix B to tex file
Added text to chapter 06
Finished chapter 07
Finished chapter 01
Added first part of English abstract and cleaned tex files

@@
   \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
+  \acro{AI}{artificial intelligence}
   \acro{AOGCM}{atmosphere-ocean general circulation model}
@@
   \acro{DJF}{December-January-February}
-  \acro{DLR}{Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace %
-    Center)}
   \acro{ECS}{effective climate sensitivity}
@@
   \acro{GHG}{greenhouse gas}
+  \acrodefplural{GHG}{greenhouse gases}
   \acro{GPP}{gross primary production}
@@
   \acro{IAM}{integrated assessment model}
-  \acro{IPA}{Institut für Physik der Atmosphäre (Institute of Atmospheric %
-    Physics)}
   \acro{IPCC}{Intergovernmental Panel on Climate Change}
@@
   \acro{MMM}{multi-model mean}
+  \acro{N2O}[N\textsubscript{2}O]{nitrous oxide}
   \acro{NAO}{North Atlantic Oscillation}
@@
   \acro{RCP}{Representative Concentration Pathway}
-  \acro{RMSE}{root-mean-square error}
-  \acro{RMSEP}{root-mean-square error of prediction}
-  \acro{RSDS}{downwelling solar radiation at surface}
+  \acro{RMSE}{root mean square error}
+  \acro{RMSEP}{root mean square error of prediction}
+  \acro{RSDS}{incoming solar radiation at surface}
   \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
+  \acro{SPE}{standard prediction error}
   \acro{SSP}{Shared Socioeconomic Pathway}

### 2021-03-02 end_appendix/appendix.tex
Added overall summary
Added first part of English abstract and cleaned tex files

@@
 
-\chapter{Appendix}
+\addchap{Appendix}
 
@@
     hus & Specific humidity \\
-    pr & Precipitation \\
+    pr & \Acl{PR} \\
     rsdt & \Acf{TOA} incoming shortwave radiation \\

### 2021-03-03 02_abstract/abstract_english.tex
Added English abstract and first part of acknowledgements

@@
 optimal mitigation and adaptation strategies. Valuable tools in this context
-are \acp{ESM}, which are state-of-the-art climate models that allow numerical
-simulations of the complex physical, biological and chemical processes of the
-Earth system. Of particular relevance are projections of the future climate,
-which enable an assessment of future climate change. Due to the chaotic nature
-of the climate system, unknowns in future emission pathways and uncertainties
-in the climate models, projections of the future climate are associated with
-large uncertainties. In this thesis, we analyze future climate projections from
-\acp{ESM} participating in the \ac{CMIP} and reduce corresponding uncertainties
-with observations.
+are \acfpAbstract{ESM}, which are state-of-the-art climate models that allow
+numerical simulations of the complex physical, biological and chemical
+processes of the Earth system. Of particular relevance are projections of the
+future climate, which enable an assessment of future climate change. Due to the
+chaotic nature of the climate system, unknowns in future emission pathways and
+uncertainties in the climate models, projections of the future climate are
+associated with large uncertainties. In this thesis, we analyze future climate
+projections from \acsp{ESM} participating in the \acfAbstract{CMIP} and reduce
+corresponding uncertainties with observations.
+
+In a first step, the climate sensitivity (\ie{} the temperature response of the
+climate system to an external forcing) in the latest generation of \acsp{ESM}
+from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, the equilibrium
+temperature response that follows a doubling of the atmospheric
+\acfAbstract{CO2} concentration, we find a \acfAbstract{MMM} and multi-model
+range of $3.74 \unit{K}$ and $\rangeunit{1.8}{5.6}{K}$, respectively. These
+values are higher than in any previous \acs{CMIP} ensemble before. Moreover, a
+third of the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely
+\acs{ECS} range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR}
+(\acs{AR}5) of the \acl{IPCC} from 2013. Similarly, the transient response of
+the climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
+shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is
+again higher than the likely range assessed in \acs{AR}5 of
+$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
+sensitivity in many \acs{CMIP}6 models is a change in the microphysical
+representation of mixed-phase clouds above the Southern Ocean, which reduces
+the strong negative shortwave cloud phase change feedback in this region that
+is present in climate models from previous \acs{CMIP} generations.
+
+To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models, we
+systematically analyze eleven published emergent constraints on \acs{ECS} that
+have mostly been derived from models participating in \acs{CMIP}5, the
+predecessor generation of \acs{CMIP}6. Emergent constraints are potentially
+promising approaches to reduce uncertainties in climate model projections by
+combining observations and output from \acsp{ESM}. The focus of this study is
+on testing if these emergent constraints hold for \acsp{ESM} participating
+\acs{CMIP}6. Since none of the emergent constraints considered here have been
+derived on the \acs{CMIP}6 ensemble, the \acs{CMIP}6 models can be used for
+cross-checking of the emergent constraints on a new model ensemble. The
+application of the emergent constraints to \acs{CMIP}6 data shows a decrease in
+skill and statistical significance of the emergent relationship for nearly all
+constraints, with this decrease being large in many cases. Consequently, the
+size of the constrained \acs{ECS} ranges ($66 \unit{\%}$ confidence intervals)
+widens by $51 \unit{\%}$ on average in \acs{CMIP}6 compared to \acs{CMIP}5.
+This is likely related to the increased multi-model spread of \acs{ECS} in
+\acs{CMIP}6, but may in some cases also be due to spurious statistical
+relationships or a too small number of models in the ensemble that the emergent
+constraint was originally derived from. The corresponding best estimates given
+by the emergent constraints also increase from \acs{CMIP}5 to \acs{CMIP}6 by
+$12 \unit{\%}$ on average. This can be at least partly explained by the
+increased number of high-\acs{ECS} models in \acs{CMIP}6 without a
+corresponding change in the constraint predictors, suggesting the emergence of
+new feedback processes rather than changes in strength of those previously
+dominant. Our results support previous studies concluding that emergent
+constraints should be based on an independently verifiable physical mechanism
+and that process-based emergent constraints on \acs{ECS} should rather be
+thought of as constraints for the process or feedback they are actually
+targeting.
+
+To overcome these issues of single-process-oriented emergent constraints, we
+introduce an alternative approach based on \acfAbstract{ML}. Since this new
+technique relies on a large number of data points in order to train the
+\acs{ML} algorithm, the scalar climate sensitivity expressed as \acs{ECS} or
+\acs{TCR} is not an appropriate target variable. As an alternative, we study
+projections of the future \acfAbstract{GPP}, which is the largest flux of the
+terrestrial carbon uptake and slows down global warming by removing \acs{CO2}
+from the atmosphere. In this analysis, we combine an existing emergent
+constraint on \acs{CO2} fertilization with a \acs{ML} approach to constrain the
+spatial variations of multi-model \acs{GPP} projections. In the first step of
+our two-step approach, we use observed changes in the \acs{CO2} seasonal cycle
+at \acl{KUM} to constrain the global mean \acs{GPP} at the end of the \nth{21}
+century (\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM}
+participating in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
+the unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second
+step, we use a \acs{ML} model to constrain gridded future absolute \acs{GPP}
+and gridded fractional \acs{GPP} change in two independent approaches. For
+this, observational data is fed into the \acs{ML} algorithm that has been
+trained on \acs{CMIP}5 data to learn relationships between present-day
+physically-relevant diagnostics and the target variable. In a
+leave-one-model-out \acl{CV} approach, the \acs{ML} model shows superior
+performance to the \acs{CMIP}5 \acs{MMM}. Our approach predicts an increased
+\acs{GPP} change in northern high latitudes compared to regions closer to the
+equator.

### 2021-03-03 ch07_conclusion/conclusion.tex
Added English abstract and first part of acknowledgements

@@
 \ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
-bound of $4.5 \unit{K}$ is exceeded by about a third of the \acs{CMIP}6 models,
-with many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the
-model range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
+bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
+many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
+range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
 \acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
@@
 \ac{GPP} response to \ac{CO2}, which shows a range of
-$\rangeunit{156}{248}{\GtCyr{}}$ in the raw multi-model ensemble. In the second
+$\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In the second
 step, a \ac{ML}-based climate model weighting approach is used to further

### 2021-03-03 end_acknowledgments/acknowledgments.tex
Cleaned tex files and added text to chapter 4
Added English abstract and first part of acknowledgements

--- 2020-11-13 end_statutory_declaration/statutory_declaration.tex
@@
 %=============================================================================%
-%                            Statutory declaration                            %
+%                               Acknowledgments                               %
 %=============================================================================%
 
-\addchap{Statutory Declaration}
+\chapter{Acknowledgments}
 
-I assure that this thesis is a result of my personal work and that no other than the indicated aids have been used for its completion.
-Furthermore I assure that all quotations and statements that have been inferred literally or in a general manner from published or unpublished writings are marked as such.
-Beyond this I assure that the work has not been used, neither completely nor in parts, to pass any previous examination.
+This work has been supported by the European Union's Horizon 2020 project
+\emph{Coordinated Research in Earth Systems and Climate: Experiments,
+  kNowledge, Dissemination and Outreach (CRESCENDO)} under grant agreement
+number 641816, the European Union's Horizon 2020 project \emph{Climate-Carbon
+  Interactions in the Coming Century (4C)} under grant agreement number 821003,
+the European Research Council Synergy Grant \emph{Understanding and Modelling
+  the Earth System with Machine Learning (USMILE)} under grant agreement number
+855187 and the \emph{Advanced Earth System Model Evaluation for CMIP
+  (EVal4CMIP)} project funded by the Helmholtz Society. I acknowledge the
+\ac{WCRP}'s \ac{WGCM}, which promoted \acs{CMIP}5 and \acs{CMIP}6. I thank the
+climate modeling groups for producing and making available their model output,
+the \ac{ESGF} for archiving the data and providing access and the multiple
+funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational
+resources of the Deutsches Klima Rechenzentrum (DKRZ, Hamburg, Germany) where
+the \ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
+kindly acknowledged.
 
-\vspace{20mm}
-
-\noindent \thePlace{}, \theDate{}
-
-\begin{flushright}
-  $\overline{~~~~~~~~~~~~~\raisebox{0mm}[5mm]{\theAuthor{}}~~~~~~~~~~~~~}$
-\end{flushright}
+Personal acknowledgments...

### 2021-03-03 ch01_introduction/introduction.tex
Restructured thesis and removed unnecessary parts

@@
 %=============================================================================%
-%                                Dissertation                                 %
+%                            Doctoral Dissertation                            %
 %                               Manuel Schlund                                %
-%                                  (c) 2020                                   %
+%                                  (c) 2021                                   %
 %=============================================================================%
@@
 Parts of this thesis are already published in multiple peer-reviewed
-publications (two first-author studies and six co-author studies). Wherever
-material from these studies is presented in this thesis, the pronoun
-\enquote{we} is used to increase readability by avoiding the passive voice and
-to acknowledge all involved contributors. However, unless stated otherwise, all
-contents from these publications (text, figures and tables) shown in this
-thesis originate from the author of this thesis. A detailed list of
-contributions to these studies (including their full reference) is given in the
-corresponding \namecrefs{ch:03:esmvaltool}.
+publications (two first-author studies and six co-author studies). A complete
+list of these is given on \cpageref{ch:bib:own}. Wherever material from these
+studies is presented in this thesis, the pronoun \enquote{we} is used to
+increase readability by avoiding the passive voice and to acknowledge all
+involved contributors. However, unless stated otherwise, all contents from
+these publications (text, figures and tables) shown in this thesis originate
+from the author of this thesis. A detailed list of contributions to these
+studies (including their full reference) is given in the corresponding
+\namecrefs{ch:03:esmvaltool}.
 

### 2021-03-04 ch02_scientific_background/scientific_background.tex
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables

@@
     ch02_scientific_background/figs/esms_historical_evolution.pdf}
-  \caption{Historical evolution of coupled climate models over the last 45
-    years. In early days, these models were so-called \acfp{AOGCM}, which only
-    included three components: the atmosphere, the land surface and the ocean.
-    Over the time, the individual components grew in complexity and included a
-    wider range of processes (illustrated by the growing cylinders).
-    Eventually, more and more components (aerosols, carbon cycle, \etc{}) were
-    added to the coupled system, forming the modern \acfp{ESM}.
-    \IPCCReproducedFrom{Cubasch2013}{their figure 1.13}.}
+  \caption[
+    Historical evolution of coupled climate models over the last 45 years.
+  ]{
+    Historical evolution of coupled climate models over the last 45 years. In
+    early days, these models were so-called \acfp{AOGCM}, which only included
+    three components: the atmosphere, the land surface and the ocean. Over the
+    time, the individual components grew in complexity and included a wider
+    range of processes (illustrated by the growing cylinders). Eventually, more
+    and more components (aerosols, carbon cycle, \etc{}) were added to the
+    coupled system, forming the modern \acfp{ESM}.
+    \IPCCReproducedFrom{Cubasch2013}{their figure 1.13}.
+  }
   \label{fig:02:esms_historical_evolution}
@@
     ch02_scientific_background/figs/cmip6.pdf}
-  \caption{Schematic of the experiment design of Phase 6 of the \acl{CMIP}
+  \caption[
+    Schematic of the experiment design of Phase 6 of the \acl{CMIP}
+    (\acs{CMIP}6).
+  ]{
+    Schematic of the experiment design of Phase 6 of the \acl{CMIP}
     (\acs{CMIP}6). The center of the circle illustrates the four \acs{DECK}
@@
     through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}.
-    \SpringerAdaptedFrom{Simpkins2017}.}
+    \SpringerAdaptedFrom{Simpkins2017}.
+  }
   \label{fig:02:cmip6}
@@
     ch02_scientific_background/figs/sources_of_uncertainty.pdf}
-  \caption{Schematic illustrating the importance of different sources of
+  \caption[
+    Schematic illustrating the importance of different sources of
+    uncertainties in climate model projections and their evolution in time.
+  ]{
+    Schematic illustrating the importance of different sources of
     uncertainties in climate model projections and their evolution in time. (a)
@@
     improvements and observational constraints.
-    \IPCCReproducedFrom{Cubasch2013}{their FAQ 1.1, figure 1}.}
+    \IPCCReproducedFrom{Cubasch2013}{their FAQ 1.1, figure 1}.
+  }
   \label{fig:02:sources_of_uncertainty}
@@
     ch02_scientific_background/figs/feedbacks_overview.pdf}
-  \caption{Climate feedbacks and corresponding time scales. \enquote{+} refers
-    to positive feedbacks, which amplify the effect of the external forcing
-    (\eg{} the water vapor feedback). \enquote{-} refers to negative feedbacks,
-    which diminish the effect of the external forcing (\eg{} the longwave
-    radiation feedback). \enquote{+/-} refers to feedbacks which might be
-    either positive or negative (\eg{} the cloud feedback). The smaller box
-    highlights the large differences in time scales for the various feedbacks.
-    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.}
+  \caption[
+    Climate feedbacks and corresponding time scales.
+  ]{
+    Climate feedbacks and corresponding time scales. \enquote{+} refers to
+    positive feedbacks, which amplify the effect of the external forcing (\eg{}
+    the water vapor feedback). \enquote{-} refers to negative feedbacks, which
+    diminish the effect of the external forcing (\eg{} the longwave radiation
+    feedback). \enquote{+/-} refers to feedbacks which might be either positive
+    or negative (\eg{} the cloud feedback). The smaller box highlights the
+    large differences in time scales for the various feedbacks.
+    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.
+  }
   \label{fig:02:feedbacks_overview}
@@
     ch02_scientific_background/figs/gregory_regression.pdf}
-  \caption{Global and annual mean net \acf{TOA} radiation $N$ versus the change
-    in global and annual mean \acl{T} $\Delta T$ for 150 years of a simulation
+  \caption[
+    Gregory regression for the \acs{CMIP}5 \acl{MMM} illustrating the
+    definition of the \acf{ECS}.
+  ]{
+    Global and annual mean net \acf{TOA} radiation $N$ versus the change in
+    global and annual mean \acl{T} $\Delta T$ for 150 years of a simulation
     with an abrupt quadrupling of the atmospheric \acs{CO2} concentration
@@
     equivalently given by the \xintercept{} of the linear regression line
-    divided by $2$. Here, $\mathtxt{ECS} = 3.14 \unit{K}$.}
+    divided by $2$. Here, $\mathtxt{ECS} = 3.14 \unit{K}$.
+  }
   \label{fig:02:gregory_regression}
@@
     ch02_scientific_background/figs/gregory_regression_different_years.png}
-  \caption{As in \cref{fig:02:gregory_regression} but for different time
-    periods considered in the Gregory regression resulting in different values
-    for the \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray,
-    blue, red and orange colors correspond to the \acf{CESM} of the \acf{NCAR}
-    for different time periods of a long running simulation (see legend). The
+  \caption[
+    Gregory regression for different time periods.
+  ]{
+    As in \cref{fig:02:gregory_regression} but for different time periods
+    considered in the Gregory regression resulting in different values for the
+    \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
+    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
+    different time periods of a long running simulation (see legend). The
     initial years are simulated many times for different initial conditions.
     Light gray colors correspond to the \acs{CMIP}5 ensemble (150 years each).
-    \SpringerAdaptedFrom{Knutti2017}.}
+    \SpringerAdaptedFrom{Knutti2017}.
+  }
   \label{fig:02:gregory_regression_different_years}
@@
     ch02_scientific_background/figs/tcr_calculation.pdf}
-  \caption{Global and annual mean \acl{T} change $\Delta T$ for 140 years of a
+  \caption[
+    Illustration of the definition of the \acf{TCR}.
+  ]{
+    Global and annual mean \acl{T} change $\Delta T$ for 140 years of a
     simulation with a $1 \unit{\%}$ per year increase of the atmospheric
     \acs{CO2} concentration (\onepctcotwo{}) for the \acf{CESM} of the
-    \acf{NCAR} (blue circles). The temperature change is calculated relative
-    to a corresponding pre-industrial control simulation smoothed with a
-    linear fit over all 140 years. The \acf{TCR} is defined as the temperature
-    change $\Delta T$ at the time of \acs{CO2} doubling averaged over a
-    20-year period (illustrated by the vertical dashed lines). Here,
-    $\mathtxt{TCR} = 2.35 \unit{K}$ (horizontal dashed line).}
+    \acf{NCAR} (blue circles). The temperature change is calculated relative to
+    a corresponding pre-industrial control simulation smoothed with a linear
+    fit over all 140 years. The \acf{TCR} is defined as the temperature change
+    $\Delta T$ at the time of \acs{CO2} doubling averaged over a 20-year period
+    (illustrated by the vertical dashed lines). Here, $\mathtxt{TCR} = 2.35
+    \unit{K}$ (horizontal dashed line).
+  }
   \label{fig:02:tcr}
@@
     ch02_scientific_background/figs/carbon_cycle_schematic.png}
-  \caption{Simplified schematic of the global carbon cycle including the
-    typical turnover time scales for carbon transfers through the major
-    reservoirs (atmosphere, land surface and ocean).
-    \IPCCReproducedFrom{Ciais2013}{their FAQ 6.2, figure 1}.}
+  \caption[
+    Simplified schematic of the global carbon cycle.
+  ]{
+    Simplified schematic of the global carbon cycle including the typical
+    turnover time scales for carbon transfers through the major reservoirs
+    (atmosphere, land surface and ocean). \IPCCReproducedFrom{Ciais2013}{their
+      FAQ 6.2, figure 1}.
+  }
   \label{fig:02:carbon_cycle_schematic}
@@
     ch02_scientific_background/figs/keeling_curve.pdf}
-  \caption{The Keeling Curve: monthly mean atmospheric \acs{CO2} concentration
-    at the Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
+  \caption[
+    The Keeling Curve.
+  ]{
+    The Keeling Curve: monthly mean atmospheric \acs{CO2} concentration at the
+    Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
     \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
@@
     concentration is superimposed with a seasonal oscillation caused by the
-    seasonal \acs{CO2} cycle (see \cref{subsec:02:carbon_cycle_overview}).}
+    seasonal \acs{CO2} cycle (see \cref{subsec:02:carbon_cycle_overview}).
+  }
   \label{fig:02:keeling_curve}
@@
     ch02_scientific_background/figs/carbon_cycle_perturbation.png}
-  \caption{Schematic representation of the overall perturbation of the global
-    carbon cycle caused by anthropogenic activities, averaged globally for the
-    decade \range{2010}{2019}. Arrows represent carbon exchange fluxes; circles
-    carbon reservoirs. More details are given in the legend of this figure.
-    \AdaptedFrom{Friedlingstein2020}.}
+  \caption[
+    Schematic representation of the overall perturbation of the global carbon
+    cycle caused by anthropogenic activities.
+  ]{
+    Schematic representation of the overall perturbation of the global carbon
+    cycle caused by anthropogenic activities, averaged globally for the decade
+    \range{2010}{2019}. Arrows represent carbon exchange fluxes; circles carbon
+    reservoirs. More details are given in the legend of this figure.
+    \AdaptedFrom{Friedlingstein2020}.
+  }
   \label{fig:02:carbon_cycle_perturbation}
@@
     ch02_scientific_background/figs/emergent_constraint.pdf}
-  \caption{Schematic illustration of the emergent constraint approach. Basis of
-    every emergent constraint is a robust and physically-interpretable emergent
+  \caption[
+    Schematic illustration of the emergent constraint approach.
+  ]{
+    Schematic illustration of the emergent constraint approach. Basis of every
+    emergent constraint is a robust and physically-interpretable emergent
     relationship (red line) between a target variable $y$ (\eg{} Earth system
@@
     shaded area) and uncertainties in the emergent relationship (red shaded
-    area). \SpringerAdaptedFrom{Eyring2019}.}
+    area). \SpringerAdaptedFrom{Eyring2019}.
+  }
   \label{fig:02:emergent_constraint}
@@
     ch02_scientific_background/figs/mder_schematic.pdf}
-  \caption{Schematic illustration of the \acf{MDER} approach
-    \autocite{Karpechko2013}. First, inter-model relationships between a target
-    variable $y$ and multiple process-based predictors $x^{(k)}$ (here: two
-    predictors $x^{(1)}$ and $x^{(2)}$) are used to fit a multivariate linear
-    regression model (gray surface). Second, observations of the predictors
-    (horizontal black lines) are fed into the regression model to calculate an
-    observation-based best estimate of the target variable $\hat{y}_0$
-    following \cref{eq:02:mder_best_estimate}. The black circle indicates the
-    best estimate for the target variable $y$ given by the observed values of
-    $x^{(1)}$ and $x^{(2)}$. Each of the remaining colored circles represents
-    a single climate model of the multi-model ensemble. The vertical dashed
-    lines visualize the distance between the climate model data and the linear
+  \caption[
+    Schematic illustration of the \acf{MDER} approach.
+  ]{
+    Schematic illustration of the \acf{MDER} approach \autocite{Karpechko2013}.
+    First, inter-model relationships between a target variable $y$ and multiple
+    process-based predictors $x^{(k)}$ (here: two predictors $x^{(1)}$ and
+    $x^{(2)}$) are used to fit a multivariate linear regression model (gray
+    surface). Second, observations of the predictors (horizontal black lines)
+    are fed into the regression model to calculate an observation-based best
+    estimate of the target variable $\hat{y}_0$ following
+    \cref{eq:02:mder_best_estimate}. The black circle indicates the best
+    estimate for the target variable $y$ given by the observed values of
+    $x^{(1)}$ and $x^{(2)}$. Each of the remaining colored circles represents a
+    single climate model of the multi-model ensemble. The vertical dashed lines
+    visualize the distance between the climate model data and the linear
     regression surface, which represents the noise term $\bm{\varepsilon}$ in
-    \cref{eq:02:multivariate_linear_model}.}
+    \cref{eq:02:multivariate_linear_model}.
+  }
   \label{fig:02:mder}

### 2021-03-04 ch03_esmvaltool/esmvaltool.tex
Restyled title page and renamed final chapter
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables

@@
 
-In order to answer the key science questions of this thesis posed in
+In order to answer the \namecrefs{enum:01:question_1} of this thesis posed in
 \cref{sec:01:key_science_questions}, a reliable and efficient tool to read,
@@
     ch03_esmvaltool/figs/esmvaltool_schematic.pdf}
-  \caption{Schematic representation of the \acf{ESMValTool}. First, input data
-    is preprocessed by the \acf{ESMValCore}. Second, the preprocessed data is
-    read by diagnostic scripts, which create the final output (\eg{} plots).
-    The user can specify the desired input datasets, preprocessing operations
-    and diagnostic scripts in the recipe, which is a configuration file that
-    controls the main workflow of the \acs{ESMValTool}.
-    \AdaptedFrom{Righi2020}.}
+  \caption[
+    Schematic representation of the structure of the \acf{ESMValTool}.
+  ]{
+    Schematic representation of the structure of the \acf{ESMValTool}. First,
+    input data is preprocessed by the \acf{ESMValCore}. Second, the
+    preprocessed data is read by diagnostic scripts, which create the final
+    output (\eg{} plots). The user can specify the desired input datasets,
+    preprocessing operations and diagnostic scripts in the recipe, which is a
+    configuration file that controls the main workflow of the \acs{ESMValTool}.
+    \AdaptedFrom{Righi2020}.
+  }
   \label{fig:03:esmvaltool_schematic}
@@
   \end{tabulary}
-  \caption{Summary of new preprocessor functions and variable derivation
-    scripts contributed to the \acf{ESMValCore} by the author of this thesis.}
+  \caption[
+    Summary of new preprocessor functions and variable derivation scripts
+    contributed to the \acf{ESMValCore} by the author of this thesis.
+  ]{
+    Summary of new preprocessor functions and variable derivation scripts
+    contributed to the \acf{ESMValCore} by the author of this thesis.
+  }
   \label{tab:03:changes_to_esmvalcore}
@@
   \end{subfigure}
-  \caption{(a) \acs{CMIP}6 \acf{MMM} of the atmospheric \acs{CO2} concentration
-    at surface (\emph{co2s}) averaged over the months \acf{JJA} of the year
-    2014 in the emission-driven historical simulation. The \acs{CMIP}6
-    \acs{MMM} includes the four climate models that provide all data needed
-    for the calculation of \emph{co2s} (see legend in (b)). (b) Monthly mean
+  \caption[
+    Illustration of the atmospheric \acs{CO2} concentration at surface
+    (\emph{co2}) for \acs{CMIP}6 models.
+  ]{
+    (a) \acs{CMIP}6 \acf{MMM} of the atmospheric \acs{CO2} concentration at
+    surface (\emph{co2s}) averaged over the months \acf{JJA} of the year 2014
+    in the emission-driven historical simulation. The \acs{CMIP}6 \acs{MMM}
+    includes the four climate models that provide all data needed for the
+    calculation of \emph{co2s} (see legend in (b)). (b) Monthly mean
     \emph{co2s} at Mauna Loa, Hawaii ($19.5 \unit{\degree N}$, $155.6
@@
     show emission-driven historical simulations from individual \acs{CMIP}6
-    models.}
+    models.
+  }
   \label{fig:03:co2s}
@@
   \end{tabulary}
-  \caption{Summary of new recipes and \acs{CMOR}ization scripts for
-    observational data contributed to the \acf{ESMValTool} by the author of
-    this thesis. The corresponding file names of the recipes in the
-    \acs{ESMValTool} repository are given by
-    \emph{recipe\_(name\_in\_table).yml}.}
+  \caption[
+    Summary of new recipes and \acs{CMOR}ization scripts for observational data
+    contributed to the \acf{ESMValTool} by the author of this thesis.
+  ]{
+    Summary of new recipes and \acs{CMOR}ization scripts for observational data
+    contributed to the \acf{ESMValTool} by the author of this thesis. The
+    corresponding file names of the recipes in the \acs{ESMValTool} repository
+    are given by \emph{recipe\_(name\_in\_table).yml}.
+  }
   \label{tab:03:changes_to_esmvaltool}
@@
   \end{subfigure}
-  \caption{Emergent constraint on the \acf{ECS} from global temperature
-    variability \autocite{Cox2018}. Magenta colors indicate high-sensitivity
-    climate models ($\lambda < 1 \unit{W m^{-2} K^{-1}}$), and green colors
-    indicate low sensitivity climate models ($\lambda > 1 \unit{W m^{-2}
-      K^{-1}}$). Both panels have been created with the \acs{ESMValTool} using
+  \caption[
+    Emergent constraint on the \acf{ECS} from global temperature variability.
+  ]{
+    Emergent constraint on the \acf{ECS} from global temperature variability
+    \autocite{Cox2018}. Magenta colors indicate high-sensitivity climate models
+    ($\lambda < 1 \unit{W m^{-2} K^{-1}}$), and green colors indicate low
+    sensitivity climate models ($\lambda > 1 \unit{W m^{-2} K^{-1}}$). Both
+    panels have been created with the \acs{ESMValTool} using
     \emph{recipe\_cox18nature.yml} and are similar to \textcite{Cox2018}. (a)
@@
     line) and standard deviation (dashed lines). The letters represent
-    individual \acs{CMIP}5 models (see \textcite{Cox2018} for details).}
+    individual \acs{CMIP}5 models (see \textcite{Cox2018} for details).
+  }
   \label{fig:03:cox18}
@@
   \end{subfigure}
-  \caption{Relationships between climate metrics within the \acs{CMIP}5 climate
-    model ensemble. Both panels have been created with the \acs{ESMValTool}
-    using \emph{recipe\_flato13ipcc.yml} and are similar to
-    \textcite{Flato2013}. (a) \Acf{GSAT} versus the \acf{ECS} for \acs{CMIP}5
-    models for the period \range{1961}{1990} (larger symbols) and for
-    pre-industrial control runs (smaller symbols). (b) \acf{TCR} versus
-    \acs{ECS} for \acs{CMIP}5 models. The black line shows a linear fit
-    between \acs{TCR} and \acs{ECS}.}
+  \caption[
+    Relationships between climate metrics within the \acs{CMIP}5 climate model
+    ensemble.
+  ]{
+    Relationships between climate metrics within the \acs{CMIP}5 climate model
+    ensemble. Both panels have been created with the \acs{ESMValTool} using
+    \emph{recipe\_flato13ipcc.yml} and are similar to \textcite{Flato2013}. (a)
+    \Acf{GSAT} versus the \acf{ECS} for \acs{CMIP}5 models for the period
+    \range{1961}{1990} (larger symbols) and for pre-industrial control runs
+    (smaller symbols). (b) \acf{TCR} versus \acs{ECS} for \acs{CMIP}5 models.
+    The black line shows a linear fit between \acs{TCR} and \acs{ECS}.
+  }
   \label{fig:03:flato13}

### 2021-03-04 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables

@@
     ch04_papers_ecs_tcr_assessment/figs/gregory_regression_cmip6_mmm.pdf}
-  \caption{Gregory regression for the \acs{CMIP}6 \acf{MMM} following
+  \caption[
+    Gregory regression for the \acs{CMIP}6 \acl{MMM}.
+  ]{
+    Gregory regression for the \acs{CMIP}6 \acf{MMM} following
     \textcite{Gregory2004}: Global and annual mean net \acl{TOA} radiation $N$
@@
     \acf{ECS} according to \cref{eq:02:ecs} with the radiative forcing
-    $F_\mathtxt{4x}$ (\yintercept) and the climate feedback parameter
-    $\lambda$ (slope) for all 150 years (black line), which results in
-    $\mathtxt{ECS} = 3.74 \unit{K}$. Due to non-linear effects, the exact
-    value of \acs{ECS} depends on the years considered for the Gregory
-    regression (see \cref{subsec:02:ecs} and
-    \cref{fig:02:gregory_regression_different_years}): Using only the first 20
-    years of the simulation (blue circles and line) yields a considerably
-    smaller \acs{ECS} than using only the last 130 years of the simulation
-    (orange circles and line). \AdaptedFrom{Meehl2020}.}
+    $F_\mathtxt{4x}$ (\yintercept) and the climate feedback parameter $\lambda$
+    (slope) for all 150 years (black line), which results in $\mathtxt{ECS} =
+    3.74 \unit{K}$. Due to non-linear effects, the exact value of \acs{ECS}
+    depends on the years considered for the Gregory regression (see
+    \cref{subsec:02:ecs} and \cref{fig:02:gregory_regression_different_years}):
+    Using only the first 20 years of the simulation (blue circles and line)
+    yields a considerably smaller \acs{ECS} than using only the last 130 years
+    of the simulation (orange circles and line). \AdaptedFrom{Meehl2020}.
+  }
   \label{fig:04:gregory_regression_cmip6_mmm}
@@
     \dataset & $\idx$ & $\ecs$ & $\tcr$}
-  \caption{\Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
-    Details on the calculation of \acs{ECS} and \acs{TCR} are given in
+  \caption[
+    \Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
+  ]{
+    \Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models. Details on
+    the calculation of \acs{ECS} and \acs{TCR} are given in
     \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively. The \acf{MMM}
@@
     number of models). Corresponding references for each model are given in
-    \cref{tab:app:a:cmip5_models}. \AdaptedFrom{Meehl2020}.}
+    \cref{tab:app:a:cmip5_models}. \AdaptedFrom{Meehl2020}.
+  }
   \label{tab:04:ecs_tcr_cmip5}
@@
     \dataset & $\idx$ & $\ecs$ & $\tcr$}
-  \caption{As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.
+  \caption[
+    As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.
+  ]{
+    As in \cref{tab:04:ecs_tcr_cmip5} but for the \acs{CMIP}6 models.
     Corresponding references for each model are given in
-    \cref{tab:app:a:cmip6_models}. \AdaptedFrom{Meehl2020}.}
+    \cref{tab:app:a:cmip6_models}. \AdaptedFrom{Meehl2020}.
+  }
   \label{tab:04:ecs_tcr_cmip6}
@@
     ch04_papers_ecs_tcr_assessment/figs/historical_ecs_tcr_ranges.pdf}
-  \caption{Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red
-    bars) over the years in the Charney report from 1979 \autocite{Charney1979}
-    and the subsequent \acfp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
+  \caption[
+    Historical values of the \acf{ECS} and the \acf{TCR} for different
+    \aclp{AR} and climate model ensembles.
+  ]{
+    Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red bars)
+    over the years in the Charney report from 1979 \autocite{Charney1979} and
+    the subsequent \acfp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
       Kattenberg1996, Albritton2001, Solomon2007, Stocker2013}. Orange and
@@
     numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
-    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}).
-    \AdaptedFrom{Meehl2020}.}
+    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). \AdaptedFrom{Meehl2020}.
+  }
   \label{fig:04:historical_ecs_tcr_ranges}
@@
     ch04_papers_ecs_tcr_assessment/figs/cloud_phase_change_feedback.pdf}
-  \caption{Schematic illustration of the strong negative shortwave cloud phase
-    change feedback over the Southern Ocean, which is present in earlier
-    \acs{CMIP} model generations like \acs{CMIP}5 (green arrows). Due to
-    substantial improvements in the microphysical representation of mixed-phase
-    clouds in many \acs{CMIP}6 models, the predominantly ice clouds over the
-    Southern Ocean in the present-day climate of previous climate model
-    generations have been replaced with clouds that predominantly consist of
-    (supercooled) liquid clouds \autocite{BodasSalcedo2019, Gettelman2019,
-      Mulcahy2020, Williams2020}. Thus, in these models a cloud phase change
-    due to global warming is no longer possible, which removes (or
-    significantly reduces) the strong negative cloud phase change feedback
-    (orange arrows). This leads to an increase in the global net cloud
-    feedback \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS} in the
-    corresponding \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019},
-    which is a possible explanation for the high \acs{ECS} values found in
-    many \acs{CMIP}6 models.}
+  \caption[
+    Schematic illustration of the strong negative shortwave cloud phase change
+    feedback over the Southern Ocean.
+  ]{
+    Schematic illustration of the strong negative shortwave cloud phase change
+    feedback over the Southern Ocean, which is present in earlier \acs{CMIP}
+    model generations like \acs{CMIP}5 (green arrows). Due to substantial
+    improvements in the microphysical representation of mixed-phase clouds in
+    many \acs{CMIP}6 models, the predominantly ice clouds over the Southern
+    Ocean in the present-day climate of previous climate model generations have
+    been replaced with clouds that predominantly consist of (supercooled)
+    liquid clouds \autocite{BodasSalcedo2019, Gettelman2019, Mulcahy2020,
+      Williams2020}. Thus, in these models a cloud phase change due to global
+    warming is no longer possible, which removes (or significantly reduces) the
+    strong negative cloud phase change feedback (orange arrows). This leads to
+    an increase in the global net cloud feedback \autocite{BodasSalcedo2019,
+      Tan2016} and the \acf{ECS} in the corresponding \acs{CMIP}6 models
+    \autocite{Andrews2019, Gettelman2019}, which is a possible explanation for
+    the high \acs{ECS} values found in many \acs{CMIP}6 models.
+  }
   \label{fig:04:cloud_phase_change_feedback}
@@
   \end{subfigure}
-  \caption{Geographical distributions of the net (a), shortwave (c) and
-    longwave (e) \acf{CRE} feedback parameters for the \acs{CMIP}6 \acf{MMM}
-    and the corresponding differences to the \acs{CMIP}5 \acs{MMM}, again for
-    the net (b), shortwave (d) and longwave (f) components. The \acs{CRE}
-    feedback parameters for each grid cell are calculated with a Gregory
-    regression of the grid cell's \acs{CRE} versus the change in the
-    \acl{GSAT} (see \cref{subsec:02:cloud_feedback_parameters}).
-    \AdaptedFrom{Bock2020}.}
+  \caption[
+    Geographical distributions of the \acl{CRE} feedback parameters for the
+    \acs{CMIP}5 and \acs{CMIP}6 \aclp{MMM}.
+  ]{
+    Geographical distributions of the net (a), shortwave (c) and longwave (e)
+    \acf{CRE} feedback parameters for the \acs{CMIP}6 \acf{MMM} and the
+    corresponding differences to the \acs{CMIP}5 \acs{MMM}, again for the net
+    (b), shortwave (d) and longwave (f) components. The \acs{CRE} feedback
+    parameters for each grid cell are calculated with a Gregory regression of
+    the grid cell's \acs{CRE} versus the change in the \acl{GSAT} (see
+    \cref{subsec:02:cloud_feedback_parameters}). \AdaptedFrom{Bock2020}.
+  }
   \label{fig:04:cloud_feedback_parameters}

### 2021-03-04 ch05_paper_ecs/paper_ecs.tex
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables

@@
   \end{tabular}
-  \caption{Overview of the eleven emergent constraints on the \acf{ECS}
-    evaluated in this \namecref{ch:05:paper_ecs}. Detailed descriptions of the
-    variables used to calculate the \xaxis{} values of the emergent
-    constraints and the references for the corresponding observations (fourth
-    column) are given in
+  \caption[
+    Overview of the eleven emergent constraints on the \acf{ECS} evaluated in
+    \cref{ch:05:paper_ecs}.
+  ]{
+    Overview of the eleven emergent constraints on the \acf{ECS} evaluated in
+    this \namecref{ch:05:paper_ecs}. Detailed descriptions of the variables
+    used to calculate the \xaxis{} values of the emergent constraints and the
+    references for the corresponding observations (fourth column) are given in
     \cref{tab:app:a:overview_variables,tab:app:a:observations}, respectively.
@@
     constraint. For all others, the observational value is computed with the
-    \acs{ESMValTool}. \AdaptedFrom{Schlund2020a}.}
+    \acs{ESMValTool}. \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:05:overview_emergent_constraints}
@@
   \end{subfigure}
-  \caption{Emergent constraints BRI, COX and LIP applied to the \acs{CMIP}5
-    (blue) and \acs{CMIP}6 ensemble (orange). (a), (c), (e) Emergent
-    relationships (solid lines) for the \acs{CMIP} models (numbers; see
+  \caption[
+    Emergent constraints BRI, COX and LIP applied to the \acs{CMIP}5 and
+    \acs{CMIP}6 ensemble.
+  ]{
+    Emergent constraints BRI, COX and LIP applied to the \acs{CMIP}5 (blue) and
+    \acs{CMIP}6 ensemble (orange). (a), (c), (e) Emergent relationships (solid
+    lines) for the \acs{CMIP} models (numbers; see
     \cref{tab:app:a:cmip5_emergent_constraints_part1,%
@@
     the \aclp{MMM} for \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange). (b), (d),
-    (f) \Aclp{PDF} for \acs{ECS} of the emergent constraints (solid lines; see
-    \cref{eq:02:pdf_y_given_x0}) and the unconstrained model ensembles
+    (f) \Aclp{PDF} for \acs{ECS} of the emergent constraints (solid lines;
+    see \cref{eq:02:pdf_y_given_x0}) and the unconstrained model ensembles
     (histograms). Due to the availability of data, the histograms may differ
-    for the different emergent constraints. \AdaptedFrom{Schlund2020a}.}
+    for the different emergent constraints. \AdaptedFrom{Schlund2020a}.
+  }
   \label{fig:05:bri_cox_lip}
@@
   \end{subfigure}
-  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints
-    SHD, SHL and SHS (see \cref{tab:app:a:cmip5_emergent_constraints_part1,%
+  \caption[
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SHD, SHL
+    and SHS.
+  ]{
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SHD, SHL
+    and SHS (see \cref{tab:app:a:cmip5_emergent_constraints_part1,%
       tab:app:a:cmip6_emergent_constraints_part1} for details).
-    \AdaptedFrom{Schlund2020a}.}
+    \AdaptedFrom{Schlund2020a}.
+  }
   \label{fig:05:shd_shl_shs}
@@
   \end{subfigure}
-  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU,
-    TIH and TII (see \cref{tab:app:a:cmip5_emergent_constraints_part2,%
+  \caption[
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU, TIH
+    and TII.
+  ]{
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU, TIH
+    and TII (see \cref{tab:app:a:cmip5_emergent_constraints_part2,%
       tab:app:a:cmip6_emergent_constraints_part2} for details).
-    \AdaptedFrom{Schlund2020a}.}
+    \AdaptedFrom{Schlund2020a}.
+  }
   \label{fig:05:su_tih_tii}
@@
   \end{subfigure}
-  \caption{As in \cref{fig:05:bri_cox_lip} but for the emergent constraints VOL
-    and ZHA (see \cref{tab:app:a:cmip5_emergent_constraints_part2,%
+  \caption[
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints VOL and
+    ZHA.
+  ]{
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints VOL and
+    ZHA (see \cref{tab:app:a:cmip5_emergent_constraints_part2,%
       tab:app:a:cmip6_emergent_constraints_part2} for details).
-    \AdaptedFrom{Schlund2020a}.}
+    \AdaptedFrom{Schlund2020a}.
+  }
   \label{fig:05:vol_zha}
@@
   \includegraphics[width=\LargeFigureWidth{}]{ch05_paper_ecs/figs/zha.pdf}
-  \caption{Emergent relationship ZHA \autocite{Zhai2015} for different subsets
-    of \acs{CMIP}5 models. Blue circles show the 15 \acs{CMIP}5 models used in
-    the original publication (except for CESM1-CAM5). The solid blue line and
-    blue shaded area show the emergent relationships evaluated on these models
+  \caption[
+    Emergent relationship ZHA for different subsets of \acs{CMIP}5 models.
+  ]{
+    Emergent relationship ZHA \autocite{Zhai2015} for different subsets of
+    \acs{CMIP}5 models. Blue circles show the 15 \acs{CMIP}5 models used in the
+    original publication (except for CESM1-CAM5). The solid blue line and blue
+    shaded area show the emergent relationships evaluated on these models
     including the uncertainty range. In our study, we added eleven more
@@
     show the corresponding constraints on the \acf{ECS} using this observation.
-    \AdaptedFrom{Schlund2020a}.}
+    \AdaptedFrom{Schlund2020a}.
+  }
   \label{fig:05:zha}
@@
   \end{tabular}
-  \caption{Overview of the constrained \acf{ECS} ranges and $p$-values for all
-    eleven analyzed emergent constraints. If not further specified, the
-    uncertainty ranges correspond to the $66 \unit{\%}$ confidence intervals
+  \caption[
+    Overview of the constrained \acf{ECS} ranges and $p$-values for all eleven
+    analyzed emergent constraints.
+  ]{
+    Overview of the constrained \acf{ECS} ranges and $p$-values for all eleven
+    analyzed emergent constraints. If not further specified, the uncertainty
+    ranges correspond to the $66 \unit{\%}$ confidence intervals
     ($\rangeunit{17}{83}{\%}$). For \acs{CMIP}5 and \acs{CMIP}6, these are
@@
     \cref{subsec:02:emergent_constraints} for details).
-    \AdaptedFrom{Schlund2020a}.}
+    \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:05:overview_results}
@@
     ch05_paper_ecs/figs/constrained_ranges_summary.pdf}
-  \caption{Overview of the constrained \acf{ECS} ranges given by the eleven
-    analyzed emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6
-    (orange) ensemble. The bars correspond to the $66 \unit{\%}$ confidence
-    intervals ($\rangeunit{17}{83}{\%}$) calculated from the constrained
-    \acl{PDF} of \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical
-    black lines illustrate the corresponding best estimates of the
-    distributions. For almost all emergent constraints, the constrained best
-    estimate and range of \ac{ECS} is higher in the \acs{CMIP}6 ensemble.}
+  \caption[
+    Overview of the constrained \acf{ECS} ranges given by the eleven analyzed
+    emergent constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensemble.
+  ]{
+    Overview of the constrained \acf{ECS} ranges given by the eleven analyzed
+    emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange)
+    ensemble. The bars correspond to the $66 \unit{\%}$ confidence intervals
+    ($\rangeunit{17}{83}{\%}$) calculated from the constrained \acl{PDF} of
+    \acs{ECS} following \cref{eq:02:pdf_y_given_x0}. The vertical black lines
+    illustrate the corresponding best estimates of the distributions. For
+    almost all emergent constraints, the constrained best estimate and range of
+    \ac{ECS} is higher in the \acs{CMIP}6 ensemble.
+  }
   \label{fig:05:overview_results}

### 2021-03-04 ch06_paper_gpp/paper_gpp.tex
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables

@@
   \includegraphics[width=\LargeFigureWidth{}]{ch06_paper_gpp/figs/1.pdf}
-  \caption{Schematic illustration of our two-step approach. In step 1 an
-    emergent constraint by \textcite{Wenzel2016} is used to constrain the
-    global mean fractional change in \acf{GPP} over the \nth{21} century.
-    Moreover, this constraint is used to rescale two different gridded target
-    variables: absolute \acs{GPP} at the end of the \nth{21} century and
-    fractional \acs{GPP} change over the \nth{21} century. In step 2, a
-    \acl{ML} model is used to constrain these two target variables (step 2a:
-    absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
-    independent approaches. \AdaptedFrom{Schlund2020}.}
+  \caption[
+    Schematic illustration of our two-step approach.
+  ]{
+    Schematic illustration of our two-step approach. In step 1 an emergent
+    constraint by \textcite{Wenzel2016} is used to constrain the global mean
+    fractional change in \acf{GPP} over the \nth{21} century. Moreover, this
+    constraint is used to rescale two different gridded target variables:
+    absolute \acs{GPP} at the end of the \nth{21} century and fractional
+    \acs{GPP} change over the \nth{21} century. In step 2, a \acl{ML} model is
+    used to constrain these two target variables (step 2a: absolute \acs{GPP};
+    step 2b: fractional \acs{GPP} change) in two independent approaches.
+    \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:schematic_steps}
@@
   \includegraphics[width=\LargerFigureWidth{}]{ch06_paper_gpp/figs/2.pdf}
-  \caption{Schematic illustration of our \acf{ML} approach to constrain
-    projected absolute \acf{GPP} in step 2a. (a) In the training phase of the
-    algorithm, the model is fitted to the training data interpolating the
-    empirical (non-linear) relationship between two process-oriented
-    diagnostics of the past climate $\left\{ x^{(1)}, x^{(2)} \right\}$ and
-    (rescaled) future \acs{GPP} (gray surface). The points show the training
-    points for the supervised \acs{ML} algorithm, each of them representing a
-    single grid cell/month of a single climate model (the different colors
-    correspond to different climate models). (b) In the prediction phase,
-    observation-based values of the diagnostic (black points) are fed into the
-    trained \acs{ML} model to constrain \acs{GPP} for every grid cell/month to
-    a value which best agrees with the observations. (c) For an independent
-    validation of our method, we use an out-of-sample testing setup based on a
-    leave-one-model-out \acl{CV} approach (see
-    \cref{subsec:02:model_weighting} for details). The schematic illustration
-    of step 2b differs only in the target variable used (fractional \acs{GPP}
-    change instead of absolute \acs{GPP}). \AdaptedFrom{Schlund2020}.}
+  \caption[
+    Schematic illustration of our \acl{ML} approach to constrain projected
+    absolute \acf{GPP} at the end of the \nth{21} century in step 2a.
+  ]{
+    Schematic illustration of our \acf{ML} approach to constrain projected
+    absolute \acf{GPP} in step 2a. (a) In the training phase of the algorithm,
+    the model is fitted to the training data interpolating the empirical
+    (non-linear) relationship between two process-oriented diagnostics of the
+    past climate $\left\{ x^{(1)}, x^{(2)} \right\}$ and (rescaled) future
+    \acs{GPP} (gray surface). The points show the training points for the
+    supervised \acs{ML} algorithm, each of them representing a single grid
+    cell/month of a single climate model (the different colors correspond to
+    different climate models). (b) In the prediction phase, observation-based
+    values of the diagnostic (black points) are fed into the trained \acs{ML}
+    model to constrain \acs{GPP} for every grid cell/month to a value which
+    best agrees with the observations. (c) For an independent validation of our
+    method, we use an out-of-sample testing setup based on a
+    leave-one-model-out \acl{CV} approach (see \cref{subsec:02:model_weighting}
+    for details). The schematic illustration of step 2b differs only in the
+    target variable used (fractional \acs{GPP} change instead of absolute
+    \acs{GPP}). \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:mlr_concept}
@@
   \end{tabular}
-  \caption{Process-oriented diagnostics (also called \emph{predictors} or
+  \caption[
+    Process-oriented diagnostics used in the \acf{GBRT} model to predict the
+    target variables.
+  ]{
+    Process-oriented diagnostics (also called \emph{predictors} or
     \emph{features}) used in the \acf{GBRT} model to predict the target
@@
     historical climate. For step 2b (target variable: fractional \acs{GPP}
-    change), the temporal mean over the specified time ranges is calculated
-    for all variables. \AdaptedFrom{Schlund2020}.}
+    change), the temporal mean over the specified time ranges is calculated for
+    all variables. \AdaptedFrom{Schlund2020}.
+  }
   \label{tab:06:predictors}
@@
   \end{subfigure}
-  \caption{Emergent relationship between the global mean fractional change in
+  \caption[
+    Emergent relationship between the global mean fractional change in
+    \acf{GPP} and the sensitivity of the \acs{CO2} seasonal cycle amplitude to
+    rising atmospheric \acs{CO2} concentrations observed at \acl{KUM}.
+  ]{
+    Emergent relationship between the global mean fractional change in
     \acf{GPP} and the sensitivity of the \acs{CO2} seasonal cycle amplitude to
@@
     \acs{1BGC} run is $\pmrangeunit{30}{9}{\%}$. (b) Fractional global mean
-    \acs{GPP} change over the \nth{21} century calculated from the 10-year
-    mean \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
-    the emission-driven fully-coupled \acs{RCP}8.5 simulations and the 10-year
-    mean \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in
-    the emission‐driven fully-coupled historical run (see \cref{eq:06:f}). In
+    \acs{GPP} change over the \nth{21} century calculated from the 10-year mean
+    \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
+    emission-driven fully-coupled \acs{RCP}8.5 simulations and the 10-year mean
+    \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in the
+    emission‐driven fully-coupled historical run (see \cref{eq:06:f}). In
     contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
@@
     \acs{CMIP}5 simulations are extended with the \acs{RCP}8.5 simulations for
-    the years \range{2006}{2019}). The constrained global mean \acs{GPP}
-    change over the \nth{21} century is $\pmrangeunit{39}{7}{\%}$.
-    \AdaptedFrom{Schlund2020}.}
+    the years \range{2006}{2019}). The constrained global mean \acs{GPP} change
+    over the \nth{21} century is $\pmrangeunit{39}{7}{\%}$.
+    \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:step1}
@@
   \end{subfigure}
-  \caption{(a) Box plot of the \acf{RMSEP} distributions for six different
-    statistical models used to predict future absolute \acf{GPP} in step 2a
-    using a leave-one-model-out \acl{CV} approach. The distribution for each
+  \caption[
+    Prediction error in a leave-one-model-out \acl{CV} setup and feature
+    importance of our \acl{ML} approach to constrain projected absolute
+    \acf{GPP} at the end of the \nth{21} century in step 2a.
+  ]{
+    (a) Box plot of the \acf{RMSEP} distributions for six different statistical
+    models used to predict future absolute \acf{GPP} in step 2a using a
+    leave-one-model-out \acl{CV} approach. The distribution for each
     statistical model contains seven elements (black points; one for each
@@
     \unit{\%}$ for the full \acs{GBRT} model (using all predictors). (b)
-    Relative global feature importance for the different predictors used in
-    the \acs{GBRT} model to predict future absolute \acs{GPP} (step 2a). The
-    red bars correspond to positive Pearson correlation coefficients between
-    all predictors and the target variable. Due to its strong positive linear
+    Relative global feature importance for the different predictors used in the
+    \acs{GBRT} model to predict future absolute \acs{GPP} (step 2a). The red
+    bars correspond to positive Pearson correlation coefficients between all
+    predictors and the target variable. Due to its strong positive linear
     relationship with the future \acs{GPP}, the historical \acs{GPP} is by far
@@
     using \acf{LIME} is not shown here because \acs{GPP} is the dominant
-    predictor for all grid cells. \AdaptedFrom{Schlund2020}.}
+    predictor for all grid cells. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:step2a_ml_inference}
@@
   \end{subfigure}
-  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the absolute
+  \caption[
+    Illustration of our \acl{ML} approach to constrain projected absolute
+    \acf{GPP} at the end of the \nth{21} century with observations in step 2a.
+  ]{
+    (a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the absolute
     \acf{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
     \acs{RCP}8.5 scenario using \cref{eq:06:y} (r\acs{MMM}) and its unweighted
-    version (\acs{MMM}). The plot shows an almost constant value over the
-    whole globe with a mean of $0.92$, which corresponds to the ratio of the
+    version (\acs{MMM}). The plot shows an almost constant value over the whole
+    globe with a mean of $0.92$, which corresponds to the ratio of the
     constrained global mean \acs{GPP} change over the \nth{21} century from
@@
     correct the historical bias of the \acs{CMIP}5 ensemble relative to the
-    FLUXNET‐MTE product. \AdaptedFrom{Schlund2020}.}
+    FLUXNET‐MTE product. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:step2a_results}
@@
   \end{subfigure}
-  \caption{Difference (a) and ratio (b) of the both biases shown in
+  \caption[
+    Further illustrations of our \acl{ML} approach to constrain projected
+    absolute \acf{GPP} at the end of the \nth{21} century with observations in
+    step 2a and comparison to other statistical models.
+  ]{
+    Difference (a) and ratio (b) of the both biases shown in
     \cref{fig:06:step2a_results:b,fig:06:step2a_results:c}. For panel (b), all
     values close to $0$ for the dataset in the denominator have been masked to
-    avoid divisions by $0$. Both panels show that our approach is only to
-    first order a bias correction, in which case both plots would only show
-    constant values. (c) Comparison of the \acf{GBRT} versus \acs{GBRT}-1D
-    projections of future \acf{GPP}. This panel indicates a clear difference
-    between the full \acs{GBRT} model using all predictors and the \acs{GBRT}
-    model using only historical \acs{GPP} as single predictor. (d) Comparison
-    of the \acs{GBRT} versus the \acf{LASSO} projections of future \acs{GPP}.
-    This panel shows that there is a clear difference between using the
-    non-linear \acs{GBRT} model and the linear \acs{LASSO} model. The results
-    of the \acs{LASSO}-1D model are not shown here because they are very
-    similar to the results of the \acs{LASSO} model.
-    \AdaptedFrom{Schlund2020}.}
+    avoid divisions by $0$. Both panels show that our approach is only to first
+    order a bias correction, in which case both plots would only show constant
+    values. (c) Comparison of the \acf{GBRT} versus \acs{GBRT}-1D projections
+    of future \acf{GPP}. This panel indicates a clear difference between the
+    full \acs{GBRT} model using all predictors and the \acs{GBRT} model using
+    only historical \acs{GPP} as single predictor. (d) Comparison of the
+    \acs{GBRT} versus the \acf{LASSO} projections of future \acs{GPP}. This
+    panel shows that there is a clear difference between using the non-linear
+    \acs{GBRT} model and the linear \acs{LASSO} model. The results of the
+    \acs{LASSO}-1D model are not shown here because they are very similar to
+    the results of the \acs{LASSO} model. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:step2a_results_biases}
@@
   \end{subfigure}
-  \caption{(a) Box plot of the \acf{RMSEP} distributions for four different
+  \caption[
+    Prediction error in a leave-one-model-out \acl{CV} setup and feature
+    importance of our \acl{ML} approach to constrain the fractional change in
+    \acf{GPP} over the \nth{21} century in step 2b.
+  ]{
+    (a) Box plot of the \acf{RMSEP} distributions for four different
     statistical models used to predict the fractional change in \acf{GPP} over
@@
     \acs{GBRT} model is reduced by more than $16 \unit{\%}$, $9 \unit{\%}$, and
-    $3 \unit{\%}$, respectively. (b) Relative global feature importance for
-    the different features used in the \acs{LASSO} model to predict the
-    fractional \acs{GPP} change (step 2b). The feature importance for the
-    \acs{LASSO} model is given by the normalized linear coefficients of the
-    model. (c) Local feature importance for the \acs{GBRT} model used to
-    predict fractional \acs{GPP} change (step 2b) calculated using the
-    \acf{LIME} approach \autocite{Ribeiro2016} for the three dominant features
-    \acf{T}, \acf{LAI} and \acs{GPP}. The relative influence of these three
-    features (ignoring all other features) is color-coded according to the
-    triangle in the lower left corner. Over large parts of the globe, \acs{T}
-    is the dominant feature. (d) Relative global feature importance for the
-    different features used in the \acs{GBRT} model to predict the fractional
-    \acs{GPP} change (step 2b). The blue bars for the two global feature
-    importance plots (b) and (d) correspond to negative Pearson correlation
-    coefficients between all predictors and the target variable.
-    \AdaptedFrom{Schlund2020}.}
+    $3 \unit{\%}$, respectively. (b) Relative global feature importance for the
+    different features used in the \acs{LASSO} model to predict the fractional
+    \acs{GPP} change (step 2b). The feature importance for the \acs{LASSO}
+    model is given by the normalized linear coefficients of the model. (c)
+    Local feature importance for the \acs{GBRT} model used to predict
+    fractional \acs{GPP} change (step 2b) calculated using the \acf{LIME}
+    approach \autocite{Ribeiro2016} for the three dominant features \acf{T},
+    \acf{LAI} and \acs{GPP}. The relative influence of these three features
+    (ignoring all other features) is color-coded according to the triangle in
+    the lower left corner. Over large parts of the globe, \acs{T} is the
+    dominant feature. (d) Relative global feature importance for the different
+    features used in the \acs{GBRT} model to predict the fractional \acs{GPP}
+    change (step 2b). The blue bars for the two global feature importance plots
+    (b) and (d) correspond to negative Pearson correlation coefficients between
+    all predictors and the target variable. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:step2b_ml_inference}
@@
   \end{subfigure}
-  \caption{Fractional change in \acf{GPP} over the \nth{21} century (2100
-    versus 2000) in the \acs{RCP}8.5 scenario for different statistical models
-    (step 2b): (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change,
-    (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:y}, (c)
-    linear \acf{LASSO} model and (d) \acf{GBRT} model. The geographical
-    patterns from the different statistical models are very similar and show a
-    higher \acs{GPP} increase in high latitudes and a lower \acs{GPP} closer
-    to the equator. \AdaptedFrom{Schlund2020}.}
+  \caption[
+    Illustration of our \acl{ML} approach to constrain the fractional change in
+    \acf{GPP} over the \nth{21} century with observations in step 2b.
+  ]{
+    Fractional change in \acf{GPP} over the \nth{21} century (2100 versus 2000)
+    in the \acs{RCP}8.5 scenario for different statistical models (step 2b):
+    (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change, (b) rescaled
+    \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:y}, (c) linear
+    \acf{LASSO} model and (d) \acf{GBRT} model. The geographical patterns from
+    the different statistical models are very similar and show a higher
+    \acs{GPP} increase in high latitudes and a lower \acs{GPP} closer to the
+    equator. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:step2b_results}
@@
   \end{subfigure}
-  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the
-    fractional change in \acf{GPP} over the \nth{21} century using
-    \cref{eq:06:y} (r\acs{MMM}; see \cref{fig:06:step2b_results:b}) and its
-    unweighted version (\acs{MMM}; see \cref{fig:06:step2b_results:a}). The
-    plot shows an almost constant value over the whole globe with a mean of
-    $0.91$, which corresponds to the ratio of the constrained global mean
-    \acs{GPP} change over the \nth{21} century from step 1 ($39 \unit{\%}$)
-    and the \acs{CMIP}5 ensemble mean global mean GPP change ($43 \unit{\%}$).
-    All values close to $0$ for the dataset in the denominator have been
-    masked to avoid divisions by $0$. Absolute (b) and relative (c)
-    differences between the fractional \acs{GPP} change over the \nth{21}
-    century predicted by the \acf{GBRT} model (see
-    \cref{fig:06:step2b_results:d}) and r\acs{MMM} (see
+  \caption[
+    Further illustrations of our \acl{ML} approach to constrain the fractional
+    change in \acf{GPP} over the \nth{21} century with observations in step 2b.
+  ]{
+    (a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the fractional
+    change in \acf{GPP} over the \nth{21} century using \cref{eq:06:y}
+    (r\acs{MMM}; see \cref{fig:06:step2b_results:b}) and its unweighted version
+    (\acs{MMM}; see \cref{fig:06:step2b_results:a}). The plot shows an almost
+    constant value over the whole globe with a mean of $0.91$, which
+    corresponds to the ratio of the constrained global mean \acs{GPP} change
+    over the \nth{21} century from step 1 ($39 \unit{\%}$) and the \acs{CMIP}5
+    ensemble mean global mean GPP change ($43 \unit{\%}$). All values close to
+    $0$ for the dataset in the denominator have been masked to avoid divisions
+    by $0$. Absolute (b) and relative (c) differences between the fractional
+    \acs{GPP} change over the \nth{21} century predicted by the \acf{GBRT}
+    model (see \cref{fig:06:step2b_results:d}) and r\acs{MMM} (see
     \cref{fig:06:step2b_results:b}). For panel (c), all values close to $0$ for
@@
     Australia and western parts of the United States.
-    \AdaptedFrom{Schlund2020}.}
+    \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:step2b_results_biases}
@@
   \end{subfigure}
-  \caption{Top row: Absolute future \acf{GPP} at the end of the \nth{21}
-    century (\range{2091}{2100}) calculated using step 2a (a) and step 2b (b).
-    Both approaches give similar results with global averages of $169
+  \caption[
+    Comparison of our two independent approaches from step 2a and step 2b.
+  ]{
+    Top row: Absolute future \acf{GPP} at the end of the \nth{21} century
+    (\range{2091}{2100}) calculated using step 2a (a) and step 2b (b). Both
+    approaches give similar results with global averages of $169
     \unit{\GtCyr{}}$ (a) and $175 \unit{\GtCyr{}}$ (b), which are both
@@
     approaches is $R^2 = 0.97$. Bottom row: Absolute (c) and relative (d)
-    differences between panels (a) and (b). \AdaptedFrom{Schlund2020}.}
+    differences between panels (a) and (b). \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:06:comparison_step2a_step2b}

### 2021-03-04 end_appendix/appendix.tex
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables

@@
   \end{tabulary}
-  \caption{Overview of the variables used in \cref{ch:05:paper_ecs}
+  \caption[
+    Overview of the variables used in \cref{ch:05:paper_ecs}.
+  ]{
+    Overview of the variables used in \cref{ch:05:paper_ecs}
     (\nameref*{ch:05:paper_ecs}). More details are given in
-    \cref{tab:05:overview_emergent_constraints}, which lists the variables
-    used for each emergent constraint.}
+    \cref{tab:05:overview_emergent_constraints}, which lists the variables used
+    for each emergent constraint.
+  }
   \label{tab:app:a:overview_variables}
@@
   \end{tabulary}
-  \caption{References for all observational datasets used in
-    \cref{ch:05:paper_ecs} (\nameref*{ch:05:paper_ecs}). More details are
-    given in \cref{tab:05:overview_emergent_constraints}, which lists the
-    observational datasets used for each emergent constraint.}
+  \caption[
+    References for all observational datasets used in \cref{ch:05:paper_ecs}.
+  ]{
+    References for all observational datasets used in \cref{ch:05:paper_ecs}
+    (\nameref*{ch:05:paper_ecs}). More details are given in
+    \cref{tab:05:overview_emergent_constraints}, which lists the observational
+    datasets used for each emergent constraint.
+  }
   \label{tab:app:a:observations}
@@
   \end{tabulary}
-  \caption{List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
+  \caption[
+    List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}.
+  ]{
+    List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
     (\nameref*{ch:05:paper_ecs}) alongside the main reference and the index
-    used in the corresponding figures. \AdaptedFrom{Schlund2020a}.}
+    used in the corresponding figures. \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:app:a:cmip5_models}
@@
   \end{tabulary}
-  \caption{As in \cref{tab:app:a:cmip5_models} but for the \acs{CMIP}6 models.
-    \AdaptedFrom{Schlund2020a}.}
+  \caption[
+    As in \cref{tab:app:a:cmip5_models} but for the \acs{CMIP}6 models.
+  ]{
+    As in \cref{tab:app:a:cmip5_models} but for the \acs{CMIP}6 models.
+    \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:app:a:cmip6_models}
@@
     $\SHS$}
-  \caption{All \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
+  \caption[
+    All \acs{CMIP}5 models used in \cref{ch:05:paper_ecs} including their
+    \acf{ECS} and \xaxis{} values for the emergent constraints BRI, COX, LIP,
+    SHD, SHL and SHS.
+  ]{
+    All \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
     (\nameref*{ch:05:paper_ecs}) including their \acf{ECS} and \xaxis{} values
@@
     corresponds to the index used in the associated figures.
-    \AdaptedFrom{Schlund2020a}.}
+    \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:app:a:cmip5_emergent_constraints_part1}
@@
     \dataset & $\idx$ & $\ECS$ & $\SU$ & $\TIH$ & $\TII$ & $\VOL$ & $\ZHA$}
-  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
+  \caption[
+    As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
     emergent constraints SU, TIH, TII, VOL and ZHA.
-    \AdaptedFrom{Schlund2020a}.}
+  ]{
+    As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
+    emergent constraints SU, TIH, TII, VOL and ZHA. \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:app:a:cmip5_emergent_constraints_part2}
@@
     $\SHS$}
-  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
-    \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.}
+  \caption[
+    As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
+    \acs{CMIP}6 models.
+  ]{
+    As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
+    \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:app:a:cmip6_emergent_constraints_part1}
@@
     \dataset & $\idx$ & $\ECS$ & $\SU$ & $\TIH$ & $\TII$ & $\VOL$ & $\ZHA$}
-  \caption{As in \cref{tab:app:a:cmip5_emergent_constraints_part2} but for the
-  \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.}
+  \caption[
+    As in \cref{tab:app:a:cmip5_emergent_constraints_part2} but for the
+    \acs{CMIP}6 models.
+  ]{
+    As in \cref{tab:app:a:cmip5_emergent_constraints_part2} but for the
+    \acs{CMIP}6 models. \AdaptedFrom{Schlund2020a}.
+  }
   \label{tab:app:a:cmip6_emergent_constraints_part2}
@@
   \end{tabulary}
-  \caption{List of the seven \acs{CMIP}5 models used in \cref{ch:06:paper_gpp}
-    (\nameref*{ch:06:paper_gpp}) alongside the main reference. More details are
-    given by \textcite{Anav2013}. We chose all \acs{CMIP}5 models which provide
-    all necessary variables (\emph{co2}, \emph{gpp}, \emph{lai}, \emph{pr},
-    \emph{rsds} and \emph{tas}) for all used experiments (\emph{esmHistorical},
-    \emph{esmrcp85} and \emph{esmFixClim1}). For all models, we only use the
-    first ensemble member available. \AdaptedFrom{Schlund2020}.}
+  \caption[
+    List of the seven \acs{CMIP}5 models used in \cref{ch:06:paper_gpp}
+    alongside the main references.
+  ]{
+    List of the seven \acs{CMIP}5 models used in \cref{ch:06:paper_gpp}
+    (\nameref*{ch:06:paper_gpp}) alongside the main references. More details
+    are given by \textcite{Anav2013}. We chose all \acs{CMIP}5 models which
+    provide all necessary variables (\emph{co2}, \emph{gpp}, \emph{lai},
+    \emph{pr}, \emph{rsds} and \emph{tas}) for all used experiments
+    (\emph{esmHistorical}, \emph{esmrcp85} and \emph{esmFixClim1}). For all
+    models, we only use the first ensemble member available.
+    \AdaptedFrom{Schlund2020}.
+  }
   \label{tab:app:b:cmip5_models}
@@
   \end{subfigure}
-  \caption{(a) Monthly mean atmospheric \acs{CO2} concentration at \acf{KUM}
-    from 1979 to 2019. The thin colored lines show the individual \acs{CMIP}5
-    models (emission-driven historical simulations for the years
-    \range{1979}{2005} and emission-driven \acs{RCP}8.5 simulations for the
-    years \range{2006}{2019}; the latter is not available for HadGEM2-ES). The
-    thick black line shows the observations. For the \acs{CMIP}5 models, the
-    grid cell closest to \acs{KUM} is considered. The curves show an increase
-    of the atmospheric \acs{CO2} concentration superimposed by a pronounced
-    seasonal cycle (see \cref{subsec:02:carbon_cycle_perturbations}). (b)
-    Annual amplitude of the seasonal cycle of \acs{CO2} (defined as the
-    difference between the maximum and the minimum monthly mean atmospheric
-    \acs{CO2} concentration for each year) against the annual mean atmospheric
-    \acs{CO2} concentration at \acs{KUM}. Colored points show the \acs{CMIP}5
-    models (similar time ranges as in (a)), and thick black points show the
+  \caption[
+    Monthly mean atmospheric \acs{CO2} concentration at \acl{KUM} from 1979 to
+    2019 and the corresponding annual amplitudes of the seasonal cycle of
+    \acs{CO2}.
+  ]{
+    (a) Monthly mean atmospheric \acs{CO2} concentration at \acf{KUM} from 1979
+    to 2019. The thin colored lines show the individual \acs{CMIP}5 models
+    (emission-driven historical simulations for the years \range{1979}{2005}
+    and emission-driven \acs{RCP}8.5 simulations for the years
+    \range{2006}{2019}; the latter is not available for HadGEM2-ES). The thick
+    black line shows the observations. For the \acs{CMIP}5 models, the grid
+    cell closest to \acs{KUM} is considered. The curves show an increase of the
+    atmospheric \acs{CO2} concentration superimposed by a pronounced seasonal
+    cycle (see \cref{subsec:02:carbon_cycle_perturbations}). (b) Annual
+    amplitude of the seasonal cycle of \acs{CO2} (defined as the difference
+    between the maximum and the minimum monthly mean atmospheric \acs{CO2}
+    concentration for each year) against the annual mean atmospheric \acs{CO2}
+    concentration at \acs{KUM}. Colored points show the \acs{CMIP}5 models
+    (similar time ranges as in (a)), and thick black points show the
     observations. The lines show the corresponding linear regression fits for
-    each dataset. The slopes of these linear fits define the sensitivity of
-    the seasonal \acs{CO2} cycle amplitude to atmospheric \acs{CO2}
-    concentrations, which is used as predictor for the emergent constraint
-    step of our approach (step 1). \AdaptedFrom{Schlund2020}.}
+    each dataset. The slopes of these linear fits define the sensitivity of the
+    seasonal \acs{CO2} cycle amplitude to atmospheric \acs{CO2} concentrations,
+    which is used as predictor for the emergent constraint step of our approach
+    (step 1). \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:app:b:co2}
@@
   \end{subfigure}
-  \caption{Distribution of the residuals of the \acf{GBRT} model for the two
-    different target variables used in step 2a (absolute \acf{GPP} at the end
-    of the \nth{21} century) and step 2b (fractional \acs{GPP} change over the
+  \caption[
+    Distribution of the residuals of the \acf{GBRT} model for the two different
+    target variables used in step 2a and step 2b.
+  ]{
+    Distribution of the residuals of the \acf{GBRT} model for the two different
+    target variables used in step 2a (absolute \acf{GPP} at the end of the
+    \nth{21} century) and step 2b (fractional \acs{GPP} change over the
     \nth{21} century). The distributions are derived by kernel density
@@
     datasets, which are very similar to each other. This indicates that the
-    \acl{ML} model does not overfit the data. \AdaptedFrom{Schlund2020}.}
+    \acl{ML} model does not overfit the data. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:app:b:residuals}
@@
   \end{subfigure}
-  \caption{Geographical distributions of the historical \acf{GPP} averaged
-    between 1991 and 2000. (a) C\acs{MIP}5 \acf{MMM}. (b) FLUXNET-MTE product
-    \autocite{Jung2011}. \AdaptedFrom{Schlund2020}.}
+  \caption[
+    Geographical distributions of the historical \acf{GPP} averaged between
+    1991 and 2000.
+  ]{
+    Geographical distributions of the historical \acf{GPP} averaged between
+    1991 and 2000. (a) C\acs{MIP}5 \acf{MMM}. (b) FLUXNET-MTE product
+    \autocite{Jung2011}. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:app:b:cmip5_hist_mte}
@@
   \end{subfigure}
-  \caption{Geographical distributions of the absolute \acf{GPP} at the end of
-    the \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for different
+  \caption[
+    Geographical distributions of the absolute \acf{GPP} at the end of the
+    \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for different
+    statistical models.
+  ]{
+    Geographical distributions of the absolute \acf{GPP} at the end of the
+    \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for different
     statistical models. (a) \acs{CMIP}5 \acf{MMM}. (b) Rescaled \acs{CMIP}5
@@
     all predictors. (f) \acs{GBRT} model using all predictors.
-    \AdaptedFrom{Schlund2020}.}
+    \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:app:b:step2a_results}
@@
   \end{subfigure}
-  \caption{Geographical distributions of the \acfp{SPE} of the absolute
-    \acf{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario
-    (step 2a) for different statistical models. Details on the calculation of
-    the \acs{SPE} are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5
-    \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c)
-    \Acf{LASSO} model using only the historical \acs{GPP} as single predictor.
-    (d) \Acf{GBRT} model using only the historical \acs{GPP} as single
-    predictor. (e) \acs{LASSO} model using all predictors. (f) \acs{GBRT}
-    model using all predictors. The \acs{SPE} is minimal for the \acs{GBRT}
-    model using all predictors. \AdaptedFrom{Schlund2020}.}
+  \caption[
+    Geographical distributions of the \aclp{SPE} of the absolute \acf{GPP} at
+    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for
+    different statistical models.
+  ]{
+    Geographical distributions of the \acfp{SPE} of the absolute \acf{GPP} at
+    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for
+    different statistical models. Details on the calculation of the \acs{SPE}
+    are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5 \acf{MMM}.
+    (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO}
+    model using only the historical \acs{GPP} as single predictor. (d)
+    \Acf{GBRT} model using only the historical \acs{GPP} as single predictor.
+    (e) \acs{LASSO} model using all predictors. (f) \acs{GBRT} model using all
+    predictors. The \acs{SPE} is minimal for the \acs{GBRT} model using all
+    predictors. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:app:b:step2a_results_errors}
@@
   \end{subfigure}
-  \caption{Geographical distributions of the \acfp{SPE} of the fractional
-    change in \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario
-    (step 2b) for different statistical models. Details on the calculation of
-    the \acs{SPE} are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5
+  \caption[
+    Geographical distributions of the \aclp{SPE} of the fractional change in
+    \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario (step 2b)
+    for different statistical models.
+  ]{
+    Geographical distributions of the \acfp{SPE} of the fractional change in
+    \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario (step 2b)
+    for different statistical models. Details on the calculation of the
+    \acs{SPE} are given in \cref{subsec:app:b:uncertainty}. (a) \acs{CMIP}5
     \acf{MMM}. (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c)
     \Acf{LASSO} model. (d) \Acf{GBRT} model. The \acs{SPE} is minimal for the
-    \acs{GBRT} model. \AdaptedFrom{Schlund2020}.}
+    \acs{GBRT} model. \AdaptedFrom{Schlund2020}.
+  }
   \label{fig:app:b:step2b_results_errors}

### 2021-03-12 02_abstract/abstract.tex
Restructured thesis and removed unnecessary parts
Restructured thesis and added more PDF bookmarks
Optimized chapter 1, 2 and 7

--- 2021-03-03 02_abstract/abstract_english.tex
@@
 %=============================================================================%
-%                                Dissertation                                 %
+%                            Doctoral Dissertation                            %
 %                               Manuel Schlund                                %
-%                                  (c) 2020                                   %
+%                                  (c) 2021                                   %
 %=============================================================================%
-%                             Abstract (English)                              %
+%                                  Abstract                                   %
 %=============================================================================%
 
-\chapter*{Abstract (English version)}
+\chapter{\abstractname}
 
@@
 uncertainties in the climate models, projections of the future climate are
-associated with large uncertainties. In this thesis, we analyze future climate
-projections from \acsp{ESM} participating in the \acfAbstract{CMIP} and reduce
-corresponding uncertainties with observations.
+associated with large uncertainties. The main focus of this thesis is the
+analysis of future climate projections from \acsp{ESM} participating in the
+\acfAbstract{CMIP} and the reduction of corresponding uncertainties with
+observations.
 
@@
 climate system to an external forcing) in the latest generation of \acsp{ESM}
-from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, the equilibrium
-temperature response that follows a doubling of the atmospheric
-\acfAbstract{CO2} concentration, we find a \acfAbstract{MMM} and multi-model
-range of $3.74 \unit{K}$ and $\rangeunit{1.8}{5.6}{K}$, respectively. These
-values are higher than in any previous \acs{CMIP} ensemble before. Moreover, a
-third of the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely
-\acs{ECS} range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR}
-(\acs{AR}5) of the \acl{IPCC} from 2013. Similarly, the transient response of
-the climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
-shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is
-again higher than the likely range assessed in \acs{AR}5 of
+from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, which is defined as
+the equilibrium temperature response that follows a doubling of the atmospheric
+\acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74 \unit{K}$ and a
+multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These values are
+higher than in any previous \acs{CMIP} ensemble before. Moreover, a third of
+the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely \acs{ECS}
+range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR} (\acs{AR}5)
+of the \acl{IPCC} from 2013. Similarly, the transient response of the climate
+system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR}, shows an
+inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is again
+higher than the likely range assessed in \acs{AR}5 of
 $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
@@
 
-To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models, we
-systematically analyze eleven published emergent constraints on \acs{ECS} that
-have mostly been derived from models participating in \acs{CMIP}5, the
-predecessor generation of \acs{CMIP}6. Emergent constraints are potentially
-promising approaches to reduce uncertainties in climate model projections by
-combining observations and output from \acsp{ESM}. The focus of this study is
-on testing if these emergent constraints hold for \acsp{ESM} participating
+To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models,
+eleven published emergent constraints on \acs{ECS} (mostly derived from models
+participating in \acs{CMIP}5, the predecessor generation of \acs{CMIP}6) are
+systematically analyzed. Emergent constraints are potentially promising
+approaches to reduce uncertainties in climate model projections by combining
+observations and output from \acsp{ESM}. The focus of this analysis is on
+testing if these emergent constraints hold for \acsp{ESM} participating
 \acs{CMIP}6. Since none of the emergent constraints considered here have been
@@
 new feedback processes rather than changes in strength of those previously
-dominant. Our results support previous studies concluding that emergent
+dominant. The results support previous studies concluding that emergent
 constraints should be based on an independently verifiable physical mechanism
@@
 
-To overcome these issues of single-process-oriented emergent constraints, we
-introduce an alternative approach based on \acfAbstract{ML}. Since this new
+To overcome these issues of single-process-oriented emergent constraints, an
+alternative approach based on \acfAbstract{ML} is introduced. Since this new
 technique relies on a large number of data points in order to train the
 \acs{ML} algorithm, the scalar climate sensitivity expressed as \acs{ECS} or
-\acs{TCR} is not an appropriate target variable. As an alternative, we study
-projections of the future \acfAbstract{GPP}, which is the largest flux of the
-terrestrial carbon uptake and slows down global warming by removing \acs{CO2}
-from the atmosphere. In this analysis, we combine an existing emergent
-constraint on \acs{CO2} fertilization with a \acs{ML} approach to constrain the
-spatial variations of multi-model \acs{GPP} projections. In the first step of
-our two-step approach, we use observed changes in the \acs{CO2} seasonal cycle
-at \acl{KUM} to constrain the global mean \acs{GPP} at the end of the \nth{21}
-century (\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM}
-participating in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
-the unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second
-step, we use a \acs{ML} model to constrain gridded future absolute \acs{GPP}
-and gridded fractional \acs{GPP} change in two independent approaches. For
-this, observational data is fed into the \acs{ML} algorithm that has been
-trained on \acs{CMIP}5 data to learn relationships between present-day
-physically-relevant diagnostics and the target variable. In a
-leave-one-model-out \acl{CV} approach, the \acs{ML} model shows superior
-performance to the \acs{CMIP}5 \acs{MMM}. Our approach predicts an increased
-\acs{GPP} change in northern high latitudes compared to regions closer to the
-equator.
+\acs{TCR} is not an appropriate target variable. Therefore, \acfAbstract{GPP}
+as a process that contributes to climate sensitivity is studied as an
+alternative. \acs{GPP} is the largest flux of the terrestrial carbon uptake and
+slows down global warming by removing \acs{CO2} from the atmosphere. In this
+analysis, an existing emergent constraint on \acs{CO2} fertilization is
+combined with a \acs{ML} approach to constrain the spatial variations of
+multi-model \acs{GPP} projections. In the first step of the two-step approach,
+observed changes in the \acs{CO2} seasonal cycle at \acl{KUM} are used to
+constrain the global mean \acs{GPP} at the end of the \nth{21} century
+(\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM} participating
+in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the
+unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second step, a
+\acs{ML} model is used to constrain gridded future absolute \acs{GPP} and
+gridded fractional \acs{GPP} change in two independent approaches. For this,
+observational data is fed into the \acs{ML} algorithm that has been trained on
+\acs{CMIP}5 data to learn relationships between present-day physically-relevant
+diagnostics and the target variable. In a leave-one-model-out \acl{CV}
+approach, the \acs{ML} model shows superior performance to the \acs{CMIP}5
+\acs{MMM}. The new approach predicts a higher \acs{GPP} increase in high
+latitudes and a lower \acs{GPP} increase in regions closer to the equator.

### 2021-03-12 ch01_introduction/introduction.tex
Restructured thesis and added more PDF bookmarks
Optimized chapter 1, 2 and 7

@@
   development} and amplifies \enquote{risks for livelihoods and for food and
-  human security} \autocite{IPCC2014}. Potential drivers for climate change
-are all natural and anthropogenic substances and processes that may alter the
-Earth's energy budget. The human influence on the climate system is clear: Ever
-increasing emissions of \acp{GHG} since the end of the pre-industrial era
-largely driven by economic and population growth led to atmospheric
+human security} \autocite{IPCC2014}. Potential drivers for climate change are
+all natural and anthropogenic substances and processes that may alter the
+Earth's energy budget. The human influence on the climate system is clear:
+Ever increasing emissions of \acp{GHG} since the end of the pre-industrial
+era largely driven by economic and population growth led to atmospheric
 concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O} that are
-\enquote{unprecedented in at least the last 800000 years} \autocite{IPCC2014}.
-The effects of these \ac{GHG} emissions and other anthropogenic drivers have
-been \enquote{detected throughout the climate system and are extremely likely
-  to have been the dominant cause of the observed warming since the
-  mid-\nth{20} century} \autocite{IPCC2014}. The greenhouse effect is based on
-the optical properties of the \acp{GHG}: While the atmosphere is mostly
-transparent for the incoming solar (shortwave) radiation, the outgoing
-infrared (longwave) radiation that is reflected from the planet's surface is
-partly absorbed by the corresponding \ac{GHG} molecules through an excitation
-of their corresponding vibrational modes and re-emitted isotropically. This
-process traps energy near the surface and leads to a warming of the Earth's
-surface and the lower atmosphere. The individual impact of the different
-drivers of climate change measured with the so-called \emph{radiative
-forcing}, which quantifies the change in energy fluxes caused by changes in
-these drivers relative to pre-industrial conditions \autocite{IPCC2013}. The
-total radiative forcing is positive (which corresponds to a warming) and its
-largest contribution is caused by the increase in the atmospheric
-concentration of \ac{CO2} due to fossil fuel emissions since the year 1750
-\autocite{IPCC2014}. Apart from their physical warming effect on the climate,
-carbon-based \acp{GHG} like \ac{CO2} also directly influence the global
-carbon cycle, an important biogeochemical cycle of the Earth. A crucial flux
-of the carbon cycle is the \emph{\ac{GPP}}, which describes the carbon uptake
-of the terrestrial biosphere due to photosynthesis. Since this land carbon
-uptake absorbs about $30 \unit{\%}$ of the anthropogenic \ac{CO2} emissions in
-today's climate \autocite{Friedlingstein2020}, this process substantially slows
-down global warming. Other important anthropogenic drivers of climate change
-are the emission of aerosols and land use/land cover changes. Apart from
-anthropogenic drivers, there are also natural processes which impact the
-climate system like changes in the solar activity or volcanic eruptions.
-However, there is clear evidence that these natural drives alone cannot explain
-the observed climate change \autocite{Haustein2017}.
+\enquote{unprecedented in at least the last 800000 years}
+\autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other
+anthropogenic drivers have been \enquote{detected throughout the climate
+system and are extremely likely to have been the dominant cause of the
+observed warming since the mid-\nth{20} century} \autocite{IPCC2014}. The
+greenhouse effect is based on the optical properties of the \acp{GHG}: While
+the atmosphere is mostly transparent for the incoming solar (shortwave)
+radiation, the outgoing infrared (longwave) radiation that is reflected from
+the planet's surface is partly absorbed by the corresponding \ac{GHG}
+molecules through an excitation of their corresponding vibrational modes and
+re-emitted isotropically. This process traps energy near the surface and
+leads to a warming of the Earth's surface and the lower atmosphere. The
+individual impact of the different drivers of climate change measured with
+the so-called \emph{radiative forcing}, which quantifies the change in energy
+fluxes caused by changes in these drivers relative to pre-industrial
+conditions \autocite{IPCC2013}. The total radiative forcing is positive
+(which corresponds to a warming) and its largest contribution is caused by
+the increase in the atmospheric concentration of \ac{CO2} due to fossil fuel
+emissions since the year 1750 \autocite{IPCC2014}. Apart from their physical
+warming effect on the climate, carbon-based \acp{GHG} like \ac{CO2} also
+directly influence the global carbon cycle, an important biogeochemical cycle
+of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}}, which
+describes the carbon uptake of the terrestrial biosphere due to photosynthesis.
+Since this land carbon uptake absorbs about $30 \unit{\%}$ of the anthropogenic
+\ac{CO2} emissions in today's climate \autocite{Friedlingstein2020}, this
+process substantially slows down global warming and directly contributes to the
+magnitude of the climate sensitivity. Other important anthropogenic drivers of
+climate change are the emission of aerosols and land use/land cover changes.
+Apart from anthropogenic drivers, there are also natural processes which impact
+the climate system like changes in the solar activity or volcanic eruptions.
+However, there is clear evidence that these natural drivers alone cannot
+explain the observed climate change \autocite{Haustein2017}.
 
@@
 the \ac{IPCC}, and their predecessor models from \acs{CMIP}5
-\autocite{Taylor2012} have been successfully evaluated in the Fifth \acl{AR}
+\autocite{Taylor2012} have been successfully analyzed in the Fifth \acl{AR}
 (\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
@@
 response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
-have been assessed with high ranges of $\rangeunit{1.5}{4.5}{K}$ and
+have been assessed with large ranges of $\rangeunit{1.5}{4.5}{K}$ and
 $\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
@@
 from the author of this thesis. A detailed list of contributions to these
-studies (including their full reference) is given in the corresponding
-\namecrefs{ch:03:esmvaltool}.
+studies is given in the corresponding \namecrefs{ch:03:esmvaltool}.
 

### 2021-03-12 ch02_scientific_background/scientific_background.tex
Optimized chapter 1, 2 and 7

@@
 
-In contrast to other fields of science, researching the future evolution of the
-Earth's climate cannot be purely done by performing experiments in a
+In contrast to many other fields of science, researching the future evolution
+of the Earth's climate cannot be purely done by performing experiments in a
 laboratory. Due to the immense complexity of the Earth system (including
@@
 \acp{AOGCM} \autocite{Flato2013} is the numerical solving of the differential
-equations describing the exchange of energy and matter between these physical
-components.
+equations describing the exchange of energy, momentum and matter between and
+within these physical components.
 
@@
 
-\subsection{The \acf{CMIP}}
+\subsection{The \acfAbstract{CMIP}}
 \label{subsec:02:cmip}
@@
 consist of hundreds of thousands of lines of computer code. Thus, a
-standardization to a certain degree is crucial for the various research groups
-developing \acp{ESM} all around the world in order to obtain comparable output
-and to facilitate model analysis. For this reason, the \ac{WGCM} of the
-\ac{WCRP} initiated the \acf{CMIP} in 1995, with the objective to
-\enquote{better understand past, present and future climate changes arising
-  from natural, unforced variability or in response to changes in radiative
-  forcing in a multi-model context} \autocite{WCRP2020}. One major element of
-\ac{CMIP} is to establish common standards, coordination, infrastructure and
-documentation in order to facilitate the distribution of climate model output
-\autocite{Eyring2016, Juckes2020}.
+standardization of the experimental setup and model output to a certain degree
+is crucial for the various research groups developing \acp{ESM} all around the
+world in order to obtain comparable output and to facilitate model analysis.
+For this reason, the \ac{WGCM} of the \ac{WCRP} initiated the \acf{CMIP} in
+1995, with the objective to \enquote{better understand past, present and future
+  climate changes arising from natural, unforced variability or in response to
+  changes in radiative forcing in a multi-model context} \autocite{WCRP2020}.
+One major element of \ac{CMIP} is to establish common standards, coordination,
+infrastructure and documentation in order to facilitate the distribution of
+climate model output \autocite{Eyring2016, Juckes2020}.
 
@@
 other scientific experiment, climate model projections suffer from associated
-uncertainties. There are three major sources of climate model projections we
-can distinguish: natural variability, climate response uncertainty and emission
-uncertainty \autocite{Hawkins2009, Hawkins2010}.
+uncertainties. There are three major sources of uncertainties in climate model
+projections we can distinguish: natural variability, climate response
+uncertainty and emission uncertainty \autocite{Hawkins2009, Hawkins2010}.
 \Cref{fig:02:sources_of_uncertainty} shows these three sources for the
@@
 natural variability on longer time scales come from phenomena like the
-\ac{ENSO} or the \ac{NAO} and from externally (and thus explainable) events
-like volcanic eruptions and variations in the solar activity. Natural
-variability can be seen as the \enquote{noise} in the climate record as opposed
-to the anthropogenic \enquote{signal} \autocite{Cubasch2013}. As illustrated by
-\cref{fig:02:sources_of_uncertainty}, the uncertainty associated with natural
-variability is constant over time.
+\ac{ENSO} or the \ac{NAO} and from other events like volcanic eruptions and
+variations in the solar activity. Natural variability can be seen as the
+\enquote{noise} in the climate record as opposed to the anthropogenic
+\enquote{signal} \autocite{Cubasch2013}.
 
@@
 
-Climate sensitivity refers to the change in \ac{GSAT} that results from a
-change in the radiative forcing. In other words, it describes how sensitive the
-climate system is to an external forcing. The source of this forcing might
-either be natural (changes in the solar activity, volcanic eruptions, \etc{})
-or anthropogenic (emissions of \acp{GHG}, land use changes, \etc{}).
+An important policy-relevant metric of the Earth system that can be assessed
+with numerical climate model simulations is the climate sensitivity. Climate
+sensitivity refers to the change in \ac{GSAT} that results from a change in the
+radiative forcing. In other words, it describes how sensitive the climate
+system is to an external forcing. The source of this forcing might either be
+natural (changes in the solar activity, volcanic eruptions, \etc{}) or
+anthropogenic (emissions of \acp{GHG}, land use changes, \etc{}). Thus, the
+assessment of climate sensitivity is essential for a precise quantification of
+the human-made climate change in order to determine optimal mitigation and
+adaptation strategies.
 
@@
 amplified or diminished by climate feedbacks. Thus, feedback processes play a
-crucial role in the evaluation of climate sensitivity.
+crucial role determining the magnitude of the climate sensitivity.
 \Cref{fig:02:feedbacks_overview} shows an overview of important feedbacks in
@@
 negative, since a variety of different mechanisms is involved. An example for
-this is the cloud feedback. Changes in clouds induced by climate change can
-cause both longwave (greenhouse warming) and shortwave (reflective cooling)
-effects, which both need to be considered for the overall cloud feedback
-\autocite{Boucher2013}. Relevant cloud properties that may change as a response
-to an external forcing and that may alter the Earth's radiative budget are
-cloud cover, cloud optical thickness, cloud altitude and the geographical
-distribution of clouds. Examples for robust cloud feedback processes are the
-rise of high clouds in a warming climate which traps longwave radiation and
-enhances global warming and the reduction in mid- and low-level cloud cover
-which diminishes the reflection of incoming solar radiation and also increases
-the surface warming \autocite{Boucher2013}. In global climate model ensembles,
-the overall cloud feedback shows a large range with positive and negative
-values, but tends to be slightly positive on average \autocite{Soden2006,
-  Dufresne2008, Vial2013, Zelinka2020}. This large uncertainty in the cloud
-feedback is a major reason for uncertainties in the climate sensitivity of
-climate models \autocite{Boucher2013, Flato2013}.
+this is the cloud feedback. Changes in the clouds induced by climate change can
+cause changes in their longwave (greenhouse warming) and shortwave (reflective
+cooling) effects on the Earth's radiation budget, which both need to be
+considered for the overall cloud feedback \autocite{Boucher2013}. Relevant
+cloud properties that may change as a response to an external forcing and that
+may alter the Earth's radiative budget are cloud cover, cloud optical
+thickness, cloud top and cloud base height, vertical extent and the
+geographical distribution of clouds. Examples for robust cloud feedback
+processes are the increase in cloud top height of high-level clouds in a
+warming climate which traps longwave radiation and enhances global warming and
+the reduction in mid- and low-level cloud cover which diminishes the reflection
+of incoming solar radiation and also increases the surface warming
+\autocite{Boucher2013}. In global climate model ensembles, the overall cloud
+feedback shows a large range with positive and negative values, but tends to be
+slightly positive on average \autocite{Soden2006, Dufresne2008, Vial2013,
+  Zelinka2020}. This large uncertainty in the cloud feedback is a major reason
+for uncertainties in the climate sensitivity of climate models
+\autocite{Boucher2013, Flato2013}.
 
@@
 
-The foundation of a basic mathematical framework for climate feedback analysis
-is a simple energy balance model \autocite{Gregory2009, Roe2009}. Anthropogenic
-activities in the Earth system like the emissions of \acp{GHG} or aerosols
-introduce an external forcing to the climate system, which is quantified with a
-radiative forcing $F$ measured in $\si{W m^{-2}}$. To restore a stable state,
-the climate system opposes this forcing with a climate response $R$, leading to
-a net energy flux of
+For a precise quantification of climate feedbacks and thus climate sensitivity,
+a mathematical framework for climate feedback analysis necessary. One possible
+approach for this is based on a simple energy balance model
+\autocite{Gregory2009, Roe2009}. Anthropogenic activities in the Earth system
+like the emissions of \acp{GHG} or aerosols introduce an external forcing to
+the climate system, which is quantified with a radiative forcing $F$ measured
+in $\si{W m^{-2}}$. To restore a stable state, the climate system opposes this
+forcing with a climate response $R$, leading to a net energy flux of
 \begin{equation}
@@
 oldest metrics describing climate change. However, in practice this traditional
-definition of is not useful. Running fully-coupled \acp{ESM} into equilibrium
-is computationally expensive as it would require thousands of model years
-\autocite{Rugenstein2020}.
+definition of is not always useful. Running fully-coupled \acp{ESM} into
+equilibrium is computationally expensive as it would require thousands of model
+years \autocite{Rugenstein2020}.
 
@@
 accurate as the deeper ocean warms up and cannot be applied to simulations with
-a steady state climate change ($N \to 0$) where the external forcing tends
-stabilize \autocite{Gregory2009}. By applying the definition of \ac{TCR}
-(transient \ac{GSAT} change $\Delta T$ at the time of \ac{CO2} doubling) and
-combining \cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be
-estimated as
+a steady state climate change in which $N \to 0$ \autocite{Gregory2009}. By
+applying the definition of \ac{TCR} (transient \ac{GSAT} change $\Delta T$ at
+the time of \ac{CO2} doubling) and combining
+\cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be estimated as
 \begin{equation}
@@
 
-Since one study presented in this thesis aims to reduce uncertainties in carbon
-cycle--related processes \commentcite{Schlund2020}{see \cref{ch:06:paper_gpp}},
-this \namecref{sec:02:carbon_cycle} introduces the scientific background of the
-global carbon cycle.
+Currently, only about $50 \unit{\%}$ of the emitted \ac{CO2} by humankind
+remains in the atmosphere \autocite{Friedlingstein2020}. The residual part is
+absorbed by the two other main carbon sinks of the planet, the terrestrial
+biosphere and the ocean. These exchange fluxes of carbon are part of the global
+carbon cycle of planet Earth. Since only atmospheric \ac{CO2} can act as
+\ac{GHG}, this removal of \ac{CO2} from the atmosphere substantially slows down
+global warming. Moreover, since the emission of carbon-based \acp{GHG}
+introduces additional carbon into the Earth system, the carbon exchange fluxes
+might change under global warming, which in turn directly influences the
+magnitude of the climate sensitivity. However, the two idealized climate
+sensitivity metrics \ac{ECS} and \ac{TCR} introduced in the previous
+\namecref{sec:02:climate_sensitivity} are by definition independent from carbon
+cycle--related effects, since both of them are defined in terms of a doubling
+of the atmospheric \ac{CO2} concentration, not in terms of \ac{CO2} emissions.
+Nevertheless, in order to assess policy-relevant metrics like the allowable
+fossil fuel emissions to meet particular warming targets, for example the $1.5
+\unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015}, it is
+crucial to take the global carbon cycle into account. This
+\namecref{sec:02:carbon_cycle} introduces the scientific background of the
+global carbon cycle and its current anthropogenic perturbations.
 

### 2021-03-12 ch03_esmvaltool/esmvaltool.tex
Optimized chapter 1, 2 and 7

@@
 
-\section{The \acf{ESMValTool}}
+\section{The \acfAbstract{ESMValTool}}
 \label{sec:03:esmvaltool}
@@
     \autocite{Cox2018}. Magenta colors indicate high-sensitivity climate models
-    ($\lambda < 1 \unit{W m^{-2} K^{-1}}$), and green colors indicate low
-    sensitivity climate models ($\lambda > 1 \unit{W m^{-2} K^{-1}}$). Both
-    panels have been created with the \acs{ESMValTool} using
+    ($\abs*{\lambda} < 1 \unit{W m^{-2} K^{-1}}$), and green colors indicate
+    low-sensitivity climate models ($\abs*{\lambda} > 1 \unit{W m^{-2}
+      K^{-1}}$). Both panels have been created with the \acs{ESMValTool} using
     \emph{recipe\_cox18nature.yml} and are similar to \textcite{Cox2018}. (a)
@@
 models and observations form HadCRUT4 \autocite{Morice2012}. Since
-high-sensitivity climate models exhibit higher values of $\Psi$ than low
-sensitivity models, an emergent relationship between \ac{ECS} and $\Psi$ can be
-established (\cref{fig:03:cox18:b}). With observations of $\Psi$, \ac{ECS} can
-eventually be constrained to $\left( 2.8 \pm 0.6 \right) \unit{K}$ ($66
-\unit{\%}$ confidence range). More details on this including the calculation of
-$\Psi$ are given in \cref{subsec:05:cox}.
+high-sensitivity climate models exhibit higher values of $\Psi$ than
+low-sensitivity models, an emergent relationship between \ac{ECS} and $\Psi$
+can be established (\cref{fig:03:cox18:b}). With observations of $\Psi$,
+\ac{ECS} can eventually be constrained to $\left( 2.8 \pm 0.6 \right) \unit{K}$
+($66 \unit{\%}$ confidence range). More details on this including the
+calculation of $\Psi$ are given in \cref{subsec:05:cox}.
 

### 2021-03-12 ch06_paper_gpp/paper_gpp.tex
Optimized chapter 1, 2 and 7

@@
 technique relies on a large number of data points in order to train the \ac{ML}
-algorithm, \ac{ECS} as a single scalar metric for each climate model is not an
-appropriate target variable. As an alternative, we analyze projections of
-future \acf{GPP} in this study, which constitutes the largest flux of the
-terrestrial carbon uptake. As described in detail in
-\cref{subsec:02:carbon_cycle_perturbations}, the global carbon cycle actively
-slows down global warming by removing \ac{CO2} from the atmosphere. In
+algorithm, \ac{ECS} as a scalar climate sensitivity metric is not an
+appropriate target variable. Thus, this analysis does not focus on reducing
+uncertainties in climate sensitivity itself, but rather on a selected process
+that contributes to it: \ac{GPP}. As described in detail in
+\cref{subsec:02:carbon_cycle_overview,subsec:02:carbon_cycle_perturbations},
+\ac{GPP} constitutes the largest flux of the terrestrial carbon uptake and
+actively slows down global warming by removing \ac{CO2} from the atmosphere. In
 particular, the terrestrial biosphere absorbs about $30 \unit{\%}$ of the total
@@
     the different statistical models are very similar and show a higher
-    \acs{GPP} increase in high latitudes and a lower \acs{GPP} closer to the
-    equator. \AdaptedFrom{Schlund2020}.
+    \acs{GPP} increase in high latitudes and a lower \acs{GPP} increase in
+    regions closer to the equator. \AdaptedFrom{Schlund2020}.
   }
@@
 approaches give consistent results (see
-\cref{fig:06:comparison_step2a_step2b}). The latter quantity shows an increased
-\ac{GPP} change in the high latitudes compared to regions closer to the
+\cref{fig:06:comparison_step2a_step2b}). The latter quantity shows a higher
+\ac{GPP} increase in high latitudes and a lower \ac{GPP} increase closer to the
 equator, which might be attributed to an additional greening trend in the high

### 2021-03-12 ch07_conclusion/conclusion.tex
Restructured thesis and removed unnecessary parts
Optimized chapter 1, 2 and 7

@@
 
-In a first analysis, we evaluate the climate sensitivity metrics \ac{ECS} and
-\ac{TCR} for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
+In a first analysis, the climate sensitivity metrics \ac{ECS} and \ac{TCR} are
+evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
 which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
@@
 \cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
-\ac{ECS}, we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which
-is higher on the upper and lower end than any model range from previous
+\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
+which is higher on the upper and lower end than any model range from previous
 \ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
@@
 representation of mixed-phase clouds above the Southern Ocean. For one thing,
-this change leads to an improved shortwave \ac{CRE} in these models when
-compared to observations. However, this change also substantially reduces the
-strong negative shortwave cloud feedback over the Southern Ocean that is
-present in previous \ac{CMIP} generations and that results from a cloud phase
-change from ice clouds in the present-day to liquid clouds in the future. In
-the affected \acs{CMIP}6 models, this cloud phase change is not possible
+this change leads to an improved simulation of the shortwave \ac{CRE} in these
+models when compared to observations. However, this change also substantially
+reduces the strong negative shortwave cloud feedback over the Southern Ocean
+that is present in previous \ac{CMIP} generations and that results from a cloud
+phase change from ice clouds in the present-day to liquid clouds in the future.
+In the affected \acs{CMIP}6 models, this cloud phase change is not possible
 anymore since the models predominantly simulate liquid clouds over the
@@
 In order to reduce this large range of \ac{ECS} in the latest generation of
-climate  models, we evaluate suitable emergent constraints on the \acs{CMIP}6
-ensemble. Emergent constraints use a physically-based inter-model relationship
+climate  models, suitable emergent constraints on the \acs{CMIP}6 ensemble are
+evaluated. Emergent constraints use a physically-based inter-model relationship
 between an observable quantity of the Earth system and a target variable to
 reduce uncertainties in the target variable with observations
-\autocite{Allen2002}. In total we assess eleven emergent constraints on
-\ac{ECS}, which are mostly related to cloud feedbacks since these constitute
+\autocite{Allen2002}. In total eleven emergent constraints on \ac{ECS} are
+assessed, which are mostly related to cloud feedbacks since these constitute
 the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
   Flato2013}. Since all of the evaluated emergent constraints have been derived
-on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, we can perform out-of-sample tests
-on the emergent constraints by assessing whether they still hold for the
+on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, out-of-sample tests on the emergent
+constraints can be performed by assessing whether they still hold for the
 \acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
-already published in \textcite{Schlund2020a}, we find a substantial reduction
-of skill for the majority of emergent constraints when applied to the
+already published in \textcite{Schlund2020a}, a substantial reduction of skill
+for the majority of emergent constraints is found when applied to the
 \acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
@@
 algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
-not an appropriate target variable. As an alternative, we study projections of
-the future \ac{GPP}, which is the largest flux of the terrestrial carbon uptake
-and slows down global warming by removing \ac{CO2} from the atmosphere. In the
-first step of our new two-step approach, we constrain the global mean \ac{GPP}
-at the end of the \nth{21} century in the \acs{CMIP}5 \acs{RCP}8.5 scenario to
-$\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
-\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models'
-\ac{GPP} response to \ac{CO2}, which shows a range of
-$\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In the second
-step, a \ac{ML}-based climate model weighting approach is used to further
-constrain the gridded \ac{GPP} based on present-day predictors that are
-relevant for the simulation of \ac{GPP} in the \acp{ESM}. Our \ac{ML} approach
+not an appropriate target variable. Therefore, this analysis does not focus on
+reducing uncertainties in climate sensitivity itself but rather on a selected
+process that contributes to it: \ac{GPP}. \ac{GPP} is the largest flux of the
+terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
+from the atmosphere. In the first step of the new two-step approach, the global
+mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
+\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
+a published emergent constraint by \textcite{Wenzel2016}. This first step
+corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2}, which shows a
+range of $\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In
+the second step, a \ac{ML}-based climate model weighting approach is used to
+further constrain the gridded \ac{GPP} based on present-day predictors that are
+relevant for the simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach
 is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
@@
 (gridded) target variables and non-linear relationships between the predictors
-and the target variable. We establish a relationship between process‐oriented
-predictors and future projections of \ac{GPP} and then utilize this to project
-today's observed conditions into the future. The prediction phase of our method
-can be interpreted as an implicit performance weighting. However, due to the
-complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
-to extract specific values for the individual weights. We consider two target
-variables: the gridded monthly climatologies of absolute \ac{GPP}
-(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
-\nth{21} century (2100 versus 2000). The latter quantity shows an increased
+and the target variable. A a relationship between process‐oriented predictors
+and future projections of \ac{GPP} is established and then utilized to project
+today's observed conditions into the future. The prediction phase of the new
+method can be interpreted as an implicit performance weighting. However, due to
+the complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not
+possible to extract specific values for the individual weights. Two target
+variables are considered: the gridded monthly climatologies of absolute
+\ac{GPP} (\range{2091}{2100}) and the gridded fractional \ac{GPP} change over
+the \nth{21} century (2100 versus 2000). The latter quantity shows an increased
 \acs{GPP} change in the high latitudes compared to regions closer to the
 equator. The results of both approaches are consistent with each other and with
-the global constraint of the first step. We validate our approach by comparing
-it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
+the global constraint of the first step. The new approach is validated by
+comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
 \ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
-(\ac{LASSO}), we find a reduction of the resulting mean \ac{RMSEP} of up to $48
-\unit{\%}$ ($3 \unit{\%}$) when using our \ac{ML} approach. Moreover, the
-evaluation of the global and local feature importance allows us to get further
+(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
+\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
+the evaluation of the global and local feature importance allows further
 insights into the \ac{ML} model. For the first target variable (absolute
 \ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
-can be explained with a correction of the historical bias in \ac{GPP} by our
-approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
+can be explained with a correction of the historical bias in \ac{GPP} by the
+new approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
 features. This study directly addresses \cref{enum:01:question_3}
-(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can confidently answer
-with \enquote{yes} based on the results we find.
+(\enquote{\emph{\KeyScienceQuestionThree{}}}), which can be confidently
+answered with \enquote{yes} based on the results found.
 
@@
 
-Due to the flexibility of our \ac{ML}-based climate model weighting approach
-presented in \cref{ch:06:paper_gpp}, one natural way forward is to apply this
-technique to other target variables. While the scalar metrics \ac{ECS} or
-\ac{TCR} with one value per climate model do not offer enough training points
-for this technique, a possible alternative could be the gridded \acl{T} over
-the \nth{21} century in the \ac{SSP} scenarios. A relevant study in this
-context is given by \textcite{Brunner2020}, who use performance and
+Due to the flexibility of the new \ac{ML}-based climate model weighting
+approach presented in \cref{ch:06:paper_gpp}, one natural way forward is to
+apply this technique to other target variables. While the scalar metrics
+\ac{ECS} or \ac{TCR} with one value per climate model do not offer enough
+training points for this technique, a possible alternative could be the gridded
+\acl{T} over the \nth{21} century in the \ac{SSP} scenarios. A relevant study
+in this context is given by \textcite{Brunner2020}, who use performance and
 interdependence-based climate model weighting to constrain the \acl{T} over the
 \nth{21} century based on observable process-based predictors of today's
-climate and which could serve as a possible baseline for our new \ac{ML}-based
-weighting approach. An exciting and valuable addition to our method could be
+climate and which could serve as a possible baseline for the new \ac{ML}-based
+weighting approach. An exciting and valuable addition to the method could be
 causal inference \autocite{Nowack2020, Runge2019}. In the current approach, the
@@
 predictor at a specific location for the prediction of the target variable by
-the \ac{ML} model. By integrating causal networks into our approach, it might
-be possible to explore and utilize true causal connections instead.
+the \ac{ML} model. By integrating causal networks into the new approach, it
+might be possible to explore and utilize true causal connections instead.
 
@@
 climate science involve for example \ac{ML}-based parametrizations that are
-learned from high-resolution climate models \autocite{Rasp2018}, the
-\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
-  Mansfield2020} or the learning of entire \acp{ESM} from observational
-products \autocite{Geer2021}. Therefore, the foundation for innovative and
+learned from high-resolution climate models \autocite{Rasp2018}, \ac{ML}-based
+analysis and prediction of forcing patterns \autocite{Barnes2019,
+  Mansfield2020} or learning of entire \acp{ESM} from observational products
+\autocite{Geer2021}. Therefore, the foundation for innovative and
 groundbreaking research in climate science for the near future has been laid,

### 2021-03-15 ch02_scientific_background/scientific_background.tex
Optimized chapters 3, 4 and 7

@@
   \label{eq:02:delta_t_eq}
-\end{equation}
-Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
-from the external forcing $F$ and the climate feedback parameter $\lambda$. The
-steady state values for $F$ and $\lambda$ can be estimated from a \nxcotwo{4}
-simulation that is not in equilibrium by extrapolation with a linear regression
-\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
-  regression}, the global and annual mean net \ac{TOA} radiation $N$ versus the
-change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
-\nxcotwo{4} run are plotted (see \cref{fig:02:gregory_regression}). To account
-for energy leakage and remove any model drift present in the control climate, a
-linear fit of the corresponding pre-industrial control run is subtracted from
-the \nxcotwo{4} simulation beforehand \autocite{Andrews2012}. Since the
-combination of \cref{eq:02:N,eq:02:linear_response} yields
+\end{equation} Thus, the change in \ac{GSAT} in radiative equilibrium can be
+easily calculated from the external forcing $F$ and the climate feedback
+parameter $\lambda$. The steady state values for $F$ and $\lambda$ can be
+estimated from a \nxcotwo{4} simulation that is not in equilibrium by
+extrapolation with a linear regression \autocite{Gregory2004}. For this
+so-called \emph{Gregory regression} (see \cref{fig:02:gregory_regression}), the
+global and annual mean net \ac{TOA} radiation $N$ versus the change in the
+annual mean \ac{GSAT} $\Delta T$ for all 150 years of the \nxcotwo{4} run are
+plotted. To account for energy leakage and remove any model drift present in
+the control climate, a linear fit of the corresponding pre-industrial control
+run is subtracted from the \nxcotwo{4} simulation beforehand
+\autocite{Andrews2012}. Since the combination of
+\cref{eq:02:N,eq:02:linear_response} yields
 \begin{equation}
@@
 whereas here a simulation with an abrupt quadrupling is considered.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\FigureWidth{}]{
+    ch02_scientific_background/figs/gregory_regression_different_years.png}
+  \caption[
+  Gregory regression for different time periods.
+  ]{
+    As in \cref{fig:02:gregory_regression} but for different time periods
+    considered in the Gregory regression resulting in different values for the
+    \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
+    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
+    different time periods of a long running simulation (see legend). The
+    initial years are simulated many times for different initial conditions.
+    Light gray colors correspond to the \acs{CMIP}5 ensemble (150 years each).
+    \SpringerAdaptedFrom{Knutti2017}.
+  }
+  \label{fig:02:gregory_regression_different_years}
+\end{figure}
 
@@
 concentration \autocite{Huang2014}.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\FigureWidth{}]{
-    ch02_scientific_background/figs/gregory_regression_different_years.png}
-  \caption[
-    Gregory regression for different time periods.
-  ]{
-    As in \cref{fig:02:gregory_regression} but for different time periods
-    considered in the Gregory regression resulting in different values for the
-    \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
-    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
-    different time periods of a long running simulation (see legend). The
-    initial years are simulated many times for different initial conditions.
-    Light gray colors correspond to the \acs{CMIP}5 ensemble (150 years each).
-    \SpringerAdaptedFrom{Knutti2017}.
-  }
-  \label{fig:02:gregory_regression_different_years}
-\end{figure}
 
@@
 population centers. Moreover, its elevation of more than $3000 \unit{m}$
-provides access to the well-mixed air of the Pacific Ocean in high altitudes,
-which prevents any interference from the vegetation present on the Hawaiian
-Islands.
+provides access to the free troposphere where \ac{CO2} is well mixed, which
+prevents any interference from the vegetation present on the Hawaiian Islands.
 
@@
     \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
-    concentration is superimposed with a seasonal oscillation caused by the
-    seasonal \acs{CO2} cycle (see \cref{subsec:02:carbon_cycle_overview}).
+    concentration is superimposed with a pronounced seasonal oscillation caused
+    by the seasonal \acs{CO2} cycle (see
+    \cref{subsec:02:carbon_cycle_overview}).
   }
@@
 anthropogenically emitted \ac{CO2} directly influences the carbon exchange
-fluxes of the global carbon cycle. Due to the excessive carbon in the
-atmosphere, there is now a net carbon flux from the atmosphere into the land
-and ocean reservoirs (see \cref{fig:02:carbon_cycle_perturbation}). Thus, the
-carbon cycle is not in a steady state anymore. In the decade
-\range{2010}{2019}, anthropogenic activities caused net carbon fluxes of $3.4
-\unit{\GtCyr{}}$ from the atmosphere into the terrestrial biosphere due to
-increased plant photosynthesis and $2.5 \unit{\GtCyr{}}$ from the atmosphere
-into the ocean due to increase dissolution of \ac{CO2} into the sea
-\autocite{Friedlingstein2020}. In the same time, the amount of carbon in the
-atmosphere reservoir increased with a rate of $5.1 \unit{\GtCyr{}}$, indicating
-that only about half of the anthropogenic \ac{CO2} emissions in the last decade
-remained in the atmosphere \autocite{Friedlingstein2020} where they can act as
-\ac{GHG}.
+fluxes of the global carbon cycle. Due to the excess carbon in the atmosphere,
+there is now a net carbon flux from the atmosphere into the land and ocean
+reservoirs (see \cref{fig:02:carbon_cycle_perturbation}). Thus, the carbon
+cycle is not in a steady state anymore. In the decade \range{2010}{2019},
+anthropogenic activities caused net carbon fluxes of $3.4 \unit{\GtCyr{}}$ from
+the atmosphere into the terrestrial biosphere due to increased plant
+photosynthesis and $2.5 \unit{\GtCyr{}}$ from the atmosphere into the ocean due
+to increase dissolution of \ac{CO2} into the sea \autocite{Friedlingstein2020}.
+In the same time, the amount of carbon in the atmosphere reservoir increased
+with a rate of $5.1 \unit{\GtCyr{}}$, indicating that only about half of the
+anthropogenic \ac{CO2} emissions in the last decade remained in the atmosphere
+\autocite{Friedlingstein2020} where they can act as \ac{GHG}.
 
@@
 thesis, the most relevant source of uncertainty is the climate response
-uncertainty. It originates from our imperfect knowledge on how the climate
-system will respond to external forcing. In a multi-model ensemble (\eg{} from
-\ac{CMIP}), this is expressed as the different responses of the different
-climate models to a given forcing. A common approach to distill information
-about a projected quantity from multi-model ensembles is to treat the
-arithmetic \ac{MMM} and the multi-model range as best estimate and uncertainty
-measure of this quantity \autocite{Collins2013}. This \emph{model democracy}
-approach basically assumes that all climate models are independent, equally
-plausible, distributed around reality and that the projected multi-model range
-is representative for the uncertainty in the projected quantity
+uncertainty. It originates from necessary simplifications that have to be
+implemented into the climate models due to limited computational resources  and
+from our imperfect knowledge on how the climate system will respond to external
+forcing. In a multi-model ensemble (\eg{} from \ac{CMIP}), the climate response
+uncertainty is expressed in the different responses of the different climate
+models to a given forcing. A common approach to distill information about a
+projected quantity from multi-model ensembles is to treat the arithmetic
+\ac{MMM} and the multi-model range as best estimate and uncertainty measure of
+this quantity \autocite{Collins2013}. This \emph{model democracy} approach
+basically assumes that all climate models are independent, equally plausible,
+distributed around reality and that the projected multi-model range is
+representative for the uncertainty in the projected quantity
 \autocite{Knutti2017a}. However, since the \ac{CMIP} ensembles, sometimes

### 2021-03-15 ch03_esmvaltool/esmvaltool.tex
Optimized chapters 3, 4 and 7

@@
 contributions that are not documented in other
-\namecrefs{ch:04:papers_ecs_tcr_assessment} of this thesis
+\namecrefs{ch:04:papers_ecs_tcr_assessment} of this thesis and puts them into
+context
 (\cref{sec:03:contributions_to_esmvalcore,sec:03:contributions_to_esmvaltool}).
@@
 the \ac{ESMValTool} are given in its extensive documentation
-(\url{https://esmvaltool.readthedocs.io}).
+(\url{https://docs.esmvaltool.org}).
 
@@
     \cmidrule(l){2-3}
-    & rsntcs & Clear-sky net \ac{TOA} shortwave radiation \\
+    & rsntcs & Clear-sky net \acs{TOA} shortwave radiation \\
     \cmidrule(l){2-3}
@@
 an arbitrary coordinate, which is usually time. A common application of this
-preprocessor is the calculation of a variable's diurnal or seasonal cycle
-amplitude. The \emph{land/sea fraction weighting} preprocessor weights fields
-with the land or sea fraction of the respective grid cells. For example, this
-is necessary for the spatial integration of flux-related variables which are
-reported in units of \enquote{per square meter of land/sea} and not in
+preprocessor is the computation of a variable's diurnal or seasonal cycle
+amplitude, which is for example necessary for the calculation of the predictor
+variable (\ac{CO2} seasonal cycle amplitude) of the emergent constraint on
+\ac{CO2} fertilization by \textcite{Wenzel2016} that is also used in
+\cref{ch:06:paper_gpp} to constrain the global mean \ac{GPP} at the end of the
+\nth{21} century. The \emph{land/sea fraction weighting} preprocessor weights
+fields with the land or sea fraction of the respective grid cells. For example,
+this is necessary for the spatial integration of flux-related variables which
+are reported in units of \enquote{per square meter of land/sea} and not in
 \enquote{per square meter of grid cell}. After the weighting, the grid cell
 areas can be used to integrate the flux-related variable over a desired region.
-The \emph{trend} preprocessor calculates the linear trend of a variable along
-an arbitrary coordinate. The linear trend is defined as the slope of an
-ordinary least squares linear regression of the variable against the selected
-coordinate. For example, this can be used to calculate the temporal trend of
-the \ac{GSAT} over the \nth{20} century.
+In this thesis, \emph{land/sea fraction weighting} is applied to \ac{GPP}
+fields given by observational products and models in order to compute the
+corresponding global mean results (see \cref{ch:06:paper_gpp}). The
+\emph{trend} preprocessor calculates the linear trend of a variable along an
+arbitrary coordinate. The linear trend is defined as the slope of an ordinary
+least squares linear regression of the variable against the selected
+coordinate. For example, this can be used to calculate the historical trend of
+the \ac{GSAT}, which is used as a predictor for \ac{ECS} in several
+recently-published emergent constraints \autocite{JimenezdelaCuesta2019,
+  Nijsse2020, Tokarska2020}.
 
@@
 
-Apart from new preprocessor functions, six derivation scripts for the
+In addition to new preprocessor functions, six derivation scripts for the
 non-\ac{CMOR} variables \emph{asr}, \emph{co2s}, \emph{et}, \emph{rlntcs},
-\emph{rsntcs} and \emph{uajet} have been added. The absorbed shortwave
-radiation (\emph{asr}) is defined as the difference in the incoming \ac{TOA}
-shortwave radiation and the outgoing \ac{TOA} shortwave radiation. The
-atmospheric \ac{CO2} concentration at surface (\emph{co2s}) can be calculated
-from the pressure level--dependent atmospheric \ac{CO2} concentration and the
-surface air pressure using interpolation. An example illustrating \emph{co2s}
-is given in \cref{fig:03:co2s}. \Cref{fig:03:co2s:a} shows the \acs{CMIP}6
-\ac{MMM} of the global \emph{co2s} for the months \acf{JJA} of the year 2014 in
-the emission-driven historical simulation. As expected, the map shows high
-atmospheric \ac{CO2} concentrations over large metropolitan areas with high
-\ac{CO2} emissions (\eg{} over North America, the Arabian Peninsula and East
-Asia). Moreover, regions with high photosynthetic activity (\eg{} the boreal
-forests in the Northern Hemisphere summer and the tropical rainforests in South
-America) exhibit smaller \ac{CO2} concentrations. \Cref{fig:03:co2s:b} shows
-the monthly mean \emph{co2s} at Mauna Loa, Hawaii from 1958 to 2014 for
-observations performed at the Mauna Loa Observatory \autocite{Keeling2005}
-(thick black line) and emission-driven historical simulations from four
-\acs{CMIP}6 models (remaining colored lines). Apart from the model CNRM-ESM2-1,
-the simulated atmospheric \ac{CO2} concentrations show the expected seasonal
-cycle present in the Keeling Curve (see \cref{fig:02:keeling_curve}). All
-\acp{ESM} correctly simulate the increase in \emph{co2s} over the years.
-However, there are some differences in the absolute values with the model
-BCC-CSM2-MR showing the largest deviations.
+\emph{rsntcs} and \emph{uajet} have been implemented into the \ac{ESMValCore}.
+Apart from the atmospheric \ac{CO2} concentration at surface (\emph{co2s}),
+which is used in \cref{ch:06:paper_gpp} to derive the predictor variable of the
+emergent constraint by \textcite{Wenzel2016}, these variables are not directly
+used in this thesis but have been added to reproduce the \ac{MDER} analysis of
+\textcite{Wenzel2016a} (\emph{asr} and \emph{uajet}) or have been used in
+earlier versions of the studies presented in this thesis (\emph{et} for the
+study presented in \cref{ch:06:paper_gpp}; \emph{rlntcs} and \emph{rsntcs} for
+the studies presented in \cref{ch:04:papers_ecs_tcr_assessment}).
+
+The absorbed shortwave radiation (\emph{asr}) is defined as the difference in
+the incoming \ac{TOA} shortwave radiation and the outgoing \ac{TOA} shortwave
+radiation. The atmospheric \ac{CO2} concentration at surface (\emph{co2s}) can
+be calculated from the pressure level--dependent atmospheric \ac{CO2}
+concentration and the surface air pressure using interpolation. An example
+illustrating \emph{co2s} is given in \cref{fig:03:co2s}. \Cref{fig:03:co2s:a}
+shows the \acs{CMIP}6 \ac{MMM} of the global \emph{co2s} for the months
+\acf{JJA} of the year 2014 in the emission-driven historical simulation. As
+expected, the map shows high atmospheric \ac{CO2} concentrations over large
+metropolitan areas with high \ac{CO2} emissions (\eg{} over North America, the
+Arabian Peninsula and East Asia). Moreover, regions with high photosynthetic
+activity (\eg{} the boreal forests in the Northern Hemisphere summer and the
+tropical rainforests in South America) exhibit smaller \ac{CO2} concentrations.
+\Cref{fig:03:co2s:b} shows the monthly mean \emph{co2s} at Mauna Loa, Hawaii
+from 1958 to 2014 for observations performed at the Mauna Loa Observatory
+\autocite{Keeling2005} (thick black line) and emission-driven historical
+simulations from four \acs{CMIP}6 models (remaining colored lines). Apart from
+the model CNRM-ESM2-1, the simulated atmospheric \ac{CO2} concentrations show
+the expected seasonal cycle present in the Keeling Curve (see
+\cref{fig:02:keeling_curve}). All \acp{ESM} correctly simulate the increase in
+\emph{co2s} over the years. However, there are some differences in the absolute
+values with the model BCC-CSM2-MR showing the largest deviations.
 
@@
 latent heat vaporization constant. The longwave/shortwave clear-sky net
-\acf{TOA} radiations (\emph{rlntcs}/\emph{rsntcs}) is defined as the difference
+\ac{TOA} radiations (\emph{rlntcs}/\emph{rsntcs}) is defined as the difference
 between the longwave/shortwave incoming \ac{TOA} radiation assuming clear-sky
@@
 
-\begin{table}[!b]
+\begin{table}[t]
   \centering
@@
     & MLS-Aura & \autocite{Read2015} \\
-    & MTE & \autocite{Jung2011} \\
+    & FLUXNET-MTE & \autocite{Jung2011} \\
     & NDP & \autocite{Gibbs2006} \\
@@
 scripts can be used to apply the \ac{MDER} method to arbitrary target variables
-and predictors.
+and predictors. Both recipes include analyses that contributed ideas and
+methods for the study presented in \cref{ch:06:paper_gpp}, in which the
+\ac{MDER} method is expanded with a \ac{ML} approach and applied to a carbon
+cycle--related target variable (\ac{GPP}).
 
@@
 (\cref{ch:06:paper_gpp}). \emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml}
-contain diagnostic scripts that are able to calculate the climate metrics
+contain diagnostic scripts that allow the calculation of the climate metrics
 \ac{ECS} and \ac{TCR} for arbitrary model output (see
-\cref{sec:02:climate_sensitivity}). Corresponding plots are shown in
-\cref{fig:02:gregory_regression,fig:02:tcr}. \emph{recipe\_cox18nature.yml}
-reproduces the analysis of \textcite{Cox2018}, who introduce an emergent
-constraint in \ac{ECS} based on a global temperature variability metric $\Psi$.
-\Cref{fig:03:cox18:a} shows $\Psi$ over the past 85 years for \acs{CMIP}5
-models and observations form HadCRUT4 \autocite{Morice2012}. Since
-high-sensitivity climate models exhibit higher values of $\Psi$ than
-low-sensitivity models, an emergent relationship between \ac{ECS} and $\Psi$
-can be established (\cref{fig:03:cox18:b}). With observations of $\Psi$,
-\ac{ECS} can eventually be constrained to $\left( 2.8 \pm 0.6 \right) \unit{K}$
-($66 \unit{\%}$ confidence range). More details on this including the
-calculation of $\Psi$ are given in \cref{subsec:05:cox}.
+\cref{sec:02:climate_sensitivity}), which is also used in
+\cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs}. Corresponding plots are
+shown in \cref{fig:02:gregory_regression,fig:02:tcr}.
+\emph{recipe\_cox18nature.yml} reproduces the analysis of \textcite{Cox2018},
+who introduce an emergent constraint in \ac{ECS} based on a global temperature
+variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
+years for \acs{CMIP}5 models and observations form HadCRUT4
+\autocite{Morice2012}. Since high-sensitivity climate models exhibit higher
+values of $\Psi$ than low-sensitivity models, an emergent relationship between
+\ac{ECS} and $\Psi$ can be established (\cref{fig:03:cox18:b}). With
+observations of $\Psi$, \ac{ECS} can eventually be constrained to $\left( 2.8
+\pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). The emergent
+constraint by \textcite{Cox2018} is, together with ten other emergent
+constraint on \ac{ECS}, analyzed in great detail for the \acs{CMIP}5 and
+\acs{CMIP}6 ensemble in \cref{ch:05:paper_ecs}, which also includes a
+description of the calculation of $\Psi$ in \cref{subsec:05:cox}.
 
@@
 figure only shows a linear regression fit between the two variables and not the
-expected reciprocal relation.
+expected reciprocal relation. The output of \emph{recipe\_flato13ipcc.yml}
+provided useful input for the analysis of the climate sensitivity in
+\acs{CMIP}6 performed in \textcite{Meehl2020}, which is also presented in
+\cref{ch:04:papers_ecs_tcr_assessment}.
 
@@
 the corresponding observational datasets is shown in
-\cref{tab:03:changes_to_esmvaltool}.
+\cref{tab:03:changes_to_esmvaltool}. The majority of these datasets is used for
+analysis presented in this thesis (\eg{} see
+\cref{tab:05:overview_emergent_constraints,tab:06:predictors}).

### 2021-03-15 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Optimized chapters 3, 4 and 7

@@
   \caption[
-    \Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
+    \Acf{ECS} and \acf{TCR} calculated for the \acs{CMIP}5 models.
   ]{
-    \Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models. Details on
-    the calculation of \acs{ECS} and \acs{TCR} are given in
-    \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively. The \acf{MMM}
-    is calculated from the Gregory regression method using the \acs{MMM} net
-    \acf{TOA} radiation and the \acs{MMM} change in \acf{GSAT} similar to
-    \cref{fig:04:gregory_regression_cmip6_mmm}. The multi-model standard
-    deviation is given by the sample standard deviation of \acs{ECS} evaluated
-    over all climate models (using the normalization $1 / M$, where $M$ is the
-    number of models). Corresponding references for each model are given in
-    \cref{tab:app:a:cmip5_models}. \AdaptedFrom{Meehl2020}.
+    \Acf{ECS} and \acf{TCR} calculated for the \acs{CMIP}5 models. Details on
+    this are given in \cref{subsec:02:ecs} and \cref{subsec:02:tcr},
+    respectively. The \acf{MMM} is calculated from the Gregory regression
+    method using the \acs{MMM} net \acf{TOA} radiation and the \acs{MMM} change
+    in \acf{GSAT} similar to \cref{fig:04:gregory_regression_cmip6_mmm}. The
+    multi-model standard deviation is given by the sample standard deviation of
+    \acs{ECS} evaluated over all climate models (using the normalization $1 /
+    M$, where $M$ is the number of models). Corresponding references for each
+    model are given in \cref{tab:app:a:cmip5_models}. \AdaptedFrom{Meehl2020}.
   }
@@
 assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
-from the assessed range are insignificant to the ones present in the
+from the assessed range are small compared to the ones present in the
 \acs{CMIP}6 ensemble, which have become evident over the course of 2020. As
@@
     liquid clouds \autocite{BodasSalcedo2019, Gettelman2019, Mulcahy2020,
-      Williams2020}. Thus, in these models a cloud phase change due to global
-    warming is no longer possible, which removes (or significantly reduces) the
-    strong negative cloud phase change feedback (orange arrows). This leads to
-    an increase in the global net cloud feedback \autocite{BodasSalcedo2019,
-      Tan2016} and the \acf{ECS} in the corresponding \acs{CMIP}6 models
-    \autocite{Andrews2019, Gettelman2019}, which is a possible explanation for
-    the high \acs{ECS} values found in many \acs{CMIP}6 models.
+      Williams2020}. Thus, in the affected \acs{CMIP}6 models, this cloud phase
+    change due to warming is reduced by allowing for supercooled cloud liquid,
+    which substantially reduces the strong negative cloud phase change feedback
+    (orange arrows). This leads to an increase in the global net cloud feedback
+    \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS} in the corresponding
+    \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019}, which is a
+    possible explanation for the high \acs{ECS} values found in many
+    \acs{CMIP}6 models.
   }
@@
 described above that results from a cloud phase change from ice clouds in the
-present-day to liquid clouds in the future is removed (or significantly
-reduced) in those \acs{CMIP}6 models that simulate predominantly liquid clouds
-over the present-day Southern Ocean. This is illustrated in
+present-day to liquid clouds in the future is substantially reduced in those
+\acs{CMIP}6 models that simulate predominantly liquid clouds over the
+present-day Southern Ocean. This is illustrated in
 \cref{fig:04:cloud_phase_change_feedback} (orange arrows). Moreover, this
@@
 the \ac{ECS} of the affected \acs{CMIP}6 models \autocite{Andrews2019,
-  Gettelman2019}, offering a possible explanation for the high \ac{ECS} values
-in some \acs{CMIP}6 models.
+  Gettelman2019}, offering a possible explanation for the high \ac{ECS} values in
+some \acs{CMIP}6 models.
 
@@
 \unit{\degree S}$ in the shortwave \ac{CRE} feedback parameter (see
-\cref{fig:04:cloud_feedback_parameters:b}) shows where the climate models are
-switching from simulating almost purely liquid clouds (north of $60
-\unit{\degree S}$) to clouds with an increasing ice component (south of $60
-\unit{\degree S}$) in their pre-industrial and present-day experiments. The
-increasing fraction of ice in the simulated mixed-phase clouds when moving
-south supports a negative shortwave cloud feedback on warming in these areas.
+\cref{fig:04:cloud_feedback_parameters:c}) roughly shows where the climate
+models simulate a high fraction of ice clouds (south of $60 \unit{\degree
+  S}$) for which possible phase changes under global warming are potentially
+important. Such possible phase changes (from ice to liquid) contribute to a
+negative shortwave cloud feedback in this region as liquid clouds tend to be
+more reflective than ice clouds with a similar cloud water content.
 \Cref{fig:04:cloud_feedback_parameters:b,fig:04:cloud_feedback_parameters:d}

### 2021-03-15 ch07_conclusion/conclusion.tex
Optimized chapters 3, 4 and 7

@@
 the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015},
-this scientific question is not only relevant to climate scientists but to
-policymakers and the whole human society. In the light of the large spread of
-climate sensitivity in the most recent generation of climate models from
-\acs{CMIP}6 \autocite{Meehl2020}, a careful statistical evaluation and
+research in this field of science is not only relevant to climate scientists
+but to policymakers and the whole human society. In the light of the large
+spread of climate sensitivity in the most recent generation of climate models
+from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical evaluation and
 refinement of the output of multi-model climate projections is as relevant as
@@
 phase change from ice clouds in the present-day to liquid clouds in the future.
-In the affected \acs{CMIP}6 models, this cloud phase change is not possible
-anymore since the models predominantly simulate liquid clouds over the
-present-day Southern Ocean.
+In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
+reduced by allowing for supercooled cloud liquid. As a consequence, these
+models simulate less cloud ice over the present-day Southern Ocean than their
+predecessor versions with no supercooled cloud liquid.
 
@@
 \cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
-be answered with a \enquote{no} for the \acs{CMIP}6 ensemble. The increased
-best estimates and spreads resulting from the emergent constraints in
+be answered with a \enquote{not very well} for the \acs{CMIP}6 ensemble. The
+increased best estimates and spreads resulting from the emergent constraints in
 \acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
@@
 
-\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for the
-increased climate sensitivity in the \acs{CMIP}6 models (changes in the
-aerosol-cloud interaction and changes in the shortwave cloud phase change
-feedback over the Southern Ocean). While these are likely explanations, there
-might be additional relevant aspects that have not been analyzed in detail yet.
-This could include additional feedback processes that are not present in older
-\ac{CMIP} generations or missing feedback processes that appear in previous
-\ac{CMIP} models but are absent in the \acs{CMIP}6 ensemble. Obvious candidates
-for these feedback processes are cloud-related feedbacks, which pose a major
-source of uncertainty in climate sensitivity in modern-day \acp{ESM}
-\autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
-can further help to gain a better understanding of the \acs{CMIP}6 models and
-potentially to reduce associated uncertainties in the entire multi-model
-ensemble.
+Climate sensitivity is a policy-relevant and easy-to-use metric to assess the
+strength of climate change. The range of model results for this important
+climate metric, however, remains large and has not been narrowed down over the
+last decades. In the latest generation of climate models contributing to
+\acs{CMIP}6, the range has even increased.
+\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
+changes in the aerosol-cloud interaction and changes in the shortwave cloud
+phase change feedback over the Southern Ocean. While these are likely
+explanations, there might be additional relevant aspects that have not been
+analyzed in detail yet. This could include additional feedback processes that
+are not present in older \ac{CMIP} generations or missing feedback processes
+that appear in previous \ac{CMIP} models but are absent in the \acs{CMIP}6
+ensemble. Obvious candidates for these feedback processes are cloud-related
+feedbacks, which pose a major source of uncertainty in climate sensitivity in
+modern-day \acp{ESM} \autocite{Boucher2013}. Identifying and quantifying such
+feedback mechanisms can further help to gain a better understanding of the
+\acs{CMIP}6 models and potentially to reduce associated uncertainties in the
+entire multi-model ensemble.
 
-Due to the flexibility of the new \ac{ML}-based climate model weighting
-approach presented in \cref{ch:06:paper_gpp}, one natural way forward is to
-apply this technique to other target variables. While the scalar metrics
-\ac{ECS} or \ac{TCR} with one value per climate model do not offer enough
-training points for this technique, a possible alternative could be the gridded
-\acl{T} over the \nth{21} century in the \ac{SSP} scenarios. A relevant study
-in this context is given by \textcite{Brunner2020}, who use performance and
-interdependence-based climate model weighting to constrain the \acl{T} over the
-\nth{21} century based on observable process-based predictors of today's
-climate and which could serve as a possible baseline for the new \ac{ML}-based
-weighting approach. An exciting and valuable addition to the method could be
-causal inference \autocite{Nowack2020, Runge2019}. In the current approach, the
-relationships between the different predictors and the target variable are
-based on statistical correlation alone. Moreover, the local feature importance
-maps do not reveal true causal relationships between the predictors and the
-target variable, but only show the relative weight that is given to a specific
-predictor at a specific location for the prediction of the target variable by
-the \ac{ML} model. By integrating causal networks into the new approach, it
-might be possible to explore and utilize true causal connections instead.
+A promising way forward to reduce uncertainties in multi-model climate
+projections is to apply the new flexible \ac{ML}-based climate model weighting
+approach introduced in \cref{ch:06:paper_gpp} to other target variables. While
+the scalar metrics \ac{ECS} or \ac{TCR} with one value per climate model do not
+offer enough training points for this technique, a possible alternative could
+be the gridded \acl{T} over the \nth{21} century in the \ac{SSP} scenarios. A
+relevant study in this context is given by \textcite{Brunner2020}, who use
+performance and interdependence-based climate model weighting to constrain the
+\acl{T} over the \nth{21} century based on observable process-based predictors
+of today's climate and which could serve as a possible baseline for the new
+\ac{ML}-based weighting approach. An exciting and valuable addition to the
+method could be causal inference \autocite{Nowack2020, Runge2019}. In the
+current approach, the relationships between the different predictors and the
+target variable are based on statistical correlation alone. Moreover, the local
+feature importance maps do not reveal true causal relationships between the
+predictors and the target variable, but only show the relative weight that is
+given to a specific predictor at a specific location for the prediction of the
+target variable by the \ac{ML} model. By integrating causal networks into the
+new approach, it might be possible to explore and utilize true causal
+connections instead.
 

### 2021-03-22 02_abstract/abstract.tex
Cleaned abstract, introduction, scientific background and conclusion

@@
 A precise quantification of climate change is crucial in order to assess
-optimal mitigation and adaptation strategies. Valuable tools in this context
-are \acfpAbstract{ESM}, which are state-of-the-art climate models that allow
-numerical simulations of the complex physical, biological and chemical
-processes of the Earth system. Of particular relevance are projections of the
-future climate, which enable an assessment of future climate change. Due to the
-chaotic nature of the climate system, unknowns in future emission pathways and
-uncertainties in the climate models, projections of the future climate are
-associated with large uncertainties. The main focus of this thesis is the
-analysis of future climate projections from \acsp{ESM} participating in the
-\acfAbstract{CMIP} and the reduction of corresponding uncertainties with
-observations.
+optimal mitigation and adaptation strategies. \AcfpAbstract{ESM}, which are
+state-of-the-art climate models that allow numerical simulations of the complex
+physical, biological and chemical processes of the Earth system, are common
+tools to understand and project climate change. Due to the chaotic nature of
+the climate system, unknowns in future emission pathways and uncertainties in
+the climate models, projections of the future climate are associated with large
+uncertainties. The main focus of this thesis is the analysis of future climate
+projections from \acsp{ESM} participating in the \acfAbstract{CMIP} and the
+reduction of uncertainties in climate projections with observations.
 
@@
 the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely \acs{ECS}
-range of $\rangeunit{1.5}{4.5}{K}$ assessed by the Fifth \acl{AR} (\acs{AR}5)
-of the \acl{IPCC} from 2013. Similarly, the transient response of the climate
-system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR}, shows an
-inter-model range of $\rangeunit{1.3}{3.0}{K}$, whose upper bound is again
-higher than the likely range assessed in \acs{AR}5 of
+range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC} Fifth
+\acl{AR} (\acs{AR}5) from 2013. Similarly, the transient response of the
+climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
+shows an inter-model range of $\rangeunit{1.3}{3.0}{K}$ with an upper bound
+again higher than the likely range assessed in \acs{AR}5 of
 $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
 sensitivity in many \acs{CMIP}6 models is a change in the microphysical
-representation of mixed-phase clouds above the Southern Ocean, which reduces
-the strong negative shortwave cloud phase change feedback in this region that
-is present in climate models from previous \acs{CMIP} generations.
+representation of mixed-phase clouds over the Southern Ocean, which reduces the
+strong negative shortwave cloud phase change feedback in this region that is
+present in climate models from previous \acs{CMIP} generations.
 

### 2021-03-22 ch01_introduction/introduction.tex
Cleaned abstract, introduction, scientific background and conclusion

@@
 
-Climate change is one of the greatest challenges for humankind today with
-far-reaching consequences for all human beings on planet Earth. The warming of
-the climate system is \enquote{unequivocal} and \enquote{many of the observed
-  changes are unprecedented over decades to millennia} \autocite{IPCC2014}. The
-changing climate increases the \enquote{likelihood of severe, pervasive and
-  irreversible impacts for people, species and ecosystems} with \enquote{mostly
-  negative impacts for biodiversity, ecosystem services and economic
-  development} and amplifies \enquote{risks for livelihoods and for food and
-human security} \autocite{IPCC2014}. Potential drivers for climate change are
-all natural and anthropogenic substances and processes that may alter the
-Earth's energy budget. The human influence on the climate system is clear:
-Ever increasing emissions of \acp{GHG} since the end of the pre-industrial
-era largely driven by economic and population growth led to atmospheric
-concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O} that are
-\enquote{unprecedented in at least the last 800000 years}
+Climate change is one of the greatest challenges for humankind today. The
+warming of the climate system is \enquote{unequivocal} and \enquote{many of the
+  observed changes are unprecedented over decades to millennia}
+\autocite{IPCC2014}. The changing climate increases the \enquote{likelihood of
+  severe, pervasive and irreversible impacts for people, species and
+  ecosystems} with \enquote{mostly negative impacts for biodiversity, ecosystem
+services and economic development} and amplifies \enquote{risks for
+livelihoods and for food and human security} \autocite{IPCC2014}. Potential
+drivers for climate change are all natural and anthropogenic substances and
+processes that may alter the Earth's energy budget. The human influence on
+the climate system is clear: Ever increasing emissions of \acp{GHG} since the
+end of the pre-industrial era largely driven by economic and population
+growth led to atmospheric concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O}
+that are \enquote{unprecedented in at least the last 800000 years}
 \autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other
@@
 directly influence the global carbon cycle, an important biogeochemical cycle
-of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}}, which
-describes the carbon uptake of the terrestrial biosphere due to photosynthesis.
-Since this land carbon uptake absorbs about $30 \unit{\%}$ of the anthropogenic
-\ac{CO2} emissions in today's climate \autocite{Friedlingstein2020}, this
-process substantially slows down global warming and directly contributes to the
-magnitude of the climate sensitivity. Other important anthropogenic drivers of
-climate change are the emission of aerosols and land use/land cover changes.
-Apart from anthropogenic drivers, there are also natural processes which impact
-the climate system like changes in the solar activity or volcanic eruptions.
-However, there is clear evidence that these natural drivers alone cannot
-explain the observed climate change \autocite{Haustein2017}.
+of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}},
+which describes the carbon uptake of the terrestrial biosphere due to
+photosynthesis. Since this land carbon uptake absorbs about $30 \unit{\%}$ of
+the anthropogenic \ac{CO2} emissions in today's climate
+\autocite{Friedlingstein2020}, this process substantially slows down global
+warming and directly contributes to the magnitude of the climate sensitivity.
+Other important anthropogenic drivers of climate change are the emission of
+aerosols and land use/land cover changes. Apart from anthropogenic drivers,
+there are also natural processes which impact the climate system like changes
+in the solar activity or volcanic eruptions. However, there is clear evidence
+that these natural drivers alone cannot explain the observed climate change
+\autocite{Haustein2017}.
 
@@
 from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
-  as \acs{CMIP}6} will be used in the upcoming Sixth \acl{AR} (\acs{AR}6) of
-the \ac{IPCC}, and their predecessor models from \acs{CMIP}5
-\autocite{Taylor2012} have been successfully analyzed in the Fifth \acl{AR}
+  as \acs{CMIP}6} support the assessment of the upcoming \ac{IPCC} Sixth
+\acl{AR} (\acs{AR}6), and their predecessor models from \acs{CMIP}5
+\autocite{Taylor2012} have been assessed as part of the Fifth \acl{AR}
 (\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which

### 2021-03-22 ch02_scientific_background/scientific_background.tex
Cleaned abstract, introduction, scientific background and conclusion

@@
     circular sectors show additional science themes that can be explored
-    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}.
+    through the 23 \acs{CMIP}6-Endorsed \acfp{MIP}.
     \SpringerAdaptedFrom{Simpkins2017}.
@@
 can participate in the so-called \acs{CMIP}6-Endorsed \acp{MIP}, of which
-\acs{CMIP}6 offers 21 (see circular sectors in \cref{fig:02:cmip6}). Some
+\acs{CMIP}6 offers 23 (see circular sectors in \cref{fig:02:cmip6}). Some
 \acp{MIP} offer additional experiments to explore specific aspects of the Earth
@@
 
-Simulations from climate model ensembles of \ac{CMIP} allow us to assess the
-impact of future climate change in a consistent and transparent way. Especially
-the \ac{ScenarioMIP} experiments can give valuable insights into the upcoming
-development of the Earth system by providing \emph{projections} of the future
+Simulations from climate model ensembles of \ac{CMIP} allow us to assess future
+climate change in a consistent and transparent way. Especially the
+\ac{ScenarioMIP} experiments can give valuable insights into possible
+developments of the Earth system by providing \emph{projections} of the future
 climate. In contrast to climate predictions, climate projections run over

### 2021-03-22 ch07_conclusion/conclusion.tex
Cleaned abstract, introduction, scientific background and conclusion

@@
 The analysis of future climate projections from numerical climate model
-simulations is of paramount importance for an accurate quantification of the
-future climate change. Since this involves metrics of public interest like the
-allowable fossil fuel emissions to meet particular warming targets, for example
-the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015},
-research in this field of science is not only relevant to climate scientists
-but to policymakers and the whole human society. In the light of the large
-spread of climate sensitivity in the most recent generation of climate models
-from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical evaluation and
-refinement of the output of multi-model climate projections is as relevant as
-ever. This thesis quantifies the associated uncertainties, presents the
-evaluation of established methods to reduce these uncertainties in the new
-climate model ensemble and finally introduces a novel alternative technique
-based on supervised \ac{ML}.
+simulations is of paramount importance to assess future climate change under
+different forcing scenarios. Since this involves metrics of public interest
+like the allowable fossil fuel emissions to meet particular warming targets,
+for example the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
+\autocite{UNFCCC2015}, research in this field of science is not only relevant
+to climate scientists but to policymakers and the whole human society. In the
+light of the large spread of climate sensitivity in the most recent generation
+of climate models from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical
+evaluation and refinement of the output of multi-model climate projections is
+as relevant as ever. This thesis quantifies the associated uncertainties,
+presents the evaluation of established methods to reduce these uncertainties in
+the new climate model ensemble and finally introduces a novel alternative
+technique based on supervised \ac{ML}.
 
@@
 performed with the \ac{ESMValTool}. Apart from that, further substantial
-changes and additions to the code base of the tool have been implemented (see
-\cref{ch:03:esmvaltool}), which lead to co-authorship in the scientific
-documentation of the \ac{ESMValTool} \autocite{Eyring2020, Lauer2020,
-  Righi2020, Weigel2020}. Since the open-source tool is freely available for
-everyone, these improvements are beneficial for the entire scientific
-community.
+changes and additions to the code base of the tool have been implemented as
+part of this thesis (see \cref{ch:03:esmvaltool}), which led to co-authorship
+in the technical and scientific documentation of the \ac{ESMValTool}
+\autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}. Since the open-source
+tool is freely available, the code that has been implemented as part of this
+thesis is beneficial for the entire scientific community.
 
-In a first analysis, the climate sensitivity metrics \ac{ECS} and \ac{TCR} are
-evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
-which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
-published in \textcite{Bock2020, Meehl2020}, directly addresses
+In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
+\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
+This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
+already published in \textcite{Bock2020, Meehl2020}, directly addresses
 \cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
@@
 climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
-representation of mixed-phase clouds above the Southern Ocean. For one thing,
-this change leads to an improved simulation of the shortwave \ac{CRE} in these
-models when compared to observations. However, this change also substantially
-reduces the strong negative shortwave cloud feedback over the Southern Ocean
-that is present in previous \ac{CMIP} generations and that results from a cloud
-phase change from ice clouds in the present-day to liquid clouds in the future.
-In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
-reduced by allowing for supercooled cloud liquid. As a consequence, these
-models simulate less cloud ice over the present-day Southern Ocean than their
-predecessor versions with no supercooled cloud liquid.
+representation of mixed-phase clouds above the Southern Ocean. This change
+leads to an improved simulation of the shortwave \ac{CRE} in these models when
+compared to observations. However, this change also substantially reduces the
+strong negative shortwave cloud feedback over the Southern Ocean that is
+present in previous \ac{CMIP} generations and that results from a cloud phase
+change from ice clouds in the present-day to liquid clouds in the future. In
+the affected \acs{CMIP}6 models, this cloud phase change due to warming is
+reduced since these models simulate less cloud ice over the present-day
+Southern Ocean than their predecessor versions with no supercooled cloud
+liquid.
 
 In order to reduce this large range of \ac{ECS} in the latest generation of
-climate  models, suitable emergent constraints on the \acs{CMIP}6 ensemble are
-evaluated. Emergent constraints use a physically-based inter-model relationship
-between an observable quantity of the Earth system and a target variable to
-reduce uncertainties in the target variable with observations
-\autocite{Allen2002}. In total eleven emergent constraints on \ac{ECS} are
-assessed, which are mostly related to cloud feedbacks since these constitute
-the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
-  Flato2013}. Since all of the evaluated emergent constraints have been derived
-on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, out-of-sample tests on the emergent
-constraints can be performed by assessing whether they still hold for the
-\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
-already published in \textcite{Schlund2020a}, a substantial reduction of skill
-for the majority of emergent constraints is found when applied to the
-\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
-skill is expressed as a decrease of the coefficient of determination $R^2$ of
-the emergent relationship and a decrease of the statistical significance using
-the null hypothesis that there is no correlation between the predictor and
-\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
-show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
-confidence) for almost all emergent constraints, resulting in values of
-$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
-$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
-an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
-the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
-\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
-increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
-\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
-be answered with a \enquote{not very well} for the \acs{CMIP}6 ensemble. The
-increased best estimates and spreads resulting from the emergent constraints in
-\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
-of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
-emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
+climate  models, already-published emergent constraints that have been derived
+on models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5
+are evaluated for their skill in the \acs{CMIP}6 ensemble. Emergent constraints
+use a physically-based inter-model relationship between an observable quantity
+of the Earth system and a target variable to reduce uncertainties in the target
+variable with observations \autocite{Allen2002}. In total eleven emergent
+constraints on \ac{ECS} are assessed, which are mostly related to cloud
+feedbacks since these constitute the most important source of uncertainty for
+\ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated emergent
+constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5 ensemble,
+out-of-sample tests on the emergent constraints can be performed by assessing
+whether they still hold for the \acs{CMIP}6 models. In this study, which is
+shown in \cref{ch:05:paper_ecs} and already published in
+\textcite{Schlund2020a}, a substantial reduction of skill for the majority of
+emergent constraints is found when applied to the \acs{CMIP}6 ensemble in
+comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
+decrease of the coefficient of determination $R^2$ of the emergent relationship
+and a decrease of the statistical significance using the null hypothesis that
+there is no correlation between the predictor and \ac{ECS}. Moreover, the
+corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
+\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
+constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
+(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
+constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
+\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
+$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
+\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
+emergent constraints. Thus, \cref{enum:01:question_2}
+(\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to be answered with a
+\enquote{not very well} for the \acs{CMIP}6 ensemble. The increased best
+estimates and spreads resulting from the emergent constraints in \acs{CMIP}6
+are likely related to the increased \ac{MMM} and multi-model spread of \ac{ECS}
+in \acs{CMIP}6. A possible reason for the reduced skill of the emergent
+constraints when applied to the \acs{CMIP}6 ensemble is the increased
 complexity of the \acs{CMIP}6 models: A basic assumption for these
@@
 can be explained with a correction of the historical bias in \ac{GPP} by the
-new approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
-features. This study directly addresses \cref{enum:01:question_3}
-(\enquote{\emph{\KeyScienceQuestionThree{}}}), which can be confidently
-answered with \enquote{yes} based on the results found.
+new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
+and \ac{LAI} are the dominant features. This study directly addresses
+\cref{enum:01:question_3} (\enquote{\emph{\KeyScienceQuestionThree{}}}), which
+can be confidently answered with \enquote{yes} based on the results found.
 
@@
 
-Climate change is a global phenomenon affecting the entire human society. Due
-to the vastly complex nature of the Earth system, accurately quantifying future
-climate change is a highly non-trivial task. This thesis covers several
-important aspects of the climate system and metrics describing the changing
-climate, but it is by no means complete and can be extended in an arbitrary
-number of ways. This final \namecref{sec:07:outlook} presents several
-suggestions on possible ways forward.
-
 Climate sensitivity is a policy-relevant and easy-to-use metric to assess the
 strength of climate change. The range of model results for this important
-climate metric, however, remains large and has not been narrowed down over the
-last decades. In the latest generation of climate models contributing to
-\acs{CMIP}6, the range has even increased.
+climate metric, however, remains large and has not been narrowed over the last
+decades. In the latest generation of climate models contributing to
+\acs{CMIP}6, the range has even increased \autocite{Meehl2020}.
 \Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
@@
 analyzed in detail yet. This could include additional feedback processes that
-are not present in older \ac{CMIP} generations or missing feedback processes
-that appear in previous \ac{CMIP} models but are absent in the \acs{CMIP}6
-ensemble. Obvious candidates for these feedback processes are cloud-related
-feedbacks, which pose a major source of uncertainty in climate sensitivity in
-modern-day \acp{ESM} \autocite{Boucher2013}. Identifying and quantifying such
-feedback mechanisms can further help to gain a better understanding of the
-\acs{CMIP}6 models and potentially to reduce associated uncertainties in the
-entire multi-model ensemble.
+are not present in older \ac{CMIP} generations. Obvious candidates for these
+feedback processes are cloud-related feedbacks, which pose a major source of
+uncertainty in climate sensitivity in modern-day \acp{ESM}
+\autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
+can further help to gain a better understanding of the \acs{CMIP}6 models and
+potentially to reduce associated uncertainties in the entire multi-model
+ensemble. A potentially interesting process relevant in this context is the
+influence of the midlatitude jet position on clouds and \acp{CRE}
+\autocite{Grise2016}.
 
@@
 climate science involve for example \ac{ML}-based parametrizations that are
-learned from high-resolution climate models \autocite{Rasp2018}, \ac{ML}-based
-analysis and prediction of forcing patterns \autocite{Barnes2019,
+learned from high-resolution climate models \autocite{Gentine2018, Rasp2018},
+\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
   Mansfield2020} or learning of entire \acp{ESM} from observational products
@@
 groundbreaking research in climate science for the near future has been laid,
-which will hopefully help us in further understanding and fighting one of the
+which will help to further understand the Earth system and fight one of the
 greatest challenges for humankind today: climate change.

### 2021-03-22 ch02_scientific_background/scientific_background.tex
Added paragraph with details on numerical climate modelling and parameterizations

@@
 A possible way out is given by numerical climate models, which offer the
-possibility to simulate the Earth's climate on a computer. The first numerical
-climate models came up in the 1960s and were based on weather prediction models
-\autocite{Flato2011}. Early models from the 1970s simulated only the physical
-components of the climate system: atmosphere, land surface, ocean and sea ice
-(see \cref{fig:02:esms_historical_evolution}). The basis of these so-called
-\acp{AOGCM} \autocite{Flato2013} is the numerical solving of the differential
-equations describing the exchange of energy, momentum and matter between and
-within these physical components.
+possibility to simulate the Earth's climate on a computer. To efficiently
+replicate the vastly complex Earth system with finite computational resources,
+climate models divide the Earth into a set of \emph{grid cells}. The typical
+size of an atmospheric grid cell in a modern-day global climate model is about
+$100 \unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
+\unit{km}$ along the vertical dimension (pressure level or height). In
+addition, the temporal evolution of the Earth system is described using
+discrete time steps, which are about $30 \unit{min}$ long in a modern-day
+global climate model. For each grid cell and time step, usually a single value
+per model variable is given. Examples for such variables provided by a climate
+model are the atmospheric prognostic variables velocity (horizontal and
+vertical), temperature, specific humidity, pressure and density. These
+prognostic variables are related to each other via the \emph{primitive
+  equations}, which is a set of partial differential equations that can be
+derived from the conservation of momentum, mass, energy and moisture
+\autocite{Holton2004}. To progress further in time, these primitive equations
+are solved numerically by the dynamical core of the climate models, which
+eventually describes the Earth's large-scale atmospheric motion. Similar to
+this, many other processes of the Earth system are simulated by the climate
+models using other fundamental physical laws and principles. As opposed to
+these calculated variables that form the \emph{output} of climate models, the
+corresponding \emph{input} is mainly given by the radiative forcing.
 
@@
 
-Over the course of the years, climate models became more and more complex by
-including a wider range of processes within the components, but also by
-introducing new components to the coupled system. Examples of these are
-aerosols, the carbon cycle, a dynamic vegetation, atmospheric chemistry and
-land ice (see \cref{fig:02:esms_historical_evolution}). \acp{AOGCM} coupled to
-these additional components are called \acfp{ESM}, which are the current
+Many processes in the Earth system occur at spatial scales much smaller than
+the size of a typical grid cell. Illustrative examples for this are clouds,
+which are usually much smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
+cell, but still play an important role in the overall climate system by
+reflecting incoming and outgoing radiation \autocite{Boucher2013}. In order to
+approximate these subgrid-scale processes, a concept called
+\emph{parameterization} is used. Instead of simulating a process exactly,
+parameterizations aim to represent the effect of that process at the grid scale
+of the climate model by generating the appropriate forcing terms for the rest
+of the system and the rest of the processes \autocite{Gettelman2016}.
+Parameterizations are necessary to simulate many processes of the Earth system,
+such as surface heat and moisture fluxes, moist convection, turbulent mixing
+and radiation \autocite{Holton2004}.
+
+The first numerical climate models came up in the 1960s and were based on
+weather prediction models \autocite{Flato2011}. Early models from the 1970s
+simulated only the physical components of the climate system: atmosphere, land
+surface, ocean and sea ice (see \cref{fig:02:esms_historical_evolution}). These
+models are called \emph{\acp{AOGCM}} \autocite{Flato2013}. Over the course of
+the years, climate models became more and more complex by including a wider
+range of processes within the components, but also by introducing new
+components to the coupled system. Examples of these are aerosols, the carbon
+cycle, a dynamic vegetation, atmospheric chemistry and land ice (see
+\cref{fig:02:esms_historical_evolution}). \acp{AOGCM} coupled to these
+additional components are called \emph{\acfp{ESM}}, which are the current
 state-of-the-art models that allow the most sophisticated simulations of the
@@
 Even though all climate models are built on the same physical principles, they
-differ in terms of spatial resolution, processes included and parametrizations
-of unresolved processes. The latter can be thought of as different
-approximations that are necessary to represent processes that take place on
-scales smaller than the common size of grid boxes in modern global climate
-models (about $\ang{1} \times \ang{1}$ horizontally), which is for example the
-case for cloud processes and vegetation processes.
+differ in terms of spatial resolution, processes included and parameterizations
+of unresolved processes (see \cref{subsec:02:numerical_climate_modeling}).
 

### 2021-03-22 ch07_conclusion/conclusion.tex
Added paragraph with details on numerical climate modelling and parameterizations

@@
 climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
-representation of mixed-phase clouds above the Southern Ocean. This change
-leads to an improved simulation of the shortwave \ac{CRE} in these models when
+representation of mixed-phase clouds over the Southern Ocean. This change leads
+to an improved simulation of the shortwave \ac{CRE} in these models when
 compared to observations. However, this change also substantially reduces the
@@
 weighting approach, further possible applications of \ac{AI} and \ac{ML} in
-climate science involve for example \ac{ML}-based parametrizations that are
+climate science involve for example \ac{ML}-based parameterizations that are
 learned from high-resolution climate models \autocite{Gentine2018, Rasp2018},

### 2021-03-23 ch02_scientific_background/scientific_background.tex
Added details on carbon cycle models to chapter 02

@@
 climate models divide the Earth into a set of \emph{grid cells}. The typical
-size of an atmospheric grid cell in a modern-day global climate model is about
+size of an atmospheric grid cell in modern-day global climate models is about
 $100 \unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
 \unit{km}$ along the vertical dimension (pressure level or height). In
-addition, the temporal evolution of the Earth system is described using
-discrete time steps, which are about $30 \unit{min}$ long in a modern-day
-global climate model. For each grid cell and time step, usually a single value
-per model variable is given. Examples for such variables provided by a climate
+addition, also the temporal evolution of the Earth system is discretized using
+time steps that are typically about $30 \unit{min}$ long in modern-day global
+climate models. For each grid cell and time step, usually a single value per
+model variable is given. Examples for such variables provided by a climate
 model are the atmospheric prognostic variables velocity (horizontal and
@@
 prognostic variables are related to each other via the \emph{primitive
-  equations}, which is a set of partial differential equations that can be
-derived from the conservation of momentum, mass, energy and moisture
-\autocite{Holton2004}. To progress further in time, these primitive equations
-are solved numerically by the dynamical core of the climate models, which
-eventually describes the Earth's large-scale atmospheric motion. Similar to
-this, many other processes of the Earth system are simulated by the climate
-models using other fundamental physical laws and principles. As opposed to
-these calculated variables that form the \emph{output} of climate models, the
-corresponding \emph{input} is mainly given by the radiative forcing.
+  equations}, a set of partial differential equations that can be derived from
+the conservation of momentum, mass, energy and moisture \autocite{Holton2004}.
+To progress further in time, these primitive equations are solved numerically
+by the dynamical core of the climate models, which eventually describes the
+Earth's large-scale atmospheric motions. Similar to this, many other processes
+of the Earth system are simulated by the climate models using other fundamental
+physical laws and principles. As opposed to these calculated variables that
+form the output of climate models, the corresponding input is mainly given by
+the radiative forcing.
 
@@
 the size of a typical grid cell. Illustrative examples for this are clouds,
-which are usually much smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
+which are usually smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
 cell, but still play an important role in the overall climate system by
 reflecting incoming and outgoing radiation \autocite{Boucher2013}. In order to
-approximate these subgrid-scale processes, a concept called
+reasonably approximate these subgrid-scale processes, a concept called
 \emph{parameterization} is used. Instead of simulating a process exactly,
@@
 Parameterizations are necessary to simulate many processes of the Earth system,
-such as surface heat and moisture fluxes, moist convection, turbulent mixing
-and radiation \autocite{Holton2004}.
+for example surface heat and moisture fluxes, moist convection, turbulent
+mixing and radiation \autocite{Holton2004}.
 
@@
 
+\subsection{Representation in \aclp{ESM}}
+\label{subsec:02:carbon_cycle_representation}
+
+In modern \acp{ESM}, the carbon cycle is usually represented by a land carbon
+cycle model and an ocean carbon cycle model that are both coupled to the other
+components of the \ac{ESM}. An overarching principle is the conservation of the
+total carbon mass in the Earth system \autocite{Gregory2009}, \ie{}
+\begin{equation}
+  \CE{} = \CA{} + \CL{} + \CO{}.
+  \label{eq:02:carbon_balance}
+\end{equation}
+Here, $\CE{}$ represents the cumulative anthropogenic carbon emissions that are
+distributed among the three main carbon reservoirs of the Earth system.
+$\CA{}$, $\CL{}$, and $\CO{}$ describe the corresponding changes in these
+carbon stores, namely the atmosphere, the terrestrial biosphere (land) and the
+ocean, respectively.
+
+Similar to many cloud processes, most carbon cycle--related processes need to
+be parameterized in \acp{ESM} since they occur on spatial scales much smaller
+than the typical grid cell sizes (see
+\cref{subsec:02:numerical_climate_modeling}). For the terrestrial carbon cycle,
+the major carbon exchange processes that need to be considered are
+photosynthesis, respiration and disturbances, which connect the atmospheric
+carbon reservoir and the land carbon reservoir. The latter mainly consists of
+the vegetation biomass, litter and soil. The terrestrial carbon exchange
+processes are commonly described with \ac{GPP}, \ac{NPP}, \ac{NEP} and
+\ac{NBP}, which are all illustrated in \cref{fig:02:carbon_cycle_processes}.
+\ac{GPP} describes the total carbon uptake of the vegetation by photosynthesis,
+while \ac{NPP} refers to the net carbon gain of the plants after accounting for
+autotrophic respiration ($\Ra{}$). \Ac{NEP} additionally considers
+heterotrophic respiration ($\Rh{}$) and describes the carbon uptake/release of
+the entire ecosystem. Finally, \ac{NBP} describes changes in the long-term
+carbon storage of the terrestrial biosphere by including disturbances like
+fires \autocite{Ciais2013}.
+
+The different land carbon cycle models integrated into modern \acp{ESM} use a
+variety of different parameterizations for the different processes
+\autocite{Arora2020}. Examples for the photosynthesis (\ac{GPP}) are given by
+\textcite{Farquhar1980} for \ch{C3} plants and \textcite{Collatz1992} for
+\ch{C4} plants. The terms \enquote{\ch{C3}} and \enquote{\ch{C4}} refer to the
+different metabolic pathways used in the carbon fixation during photosynthesis.
+In both cases, the enzyme \ac{RuBisCO} is an important catalyst for the
+corresponding chemical reactions. Both parameterizations utilize variables that
+are known to affect photosynthesis, such as the availability of light, the
+\ac{CO2} concentration, the soil moisture and the temperature.
+
 \section{Reducing Uncertainties in Multi-Model Climate Projections

### 2021-03-23 ch06_paper_gpp/paper_gpp.tex
Added details on carbon cycle models to chapter 02

@@
 
-\subsection{Constraining the \texorpdfstring{\acs{CO2}}{CO2} Fertilization
+\subsection{Constraining the \texorpdfstring{\acs{CO2}}{\ch{CO2}} Fertilization
   Effect (Step 1)}
@@
 effect could be different for tropical and extratropical ecosystems due to
-nutrient limitations and the temperature dependence of Rubisco kinetics
+nutrient limitations and the temperature dependence of \ac{RuBisCO} kinetics
 \autocite{CraftsBrandner2000} in the tropics. Nevertheless, we argue that we

### 2021-03-23 ch02_scientific_background/scientific_background.tex
Added figure describing the land carbon uptake processes

@@
   \caption[
-    Schematic illustrating the importance of different sources of
+    Schematic illustration of the importance of different sources of
     uncertainties in climate model projections and their evolution in time.
   ]{
-    Schematic illustrating the importance of different sources of
+    Schematic illustration of the importance of different sources of
     uncertainties in climate model projections and their evolution in time. (a)
     Time series of the anomaly of the decadal and global mean surface
-    temperature relative to the period \range{1961}{1980}. The black line
-    shows the historical observations with estimates of uncertainty from
-    climate models (gray). The remaining colors show different sources of
-    uncertainty in future climate projections: Natural variability (orange),
-    climate response uncertainty (blue) and emission uncertainty (green)
+    temperature relative to the period \range{1961}{1980}. The black line shows
+    the historical observations with estimates of uncertainty from climate
+    models (gray). The remaining colors show different sources of uncertainty
+    in future climate projections: Natural variability (orange), climate
+    response uncertainty (blue) and emission uncertainty (green)
     \autocite{Hawkins2009, Hawkins2010}. Climate response uncertainty can
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch02_scientific_background/figs/carbon_cycle_processes.pdf}
+  \caption[
+    Schematic illustration of the processes of the terrestrial carbon uptake.
+  ]{
+    Schematic illustration of the processes of the terrestrial carbon uptake.
+    The \acf{GPP} describes the total carbon uptake of the vegetation by
+    photosynthesis, while the \acf{NPP} refers to the net carbon gain of the
+    plants after accounting for autotrophic respiration ($\Ra{}$). The
+    \acf{NEP} additionally considers heterotrophic respiration ($\Rh{}$) and
+    describes the carbon uptake/release of the entire ecosystem. Finally, the
+    \acf{NBP} describes changes in the long-term carbon storage of the
+    terrestrial biosphere by including disturbance losses ($\Ld{}$).
+  }
+  \label{fig:02:carbon_cycle_processes}
+\end{figure}
+
 Similar to many cloud processes, most carbon cycle--related processes need to
@@
 carbon storage of the terrestrial biosphere by including disturbances like
-fires \autocite{Ciais2013}.
+fires ($\Ld{}$).
 

### 2021-03-24 02_abstract/abstract.tex
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

@@
 again higher than the likely range assessed in \acs{AR}5 of
-$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
-sensitivity in many \acs{CMIP}6 models is a change in the microphysical
-representation of mixed-phase clouds over the Southern Ocean, which reduces the
+$\rangeunit{1.0}{2.5}{K}$. Possible reasons for the increased climate
+sensitivity in many \acs{CMIP}6 models are aerosol-cloud interactions in the
+simulations due to changed aerosol schemes and changes in the microphysical
+representation of mixed-phase clouds over the Southern Ocean, which reduce the
 strong negative shortwave cloud phase change feedback in this region that is

### 2021-03-24 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

@@
 quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
-on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we use it to assess
-and analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to
-corresponding results of predecessor model generations and international
+on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), it is used here to
+assess and analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare
+it to corresponding results of predecessor model generations and international
 climate assessments. This work is already published in two scientific papers:
@@
 a \onepctcotwo{} run \commentcite{Bindoff2013}{see \cref{subsec:02:tcr}}. Using
-these two definitions and the \ac{ESMValTool}, we assess \ac{ECS} and \ac{TCR}
-for all \acs{CMIP}5 and \acs{CMIP}6 models where the necessary temperature and
-radiation data is available (as of \TheMonth{}).
+these two definitions and the \ac{ESMValTool}, \ac{ECS} and \ac{TCR} are
+assessed for all \acs{CMIP}5 and \acs{CMIP}6 models where the necessary
+temperature and radiation data is available (as of \TheMonth{}).
 \Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an overview over the
@@
 exceeds this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third
-(14 models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
-several modeling centers independently confirmed high \ac{ECS} values in their
-models \autocite{Andrews2019, Gettelman2019, Wyser2020}. A list of possible
-reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble
-including aerosol-cloud interactions and changes in the shortwave cloud
-feedback over the Southern Ocean is given in
+(14 models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to the analysis
+presented here, several modeling centers independently confirmed high \ac{ECS}
+values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
+  Wyser2020}. A list of possible reasons for the increased climate sensitivity
+in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
+the shortwave cloud feedback over the Southern Ocean is given in
 \cref{sec:04:possible_reasons_high_ecs_cmip6}.
@@
 responses to aerosols over the different periods of the \nth{20} century
-\autocite{Dittus2020}.
+\autocite{Dittus2020}. An example for a \acs{CMIP}6 model with high \ac{ECS}
+and a new aerosol scheme is the \ac{HadGEM} version 3. The new aerosol scheme
+in this model suppresses a strong negative cloud feedback over the Southern
+Ocean \autocite{BodasSalcedo2019}, which is one of the most pristine aerosol
+regions in the world \autocite{Hamilton2014}, where aerosols are predominantly
+of natural origin (\eg{} marine emissions of gaseous dimethyl sulfide). Earlier
+\ac{HadGEM} versions show a large reduction in cloud droplet size with warming
+over the Southern Ocean, which produces a negative shortwave cloud feedback.
+However, in \acs{HadGEM}3, this negative feedback is completely suppressed or
+even positive due to a small increase of the cloud droplet size with warming,
+which eventually increases \ac{ECS} due to an increased net cloud feedback
+\autocite{BodasSalcedo2019}.
 
@@
     been replaced with clouds that predominantly consist of (supercooled)
-    liquid clouds \autocite{BodasSalcedo2019, Gettelman2019, Mulcahy2020,
-      Williams2020}. Thus, in the affected \acs{CMIP}6 models, this cloud phase
-    change due to warming is reduced by allowing for supercooled cloud liquid,
-    which substantially reduces the strong negative cloud phase change feedback
-    (orange arrows). This leads to an increase in the global net cloud feedback
+    liquid clouds \autocite{BodasSalcedo2019, Gettelman2019}. Thus, in the
+    affected \acs{CMIP}6 models, this cloud phase change due to warming is
+    reduced by allowing for supercooled cloud liquid, which substantially
+    reduces the strong negative cloud phase change feedback (orange arrows).
+    This leads to an increase in the global net cloud feedback
     \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS} in the corresponding
@@
 Southern Ocean in these earlier climate models are biased towards too large
-amounts of ice crystals and too little amounts of supercooled liquid water
-\autocite{BodasSalcedo2016, McCoy2015, McCoy2016, Tan2016}. For this reason,
-the microphysical representation of mixed-phase clouds has been substantially
-improved in many \acs{CMIP}6 models \autocite{BodasSalcedo2019, Gettelman2019,
-  Mulcahy2020, Williams2020}, leading for example to a better agreement of the
-simulated \ac{TOA} shortwave \ac{CRE} with corresponding observations
-\autocite{Bock2020}. As a consequence, the strong negative cloud feedback
-described above that results from a cloud phase change from ice clouds in the
-present-day to liquid clouds in the future is substantially reduced in those
-\acs{CMIP}6 models that simulate predominantly liquid clouds over the
-present-day Southern Ocean. This is illustrated in
-\cref{fig:04:cloud_phase_change_feedback} (orange arrows). Moreover, this
-reduction of the negative shortwave cloud phase change feedback increases the
-global net cloud feedback \autocite{BodasSalcedo2019, Tan2016} and eventually
-the \ac{ECS} of the affected \acs{CMIP}6 models \autocite{Andrews2019,
-  Gettelman2019}, offering a possible explanation for the high \ac{ECS} values in
-some \acs{CMIP}6 models.
+amounts of ice crystals and too little amounts of supercooled liquid water when
+compared to satellite observations \autocite{BodasSalcedo2016}. Thus, the size
+of the resulting strong negative cloud phase change feedback in these models
+has long been questioned \autocite{McCoy2015, Tan2016}. Improvements in the
+representation of mixed-phase clouds are known to reduce this long-standing
+cloud phase bias over the Southern Ocean \autocite{BodasSalcedo2016, McCoy2016}
+and to improve the representation of both cloud microphysical structure and
+cloud radiative impacts in this region \autocite{Hyder2018, Kay2016}. For this
+reason, the microphysical representation of mixed-phase clouds has been updated
+in many \acs{CMIP}6 models. For example, in \acs{CESM}2, a new mixed-phase ice
+nucleation scheme by \textcite{Hoose2010} replaces the \textcite{Meyers1992}
+empirical scheme used in predecessor models \autocite{Gettelman2019}.
+Simulations with the \textcite{Meyers1992} scheme show almost no supercooled
+liquid water in high-latitude clouds and a strong negative cloud phase feedback
+when ice turns to liquid. On the contrary, simulations with the new ice
+nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
+feedback and are more consistent with observations of the current climate
+\autocite{Gettelman2019}. Another example is the \acs{HadGEM}3 model, for which
+a new cloud scheme that accounts for turbulent production of liquid water in
+mixed-phase clouds has been implemented \autocite{BodasSalcedo2019}. The new
+scheme parameterizes the role of subgrid-scale turbulence in the production and
+maintenance of supercooled liquid water \autocite{Furtado2016}. As a
+consequence, the cloud liquid water content over the Southern ocean is
+increased, which is in closer agreement to observations compared to earlier
+\ac{HadGEM} versions. Moreover, the increase in the cloud liquid water content
+with warming is reduced in \acs{HadGEM}3, leading to a reduction in the
+negative cloud phase change feedback over the Southern Ocean
+\autocite{BodasSalcedo2019}. By evaluating short-term forecasts created with
+\acs{HadGEM}3, \textcite{Williams2020} additionally show that the new climate
+model version improves in traditional metrics used for numerical weather
+prediction, which indicates an improvement in the physical processes simulated
+by the climate model.
+
+Overall, these changes in many of the \acs{CMIP}6 models lead to a better
+agreement of the \acs{CMIP}6 \ac{MMM} of the \ac{TOA} shortwave \ac{CRE} with
+corresponding observations \autocite{Bock2020}. Moreover, the strong negative
+cloud feedback described above that results from a cloud phase change from ice
+clouds in the present-day to liquid clouds in the future is substantially
+reduced in those \acs{CMIP}6 models that simulate predominantly liquid clouds
+over the present-day Southern Ocean. This is illustrated in
+\cref{fig:04:cloud_phase_change_feedback} (orange arrows). As a consequence,
+this reduction of the negative shortwave cloud phase change feedback increases
+the global net cloud feedback \autocite{BodasSalcedo2019, Tan2016} and
+eventually the \ac{ECS} of the affected \acs{CMIP}6 models
+\autocite{Andrews2019, Gettelman2019}, offering a possible explanation for the
+high \ac{ECS} values found in many \acs{CMIP}6 models.
 
@@
 
-To further confirm this hypothesis, we show the global distributions of
-relevant cloud feedback parameters (expressed by the net, shortwave and
-longwave \ac{CRE} feedback parameters; see
-\cref{subsec:02:cloud_feedback_parameters}) for the \acs{CMIP}6 \ac{MMM} and
-its difference to the \acs{CMIP}5 \ac{MMM} in
+To further confirm this hypothesis, the global distributions of relevant cloud
+feedback parameters (expressed by the net, shortwave and longwave \ac{CRE}
+feedback parameters; see \cref{subsec:02:cloud_feedback_parameters}) are shown
+for the \acs{CMIP}6 \ac{MMM} and its difference to the \acs{CMIP}5 \ac{MMM} in
 \cref{fig:04:cloud_feedback_parameters}. First of all, the comparison of

### 2021-03-24 ch05_paper_ecs/paper_ecs.tex
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

@@
     \acs{CMIP}5 models. Blue circles show the 15 \acs{CMIP}5 models used in the
-    original publication (except for CESM1-CAM5). The solid blue line and blue
-    shaded area show the emergent relationships evaluated on these models
+    original publication (except for \acs{CESM}1-CAM5). The solid blue line and
+    blue shaded area show the emergent relationships evaluated on these models
     including the uncertainty range. In our study, we added eleven more
@@
 CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and
-MPI-ESM-P). Due to a lack of publicly available data, the model CESM1-CAM5 that
-is used in the original publication is not included in our analysis. The effect
-of choosing different subsets of \acs{CMIP}5 models on the emergent
-relationship is illustrated in \cref{fig:05:zha}. Using the original
-\acs{CMIP}5 models from the original publication gives a considerably higher
-correlation ($R^2 = 0.38$) than using all available \acs{CMIP}5 models ($R^2 =
-0.05$). This result shows a strong dependency of this emergent constraint on
-the subset of climate models used. Nonetheless, the performance on \acs{CMIP}6
-models is, surprisingly, the best of all the constraints and much better than
-on either subset of \acs{CMIP}5 models.
+MPI-ESM-P). Due to a lack of publicly available data, the model
+\acs{CESM}1-CAM5 that is used in the original publication is not included in
+our analysis. The effect of choosing different subsets of \acs{CMIP}5 models on
+the emergent relationship is illustrated in \cref{fig:05:zha}. Using the
+original \acs{CMIP}5 models from the original publication gives a considerably
+higher correlation ($R^2 = 0.38$) than using all available \acs{CMIP}5 models
+($R^2 = 0.05$). This result shows a strong dependency of this emergent
+constraint on the subset of climate models used. Nonetheless, the performance
+on \acs{CMIP}6 models is, surprisingly, the best of all the constraints and
+much better than on either subset of \acs{CMIP}5 models.
 

### 2021-03-24 ch07_conclusion/conclusion.tex
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

@@
 \acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
-\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
-climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
-representation of mixed-phase clouds over the Southern Ocean. This change leads
-to an improved simulation of the shortwave \ac{CRE} in these models when
-compared to observations. However, this change also substantially reduces the
-strong negative shortwave cloud feedback over the Southern Ocean that is
-present in previous \ac{CMIP} generations and that results from a cloud phase
-change from ice clouds in the present-day to liquid clouds in the future. In
-the affected \acs{CMIP}6 models, this cloud phase change due to warming is
+\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased
+climate sensitivity in many \acs{CMIP}6 models is the addition of prognostic
+aerosol schemes that include aerosol-cloud interactions which could have
+produced overly large negative radiative forcing in some models. To correctly
+reproduce the observed historical temperature trend, a stronger model response
+to \ac{GHG} forcing is required, resulting in a higher climate sensitivity. A
+further reason for the high climate sensitivity in \acs{CMIP}6 is a change in
+the microphysical representation of mixed-phase clouds over the Southern Ocean
+in some models. This change was mainly implemented to address a long-standing
+cloud phase bias over the Southern Ocean in the climate models, which has been
+successfully reduced. Moreover, the \acs{CMIP}6 \ac{MMM} shows an improved
+simulation of the shortwave \ac{CRE} when compared to observations. However,
+the change in the representation of mixed-phase clouds also substantially
+reduces the strong negative shortwave cloud feedback over the Southern Ocean
+that is present in previous \ac{CMIP} generations and that results from a cloud
+phase change from ice clouds in the present-day to liquid clouds in the future.
+In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
 reduced since these models simulate less cloud ice over the present-day

### 2021-03-24 end_abbreviations/abbreviations.tex
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables
Added details on carbon cycle models to chapter 02
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

--- 2021-03-02 end_acronyms/acronyms.tex
@@
 %=============================================================================%
-%                                Dissertation                                 %
+%                            Doctoral Dissertation                            %
 %                               Manuel Schlund                                %
-%                                  (c) 2020                                   %
+%                                  (c) 2021                                   %
 %=============================================================================%
-%                                  Acronyms                                   %
+%                                Abbreviations                                %
 %=============================================================================%
 
-\chapter{List of Acronyms}
+\chapter{List of Abbreviations}
 
@@
   \acro{1BGC}[1\%BGC]{biogeochemically-coupled simulation where the %
-    atmospheric CO\textsubscript{2} concentration increases by $1 \unit{\%}$ %
-    per year}
+    atmospheric \ch{CO2} concentration increases by $1 \unit{\%}$ per year}
   \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
@@
   \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
+  \acro{CH4}[\ch{CH4}]{methane}
   \acro{CMIP}{Coupled Model Intercomparison Project}
-  \acro{CH4}[CH\textsubscript{4}]{methane}
   \acro{CMOR}{Climate Model Output Rewriter}
   \acro{CO}{carbon monoxide}
-  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
+  \acro{CO2}[\ch{CO2}]{carbon dioxide}
   \acro{CRE}{cloud radiative effect}
@@
   \acro{GtC}{gigatonnes of carbon}
+  \acro{HadGEM}{Hadley Centre Global Environmental Model}
   \acro{IAM}{integrated assessment model}
@@
   \acro{MMM}{multi-model mean}
-  \acro{N2O}[N\textsubscript{2}O]{nitrous oxide}
+  \acro{N2O}[\ch{N2O}]{nitrous oxide}
   \acro{NAO}{North Atlantic Oscillation}
+  \acro{NBP}{net biome production}
   \acro{NCAR}{National Center for Atmospheric Research}
+  \acro{NEP}{net ecosystem production}
   \acro{netCDF}{Network Common Data Form}
+  \acro{NPP}{net primary production}
   \acro{PDF}{probability density function}
@@
   \acro{RSDS}{incoming solar radiation at surface}
+  \acro{RuBisCO}{ribulose-1,5-bisphosphate carboxylase/oxygenase}
   \acro{ScenarioMIP}{Scenario Model Intercomparison Project}

### 2021-03-24 end_appendix/appendix.tex
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

@@
     GISS-E2-R & 14 & \textcite{Schmidt2006} \\
-    HadGEM2-ES & 15 & \textcite{Collins2011} \\
+    \acs{HadGEM}2-ES & 15 & \textcite{Collins2011} \\
     IPSL-CM5A-LR & 16 & \textcite{Dufresne2013} \\
@@
     CAS-ESM2-0 & 35 & \textcite{Wang2020} \\
-    CESM2 & 36 & \textcite{Danabasoglu2020} \\
-    CESM2-FV2 & 37 & \textcite{Danabasoglu2020} \\
-    CESM2-WACCM & 38 & \textcite{Danabasoglu2020, Gettelman2019a} \\
-    CESM2-WACCM-FV2 & 39 & \textcite{Danabasoglu2020, Gettelman2019a} \\
+    \acs{CESM}2 & 36 & \textcite{Danabasoglu2020} \\
+    \acs{CESM}2-FV2 & 37 & \textcite{Danabasoglu2020} \\
+    \acs{CESM}2-WACCM & 38 & \textcite{Danabasoglu2020, Gettelman2019a} \\
+    \acs{CESM}2-WACCM-FV2 & 39 & \textcite{Danabasoglu2020, Gettelman2019a} \\
     CMCC-CM2-SR5 & 40 & \textcite{Cherchi2019} \\
@@
     GISS-E2-1-H & 50 & \textcite{Rind2020} \\
-    HadGEM3-GC31-LL & 51 & \textcite{Kuhlbrodt2018} \\
-    HadGEM3-GC31-MM & 52 & \textcite{Williams2018} \\
+    \acs{HadGEM}3-GC31-LL & 51 & \textcite{Kuhlbrodt2018} \\
+    \acs{HadGEM}3-GC31-MM & 52 & \textcite{Williams2018} \\
     INM-CM4-8 & 53 & \textcite{Volodin2017,Volodin2017a} \\
@@
     \midrule
-    CESM1-BGC & CLM4 & \textcite{Gent2011} \\
+    \acs{CESM}1-BGC & CLM4 & \textcite{Gent2011} \\
     CanESM2 & CLASS2.7 + CTEM1 & \textcite{Arora2011} \\
     GFDL-ESM2M & LM3 & \textcite{Dunne2012} \\
-    HadGEM2-ES & JULES + TRIFFID & \textcite{Collins2011} \\
+    \acs{HadGEM}2-ES & JULES + TRIFFID & \textcite{Collins2011} \\
     MIROC-ESM & MATSIRO + SEIB-DGVM & \textcite{Watanabe2011} \\
@@
     and emission-driven \acs{RCP}8.5 simulations for the years
-    \range{2006}{2019}; the latter is not available for HadGEM2-ES). The thick
-    black line shows the observations. For the \acs{CMIP}5 models, the grid
-    cell closest to \acs{KUM} is considered. The curves show an increase of the
-    atmospheric \acs{CO2} concentration superimposed by a pronounced seasonal
-    cycle (see \cref{subsec:02:carbon_cycle_perturbations}). (b) Annual
-    amplitude of the seasonal cycle of \acs{CO2} (defined as the difference
-    between the maximum and the minimum monthly mean atmospheric \acs{CO2}
-    concentration for each year) against the annual mean atmospheric \acs{CO2}
-    concentration at \acs{KUM}. Colored points show the \acs{CMIP}5 models
-    (similar time ranges as in (a)), and thick black points show the
+    \range{2006}{2019}; the latter is not available for \acs{HadGEM}2-ES). The
+    thick black line shows the observations. For the \acs{CMIP}5 models, the
+    grid cell closest to \acs{KUM} is considered. The curves show an increase
+    of the atmospheric \acs{CO2} concentration superimposed by a pronounced
+    seasonal cycle (see \cref{subsec:02:carbon_cycle_perturbations}). (b)
+    Annual amplitude of the seasonal cycle of \acs{CO2} (defined as the
+    difference between the maximum and the minimum monthly mean atmospheric
+    \acs{CO2} concentration for each year) against the annual mean atmospheric
+    \acs{CO2} concentration at \acs{KUM}. Colored points show the \acs{CMIP}5
+    models (similar time ranges as in (a)), and thick black points show the
     observations. The lines show the corresponding linear regression fits for

### 2021-03-24 02_abstract/abstract.tex
Further improvements of text in chapter 04

--- 2021-03-22 02_abstract/abstract.tex
@@
 again higher than the likely range assessed in \acs{AR}5 of
-$\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased climate
-sensitivity in many \acs{CMIP}6 models is a change in the microphysical
-representation of mixed-phase clouds over the Southern Ocean, which reduces the
-strong negative shortwave cloud phase change feedback in this region that is
-present in climate models from previous \acs{CMIP} generations.
+$\rangeunit{1.0}{2.5}{K}$. Possible reasons for the increased climate
+sensitivity in many \acs{CMIP}6 models are the addition of new prognostic
+aerosol schemes that include aerosol-cloud interactions and changes in the
+microphysical representation of mixed-phase clouds over the Southern Ocean,
+which reduce the strong negative shortwave cloud phase change feedback in this
+region that is present in climate models from previous \acs{CMIP} generations.
 

### 2021-03-24 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Further improvements of text in chapter 04

@@
 responses to aerosols over the different periods of the \nth{20} century
-\autocite{Dittus2020}. An example for a \acs{CMIP}6 model with high \ac{ECS}
-and a new aerosol scheme is the \ac{HadGEM} version 3. The new aerosol scheme
-in this model suppresses a strong negative cloud feedback over the Southern
-Ocean \autocite{BodasSalcedo2019}, which is one of the most pristine aerosol
-regions in the world \autocite{Hamilton2014}, where aerosols are predominantly
-of natural origin (\eg{} marine emissions of gaseous dimethyl sulfide). Earlier
-\ac{HadGEM} versions show a large reduction in cloud droplet size with warming
-over the Southern Ocean, which produces a negative shortwave cloud feedback.
-However, in \acs{HadGEM}3, this negative feedback is completely suppressed or
-even positive due to a small increase of the cloud droplet size with warming,
-which eventually increases \ac{ECS} due to an increased net cloud feedback
-\autocite{BodasSalcedo2019}.
+\autocite{Dittus2020}. A further type of aerosol-cloud interactions is for
+example present in the \ac{HadGEM} version 3. The new aerosol scheme in this
+model suppresses a strong negative cloud feedback over the Southern Ocean that
+is found in the model's predecessor versions \autocite{BodasSalcedo2019}. The
+Southern Ocean is one of the most pristine aerosol regions in the world
+\autocite{Hamilton2014}, where aerosols predominantly originate from natural
+sources like marine emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM}
+versions show a large reduction in cloud droplet size with warming over the
+Southern Ocean, which increases the cloud albedo and thus produces a negative
+shortwave cloud feedback. On the contrary, in \acs{HadGEM}3, this negative
+feedback is completely suppressed or even positive due to a small increase of
+the cloud droplet size with warming, which eventually increases \ac{ECS} due to
+an increased net cloud feedback \autocite{BodasSalcedo2019}.
 
@@
 maintenance of supercooled liquid water \autocite{Furtado2016}. As a
-consequence, the cloud liquid water content over the Southern ocean is
+consequence, the cloud liquid water content over the Southern Ocean is
 increased, which is in closer agreement to observations compared to earlier
@@
 negative cloud phase change feedback over the Southern Ocean
-\autocite{BodasSalcedo2019}. By evaluating short-term forecasts created with
-\acs{HadGEM}3, \textcite{Williams2020} additionally show that the new climate
-model version improves in traditional metrics used for numerical weather
-prediction, which indicates an improvement in the physical processes simulated
-by the climate model.
+\autocite{BodasSalcedo2019}. By evaluating traditional metrics used for
+numerical weather prediction, \textcite{Williams2020} additionally demonstrate
+that short-term forecasts created with \acs{HadGEM}3 show a higher performance
+than comparable simulations of predecessor model versions, which indicates an
+improvement of the physical processes in the new climate model.
 

### 2021-03-24 ch07_conclusion/conclusion.tex
Further improvements of text in chapter 04

@@
 climate sensitivity in many \acs{CMIP}6 models is the addition of prognostic
-aerosol schemes that include aerosol-cloud interactions which could have
-produced overly large negative radiative forcing in some models. To correctly
-reproduce the observed historical temperature trend, a stronger model response
-to \ac{GHG} forcing is required, resulting in a higher climate sensitivity. A
-further reason for the high climate sensitivity in \acs{CMIP}6 is a change in
+aerosol schemes that include aerosol-cloud interactions, which could have
+produced overly large negative radiative forcing. To correctly reproduce the
+observed historical temperature trend, a stronger response to \ac{GHG} forcing
+is required, resulting in a higher climate sensitivity in the corresponding
+models. A further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in
 the microphysical representation of mixed-phase clouds over the Southern Ocean
-in some models. This change was mainly implemented to address a long-standing
-cloud phase bias over the Southern Ocean in the climate models, which has been
-successfully reduced. Moreover, the \acs{CMIP}6 \ac{MMM} shows an improved
-simulation of the shortwave \ac{CRE} when compared to observations. However,
-the change in the representation of mixed-phase clouds also substantially
-reduces the strong negative shortwave cloud feedback over the Southern Ocean
-that is present in previous \ac{CMIP} generations and that results from a cloud
-phase change from ice clouds in the present-day to liquid clouds in the future.
-In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
-reduced since these models simulate less cloud ice over the present-day
-Southern Ocean than their predecessor versions with no supercooled cloud
-liquid.
+in some models. This change was mainly implemented to reduce a long-standing
+cloud phase bias over the Southern Ocean that is present in models from
+previous \ac{CMIP} generations. Moreover, the \acs{CMIP}6 \ac{MMM} shows an
+improved simulation of the shortwave \ac{CRE} when compared to observations.
+However, the change in the representation of mixed-phase clouds also
+substantially reduces the strong negative shortwave cloud feedback over the
+Southern Ocean that is present in previous \ac{CMIP} generations and that
+results from a cloud phase change from ice clouds in the present-day to liquid
+clouds in the future. In the affected \acs{CMIP}6 models, this cloud phase
+change due to warming is reduced since these models simulate less cloud ice
+over the present-day Southern Ocean than their predecessor versions with no
+supercooled cloud liquid.
 

### 2021-03-25 ch02_scientific_background/scientific_background.tex
Further improvements of text

@@
 
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch02_scientific_background/figs/feedbacks_overview.pdf}
+  \caption[
+    Climate feedbacks and corresponding time scales.
+  ]{
+    Climate feedbacks and corresponding time scales. The different sings refer
+    to the signs of the different feedbacks. Positive feedbacks amplify the
+    effect of the external forcing (\eg{} the water vapor feedback) and
+    negative feedbacks diminish the effect of the external forcing (\eg{} the
+    longwave radiation feedback). An example for a feedback that might be
+    either positive or negative is the cloud feedback. The smaller box
+    highlights the large differences in time scales for the various feedbacks.
+    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.
+  }
+  \label{fig:02:feedbacks_overview}
+\end{figure}
+
 As already described in \cref{subsec:02:source_of_uncertainties}, the effects
@@
 darker and more absorbing surfaces below \autocite{Cubasch2013}.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
-    ch02_scientific_background/figs/feedbacks_overview.pdf}
-  \caption[
-    Climate feedbacks and corresponding time scales.
-  ]{
-    Climate feedbacks and corresponding time scales. \enquote{+} refers to
-    positive feedbacks, which amplify the effect of the external forcing (\eg{}
-    the water vapor feedback). \enquote{-} refers to negative feedbacks, which
-    diminish the effect of the external forcing (\eg{} the longwave radiation
-    feedback). \enquote{+/-} refers to feedbacks which might be either positive
-    or negative (\eg{} the cloud feedback). The smaller box highlights the
-    large differences in time scales for the various feedbacks.
-    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.
-  }
-  \label{fig:02:feedbacks_overview}
-\end{figure}
 
@@
   \centering
-  \includegraphics[width=\FigureWidth{}]{
+  \includegraphics[width=\EvenLargerFigureWidth{}]{
     ch02_scientific_background/figs/keeling_curve.pdf}
@@
 
+Similar to many cloud processes, most carbon cycle--related processes need to
+be parameterized in \acp{ESM} since they occur on spatial scales much smaller
+than the typical grid cell sizes (see
+\cref{subsec:02:numerical_climate_modeling}). For the terrestrial carbon cycle,
+the major carbon exchange processes that need to be considered are
+photosynthesis, respiration and disturbances, which connect the atmospheric
+carbon reservoir and the land carbon reservoir. The latter mainly consists of
+the vegetation biomass, litter and soil. The terrestrial carbon exchange
+processes are commonly described with \ac{GPP}, \ac{NPP}, \ac{NEP} and
+\ac{NBP}, which are all illustrated in \cref{fig:02:carbon_cycle_processes}.
+\ac{GPP} describes the total carbon uptake of the vegetation by photosynthesis,
+while \ac{NPP} refers to the net carbon gain of the plants after accounting for
+autotrophic respiration ($\Ra{}$). \Ac{NEP} additionally considers
+heterotrophic respiration ($\Rh{}$) and describes the carbon uptake/release of
+the entire ecosystem. Finally, \ac{NBP} describes changes in the long-term
+carbon storage of the terrestrial biosphere by including disturbances like
+fires ($\Ld{}$).
+
+The different land carbon cycle models integrated into modern \acp{ESM} use a
+variety of different parameterizations for the different processes
+\autocite{Arora2020}. Examples for the photosynthesis (\ac{GPP}) are given by
+\textcite{Farquhar1980} for \ch{C3} plants and \textcite{Collatz1992} for
+\ch{C4} plants. The terms \enquote{\ch{C3}} and \enquote{\ch{C4}} refer to the
+different metabolic pathways used in the carbon fixation during photosynthesis.
+In both cases, the enzyme \ac{RuBisCO} is an important catalyst for the
+corresponding chemical reactions. Both parameterizations utilize variables that
+are known to affect photosynthesis, such as the availability of light, the
+\ac{CO2} concentration, the soil moisture and the temperature.
+
 \begin{figure}[t]
@@
 \end{figure}
-
-Similar to many cloud processes, most carbon cycle--related processes need to
-be parameterized in \acp{ESM} since they occur on spatial scales much smaller
-than the typical grid cell sizes (see
-\cref{subsec:02:numerical_climate_modeling}). For the terrestrial carbon cycle,
-the major carbon exchange processes that need to be considered are
-photosynthesis, respiration and disturbances, which connect the atmospheric
-carbon reservoir and the land carbon reservoir. The latter mainly consists of
-the vegetation biomass, litter and soil. The terrestrial carbon exchange
-processes are commonly described with \ac{GPP}, \ac{NPP}, \ac{NEP} and
-\ac{NBP}, which are all illustrated in \cref{fig:02:carbon_cycle_processes}.
-\ac{GPP} describes the total carbon uptake of the vegetation by photosynthesis,
-while \ac{NPP} refers to the net carbon gain of the plants after accounting for
-autotrophic respiration ($\Ra{}$). \Ac{NEP} additionally considers
-heterotrophic respiration ($\Rh{}$) and describes the carbon uptake/release of
-the entire ecosystem. Finally, \ac{NBP} describes changes in the long-term
-carbon storage of the terrestrial biosphere by including disturbances like
-fires ($\Ld{}$).
-
-The different land carbon cycle models integrated into modern \acp{ESM} use a
-variety of different parameterizations for the different processes
-\autocite{Arora2020}. Examples for the photosynthesis (\ac{GPP}) are given by
-\textcite{Farquhar1980} for \ch{C3} plants and \textcite{Collatz1992} for
-\ch{C4} plants. The terms \enquote{\ch{C3}} and \enquote{\ch{C4}} refer to the
-different metabolic pathways used in the carbon fixation during photosynthesis.
-In both cases, the enzyme \ac{RuBisCO} is an important catalyst for the
-corresponding chemical reactions. Both parameterizations utilize variables that
-are known to affect photosynthesis, such as the availability of light, the
-\ac{CO2} concentration, the soil moisture and the temperature.
 
@@
 target variable. Since cloud feedbacks are a major source of uncertainty for
-climate sensitivity, a variety of papers focus on constraining \ac{ECS} with
-cloud-related processes \autocite{Brient2015, Brient2016, Fasullo2012,
-  Lipat2017, Sherwood2014, Su2014, Tian2015, Volodin2008, Qu2013, Zhai2015},
-which are discussed in detail in
+climate sensitivity, a variety of scientific papers focus on constraining
+\ac{ECS} with cloud-related processes \autocite{Brient2015, Brient2016,
+  Fasullo2012, Lipat2017, Sherwood2014, Su2014, Tian2015, Volodin2008, Qu2013,
+  Zhai2015}, which are discussed in detail in
 \cref{sec:05:comparison_of_emergent_constraints}. More recent studies aim to

### 2021-03-25 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Further improvements of text

@@
 responses to aerosols over the different periods of the \nth{20} century
-\autocite{Dittus2020}. A further type of aerosol-cloud interactions is for
-example present in the \ac{HadGEM} version 3. The new aerosol scheme in this
-model suppresses a strong negative cloud feedback over the Southern Ocean that
-is found in the model's predecessor versions \autocite{BodasSalcedo2019}. The
-Southern Ocean is one of the most pristine aerosol regions in the world
-\autocite{Hamilton2014}, where aerosols predominantly originate from natural
-sources like marine emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM}
-versions show a large reduction in cloud droplet size with warming over the
-Southern Ocean, which increases the cloud albedo and thus produces a negative
-shortwave cloud feedback. On the contrary, in \acs{HadGEM}3, this negative
-feedback is completely suppressed or even positive due to a small increase of
-the cloud droplet size with warming, which eventually increases \ac{ECS} due to
-an increased net cloud feedback \autocite{BodasSalcedo2019}.
+\autocite{Dittus2020}. Further aerosol-cloud interactions that are relevant in
+the context of high climate sensitivity in \acs{CMIP}6 are found for example in
+the \ac{HadGEM} version 3. The new aerosol scheme in this model suppresses a
+strong negative cloud feedback over the Southern Ocean that is found in the
+model's predecessor versions \autocite{BodasSalcedo2019}. The Southern Ocean is
+one of the most pristine aerosol regions in the world \autocite{Hamilton2014},
+where aerosols predominantly originate from natural sources like marine
+emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM} versions show a
+large reduction in cloud droplet size with warming over the Southern Ocean,
+which increases the cloud albedo and thus produces a negative shortwave cloud
+feedback. On the contrary, in \acs{HadGEM}3, this negative feedback is
+completely suppressed or even positive due to a small increase of the cloud
+droplet size with warming, which eventually increases \ac{ECS} due to an
+increased net cloud feedback \autocite{BodasSalcedo2019}.
 
@@
 nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
-feedback and are more consistent with observations of the current climate
-\autocite{Gettelman2019}. Another example is the \acs{HadGEM}3 model, for which
-a new cloud scheme that accounts for turbulent production of liquid water in
-mixed-phase clouds has been implemented \autocite{BodasSalcedo2019}. The new
-scheme parameterizes the role of subgrid-scale turbulence in the production and
-maintenance of supercooled liquid water \autocite{Furtado2016}. As a
-consequence, the cloud liquid water content over the Southern Ocean is
-increased, which is in closer agreement to observations compared to earlier
-\ac{HadGEM} versions. Moreover, the increase in the cloud liquid water content
-with warming is reduced in \acs{HadGEM}3, leading to a reduction in the
-negative cloud phase change feedback over the Southern Ocean
-\autocite{BodasSalcedo2019}. By evaluating traditional metrics used for
-numerical weather prediction, \textcite{Williams2020} additionally demonstrate
-that short-term forecasts created with \acs{HadGEM}3 show a higher performance
-than comparable simulations of predecessor model versions, which indicates an
-improvement of the physical processes in the new climate model.
+feedback. Since the new scheme also reduces significant cloud phase biases over
+the Southern Ocean that are present in the old scheme, the reduced negative
+cloud phase feedback is thought to be more consistent with observations of the
+current climate \autocite{Gettelman2019}. Another example is the \acs{HadGEM}3
+model, for which a new cloud scheme that accounts for turbulent production of
+liquid water in mixed-phase clouds has been implemented
+\autocite{BodasSalcedo2019}. The new scheme parameterizes the role of
+subgrid-scale turbulence in the production and maintenance of supercooled
+liquid water \autocite{Furtado2016}. As a consequence, the cloud liquid water
+path over the Southern Ocean is increased, which is more in line with
+observations from microwave satellite retrievals. Moreover, the increase in the
+cloud liquid water path with warming is reduced in \acs{HadGEM}3, leading to a
+reduction in the negative cloud phase change feedback over the Southern Ocean
+\autocite{BodasSalcedo2019}. In addition, \textcite{Williams2020} show that
+short-term forecasts created with \acs{HadGEM}3 are more accurate than
+comparable simulations from predecessor model versions (using traditional
+metrics from numerical weather prediction), which indicates an improvement of
+the physical processes in the new climate model.
 

### 2021-03-25 ch05_paper_ecs/paper_ecs.tex
Further improvements of text

@@
   \caption[
-    Overview of the constrained \acf{ECS} ranges given by the eleven analyzed
-    emergent constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensemble.
+    Overview of the constrained \acf{ECS} ranges given by the eleven emergent
+    constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensemble.
   ]{
@@
 
-This paper assesses eleven different emergent constraints on \ac{ECS}, of which
-most are directly or indirectly related to cloud feedbacks, by applying them to
-results from \acp{ESM} contributing to \acs{CMIP}5 and \acs{CMIP}6. Of
-particular interest are the results from \acs{CMIP}6, since all analyzed
-emergent constraints have been published prior to the availability of
-\acs{CMIP}6 data. In summary, the best estimate of \ac{ECS} averaged over all
-emergent constraints increases by $12 \unit{\%}$ when moving from relationships
-trained on \acs{CMIP}5 to those trained on \acs{CMIP}6. Some increase is
-predicted by almost every constraint we analyze (with TII as the only
-exception) and can be at least partly explained by the increased multi-model
-mean \ac{ECS} of \acs{CMIP}6, which has not been accompanied by systematic
-changes in the constraint variables that could explain this increase, leading
-to regression fits with higher intercept values at observed constraint values.
-This is also illustrated by the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM} in the
-left columns of
+In this \namecref{ch:05:paper_ecs}, eleven different emergent constraints on
+\ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,
+are assessed by applying them to results from \acp{ESM} contributing to
+\acs{CMIP}5 and \acs{CMIP}6. Of particular interest are the results from
+\acs{CMIP}6, since all analyzed emergent constraints have been published prior
+to the availability of \acs{CMIP}6 data. In summary, the best estimate of
+\ac{ECS} averaged over all emergent constraints increases by $12 \unit{\%}$
+when moving from relationships trained on \acs{CMIP}5 to those trained on
+\acs{CMIP}6. Some increase is predicted by almost every constraint we analyze
+(with TII as the only exception) and can be at least partly explained by the
+increased multi-model mean \ac{ECS} of \acs{CMIP}6, which has not been
+accompanied by systematic changes in the constraint variables that could
+explain this increase, leading to regression fits with higher intercept values
+at observed constraint values. This is also illustrated by the \acs{CMIP}5 and
+\acs{CMIP}6 \acp{MMM} in the left columns of
 \cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}

### 2021-03-25 ch06_paper_gpp/paper_gpp.tex
Further improvements of text

@@
 
-In this paper, we develop a two-step approach to constrain the projected
-\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
-\acs{RCP}8.5 scenario. In the first step, we constrain the global mean \ac{GPP}
-to $\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint
-approach which we assume can also be applied to the global mean \ac{GPP} at the
-end of the \nth{21} century in the \acs{RCP}8.5 scenario (see discussion in
-\cref{subsec:06:step_1}). This step corresponds to the correction of the
-\acp{ESM}' biases in the response of future \ac{GPP} to rising atmospheric
-\ac{CO2} concentration, which is the main source of uncertainty in future
-\ac{GPP} projections \autocite{Arora2013, Haverd2020, Rogers2017}. As already
-discussed in \cref{subsec:06:step_1}, due to other than \ac{CO2} fertilization
-drivers of the increase in the \ac{CO2} seasonal cycle amplitude
-\autocite{Bastos2019, Forkel2016, Piao2018, Zhao2016} this emergent constraint
-is not undisputed and can be replaced with another emergent constraint in
-future studies. In the second step, a \ac{ML} approach is used to further
-constrain the gridded \ac{GPP} based on process-based present-day predictors.
-We consider two target variables: first (step 2a) the gridded monthly
-climatologies of absolute \ac{GPP} (\range{2091}{2100}) and second (step 2b)
-the gridded fractional \ac{GPP} change over the \nth{21} century (years
-\range{2091}{2100} versus years \range{1991}{2000}; see \cref{eq:06:f}). Both
-approaches give consistent results (see
+In this \namecref{ch:06:paper_gpp}, a two-step approach to constrain the
+projected \ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
+the \acs{RCP}8.5 scenario is developed. In the first step, we constrain the
+global mean \ac{GPP} to $\pmrangeunit{171}{12}{\GtCyr{}}$ using a published
+emergent constraint approach which we assume can also be applied to the global
+mean \ac{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario
+(see discussion in \cref{subsec:06:step_1}). This step corresponds to the
+correction of the \acp{ESM}' biases in the response of future \ac{GPP} to
+rising atmospheric \ac{CO2} concentration, which is the main source of
+uncertainty in future \ac{GPP} projections \autocite{Arora2013, Haverd2020,
+  Rogers2017}. As already discussed in \cref{subsec:06:step_1}, due to other
+than \ac{CO2} fertilization drivers of the increase in the \ac{CO2} seasonal
+cycle amplitude \autocite{Bastos2019, Forkel2016, Piao2018, Zhao2016} this
+emergent constraint is not undisputed and can be replaced with another emergent
+constraint in future studies. In the second step, a \ac{ML} approach is used to
+further constrain the gridded \ac{GPP} based on process-based present-day
+predictors. We consider two target variables: first (step 2a) the gridded
+monthly climatologies of absolute \ac{GPP} (\range{2091}{2100}) and second
+(step 2b) the gridded fractional \ac{GPP} change over the \nth{21} century
+(years \range{2091}{2100} versus years \range{1991}{2000}; see \cref{eq:06:f}).
+Both approaches give consistent results (see
 \cref{fig:06:comparison_step2a_step2b}). The latter quantity shows a higher

### 2021-03-25 ch07_conclusion/conclusion.tex
Further improvements of text

@@
 light of the large spread of climate sensitivity in the most recent generation
-of climate models from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical
-evaluation and refinement of the output of multi-model climate projections is
-as relevant as ever. This thesis quantifies the associated uncertainties,
-presents the evaluation of established methods to reduce these uncertainties in
-the new climate model ensemble and finally introduces a novel alternative
-technique based on supervised \ac{ML}.
+of climate models from \acs{CMIP}6, that has been demonstrated in
+\cref{ch:04:papers_ecs_tcr_assessment}, a careful statistical evaluation and
+refinement of the output of multi-model climate projections is as relevant as
+ever. This thesis quantifies the associated uncertainties, presents the
+evaluation of established methods to reduce these uncertainties in the new
+climate model ensemble and finally introduces a novel alternative technique
+based on supervised \ac{ML}.
 
@@
 \acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
-a published emergent constraint by \textcite{Wenzel2016}. This first step
-corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2}, which shows a
-range of $\rangeunit{156}{247}{\GtCyr{}}$ in the raw multi-model ensemble. In
-the second step, a \ac{ML}-based climate model weighting approach is used to
-further constrain the gridded \ac{GPP} based on present-day predictors that are
-relevant for the simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach
-is mathematically similar to the \ac{MDER} approach \autocite{Karpechko2013,
-  Senftleben2020, Wenzel2016a}, but additionally considers multi-dimensional
-(gridded) target variables and non-linear relationships between the predictors
-and the target variable. A a relationship between process‐oriented predictors
-and future projections of \ac{GPP} is established and then utilized to project
-today's observed conditions into the future. The prediction phase of the new
-method can be interpreted as an implicit performance weighting. However, due to
-the complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not
-possible to extract specific values for the individual weights. Two target
-variables are considered: the gridded monthly climatologies of absolute
-\ac{GPP} (\range{2091}{2100}) and the gridded fractional \ac{GPP} change over
-the \nth{21} century (2100 versus 2000). The latter quantity shows an increased
+a published emergent constraint by \textcite{Wenzel2016}. In contrast to that,
+the corresponding multi-model range of \ac{GPP} in the raw \acs{CMIP}5 ensemble
+is $\rangeunit{156}{247}{\GtCyr{}}$. This first step corrects the \acs{CMIP}5
+models' \ac{GPP} response to \ac{CO2}, which is primarily driven by the
+\ac{CO2} fertilization. In the second step, a \ac{ML}-based climate model
+weighting approach is used to further constrain the gridded \ac{GPP} based on
+present-day predictors that are relevant for the simulation of \ac{GPP} in the
+\acp{ESM}. The \ac{ML} approach is mathematically similar to the \ac{MDER}
+approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}, but
+additionally considers multi-dimensional (gridded) target variables and
+non-linear relationships between the predictors and the target variable. A a
+relationship between process‐oriented predictors and future projections of
+\ac{GPP} is established and then utilized to project today's observed
+conditions into the future. The prediction phase of the new method can be
+interpreted as an implicit performance weighting. However, due to the complex
+structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
+extract specific values for the individual weights. Two target variables are
+considered: the gridded monthly climatologies of absolute \ac{GPP}
+(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
+\nth{21} century (2100 versus 2000). The latter quantity shows an increased
 \acs{GPP} change in the high latitudes compared to regions closer to the
@@
 new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
-and \ac{LAI} are the dominant features. This study directly addresses
-\cref{enum:01:question_3} (\enquote{\emph{\KeyScienceQuestionThree{}}}), which
-can be confidently answered with \enquote{yes} based on the results found.
+and \ac{LAI} are the most important features. A corresponding feature map shows
+that \ac{T} dominates in the tropics and \ac{LAI} dominates in the northern
+high latitudes, which might be attributed to an additional greening trend in
+the high latitudes that enhances the \ac{GPP} increase in this regions compared
+to regions closer to the equator through an extension of the growing season.
+This study directly addresses \cref{enum:01:question_3}
+(\enquote{\emph{\KeyScienceQuestionThree{}}}), which can be confidently
+answered with \enquote{yes} based on the results found.
 

### 2021-03-25 utils/add_headers.py
Updated license information in every file

@@ -1,0 +1,56 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+# This file is used to produce materials for the following work:
+
+# Doctoral Dissertation
+# (c) by Manuel Schlund
+
+# This work is licensed under a
+# Creative Commons Attribution 4.0 International License.
+
+# You should have received a copy of the license along with this
+# work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
+
+from pathlib import Path
+
+
+ROOT = Path(__file__).parents[1].resolve()
+
+
+def add_header(filename, header):
+    """Add header to a single file."""
+    with open(header, 'r') as f_in:
+        header_str = f_in.read()
+    with open(filename, 'r+') as f_out:
+        file_contents = f_out.read()
+        if file_contents.startswith(header_str):
+            print(f"    Skipping {filename} which alread contains header "
+                  f"{header}")
+            return
+        f_out.seek(0, 0)
+        f_out.write(header_str)
+        f_out.write(file_contents)
+    print(f"    Wrote header {header} to file {filename}")
+
+
+def add_headers(pattern, header):
+    """Add header to all files that match a pattern."""
+    print(f"Writing header {header} to all files matching {ROOT / pattern}")
+    files = ROOT.rglob(pattern)
+    for filename in files:
+        add_header(filename, header)
+    print()
+
+
+def main():
+    """Main function."""
+    print(f"Using root path {ROOT}")
+    print()
+    add_headers('*.py', 'HEADER_PY')
+    add_headers('*.tex', 'HEADER_TEX')
+    add_headers('*.yml', 'HEADER_YML')
+
+
+if __name__ == '__main__':
+    main()

### 2021-03-26 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Updated license information in every file
Further improved text on reasons for high ECS in CMIP6

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                        Papers ECS and TCR assessment                        %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
     feedback over the Southern Ocean, which is present in earlier \acs{CMIP}
-    model generations like \acs{CMIP}5 (green arrows). Due to substantial
-    improvements in the microphysical representation of mixed-phase clouds in
-    many \acs{CMIP}6 models, the predominantly ice clouds over the Southern
-    Ocean in the present-day climate of previous climate model generations have
-    been replaced with clouds that predominantly consist of (supercooled)
-    liquid clouds \autocite{BodasSalcedo2019, Gettelman2019}. Thus, in the
-    affected \acs{CMIP}6 models, this cloud phase change due to warming is
-    reduced by allowing for supercooled cloud liquid, which substantially
-    reduces the strong negative cloud phase change feedback (orange arrows).
-    This leads to an increase in the global net cloud feedback
-    \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS} in the corresponding
-    \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019}, which is a
-    possible explanation for the high \acs{ECS} values found in many
-    \acs{CMIP}6 models.
+    model generations like \acs{CMIP}5 (green arrows). Due to improvements in
+    the microphysical representation of mixed-phase clouds, some \acs{CMIP}6
+    models simulate a higher fraction of (supercooled) liquid water in the
+    clouds over the present-day Southern Ocean than their corresponding
+    predecessor models \autocite{BodasSalcedo2019, Gettelman2019}. Thus, in the
+    affected \acs{CMIP}6 models, the strong negative cloud feedback that
+    results from a cloud phase change from ice clouds in the present-day to
+    liquid clouds in the future is substantially reduced (orange arrows). This
+    leads to an increase in the global net cloud feedback
+    \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS}
+    \autocite{Andrews2019, Gettelman2019}, which is a possible explanation for
+    the high \acs{ECS} values found in some \acs{CMIP}6 models.
   }
@@
 reason, the microphysical representation of mixed-phase clouds has been updated
-in many \acs{CMIP}6 models. For example, in \acs{CESM}2, a new mixed-phase ice
+in some \acs{CMIP}6 models. For example, in \acs{CESM}2, a new mixed-phase ice
 nucleation scheme by \textcite{Hoose2010} replaces the \textcite{Meyers1992}
@@
 Simulations with the \textcite{Meyers1992} scheme show almost no supercooled
-liquid water in high-latitude clouds and a strong negative cloud phase feedback
-when ice turns to liquid. On the contrary, simulations with the new ice
-nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
+liquid water in high-latitude clouds and a strong negative cloud phase change
+feedback when ice turns to liquid. On the contrary, simulations with the new
+ice nucleation scheme by \textcite{Hoose2010} show a strong reduction in this
 feedback. Since the new scheme also reduces significant cloud phase biases over
 the Southern Ocean that are present in the old scheme, the reduced negative
-cloud phase feedback is thought to be more consistent with observations of the
-current climate \autocite{Gettelman2019}. Another example is the \acs{HadGEM}3
-model, for which a new cloud scheme that accounts for turbulent production of
-liquid water in mixed-phase clouds has been implemented
+cloud phase change feedback is thought to be more consistent with observations
+of the current climate \autocite{Gettelman2019}. Another example is the
+\acs{HadGEM}3 model, for which a new cloud scheme that accounts for turbulent
+production of liquid water in mixed-phase clouds has been implemented
 \autocite{BodasSalcedo2019}. The new scheme parameterizes the role of
@@
 
-Overall, these changes in many of the \acs{CMIP}6 models lead to a better
+Overall, such changes in the \acs{CMIP}6 models contribute to a better
 agreement of the \acs{CMIP}6 \ac{MMM} of the \ac{TOA} shortwave \ac{CRE} with
-corresponding observations \autocite{Bock2020}. Moreover, the strong negative
-cloud feedback described above that results from a cloud phase change from ice
-clouds in the present-day to liquid clouds in the future is substantially
-reduced in those \acs{CMIP}6 models that simulate predominantly liquid clouds
-over the present-day Southern Ocean. This is illustrated in
-\cref{fig:04:cloud_phase_change_feedback} (orange arrows). As a consequence,
-this reduction of the negative shortwave cloud phase change feedback increases
-the global net cloud feedback \autocite{BodasSalcedo2019, Tan2016} and
-eventually the \ac{ECS} of the affected \acs{CMIP}6 models
-\autocite{Andrews2019, Gettelman2019}, offering a possible explanation for the
-high \ac{ECS} values found in many \acs{CMIP}6 models.
+observations resulting in higher average pattern correlations of the model
+output with satellite observations \autocite{Bock2020}. Moreover, the strong
+negative cloud feedback described above that results from a cloud phase change
+from ice clouds in the present-day to liquid clouds in the future is
+substantially reduced in those \acs{CMIP}6 models that simulate a higher
+fraction of (supercooled) liquid water in the clouds over the present-day
+Southern Ocean than their corresponding \acs{CMIP}5 predecessors. This is
+illustrated in \cref{fig:04:cloud_phase_change_feedback} (orange arrows). The
+reduction of the negative shortwave cloud phase change feedback over the
+Southern Ocean increases the global net cloud feedback
+\autocite{BodasSalcedo2019, Tan2016} and eventually the \ac{ECS} of the
+affected \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019}, offering a
+possible explanation for the high \ac{ECS} values found in some \acs{CMIP}6
+models.
 

### 2021-03-26 ch07_conclusion/conclusion.tex
Updated license information in every file
Further improved text on reasons for high ECS in CMIP6

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                 Conclusion                                  %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 \acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased
-climate sensitivity in many \acs{CMIP}6 models is the addition of prognostic
-aerosol schemes that include aerosol-cloud interactions, which could have
-produced overly large negative radiative forcing. To correctly reproduce the
-observed historical temperature trend, a stronger response to \ac{GHG} forcing
-is required, resulting in a higher climate sensitivity in the corresponding
-models. A further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in
-the microphysical representation of mixed-phase clouds over the Southern Ocean
-in some models. This change was mainly implemented to reduce a long-standing
-cloud phase bias over the Southern Ocean that is present in models from
-previous \ac{CMIP} generations. Moreover, the \acs{CMIP}6 \ac{MMM} shows an
-improved simulation of the shortwave \ac{CRE} when compared to observations.
-However, the change in the representation of mixed-phase clouds also
-substantially reduces the strong negative shortwave cloud feedback over the
-Southern Ocean that is present in previous \ac{CMIP} generations and that
-results from a cloud phase change from ice clouds in the present-day to liquid
-clouds in the future. In the affected \acs{CMIP}6 models, this cloud phase
-change due to warming is reduced since these models simulate less cloud ice
-over the present-day Southern Ocean than their predecessor versions with no
-supercooled cloud liquid.
+climate sensitivity in some \acs{CMIP}6 models is the addition of prognostic
+aerosol schemes that include aerosol-cloud interactions, which might result in
+overly large negative radiative forcing. In this case, a stronger response to
+\ac{GHG} forcing is required to correctly reproduce the observed historical
+temperature trend, resulting in a higher climate sensitivity in the
+corresponding models. A further reason for the high \ac{ECS} in \acs{CMIP}6 is
+a change in the microphysical representation of mixed-phase clouds in some
+\acs{CMIP}6 models. This change was implemented to improve the representation
+of the fractioning of cloud ice and cloud liquid in those models, \eg{} by
+allowing for (more) supercooled cloud liquid water. In addition, the
+\acs{CMIP}6 \ac{MMM} shows an improved simulation of the shortwave \ac{CRE}
+when compared to satellite observations. The change in the representation of
+mixed-phase clouds in some \acs{CMIP}6 models reduces the strong negative
+shortwave cloud feedback over the Southern Ocean that is present in predecessor
+versions of these models and that results from a cloud phase change from ice
+clouds in the present-day to liquid clouds in the future. In the affected
+\acs{CMIP}6 models, this cloud phase change due to warming is reduced since
+these models simulate less cloud ice over the present-day Southern Ocean than
+their predecessor versions with no supercooled cloud liquid.
 

### 2021-03-26 ch02_scientific_background/scientific_background.tex
Updated license information in every file
Fixed acronyms at beginning of sentence

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                            Scientific Background                            %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 cycle, a dynamic vegetation, atmospheric chemistry and land ice (see
-\cref{fig:02:esms_historical_evolution}). \acp{AOGCM} coupled to these
+\cref{fig:02:esms_historical_evolution}). \Acp{AOGCM} coupled to these
 additional components are called \emph{\acfp{ESM}}, which are the current
@@
 group of exchange processes in the fast domain connects the atmosphere and the
-terrestrial biosphere. \ac{CO2} is removed from the atmosphere by plant
+terrestrial biosphere. \Ac{CO2} is removed from the atmosphere by plant
 photosynthesis with about $120 \unit{\GtCyr{}}$ \autocite{Ciais2013}. This

### 2021-03-26 ch06_paper_gpp/paper_gpp.tex
Updated license information in every file
Fixed acronyms at beginning of sentence

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                  Paper GPP                                  %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 \emph{\ac{GBRT}} algorithm is able to handle multiple predictors and copes with
-non-linearities in the data. \ac{GBRT} is a well-known and successful tool used
+non-linearities in the data. \Ac{GBRT} is a well-known and successful tool used
 for interpolation, classification and prediction in other fields of data

### 2021-03-27 ch02_scientific_background/scientific_background.tex
Fixed further acronyms, refs and cites at beginning of sentence

@@
 
-\subsection{Representation in \aclp{ESM}}
+\subsection{Representation in Earth System Models}
 \label{subsec:02:carbon_cycle_representation}
@@
 \ac{NBP}, which are all illustrated in \cref{fig:02:carbon_cycle_processes}.
-\ac{GPP} describes the total carbon uptake of the vegetation by photosynthesis,
+\Ac{GPP} describes the total carbon uptake of the vegetation by photosynthesis,
 while \ac{NPP} refers to the net carbon gain of the plants after accounting for

### 2021-03-27 ch05_paper_ecs/paper_ecs.tex
Updated license information in every file
Fixed further acronyms, refs and cites at beginning of sentence

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                  Paper ECS                                  %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 ($R^2 = 0.31$) and $\pmrangeunit{3.71}{1.09}{K}$ for \acs{CMIP}6 ($R^2 =
-0.01$). \textcite{Cox2018} derive a $66 \unit{\%}$ range of
+0.01$). \Textcite{Cox2018} derive a $66 \unit{\%}$ range of
 $\pmrangeunit{2.8}{0.6}{K}$ from a different subset of \acs{CMIP}5 models but
@@
 function calculated from \ac{DJF} means of the meridional wind field changes
-sign from negative to positive. \textcite{Lipat2017} explain this correlation
+sign from negative to positive. \Textcite{Lipat2017} explain this correlation
 by tying it to the observed correlation of the interannual variability in
@@
 $\pmrangeunit{3.77}{1.06}{K}$ for \acs{CMIP}6 ($R^2 = 0.03$).
-\textcite{Sherwood2014} do not give a best estimate for \ac{ECS} based on the
+\Textcite{Sherwood2014} do not give a best estimate for \ac{ECS} based on the
 large-scale component of mixing $D$ or its small-scale counterpart $S$ (see
@@
 complementary components of the total mixing phenomenon.
-\textcite{Sherwood2014} argue that the increase in dehydration depends on
+\Textcite{Sherwood2014} argue that the increase in dehydration depends on
 initial mixing linking it to cloud feedbacks and thus also to \ac{ECS}. For
@@
 $\pmrangeunit{3.67}{1.06}{K}$ for \acs{CMIP}6 ($R^2 = 0.16$).
-\textcite{Sherwood2014} give a best estimate of about $4 \unit{K}$ with a lower
+\Textcite{Sherwood2014} give a best estimate of about $4 \unit{K}$ with a lower
 limit of $3 \unit{K}$. Similar to both other constraints by
@@
 ($R^2 = 0.24$) and $\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 =
-0.06$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
+0.06$). \Textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
 significance of the emergent relationship drops massively from $p = 0.0089$ in
@@
 in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ}, a common
-problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds a link
+problem in many \acs{CMIP}5 climate models. \Textcite{Tian2015} finds a link
 between double-\ac{ITCZ} bias and simulated moisture, \acl{PR}, clouds, and
@@
 0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 < 0.01$).
-\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The emergent
+\Textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The emergent
 relationship shows a much lower statistical significance in \acs{CMIP}6 ($p =
@@
 \acs{CMIP}5 and \acs{CMIP}6 are out-of-sample tests for it.
-\textcite{Volodin2008} shows that high \ac{ECS} models tend to simulate a
+\Textcite{Volodin2008} shows that high \ac{ECS} models tend to simulate a
 higher total cloud cover over the southern midlatitudes and a lower total cloud
@@
 
-\ac{ECS} is the product of the complex interactions of the many components and
+\Ac{ECS} is the product of the complex interactions of the many components and
 feedbacks. Thus, constraining \ac{ECS} with a single physical process might

### 2021-03-27 ch06_paper_gpp/paper_gpp.tex
Fixed further acronyms, refs and cites at beginning of sentence

@@
 improvements upon this constraint that could be used in the future.
-\textcite{Wenzel2016} show a correlation between the increase in the \ac{CO2}
+\Textcite{Wenzel2016} show a correlation between the increase in the \ac{CO2}
 seasonal cycle amplitude simulated by models and the projected \ac{CO2}

### 2021-03-27 utils/search_in_tex.py

@@ -1,0 +1,75 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+# This file is used to produce materials for the following work:
+
+# Doctoral Dissertation
+# (c) by Manuel Schlund
+
+# This work is licensed under a
+# Creative Commons Attribution 4.0 International License.
+
+# You should have received a copy of the license along with this
+# work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
+
+"""Search regex in all ``tex`` files ignoring line breaks.
+
+Examples
+--------
+python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}cite{.{25}'
+python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}cite{.{25}'
+
+python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}ac.{0,12}{.{25}'
+python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}ac.{0,12}{.{25}'
+
+python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}cref.{0,12}{.{25}'
+python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}cref.{0,12}{.{25}'
+
+"""
+
+
+import re
+import sys
+from pathlib import Path
+
+
+ROOT = Path(__file__).parents[1].resolve()
+
+
+def get_file_as_str(filename):
+    """Read file and return ``str``."""
+    with open(filename, 'r') as f_in:
+        file_contents = f_in.read()
+    return file_contents
+
+
+def search(filename, pattern, replace_linebreak_char=' '):
+    """Search for pattern in filename."""
+    file_str = get_file_as_str(filename)
+    file_str_no_lb = file_str.replace('\n', replace_linebreak_char)
+    matches = re.findall(pattern, file_str_no_lb)
+    if not matches:
+        return
+    print(f"In file '{filename}', found {len(matches):d} matche(s)")
+    for match in matches:
+        print(f"    {match}")
+
+
+def main():
+    """Main function."""
+    if len(sys.argv) != 2:
+        raise ValueError(f"Expected exactly one command line argument for "
+                         f"regex pattern, got {len(sys.argv) - 1:d}")
+    pattern = sys.argv[1]
+    print(f"Using root path '{ROOT}'")
+    print(f"Using regex pattern '{pattern}'")
+    print()
+
+    # Iterate over all tex files
[…]

### 2021-03-29 ch02_scientific_background/scientific_background.tex
Cleaned tex files (hyphens, acronyms, etc.)

@@
 surface, ocean and sea ice (see \cref{fig:02:esms_historical_evolution}). These
-models are called \emph{\acp{AOGCM}} \autocite{Flato2013}. Over the course of
+models are called \emph{\acfp{AOGCM}} \autocite{Flato2013}. Over the course of
 the years, climate models became more and more complex by including a wider
@@
 per year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}).
-For this, the annual mean \acp{GSAT} are averaged over a 20 year period
+For this, the annual mean \acp{GSAT} are averaged over a 20-year period
 centered at the time of the \ac{CO2} doubling (years \range{61}{80} when the

### 2021-03-29 ch06_paper_gpp/paper_gpp.tex
Cleaned tex files (hyphens, acronyms, etc.)

@@
 which is calculated from the 10-year mean \ac{GPP} at the end of the \nth{21}
-century in the emission‐driven fully-coupled \acs{RCP}8.5 simulations
+century in the emission-driven fully-coupled \acs{RCP}8.5 simulations
 (\range{2091}{2100}) and the 10-year mean \ac{GPP} at the end of the \nth{20}
@@
 The whole procedure is repeated until the loss function does not improve
-anymore, as evaluated on an independent validation data set. Relevant
+anymore, as evaluated on an independent validation dataset. Relevant
 hyperparameters of this algorithm (maximum number of decision trees, maximum
@@
     \toprule
-    Predictor & \makecell{Observation‐driven \\ data} & Used time range &
+    Predictor & \makecell{Observation-driven \\ data} & Used time range &
     Physical connection to \acs{GPP} \\
@@
     \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in the
-    emission‐driven fully-coupled historical run (see \cref{eq:06:f}). In
+    emission-driven fully-coupled historical run (see \cref{eq:06:f}). In
     contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
@@
 
-\subsubsection{Prediction Error in Leave‐One‐Model‐Out \acs{CV} Approach and
+\subsubsection{Prediction Error in Leave-One-Model-Out \acs{CV} Approach and
   Feature Importance}
@@
 rather small for the gridded values. A far larger reduction of the \ac{RMSEP}
-can be achieved by using the regression models \acs{LASSO}‐1D, \ac{LASSO},
-\acs{GBRT}‐1D and \ac{GBRT}. All of them share similar \ac{RMSEP}
+can be achieved by using the regression models \acs{LASSO}-1D, \ac{LASSO},
+\acs{GBRT}-1D and \ac{GBRT}. All of them share similar \ac{RMSEP}
 distributions, which can be explained as follows: The historical \ac{GPP} is
@@
 the mean \ac{RMSEP} even further compared to the linear \ac{LASSO} models by
-about $2 \unit{\%}$ for \acs{GBRT}‐1D and $3 \unit{\%}$ for the full \ac{GBRT}
+about $2 \unit{\%}$ for \acs{GBRT}-1D and $3 \unit{\%}$ for the full \ac{GBRT}
 model. This leads to the conclusion that in addition to the strong linear
@@
 century. For this, we feed observation-based data (see
-\cref{tab:06:predictors}) into the regression models \ac{GBRT}, \acs{GBRT}‐1D,
-\ac{LASSO} and \acs{LASSO}‐1D.
+\cref{tab:06:predictors}) into the regression models \ac{GBRT}, \acs{GBRT}-1D,
+\ac{LASSO} and \acs{LASSO}-1D.
 
@@
     between the modeled \acs{GPP} in the \acs{CMIP}5 \acs{MMM} of the
-    historical simulation and the FLUXNET-MTE observation‐based estimate of
+    historical simulation and the FLUXNET-MTE observation-based estimate of
     \acs{GPP} \autocite{Jung2011} averaged between 1991 and 2000. Over large
@@
     correct the historical bias of the \acs{CMIP}5 ensemble relative to the
-    FLUXNET‐MTE product. \AdaptedFrom{Schlund2020}.
+    FLUXNET-MTE product. \AdaptedFrom{Schlund2020}.
   }
@@
 To further compare our \ac{GBRT} approach to other statistical models, we also
-fed the observation-driven data into the \acs{GBRT}‐1D, \ac{LASSO}, and
-\acs{LASSO}‐1D models to get a constrained projection of the \acs{RCP}8.5
+fed the observation-driven data into the \acs{GBRT}-1D, \ac{LASSO}, and
+\acs{LASSO}-1D models to get a constrained projection of the \acs{RCP}8.5
 \ac{GPP}. \Cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}
@@
 panels show values close to $0$ for most of the globe (\ie{} the differences
-between the \ac{GBRT} and \acs{GBRT}‐1D/\ac{LASSO} models are small),
+between the \ac{GBRT} and \acs{GBRT}-1D/\ac{LASSO} models are small),
 indicating that the bias-correcting property of our approach is also present in
@@
 improves the final result. Moreover, the similar patterns in both panels
-demonstrate that the \acs{GBRT}‐1D model emulates the strong linear relation of
+demonstrate that the \acs{GBRT}-1D model emulates the strong linear relation of
 the historical \ac{GPP} to future \ac{GPP} and performs equally well as the
-linear models. The global estimates of $167 \unit{\GtCyr{}}$ (\acs{GBRT}‐1D),
-$163 \unit{\GtCyr{}}$ (\ac{LASSO}), and $162 \unit{\GtCyr{}}$ (\acs{LASSO}‐1D)
+linear models. The global estimates of $167 \unit{\GtCyr{}}$ (\acs{GBRT}-1D),
+$163 \unit{\GtCyr{}}$ (\ac{LASSO}), and $162 \unit{\GtCyr{}}$ (\acs{LASSO}-1D)
 are all consistent with the global constraint of
@@
 heavily dominates the feature importance of the \ac{GBRT} model. Thus, using
-single-predictor models (\acs{GBRT}‐1D and \acs{LASSO}‐1D) is not necessary in
+single-predictor models (\acs{GBRT}-1D and \acs{LASSO}-1D) is not necessary in
 step 2b. \Cref{fig:06:step2b_ml_inference:b,fig:06:step2b_ml_inference:d} show
@@
     statistical models used to predict the fractional change in \acf{GPP} over
-    the \nth{21} century (step 2b) using the leave‐one‐model‐out \acl{CV}
+    the \nth{21} century (step 2b) using the leave-one-model-out \acl{CV}
     approach. A detailed description of the representation of these
@@
 
-In a last step, we use the FLUXNET‐MTE observational product to constrain the
+In a last step, we use the FLUXNET-MTE observational product to constrain the
 absolute \ac{GPP} increase at the end of the \nth{21} century

### 2021-03-29 ch07_conclusion/conclusion.tex
Fixed acronyms at beginning of sentence
Cleaned tex files (hyphens, acronyms, etc.)

@@
 reducing uncertainties in climate sensitivity itself but rather on a selected
-process that contributes to it: \ac{GPP}. \ac{GPP} is the largest flux of the
+process that contributes to it: \ac{GPP}. \Ac{GPP} is the largest flux of the
 terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
@@
 non-linear relationships between the predictors and the target variable. A a
-relationship between process‐oriented predictors and future projections of
+relationship between process-oriented predictors and future projections of
 \ac{GPP} is established and then utilized to project today's observed
@@
 comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
-\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
+\ac{LASSO} model) in a leave-one-model-out \ac{CV} setup. Compared to \ac{MMM}
 (\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48

### 2021-03-29 utils/search_in_tex.py

@@ -16,14 +16,16 @@
 
 Examples
 --------
-python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}cite{.{25}'
-python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}cite{.{25}'
+python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}cite\{.{25}'
+python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}cite\{.{25}'
 
-python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}ac.{0,12}{.{25}'
-python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}ac.{0,12}{.{25}'
+python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}ac.{0,12}\{.{25}'
+python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}ac.{0,12}\{.{25}'
 
-python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}cref.{0,12}{.{25}'
-python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}cref.{0,12}{.{25}'
+python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}ref.{0,12}\{.{25}'
+python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}ref.{0,12}\{.{25}'
+
+python search_in_tex.py '(?i)\\caption\[.*?\]\{.*?\\label' | egrep '\\ac\{'
 
 """
 
@@ -53,6 +55,7 @@ def search(filename, pattern, replace_linebreak_char=' '):
     print(f"In file '{filename}', found {len(matches):d} matche(s)")
     for match in matches:
         print(f"    {match}")
+    print()
 
 
 def main():

### 2021-03-30 02_abstract/abstract.tex
Updated license information in every file
Added Oxford comma for series with 'and'

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                  Abstract                                   %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 state-of-the-art climate models that allow numerical simulations of the complex
-physical, biological and chemical processes of the Earth system, are common
+physical, biological, and chemical processes of the Earth system, are common
 tools to understand and project climate change. Due to the chaotic nature of
-the climate system, unknowns in future emission pathways and uncertainties in
+the climate system, unknowns in future emission pathways, and uncertainties in
 the climate models, projections of the future climate are associated with large

### 2021-03-30 ch01_introduction/introduction.tex
Updated license information in every file
Cleaned tex files (hyphens, acronyms, etc.)
Added Oxford comma for series with 'and'

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                Introduction                                 %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 Climate change is one of the greatest challenges for humankind today. The
-warming of the climate system is \enquote{unequivocal} and \enquote{many of the
-  observed changes are unprecedented over decades to millennia}
+warming of the climate system is \enquote{unequivocal}, and \enquote{many of
+  the observed changes are unprecedented over decades to millennia}
 \autocite{IPCC2014}. The changing climate increases the \enquote{likelihood of
@@
   ecosystems} with \enquote{mostly negative impacts for biodiversity, ecosystem
-services and economic development} and amplifies \enquote{risks for
+services and economic development}, and amplifies \enquote{risks for
 livelihoods and for food and human security} \autocite{IPCC2014}. Potential
@@
 end of the pre-industrial era largely driven by economic and population
-growth led to atmospheric concentrations of \ac{CO2}, \ac{CH4} and \ac{N2O}
+growth led to atmospheric concentrations of \ac{CO2}, \ac{CH4}, and \ac{N2O}
 that are \enquote{unprecedented in at least the last 800000 years}
@@
 directly influence the global carbon cycle, an important biogeochemical cycle
-of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}},
-which describes the carbon uptake of the terrestrial biosphere due to
+of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}}, which
+describes the carbon uptake of the terrestrial biosphere due to
 photosynthesis. Since this land carbon uptake absorbs about $30 \unit{\%}$ of
@@
 equilibrium response of the climate system after a doubling of the atmospheric
-\ac{CO2} concentration and the \emph{\ac{TCR}}, which describes the transient
+\ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
 response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
@@
 involved contributors. However, unless stated otherwise, all contents from
-these publications (text, figures and tables) shown in this thesis originate
+these publications (text, figures, and tables) shown in this thesis originate
 from the author of this thesis. A detailed list of contributions to these
@@
 evaluation of \acp{ESM} which is useful for the entire scientific community and
-lead to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
+led to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
   Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:papers_ecs_tcr_assessment}

### 2021-03-30 ch02_scientific_background/scientific_background.tex
Added Oxford comma for series with 'and'

@@
 laboratory. Due to the immense complexity of the Earth system (including
-physical, biological and chemical processes on various temporal and spatial
+physical, biological, and chemical processes on various temporal and spatial
 scales and their mutual interactions), we do not have access to a tiny replica
@@
 model are the atmospheric prognostic variables velocity (horizontal and
-vertical), temperature, specific humidity, pressure and density. These
+vertical), temperature, specific humidity, pressure, and density. These
 prognostic variables are related to each other via the \emph{primitive
   equations}, a set of partial differential equations that can be derived from
-the conservation of momentum, mass, energy and moisture \autocite{Holton2004}.
+the conservation of momentum, mass, energy, and moisture \autocite{Holton2004}.
 To progress further in time, these primitive equations are solved numerically
@@
     early days, these models were so-called \acfp{AOGCM}, which only included
-    three components: the atmosphere, the land surface and the ocean. Over the
+    three components: the atmosphere, the land surface, and the ocean. Over the
     time, the individual components grew in complexity and included a wider
@@
 for example surface heat and moisture fluxes, moist convection, turbulent
-mixing and radiation \autocite{Holton2004}.
+mixing, and radiation \autocite{Holton2004}.
 
@@
 simulated only the physical components of the climate system: atmosphere, land
-surface, ocean and sea ice (see \cref{fig:02:esms_historical_evolution}). These
-models are called \emph{\acfp{AOGCM}} \autocite{Flato2013}. Over the course of
-the years, climate models became more and more complex by including a wider
-range of processes within the components, but also by introducing new
+surface, ocean, and sea ice (see \cref{fig:02:esms_historical_evolution}).
+These models are called \emph{\acfp{AOGCM}} \autocite{Flato2013}. Over the
+course of the years, climate models became more and more complex by including a
+wider range of processes within the components, but also by introducing new
 components to the coupled system. Examples of these are aerosols, the carbon
-cycle, a dynamic vegetation, atmospheric chemistry and land ice (see
+cycle, a dynamic vegetation, atmospheric chemistry, and land ice (see
 \cref{fig:02:esms_historical_evolution}). \Acp{AOGCM} coupled to these
@@
 One major element of \ac{CMIP} is to establish common standards, coordination,
-infrastructure and documentation in order to facilitate the distribution of
+infrastructure, and documentation in order to facilitate the distribution of
 climate model output \autocite{Eyring2016, Juckes2020}.
@@
 historical Atmospheric \ac{MIP} simulation (\emph{amip}), a simulation forced
-with an abrupt quadrupling of \ac{CO2} (\emph{abrupt-4xCO2}) and a simulation
+with an abrupt quadrupling of \ac{CO2} (\emph{abrupt-4xCO2}), and a simulation
 forced with a $1 \unit{\%}$ per year increase of the atmospheric \ac{CO2}
@@
 projections we can distinguish: natural variability, climate response
-uncertainty and emission uncertainty \autocite{Hawkins2009, Hawkins2010}.
+uncertainty, and emission uncertainty \autocite{Hawkins2009, Hawkins2010}.
 \Cref{fig:02:sources_of_uncertainty} shows these three sources for the
@@
     in future climate projections: Natural variability (orange), climate
-    response uncertainty (blue) and emission uncertainty (green)
+    response uncertainty (blue), and emission uncertainty (green)
     \autocite{Hawkins2009, Hawkins2010}. Climate response uncertainty can
@@
 system that arises from complex interactions between the ocean, atmosphere,
-land, biosphere and cryosphere \autocite{Cubasch2013}. It constitutes a
+land, biosphere, and cryosphere \autocite{Cubasch2013}. It constitutes a
 fundamental limit of how precisely we can project the future climate since it
@@
 Even though all climate models are built on the same physical principles, they
-differ in terms of spatial resolution, processes included and parameterizations
-of unresolved processes (see \cref{subsec:02:numerical_climate_modeling}).
+differ in terms of spatial resolution, processes included, and
+parameterizations of unresolved processes (see
+\cref{subsec:02:numerical_climate_modeling}).
 
@@
 is the largest contributor to the natural greenhouse effect. Since its amount
-in the atmosphere is mainly controlled by the air temperature and anthropogenic
-emissions of water vapor are negligible, the influence of water vapor on the
-climate system is described as a feedback mechanism and not as an external
-forcing \autocite{Myhre2013}. Basis of this feedback is the enhanced
+in the atmosphere is mainly controlled by the air temperature, and
+anthropogenic emissions of water vapor are negligible, the influence of water
+vapor on the climate system is described as a feedback mechanism and not as an
+external forcing \autocite{Myhre2013}. Basis of this feedback is the enhanced
 evaporation of water with increasing air temperatures. Each degree of warming
@@
 may alter the Earth's radiative budget are cloud cover, cloud optical
-thickness, cloud top and cloud base height, vertical extent and the
+thickness, cloud top and cloud base height, vertical extent, and the
 geographical distribution of clouds. Examples for robust cloud feedback
 processes are the increase in cloud top height of high-level clouds in a
-warming climate which traps longwave radiation and enhances global warming and
+warming climate which traps longwave radiation and enhances global warming, and
 the reduction in mid- and low-level cloud cover which diminishes the reflection
@@
 \end{equation}
-into the system. Positive values of $N$, $F$ and $R$ indicate incoming fluxes;
+into the system. Positive values of $N$, $F$, and $R$ indicate incoming fluxes;
 usually $F > 0$ and $R < 0$. On long time scales (multiple years), the net
@@
 $\lambda_\mathtxt{WV}$ refers to the water vapor feedback,
-$\lambda_\mathtxt{Albedo}$ to the snow/ice albedo feedback and
+$\lambda_\mathtxt{Albedo}$ to the snow/ice albedo feedback, and
 $\lambda_\mathtxt{Cloud}$ to the cloud feedback. Thus, the overall climate
@@
     \acf{ECS} and the climate feedback parameter $\lambda$. Dark gray, blue,
-    red and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
+    red, and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
     different time periods of a long running simulation (see legend). The
@@
 dependencies of the feedbacks, atmospheric and oceanic adjustments over time,
-changing warming patterns over time, non-additive feedbacks and dependencies on
-the type and magnitude of the external forcings \autocite{Knutti2017}. All in
-all, this demonstrates the limits of the linear feedback framework introduced
-in \cref{subsec:02:mathematical_framework_feedbacks} that is not capable of
-describing non-linear effects. A second major reason for the discrepancies
-between the equilibrium climate sensitivity and \ac{ECS} is the use of a
-\nxcotwo{4} instead of a \nxcotwo{2} simulation. The factor of $2$ in the
-denominator of \cref{eq:02:ecs} only partly compensates this difference since
-the radiative forcing logarithmically depends on the atmospheric \ac{CO2}
+changing warming patterns over time, non-additive feedbacks, and dependencies
+on the type and magnitude of the external forcings \autocite{Knutti2017}. All
+in all, this demonstrates the limits of the linear feedback framework
+introduced in \cref{subsec:02:mathematical_framework_feedbacks} that is not
+capable of describing non-linear effects. A second major reason for the
+discrepancies between the equilibrium climate sensitivity and \ac{ECS} is the
+use of a \nxcotwo{4} instead of a \nxcotwo{2} simulation. The factor of $2$ in
+the denominator of \cref{eq:02:ecs} only partly compensates this difference
+since the radiative forcing logarithmically depends on the atmospheric \ac{CO2}
 concentration \autocite{Huang2014}.
@@
 \emph{\ac{CRE} feedback parameters} $\lambda_\mathtxt{SWCRE}$,
-$\lambda_\mathtxt{LWCRE}$ and $\lambda_\mathtxt{CRE}$, which quantify the
+$\lambda_\mathtxt{LWCRE}$, and $\lambda_\mathtxt{CRE}$, which quantify the
 change in \ac{CRE} as a response to increasing \acp{GSAT}.
@@
 in rocks and sediments which are connected to the rapid domain of the carbon
-cycle through volcanic emissions of \ac{CO2}, chemical weathering, erosion and
+cycle through volcanic emissions of \ac{CO2}, chemical weathering, erosion, and
 sediment formation on the sea floor. These natural exchange fluxes between the
@@
 The fast domain of the global carbon cycle consists of three main carbon
-reservoirs: the atmosphere, the terrestrial biosphere and the ocean. In the
+reservoirs: the atmosphere, the terrestrial biosphere, and the ocean. In the
 atmosphere, carbon is mainly stored in trace gases, with \ac{CO2} as the major
@@
 carbon content are the trace gas \ac{CH4}, the trace gas \ac{CO}, hydrocarbons,
-black carbon aerosols and organic compounds \autocite{Ciais2013}. Carbon in the
-terrestrial biosphere is mainly stored as organic compounds, with about
+black carbon aerosols, and organic compounds \autocite{Ciais2013}. Carbon in
+the terrestrial biosphere is mainly stored as organic compounds, with about
 $\rangeunit{450}{650}{GtC}$ in the living vegetation biomass,
-$\rangeunit{1500}{2400}{GtC}$ in dead organic matter in litter and soils and
+$\rangeunit{1500}{2400}{GtC}$ in dead organic matter in litter and soils, and
 about $1700 \unit{GtC}$ in permafrost soils \autocite{Ciais2013}. The main
 component of the oceanic carbon reservoir is dissolved inorganic carbon
-(carbonic acid, bicarbonate ions and carbonate ions) with about $38000
+(carbonic acid, bicarbonate ions, and carbonate ions) with about $38000
 \unit{GtC}$. Further carbon is stored as dissolved organic carbon (about $700
-\unit{GtC}$), in surface sediments (about $1750 \unit{GtC}$) and in marine
+\unit{GtC}$), in surface sediments (about $1750 \unit{GtC}$), and in marine
 biota (about $3 \unit{GtC}$, predominantly phytoplankton and other
@@
     turnover time scales for carbon transfers through the major reservoirs
-    (atmosphere, land surface and ocean). \IPCCReproducedFrom{Ciais2013}{their
+    (atmosphere, land surface, and ocean). \IPCCReproducedFrom{Ciais2013}{their
       FAQ 6.2, figure 1}.
@@
 photosynthesis and $2.5 \unit{\GtCyr{}}$ from the atmosphere into the ocean due
-to increase dissolution of \ac{CO2} into the sea \autocite{Friedlingstein2020}.
-In the same time, the amount of carbon in the atmosphere reservoir increased
-with a rate of $5.1 \unit{\GtCyr{}}$, indicating that only about half of the
-anthropogenic \ac{CO2} emissions in the last decade remained in the atmosphere
-\autocite{Friedlingstein2020} where they can act as \ac{GHG}.
+to increased dissolution of \ac{CO2} into the sea
+\autocite{Friedlingstein2020}. In the same time, the amount of carbon in the
+atmosphere reservoir increased with a rate of $5.1 \unit{\GtCyr{}}$, indicating
+that only about half of the anthropogenic \ac{CO2} emissions in the last decade
+remained in the atmosphere \autocite{Friedlingstein2020} where they can act as
+\ac{GHG}.
 
@@
 $\CA{}$, $\CL{}$, and $\CO{}$ describe the corresponding changes in these
-carbon stores, namely the atmosphere, the terrestrial biosphere (land) and the
+carbon stores, namely the atmosphere, the terrestrial biosphere (land), and the
 ocean, respectively.
@@
 the major carbon exchange processes that need to be considered are
-photosynthesis, respiration and disturbances, which connect the atmospheric
+photosynthesis, respiration, and disturbances, which connect the atmospheric
 carbon reservoir and the land carbon reservoir. The latter mainly consists of
-the vegetation biomass, litter and soil. The terrestrial carbon exchange
-processes are commonly described with \ac{GPP}, \ac{NPP}, \ac{NEP} and
+the vegetation biomass, litter, and soil. The terrestrial carbon exchange
+processes are commonly described with \ac{GPP}, \ac{NPP}, \ac{NEP}, and
 \ac{NBP}, which are all illustrated in \cref{fig:02:carbon_cycle_processes}.
@@
 are known to affect photosynthesis, such as the availability of light, the
-\ac{CO2} concentration, the soil moisture and the temperature.
+\ac{CO2} concentration, the soil moisture, and the temperature.
 
@@
 basically assumes that all climate models are independent, equally plausible,
-distributed around reality and that the projected multi-model range is
+distributed around reality, and that the projected multi-model range is
 representative for the uncertainty in the projected quantity
@@
 represent the observed past and present-day climate \autocite{Gleckler2008,
-  Knutti2013} and that models might suffer from common structural limitations
+  Knutti2013}, and that models might suffer from common structural limitations
 like missing processes \autocite{Knutti2017a}.
@@
 approaches based on the Kullback-Leibler divergence between the models'
-\acp{PDF} of $x$ and the observational \ac{PDF} \autocite{Brient2016} and
+\acp{PDF} of $x$ and the observational \ac{PDF} \autocite{Brient2016}, and
 linear regression models based on hierarchical Bayesian models
@@
 sample Pearson correlation coefficient greater than $\abs*{r}$ if the null
-hypothesis is true. Smaller $p$-values indicate higher a higher statistical
+hypothesis is true. Smaller $p$-values indicate a higher statistical
 significance and vice versa.
@@
 \autocite{Amos2020}, North American maximum temperature \autocite{Lorenz2018},
-European temperature and \acl{PR} \autocite{Brunner2019, Merrifield2020} and
+European temperature and \acl{PR} \autocite{Brunner2019, Merrifield2020}, and
 global warming over the \nth{21} century \autocite{Brunner2020, Liang2020a}.
@@
 variable $y$ for emergent constraints. In \cref{fig:02:mder}, the observations
-of the predictors $\bm{x}_0$ are illustrated with horizontal black lines and
+of the predictors $\bm{x}_0$ are illustrated with horizontal black lines, and
 the best estimate $\hat{y}_0$ is shown as a black circle. By combining
@@
 the observational data, the limitation to linear relationships between the
-process-relevant diagnostics and the target variable and the limitation to a
+process-relevant diagnostics and the target variable, and the limitation to a
 single data point per climate model. Despite these, \ac{MDER} has been
@@
 projections \autocite{Karpechko2013}, in the projected change of the austral
-jet position \autocite{Wenzel2016a} and in projections of the Arctic sea ice
+jet position \autocite{Wenzel2016a}, and in projections of the Arctic sea ice
 extent \autocite{Senftleben2020}.

### 2021-03-30 ch03_esmvaltool/esmvaltool.tex
Updated license information in every file
Fixed acronyms at beginning of sentence
Added Oxford comma for series with 'and'

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                 ESMValTool                                  %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 \cref{sec:01:key_science_questions}, a reliable and efficient tool to read,
-process and evaluate climate model output and observational data is necessary.
+process, and evaluate climate model output and observational data is necessary.
 A valuable software that fits these criteria is the \ac{ESMValTool}. The
@@
 been contributed that improve the routine evaluation of climate models, which
-is beneficial for the entire scientific community. This lead to co-authorship
-in the scientific documentation of the \ac{ESMValTool}, which is published in
-four peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020,
-  Weigel2020}. After a brief overview of the \ac{ESMValTool}'s structure
+is beneficial for the entire scientific community. This led to co-authorship in
+the scientific documentation of the \ac{ESMValTool}, which is published in four
+peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
+After a brief overview of the \ac{ESMValTool}'s structure
 (\cref{sec:03:esmvaltool}), this \namecref{ch:03:esmvaltool} presents all these
@@
     output (\eg{} plots). The user can specify the desired input datasets,
-    preprocessing operations and diagnostic scripts in the recipe, which is a
+    preprocessing operations, and diagnostic scripts in the recipe, which is a
     configuration file that controls the main workflow of the \acs{ESMValTool}.
@@
 horizontal and vertical \emph{interpolation}, land/sea/ice \emph{masking},
-\emph{unit conversion}, or the calculation of \emph{multi-model statistics}. A
+\emph{unit conversion}, and the calculation of \emph{multi-model statistics}. A
 further example is the \emph{variable derivation}, which can be used to derive
@@
 be written in multiple programming languages. Currently, the languages Python,
-NCL, R and Julia are supported. The diagnostic scripts, which contain the code
+NCL, R, and Julia are supported. The diagnostic scripts, which contain the code
 that runs the actual scientific evaluation, read the preprocessed datasets from
@@
 which is a configuration file in which the user can specify the desired input
-datasets, preprocessing operations and diagnostic scripts. Implementing a new
+datasets, preprocessing operations, and diagnostic scripts. Implementing a new
 scientific assessment into the \ac{ESMValTool} usually includes writing a new
@@
 In total, three new preprocessor functions have been added: \emph{amplitude},
-\emph{land/sea fraction weighting} and \emph{trend}. The \emph{amplitude}
+\emph{land/sea fraction weighting}, and \emph{trend}. The \emph{amplitude}
 preprocessor calculates the peak-to-peak amplitude of periodic phenomena along
@@
 non-\ac{CMOR} variables \emph{asr}, \emph{co2s}, \emph{et}, \emph{rlntcs},
-\emph{rsntcs} and \emph{uajet} have been implemented into the \ac{ESMValCore}.
+\emph{rsntcs}, and \emph{uajet} have been implemented into the \ac{ESMValCore}.
 Apart from the atmospheric \ac{CO2} concentration at surface (\emph{co2s}),
@@
 metropolitan areas with high \ac{CO2} emissions (\eg{} over North America, the
-Arabian Peninsula and East Asia). Moreover, regions with high photosynthetic
+Arabian Peninsula, and East Asia). Moreover, regions with high photosynthetic
 activity (\eg{} the boreal forests in the Northern Hemisphere summer and the
@@
 diagnostics \autocite{Eyring2020}, diagnostics for emergent constraints and
-future projections \autocite{Lauer2020} and diagnostics for extreme events and
+future projections \autocite{Lauer2020}, and diagnostics for extreme events and
 regional evaluation \autocite{Weigel2020}.
@@
 analysis of \textcite{Anav2013}, who evaluate carbon cycle--related variables
-for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends
+for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends,
 and variabilities of important quantities of the terrestrial and oceanic carbon
-cycle like carbon fluxes, carbon reservoir sizes and vegetation distributions.
+cycle like carbon fluxes, carbon reservoir sizes, and vegetation distributions.
 \emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis performed by
@@
 (\cref{ch:04:papers_ecs_tcr_assessment}), \emph{recipe\_schlund20esd.yml}
-(\cref{ch:05:paper_ecs}) and \emph{recipe\_schlund20jgr\_*.yml}
+(\cref{ch:05:paper_ecs}), and \emph{recipe\_schlund20jgr\_*.yml}
 (\cref{ch:06:paper_gpp}). \emph{recipe\_ecs.yml} and \emph{recipe\_tcr.yml}
@@
     \range{1961}{1990} (larger symbols) and for pre-industrial control runs
-    (smaller symbols). (b) \acf{TCR} versus \acs{ECS} for \acs{CMIP}5 models.
+    (smaller symbols). (b) \Acf{TCR} versus \acs{ECS} for \acs{CMIP}5 models.
     The black line shows a linear fit between \acs{TCR} and \acs{ECS}.

### 2021-03-30 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added Oxford comma for series with 'and'

@@
 author of this thesis contributed their figure 8 (a bar chart showing the
-\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5 and \acs{CMIP}6 models), their
+\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their
 figure 10 (map plots showing cloud-related feedback parameters for the
@@
 \cref{fig:04:cloud_feedback_parameters}), code to create these figures with the
-\ac{ESMValTool} and text for the manuscript (in particular a section about the
+\ac{ESMValTool}, and text for the manuscript (in particular a section about the
 evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author of this thesis
 contributed all figures and tables, code to produce these with the
-\ac{ESMValTool} and text for the manuscript. This
+\ac{ESMValTool}, and text for the manuscript. This
 \namecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
@@
 \ac{ECS}, the relative difference is about $49 \unit{\%}$ ($0.72 \unit{K}$ in
-\acs{CMIP}5 to $1.07 \unit{K}$ in \acs{CMIP}6) and for \ac{TCR}, the relative
+\acs{CMIP}5 to $1.07 \unit{K}$ in \acs{CMIP}6), and for \ac{TCR}, the relative
 difference is about $17 \unit{\%}$ ($0.36 \unit{K}$ in \acs{CMIP}5 to $0.42
@@
 temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
-\acs{AR}4 and \acs{AR}5, the corresponding climate model generations from
-\acs{CMIP}1, \acs{CMIP}3 and \acs{CMIP}5 more or less agree with this assessed
+\acs{AR}4, and \acs{AR}5, the corresponding climate model generations from
+\acs{CMIP}1, \acs{CMIP}3, and \acs{CMIP}5 more or less agree with this assessed
 range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
@@
 
-Various improvements of the underlying physical, biological and chemical
+Various improvements of the underlying physical, biological, and chemical
 processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
@@
   ]{
-    Geographical distributions of the net (a), shortwave (c) and longwave (e)
+    Geographical distributions of the net (a), shortwave (c), and longwave (e)
     \acf{CRE} feedback parameters for the \acs{CMIP}6 \acf{MMM} and the
     corresponding differences to the \acs{CMIP}5 \acs{MMM}, again for the net
-    (b), shortwave (d) and longwave (f) components. The \acs{CRE} feedback
+    (b), shortwave (d), and longwave (f) components. The \acs{CRE} feedback
     parameters for each grid cell are calculated with a Gregory regression of
@@
 To further confirm this hypothesis, the global distributions of relevant cloud
-feedback parameters (expressed by the net, shortwave and longwave \ac{CRE}
+feedback parameters (expressed by the net, shortwave, and longwave \ac{CRE}
 feedback parameters; see \cref{subsec:02:cloud_feedback_parameters}) are shown

### 2021-03-30 ch05_paper_ecs/paper_ecs.tex
Cleaned tex files (hyphens, acronyms, etc.)
Added Oxford comma for series with 'and'

@@
 This work is already published in \textcite{Schlund2020a}. For this paper, the
-author of this thesis lead the writing and the analysis and implemented the
-code to reproduce this study with all figures and tables using the
-\ac{ESMValTool}. In this \namecref{ch:05:paper_ecs}, we first introduce the
-data (climate model output and observations) and methods used in this study
+author of this thesis led the writing and the analysis and implemented the code
+to reproduce this study with all figures and tables using the \ac{ESMValTool}.
+In this \namecref{ch:05:paper_ecs}, we first introduce the data (climate model
+output and observations) and methods used in this study
 (\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
@@
 relationship, the statistical significance using a two-sided $t$-test based on
-the Pearson correlation coefficient $r$ (see \cref{eq:02:t}) and the
+the Pearson correlation coefficient $r$ (see \cref{eq:02:t}), and the
 constrained range of \ac{ECS} based on the constrained \ac{PDF} (see
@@
 \autocite{Rayner2003}, ERA-Interim for $500 \unit{hPa}$ relative humidity
-\autocite{Dee2011} and CERES-EBAF \autocite{Loeb2018} for the \ac{TOA}
+\autocite{Dee2011}, and CERES-EBAF \autocite{Loeb2018} for the \ac{TOA}
 radiative fluxes over the time period \range{2001}{2005}. In the original
@@
   \caption[
-    Emergent constraints BRI, COX and LIP applied to the \acs{CMIP}5 and
+    Emergent constraints BRI, COX, and LIP applied to the \acs{CMIP}5 and
     \acs{CMIP}6 ensemble.
   ]{
-    Emergent constraints BRI, COX and LIP applied to the \acs{CMIP}5 (blue) and
-    \acs{CMIP}6 ensemble (orange). (a), (c), (e) Emergent relationships (solid
-    lines) for the \acs{CMIP} models (numbers; see
+    Emergent constraints BRI, COX, and LIP applied to the \acs{CMIP}5 (blue)
+    and \acs{CMIP}6 ensemble (orange). (a), (c), (e) Emergent relationships
+    (solid lines) for the \acs{CMIP} models (numbers; see
     \cref{tab:app:a:cmip5_emergent_constraints_part1,%
@@
 height regions: $850$ and $700 \unit{hPa}$ for shallow overturning and $600$,
-$500$ and $400 \unit{hPa}$ for deep overturning. Both quantities are averaged
+$500$, and $400 \unit{hPa}$ for deep overturning. Both quantities are averaged
 over parts of the tropical ocean region away from the regions of highest
@@
   \caption[
-    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SHD, SHL
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SHD, SHL,
     and SHS.
   ]{
-    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SHD, SHL
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SHD, SHL,
     and SHS (see \cref{tab:app:a:cmip5_emergent_constraints_part1,%
@@
   \caption[
-    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU, TIH
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU, TIH,
     and TII.
   ]{
-    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU, TIH
+    As in \cref{fig:05:bri_cox_lip} but for the emergent constraints SU, TIH,
     and TII (see \cref{tab:app:a:cmip5_emergent_constraints_part2,%
@@
 with observations of cloud fraction from Cloudsat/CALIPSO \autocite{Mace2009},
-\ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and vertical velocity from
+\ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011}, and vertical velocity from
 ERA-Interim \autocite{Dee2011}, the seasonal response of \ac{MBLC} fraction to
@@
 in the original publication (ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m,
-CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and
+CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR, and
 MPI-ESM-P). Due to a lack of publicly available data, the model
@@
       tab:app:a:cmip6_emergent_constraints_part2} for details). The $p$-values
-    describing the significance of the emergent relationships are defined as
+    describing the significances of the emergent relationships are defined as
     the probability to obtain an absolute correlation coefficient $\abs*{r}$
@@
     almost all emergent constraints, the constrained best estimate and range of
-    \ac{ECS} is higher in the \acs{CMIP}6 ensemble.
+    \acs{ECS} is higher in the \acs{CMIP}6 ensemble.
   }
@@
 would pass this test on \acs{CMIP}5 model data but only five (BRI, SHL, SHS,
-VOL and ZHA) would pass this test on \acs{CMIP}6. This is still much better
+VOL, and ZHA) would pass this test on \acs{CMIP}6. This is still much better
 than would be expected purely by chance. Hence, there is still skill in at
@@
 during the Last Glacial Maximum \autocite{Hargreaves2012}, tropical temperature
-anomalies during the mid-Pliocene Warm Period \autocite{Hargreaves2016} and
+anomalies during the mid-Pliocene Warm Period \autocite{Hargreaves2016}, and
 post-1970s warming \autocite{JimenezdelaCuesta2019}. This seems to be supported
@@
 from multiple emergent constraints on \ac{ECS} using a multivariate Gaussian
-and multiple linear regression. For the \acs{CMIP}3 and \acs{CMIP}5 ensembles,
-they find an increased best estimate relative to the unweighted ensemble mean
-similar to the participating individual emergent constraints, but with lower
-uncertainty range. Moreover, new \ac{ML} techniques are a promising avenue
-forward for such multivariate approaches and for constraining uncertainties in
-multi-model projections \commentcite{Schlund2020}{see \cref{ch:06:paper_gpp}}
-with the aim of further improving climate modeling and analysis
-\autocite{Reichstein2019}.
+and multivariate linear regression. For the \acs{CMIP}3 and \acs{CMIP}5
+ensembles, they find an increased best estimate relative to the unweighted
+ensemble mean similar to the participating individual emergent constraints, but
+with lower uncertainty range. Moreover, new \ac{ML} techniques are a promising
+avenue forward for such multivariate approaches and for constraining
+uncertainties in multi-model projections \commentcite{Schlund2020}{see
+  \cref{ch:06:paper_gpp}} with the aim of further improving climate modeling
+and analysis \autocite{Reichstein2019}.

### 2021-03-30 ch06_paper_gpp/paper_gpp.tex
Added Oxford comma for series with 'and'

@@
 emergent constraints, this approach considers (1) multiple process-based
-predictors, (2) multi-dimensional (gridded) target variables and (3) non-linear
-relationships between the predictors and the target variable. Since the new
-technique relies on a large number of data points in order to train the \ac{ML}
-algorithm, \ac{ECS} as a scalar climate sensitivity metric is not an
+predictors, (2) multi-dimensional (gridded) target variables, and (3)
+non-linear relationships between the predictors and the target variable. Since
+the new technique relies on a large number of data points in order to train the
+\ac{ML} algorithm, \ac{ECS} as a scalar climate sensitivity metric is not an
 appropriate target variable. Thus, this analysis does not focus on reducing
@@
 non-linearities in the data. \Ac{GBRT} is a well-known and successful tool used
-for interpolation, classification and prediction in other fields of data
+for interpolation, classification, and prediction in other fields of data
 science and engineering \autocite{Death2007, Elith2008}. In the context of
@@
 This work is already published in \textcite{Schlund2020}. For this paper, the
-author of this thesis lead the writing and the analysis and implemented the
-code to reproduce this study with all figures and tables using the
-\ac{ESMValTool}. \Cref{sec:06:data_and_methods} of this
-\namecref{ch:06:paper_gpp} provides an overview of the data and methods used in
-this paper. The results are presented in \cref{sec:06:results} and
-\cref{sec:06:summary_and_discussion} closes with a summary and discussion.
+author of this thesis led the writing and the analysis and implemented the code
+to reproduce this study with all figures and tables using the \ac{ESMValTool}.
+\Cref{sec:06:data_and_methods} of this \namecref{ch:06:paper_gpp} provides an
+overview of the data and methods used in this paper. The results are presented
+in \cref{sec:06:results} and \cref{sec:06:summary_and_discussion} closes with a
+summary and discussion.
 
@@
 two remaining models equally attribute this to \ac{CO2} fertilization, climate
-change and land use changes. Moreover, \textcite{Winkler2019a} discuss
+change, and land use changes. Moreover, \textcite{Winkler2019a} discuss
 limitations of carbon cycle emergent constraints related to the calculation of
@@
 This variable is given as a percentage, where negative values correspond to a
-decrease in \ac{GPP}, positive values to an increase in \ac{GPP} and a value of
-$0 \unit{\%}$ to no \ac{GPP} change. The \ac{CO2} amplitude sensitivity is
+decrease in \ac{GPP}, positive values to an increase in \ac{GPP}, and a value
+of $0 \unit{\%}$ to no \ac{GPP} change. The \ac{CO2} amplitude sensitivity is
 calculated for the years \range{1979}{2019}, which is the full available time
@@
 global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
-$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change
+$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change,
 and $y_{mi}$ the target variable of climate model $m$. Then, we define the
@@
 hyperparameters of this algorithm (maximum number of decision trees, maximum
-depth of the decision trees, learning rate and the subsample ratio) are
+depth of the decision trees, learning rate, and the subsample ratio) are
 optimized by $5$-fold \acf{CV} using the training data. More details on the
@@
 three sources of uncertainty: the uncertainty of the \ac{ML} model itself, the
-uncertainty in the rescaling of the target variable and the uncertainty in the
+uncertainty in the rescaling of the target variable, and the uncertainty in the
 input data, that is, errors in the observation-based products. First, the
@@
 the internal variability of the climate system, the different model responses
-within the climate model ensemble to forcing and the incomplete description of
+within the climate model ensemble to forcing, and the incomplete description of
 the climate system with the limited number of variables used as input features.
@@
 defined in \cref{tab:06:predictors}, a single-predictor \ac{LASSO} model
-(\acs{LASSO}-1D) and a single-predictor \ac{GBRT} model (\acs{GBRT}-1D). Both
+(\acs{LASSO}-1D), and a single-predictor \ac{GBRT} model (\acs{GBRT}-1D). Both
 single-predictor models use the historical \ac{GPP} as only feature. The
@@
 can be achieved by using the regression models \acs{LASSO}-1D, \ac{LASSO},
-\acs{GBRT}-1D and \ac{GBRT}. All of them share similar \ac{RMSEP}
+\acs{GBRT}-1D, and \ac{GBRT}. All of them share similar \ac{RMSEP}
 distributions, which can be explained as follows: The historical \ac{GPP} is
@@
 \cref{tab:06:predictors}) into the regression models \ac{GBRT}, \acs{GBRT}-1D,
-\ac{LASSO} and \acs{LASSO}-1D.
+\ac{LASSO}, and \acs{LASSO}-1D.
 
@@
 we present in \cref{subsec:06:uncertainty_and_skill} (error in the \ac{GBRT}
-model, error in the rescaling and error in the observational-driven
+model, error in the rescaling, and error in the observational-driven
 products). In contrast to that, the error derived from the
@@
 \ac{GPP}). The remaining setup including the predictors, the \ac{GBRT} model
-itself and the datasets are similar.
+itself, and the datasets are similar.
 
@@
 \ac{GPP} change, its rescaled version using \cref{eq:06:y} (r\acs{MMM}), a
-linear \ac{LASSO} regression model and the \ac{GBRT} model. Similar to step 2a,
-the \ac{GBRT} approach shows the smallest mean \ac{RMSEP} values of all
+linear \ac{LASSO} regression model, and the \ac{GBRT} model. Similar to step
+2a, the \ac{GBRT} approach shows the smallest mean \ac{RMSEP} values of all
 statistical models. Compared to \ac{MMM} and r\acs{MMM} the mean is reduced by
@@
     approach \autocite{Ribeiro2016} for the three dominant features \acf{T},
-    \acf{LAI} and \acs{GPP}. The relative influence of these three features
+    \acf{LAI}, and \acs{GPP}. The relative influence of these three features
     (ignoring all other features) is color-coded according to the triangle in
@@
     \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:y}, (c) linear
-    \acf{LASSO} model and (d) \acf{GBRT} model. The geographical patterns from
+    \acf{LASSO} model, and (d) \acf{GBRT} model. The geographical patterns from
     the different statistical models are very similar and show a higher
@@
     America, South Africa, the west coast of Africa, the Middle East, parts of
-    Australia and western parts of the United States.
+    Australia, and western parts of the United States.
     \AdaptedFrom{Schlund2020}.
@@
 equator. For most parts of the remaining globe, \ac{T} is the dominant
-predictor, with some areas in northern Africa, the Middle East, India and
+predictor, with some areas in northern Africa, the Middle East, India, and
 Australia showing an increased relative importance of the historical \ac{GPP}.
@@
 the global pattern of the fractional \ac{GPP} change: The high latitudes with
-lower \ac{T}, lower \ac{LAI} and lower historical \ac{GPP} show higher \ac{GPP}
-changes, whereas regions closer to the equator with higher \ac{T}, higher
-\ac{LAI} and higher historical \ac{GPP} show smaller \ac{GPP} changes. At this
-point we want to emphasize that it is not possible to infer global or other
-large-scale averages of the fractional \ac{GPP} change from the geographical
-distributions in \cref{fig:06:step2b_results} since division operations and
-averaging operations do not commute: in general, $\sum_i \left( a_i / b_i
-\right) \neq \left( \sum_i a_i \right) / \left( \sum_i b_i \right)$. Thus, a
-direct comparison to other studies that give locally or globally aggregated
-results is not possible.
+lower \ac{T}, lower \ac{LAI}, and lower historical \ac{GPP} show higher
+\ac{GPP} changes, whereas regions closer to the equator with higher \ac{T},
+higher \ac{LAI}, and higher historical \ac{GPP} show smaller \ac{GPP} changes.
+At this point we want to emphasize that it is not possible to infer global or
+other large-scale averages of the fractional \ac{GPP} change from the
+geographical distributions in \cref{fig:06:step2b_results} since division
+operations and averaging operations do not commute: in general, $\sum_i \left(
+a_i / b_i \right) \neq \left( \sum_i a_i \right) / \left( \sum_i b_i \right)$.
+Thus, a direct comparison to other studies that give locally or globally
+aggregated results is not possible.
 
@@
 in particular South America, South Africa, the west coast of Africa, the Middle
-East, parts of Australia and western parts of the United States. However, all
+East, parts of Australia, and western parts of the United States. However, all
 these correspond to low absolute \ac{GPP} changes.

### 2021-03-30 ch07_conclusion/conclusion.tex
Added Oxford comma for series with 'and'

@@
 evaluation of established methods to reduce these uncertainties in the new
-climate model ensemble and finally introduces a novel alternative technique
+climate model ensemble, and finally introduces a novel alternative technique
 based on supervised \ac{ML}.
@@
 relevant study in this context is given by \textcite{Brunner2020}, who use
-performance and interdependence-based climate model weighting to constrain the
+performance- and interdependence-based climate model weighting to constrain the
 \acl{T} over the \nth{21} century based on observable process-based predictors
@@
 \ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
-  Mansfield2020} or learning of entire \acp{ESM} from observational products
+  Mansfield2020}, or learning of entire \acp{ESM} from observational products
 \autocite{Geer2021}. Therefore, the foundation for innovative and

### 2021-03-30 end_acknowledgments/acknowledgments.tex
Restructured thesis and removed unnecessary parts
Updated license information in every file
Added Oxford comma for series with 'and'

@@
 %=============================================================================%
-%                                Dissertation                                 %
-%                               Manuel Schlund                                %
-%                                  (c) 2020                                   %
+%                            Doctoral Dissertation                            %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                               Acknowledgments                               %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
   the Earth System with Machine Learning (USMILE)} under grant agreement number
-855187 and the \emph{Advanced Earth System Model Evaluation for CMIP
+855187, and the \emph{Advanced Earth System Model Evaluation for CMIP
   (EVal4CMIP)} project funded by the Helmholtz Society. I acknowledge the
@@
 climate modeling groups for producing and making available their model output,
-the \ac{ESGF} for archiving the data and providing access and the multiple
+the \ac{ESGF} for archiving the data and providing access, and the multiple
 funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational

### 2021-03-30 end_appendix/appendix.tex
Updated license information in every file
Added Oxford comma for series with 'and'

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                  Appendix                                   %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
     \acf{ECS} and \xaxis{} values for the emergent constraints BRI, COX, LIP,
-    SHD, SHL and SHS.
+    SHD, SHL, and SHS.
   ]{
@@
     (\nameref*{ch:05:paper_ecs}) including their \acf{ECS} and \xaxis{} values
-    for the emergent constraints BRI, COX, LIP, SHD, SHL and SHS. More details
+    for the emergent constraints BRI, COX, LIP, SHD, SHL, and SHS. More details
     on the emergent constraints are given in
@@
     As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
-    emergent constraints SU, TIH, TII, VOL and ZHA.
+    emergent constraints SU, TIH, TII, VOL, and ZHA.
   ]{
     As in \cref{tab:app:a:cmip5_emergent_constraints_part1} but for the
-    emergent constraints SU, TIH, TII, VOL and ZHA. \AdaptedFrom{Schlund2020a}.
+    emergent constraints SU, TIH, TII, VOL, and ZHA.
+    \AdaptedFrom{Schlund2020a}.
   }
@@
 where $H$ is the total number of decision trees, $\beta_j$ the expansion
-coefficient for tree $j$ and $\bm{\alpha}_j$ the hyperparameters for tree $j$.
+coefficient for tree $j$, and $\bm{\alpha}_j$ the hyperparameters for tree $j$.
 Using all $N$ training data points $\left\{ \left( \bm{x}_i, y_i \right) \mid i
@@
 complexity of the individual trees (for example measured by the maximum tree
-depth) and the learning rate $\nu \ll 1$. The latter parameter is used for
+depth), and the learning rate $\nu \ll 1$. The latter parameter is used for
 regularization and dramatically reduces the risk of overfitting by scaling down
@@
 the standard error in the global mean fractional \acs{GPP} change given by the
-emergent constraint from step 1 and $\bar{f}$ the \acs{CMIP}5 \ac{MMM} global
+emergent constraint from step 1, and $\bar{f}$ the \acs{CMIP}5 \ac{MMM} global
 mean fractional \ac{GPP} change over the \nth{21} century.
@@
     provide all necessary variables (\emph{co2}, \emph{gpp}, \emph{lai},
-    \emph{pr}, \emph{rsds} and \emph{tas}) for all used experiments
-    (\emph{esmHistorical}, \emph{esmrcp85} and \emph{esmFixClim1}). For all
+    \emph{pr}, \emph{rsds}, and \emph{tas}) for all used experiments
+    (\emph{esmHistorical}, \emph{esmrcp85}, and \emph{esmFixClim1}). For all
     models, we only use the first ensemble member available.

### 2021-03-31 ch01_introduction/introduction.tex
Standardized the usage of 'emph' and 'enquote'

@@
 dynamics of the physical components of the Earth system, are also known as
-\acp{ESM} and provide the most sophisticated simulations of the Earth's
+\emph{\acp{ESM}} and provide the most sophisticated simulations of the Earth's
 climate. In this thesis, the terms \enquote{climate model} and
@@
 
-Simulations that extrapolate state of the climate system into the future are
-called \emph{projections} of the future climate. These include idealized
+Simulations that extrapolate the state of the climate system into the future
+are called \emph{climate model projections}. These include idealized
 simulations with only a prescribed change in the atmospheric \ac{CO2}

### 2021-03-31 ch02_scientific_background/scientific_background.tex
Standardized the usage of 'emph' and 'enquote'

@@
 replicate the vastly complex Earth system with finite computational resources,
-climate models divide the Earth into a set of \emph{grid cells}. The typical
-size of an atmospheric grid cell in modern-day global climate models is about
-$100 \unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
+climate models divide the Earth into a set of grid cells. The typical size of
+an atmospheric grid cell in modern-day global climate models is about $100
+\unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
 \unit{km}$ along the vertical dimension (pressure level or height). In
@@
 \ac{ScenarioMIP} experiments can give valuable insights into possible
-developments of the Earth system by providing \emph{projections} of the future
-climate. In contrast to climate predictions, climate projections run over
-multiple decades and depend upon the future scenario considered, which are
+developments of the Earth system by providing future \emph{climate
+  projections}. In contrast to climate predictions, climate projections run
+over multiple decades and depend upon the future scenario considered, which are
 based on assumptions that may or may not turn out to be correct. On the
@@
 variations in the solar activity. Natural variability can be seen as the
-\enquote{noise} in the climate record as opposed to the anthropogenic
-\enquote{signal} \autocite{Cubasch2013}.
+\emph{noise} in the climate record as opposed to the anthropogenic
+\emph{signal} \autocite{Cubasch2013}.
 
@@
 expert judgment about relevant processes, on emergent relationships (see
-\cref{subsec:02:emergent_constraints}) or on multivariate regression models
+\cref{subsec:02:emergent_constraints}), or on multivariate regression models
 \commentcite{Karpechko2013, Sanderson2015a}{see \cref{subsec:02:mder}}. It
@@
 optimal parameters is the \emph{leave-one-model-out \ac{CV}} approach, which is
-also known as \emph{pseudo-reality}, \emph{model-as-truth} or \emph{perfect
+also known as \emph{pseudo-reality}, \emph{model-as-truth}, or \emph{perfect
   model} setup \autocite{Elia2002, Karpechko2013}. For this, a single climate
@@
 \end{equation}
-where the exponent \enquote{$-1$} denotes the inverse matrix. Since this
-definition works for any number of diagnostics $K$, it can also be used to
-calculate the linear coefficients $\hat{b}_0$ (intercept) and $\hat{b}_1$
-(slope) that define emergent relationships (see
-\cref{eq:02:linear_regression_y}).
+where the exponent $-1$ denotes the inverse matrix. Since this definition works
+for any number of diagnostics $K$, it can also be used to calculate the linear
+coefficients $\hat{b}_0$ (intercept) and $\hat{b}_1$ (slope) that define
+emergent relationships (see \cref{eq:02:linear_regression_y}).
 

### 2021-03-31 ch06_paper_gpp/paper_gpp.tex
Standardized the usage of 'emph' and 'enquote'

@@
 cell level, it is able to adapt to regional climate characteristics and thus
-applies an implicit \emph{local} and \emph{climate-dependent} weighting.
-Moreover, the algorithm adds another level of weighting by assigning a relative
-relevance score to the different predictors. We emphasize that this is only an
-implicit weighting, since it is not possible to extract specific values for the
+applies an implicit local and climate-dependent weighting. Moreover, the
+algorithm adds another level of weighting by assigning a relative relevance
+score to the different predictors. We emphasize that this is only an implicit
+weighting, since it is not possible to extract specific values for the
 individual weights due to the complex structure of the \ac{GBRT} model.

### 2021-03-31 ch07_conclusion/conclusion.tex
Standardized the usage of 'emph' and 'enquote'

@@
 already published in \textcite{Bock2020, Meehl2020}, directly addresses
-\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
-\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
-which is higher on the upper and lower end than any model range from previous
-\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
-\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
-\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
-upper bound of the \acs{CMIP}6 models range is considerably larger than in the
-assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
-\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
-bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
-many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
-range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
-\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
-\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased
-climate sensitivity in some \acs{CMIP}6 models is the addition of prognostic
-aerosol schemes that include aerosol-cloud interactions, which might result in
-overly large negative radiative forcing. In this case, a stronger response to
-\ac{GHG} forcing is required to correctly reproduce the observed historical
-temperature trend, resulting in a higher climate sensitivity in the
-corresponding models. A further reason for the high \ac{ECS} in \acs{CMIP}6 is
-a change in the microphysical representation of mixed-phase clouds in some
-\acs{CMIP}6 models. This change was implemented to improve the representation
-of the fractioning of cloud ice and cloud liquid in those models, \eg{} by
-allowing for (more) supercooled cloud liquid water. In addition, the
-\acs{CMIP}6 \ac{MMM} shows an improved simulation of the shortwave \ac{CRE}
-when compared to satellite observations. The change in the representation of
-mixed-phase clouds in some \acs{CMIP}6 models reduces the strong negative
-shortwave cloud feedback over the Southern Ocean that is present in predecessor
-versions of these models and that results from a cloud phase change from ice
-clouds in the present-day to liquid clouds in the future. In the affected
-\acs{CMIP}6 models, this cloud phase change due to warming is reduced since
-these models simulate less cloud ice over the present-day Southern Ocean than
-their predecessor versions with no supercooled cloud liquid.
+\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
+\acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
+on the upper and lower end than any model range from previous \ac{CMIP}
+generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of
+\ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus
+$3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
+\acs{CMIP}6 models range is considerably larger than in the assessed range of
+$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
+\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
+\unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with many models
+showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model range of
+\acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the \acs{CMIP}5
+range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from \acs{AR}5 of
+$\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased climate
+sensitivity in some \acs{CMIP}6 models is the addition of prognostic aerosol
+schemes that include aerosol-cloud interactions, which might result in overly
+large negative radiative forcing. In this case, a stronger response to \ac{GHG}
+forcing is required to correctly reproduce the observed historical temperature
+trend, resulting in a higher climate sensitivity in the corresponding models. A
+further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in the
+microphysical representation of mixed-phase clouds in some \acs{CMIP}6 models.
+This change was implemented to improve the representation of the fractioning of
+cloud ice and cloud liquid in those models, \eg{} by allowing for (more)
+supercooled cloud liquid water. In addition, the \acs{CMIP}6 \ac{MMM} shows an
+improved simulation of the shortwave \ac{CRE} when compared to satellite
+observations. The change in the representation of mixed-phase clouds in some
+\acs{CMIP}6 models reduces the strong negative shortwave cloud feedback over
+the Southern Ocean that is present in predecessor versions of these models and
+that results from a cloud phase change from ice clouds in the present-day to
+liquid clouds in the future. In the affected \acs{CMIP}6 models, this cloud
+phase change due to warming is reduced since these models simulate less cloud
+ice over the present-day Southern Ocean than their predecessor versions with no
+supercooled cloud liquid.
 
@@
 emergent constraints. Thus, \cref{enum:01:question_2}
-(\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to be answered with a
-\enquote{not very well} for the \acs{CMIP}6 ensemble. The increased best
-estimates and spreads resulting from the emergent constraints in \acs{CMIP}6
-are likely related to the increased \ac{MMM} and multi-model spread of \ac{ECS}
-in \acs{CMIP}6. A possible reason for the reduced skill of the emergent
+(\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{not
+  very well} for the \acs{CMIP}6 ensemble. The increased best estimates and
+spreads resulting from the emergent constraints in \acs{CMIP}6 are likely
+related to the increased \ac{MMM} and multi-model spread of \ac{ECS} in
+\acs{CMIP}6. A possible reason for the reduced skill of the emergent
 constraints when applied to the \acs{CMIP}6 ensemble is the increased
@@
 This study directly addresses \cref{enum:01:question_3}
-(\enquote{\emph{\KeyScienceQuestionThree{}}}), which can be confidently
-answered with \enquote{yes} based on the results found.
+(\enquote{\KeyScienceQuestionThree{}}), which can be confidently answered with
+\enquote{yes} based on the results found.
 

### 2021-03-31 02_abstract/abstract.tex
Further optimizations of abstract, introduction and conclusion

@@
 uncertainties. The main focus of this thesis is the analysis of future climate
-projections from \acsp{ESM} participating in the \acfAbstract{CMIP} and the
-reduction of uncertainties in climate projections with observations.
+projections from \acsp{ESM} participating in the \acfAbstract{CMIP} with the
+aim to reduce uncertainties in climate projections with observations.
 
 In a first step, the climate sensitivity (\ie{} the temperature response of the
-climate system to an external forcing) in the latest generation of \acsp{ESM}
-from \acs{CMIP}6 is evaluated. For the \acfAbstract{ECS}, which is defined as
-the equilibrium temperature response that follows a doubling of the atmospheric
+climate system to an external forcing) is evaluated in the latest generation of
+\acsp{ESM} from \acs{CMIP}6. For the \acfAbstract{ECS}, which is defined as the
+equilibrium temperature response that follows a doubling of the atmospheric
 \acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74 \unit{K}$ and a
@@
 higher than in any previous \acs{CMIP} ensemble before. Moreover, a third of
-the analyzed \acs{CMIP}6 models exceeds the upper bound of the likely \acs{ECS}
+the analyzed \acs{CMIP}6 models exceed the upper bound of the likely \acs{ECS}
 range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC} Fifth
@@
 aerosol schemes that include aerosol-cloud interactions and changes in the
-microphysical representation of mixed-phase clouds over the Southern Ocean,
-which reduce the strong negative shortwave cloud phase change feedback in this
-region that is present in climate models from previous \acs{CMIP} generations.
+microphysical representation of mixed-phase clouds. The changes in the
+mixed-cloud representation reduce the strong negative shortwave cloud phase
+change feedback over the Southern Ocean that is present in climate models from
+previous \acs{CMIP} generations.
 
@@
 observations and output from \acsp{ESM}. The focus of this analysis is on
-testing if these emergent constraints hold for \acsp{ESM} participating
+testing if these emergent constraints hold for \acsp{ESM} participating in
 \acs{CMIP}6. Since none of the emergent constraints considered here have been
-derived on the \acs{CMIP}6 ensemble, the \acs{CMIP}6 models can be used for
-cross-checking of the emergent constraints on a new model ensemble. The
+derived from the \acs{CMIP}6 ensemble, the \acs{CMIP}6 models can be used for
+cross-checking the emergent constraints on a new model ensemble. The
 application of the emergent constraints to \acs{CMIP}6 data shows a decrease in
@@
 relationships or a too small number of models in the ensemble that the emergent
-constraint was originally derived from. The corresponding best estimates given
-by the emergent constraints also increase from \acs{CMIP}5 to \acs{CMIP}6 by
-$12 \unit{\%}$ on average. This can be at least partly explained by the
-increased number of high-\acs{ECS} models in \acs{CMIP}6 without a
+constraint was originally derived from. The corresponding best estimates of
+\acs{ECS} given by the emergent constraints also increase from \acs{CMIP}5 to
+\acs{CMIP}6 by $12 \unit{\%}$ on average. This can at least be partly explained
+by the increased number of high-\acs{ECS} models in \acs{CMIP}6 without a
 corresponding change in the constraint predictors, suggesting the emergence of

### 2021-03-31 ch01_introduction/introduction.tex
Further optimizations of abstract, introduction and conclusion

@@
 leads to a warming of the Earth's surface and the lower atmosphere. The
-individual impact of the different drivers of climate change measured with
+individual impact of the different drivers of climate change is measured with
 the so-called \emph{radiative forcing}, which quantifies the change in energy
@@
 directly influence the global carbon cycle, an important biogeochemical cycle
-of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}}, which
-describes the carbon uptake of the terrestrial biosphere due to
+of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}},
+which describes the carbon uptake of the terrestrial biosphere due to
 photosynthesis. Since this land carbon uptake absorbs about $30 \unit{\%}$ of
@@
 \autocite{Friedlingstein2020}, this process substantially slows down global
-warming and directly contributes to the magnitude of the climate sensitivity.
+warming and directly contributes to the magnitude of climate sensitivity.
 Other important anthropogenic drivers of climate change are the emission of
-aerosols and land use/land cover changes. Apart from anthropogenic drivers,
+aerosols, and land use/land cover changes. Apart from anthropogenic drivers,
 there are also natural processes which impact the climate system like changes
@@
 conditions without having to perform (ethnically questionable) experiments in
-the real world. All around the world climate research institutes provide a
+the real world. All around the world climate research institutes develop a
 variety of different climate models. Many of them participate in the \ac{CMIP},
@@
 consider different future scenarios (\eg{} a fossil fuel--based future or a
-scenario that is based on a sustainable development). In many variables that
-are relevant for climate change, multi-model projections from \ac{CMIP} show a
-large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
+scenario that is based on a sustainable development). Multi-model projections
+from \ac{CMIP} show a large inter-model range in many variables that are
+relevant for climate change \autocite{Collins2013, Flato2013}. A crucial and
 policy-relevant example for this is the climate sensitivity, which refers to
@@
 \ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
-response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
-have been assessed with large ranges of $\rangeunit{1.5}{4.5}{K}$ and
-$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
-\autocite{Stocker2013}. The corresponding inter-model ranges from the
-\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
-a careful statistical evaluation and further refinement of the output of
-multi-model climate projections is necessary in order to reduce associated
-uncertainties. A state-of-the-art technique for this is the \emph{emergent
-  constraints} method, which uses a physically-based inter-model relationship
-between an observable quantity of the Earth system and a target variable to
-reduce uncertainties in the target variable with observations
-\autocite{Allen2002}. An alternative approach is the weighting of climate
-models based on their performance (\ie{} the distance of one model to
-observational products) and interdependence (\ie{} the distance of one model to
-other climate models) \autocite{Knutti2017a}. These techniques form the
-baseline for the new analyses and results presented in this thesis, which
-partly utilize methods from a new emerging research field in climate sciences:
-\ac{AI} and \ac{ML}.
+response of the system to a \ac{CO2} doubling. In \acs{AR}5, the assessed range
+of both quantities is large with $\rangeunit{1.5}{4.5}{K}$ for \ac{ECS} and
+$\rangeunit{1.0}{2.5}{K}$ for \ac{TCR} \autocite{Stocker2013}. The
+corresponding inter-model ranges from the \acs{CMIP}5 models show similar
+results \autocite{Flato2013}. For this reason, a careful statistical evaluation
+and further refinement of the output of multi-model climate projections is
+necessary in order to reduce associated uncertainties. A state-of-the-art
+technique for this is the \emph{emergent constraints} method, which uses a
+physically-based inter-model relationship between an observable quantity of the
+Earth system and a target variable to reduce uncertainties in the target
+variable with observations \autocite{Allen2002}. An alternative approach is the
+weighting of climate models based on their performance (\ie{} the distance of
+one model to observational products) and interdependence (\ie{} the distance of
+one model to other climate models) \autocite{Knutti2017a}. These techniques
+form the baseline for the new analyses and results presented in this thesis,
+which partly utilize methods from a new emerging research field in climate
+sciences: \ac{AI} and \ac{ML}.
 
@@
 publications (two first-author studies and six co-author studies). A complete
-list of these is given on \cpageref{ch:bib:own}. Wherever material from these
-studies is presented in this thesis, the pronoun \enquote{we} is used to
-increase readability by avoiding the passive voice and to acknowledge all
-involved contributors. However, unless stated otherwise, all contents from
+list of these publications is given on \cpageref{ch:bib:own}. Wherever material
+from these studies is presented in this thesis, the pronoun \enquote{we} is
+used to increase readability by avoiding the passive voice and to acknowledge
+all involved contributors. However, unless stated otherwise, all content from
 these publications (text, figures, and tables) shown in this thesis originate
@@
 of emergent constraints on the \ac{ECS} for these \acs{CMIP}6 models and
-compares these to results derived from \acs{CMIP}5 models. The contents of this
-\namecref{ch:05:paper_ecs} are published in \emph{Earth System Dynamics}
+compares these to results derived from \acs{CMIP}5 models. The content of this
+\namecref{ch:05:paper_ecs} is published in \emph{Earth System Dynamics}
 \autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
@@
 \autocite{Schlund2020}. Finally, \cref{ch:07:conclusion} provides a summary of
-the results of this thesis and gives an outlook of possible future works.
+the results of this thesis and gives an outlook of possible future work.

### 2021-03-31 ch02_scientific_background/scientific_background.tex
Further optimizations of abstract, introduction and conclusion

@@
 amplified or diminished by climate feedbacks. Thus, feedback processes play a
-crucial role determining the magnitude of the climate sensitivity.
+crucial role determining the magnitude of climate sensitivity.
 \Cref{fig:02:feedbacks_overview} shows an overview of important feedbacks in
@@
 might change under global warming, which in turn directly influences the
-magnitude of the climate sensitivity. However, the two idealized climate
+magnitude of climate sensitivity. However, the two idealized climate
 sensitivity metrics \ac{ECS} and \ac{TCR} introduced in the previous

### 2021-03-31 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Further optimizations of abstract, introduction and conclusion

@@
 exceeds this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third
-(14 models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to the analysis
+(14 models) have an \ac{ECS} above $4.5 \unit{K}$. In addition to the analysis
 presented here, several modeling centers independently confirmed high \ac{ECS}

### 2021-03-31 ch05_paper_ecs/paper_ecs.tex
Further optimizations of abstract, introduction and conclusion

@@
   Flato2013}{see also \cref{subsec:02:climate_feedbacks}}. Since all of the
-evaluated emergent constraints have been derived on the \acs{CMIP}3 or
+evaluated emergent constraints have been derived from the \acs{CMIP}3 or
 \acs{CMIP}5 ensemble, the \acs{CMIP}6 models offer an exciting possibility to
@@
 \acs{CMIP}6. Some increase is predicted by almost every constraint we analyze
-(with TII as the only exception) and can be at least partly explained by the
+(with TII as the only exception) and can at least be partly explained by the
 increased multi-model mean \ac{ECS} of \acs{CMIP}6, which has not been

### 2021-03-31 ch07_conclusion/conclusion.tex
Further optimizations of abstract, introduction and conclusion

@@
 \autocite{UNFCCC2015}, research in this field of science is not only relevant
-to climate scientists but to policymakers and the whole human society. In the
+to climate scientists but to policymakers and human society as a whole. In the
 light of the large spread of climate sensitivity in the most recent generation
@@
 refinement of the output of multi-model climate projections is as relevant as
-ever. This thesis quantifies the associated uncertainties, presents the
-evaluation of established methods to reduce these uncertainties in the new
-climate model ensemble, and finally introduces a novel alternative technique
-based on supervised \ac{ML}.
+ever. This thesis quantifies uncertainties in climate model projections,
+presents the evaluation of established methods to reduce these uncertainties in
+a new climate model ensemble, and finally introduces a novel alternative
+technique based on supervised \ac{ML}.
 
@@
 \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
-on the upper and lower end than any model range from previous \ac{CMIP}
-generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of
-\ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus
-$3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
+on the upper and lower bound than any model range from previous \ac{CMIP}
+generations. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of \ac{ECS}
+is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus $3.23
+\unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
 \acs{CMIP}6 models range is considerably larger than in the assessed range of
@@
 climate  models, already-published emergent constraints that have been derived
-on models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5
+from models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5
 are evaluated for their skill in the \acs{CMIP}6 ensemble. Emergent constraints
@@
 \ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated emergent
-constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5 ensemble,
+constraints have been derived from the \acs{CMIP}3 or \acs{CMIP}5 ensemble,
 out-of-sample tests on the emergent constraints can be performed by assessing
@@
 process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
-due to an increased number of processes that are included in the \acs{CMIP}6
-models.
+for the latest generation of \acp{ESM} from \acs{CMIP}6 due to an increased
+number of processes that are included in these models.
 
@@
 terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
-from the atmosphere. In the first step of the new two-step approach, the global
-mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
-\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
-a published emergent constraint by \textcite{Wenzel2016}. In contrast to that,
-the corresponding multi-model range of \ac{GPP} in the raw \acs{CMIP}5 ensemble
-is $\rangeunit{156}{247}{\GtCyr{}}$. This first step corrects the \acs{CMIP}5
-models' \ac{GPP} response to \ac{CO2}, which is primarily driven by the
-\ac{CO2} fertilization. In the second step, a \ac{ML}-based climate model
-weighting approach is used to further constrain the gridded \ac{GPP} based on
-present-day predictors that are relevant for the simulation of \ac{GPP} in the
-\acp{ESM}. The \ac{ML} approach is mathematically similar to the \ac{MDER}
-approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}, but
-additionally considers multi-dimensional (gridded) target variables and
-non-linear relationships between the predictors and the target variable. A a
-relationship between process-oriented predictors and future projections of
-\ac{GPP} is established and then utilized to project today's observed
-conditions into the future. The prediction phase of the new method can be
-interpreted as an implicit performance weighting. However, due to the complex
-structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
-extract specific values for the individual weights. Two target variables are
+from the atmosphere. In the first step of the new two-step approach, a
+published emergent constraint by \textcite{Wenzel2016} is used to constrain the
+global mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
+\acs{RCP}8.5 scenario to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
+unconstrained \acs{CMIP}5 model range of $\rangeunit{156}{247}{\GtCyr{}}$. This
+first step corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2},
+which is primarily driven by the \ac{CO2} fertilization. In the second step, a
+\ac{ML}-based climate model weighting approach is used to further constrain the
+gridded \ac{GPP} based on present-day predictors that are relevant for the
+simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach is mathematically
+similar to the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020,
+  Wenzel2016a}, but additionally considers multi-dimensional (gridded) target
+variables and non-linear relationships between the predictors and the target
+variables. A relationship between process-oriented predictors and future
+projections of \ac{GPP} is established and then utilized to project today's
+observed conditions into the future. The prediction phase of the new method can
+be interpreted as an implicit performance weighting. However, due to the
+complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
+to extract specific values for the individual weights. Two target variables are
 considered: the gridded monthly climatologies of absolute \ac{GPP}
@@
 strength of climate change. The range of model results for this important
-climate metric, however, remains large and has not been narrowed over the last
-decades. In the latest generation of climate models contributing to
+climate metric, however, remains large and could not be narrowed down over the
+last decades. In the latest generation of climate models contributing to
 \acs{CMIP}6, the range has even increased \autocite{Meehl2020}.
@@
 analyzed in detail yet. This could include additional feedback processes that
-are not present in older \ac{CMIP} generations. Obvious candidates for these
-feedback processes are cloud-related feedbacks, which pose a major source of
-uncertainty in climate sensitivity in modern-day \acp{ESM}
+are not present in older \ac{CMIP} model generations. Obvious candidates for
+these feedback processes are cloud-related feedbacks, which pose a major source
+of uncertainty in climate sensitivity in modern-day \acp{ESM}
 \autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms

### 2021-04-06 ch01_introduction/introduction.tex
Fixed capitalization in headings and subject-verb agreement for collective nouns

@@
 drivers for climate change are all natural and anthropogenic substances and
-processes that may alter the Earth's energy budget. The human influence on
-the climate system is clear: Ever increasing emissions of \acp{GHG} since the
-end of the pre-industrial era largely driven by economic and population
-growth led to atmospheric concentrations of \ac{CO2}, \ac{CH4}, and \ac{N2O}
-that are \enquote{unprecedented in at least the last 800000 years}
+processes that may alter the Earth's energy budget. Human influence on the
+climate system is clear: Ever increasing emissions of \acp{GHG} since the end
+of the pre-industrial era largely driven by economic and population growth
+led to atmospheric concentrations of \ac{CO2}, \ac{CH4}, and \ac{N2O} that
+are \enquote{unprecedented in at least the last 800000 years}
 \autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other
@@
 (which corresponds to a warming) and its largest contribution is caused by
-the increase in the atmospheric concentration of \ac{CO2} due to fossil fuel
-emissions since the year 1750 \autocite{IPCC2014}. Apart from their physical
-warming effect on the climate, carbon-based \acp{GHG} like \ac{CO2} also
-directly influence the global carbon cycle, an important biogeochemical cycle
-of the Earth. A crucial flux of the carbon cycle is the \emph{\ac{GPP}},
-which describes the carbon uptake of the terrestrial biosphere due to
-photosynthesis. Since this land carbon uptake absorbs about $30 \unit{\%}$ of
-the anthropogenic \ac{CO2} emissions in today's climate
+the increase in the atmospheric concentration of \ac{CO2} due to anthropogenic
+fossil fuel emissions since the mid-\nth{18} century \autocite{IPCC2014}. Apart
+from their physical warming effect on the climate, carbon-based \acp{GHG} like
+\ac{CO2} also directly influence the global carbon cycle, an important
+biogeochemical cycle of the Earth. A crucial flux of the carbon cycle is the
+\emph{\ac{GPP}}, which describes the carbon uptake of the terrestrial biosphere
+due to photosynthesis. Since this land carbon uptake absorbs about $30
+\unit{\%}$ of the anthropogenic \ac{CO2} emissions in today's climate
 \autocite{Friedlingstein2020}, this process substantially slows down global
-warming and directly contributes to the magnitude of climate sensitivity.
-Other important anthropogenic drivers of climate change are the emission of
-aerosols, and land use/land cover changes. Apart from anthropogenic drivers,
-there are also natural processes which impact the climate system like changes
-in the solar activity or volcanic eruptions. However, there is clear evidence
-that these natural drivers alone cannot explain the observed climate change
+warming and directly contributes to the magnitude of climate sensitivity. Other
+important anthropogenic drivers of climate change are the emission of aerosols,
+and land use/land cover changes. Apart from anthropogenic drivers, there are
+also natural processes which impact the climate system like changes in the
+solar activity or volcanic eruptions. However, there is clear evidence that
+these natural drivers alone cannot explain the observed climate change
 \autocite{Haustein2017}.

### 2021-04-06 ch02_scientific_background/scientific_background.tex
Fixed capitalization in headings and subject-verb agreement for collective nouns

@@
 of future societal development \autocite{ONeill2013, ONeill2017}. For each
-experiment, a set of emissions and land use changes is calculated from the
+experiment, specific emissions and land use changes are calculated from the
 \acp{SSP} \autocite{Riahi2017} which are then used to force the global climate
@@
 For some domains of the Earth system, feedbacks can be positive and/or
-negative, since a variety of different mechanisms is involved. An example for
+negative, since a variety of different mechanisms are involved. An example for
 this is the cloud feedback. Changes in the clouds induced by climate change can
@@
 
-\subsection{Equilibrium and Effective Climate Sensitivity}
+\subsection{Equilibrium and Effective Climate Sensitivity (\acs{ECS})}
 \label{subsec:02:ecs}
@@
 
-\subsection{Transient Climate Response}
+\subsection{Transient Climate Response (\acs{TCR})}
 \label{subsec:02:tcr}

### 2021-04-06 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Fixed capitalization in headings and subject-verb agreement for collective nouns

--- 2021-03-30 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
@@
 spread in the \acs{CMIP}6 ensemble is the existence of several models with very
-high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exists a
+high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exist a
 number of models with very low values of \ac{ECS}. Therefore, the \acs{CMIP}6
@@
 bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
-$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that
-exceeds this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third
-(14 models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to the analysis
+$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that exceed
+this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third (14
+models) have an \ac{ECS} above $4.5 \unit{K}$. In addition to the analysis
 presented here, several modeling centers independently confirmed high \ac{ECS}
 values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
-  Wyser2020}. A list of possible reasons for the increased climate sensitivity
-in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
+  Wyser2020}. A list of possible reasons for the increased climate sensitivity in
+the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
 the shortwave cloud feedback over the Southern Ocean is given in

### 2021-04-06 ch05_paper_ecs/paper_ecs.tex
Fixed capitalization in headings and subject-verb agreement for collective nouns

@@
 \chapter{Evaluation of Emergent Constraints on the Effective Climate
-  Sensitivity in \acs{CMIP}6}
+  Sensitivity (\acs{ECS}) in \acs{CMIP}6}
 \label{ch:05:paper_ecs}
@@
 associated uncertainties. For \ac{ECS}, a large variety of different emergent
-constraints has been developed over the years, which we now evaluate on the
+constraints have been developed over the years, which we now evaluate on the
 \acs{CMIP}6 ensemble in order to assess whether they still hold for these new
@@
 
-\subsection{Sensitivity of Shortwave Cloud Albedo to changes in \acl{SST}
-  (BRI)}
+\subsection{Sensitivity of Shortwave Cloud Albedo to Changes in Sea Surface
+  Temperature (BRI)}
 \label{subsec:05:bri}
@@
 
-\subsection{Southern \acs{ITCZ} index from (TII)}
+\subsection{Southern \acs{ITCZ} Index (TII)}
 \label{subsec:05:tii}
@@
 
-\subsection{Response of Marine Boundary Layer Cloud Fraction to changes in
-  \acl{SST} (ZHA)}
+\subsection{Response of Marine Boundary Layer Cloud Fraction to Changes in Sea
+  Surface Temperature (ZHA)}
 \label{subsec:05:zha}
@@
 
-\subsection{Constrained \acs{ECS} ranges and statistical significance of the
-  eleven emergent constraints}
+\subsection{Constrained \acs{ECS} Ranges and Statistical Significance of the
+  Eleven Emergent Constraints}
 \label{subsec:05:emergent_constraints_summary}
@@
 correlations of variables between past climate and \ac{ECS} just by chance,
-especially when only a small number of climate models is considered. This would
-cause the performance of the emergent constraint to be reduced on out-of-sample
-data (like the new \acs{CMIP}6 ensemble), since the emergent relationship
-appeared just by chance and not because of a physically-based mechanism.
+especially when only a few climate models are considered. This would cause the
+performance of the emergent constraint to be reduced on out-of-sample data
+(like the new \acs{CMIP}6 ensemble), since the emergent relationship appeared
+just by chance and not because of a physically-based mechanism.
 

### 2021-04-06 ch02_scientific_background/scientific_background.tex
Ran spellchecker on tex files

@@
 \ac{CMIP}, \acs{CMIP}6, climate models need to run a \emph{historical}
-simulation of the period \range{1850}{2014} and the so-called \ac{DECK}
+simulation of the period \range{1850}{2014} and the so-called \emph{\ac{DECK}}
 experiments, which include a pre-industrial control run (\emph{piControl}), a
@@
 units are \ac{ppm} for the atmospheric trace gas concentrations (dry-air mole
-fraction) and \ac{GtC} or $\si{\GtCyr{}}$ for the reservoirs masses or exchange
+fraction) and \ac{GtC} or $\si{\GtCyr{}}$ for the reservoir masses or exchange
 fluxes, respectively. The carbon exchange processes between the different
@@
 Using an observation of $x$, this emergent relationship can then be used to
-derive a emergent constraint on $y$ (gray shaded area in
+derive an emergent constraint on $y$ (gray shaded area in
 \cref{fig:02:emergent_constraint}) that considers uncertainties in the emergent
@@
 in the beginning of \cref{sec:02:reducing_uncertainties}, this is not the case
-for typical climate models ensembles, as some modeling groups provide output
-for multiple climate models and some climate models from different modeling
+for typical climate model ensembles, as some modeling groups provide output for
+multiple climate models and some climate models from different modeling
 institutions share components and code \autocite{Knutti2013}. The duplicated
@@
 physical mechanism, an emergent constraint can be considered credible. The
-reason for this are spurious relationships: Due to the large number of possible
-observables provided by modern \acp{ESM} and the comparatively small number of
-climate models, spurious relationships are possible just by chance
-\autocite{Caldwell2014}. Furthermore, out-of-sample tests offer an important
-tool to evaluate the credibility of emergent constraints \autocite{Hall2019}.
-These ensure that the existence of an emergent relationship is not limited to a
-certain climate model ensemble and might indicate that the relationship is also
-valid for the true climate system. Testing emergent constraints in different
-\ac{CMIP} generations offers a straightforward setup for out-of-sample testing
-\autocite{Caldwell2018}, which is discussed in more detail in
-\cref{ch:05:paper_ecs}, where eleven emergent constraints on \ac{ECS} are
-evaluated on the new \acs{CMIP}6 ensemble.
+reason for this is the occurrence spurious relationships: Due to the large
+number of possible observables provided by modern \acp{ESM} and the
+comparatively small number of climate models, spurious relationships are
+possible just by chance \autocite{Caldwell2014}. Furthermore, out-of-sample
+tests offer an important tool to evaluate the credibility of emergent
+constraints \autocite{Hall2019}. These ensure that the existence of an emergent
+relationship is not limited to a certain climate model ensemble and might
+indicate that the relationship is also valid for the true climate system.
+Testing emergent constraints in different \ac{CMIP} generations offers a
+straightforward setup for out-of-sample testing \autocite{Caldwell2018}, which
+is discussed in more detail in \cref{ch:05:paper_ecs}, where eleven emergent
+constraints on \ac{ECS} are evaluated on the new \acs{CMIP}6 ensemble.
 

### 2021-04-06 ch03_esmvaltool/esmvaltool.tex
Fixed capitalization in headings and subject-verb agreement for collective nouns
Ran spellchecker on tex files

@@
 observations) and \emph{save} (saving the preprocessed data), which utilize the
-\ac{netCDF} format (a self-describing, machine-independent binary file format
+\ac{NetCDF} format (a self-describing, machine-independent binary file format
 used for the storage of array-based scientific data). Further non-optional
@@
 \ac{ESMValCore} and use these to create the final output of the tool. Apart
-from \ac{netCDF} files and plots, this includes log files with provenance
+from \ac{NetCDF} files and plots, this includes log files with provenance
 information that ensure the reproducibility and transparency of the results.
@@
 \ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high
-temperatures in the historical or pre-industrial period do no necessarily
+temperatures in the historical or pre-industrial period do not necessarily
 exhibit a high equilibrium warming in the future or vice versa. On the other
@@
 the corresponding observational datasets is shown in
-\cref{tab:03:changes_to_esmvaltool}. The majority of these datasets is used for
-analysis presented in this thesis (\eg{} see
+\cref{tab:03:changes_to_esmvaltool}. The majority of these datasets are used
+for analysis presented in this thesis (\eg{} see
 \cref{tab:05:overview_emergent_constraints,tab:06:predictors}).

### 2021-04-06 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Ran spellchecker on tex files

@@
 used in \acs{AR}3 exhibit values of \acs{ECS} well above $4.5 \unit{K}$,
-resulting in a upper model range of about $0.6 \unit{K}$ higher than the
+resulting in an upper model range of about $0.6 \unit{K}$ higher than the
 assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
@@
 values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
-  Wyser2020}. A list of possible reasons for the increased climate sensitivity in
-the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
+  Wyser2020}. A list of possible reasons for the increased climate sensitivity
+in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
 the shortwave cloud feedback over the Southern Ocean is given in

### 2021-04-06 ch05_paper_ecs/paper_ecs.tex
Ran spellchecker on tex files

@@
 than those derived from earlier \ac{CMIP} models? In general, the answer to
-this question is \enquote{no}. Even the \acs{CMIP}6 models are far from fully
-simulating all aspects and processes of the full vastly complex Earth system,
-which could introduce new biases which have not been present in earlier
-\ac{CMIP} models. For example, the apparent increase in climate sensitivity in
-the \acs{CMIP}6 models might be wrongly caused by missing processes that cancel
-out the effects of the newly included processes. Thus, to assess the
-credibility of \ac{ESM} results it is vital to take other sources of
-information into account.
+this question is \enquote{not necessarily}. Even the \acs{CMIP}6 models are far
+from fully simulating all aspects and processes of the full vastly complex
+Earth system. Therefore, the addition of new processes to these models could
+introduce new biases which have not been present in earlier \ac{CMIP} models.
+For example, the apparent increase in climate sensitivity in the \acs{CMIP}6
+models might be wrongly caused by missing processes that cancel out the effects
+of the newly included processes. Thus, it is vital to take other sources of
+information into account in order to assess the credibility of \ac{ESM}
+results.
 
@@
     \midrule
-    ZHA* & \textcite{Zhai2015} & Seasonal response of \acl{MBLC} cloud
-    fraction to changes in \acs{SST} [$\si{\pctK{}}$] &
-    \makecell{\tabitem \emph{cl} (Cloudsat/CALIPSO) \\ \tabitem \emph{tos}
+    ZHA* & \textcite{Zhai2015} & Seasonal response of \acl{MBLC} fraction to
+    changes in \acs{SST} [$\si{\pctK{}}$] &
+    \makecell{\tabitem \emph{cl} (CloudSat/CALIPSO) \\ \tabitem \emph{tos}
       (AMSRE \ac{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
@@
 between a climate model's relative humidity profile and the corresponding
-observational reference. Similarly to the original publication, we use humidity
+observational reference. Similar to the original publication, we use humidity
 observations from AIRS \autocite{Aumann2003} for pressure levels greater than
@@
 with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5 models. Together
-with observations of cloud fraction from Cloudsat/CALIPSO \autocite{Mace2009},
+with observations of cloud fraction from CloudSat/CALIPSO \autocite{Mace2009},
 \ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011}, and vertical velocity from

### 2021-04-06 ch06_paper_gpp/paper_gpp.tex
Fixed capitalization in headings and subject-verb agreement for collective nouns
Ran spellchecker on tex files

@@
 
-\chapter{Constraining Future Gross Primary Production with Machine Learning}
+\chapter{Constraining Future Gross Primary Production (\acs{GPP}) with Machine
+  Learning}
 \label{ch:06:paper_gpp}
@@
 global constraint from step 1 of $\pmrangeunit{171}{12}{\GtCyr{}}$. In summary,
-our approach first corrects the \acs{CMIP}5 models response to \ac{CO2} (step
+our approach first corrects the \acs{CMIP}5 models' response to \ac{CO2} (step
 1) and second corrects the historical bias of the \acp{ESM} relative to
@@
 All four patterns of \cref{fig:06:step2b_results} show very similar
-geographical patterns. Obvious exceptions are the Sahara desert and the Arabian
+geographical patterns. Obvious exceptions are the Sahara Desert and the Arabian
 Peninsula, which show a noisy behavior for the \aclp{MMM} (\ac{MMM} and

### 2021-04-06 ch07_conclusion/conclusion.tex
Ran spellchecker on tex files

@@
 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
-\acs{CMIP}6 models range is considerably larger than in the assessed range of
+\acs{CMIP}6 model range is considerably larger than in the assessed range of
 $\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
@@
 high latitudes, which might be attributed to an additional greening trend in
-the high latitudes that enhances the \ac{GPP} increase in this regions compared
+the high latitudes that enhances the \ac{GPP} increase in this region compared
 to regions closer to the equator through an extension of the growing season.

### 2021-04-06 end_appendix/appendix.tex
Standardized the usage of 'emph' and 'enquote'
Fixed capitalization in headings and subject-verb agreement for collective nouns
Ran spellchecker on tex files

@@
     CERES-EBAF & rsdt, rsut, rsutcs & \textcite{Loeb2018} \\
-    Cloudsat/CALIPSO & cl & \textcite{Mace2009} \\
+    CloudSat/CALIPSO & cl & \textcite{Mace2009} \\
     ERA-Interim & hur, ta, va, wap & \textcite{Dee2011} \\
@@
 \in I_N \right\}$ with index set $I_N = \left\{ 1, 2, \ldots, N \right\}$
-(\enquote{classic} gradient boosting), the expansion coefficients and tree
+(\emph{classic gradient boosting}), the expansion coefficients and tree
 parameters are jointly fitted by minimizing a loss function $L \left( y,
@@
 process is repeated $n$ times so that every input point is part of the
-validation set at least once. The optimal hyperparameters are the set of
-hyperparameters with optimal performance on the validation datasets
+validation set at least once. The optimal hyperparameters are the
+hyperparameters that show optimal performance on the validation datasets
 \autocite{Elith2008}.

### 2021-04-07 ch03_esmvaltool/esmvaltool.tex
Fixed use of acronyms followed by non-whitespace

@@
 peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
-After a brief overview of the \ac{ESMValTool}'s structure
+After a brief overview of the \acs{ESMValTool}'s structure
 (\cref{sec:03:esmvaltool}), this \namecref{ch:03:esmvaltool} presents all these
@@
 recipe and one or more diagnostic scripts. A third major element of the
-\ac{ESMValTool} is the \emph{\ac{CMOR}ization} (\ie{} the reformatting of data
+\ac{ESMValTool} is the \emph{\acs{CMOR}ization} (\ie{} the reformatting of data
 so it adheres to \ac{CMOR} standards) of raw observational datasets (not shown
@@
 \cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs,ch:06:paper_gpp}, several
-more recipes and \ac{CMOR}ization scripts for observational data have been
+more recipes and \acs{CMOR}ization scripts for observational data have been
 added to the main \ac{ESMValTool} repository on GitHub. A summary of these is
@@
     \midrule
-    \ac{CMOR}izer & CRU & \autocite{Harris2014} \\
+    \acs{CMOR}izer & CRU & \autocite{Harris2014} \\
     & CT2019 & \autocite{Jacobson2020} \\
@@
 reproduces large parts of the climate model evaluation performed in chapter 9
-of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
+of the \acs{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
 recipe, the two-panel figure 9.42 has been added, which shows the relationship
@@
 
-The implemented \ac{CMOR}ization scripts (also referred to as
-\emph{\ac{CMOR}izers}) allow the processing of arbitrary observational data
+The implemented \acs{CMOR}ization scripts (also referred to as
+\emph{\acs{CMOR}izers}) allow the processing of arbitrary observational data
 with the \ac{ESMValTool}. For this, the raw observational datasets are
 reformatted in such way that they adhere to the \ac{CMOR} standards. Since the
-\ac{CMOR}ization scripts are publicly available like any other code in the
+\acs{CMOR}ization scripts are publicly available like any other code in the
 \ac{ESMValTool} and \ac{ESMValCore} repositories, the entire scientific
-community can use them to \ac{CMOR}ize the observational data. A complete list
-of all eleven newly implemented \ac{CMOR}izers including the main reference for
-the corresponding observational datasets is shown in
+community can use them to \acs{CMOR}ize the observational data. A complete list
+of all eleven newly implemented \acs{CMOR}izers including the main reference
+for the corresponding observational datasets is shown in
 \cref{tab:03:changes_to_esmvaltool}. The majority of these datasets are used

### 2021-04-07 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Fixed use of acronyms followed by non-whitespace

@@
 released every day (as of \TheMonth{}). In light of the upcoming international
-climate assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of
+climate assessment of the \acs{IPCC}'s \acs{AR}6, the evaluation of
 policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6

### 2021-04-07 ch06_paper_gpp/paper_gpp.tex
Fixed use of acronyms followed by non-whitespace

@@
 \acs{CMIP}5 \acs{RCP}8.5 scenario at the end of the \nth{21} century with a new
-\ac{ML}-based weighting technique and appropriate observations. The new
+\acs{ML}-based weighting technique and appropriate observations. The new
 two-step approach utilizes aspects of emergent constraints and the \ac{MDER}
@@
 also select diagnostics as predictor variables $\bm{x}$ that are related to
-\ac{GPP}-relevant processes (see \cref{tab:06:predictors} for detailed physical
-explanations of the links between the predictors and \ac{GPP}).
+\acs{GPP}-relevant processes (see \cref{tab:06:predictors} for detailed
+physical explanations of the links between the predictors and \ac{GPP}).
 
@@
 trajectories of \ac{GPP} without this being fully encoded in today's
-observational products. Moreover, all \ac{CO2}-related effects are only
+observational products. Moreover, all \acs{CO2}-related effects are only
 implicitly incorporated in our approach by the target period we are using and

### 2021-04-07 ch07_conclusion/conclusion.tex
Fixed use of acronyms followed by non-whitespace

@@
 which is primarily driven by the \ac{CO2} fertilization. In the second step, a
-\ac{ML}-based climate model weighting approach is used to further constrain the
-gridded \ac{GPP} based on present-day predictors that are relevant for the
+\acs{ML}-based climate model weighting approach is used to further constrain
+the gridded \ac{GPP} based on present-day predictors that are relevant for the
 simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach is mathematically
@@
 A promising way forward to reduce uncertainties in multi-model climate
-projections is to apply the new flexible \ac{ML}-based climate model weighting
+projections is to apply the new flexible \acs{ML}-based climate model weighting
 approach introduced in \cref{ch:06:paper_gpp} to other target variables. While
@@
 of today's climate and which could serve as a possible baseline for the new
-\ac{ML}-based weighting approach. An exciting and valuable addition to the
+\acs{ML}-based weighting approach. An exciting and valuable addition to the
 method could be causal inference \autocite{Nowack2020, Runge2019}. In the
@@
 supercomputers has the potential for enormous improvements in climate modeling
-and analysis \autocite{Reichstein2019}. Apart from the presented \ac{ML}-based
+and analysis \autocite{Reichstein2019}. Apart from the presented \acs{ML}-based
 weighting approach, further possible applications of \ac{AI} and \ac{ML} in
-climate science involve for example \ac{ML}-based parameterizations that are
+climate science involve for example \acs{ML}-based parameterizations that are
 learned from high-resolution climate models \autocite{Gentine2018, Rasp2018},
-\ac{ML}-based analysis and prediction of forcing patterns \autocite{Barnes2019,
-  Mansfield2020}, or learning of entire \acp{ESM} from observational products
-\autocite{Geer2021}. Therefore, the foundation for innovative and
-groundbreaking research in climate science for the near future has been laid,
-which will help to further understand the Earth system and fight one of the
-greatest challenges for humankind today: climate change.
+\acs{ML}-based analysis and prediction of forcing patterns
+\autocite{Barnes2019, Mansfield2020}, or learning of entire \acp{ESM} from
+observational products \autocite{Geer2021}. Therefore, the foundation for
+innovative and groundbreaking research in climate science for the near future
+has been laid, which will help to further understand the Earth system and fight
+one of the greatest challenges for humankind today: climate change.

### 2021-04-07 utils/search_in_tex.py

@@ -27,6 +27,8 @@ python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}ref.{0,12}\{.{25}'
 
 python search_in_tex.py '(?i)\\caption\[.*?\]\{.*?\\label' | egrep '\\ac\{'
 
+python utils/search_in_tex.py '.{20}\\ac\{[^\}]+\}[^\s,\.\)\}\:]+?.{20}'
+
 """
 
 

### 2021-04-07 ch03_esmvaltool/esmvaltool.tex
Fixed further acronyms

@@
 peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
-After a brief overview of the \acs{ESMValTool}'s structure
+After a brief overview of the \ac{ESMValTool}'s structure
 (\cref{sec:03:esmvaltool}), this \namecref{ch:03:esmvaltool} presents all these
@@
     \cmidrule(l){2-3}
-    & co2s & Atmospheric \ac{CO2} concentration at
-    surface \\
+    & co2s & Atmospheric \acs{CO2} concentration at surface \\
     \cmidrule(l){2-3}
@@
     & NDP & \autocite{Gibbs2006} \\
-    & Scripps-\ac{CO2} & \autocite{Keeling2005} \\
+    & Scripps-\acs{CO2} & \autocite{Keeling2005} \\
     \bottomrule
@@
 reproduces large parts of the climate model evaluation performed in chapter 9
-of the \acs{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
+of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
 recipe, the two-panel figure 9.42 has been added, which shows the relationship

### 2021-04-07 end_appendix/appendix.tex
Standardized list of abbreviations
Fixed further acronyms

@@
 \numberwithin{table}{section}
-\renewcommand{\thesection}{\Alph{section}}
+\renewcommand*{\thesection}{\Alph{section}}
 
@@
     AIRS & hur, hus & \textcite{Aumann2003} \\
-    AMSRE \ac{SST} & tos & \textcite{AMSRE2011} \\
+    AMSRE \acs{SST} & tos & \textcite{AMSRE2011} \\
     CERES-EBAF & rsdt, rsut, rsutcs & \textcite{Loeb2018} \\

### 2021-04-07 utils/search_in_tex.py

@@ -16,18 +16,14 @@
 
 Examples
 --------
-python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}cite\{.{25}'
-python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}cite\{.{25}'
+python search_in_tex.py '(?i).{25}[\.\)]\s*\\.{0,12}cite\{.{25}'
+python search_in_tex.py '(?i).{25}[\.\)]\s*\\.{0,12}ac.{0,12}\{.{25}'
+python search_in_tex.py '(?i).{25}[\.\)]\s*\\.{0,12}ref.{0,12}\{.{25}'
 
-python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}ac.{0,12}\{.{25}'
-python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}ac.{0,12}\{.{25}'
+python search_in_tex.py '(?i)\\caption\[.*?\]\{.*?\\label' | egrep -iC 999 '\\ac\{'
+python search_in_tex.py '(?i)\\begin\{tabular.*?\}.*?\\end\{tabular.*?\}' | egrep -iC 999 '\\ac\{'
 
-python search_in_tex.py '(?i).{25}\.\s*\\.{0,12}ref.{0,12}\{.{25}'
-python search_in_tex.py '(?i).{25}\)\s*\\.{0,12}ref.{0,12}\{.{25}'
-
-python search_in_tex.py '(?i)\\caption\[.*?\]\{.*?\\label' | egrep '\\ac\{'
-
-python utils/search_in_tex.py '.{20}\\ac\{[^\}]+\}[^\s,\.\)\}\:]+?.{20}'
+python search_in_tex.py '.{20}\\[aA]{1}c[fl]?[p]?\{[^\}]+\}[^\s,\.\)\}\:;'']+?.{20}'
 
 """
 

### 2021-04-08 02_abstract/abstract.tex
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 
-A precise quantification of climate change is crucial in order to assess
-optimal mitigation and adaptation strategies. \AcfpAbstract{ESM}, which are
+A precise quantification of climate change is crucial to assess optimal
+mitigation and adaptation strategies. \AcfpAbstract{ESM}, which are
 state-of-the-art climate models that allow numerical simulations of the complex
@@
 \acs{CMIP}6, but may in some cases also be due to spurious statistical
-relationships or a too small number of models in the ensemble that the emergent
+relationships or a too-small number of models in the ensemble that the emergent
 constraint was originally derived from. The corresponding best estimates of
@@
 corresponding change in the constraint predictors, suggesting the emergence of
-new feedback processes rather than changes in strength of those previously
+new feedback processes rather than changes in the strength of those previously
 dominant. The results support previous studies concluding that emergent
@@
 alternative approach based on \acfAbstract{ML} is introduced. Since this new
-technique relies on a large number of data points in order to train the
-\acs{ML} algorithm, the scalar climate sensitivity expressed as \acs{ECS} or
-\acs{TCR} is not an appropriate target variable. Therefore, \acfAbstract{GPP}
-as a process that contributes to climate sensitivity is studied as an
-alternative. \acs{GPP} is the largest flux of the terrestrial carbon uptake and
-slows down global warming by removing \acs{CO2} from the atmosphere. In this
-analysis, an existing emergent constraint on \acs{CO2} fertilization is
-combined with a \acs{ML} approach to constrain the spatial variations of
-multi-model \acs{GPP} projections. In the first step of the two-step approach,
-observed changes in the \acs{CO2} seasonal cycle at \acl{KUM} are used to
-constrain the global mean \acs{GPP} at the end of the \nth{21} century
-(\range{2091}{2100}) in \acl{RCP} 8.5 simulations with \acsp{ESM} participating
-in \acs{CMIP}5 to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the
-unconstrained model range $\rangeunit{156}{247}{\GtCyr{}}$. In a second step, a
-\acs{ML} model is used to constrain gridded future absolute \acs{GPP} and
-gridded fractional \acs{GPP} change in two independent approaches. For this,
-observational data is fed into the \acs{ML} algorithm that has been trained on
-\acs{CMIP}5 data to learn relationships between present-day physically-relevant
-diagnostics and the target variable. In a leave-one-model-out \acl{CV}
-approach, the \acs{ML} model shows superior performance to the \acs{CMIP}5
-\acs{MMM}. The new approach predicts a higher \acs{GPP} increase in high
-latitudes and a lower \acs{GPP} increase in regions closer to the equator.
+technique relies on a large number of data points to train the \acs{ML}
+algorithm, the scalar climate sensitivity expressed as \acs{ECS} or \acs{TCR}
+is not an appropriate target variable. Therefore, \acfAbstract{GPP} as a
+process that contributes to climate sensitivity is studied as an alternative.
+\acs{GPP} is the largest flux of the terrestrial carbon uptake and slows down
+global warming by removing \acs{CO2} from the atmosphere. In this analysis, an
+existing emergent constraint on \acs{CO2} fertilization is combined with a
+\acs{ML} approach to constrain the spatial variations of multi-model \acs{GPP}
+projections. In the first step of the two-step approach, observed changes in
+the \acs{CO2} seasonal cycle at \acl{KUM} are used to constrain the global mean
+\acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in \acl{RCP}
+8.5 simulations with \acsp{ESM} participating in \acs{CMIP}5 to
+$\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the unconstrained model range
+$\rangeunit{156}{247}{\GtCyr{}}$. In a second step, a \acs{ML} model is used to
+constrain gridded future absolute \acs{GPP} and gridded fractional \acs{GPP}
+change in two independent approaches. For this, observational data is fed into
+the \acs{ML} algorithm that has been trained on \acs{CMIP}5 data to learn
+relationships between present-day physically relevant diagnostics and the
+target variable. In a leave-one-model-out \acl{CV} approach, the \acs{ML} model
+shows superior performance to the \acs{CMIP}5 \acs{MMM}. The new approach
+predicts a higher \acs{GPP} increase in high latitudes and a lower \acs{GPP}
+increase in regions closer to the equator.

### 2021-04-08 ch01_introduction/introduction.tex
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 and further refinement of the output of multi-model climate projections is
-necessary in order to reduce associated uncertainties. A state-of-the-art
-technique for this is the \emph{emergent constraints} method, which uses a
-physically-based inter-model relationship between an observable quantity of the
-Earth system and a target variable to reduce uncertainties in the target
-variable with observations \autocite{Allen2002}. An alternative approach is the
-weighting of climate models based on their performance (\ie{} the distance of
-one model to observational products) and interdependence (\ie{} the distance of
-one model to other climate models) \autocite{Knutti2017a}. These techniques
-form the baseline for the new analyses and results presented in this thesis,
-which partly utilize methods from a new emerging research field in climate
-sciences: \ac{AI} and \ac{ML}.
+necessary to reduce associated uncertainties. A state-of-the-art technique for
+this is the \emph{emergent constraints} method, which uses a physically based
+inter-model relationship between an observable quantity of the Earth system and
+a target variable to reduce uncertainties in the target variable with
+observations \autocite{Allen2002}. An alternative approach is the weighting of
+climate models based on their performance (\ie{} the distance of one model to
+observational products) and interdependence (\ie{} the distance of one model to
+other climate models) \autocite{Knutti2017a}. These techniques form the
+baseline for the new analyses and results presented in this thesis, which
+partly utilize methods from a new emerging research field in climate sciences:
+\ac{AI} and \ac{ML}.
 

### 2021-04-08 ch02_scientific_background/scientific_background.tex
Fixed use of acronyms followed by non-whitespace
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 cell, but still play an important role in the overall climate system by
-reflecting incoming and outgoing radiation \autocite{Boucher2013}. In order to
+reflecting incoming and outgoing radiation \autocite{Boucher2013}. To
 reasonably approximate these subgrid-scale processes, a concept called
@@
 is crucial for the various research groups developing \acp{ESM} all around the
-world in order to obtain comparable output and to facilitate model analysis.
-For this reason, the \ac{WGCM} of the \ac{WCRP} initiated the \acf{CMIP} in
-1995, with the objective to \enquote{better understand past, present and future
-  climate changes arising from natural, unforced variability or in response to
-  changes in radiative forcing in a multi-model context} \autocite{WCRP2020}.
-One major element of \ac{CMIP} is to establish common standards, coordination,
-infrastructure, and documentation in order to facilitate the distribution of
-climate model output \autocite{Eyring2016, Juckes2020}.
+world to obtain comparable output and to facilitate model analysis. For this
+reason, the \ac{WGCM} of the \ac{WCRP} initiated the \acf{CMIP} in 1995, with
+the objective to \enquote{better understand past, present and future climate
+  changes arising from natural, unforced variability or in response to changes
+  in radiative forcing in a multi-model context} \autocite{WCRP2020}. One major
+element of \ac{CMIP} is to establish common standards, coordination,
+infrastructure, and documentation to facilitate the distribution of climate
+model output \autocite{Eyring2016, Juckes2020}.
 
@@
 of future forcing (\acp{GHG}, aerosols, land use changes, \etc{}) humankind
-might take. Examples for these are the \ac{SSP}-based experiments given by
+might take. Examples for these are the \acs{SSP}-based experiments given by
 \ac{ScenarioMIP} that include a variety of different scenarios from a
@@
 assessment of climate sensitivity is essential for a precise quantification of
-the human-made climate change in order to determine optimal mitigation and
-adaptation strategies.
+the human-made climate change to determine optimal mitigation and adaptation
+strategies.
 
@@
 oldest metrics describing climate change. However, in practice this traditional
-definition of is not always useful. Running fully-coupled \acp{ESM} into
+definition of is not always useful. Running fully coupled \acp{ESM} into
 equilibrium is computationally expensive as it would require thousands of model
@@
 of the atmospheric \ac{CO2} concentration, not in terms of \ac{CO2} emissions.
-Nevertheless, in order to assess policy-relevant metrics like the allowable
-fossil fuel emissions to meet particular warming targets, for example the $1.5
+Nevertheless, to assess policy-relevant metrics like the allowable fossil fuel
+emissions to meet particular warming targets, for example the $1.5
 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015}, it is
@@
 relationship (red line in \cref{fig:02:emergent_constraint}) is a robust and
-physically-interpretable inter-model relationship between a target variable $y$
+physically interpretable inter-model relationship between a target variable $y$
 related to the future climate and an observable $x$ of the past or present-day
@@
     Schematic illustration of the emergent constraint approach. Basis of every
-    emergent constraint is a robust and physically-interpretable emergent
+    emergent constraint is a robust and physically interpretable emergent
     relationship (red line) between a target variable $y$ (\eg{} Earth system

### 2021-04-08 ch03_esmvaltool/esmvaltool.tex
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 
-In order to answer the \namecrefs{enum:01:question_1} of this thesis posed in
+To address the \namecrefs{enum:01:question_1} of this thesis posed in
 \cref{sec:01:key_science_questions}, a reliable and efficient tool to read,
@@
 In this thesis, \emph{land/sea fraction weighting} is applied to \ac{GPP}
-fields given by observational products and models in order to compute the
-corresponding global mean results (see \cref{ch:06:paper_gpp}). The
-\emph{trend} preprocessor calculates the linear trend of a variable along an
-arbitrary coordinate. The linear trend is defined as the slope of an ordinary
-least squares linear regression of the variable against the selected
-coordinate. For example, this can be used to calculate the historical trend of
-the \ac{GSAT}, which is used as a predictor for \ac{ECS} in several
-recently-published emergent constraints \autocite{JimenezdelaCuesta2019,
-  Nijsse2020, Tokarska2020}.
+fields given by observational products and models to compute the corresponding
+global mean results (see \cref{ch:06:paper_gpp}). The \emph{trend} preprocessor
+calculates the linear trend of a variable along an arbitrary coordinate. The
+linear trend is defined as the slope of an ordinary least squares linear
+regression of the variable against the selected coordinate. For example, this
+can be used to calculate the historical trend of the \ac{GSAT}, which is used
+as a predictor for \ac{ECS} in several recently published emergent constraints
+\autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}.
 

### 2021-04-08 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

--- 2021-04-06 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
@@
 
-In order to reduce uncertainties in multi-model climate projections, a first
-important step is the assessment of the desired target variables in the
-corresponding climate model ensemble. This is particularly relevant when a new
-generation of \acp{ESM} is published that includes considerable modifications
-compared to the respective predecessor model versions. Currently, we are in a
-situation like this with the \acs{CMIP}6 ensemble, for which new data is still
-released every day (as of \TheMonth{}). In light of the upcoming international
-climate assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of
-policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6
-models is crucial since they provide vital information about the future climate
-of planet Earth. As the \ac{ESMValTool} is an excellent tool that allows for a
-quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
-on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), it is used here to
-assess and analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare
-it to corresponding results of predecessor model generations and international
-climate assessments. This work is already published in two scientific papers:
-\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
-author of this thesis contributed their figure 8 (a bar chart showing the
-\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their
-figure 10 (map plots showing cloud-related feedback parameters for the
-\acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
-\cref{fig:04:cloud_feedback_parameters}), code to create these figures with the
-\ac{ESMValTool}, and text for the manuscript (in particular a section about the
-evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author of this thesis
-contributed all figures and tables, code to produce these with the
-\ac{ESMValTool}, and text for the manuscript. This
+To reduce uncertainties in multi-model climate projections, a first important
+step is the assessment of the desired target variables in the corresponding
+climate model ensemble. This is particularly relevant when a new generation of
+\acp{ESM} is published that includes considerable modifications compared to the
+respective predecessor model versions. Currently, we are in a situation like
+this with the \acs{CMIP}6 ensemble, for which new data is still released every
+day (as of \TheMonth{}). In light of the upcoming international climate
+assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of policy-relevant
+metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6 models is crucial
+since they provide vital information about the future climate of planet Earth.
+As the \ac{ESMValTool} is an excellent tool that allows for a quick and robust
+evaluation of new \ac{CMIP} data as soon as it gets published on the \ac{ESGF}
+servers (see \cref{sec:03:esmvaltool}), it is used here to assess and analyze
+\ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to corresponding
+results of predecessor model generations and international climate assessments.
+This work is already published in two scientific papers: \textcite{Bock2020}
+and \textcite{Meehl2020}. For \textcite{Bock2020}, the author of this thesis
+contributed their figure 8 (a bar chart showing the \ac{ECS} for several
+\acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their figure 10 (map plots
+showing cloud-related feedback parameters for the \acs{CMIP}5 and \acs{CMIP}6
+\acp{MMM}; see \cref{fig:04:cloud_feedback_parameters}), code to create these
+figures with the \ac{ESMValTool}, and text for the manuscript (in particular a
+section about the evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author
+of this thesis contributed all figures and tables, code to produce these with
+the \ac{ESMValTool}, and text for the manuscript. This
 \namecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
@@
 Various improvements of the underlying physical, biological, and chemical
-processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
-represent the coupled Earth system in more detail. Since many of these
-processes directly influence the models' \ac{ECS} \autocite{Forster2020},
-determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
-highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
-possible reason is the consideration of aerosol-cloud interactions. New
-prognostic aerosol schemes added to some \acs{CMIP}6 models that include
-aerosol-cloud interactions could have produced overly large negative radiative
-forcing, which then required a stronger model response to \ac{GHG} forcing in
-order to reproduce the observed historical temperature trend. In fact,
-\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
-forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
-present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
-have also been found for previous climate model generations
-\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
-over time, these relations cannot be used to tune \ac{ECS} based on the models'
-responses to aerosols over the different periods of the \nth{20} century
-\autocite{Dittus2020}. Further aerosol-cloud interactions that are relevant in
-the context of high climate sensitivity in \acs{CMIP}6 are found for example in
-the \ac{HadGEM} version 3. The new aerosol scheme in this model suppresses a
-strong negative cloud feedback over the Southern Ocean that is found in the
-model's predecessor versions \autocite{BodasSalcedo2019}. The Southern Ocean is
-one of the most pristine aerosol regions in the world \autocite{Hamilton2014},
-where aerosols predominantly originate from natural sources like marine
-emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM} versions show a
-large reduction in cloud droplet size with warming over the Southern Ocean,
-which increases the cloud albedo and thus produces a negative shortwave cloud
-feedback. On the contrary, in \acs{HadGEM}3, this negative feedback is
-completely suppressed or even positive due to a small increase of the cloud
-droplet size with warming, which eventually increases \ac{ECS} due to an
-increased net cloud feedback \autocite{BodasSalcedo2019}.
+processes have been introduced to the \acs{CMIP}6 \acp{ESM} to represent the
+coupled Earth system in more detail. Since many of these processes directly
+influence the models' \ac{ECS} \autocite{Forster2020}, determining reasons for
+the increased \ac{ECS} in the \acs{CMIP}6 ensemble is highly non-trivial. As
+discussed in detail by \textcite{Meehl2020}, one possible reason is the
+consideration of aerosol-cloud interactions. New prognostic aerosol schemes
+added to some \acs{CMIP}6 models that include aerosol-cloud interactions could
+have produced overly large negative radiative forcing, which then required a
+stronger model response to \ac{GHG} forcing to reproduce the observed
+historical temperature trend. In fact, \textcite{Meehl2020} find a weak
+inter-model relationship between the aerosol forcing and \ac{ECS} for some
+\acs{CMIP}6 models, with larger negative present-day aerosol forcing associated
+with higher \ac{ECS}. Such relationships have also been found for previous
+climate model generations \autocite{Kiehl2007, Forster2013}. However, due to
+the varying aerosol forcing over time, these relations cannot be used to tune
+\ac{ECS} based on the models' responses to aerosols over the different periods
+of the \nth{20} century \autocite{Dittus2020}. Further aerosol-cloud
+interactions that are relevant in the context of high climate sensitivity in
+\acs{CMIP}6 are found for example in the \ac{HadGEM} version 3. The new aerosol
+scheme in this model suppresses a strong negative cloud feedback over the
+Southern Ocean that is found in the model's predecessor versions
+\autocite{BodasSalcedo2019}. The Southern Ocean is one of the most pristine
+aerosol regions in the world \autocite{Hamilton2014}, where aerosols
+predominantly originate from natural sources like marine emissions of gaseous
+dimethyl sulfide. Earlier \ac{HadGEM} versions show a large reduction in cloud
+droplet size with warming over the Southern Ocean, which increases the cloud
+albedo and thus produces a negative shortwave cloud feedback. On the contrary,
+in \acs{HadGEM}3, this negative feedback is completely suppressed or even
+positive due to a small increase of the cloud droplet size with warming, which
+eventually increases \ac{ECS} due to an increased net cloud feedback
+\autocite{BodasSalcedo2019}.
 

### 2021-04-08 ch05_paper_ecs/paper_ecs.tex
Fixed further acronyms
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 of the newly included processes. Thus, it is vital to take other sources of
-information into account in order to assess the credibility of \ac{ESM}
-results.
+information into account to assess the credibility of \ac{ESM} results.
 
@@
 constraints have been developed over the years, which we now evaluate on the
-\acs{CMIP}6 ensemble in order to assess whether they still hold for these new
-climate models, \ie{} whether they have skill to reduce uncertainties in
-\ac{ECS} in the \acs{CMIP}6 ensemble. In total we assess eleven emergent
-constraints on \ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out
-of which the most are related to cloud feedbacks since these constitute the
-most important source of uncertainty for \ac{ECS} \commentcite{Boucher2013,
-  Flato2013}{see also \cref{subsec:02:climate_feedbacks}}. Since all of the
-evaluated emergent constraints have been derived from the \acs{CMIP}3 or
-\acs{CMIP}5 ensemble, the \acs{CMIP}6 models offer an exciting possibility to
-perform out-of-sample testing using the previously unseen data.
+\acs{CMIP}6 ensemble to assess whether they still hold for these new climate
+models, \ie{} whether they have skill to reduce uncertainties in \ac{ECS} in
+the \acs{CMIP}6 ensemble. In total we assess eleven emergent constraints on
+\ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out of which the
+most are related to cloud feedbacks since these constitute the most important
+source of uncertainty for \ac{ECS} \commentcite{Boucher2013, Flato2013}{see
+  also \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
+constraints have been derived from the \acs{CMIP}3 or \acs{CMIP}5 ensemble, the
+\acs{CMIP}6 models offer an exciting possibility to perform out-of-sample
+testing using the previously unseen data.
 
@@
     \makecell{\tabitem \emph{cl} (CloudSat/CALIPSO) \\ \tabitem \emph{tos}
-      (AMSRE \ac{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
+      (AMSRE \acs{SST}) \\ \tabitem \emph{wap} (ERA-Interim)} \\
     \bottomrule
@@
 this analysis. For all emergent constraints, we use the historical simulations
-of \acs{CMIP}5 and \acs{CMIP}6 in order to ensure maximum agreement with the
+of \acs{CMIP}5 and \acs{CMIP}6 to ensure maximum agreement with the
 observational data. If necessary, the historical simulation of \acs{CMIP}5 is
@@
 (like the new \acs{CMIP}6 ensemble), since the emergent relationship appeared
-just by chance and not because of a physically-based mechanism.
+just by chance and not because of a physically based mechanism.
 

### 2021-04-08 ch06_paper_gpp/paper_gpp.tex
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 non-linear relationships between the predictors and the target variable. Since
-the new technique relies on a large number of data points in order to train the
-\ac{ML} algorithm, \ac{ECS} as a scalar climate sensitivity metric is not an
+the new technique relies on a large number of data points to train the \ac{ML}
+algorithm, \ac{ECS} as a scalar climate sensitivity metric is not an
 appropriate target variable. Thus, this analysis does not focus on reducing
@@
 which is calculated from the 10-year mean \ac{GPP} at the end of the \nth{21}
-century in the emission-driven fully-coupled \acs{RCP}8.5 simulations
+century in the emission-driven fully coupled \acs{RCP}8.5 simulations
 (\range{2091}{2100}) and the 10-year mean \ac{GPP} at the end of the \nth{20}
@@
 $\bm{x}$ of the historical climate (for every grid cell/month) are fed into the
-trained \ac{GBRT} model in order to get a spatially constrained projection of
-\ac{GPP} (see black points in \cref{fig:06:mlr_concept}{\color{BrownRed} b}).
-Similar to the \ac{MDER} algorithm, our approach assumes that there exists an
-inter-model relationship between the predictors and the target variable which
-also holds for the true climate. This assumption may seem weak at first glance,
-but compared to the assumption of traditional weighting approaches (models that
-are better in simulating the present-day climate are also better in simulating
-the future climate and vice versa), the \ac{MDER} assumption is more reasonable
-and allows correcting models instead of just weighing
+trained \ac{GBRT} model to get a spatially constrained projection of \ac{GPP}
+(see black points in \cref{fig:06:mlr_concept}{\color{BrownRed} b}). Similar to
+the \ac{MDER} algorithm, our approach assumes that there exists an inter-model
+relationship between the predictors and the target variable which also holds
+for the true climate. This assumption may seem weak at first glance, but
+compared to the assumption of traditional weighting approaches (models that are
+better in simulating the present-day climate are also better in simulating the
+future climate and vice versa), the \ac{MDER} assumption is more reasonable and
+allows correcting models instead of just weighing
 \commentcite{Karpechko2013}{see \cref{subsec:02:mder}}. Following other
@@
 evaluation of the predictive power by calculating the \ac{RMSEP}. The whole
-process is repeated for every climate model of the ensemble in order to get a
+process is repeated for every climate model of the ensemble to get a
 distribution of \acp{RMSEP}. Since this approach is not limited to \ac{GBRT},
@@
 emergent constraint can also be applied to the global mean \ac{GPP} change if
-the emergent relationship holds. In order to test this, we first use the
-idealized \acs{CMIP}5 simulations that were analyzed by \textcite{Wenzel2016}
-to calculate the emergent constraint (\cref{fig:06:step1:a}). The relationship
-is statistically significant at a $5 \unit{\%}$ significance level ($R^2 =
-0.79$, $p = 0.007$) and predicts a constrained global mean \acs{GPP} change of
+the emergent relationship holds. To test this, we first use the idealized
+\acs{CMIP}5 simulations that were analyzed by \textcite{Wenzel2016} to
+calculate the emergent constraint (\cref{fig:06:step1:a}). The relationship is
+statistically significant at a $5 \unit{\%}$ significance level ($R^2 = 0.79$,
+$p = 0.007$) and predicts a constrained global mean \acs{GPP} change of
 $\pmrangeunit{30}{9}{\%}$, which is consistent with the findings of
@@
     \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
-    emission-driven fully-coupled \acs{RCP}8.5 simulations and the 10-year mean
+    emission-driven fully coupled \acs{RCP}8.5 simulations and the 10-year mean
     \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in the
-    emission-driven fully-coupled historical run (see \cref{eq:06:f}). In
+    emission-driven fully coupled historical run (see \cref{eq:06:f}). In
     contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
@@
 climate-carbon feedback \autocite{Gregory2009}, the \ac{CO2} fertilization
-effect is still the dominant driver of \ac{GPP} change in the fully-coupled
+effect is still the dominant driver of \ac{GPP} change in the fully coupled
 simulations \autocite{Huntzinger2017}. This is also supported by
 \textcite{Wenzel2016}, who show that the \ac{GPP} increase due to rising
-atmospheric \ac{CO2} concentrations in the fully-coupled historical simulations
+atmospheric \ac{CO2} concentrations in the fully coupled historical simulations
 is proportional to the same quantity in the \ac{1BGC} run. Additional lines of
@@
 \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}: We establish a
-relationship between process-oriented, physically-relevant diagnostics and
+relationship between process-oriented, physically relevant diagnostics and
 future projections and then utilize this to project today's observed conditions
@@
 gridded observation-based data, the \ac{GBRT} algorithm can be applied to any
-variable of interest if physically-relevant diagnostics that influence the
+variable of interest if physically relevant diagnostics that influence the
 target variable are known. This opens a wide range of possibilities for

### 2021-04-08 ch07_conclusion/conclusion.tex
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 
-In order to reduce this large range of \ac{ECS} in the latest generation of
-climate  models, already-published emergent constraints that have been derived
-from models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5
-are evaluated for their skill in the \acs{CMIP}6 ensemble. Emergent constraints
-use a physically-based inter-model relationship between an observable quantity
-of the Earth system and a target variable to reduce uncertainties in the target
+To reduce this large range of \ac{ECS} in the latest generation of climate
+models, already-published emergent constraints that have been derived from
+models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5 are
+evaluated for their skill in the \acs{CMIP}6 ensemble. Emergent constraints use
+a physically based inter-model relationship between an observable quantity of
+the Earth system and a target variable to reduce uncertainties in the target
 variable with observations \autocite{Allen2002}. In total eleven emergent
@@
 This work is already published in \textcite{Schlund2020}. Since the new
-technique relies on a large number of data points in order to train the \ac{ML}
+technique relies on a large number of data points to train the \ac{ML}
 algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is

### 2021-04-08 end_abbreviations/abbreviations.tex
Updated license information in every file
Cleaned tex files (hyphens, acronyms, etc.)
Ran spellchecker on tex files
Standardized list of abbreviations
Fixed hyphen-usage in compound modifiers and removed all appearances of 'in order to'

@@
 %                            Doctoral Dissertation                            %
-%                               Manuel Schlund                                %
-%                                  (c) 2021                                   %
+%                            (c) by Manuel Schlund                            %
 %=============================================================================%
-%                                Abbreviations                                %
+
+%=============================================================================%
+% This work is licensed under a
+% Creative Commons Attribution 4.0 International License.
+%
+% You should have received a copy of the license along with this
+% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
 %=============================================================================%
@@
 
-\begingroup
-\setstretch{0.5}
-\begin{acronym}
-  \acro{1BGC}[1\%BGC]{biogeochemically-coupled simulation where the %
+\begin{acronym}[xxxxxxxxxxxxxx] % The x's define the label column width
+  \acro{1BGC}[1\%BGC]{biogeochemically coupled simulation where the %
     atmospheric \ch{CO2} concentration increases by $1 \unit{\%}$ per year}
@@
   \acro{CV}{cross-validation}
-  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
+  \acro{DECK}{Diagnostic, Evaluation, and Characterization of Klima}
   \acro{DJF}{December-January-February}
   \acro{ECS}{effective climate sensitivity}
-  \acro{ENSO}{El Niño-Southern Oscillation}
+  \acro{ENSO}{El Niño--Southern Oscillation}
   \acro{ESGF}{Earth System Grid Federation}
@@
   \acro{NEP}{net ecosystem production}
-  \acro{netCDF}{Network Common Data Form}
+  \acro{NetCDF}{Network Common Data Form}
   \acro{NPP}{net primary production}
@@
 \end{acronym}
-\endgroup

### 2021-04-08 utils/search_in_tex.py

@@ -25,6 +25,10 @@ python search_in_tex.py '(?i)\\begin\{tabular.*?\}.*?\\end\{tabular.*?\}' | egre
 
 python search_in_tex.py '.{20}\\[aA]{1}c[fl]?[p]?\{[^\}]+\}[^\s,\.\)\}\:;'']+?.{20}'
 
+python search_in_tex.py '.{20}ly-.{20}'
+
+python search_in_tex.py '.{20}\s+[Aa][n]?\s+\\[Aa]c.{20}'
+
 """
 
 

### 2021-04-08 02_abstract/abstract.tex
Standardized use of indefinite articles with acronyms

@@
 global warming by removing \acs{CO2} from the atmosphere. In this analysis, an
-existing emergent constraint on \acs{CO2} fertilization is combined with a
+existing emergent constraint on \acs{CO2} fertilization is combined with an
 \acs{ML} approach to constrain the spatial variations of multi-model \acs{GPP}
@@
 $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the unconstrained model range
-$\rangeunit{156}{247}{\GtCyr{}}$. In a second step, a \acs{ML} model is used to
-constrain gridded future absolute \acs{GPP} and gridded fractional \acs{GPP}
+$\rangeunit{156}{247}{\GtCyr{}}$. In a second step, an \acs{ML} model is used
+to constrain gridded future absolute \acs{GPP} and gridded fractional \acs{GPP}
 change in two independent approaches. For this, observational data is fed into

### 2021-04-08 ch01_introduction/introduction.tex
Standardized use of indefinite articles with acronyms

@@
 simulations with only a prescribed change in the atmospheric \ac{CO2}
-concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
+concentration (\eg{} an instantaneous doubling of \ac{CO2} or \iac{CO2}
 increase of $1 \unit{\%}$ per year) as well as more realistic projections that
@@
 \ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
-response of the system to a \ac{CO2} doubling. In \acs{AR}5, the assessed range
+response of the system to \iac{CO2} doubling. In \acs{AR}5, the assessed range
 of both quantities is large with $\rangeunit{1.5}{4.5}{K}$ for \ac{ECS} and

### 2021-04-08 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Standardized use of indefinite articles with acronyms

@@
 illustrated in \cref{fig:04:gregory_regression_cmip6_mmm} for the \acs{CMIP}6
-\ac{MMM}, which yields an \ac{ECS} of $3.74 \unit{K}$ when all 150 years of the
+\ac{MMM}, which yields \iac{ECS} of $3.74 \unit{K}$ when all 150 years of the
 run are used. However, similar to the \acs{CMIP}5 ensemble, the exact value of
@@
 years of the simulation gives a significantly lower \ac{ECS} of $3.31 \unit{K}$
-than using only the last 130 years of the simulation, which gives an \ac{ECS}
-of $4.05 \unit{K}$. As thoroughly describes in \cref{subsec:02:ecs}, the reason
+than using only the last 130 years of the simulation, which gives \iac{ECS} of
+$4.05 \unit{K}$. As thoroughly describes in \cref{subsec:02:ecs}, the reason
 for this is the state and time dependence of the climate feedback parameter,
@@
 this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third (14
-models) have an \ac{ECS} above $4.5 \unit{K}$. In addition to the analysis
+models) have \iac{ECS} above $4.5 \unit{K}$. In addition to the analysis
 presented here, several modeling centers independently confirmed high \ac{ECS}
@@
   Wyser2020}. A list of possible reasons for the increased climate sensitivity
-in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
-the shortwave cloud feedback over the Southern Ocean is given in
+in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes
+in the shortwave cloud feedback over the Southern Ocean is given in
 \cref{sec:04:possible_reasons_high_ecs_cmip6}.

### 2021-04-08 ch05_paper_ecs/paper_ecs.tex
Standardized use of indefinite articles with acronyms

@@
 $\pmrangeunit{3.75}{1.11}{K}$ for \acs{CMIP}6 ($R^2 < 0.01$). The original
-publication does not specify an \ac{ECS} range. For \acs{CMIP}6, the emergent
+publication does not specify \iac{ECS} range. For \acs{CMIP}6, the emergent
 constraint shows a much lower statistical significance ($p = 0.6791$) than for
@@
 initial mixing linking it to cloud feedbacks and thus also to \ac{ECS}. For
-this constraint, we derive an \ac{ECS} $66 \unit{\%}$ confidence range of
+this constraint, we derive \iac{ECS} $66 \unit{\%}$ confidence range of
 $\pmrangeunit{3.42}{0.65}{K}$ for \acs{CMIP}5 ($R^2 = 0.41$) and
@@
 reference data, we use observed \acl{PR} data for the years \range{1986}{2005}
-from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66 \unit{\%}$
+from GPCP \autocite{Adler2003}. We calculate \iac{ECS} $66 \unit{\%}$
 confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5 ($R^2 =
@@
 ($R^2 = 0.05$) and $\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 =
-0.62$). In their original publication, \textcite{Zhai2015} find an \ac{ECS}
-range of $\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination
-of \acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance,
-the results of the ZHA constraints are somewhat surprising: although
-\acs{CMIP}5 data (in combination with \acs{CMIP}3 data) has successfully been
-used in the original publication, our approach finds that the statistical
-significance of the emergent relationship is much higher in the unseen
-\acs{CMIP}6 ensemble ($p < 0.0001$) than in the previously available
-\acs{CMIP}5 ensemble ($p = 0.2567$). The ZHA constraint is the only emergent
-constraint analyzed here that shows this extreme behavior (only one other
-constraint, SHS, shows a slightly higher significance in \acs{CMIP}6; all other
-constraints show lower significances in \acs{CMIP}6). The reason for the
-erratic skill in \acs{CMIP}5 is the set of climate models used. For our
-analysis, we use eleven additional \acs{CMIP}5 models that have not been used
-in the original publication (ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m,
-CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR, and
-MPI-ESM-P). Due to a lack of publicly available data, the model
-\acs{CESM}1-CAM5 that is used in the original publication is not included in
-our analysis. The effect of choosing different subsets of \acs{CMIP}5 models on
-the emergent relationship is illustrated in \cref{fig:05:zha}. Using the
-original \acs{CMIP}5 models from the original publication gives a considerably
-higher correlation ($R^2 = 0.38$) than using all available \acs{CMIP}5 models
-($R^2 = 0.05$). This result shows a strong dependency of this emergent
-constraint on the subset of climate models used. Nonetheless, the performance
-on \acs{CMIP}6 models is, surprisingly, the best of all the constraints and
-much better than on either subset of \acs{CMIP}5 models.
+0.62$). In their original publication, \textcite{Zhai2015} find \iac{ECS} range
+of $\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination of
+\acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance, the
+results of the ZHA constraints are somewhat surprising: although \acs{CMIP}5
+data (in combination with \acs{CMIP}3 data) has successfully been used in the
+original publication, our approach finds that the statistical significance of
+the emergent relationship is much higher in the unseen \acs{CMIP}6 ensemble ($p
+< 0.0001$) than in the previously available \acs{CMIP}5 ensemble ($p =
+0.2567$). The ZHA constraint is the only emergent constraint analyzed here that
+shows this extreme behavior (only one other constraint, SHS, shows a slightly
+higher significance in \acs{CMIP}6; all other constraints show lower
+significances in \acs{CMIP}6). The reason for the erratic skill in \acs{CMIP}5
+is the set of climate models used. For our analysis, we use eleven additional
+\acs{CMIP}5 models that have not been used in the original publication
+(ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m, CCSM4, GFDL-ESM2G, GFDL-ESM2M,
+IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR, and MPI-ESM-P). Due to a lack of
+publicly available data, the model \acs{CESM}1-CAM5 that is used in the
+original publication is not included in our analysis. The effect of choosing
+different subsets of \acs{CMIP}5 models on the emergent relationship is
+illustrated in \cref{fig:05:zha}. Using the original \acs{CMIP}5 models from
+the original publication gives a considerably higher correlation ($R^2 = 0.38$)
+than using all available \acs{CMIP}5 models ($R^2 = 0.05$). This result shows a
+strong dependency of this emergent constraint on the subset of climate models
+used. Nonetheless, the performance on \acs{CMIP}6 models is, surprisingly, the
+best of all the constraints and much better than on either subset of
+\acs{CMIP}5 models.
 

### 2021-04-08 ch06_paper_gpp/paper_gpp.tex
Standardized use of indefinite articles with acronyms

@@
 two-step approach utilizes aspects of emergent constraints and the \ac{MDER}
-technique in combination with a \ac{ML} algorithm. In the first step, we apply
+technique in combination with \iac{ML} algorithm. In the first step, we apply
 an existing emergent constraint on \ac{CO2} fertilization \autocite{Wenzel2016}
@@
 cycle amplitude, the fractional \ac{GPP} change at the time of \ac{CO2}
-doubling relative to pre-industrial conditions in a \ac{1BGC} is constrained.
+doubling relative to pre-industrial conditions in \iac{1BGC} is constrained.
 The predictor in this study is the sensitivity of the \ac{CO2} seasonal cycle
@@
   \cref{subsec:02:model_weighting}}. For this, a single climate model is
-removed from the multi-model ensemble and considered as being the true climate.
-Our statistical model is then fitted on all remaining climate models and a
-prediction for the \enquote{true} model is created. This allows a simple
-evaluation of the predictive power by calculating the \ac{RMSEP}. The whole
-process is repeated for every climate model of the ensemble to get a
+removed from the multi-model ensemble and considered as being the true
+climate. Our statistical model is then fitted on all remaining climate models
+and a prediction for the \enquote{true} model is created. This allows a
+simple evaluation of the predictive power by calculating the \ac{RMSEP}. The
+whole process is repeated for every climate model of the ensemble to get a
 distribution of \acp{RMSEP}. Since this approach is not limited to \ac{GBRT},
-it allows a simple comparison of the predictive skill of different statistical
-models. We compare the \ac{GBRT} model to the simple \acs{CMIP}5 \acf{MMM} and
-a \ac{LASSO} model \autocite{Tibshirani1996}. The \ac{LASSO} model is an
-extended linear regression model that regularizes the least squares solution
-with an L$1$ penalty term. In contrast to the ordinary least squares
-regression, the \ac{LASSO} model promotes sparse solutions (\ie{} only a few
-non-zero coefficients/weights), which acts as an intrinsic feature selection
-and provides an effective way to combat overfitting induced by collinearity.
+it allows a simple comparison of the predictive skill of different
+statistical models. We compare the \ac{GBRT} model to the simple \acs{CMIP}5
+\acf{MMM} and \iac{LASSO} model \autocite{Tibshirani1996}. The \ac{LASSO}
+model is an extended linear regression model that regularizes the least
+squares solution with an L$1$ penalty term. In contrast to the ordinary least
+squares regression, the \ac{LASSO} model promotes sparse solutions (\ie{}
+only a few non-zero coefficients/weights), which acts as an intrinsic feature
+selection and provides an effective way to combat overfitting induced by
+collinearity.
 
@@
 using the leave-one-model-out \ac{CV} approach (see
-\cref{subsec:06:uncertainty_and_skill}). This allows us to create an \ac{RMSEP}
+\cref{subsec:06:uncertainty_and_skill}). This allows us to create \iac{RMSEP}
 distribution for each statistical approach, where the different points of the
@@
 our approach applies different corrections (in sign and magnitude) for
-different regions. On the global domain, our approach predicts a \ac{GPP} at
-the end of the century of $169 \unit{\GtCyr{}}$, which is consistent with the
+different regions. On the global domain, our approach predicts \iac{GPP} at the
+end of the century of $169 \unit{\GtCyr{}}$, which is consistent with the
 global constraint from step 1 of $\pmrangeunit{171}{12}{\GtCyr{}}$. In summary,
@@
 The global distribution of the fractional \ac{GPP} change for the different
-statistical models is shown in \cref{fig:06:step2b_results}. All panels show a
-\ac{GPP} increase over the \nth{21} century in the \acs{RCP}8.5 scenario for
+statistical models is shown in \cref{fig:06:step2b_results}. All panels show
+\iac{GPP} increase over the \nth{21} century in the \acs{RCP}8.5 scenario for
 almost all regions of the globe. Compared to the unweighted \acs{CMIP}5
@@
 emergent constraint is not undisputed and can be replaced with another emergent
-constraint in future studies. In the second step, a \ac{ML} approach is used to
+constraint in future studies. In the second step, \iac{ML} approach is used to
 further constrain the gridded \ac{GPP} based on process-based present-day

### 2021-04-08 02_abstract/abstract.tex
Spell- and grammer-checked chapter 1

@@
 
-In a first step, the climate sensitivity (\ie{} the temperature response of the
+In a first step, climate sensitivity (\ie{} the temperature response of the
 climate system to an external forcing) is evaluated in the latest generation of
-\acsp{ESM} from \acs{CMIP}6. For the \acfAbstract{ECS}, which is defined as the
-equilibrium temperature response that follows a doubling of the atmospheric
-\acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74 \unit{K}$ and a
-multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These values are
-higher than in any previous \acs{CMIP} ensemble before. Moreover, a third of
-the analyzed \acs{CMIP}6 models exceed the upper bound of the likely \acs{ECS}
-range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC} Fifth
-\acl{AR} (\acs{AR}5) from 2013. Similarly, the transient response of the
+\acsp{ESM} from \acs{CMIP}6. For the \acfAbstract{ECS}, which is an estimation
+of the equilibrium temperature response that follows a doubling of the
+atmospheric \acfAbstract{CO2} concentration, a \acfAbstract{MMM} of $3.74
+\unit{K}$ and a multi-model range of $\rangeunit{1.8}{5.6}{K}$ are found. These
+values are higher than in any previous \acs{CMIP} ensemble before. Moreover, a
+third of the analyzed \acs{CMIP}6 models exceed the upper bound of the likely
+\acs{ECS} range of $\rangeunit{1.5}{4.5}{K}$ assessed by the \acfAbstract{IPCC}
+Fifth \acl{AR} (\acs{AR}5) from 2013. Similarly, the transient response of the
 climate system to a doubling of \acs{CO2}, also known as \acfAbstract{TCR},
@@
 again higher than the likely range assessed in \acs{AR}5 of
-$\rangeunit{1.0}{2.5}{K}$. Possible reasons for the increased climate
-sensitivity in many \acs{CMIP}6 models are the addition of new prognostic
-aerosol schemes that include aerosol-cloud interactions and changes in the
-microphysical representation of mixed-phase clouds. The changes in the
-mixed-cloud representation reduce the strong negative shortwave cloud phase
-change feedback over the Southern Ocean that is present in climate models from
-previous \acs{CMIP} generations.
+$\rangeunit{1.0}{2.5}{K}$. Possible reasons for increased climate sensitivity
+in many \acs{CMIP}6 models are the addition of new prognostic aerosol schemes
+that include aerosol-cloud interactions and changes in the microphysical
+representation of mixed-phase clouds. The changes in the mixed-cloud
+representation reduce the strong negative shortwave cloud phase change feedback
+over the Southern Ocean that is present in climate models from previous
+\acs{CMIP} generations.
 
@@
 technique relies on a large number of data points to train the \acs{ML}
-algorithm, the scalar climate sensitivity expressed as \acs{ECS} or \acs{TCR}
-is not an appropriate target variable. Therefore, \acfAbstract{GPP} as a
-process that contributes to climate sensitivity is studied as an alternative.
-\acs{GPP} is the largest flux of the terrestrial carbon uptake and slows down
-global warming by removing \acs{CO2} from the atmosphere. In this analysis, an
+algorithm, scalar climate sensitivity expressed as \acs{ECS} or \acs{TCR} is
+not an appropriate target variable. Therefore, \acfAbstract{GPP} as a process
+that contributes to climate sensitivity is studied as an alternative. \acs{GPP}
+is the largest flux of the terrestrial carbon uptake and slows down global
+warming by removing \acs{CO2} from the atmosphere. In this analysis, an
 existing emergent constraint on \acs{CO2} fertilization is combined with an

### 2021-04-08 ch01_introduction/introduction.tex
Spell- and grammer-checked chapter 1

@@
 processes that may alter the Earth's energy budget. Human influence on the
-climate system is clear: Ever increasing emissions of \acp{GHG} since the end
+climate system is clear: Ever-increasing emissions of \acp{GHG} since the end
 of the pre-industrial era largely driven by economic and population growth
@@
 the planet's surface is partly absorbed by the corresponding \ac{GHG}
-molecules through an excitation of their corresponding vibrational modes and
+molecules through excitation of their corresponding vibrational modes and
 re-emitted isotropically. This process traps energy near the surface and
@@
 (which corresponds to a warming) and its largest contribution is caused by
-the increase in the atmospheric concentration of \ac{CO2} due to anthropogenic
-fossil fuel emissions since the mid-\nth{18} century \autocite{IPCC2014}. Apart
-from their physical warming effect on the climate, carbon-based \acp{GHG} like
-\ac{CO2} also directly influence the global carbon cycle, an important
-biogeochemical cycle of the Earth. A crucial flux of the carbon cycle is the
-\emph{\ac{GPP}}, which describes the carbon uptake of the terrestrial biosphere
-due to photosynthesis. Since this land carbon uptake absorbs about $30
-\unit{\%}$ of the anthropogenic \ac{CO2} emissions in today's climate
-\autocite{Friedlingstein2020}, this process substantially slows down global
-warming and directly contributes to the magnitude of climate sensitivity. Other
-important anthropogenic drivers of climate change are the emission of aerosols,
-and land use/land cover changes. Apart from anthropogenic drivers, there are
-also natural processes which impact the climate system like changes in the
-solar activity or volcanic eruptions. However, there is clear evidence that
-these natural drivers alone cannot explain the observed climate change
-\autocite{Haustein2017}.
+the increase in the atmospheric concentration of \ac{CO2} due to
+anthropogenic fossil fuel emissions since the mid-\nth{18} century
+\autocite{IPCC2014}. Apart from their physical warming effect on the climate,
+carbon-based \acp{GHG} like \ac{CO2} also directly influence the global
+carbon cycle, an important biogeochemical cycle of the Earth. A crucial flux
+of the carbon cycle is the \emph{\ac{GPP}}, which describes the carbon uptake
+of the terrestrial biosphere due to photosynthesis. Since this land carbon
+uptake absorbs about $30 \unit{\%}$ of the anthropogenic \ac{CO2} emissions
+in today's climate \autocite{Friedlingstein2020}, this process substantially
+slows down global warming and directly contributes to the magnitude of
+climate sensitivity. Other important anthropogenic drivers of climate change
+are the emission of aerosols, and land use/land cover changes. Apart from
+anthropogenic drivers, there are also natural processes that impact the climate
+system like changes in solar activity or volcanic eruptions. However, there is
+clear evidence that these natural drivers alone cannot explain the observed
+climate change \autocite{Haustein2017}.
 
@@
 consider different future scenarios (\eg{} a fossil fuel--based future or a
-scenario that is based on a sustainable development). Multi-model projections
+scenario that is based on sustainable development). Multi-model projections
 from \ac{CMIP} show a large inter-model range in many variables that are
 relevant for climate change \autocite{Collins2013, Flato2013}. A crucial and
-policy-relevant example for this is the climate sensitivity, which refers to
-the change in the \ac{GSAT} that results from a change in the radiative
-forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
+policy-relevant example for this is climate sensitivity, which refers to the
+change in the \ac{GSAT} that results from a change in the radiative forcing.
+Common metrics for this are the \emph{\ac{ECS}}, which describes the
 equilibrium response of the climate system after a doubling of the atmospheric
@@
 results \autocite{Flato2013}. For this reason, a careful statistical evaluation
-and further refinement of the output of multi-model climate projections is
+and further refinement of the output of multi-model climate projections are
 necessary to reduce associated uncertainties. A state-of-the-art technique for
@@
 
-The aim of this thesis is to reduce uncertainties in multi-model climate
+This thesis aims to reduce to reduce uncertainties in multi-model climate
 projections with observations by addressing the following three
@@
 all involved contributors. However, unless stated otherwise, all content from
-these publications (text, figures, and tables) shown in this thesis originate
+these publications (text, figures, and tables) shown in this thesis originates
 from the author of this thesis. A detailed list of contributions to these
@@
 used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
-over the contributions made to the \ac{ESMValTool}, an open-source software for
-the analysis of \acp{ESM}. These contributions helped improving the routine
-evaluation of \acp{ESM} which is useful for the entire scientific community and
-led to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
-  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:papers_ecs_tcr_assessment}
-covers the assessment of climate sensitivity metrics like the \ac{ECS} or
-\ac{TCR} in the latest generation of \acp{ESM} from \acs{CMIP}6. This work is
-already published in two scientific publications \autocite{Bock2020,
-  Meehl2020}. Since the \ac{ECS} and \ac{TCR} are considerably higher in this
-new climate model generation, \cref{ch:05:paper_ecs} describes the assessment
-of emergent constraints on the \ac{ECS} for these \acs{CMIP}6 models and
-compares these to results derived from \acs{CMIP}5 models. The content of this
-\namecref{ch:05:paper_ecs} is published in \emph{Earth System Dynamics}
-\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
-reduce uncertainties in multi-dimensional (gridded) multi-model projections of
-the future climate with observations based on \ac{ML}. As an example, the
-method is applied to \ac{GPP} at the end of the \nth{21} century, which is
-already published in the \emph{Journal of Geophysical Research: Biogeosciences}
-\autocite{Schlund2020}. Finally, \cref{ch:07:conclusion} provides a summary of
-the results of this thesis and gives an outlook of possible future work.
+of the contributions made to the \ac{ESMValTool}, an open-source software
+package for the analysis of \acp{ESM}. These contributions helped to improve
+the routine evaluation of \acp{ESM} which is useful for the entire scientific
+community and led to co-authorship in four peer-reviewed studies
+\autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
+\Cref{ch:04:papers_ecs_tcr_assessment} covers the assessment of climate
+sensitivity metrics like the \ac{ECS} or \ac{TCR} in the latest generation of
+\acp{ESM} from \acs{CMIP}6. This work is already published in two scientific
+publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and \ac{TCR}
+are considerably higher in this new climate model generation,
+\cref{ch:05:paper_ecs} describes the assessment of emergent constraints on the
+\ac{ECS} for these \acs{CMIP}6 models and compares these to results derived
+from \acs{CMIP}5 models. The content of this \namecref{ch:05:paper_ecs} is
+published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.
+\Cref{ch:06:paper_gpp} focuses on a new method to reduce uncertainties in
+multi-dimensional (gridded) multi-model projections of the future climate with
+observations based on \ac{ML}. As an example, the method is applied to \ac{GPP}
+at the end of the \nth{21} century, which is already published in the
+\emph{Journal of Geophysical Research: Biogeosciences} \autocite{Schlund2020}.
+Finally, \cref{ch:07:conclusion} provides a summary of the results of this
+thesis and gives an outlook of possible future work.

### 2021-04-08 ch02_scientific_background/scientific_background.tex
Spell- and grammer-checked chapter 1

@@
 An important policy-relevant metric of the Earth system that can be assessed
-with numerical climate model simulations is the climate sensitivity. Climate
+with numerical climate model simulations is climate sensitivity. Climate
 sensitivity refers to the change in \ac{GSAT} that results from a change in the
@@
   Zelinka2020}. This large uncertainty in the cloud feedback is a major reason
-for uncertainties in the climate sensitivity of climate models
+for uncertainties in climate sensitivity of climate models
 \autocite{Boucher2013, Flato2013}.
@@
 \end{equation}
-The factor of $2$ in the denominator accounts for the fact the traditional
+The factor of $2$ in the denominator accounts for the fact that the traditional
 equilibrium climate sensitivity is defined for an abrupt \ac{CO2} doubling,
@@
 
-A further metric describing the climate sensitivity is the \emph{\ac{TCR}}. In
+A further metric describing climate sensitivity is the \emph{\ac{TCR}}. In
 contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
@@
 \autocite{Allen2002} and the snow-albedo feedback \autocite{Hall2006}. Over the
-years, the climate sensitivity expressed by the \ac{ECS} has been a prominent
+years, climate sensitivity expressed by the \ac{ECS} has been a prominent
 target variable. Since cloud feedbacks are a major source of uncertainty for

### 2021-04-08 ch03_esmvaltool/esmvaltool.tex
Standardized use of indefinite articles with acronyms
Spell- and grammer-checked chapter 1

@@
 process, and evaluate climate model output and observational data is necessary.
-A valuable software that fits these criteria is the \ac{ESMValTool}. The
-\ac{ESMValTool} is an open-source community diagnostics and performance metrics
-tool for the routine evaluation of \ac{ESM} output, which notably facilitates
-the analysis of \ac{CMIP} models (\url{https://www.esmvaltool.org}). For this
-reason, all analyses presented in
+A valuable software package that fits these criteria is the \ac{ESMValTool}.
+The \ac{ESMValTool} is an open-source community diagnostics and performance
+metrics tool for the routine evaluation of \ac{ESM} output, which notably
+facilitates the analysis of \ac{CMIP} models
+(\url{https://www.esmvaltool.org}). For this reason, all analyses presented in
 \cref{ch:04:papers_ecs_tcr_assessment,ch:05:paper_ecs,ch:06:paper_gpp} of this
@@
 methods for the study presented in \cref{ch:06:paper_gpp}, in which the
-\ac{MDER} method is expanded with a \ac{ML} approach and applied to a carbon
+\ac{MDER} method is expanded with \iac{ML} approach and applied to a carbon
 cycle--related target variable (\ac{GPP}).
@@
 expected reciprocal relation. The output of \emph{recipe\_flato13ipcc.yml}
-provided useful input for the analysis of the climate sensitivity in
-\acs{CMIP}6 performed in \textcite{Meehl2020}, which is also presented in
+provided useful input for the analysis of climate sensitivity in \acs{CMIP}6
+performed in \textcite{Meehl2020}, which is also presented in
 \cref{ch:04:papers_ecs_tcr_assessment}.

### 2021-04-08 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Spell- and grammer-checked chapter 1

@@
 values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
-  Wyser2020}. A list of possible reasons for the increased climate sensitivity
-in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes
-in the shortwave cloud feedback over the Southern Ocean is given in
+  Wyser2020}. A list of possible reasons for increased climate sensitivity in
+the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
+the shortwave cloud feedback over the Southern Ocean is given in
 \cref{sec:04:possible_reasons_high_ecs_cmip6}.
@@
 this \namecref{sec:04:possible_reasons_high_ecs_cmip6} discusses possible
-reasons for the increased climate sensitivity in \acs{CMIP}6.
+reasons for increased climate sensitivity in \acs{CMIP}6.
 
@@
 
-Another possible reason for the high climate sensitivity in the \acs{CMIP}6
+Another possible reason for high climate sensitivity in the \acs{CMIP}6
 ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
@@
 \autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
-to cloud feedbacks immediately impact the climate models' climate sensitivity.
-The particular feedback mechanism that is thought to be relevant for the high
-\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
-Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5).
-This so-called \emph{cloud phase change feedback} is illustrated in
-\cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate warms,
-the predominantly ice clouds over the Southern Ocean in these climate models
-become liquid clouds \autocite{McCoy2015}. As a result, the cloud will get more
-reflective to the incoming solar radiation since a cloud consisting of smaller
-liquid droplets reflects more sunlight than a cloud consisting of larger ice
-crystals (for a fixed water content), which has a cooling effect on the climate
-system. In addition, predominantly liquid clouds tend to precipitate less than
-mixed-phase clouds (composed of liquid water and ice). This leads to a higher
-liquid water content of the cloud and a further amplification of the cooling
-effect. Overall, this forms a strong negative shortwave cloud feedback over the
-Southern Ocean in climate models from previous \ac{CMIP} generations.
+to cloud feedbacks immediately impact climate sensitivity of the climate
+models. The particular feedback mechanism that is thought to be relevant for
+the high \ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the
+Southern Ocean that are present in earlier \ac{CMIP} generations (\eg{}
+\acs{CMIP}5). This so-called \emph{cloud phase change feedback} is illustrated
+in \cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate
+warms, the predominantly ice clouds over the Southern Ocean in these climate
+models become liquid clouds \autocite{McCoy2015}. As a result, the cloud will
+get more reflective to the incoming solar radiation since a cloud consisting of
+smaller liquid droplets reflects more sunlight than a cloud consisting of
+larger ice crystals (for a fixed water content), which has a cooling effect on
+the climate system. In addition, predominantly liquid clouds tend to
+precipitate less than mixed-phase clouds (composed of liquid water and ice).
+This leads to a higher liquid water content of the cloud and a further
+amplification of the cooling effect. Overall, this forms a strong negative
+shortwave cloud feedback over the Southern Ocean in climate models from
+previous \ac{CMIP} generations.
 

### 2021-04-08 ch07_conclusion/conclusion.tex
Spell- and grammer-checked chapter 1

@@
 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from \acs{AR}5 of
-$\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased climate
+$\rangeunit{1.0}{2.5}{K}$. One possible reason for increased climate
 sensitivity in some \acs{CMIP}6 models is the addition of prognostic aerosol
@@
 technique relies on a large number of data points to train the \ac{ML}
-algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
-not an appropriate target variable. Therefore, this analysis does not focus on
+algorithm, scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is not
+an appropriate target variable. Therefore, this analysis does not focus on
 reducing uncertainties in climate sensitivity itself but rather on a selected
@@
 first step corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2},
-which is primarily driven by the \ac{CO2} fertilization. In the second step, a
+which is primarily driven by the \ac{CO2} fertilization. In the second step, an
 \acs{ML}-based climate model weighting approach is used to further constrain

### 2021-04-08 ch01_introduction/introduction.tex
Spell- and grammer-checked chapter 2

@@
 climate sensitivity. Other important anthropogenic drivers of climate change
-are the emission of aerosols, and land use/land cover changes. Apart from
-anthropogenic drivers, there are also natural processes that impact the climate
-system like changes in solar activity or volcanic eruptions. However, there is
-clear evidence that these natural drivers alone cannot explain the observed
-climate change \autocite{Haustein2017}.
+are the emission of aerosols and land-use/land-cover changes. Apart from
+anthropogenic drivers, there are also natural processes that impact the
+climate system like changes in solar activity or volcanic eruptions. However,
+there is clear evidence that these natural drivers alone cannot explain the
+observed climate change \autocite{Haustein2017}.
 
@@
 
-This thesis aims to reduce to reduce uncertainties in multi-model climate
-projections with observations by addressing the following three
+This thesis aims to reduce uncertainties in multi-model climate projections
+with observations by addressing the following three
 \namecrefs{enum:01:question_1}:

### 2021-04-08 ch02_scientific_background/scientific_background.tex
Spell- and grammer-checked chapter 2

@@
 the global carbon cycle (\cref{sec:02:esms}) are presented. Finally,
-state-of-the art techniques that can be used to reduce uncertainties in
+state-of-the-art techniques that can be used to reduce uncertainties in
 projections of the future climate with observations are shown
@@
 time steps that are typically about $30 \unit{min}$ long in modern-day global
-climate models. For each grid cell and time step, usually a single value per
-model variable is given. Examples for such variables provided by a climate
+climate models. For each grid cell and time step, a single value per model
+variable is usually given. Examples for such variables provided by a climate
 model are the atmospheric prognostic variables velocity (horizontal and
@@
     Historical evolution of coupled climate models over the last 45 years. In
-    early days, these models were so-called \acfp{AOGCM}, which only included
-    three components: the atmosphere, the land surface, and the ocean. Over the
-    time, the individual components grew in complexity and included a wider
-    range of processes (illustrated by the growing cylinders). Eventually, more
-    and more components (aerosols, carbon cycle, \etc{}) were added to the
-    coupled system, forming the modern \acfp{ESM}.
+    the early days, these models were so-called \acfp{AOGCM}, which only
+    included three components: the atmosphere, the land surface, and the ocean.
+    Over time, the individual components grew in complexity and included a
+    wider range of processes (illustrated by the growing cylinders).
+    Eventually, more and more components (aerosols, carbon cycle, \etc{}) were
+    added to the coupled system, forming the modern \acfp{ESM}.
     \IPCCReproducedFrom{Cubasch2013}{their figure 1.13}.
@@
 Parameterizations are necessary to simulate many processes of the Earth system,
-for example surface heat and moisture fluxes, moist convection, turbulent
+for example, surface heat and moisture fluxes, moist convection, turbulent
 mixing, and radiation \autocite{Holton2004}.
@@
 components to the coupled system. Examples of these are aerosols, the carbon
-cycle, a dynamic vegetation, atmospheric chemistry, and land ice (see
+cycle, dynamic vegetation, atmospheric chemistry, and land ice (see
 \cref{fig:02:esms_historical_evolution}). \Acp{AOGCM} coupled to these
@@
 the system (see \cref{sec:02:carbon_cycle} for details). Further examples
-include land use changes like the deforestation of tropical rainforests, which
-also directly influences several biogeochemical cycles (\eg{} carbon cycle,
+include land-use changes like the deforestation of tropical rainforests, which
+also directly influence several biogeochemical cycles (\eg{} carbon cycle,
 nitrogen cycle, phosphorus cycle, \etc{}) by altering respective sinks and
@@
 atmosphere \autocite{Friedlingstein2020}. The residual part is absorbed by the
-two other main carbon sinks of the planet, the terrestrial biosphere and the
+two other main carbon sinks of the planet: the terrestrial biosphere and the
 ocean. Since only atmospheric \ac{CO2} can act as \ac{GHG} by introducing an
@@
 world to obtain comparable output and to facilitate model analysis. For this
-reason, the \ac{WGCM} of the \ac{WCRP} initiated the \acf{CMIP} in 1995, with
-the objective to \enquote{better understand past, present and future climate
-  changes arising from natural, unforced variability or in response to changes
-  in radiative forcing in a multi-model context} \autocite{WCRP2020}. One major
-element of \ac{CMIP} is to establish common standards, coordination,
-infrastructure, and documentation to facilitate the distribution of climate
-model output \autocite{Eyring2016, Juckes2020}.
+reason, the \ac{WGCM} of the \ac{WCRP} initiated the \acf{CMIP} in 1995 to
+\enquote{better understand past, present and future climate changes arising
+  from natural, unforced variability or in response to changes in radiative
+  forcing in a multi-model context} \autocite{WCRP2020}. One major element of
+\ac{CMIP} is to establish common standards, coordination, infrastructure, and
+documentation to facilitate the distribution of climate model output
+\autocite{Eyring2016, Juckes2020}.
 
@@
 that simulate different possible futures \autocite{ONeill2016}. These
-experiments are based on the so-called \acp{SSP}, a set of alternative pathways
-of future societal development \autocite{ONeill2013, ONeill2017}. For each
-experiment, specific emissions and land use changes are calculated from the
-\acp{SSP} \autocite{Riahi2017} which are then used to force the global climate
-models. For \ac{ScenarioMIP}, five different \acp{SSP} are considered, ranging
-from \acs{SSP}1 (sustainability) to \acs{SSP}5 (fossil fuel--based
+experiments are based on the so-called \emph{\acp{SSP}}, a set of alternative
+pathways of future societal development \autocite{ONeill2013, ONeill2017}. For
+each experiment, specific emissions and land-use changes are calculated from
+the \acp{SSP} \autocite{Riahi2017} which are then used to force the global
+climate models. For \ac{ScenarioMIP}, five different \acp{SSP} are considered,
+ranging from \acs{SSP}1 (sustainability) to \acs{SSP}5 (fossil fuel--based
 development). Each \ac{SSP} is combined with a climate outcome (measured as
@@
 future with a radiative forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The
-two scenarios in the main category of \ac{ScenarioMIP}, the \emph{Tier 1}
-experiments, are the \acs{SSP}2-4.5 and \acs{SSP}3-7.0 scenarios. In contrast
-to the \ac{ScenarioMIP} experiments, the corresponding \acs{CMIP}5 counterparts
-\autocite{Taylor2012}, the so-called \acp{RCP}, only used the radiative forcing
-in 2100 as only dimension to describe the possible futures (\eg{} \acs{RCP}8.5,
-\acs{RCP}4.5, \acs{RCP}2.6, \etc{}).
+four scenarios in the main category of \ac{ScenarioMIP}, the \emph{Tier 1}
+experiments, are the \acs{SSP}1-2.6, \acs{SSP}2-4.5, \acs{SSP}3-7.0, and
+\acs{SSP}5-8.5 scenarios. In contrast to the \ac{ScenarioMIP} experiments, the
+corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012}, the so-called
+\emph{\acp{RCP}}, only used the radiative forcing in 2100 as single dimension
+to describe the possible futures (\eg{} \acs{RCP}8.5, \acs{RCP}4.5,
+\acs{RCP}2.6, \etc{}).
 
@@
 Simulations from climate model ensembles of \ac{CMIP} allow us to assess future
-climate change in a consistent and transparent way. Especially the
-\ac{ScenarioMIP} experiments can give valuable insights into possible
-developments of the Earth system by providing future \emph{climate
-  projections}. In contrast to climate predictions, climate projections run
-over multiple decades and depend upon the future scenario considered, which are
-based on assumptions that may or may not turn out to be correct. On the
-contrary, climate predictions are attempts to predict the actual evolution of
-the climate on much shorter time scales from seasons to years. Similar to any
-other scientific experiment, climate model projections suffer from associated
-uncertainties. There are three major sources of uncertainties in climate model
-projections we can distinguish: natural variability, climate response
-uncertainty, and emission uncertainty \autocite{Hawkins2009, Hawkins2010}.
-\Cref{fig:02:sources_of_uncertainty} shows these three sources for the
-projected global mean surface temperature anomaly over the \nth{21} century.
+climate change consistently and transparently. Especially the \ac{ScenarioMIP}
+experiments can give valuable insights into possible developments of the Earth
+system by providing future \emph{climate projections}. In contrast to climate
+predictions, climate projections run over multiple decades and depend upon the
+future scenario considered, which are based on assumptions that may or may not
+turn out to be correct. On the contrary, climate predictions are attempts to
+predict the actual evolution of the climate on much shorter time scales from
+seasons to years. Similar to any other scientific experiment, climate model
+projections suffer from associated uncertainties. There are three major sources
+of uncertainties in climate model projections we can distinguish: natural
+variability, climate response uncertainty, and emission uncertainty
+\autocite{Hawkins2009, Hawkins2010}. \Cref{fig:02:sources_of_uncertainty} shows
+these three sources for the projected global mean surface temperature anomaly
+over the \nth{21} century.
 
@@
   uncertainty}. This arises from the different possible trajectories in terms
-of future forcing (\acp{GHG}, aerosols, land use changes, \etc{}) humankind
+of future forcing (\acp{GHG}, aerosols, land-use changes, \etc{}) humankind
 might take. Examples for these are the \acs{SSP}-based experiments given by
@@
 to assess the climate impact of these different trajectories. Since the
-emission uncertainty strongly depends on the future development of the human
+emission uncertainty strongly depends on the future development of human
 society, it cannot be reduced by improving climate models. In contrast to
@@
 in which the increased surface temperature (caused by anthropogenic forcing)
-leads to an enhanced evaporation of water which increases the amount of water
+leads to enhanced evaporation of water which increases the amount of water
 vapor in the atmosphere. Since water vapor itself is a powerful \ac{GHG}, this
@@
 
-As sciences evolves, representations of already included processes can be
+As science evolves, representations of already included processes can be
 improved in climate models. Moreover, new geophysical and biogeochemical
@@
 \autocite{Zelinka2020}. On the other hand, the climate response uncertainty can
-decrease with additional model improvements and better understanding of the
+decrease with additional model improvements and a better understanding of the
 Earth system (see \cref{fig:02:sources_of_uncertainty}{\color{BrownRed} c}).
@@
 natural (changes in the solar activity, volcanic eruptions, \etc{}) or
-anthropogenic (emissions of \acp{GHG}, land use changes, \etc{}). Thus, the
+anthropogenic (emissions of \acp{GHG}, land-use changes, \etc{}). Thus, the
 assessment of climate sensitivity is essential for a precise quantification of
-the human-made climate change to determine optimal mitigation and adaptation
+human-made climate change to determine optimal mitigation and adaptation
 strategies.
@@
   ]{
-    Climate feedbacks and corresponding time scales. The different sings refer
+    Climate feedbacks and corresponding time scales. The different signs refer
     to the signs of the different feedbacks. Positive feedbacks amplify the
@@
     negative feedbacks diminish the effect of the external forcing (\eg{} the
-    longwave radiation feedback). An example for a feedback that might be
-    either positive or negative is the cloud feedback. The smaller box
-    highlights the large differences in time scales for the various feedbacks.
+    longwave radiation feedback). An example of a feedback that might be either
+    positive or negative is the cloud feedback. The smaller box highlights the
+    large differences in time scales for the various feedbacks.
     \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.
@@
 amplified or diminished by climate feedbacks. Thus, feedback processes play a
-crucial role determining the magnitude of climate sensitivity.
+crucial role in determining the magnitude of climate sensitivity.
 \Cref{fig:02:feedbacks_overview} shows an overview of important feedbacks in
@@
 
-An example for a positive feedback is the already mentioned \emph{water vapor
+An example of a positive feedback is the already mentioned \emph{water vapor
   feedback}. Being the primary \ac{GHG} in the Earth's atmosphere, water vapor
@@
 vapor on the climate system is described as a feedback mechanism and not as an
-external forcing \autocite{Myhre2013}. Basis of this feedback is the enhanced
-evaporation of water with increasing air temperatures. Each degree of warming
-allows the atmosphere to retain about $7 \unit{\%}$ more water vapor
+external forcing \autocite{Myhre2013}. The basis of this feedback is the
+enhanced evaporation of water with increasing air temperatures. Each degree of
+warming allows the atmosphere to retain about $7 \unit{\%}$ more water vapor
 \autocite{Myhre2013}, which closes the positive feedback loop by further
@@
 emissions) by a typical factor between $2$ and $3$, rendering water vapor a
-fundamental agent of climate change \autocite{Myhre2013}. An example for a
+fundamental agent of climate change \autocite{Myhre2013}. An example of a
 positive feedback that operates on longer time scales (several years) is the
-\emph{snow/ice albedo feedback}, in which the surface albedo decreases as
+\emph{snow/ice-albedo feedback}, in which the surface albedo decreases as
 highly reflective ice and snow surfaces melt with global warming, exposing the
@@
 In contrast to positive feedbacks, negative feedbacks diminish the effect of an
-external forcing. An example for this is the \emph{blackbody feedback} (also
+external forcing. An example of this is the \emph{blackbody feedback} (also
 known as \emph{Planck feedback} or \emph{longwave radiation feedback}), which
@@
 For some domains of the Earth system, feedbacks can be positive and/or
-negative, since a variety of different mechanisms are involved. An example for
+negative, since a variety of different mechanisms are involved. An example of
 this is the cloud feedback. Changes in the clouds induced by climate change can
@@
   Zelinka2020}. This large uncertainty in the cloud feedback is a major reason
-for uncertainties in climate sensitivity of climate models
+for uncertainties in the climate sensitivity of climate models
 \autocite{Boucher2013, Flato2013}.
@@
 decrease the atmospheric \ac{CO2} content and diminish global warming. An
-example for a positive contribution is the decreased solubility of \ac{CO2} in
+example of a positive contribution is the decreased solubility of \ac{CO2} in
 water in a warmer climate, which reduces the atmosphere-ocean \ac{CO2} flux and
@@
 incoming radiative flux at the \ac{TOA} and the net heat flux into the ocean
-are basically equal definitions of $N$ since nearly all of the Earth's heat
-capacity resides in the ocean \autocite{Gregory2009}. While $N \neq 0$, the
-climate system evolves; when $N = 0$ a new steady state has been reached.
+are equal definitions of $N$ since nearly all of the Earth's heat capacity
+resides in the ocean \autocite{Gregory2009}. While $N \neq 0$, the climate
+system evolves; when $N = 0$ a new steady state has been reached.
 
@@
 $\lambda_\mathtxt{WV}$ refers to the water vapor feedback,
-$\lambda_\mathtxt{Albedo}$ to the snow/ice albedo feedback, and
+$\lambda_\mathtxt{Albedo}$ to the snow/ice-albedo feedback, and
 $\lambda_\mathtxt{Cloud}$ to the cloud feedback. Thus, the overall climate
@@
 \autocite{Charney1979}, the equilibrium climate sensitivity is one of the
-oldest metrics describing climate change. However, in practice this traditional
-definition of is not always useful. Running fully coupled \acp{ESM} into
-equilibrium is computationally expensive as it would require thousands of model
-years \autocite{Rugenstein2020}.
+oldest metrics describing climate change. However, in practice, this
+traditional definition is not always useful. Running fully coupled \acp{ESM}
+into equilibrium is computationally expensive as it would require thousands of
+model years \autocite{Rugenstein2020}.
 
@@
 easily calculated from the external forcing $F$ and the climate feedback
-parameter $\lambda$. The steady state values for $F$ and $\lambda$ can be
+parameter $\lambda$. The steady-state values for $F$ and $\lambda$ can be
 estimated from a \nxcotwo{4} simulation that is not in equilibrium by
@@
     red, and orange colors correspond to the \acf{CESM} of the \acf{NCAR} for
-    different time periods of a long running simulation (see legend). The
+    different time periods of a long-running simulation (see legend). The
     initial years are simulated many times for different initial conditions.
@@
 \cref{fig:02:gregory_regression_different_years} shows, the slope in the
-Gregory regression is not constant, but rather decreases over time when a long
-running \nxcotwo{4} simulation with more than 1000 model years is considered.
-As a result, the climate feedback parameter $\lambda$ decreases over time,
-resulting in a higher \ac{ECS}. Major reasons for this are temperature
-dependencies of the feedbacks, atmospheric and oceanic adjustments over time,
-changing warming patterns over time, non-additive feedbacks, and dependencies
-on the type and magnitude of the external forcings \autocite{Knutti2017}. All
-in all, this demonstrates the limits of the linear feedback framework
-introduced in \cref{subsec:02:mathematical_framework_feedbacks} that is not
-capable of describing non-linear effects. A second major reason for the
-discrepancies between the equilibrium climate sensitivity and \ac{ECS} is the
-use of a \nxcotwo{4} instead of a \nxcotwo{2} simulation. The factor of $2$ in
-the denominator of \cref{eq:02:ecs} only partly compensates this difference
+Gregory regression is not constant but rather decreases over time when a
+long-running \nxcotwo{4} simulation with more than 1000 model years is
+considered. As a result, the climate feedback parameter $\lambda$ decreases
+over time, resulting in a higher \ac{ECS}. Major reasons for this are
+temperature dependencies of the feedbacks, atmospheric and oceanic adjustments
+over time, changing warming patterns over time, non-additive feedbacks, and
+dependencies on the type and magnitude of the external forcings
+\autocite{Knutti2017}. All in all, this demonstrates the limits of the linear
+feedback framework introduced in
+\cref{subsec:02:mathematical_framework_feedbacks} that is not capable of
+describing non-linear effects. A second major reason for the discrepancies
+between the equilibrium climate sensitivity and \ac{ECS} is the use of a
+\nxcotwo{4} instead of a \nxcotwo{2} simulation. The factor of $2$ in the
+denominator of \cref{eq:02:ecs} only partly compensates for this difference
 since the radiative forcing logarithmically depends on the atmospheric \ac{CO2}
@@
 
-However, despite these deficiencies the \ac{ECS} is still a practical estimate
+However, despite these deficiencies, the \ac{ECS} is still a practical estimate
 of the equilibrium climate sensitivity. With the help of climate models,
@@
 integrating climate models until a new steady state is reached. Nevertheless,
-for \acs{CMIP}6 long running simulations from the \ac{LongRunMIP}
+for \acs{CMIP}6 long-running simulations from the \ac{LongRunMIP}
 \autocite{Rugenstein2019} can be a promising way forward to estimate the true
@@
 In addition to the calculation of the external forcing $F$, the overall climate
-feedback parameter $\lambda$ and the \ac{ECS}, the Gregory regression can also
+feedback parameter $\lambda$, and the \ac{ECS}, the Gregory regression can also
 be used to estimate cloud-related feedback parameters. For this, the net
@@
 accurate as the deeper ocean warms up and cannot be applied to simulations with
-a steady state climate change in which $N \to 0$ \autocite{Gregory2009}. By
+a steady-state climate change in which $N \to 0$ \autocite{Gregory2009}. By
 applying the definition of \ac{TCR} (transient \ac{GSAT} change $\Delta T$ at
@@
 remains in the atmosphere \autocite{Friedlingstein2020}. The residual part is
-absorbed by the two other main carbon sinks of the planet, the terrestrial
+absorbed by the two other main carbon sinks of the planet: the terrestrial
 biosphere and the ocean. These exchange fluxes of carbon are part of the global
@@
 Nevertheless, to assess policy-relevant metrics like the allowable fossil fuel
-emissions to meet particular warming targets, for example the $1.5
+emissions to meet particular warming targets, for example, the $1.5
 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015}, it is
@@
 cycle through volcanic emissions of \ac{CO2}, chemical weathering, erosion, and
-sediment formation on the sea floor. These natural exchange fluxes between the
+sediment formation on the seafloor. These natural exchange fluxes between the
 slow and the fast domain are comparatively small ($< 0.3 \unit{\GtCyr{}}$) and
@@
 occurs nearly all year, the larger amount of vegetation in the Northern
-Hemisphere (due to the larger land mass) gives rise to a seasonal cycle of the
+Hemisphere (due to the larger landmass) gives rise to a seasonal cycle of the
 atmospheric \ac{CO2} concentration \autocite{Keeling1995}. This seasonal cycle
@@
 atmospheric \ac{CO2} concentration in the Northern Hemisphere winter (net
-\ac{CO2} flux into atmosphere due to respiration) and a minimum during the
+\ac{CO2} flux into the atmosphere due to respiration) and a minimum during the
 Northern Hemisphere summer (net \ac{CO2} flux into the land due to
@@
 and the ocean. Atmospheric \ac{CO2} is exchanged with the surface ocean through
-gas exchange, which is driven by the partial \ac{CO2} pressure difference
+a gas exchange, which is driven by the partial \ac{CO2} pressure difference
 between the air and the sea \autocite{Ciais2013}.
@@
 to increased dissolution of \ac{CO2} into the sea
-\autocite{Friedlingstein2020}. In the same time, the amount of carbon in the
+\autocite{Friedlingstein2020}. At the same time, the amount of carbon in the
 atmosphere reservoir increased with a rate of $5.1 \unit{\GtCyr{}}$, indicating
@@
 biosphere, the concentration-carbon feedback is connected to the \emph{\ac{CO2}
-  fertilization effect} \autocite{Walker2020}, that causes an increase of
+  fertilization effect} \autocite{Walker2020}, which causes an increase in
 photosynthesis rates when the atmospheric \ac{CO2} concentration increases,
-which in turns removes \ac{CO2} from the atmosphere, forming a negative
+which in turn removes \ac{CO2} from the atmosphere, forming a negative
 feedback. For the ocean, the concentration-carbon feedback is negative as well.
@@
 variety of different parameterizations for the different processes
-\autocite{Arora2020}. Examples for the photosynthesis (\ac{GPP}) are given by
+\autocite{Arora2020}. Examples for photosynthesis (\ac{GPP}) are given by
 \textcite{Farquhar1980} for \ch{C3} plants and \textcite{Collatz1992} for
 \ch{C4} plants. The terms \enquote{\ch{C3}} and \enquote{\ch{C4}} refer to the
-different metabolic pathways used in the carbon fixation during photosynthesis.
-In both cases, the enzyme \ac{RuBisCO} is an important catalyst for the
+different metabolic pathways used in carbon fixation during photosynthesis. In
+both cases, the enzyme \ac{RuBisCO} is an important catalyst for the
 corresponding chemical reactions. Both parameterizations utilize variables that
@@
 projected quantity from multi-model ensembles is to treat the arithmetic
-\ac{MMM} and the multi-model range as best estimate and uncertainty measure of
-this quantity \autocite{Collins2013}. This \emph{model democracy} approach
-basically assumes that all climate models are independent, equally plausible,
-distributed around reality, and that the projected multi-model range is
-representative for the uncertainty in the projected quantity
-\autocite{Knutti2017a}. However, since the \ac{CMIP} ensembles, sometimes
-referred to as \emph{ensembles of opportunity}, have not been designed to
-represent a true statistical sample of the reality composed of independent
-climate models \autocite{Tebaldi2007}, these assumptions do not hold in
-practice. The main reasons for this are that different climate models (even for
-different modeling institutions) share parts of their code
+\ac{MMM} as best estimate and the multi-model range as uncertainty measure
+\autocite{Collins2013}. This \emph{model democracy} approach assumes that all
+climate models are independent, equally plausible, distributed around reality,
+and that the projected multi-model range is representative for the uncertainty
+in the projected quantity \autocite{Knutti2017a}. However, since the \ac{CMIP}
+ensembles, sometimes referred to as \emph{ensembles of opportunity}, have not
+been designed to represent a true statistical sample of the reality composed of
+independent climate models \autocite{Tebaldi2007}, these assumptions do not
+hold in practice. The main reasons for this are that different climate models
+(even for different modeling institutions) share parts of their code
 \autocite{Abramowitz2019, Knutti2013}, that models do not equally well
@@
 requires two key components: an \emph{emergent relationship} and a
-corresponding observation of the real world \autocite{Eyring2019}. The emergent
-relationship (red line in \cref{fig:02:emergent_constraint}) is a robust and
-physically interpretable inter-model relationship between a target variable $y$
-related to the future climate and an observable $x$ of the past or present-day
-climate. Basis for the relationship is output of the different climate models
-of a multi-model ensemble (blue circles in \cref{fig:02:emergent_constraint}).
-Using an observation of $x$, this emergent relationship can then be used to
-derive an emergent constraint on $y$ (gray shaded area in
-\cref{fig:02:emergent_constraint}) that considers uncertainties in the emergent
-relationship itself (red shaded area in \cref{fig:02:emergent_constraint}) and
-uncertainties in the observation (blue shaded area in
-\cref{fig:02:emergent_constraint}).
+corresponding observation of the real world \autocite{Eyring2019}. The
+emergent relationship (red line in \cref{fig:02:emergent_constraint}) is a
+robust and physically interpretable inter-model relationship between a target
+variable $y$ related to the future climate and an observable $x$ of the past
+or present-day climate. The basis for the relationship is output from the
+different climate models of a multi-model ensemble (blue circles in
+\cref{fig:02:emergent_constraint}). Using an observation of $x$, this
+emergent relationship can then be used to derive an emergent constraint on
+$y$ (gray shaded area in \cref{fig:02:emergent_constraint}) that considers
+uncertainties in the emergent relationship itself (red shaded area in
+\cref{fig:02:emergent_constraint}) and uncertainties in the observation (blue
+shaded area in \cref{fig:02:emergent_constraint}).
 
@@
 \end{equation}
-Here, $\bar{x}$ indicates the arithmetic mean of $x$ over all climate models.
-Assuming Gaussian errors and a mean of $\hat{y}(x)$ (\ie{} the best estimate of
-the target variable $y$ is given by the regression line), \cref{eq:02:spe} can
-be used to define a conditional \acf{PDF} for predicting a target variable of
-$y$ given $x$:
+Here, $\bar{x}$ indicates the arithmetic mean of $x$ calculated over all
+climate models. Assuming Gaussian errors and a mean of $\hat{y}(x)$ (\ie{} the
+best estimate of the target variable $y$ is given by the regression line),
+\cref{eq:02:spe} can be used to define a conditional \acf{PDF} for predicting a
+target variable of $y$ given $x$:
 \begin{equation}
@@
 observable $x$. However, the observed value of $x$, called $x_0$, also has
-uncertainties associated with it (blue shaded are in
+uncertainties associated with it (blue shaded area in
 \cref{fig:02:emergent_constraint}). Assuming again a Gaussian distribution, the
@@
 statistical inference has been performed. Alternative methods that can be used
-include Bayesian frameworks \autocite{Renoult2020}, information theoretic
+include Bayesian frameworks \autocite{Renoult2020}, information-theoretic
 approaches based on the Kullback-Leibler divergence between the models'
@@
 data points from the different climate models are independent. As already noted
-in the beginning of \cref{sec:02:reducing_uncertainties}, this is not the case
+at the beginning of \cref{sec:02:reducing_uncertainties}, this is not the case
 for typical climate model ensembles, as some modeling groups provide output for
@@
 physical mechanism, an emergent constraint can be considered credible. The
-reason for this is the occurrence spurious relationships: Due to the large
+reason for this is the occurrence of spurious relationships: Due to the large
 number of possible observables provided by modern \acp{ESM} and the
@@
 Earth system have been published. Early studies tackled the hydrological cycle
-\autocite{Allen2002} and the snow-albedo feedback \autocite{Hall2006}. Over the
-years, climate sensitivity expressed by the \ac{ECS} has been a prominent
+\autocite{Allen2002} and the snow/ice-albedo feedback \autocite{Hall2006}. Over
+the years, climate sensitivity expressed by the \ac{ECS} has been a prominent
 target variable. Since cloud feedbacks are a major source of uncertainty for
@@
 \autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}. Emergent
-constraints are not only limited to physical processes, but can also be applied
+constraints are not only limited to physical processes but can also be applied
 to other domains, like the global carbon cycle \autocite{Cox2013,
@@
 
-A further technique to reduce uncertainties in climate model projections with
+Further techniques to reduce uncertainties in climate model projections with
 observations are model weighting schemes. Their basic idea is to abandon model
@@
 climate models. To address two major issues of model democracy (different
-climate models are not equally plausible and not independent; see beginning of
-\cref{sec:02:reducing_uncertainties}), \textcite{Knutti2017a} propose a
-weighting scheme based on climate model performance and interdependence with
-the following weights:
+climate models are not equally plausible and not independent),
+\textcite{Knutti2017a} propose a weighting scheme based on climate model
+performance and interdependence with the following weights:
 \begin{equation}
@@
 which are relevant for the projection of the target variable $y$. The reasoning
-for this choice of the diagnostics is similar to the one presented in the
-previous \namecref{subsec:02:model_weighting}: Weighting schemes should address
-the question \enquote{which climate model is adequate for predicting the target
+for this choice of diagnostics is similar to the one presented in the previous
+\namecref{subsec:02:model_weighting}: Weighting schemes should address the
+question \enquote{which climate model is adequate for predicting the target
   variable?} and not \enquote{which climate model is the best?}.

### 2021-04-08 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Spell- and grammer-checked chapter 2

--- 2021-04-08 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
@@
   Wyser2020}. A list of possible reasons for the increased climate sensitivity
-in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes
-in the shortwave cloud feedback over the Southern Ocean is given in
+in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
+the shortwave cloud feedback over the Southern Ocean is given in
 \cref{sec:04:possible_reasons_high_ecs_cmip6}.
@@
 \autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
-to cloud feedbacks immediately impact the climate models' climate sensitivity.
-The particular feedback mechanism that is thought to be relevant for the high
-\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
-Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5).
-This so-called \emph{cloud phase change feedback} is illustrated in
-\cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate warms,
-the predominantly ice clouds over the Southern Ocean in these climate models
-become liquid clouds \autocite{McCoy2015}. As a result, the cloud will get more
-reflective to the incoming solar radiation since a cloud consisting of smaller
-liquid droplets reflects more sunlight than a cloud consisting of larger ice
-crystals (for a fixed water content), which has a cooling effect on the climate
-system. In addition, predominantly liquid clouds tend to precipitate less than
-mixed-phase clouds (composed of liquid water and ice). This leads to a higher
-liquid water content of the cloud and a further amplification of the cooling
-effect. Overall, this forms a strong negative shortwave cloud feedback over the
-Southern Ocean in climate models from previous \ac{CMIP} generations.
+to cloud feedbacks immediately impact the climate models' climate
+sensitivities. The particular feedback mechanism that is thought to be relevant
+for the high \ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over
+the Southern Ocean that are present in earlier \ac{CMIP} generations (\eg{}
+\acs{CMIP}5). This so-called \emph{cloud phase change feedback} is illustrated
+in \cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate
+warms, the predominantly ice clouds over the Southern Ocean in these climate
+models become liquid clouds \autocite{McCoy2015}. As a result, the cloud will
+get more reflective to the incoming solar radiation since a cloud consisting of
+smaller liquid droplets reflects more sunlight than a cloud consisting of
+larger ice crystals (for a fixed water content), which has a cooling effect on
+the climate system. In addition, predominantly liquid clouds tend to
+precipitate less than mixed-phase clouds (composed of liquid water and ice).
+This leads to a higher liquid water content of the cloud and a further
+amplification of the cooling effect. Overall, this forms a strong negative
+shortwave cloud feedback over the Southern Ocean in climate models from
+previous \ac{CMIP} generations.
 

### 2021-04-08 ch03_esmvaltool/esmvaltool.tex
Spell- and grammer-checked chapter 3

@@
 applied to the input data. There are two classes of preprocessors: non-optional
-and optional ones. Examples for non-optional preprocessors are the input/output
-operations \emph{load} (loading input data from climate models and
+and optional preprocessors. Examples for non-optional preprocessors are the
+input/output operations \emph{load} (loading input data from climate models and
 observations) and \emph{save} (saving the preprocessed data), which utilize the
@@
 \emph{unit conversion}, and the calculation of \emph{multi-model statistics}. A
-further example is the \emph{variable derivation}, which can be used to derive
-non-\ac{CMOR} variables from the input data. The aim of the preprocessor is to
-facilitate the routine evaluation of climate models by providing a set of
-commonly used data operations. To reduce computation times, \ac{ESMValCore}
-allows parallel preprocessing of different datasets. However, since this leads
-to a higher memory usage, a compromise has to be found in practice to optimize
-the use of time and memory resources.
+further example is \emph{variable derivation}, which can be used to derive
+non-\ac{CMOR} variables from the input data. The preprocessing functions of
+\ac{ESMValCore} aim to facilitate the routine evaluation of climate models by
+providing a set of commonly used data operations. To reduce computation times,
+\ac{ESMValCore} allows parallel preprocessing of different datasets. However,
+since this leads to higher memory usage, a compromise has to be found in
+practice to optimize the use of time and memory resources.
 
@@
 As of \TheMonth{}, 40408 lines of code have been added and 21050 lines of code
-have been removed from to the GitHub repository of \ac{ESMValCore} by the
-author of this thesis. Apart from general improvements of the code base, these
-changes and additions mainly include new preprocessor functions that can be
-applied to the input data and derivation scripts for new non-\ac{CMOR}
-variables. \Cref{tab:03:changes_to_esmvalcore} shows a summary of these main
+have been removed from the GitHub repository of \ac{ESMValCore} by the author
+of this thesis. Apart from general improvements of the code base, these changes
+and additions mainly include new preprocessor functions that can be applied to
+the input data and derivation scripts for new non-\ac{CMOR} variables.
+\Cref{tab:03:changes_to_esmvalcore} shows a summary of these main
 contributions, which are partly published in the scientific documentation of
@@
     \cmidrule(l){2-3}
-    & co2s & Atmospheric \acs{CO2} concentration at surface \\
+    & co2s & Atmospheric \acs{CO2} concentration at the surface \\
     \cmidrule(l){2-3}
@@
 \enquote{per square meter of grid cell}. After the weighting, the grid cell
-areas can be used to integrate the flux-related variable over a desired region.
-In this thesis, \emph{land/sea fraction weighting} is applied to \ac{GPP}
-fields given by observational products and models to compute the corresponding
-global mean results (see \cref{ch:06:paper_gpp}). The \emph{trend} preprocessor
-calculates the linear trend of a variable along an arbitrary coordinate. The
-linear trend is defined as the slope of an ordinary least squares linear
-regression of the variable against the selected coordinate. For example, this
-can be used to calculate the historical trend of the \ac{GSAT}, which is used
-as a predictor for \ac{ECS} in several recently published emergent constraints
-\autocite{JimenezdelaCuesta2019, Nijsse2020, Tokarska2020}.
+areas can be used to integrate the flux-related variable over any desired
+region. In this thesis, \emph{land/sea fraction weighting} is applied to
+\ac{GPP} fields given by observational products and models to compute the
+corresponding global mean results (see \cref{ch:06:paper_gpp}). The
+\emph{trend} preprocessor calculates the linear trend of a variable along an
+arbitrary coordinate. The linear trend is defined as the slope of an ordinary
+least squares linear regression of the variable against the selected
+coordinate. For example, this can be used to calculate the historical trend of
+the \ac{GSAT}, which is used as a predictor for \ac{ECS} in several recently
+published emergent constraints \autocite{JimenezdelaCuesta2019, Nijsse2020,
+  Tokarska2020}.
 
@@
   \caption[
-    Illustration of the atmospheric \acs{CO2} concentration at surface
+    Illustration of the atmospheric \acs{CO2} concentration at the surface
     (\emph{co2}) for \acs{CMIP}6 models.
   ]{
-    (a) \acs{CMIP}6 \acf{MMM} of the atmospheric \acs{CO2} concentration at
+    (a) \acs{CMIP}6 \acf{MMM} of the atmospheric \acs{CO2} concentration at the
     surface (\emph{co2s}) averaged over the months \acf{JJA} of the year 2014
@@
 \emph{rsntcs}, and \emph{uajet} have been implemented into the \ac{ESMValCore}.
-Apart from the atmospheric \ac{CO2} concentration at surface (\emph{co2s}),
+Apart from the atmospheric \ac{CO2} concentration at the surface (\emph{co2s}),
 which is used in \cref{ch:06:paper_gpp} to derive the predictor variable of the
@@
 
-The absorbed shortwave radiation (\emph{asr}) is defined as the difference in
-the incoming \ac{TOA} shortwave radiation and the outgoing \ac{TOA} shortwave
-radiation. The atmospheric \ac{CO2} concentration at surface (\emph{co2s}) can
-be calculated from the pressure level--dependent atmospheric \ac{CO2}
-concentration and the surface air pressure using interpolation. An example
-illustrating \emph{co2s} is given in \cref{fig:03:co2s}. \Cref{fig:03:co2s:a}
-shows the \acs{CMIP}6 \ac{MMM} of the global \emph{co2s} for the months
-\acf{JJA} of the year 2014 in the emission-driven historical simulation. As
-expected, the map shows high atmospheric \ac{CO2} concentrations over large
-metropolitan areas with high \ac{CO2} emissions (\eg{} over North America, the
-Arabian Peninsula, and East Asia). Moreover, regions with high photosynthetic
-activity (\eg{} the boreal forests in the Northern Hemisphere summer and the
-tropical rainforests in South America) exhibit smaller \ac{CO2} concentrations.
-\Cref{fig:03:co2s:b} shows the monthly mean \emph{co2s} at Mauna Loa, Hawaii
-from 1958 to 2014 for observations performed at the Mauna Loa Observatory
-\autocite{Keeling2005} (thick black line) and emission-driven historical
-simulations from four \acs{CMIP}6 models (remaining colored lines). Apart from
-the model CNRM-ESM2-1, the simulated atmospheric \ac{CO2} concentrations show
-the expected seasonal cycle present in the Keeling Curve (see
-\cref{fig:02:keeling_curve}). All \acp{ESM} correctly simulate the increase in
-\emph{co2s} over the years. However, there are some differences in the absolute
-values with the model BCC-CSM2-MR showing the largest deviations.
+The absorbed shortwave radiation (\emph{asr}) is defined as the difference
+between the incoming \ac{TOA} shortwave radiation and the outgoing \ac{TOA}
+shortwave radiation. The atmospheric \ac{CO2} concentration at the surface
+(\emph{co2s}) can be calculated from the pressure level--dependent atmospheric
+\ac{CO2} concentration and the surface air pressure using interpolation. An
+example illustrating \emph{co2s} is given in \cref{fig:03:co2s}.
+\Cref{fig:03:co2s:a} shows the \acs{CMIP}6 \ac{MMM} of the global \emph{co2s}
+for the months \acf{JJA} of the year 2014 in the emission-driven historical
+simulation. As expected, the map shows high atmospheric \ac{CO2} concentrations
+over large metropolitan areas with high \ac{CO2} emissions (\eg{} over North
+America, the Arabian Peninsula, and East Asia). Moreover, regions with high
+photosynthetic activity (\eg{} the boreal forests in the Northern Hemisphere
+summer and the tropical rainforests in South America) exhibit smaller \ac{CO2}
+concentrations. \Cref{fig:03:co2s:b} shows the monthly mean \emph{co2s} at
+Mauna Loa, Hawaii from 1958 to 2014 for observations performed at the Mauna Loa
+Observatory \autocite{Keeling2005} (thick black line) and emission-driven
+historical simulations from four \acs{CMIP}6 models (remaining colored lines).
+Apart from the model CNRM-ESM2-1, the simulated atmospheric \ac{CO2}
+concentrations show the expected seasonal cycle present in the Keeling Curve
+(see \cref{fig:02:keeling_curve}). All \acp{ESM} correctly simulate the
+increase in \emph{co2s} over the years. However, there are some differences in
+the absolute values with the model BCC-CSM2-MR showing the largest deviations.
 
@@
 variability metric $\Psi$. \Cref{fig:03:cox18:a} shows $\Psi$ over the past 85
-years for \acs{CMIP}5 models and observations form HadCRUT4
+years for \acs{CMIP}5 models and observations from HadCRUT4
 \autocite{Morice2012}. Since high-sensitivity climate models exhibit higher
@@
 constraint by \textcite{Cox2018} is, together with ten other emergent
-constraint on \ac{ECS}, analyzed in great detail for the \acs{CMIP}5 and
+constraints on \ac{ECS}, analyzed in great detail for the \acs{CMIP}5 and
 \acs{CMIP}6 ensemble in \cref{ch:05:paper_ecs}, which also includes a
@@
 exhibit a high equilibrium warming in the future or vice versa. On the other
-hand, \ac{TCR} and \ac{ECS} are in fact well correlated within the \acs{CMIP}5
-ensemble as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this
-figure only shows a linear regression fit between the two variables and not the
+hand, \ac{TCR} and \ac{ECS} are well correlated within the \acs{CMIP}5 ensemble
+as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this figure
+only shows a linear regression fit between the two variables and not the
 expected reciprocal relation. The output of \emph{recipe\_flato13ipcc.yml}
@@
 with the \ac{ESMValTool}. For this, the raw observational datasets are
-reformatted in such way that they adhere to the \ac{CMOR} standards. Since the
-\acs{CMOR}ization scripts are publicly available like any other code in the
+reformatted in such a way that they adhere to the \ac{CMOR} standards. Since
+the \acs{CMOR}ization scripts are publicly available like any other code in the
 \ac{ESMValTool} and \ac{ESMValCore} repositories, the entire scientific
@@
 \cref{tab:03:changes_to_esmvaltool}. The majority of these datasets are used
-for analysis presented in this thesis (\eg{} see
+for analyses presented in this thesis (\eg{} see
 \cref{tab:05:overview_emergent_constraints,tab:06:predictors}).

### 2021-04-08 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Spell- and grammer-checked chapter 4

@@
 metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6 models is crucial
-since they provide vital information about the future climate of planet Earth.
-As the \ac{ESMValTool} is an excellent tool that allows for a quick and robust
-evaluation of new \ac{CMIP} data as soon as it gets published on the \ac{ESGF}
-servers (see \cref{sec:03:esmvaltool}), it is used here to assess and analyze
-\ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to corresponding
-results of predecessor model generations and international climate assessments.
-This work is already published in two scientific papers: \textcite{Bock2020}
-and \textcite{Meehl2020}. For \textcite{Bock2020}, the author of this thesis
-contributed their figure 8 (a bar chart showing the \ac{ECS} for several
-\acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their figure 10 (map plots
-showing cloud-related feedback parameters for the \acs{CMIP}5 and \acs{CMIP}6
-\acp{MMM}; see \cref{fig:04:cloud_feedback_parameters}), code to create these
-figures with the \ac{ESMValTool}, and text for the manuscript (in particular a
-section about the evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author
-of this thesis contributed all figures and tables, code to produce these with
-the \ac{ESMValTool}, and text for the manuscript. This
+since they provide vital information about the future climate of the planet
+Earth. As the \ac{ESMValTool} is an excellent tool that allows for a quick and
+robust evaluation of new \ac{CMIP} data as soon as it gets published on the
+\ac{ESGF} servers (see \cref{sec:03:esmvaltool}), it is used here to assess and
+analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to
+corresponding results of predecessor model generations and international
+climate assessments. This work is already published in two scientific papers:
+\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
+author of this thesis contributed their figure 8 (a bar chart showing the
+\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their
+figure 10 (map plots showing cloud-related feedback parameters for the
+\acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
+\cref{fig:04:cloud_feedback_parameters}), code to create these figures with the
+\ac{ESMValTool}, and text for the manuscript (in particular a section about the
+evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author of this thesis
+contributed all figures and tables, code to produce these with the
+\ac{ESMValTool}, and text for the manuscript. This
 \namecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
@@
 assessed for all \acs{CMIP}5 and \acs{CMIP}6 models where the necessary
-temperature and radiation data is available (as of \TheMonth{}).
-\Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an overview over the
+temperature and radiation data are available (as of \TheMonth{}).
+\Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an overview of the
 results for the \acs{CMIP}5 and \acs{CMIP}6 models, respectively.
@@
     multi-model standard deviation is given by the sample standard deviation of
-    \acs{ECS} evaluated over all climate models (using the normalization $1 /
+    \acs{ECS} calculated over all climate models (using the normalization $1 /
     M$, where $M$ is the number of models). Corresponding references for each
@@
 spread in the \acs{CMIP}6 ensemble is the existence of several models with very
-high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exist a
-number of models with very low values of \ac{ECS}. Therefore, the \acs{CMIP}6
-model ranges of \ac{ECS} and \ac{TCR} are well outside the corresponding
-assessed ranges given by the latest published \ac{AR} of the \ac{IPCC} from
-2013 \autocite{Stocker2013} with $\rangeunit{1.8}{5.6}{K}$ for \ac{ECS}
-(\acs{AR}5: $\rangeunit{1.5}{4.5}{K}$) and $\rangeunit{1.3}{3.0}{K}$
-(\acs{AR}5: $\rangeunit{1.0}{2.5}{K}$) for \ac{TCR}.
+high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exist
+some models with very low values of \ac{ECS}. Therefore, the \acs{CMIP}6 model
+ranges of \ac{ECS} and \ac{TCR} are well outside the corresponding assessed
+ranges given by the latest published \ac{AR} of the \ac{IPCC} from 2013
+\autocite{Stocker2013} with $\rangeunit{1.8}{5.6}{K}$ for \ac{ECS} (\acs{AR}5:
+$\rangeunit{1.5}{4.5}{K}$) and $\rangeunit{1.3}{3.0}{K}$ (\acs{AR}5:
+$\rangeunit{1.0}{2.5}{K}$) for \ac{TCR}.
 
@@
 years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
-  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
+  Stocker2013}, with the exception of \acs{AR}4, where the lower bound was
 temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
@@
     with the corresponding \acf{MMM} illustrated with the horizontal black
-    lines (the green circle represents output from a single climate model). The
-    numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
-    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). \AdaptedFrom{Meehl2020}.
+    lines (the green circle represents the output from a single climate
+    model). The numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6
+    models (see \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}).
+    \AdaptedFrom{Meehl2020}.
   }
@@
 \cref{eq:02:tcr_vs_ecs}, the increase of \ac{TCR} in many \acs{CMIP}6 models is
-not surprising in light of the many high \ac{ECS} models that are present the
-\acs{CMIP}6 ensemble.
+not surprising in light of the many high \ac{ECS} models that are present in
+the \acs{CMIP}6 ensemble.
 
@@
 smaller liquid droplets reflects more sunlight than a cloud consisting of
-larger ice crystals (for a fixed water content), which has a cooling effect on
+larger ice crystals (for fixed water content), which has a cooling effect on
 the climate system. In addition, predominantly liquid clouds tend to
@@
     affected \acs{CMIP}6 models, the strong negative cloud feedback that
-    results from a cloud phase change from ice clouds in the present-day to
+    results from a cloud phase change from ice clouds in the present day to
     liquid clouds in the future is substantially reduced (orange arrows). This
@@
 of the current climate \autocite{Gettelman2019}. Another example is the
-\acs{HadGEM}3 model, for which a new cloud scheme that accounts for turbulent
-production of liquid water in mixed-phase clouds has been implemented
+\acs{HadGEM}3 model, for which a new cloud scheme that accounts for the
+turbulent production of liquid water in mixed-phase clouds has been implemented
 \autocite{BodasSalcedo2019}. The new scheme parameterizes the role of
@@
 negative cloud feedback described above that results from a cloud phase change
-from ice clouds in the present-day to liquid clouds in the future is
+from ice clouds in the present day to liquid clouds in the future is
 substantially reduced in those \acs{CMIP}6 models that simulate a higher
@@
 show a higher shortwave and net \ac{CRE} feedback parameter over the Southern
-Ocean in \acs{CMIP}6 compared to \acs{CMIP}5. This is consistent to the
+Ocean in \acs{CMIP}6 compared to \acs{CMIP}5. This is consistent with the
 hypothesis presented in the paragraph above, which claims that the reduced

### 2021-04-08 ch07_conclusion/conclusion.tex
Spell- and grammer-checked chapter 2
Spell- and grammer-checked chapter 4

--- 2021-04-08 ch07_conclusion/conclusion.tex
@@
 the Southern Ocean that is present in predecessor versions of these models and
-that results from a cloud phase change from ice clouds in the present-day to
+that results from a cloud phase change from ice clouds in the present day to
 liquid clouds in the future. In the affected \acs{CMIP}6 models, this cloud

### 2021-04-09 02_abstract/abstract.tex
Fixed usage of numerals (follolwing APA style)

@@
 
-To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models,
-eleven published emergent constraints on \acs{ECS} (mostly derived from models
+To reduce uncertainties in \acs{ECS} projected by the \acs{CMIP}6 models, 11
+published emergent constraints on \acs{ECS} (mostly derived from models
 participating in \acs{CMIP}5, the predecessor generation of \acs{CMIP}6) are

### 2021-04-09 ch02_scientific_background/scientific_background.tex
Fixed usage of numerals (follolwing APA style)

@@
 straightforward setup for out-of-sample testing \autocite{Caldwell2018}, which
-is discussed in more detail in \cref{ch:05:paper_ecs}, where eleven emergent
+is discussed in more detail in \cref{ch:05:paper_ecs}, where 11 emergent
 constraints on \ac{ECS} are evaluated on the new \acs{CMIP}6 ensemble.

### 2021-04-09 ch03_esmvaltool/esmvaltool.tex
Fixed usage of numerals (follolwing APA style)

@@
 \pm 0.6 \right) \unit{K}$ ($66 \unit{\%}$ confidence range). The emergent
-constraint by \textcite{Cox2018} is, together with ten other emergent
+constraint by \textcite{Cox2018} is, together with 10 other emergent
 constraints on \ac{ECS}, analyzed in great detail for the \acs{CMIP}5 and
@@
 community can use them to \acs{CMOR}ize the observational data. A complete list
-of all eleven newly implemented \acs{CMOR}izers including the main reference
-for the corresponding observational datasets is shown in
+of all 11 newly implemented \acs{CMOR}izers including the main reference for
+the corresponding observational datasets is shown in
 \cref{tab:03:changes_to_esmvaltool}. The majority of these datasets are used

### 2021-04-09 ch05_paper_ecs/paper_ecs.tex
Fixed usage of numerals (follolwing APA style)

@@
 models, \ie{} whether they have skill to reduce uncertainties in \ac{ECS} in
-the \acs{CMIP}6 ensemble. In total we assess eleven emergent constraints on
+the \acs{CMIP}6 ensemble. In total we assess 11 emergent constraints on
 \ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out of which the
@@
   \caption[
-    Overview of the eleven emergent constraints on the \acf{ECS} evaluated in
+    Overview of the 11 emergent constraints on the \acf{ECS} evaluated in
     \cref{ch:05:paper_ecs}.
   ]{
-    Overview of the eleven emergent constraints on the \acf{ECS} evaluated in
-    this \namecref{ch:05:paper_ecs}. Detailed descriptions of the variables
-    used to calculate the \xaxis{} values of the emergent constraints and the
+    Overview of the 11 emergent constraints on the \acf{ECS} evaluated in this
+    \namecref{ch:05:paper_ecs}. Detailed descriptions of the variables used to
+    calculate the \xaxis{} values of the emergent constraints and the
     references for the corresponding observations (fourth column) are given in
@@
 output and observations) and methods used in this study
-(\cref{sec:05:data_and_methods}). After that, we present the eleven emergent
+(\cref{sec:05:data_and_methods}). After that, we present the 11 emergent
 constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
@@
 \ac{ECS} with the Gregory regression method \autocite{Gregory2004}, which is
-described in detail in \cref{subsec:02:ecs}. An overview of the eleven emergent
+described in detail in \cref{subsec:02:ecs}. An overview of the 11 emergent
 constraints on \ac{ECS} analyzed in this study including the variables required
@@
 In this \namecref{sec:05:comparison_of_emergent_constraints} we describe and
-discuss the eleven emergent constraints on \ac{ECS} summarized in
+discuss the 11 emergent constraints on \ac{ECS} summarized in
 \cref{tab:05:overview_emergent_constraints} using \acs{CMIP}5 and \acs{CMIP}6
@@
   subsec:05:zha}) and provide a best estimate for \ac{ECS} and statistical
-significance of the eleven emergent constraints in
+significance of the 11 emergent constraints in
 \cref{subsec:05:emergent_constraints_summary}. While most of these emergent
@@
     blue shaded area show the emergent relationships evaluated on these models
-    including the uncertainty range. In our study, we added eleven more
-    \acs{CMIP}5 models (red circles). The corresponding emergent relationship
-    that considers all available \acs{CMIP}5 models is shown in red colors.
-    This relationship shows a considerably lower coefficient of determination
+    including the uncertainty range. In our study, we added 11 more \acs{CMIP}5
+    models (red circles). The corresponding emergent relationship that
+    considers all available \acs{CMIP}5 models is shown in red colors. This
+    relationship shows a considerably lower coefficient of determination
     ($R^2$) and higher $p$-value than the relationship using the original
@@
 significances in \acs{CMIP}6). The reason for the erratic skill in \acs{CMIP}5
-is the set of climate models used. For our analysis, we use eleven additional
+is the set of climate models used. For our analysis, we use 11 additional
 \acs{CMIP}5 models that have not been used in the original publication
@@
 
-\subsection{Constrained \acs{ECS} Ranges and Statistical Significance of the
-  Eleven Emergent Constraints}
+\subsection{Constrained \acs{ECS} Ranges and Statistical Significance of the 11
+  Emergent Constraints}
 \label{subsec:05:emergent_constraints_summary}
@@
   \caption[
-    Overview of the constrained \acf{ECS} ranges and $p$-values for all eleven
+    Overview of the constrained \acf{ECS} ranges and $p$-values for all 11
     analyzed emergent constraints.
   ]{
-    Overview of the constrained \acf{ECS} ranges and $p$-values for all eleven
+    Overview of the constrained \acf{ECS} ranges and $p$-values for all 11
     analyzed emergent constraints. If not further specified, the uncertainty
@@
   \caption[
-    Overview of the constrained \acf{ECS} ranges given by the eleven emergent
+    Overview of the constrained \acf{ECS} ranges given by the 11 emergent
     constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensemble.
   ]{
-    Overview of the constrained \acf{ECS} ranges given by the eleven analyzed
+    Overview of the constrained \acf{ECS} ranges given by the 11 analyzed
     emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange)
@@
 in the \acs{CMIP}6 ensemble than in the \acs{CMIP}5 ensemble. If a conventional
-significance test of $p < 0.05$ was applied, eight of the eleven constraints
-would pass this test on \acs{CMIP}5 model data but only five (BRI, SHL, SHS,
-VOL, and ZHA) would pass this test on \acs{CMIP}6. This is still much better
-than would be expected purely by chance. Hence, there is still skill in at
-least a few of the constraints, but it is much lower than suggested in nearly
-all of the initial studies.
+significance test of $p < 0.05$ was applied, eight of the 11 constraints would
+pass this test on \acs{CMIP}5 model data but only five (BRI, SHL, SHS, VOL, and
+ZHA) would pass this test on \acs{CMIP}6. This is still much better than would
+be expected purely by chance. Hence, there is still skill in at least a few of
+the constraints, but it is much lower than suggested in nearly all of the
+initial studies.
 
@@
 
-In this \namecref{ch:05:paper_ecs}, eleven different emergent constraints on
+In this \namecref{ch:05:paper_ecs}, 11 different emergent constraints on
 \ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,

### 2021-04-09 ch06_paper_gpp/paper_gpp.tex
Fixed usage of numerals (follolwing APA style)

@@
 records from 26 stations and nine terrestrial ecosystem models. An evaluation
-of two atmospheric inversions and eleven land surface models shows contrasting
+of two atmospheric inversions and 11 land surface models shows contrasting
 effects of \ac{CO2} fertilization (positive) and global warming (negative) on

### 2021-04-09 ch07_conclusion/conclusion.tex
Fixed usage of numerals (follolwing APA style)

@@
 the Earth system and a target variable to reduce uncertainties in the target
-variable with observations \autocite{Allen2002}. In total eleven emergent
+variable with observations \autocite{Allen2002}. In total 11 emergent
 constraints on \ac{ECS} are assessed, which are mostly related to cloud

### 2021-04-09 utils/search_in_tex.py

@@ -29,6 +29,8 @@ python search_in_tex.py '.{20}ly-.{20}'
 
 python search_in_tex.py '.{20}\s+[Aa][n]?\s+\\[Aa]c.{20}'
 
+python search_in_tex.py '\\.{0,7}ref.?\{' | sort -u
+
 """
 
 

### 2021-04-09 utils/search_in_pdf.py
Added script to search in PDF

@@ -1,0 +1,63 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+# This file is used to produce materials for the following work:
+
+# Doctoral Dissertation
+# (c) by Manuel Schlund
+
+# This work is licensed under a
+# Creative Commons Attribution 4.0 International License.
+
+# You should have received a copy of the license along with this
+# work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
+
+"""Search regex in ``pdf`` file ignoring whitespace.
+
+Examples
+--------
+python search_in_pdf.py '.{40}\s+1[^A-Za-z0-9].{40}'
+
+"""
+
+
+import re
+import sys
+from pathlib import Path
+
+import pdftotext
+
+
+PATH = Path(__file__).parents[1].resolve() / 'Dissertation_Manuel_Schlund.pdf'
+
+
+def search(filename, pattern, replace_whitespace_char=' '):
+    """Search for pattern in filename."""
+    with open(filename, 'rb') as f_in:
+        pdf_pages = pdftotext.PDF(f_in)
+    pdf_str = '\n\n'.join(pdf_pages)
+    pdf_str = re.sub('\s+', replace_whitespace_char, pdf_str)
+    matches = re.findall(pattern, pdf_str)
+    if not matches:
+        return
+    print("Found {len(matches):d} matche(s)")
+    for match in matches:
+        print(f"    {match}")
+
+
+def main():
+    """Main function."""
+    if len(sys.argv) != 2:
+        raise ValueError(f"Expected exactly one command line argument for "
+                         f"regex pattern, got {len(sys.argv) - 1:d}")
+    pattern = sys.argv[1]
+    print(f"Searching file '{PATH}'")
+    print(f"Using regex pattern '{pattern}'")
+    print()
+
+    # Search in file
+    search(PATH, pattern)
+
+
+if __name__ == '__main__':
+    main()

### 2021-04-09 ch01_introduction/introduction.tex
Used lowercase versions of namecref(s) at correct places

@@
 with observations by addressing the following three
-\namecrefs{enum:01:question_1}:
+\lcnamecrefs{enum:01:question_1}:
 
@@
 from the author of this thesis. A detailed list of contributions to these
-studies is given in the corresponding \namecrefs{ch:03:esmvaltool}.
+studies is given in the corresponding \lcnamecrefs{ch:03:esmvaltool}.
 
@@
 \ac{ECS} for these \acs{CMIP}6 models and compares these to results derived
-from \acs{CMIP}5 models. The content of this \namecref{ch:05:paper_ecs} is
+from \acs{CMIP}5 models. The content of this \lcnamecref{ch:05:paper_ecs} is
 published in \emph{Earth System Dynamics} \autocite{Schlund2020a}.

### 2021-04-09 ch02_scientific_background/scientific_background.tex
Used lowercase versions of namecref(s) at correct places

@@
 
-This \namecref{ch:02:scientific_background} introduces the scientific
+This \lcnamecref{ch:02:scientific_background} introduces the scientific
 background of this thesis. First, basic concepts of climate model simulations
@@
 about the specific variables and experiments analyzed is given in the
-corresponding \namecrefs{ch:04:papers_ecs_tcr_assessment}.
+corresponding \lcnamecrefs{ch:04:papers_ecs_tcr_assessment}.
 
@@
 sensitivity metrics \ac{ECS} and \ac{TCR} introduced in the previous
-\namecref{sec:02:climate_sensitivity} are by definition independent from carbon
-cycle--related effects, since both of them are defined in terms of a doubling
-of the atmospheric \ac{CO2} concentration, not in terms of \ac{CO2} emissions.
-Nevertheless, to assess policy-relevant metrics like the allowable fossil fuel
-emissions to meet particular warming targets, for example, the $1.5
+\lcnamecref{sec:02:climate_sensitivity} are by definition independent from
+carbon cycle--related effects, since both of them are defined in terms of a
+doubling of the atmospheric \ac{CO2} concentration, not in terms of \ac{CO2}
+emissions. Nevertheless, to assess policy-relevant metrics like the allowable
+fossil fuel emissions to meet particular warming targets, for example, the $1.5
 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015}, it is
 crucial to take the global carbon cycle into account. This
-\namecref{sec:02:carbon_cycle} introduces the scientific background of the
+\lcnamecref{sec:02:carbon_cycle} introduces the scientific background of the
 global carbon cycle and its current anthropogenic perturbations.
@@
 Thus, more sophisticated techniques are necessary to evaluate multi-model
-climate projections. This \namecref{sec:02:reducing_uncertainties} introduces
+climate projections. This \lcnamecref{sec:02:reducing_uncertainties} introduces
 three state-of-the-art methods to assess multi-model projections and reduce
 associated uncertainties with observations. These techniques form the baseline
-of the work presented in the following \namecrefs{ch:05:paper_ecs}.
+of the work presented in the following \lcnamecrefs{ch:05:paper_ecs}.
 
@@
 
-As indicated at the beginning of this \namecref{sec:02:reducing_uncertainties},
-one main issue of multi-model ensembles is that not all participating climate
-models are equally plausible. Usually, this is quantified with some kind of
-measure of the models' \emph{performance}, \ie{} their agreement with
-observations of the real climate system. However, this model performance can
-only be evaluated against observations of the past and present-day climate,
-which does not necessarily provide insights into the quality of model
-projections of the future climate.
+As indicated at the beginning of this
+\lcnamecref{sec:02:reducing_uncertainties}, one main issue of multi-model
+ensembles is that not all participating climate models are equally plausible.
+Usually, this is quantified with some kind of measure of the models'
+\emph{performance}, \ie{} their agreement with observations of the real climate
+system. However, this model performance can only be evaluated against
+observations of the past and present-day climate, which does not necessarily
+provide insights into the quality of model projections of the future climate.
 
@@
 with normalized weights $w_m$. Similar to the notation introduced in the
-previous \namecref{subsec:02:emergent_constraints}, $y$ is a target variable
+previous \lcnamecref{subsec:02:emergent_constraints}, $y$ is a target variable
 (\eg{} a projection of the future climate) and $m$ indexes the $M$ different
@@
 for this choice of diagnostics is similar to the one presented in the previous
-\namecref{subsec:02:model_weighting}: Weighting schemes should address the
+\lcnamecref{subsec:02:model_weighting}: Weighting schemes should address the
 question \enquote{which climate model is adequate for predicting the target

### 2021-04-09 ch03_esmvaltool/esmvaltool.tex
Used lowercase versions of namecref(s) at correct places

@@
 
-To address the \namecrefs{enum:01:question_1} of this thesis posed in
+To address the \lcnamecrefs{enum:01:question_1} of this thesis posed in
 \cref{sec:01:key_science_questions}, a reliable and efficient tool to read,
@@
 After a brief overview of the \ac{ESMValTool}'s structure
-(\cref{sec:03:esmvaltool}), this \namecref{ch:03:esmvaltool} presents all these
-contributions that are not documented in other
-\namecrefs{ch:04:papers_ecs_tcr_assessment} of this thesis and puts them into
+(\cref{sec:03:esmvaltool}), this \lcnamecref{ch:03:esmvaltool} presents all
+these contributions that are not documented in other
+\lcnamecrefs{ch:04:papers_ecs_tcr_assessment} of this thesis and puts them into
 context
@@
 \cref{tab:03:changes_to_esmvaltool} for a summary). Some of these include the
-analyses performed in other \namecrefs{ch:04:papers_ecs_tcr_assessment} of this
-thesis, namely \emph{recipe\_bock20jgr.yml}
+analyses performed in other \lcnamecrefs{ch:04:papers_ecs_tcr_assessment} of
+this thesis, namely \emph{recipe\_bock20jgr.yml}
 (\cref{ch:04:papers_ecs_tcr_assessment}), \emph{recipe\_meehl20sciadv.yml}

### 2021-04-09 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Used lowercase versions of namecref(s) at correct places

@@
 \ac{ESMValTool}, and text for the manuscript. This
-\namecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
+\lcnamecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
 climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6 ensembles
@@
 To illustrate the results of the previous
-\namecref{sec:04:evaluation_ecs_and_tcr} and put them into historical context,
-\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
-\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
-and the different \acp{AR} of the \ac{IPCC} in combination with the
-corresponding modeled ranges from the different \ac{CMIP} generations. Since
-the Charney report in 1979, the assessed range of \ac{ECS} of
-$\rangeunit{1.5}{4.5}{K}$ has almost remained unchanged during the last 40
+\lcnamecref{sec:04:evaluation_ecs_and_tcr} and put them into historical
+context, \cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS}
+and \ac{TCR} ranges over the years from the Charney report
+\autocite{Charney1979} and the different \acp{AR} of the \ac{IPCC} in
+combination with the corresponding modeled ranges from the different \ac{CMIP}
+generations. Since the Charney report in 1979, the assessed range of \ac{ECS}
+of $\rangeunit{1.5}{4.5}{K}$ has almost remained unchanged during the last 40
 years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
@@
 values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
-  Wyser2020}. A list of possible reasons for the increased climate sensitivity
-in the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
+  Wyser2020}. A list of possible reasons for the increased climate sensitivity in
+the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
 the shortwave cloud feedback over the Southern Ocean is given in
@@
 important questions for the entire climate modeling community. For this reason,
-this \namecref{sec:04:possible_reasons_high_ecs_cmip6} discusses possible
+this \lcnamecref{sec:04:possible_reasons_high_ecs_cmip6} discusses possible
 reasons for the increased climate sensitivity in \acs{CMIP}6.

### 2021-04-09 ch05_paper_ecs/paper_ecs.tex
Used lowercase versions of namecref(s) at correct places

@@
 The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
-previous \namecref{ch:04:papers_ecs_tcr_assessment} begs the question about the
-interpretation and credibility of these results. Since the \acs{CMIP}6 models
-provide crucial input for the upcoming international climate assessment of the
-\ac{IPCC} (\acs{AR}6), this is not only important for the scientific community,
-but also directly affects policymakers and society. Due to the many new and
-improved processes included to the models, the \acs{CMIP}6 ensemble is arguably
-more realistic than its predecessor generations. However, does this
+previous \lcnamecref{ch:04:papers_ecs_tcr_assessment} begs the question about
+the interpretation and credibility of these results. Since the \acs{CMIP}6
+models provide crucial input for the upcoming international climate assessment
+of the \ac{IPCC} (\acs{AR}6), this is not only important for the scientific
+community, but also directly affects policymakers and society. Due to the many
+new and improved processes included to the models, the \acs{CMIP}6 ensemble is
+arguably more realistic than its predecessor generations. However, does this
 automatically make results derived from \acs{CMIP}6 models more trustworthy
@@
     Overview of the 11 emergent constraints on the \acf{ECS} evaluated in this
-    \namecref{ch:05:paper_ecs}. Detailed descriptions of the variables used to
-    calculate the \xaxis{} values of the emergent constraints and the
+    \lcnamecref{ch:05:paper_ecs}. Detailed descriptions of the variables used
+    to calculate the \xaxis{} values of the emergent constraints and the
     references for the corresponding observations (fourth column) are given in
@@
 to reproduce this study with all figures and tables using the \ac{ESMValTool}.
-In this \namecref{ch:05:paper_ecs}, we first introduce the data (climate model
-output and observations) and methods used in this study
+In this \lcnamecref{ch:05:paper_ecs}, we first introduce the data (climate
+model output and observations) and methods used in this study
 (\cref{sec:05:data_and_methods}). After that, we present the 11 emergent
@@
 
-In this \namecref{sec:05:comparison_of_emergent_constraints} we describe and
+In this \lcnamecref{sec:05:comparison_of_emergent_constraints} we describe and
 discuss the 11 emergent constraints on \ac{ECS} summarized in
@@
 
-As shown in the previous \namecrefs{sec:05:comparison_of_emergent_constraints},
-most emergent relationships show smaller coefficients of determination when
-evaluated on the new \acs{CMIP}6 ensemble compared to the \acs{CMIP}5 ensemble.
-In this \namecref{sec:05:discussion}, we discuss possible reasons for, and
+As shown in the previous
+\lcnamecrefs{sec:05:comparison_of_emergent_constraints}, most emergent
+relationships show smaller coefficients of determination when evaluated on the
+new \acs{CMIP}6 ensemble compared to the \acs{CMIP}5 ensemble. In this
+\lcnamecref{sec:05:discussion}, we discuss possible reasons for, and
 implications of, these differences. As reported by \textcite{Caldwell2014}, the
@@
 
-In this \namecref{ch:05:paper_ecs}, 11 different emergent constraints on
+In this \lcnamecref{ch:05:paper_ecs}, 11 different emergent constraints on
 \ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,

### 2021-04-09 ch06_paper_gpp/paper_gpp.tex
Used lowercase versions of namecref(s) at correct places

@@
 
-As argued in detail in the previous \namecref{ch:05:paper_ecs},
+As argued in detail in the previous \lcnamecref{ch:05:paper_ecs},
 single-process-oriented emergent constraints do not seem to be beneficial in
@@
 to reproduce this study with all figures and tables using the \ac{ESMValTool}.
-\Cref{sec:06:data_and_methods} of this \namecref{ch:06:paper_gpp} provides an
+\Cref{sec:06:data_and_methods} of this \lcnamecref{ch:06:paper_gpp} provides an
 overview of the data and methods used in this paper. The results are presented
@@
 
-In this \namecref{subsec:06:step_1}, we present the observational emergent
+In this \lcnamecref{subsec:06:step_1}, we present the observational emergent
 constraint on the \ac{CO2} fertilization effect by \textcite{Wenzel2016} which
@@
 vice versa. The rescaled target variables are then fed into the \ac{ML} model
-(step 2) presented in the following \namecref{subsec:06:step_2}. We note that
+(step 2) presented in the following \lcnamecref{subsec:06:step_2}. We note that
 this global rescaling assumes that the relative biases in the fractional
@@
 
-As already described in the previous \namecref{subsec:06:step_1}, we analyze
+As already described in the previous \lcnamecref{subsec:06:step_1}, we analyze
 two different target variables in this study: the monthly climatology of
@@
 used to rescale the gridded output of the different climate models with
-\cref{eq:06:y}. The following two \namecrefs{subsec:06:results_step2a} show the
-results for the rescaled absolute future \ac{GPP} (step 2a) and the rescaled
-fractional \ac{GPP} change (step 2b) further constrained by the \ac{GBRT}
-model.
+\cref{eq:06:y}. The following two \lcnamecrefs{subsec:06:results_step2a} show
+the results for the rescaled absolute future \ac{GPP} (step 2a) and the
+rescaled fractional \ac{GPP} change (step 2b) further constrained by the
+\ac{GBRT} model.
 
@@
 
-In this \namecref{subsubsec:06:results_step2a_obs}, we use the different
+In this \lcnamecref{subsubsec:06:results_step2a_obs}, we use the different
 statistical models to predict absolute \ac{GPP} at the end of the \nth{21}
@@
 
-In this \namecref{ch:06:paper_gpp}, a two-step approach to constrain the
+In this \lcnamecref{ch:06:paper_gpp}, a two-step approach to constrain the
 projected \ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in

### 2021-04-09 ch02_scientific_background/scientific_background.tex
Capitalized automatically generated enumerated elements

@@
 important metrics describing climate sensitivity
-(\cref{sec:02:climate_sensitivity}) and fundamental biogeochemical processes of
-the global carbon cycle (\cref{sec:02:esms}) are presented. Finally,
-state-of-the-art techniques that can be used to reduce uncertainties in
-projections of the future climate with observations are shown
+(\cref{sec:02:climate_sensitivity}) and fundamental processes of the global
+carbon cycle (\cref{sec:02:esms}) are presented. Finally, state-of-the-art
+techniques that can be used to reduce uncertainties in projections of the
+future climate with observations are shown
 (\cref{sec:02:reducing_uncertainties}). These methods form the basis for new

### 2021-04-09 ch07_conclusion/conclusion.tex
Used lowercase versions of namecref(s) at correct places
Capitalized automatically generated enumerated elements

@@
 
-To answer the \namecrefs{enum:01:question_1} posed in
+To answer the \lcnamecrefs{enum:01:question_1} posed in
 \cref{sec:01:key_science_questions} and to ensure a consistent evaluation of
@@
 already published in \textcite{Bock2020, Meehl2020}, directly addresses
-\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
+\Cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
 \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
@@
 \acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
-emergent constraints. Thus, \cref{enum:01:question_2}
+emergent constraints. Thus, \Cref{enum:01:question_2}
 (\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{not
@@
 to regions closer to the equator through an extension of the growing season.
-This study directly addresses \cref{enum:01:question_3}
+This study directly addresses \Cref{enum:01:question_3}
 (\enquote{\KeyScienceQuestionThree{}}), which can be confidently answered with

### 2021-04-09 end_appendix/appendix.tex
Removed nameref*

@@
 
-\vspace{\fill}
+\vspace{7mm}
 
@@
   ]{
-    Overview of the variables used in \cref{ch:05:paper_ecs}
-    (\nameref*{ch:05:paper_ecs}). More details are given in
-    \cref{tab:05:overview_emergent_constraints}, which lists the variables used
-    for each emergent constraint.
+    Overview of the variables used in \cref{ch:05:paper_ecs}. More details are
+    given in \cref{tab:05:overview_emergent_constraints}, which lists the
+    variables used for each emergent constraint.
   }
@@
   ]{
-    References for all observational datasets used in \cref{ch:05:paper_ecs}
-    (\nameref*{ch:05:paper_ecs}). More details are given in
-    \cref{tab:05:overview_emergent_constraints}, which lists the observational
-    datasets used for each emergent constraint.
+    References for all observational datasets used in \cref{ch:05:paper_ecs}.
+    More details are given in \cref{tab:05:overview_emergent_constraints},
+    which lists the observational datasets used for each emergent constraint.
   }
@@
   ]{
-    List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
-    (\nameref*{ch:05:paper_ecs}) alongside the main reference and the index
-    used in the corresponding figures. \AdaptedFrom{Schlund2020a}.
+    List of \acs{CMIP}5 models used in \cref{ch:05:paper_ecs} alongside the
+    main reference and the index used in the corresponding figures.
+    \AdaptedFrom{Schlund2020a}.
   }
@@
   ]{
-    All \acs{CMIP}5 models used in \cref{ch:05:paper_ecs}
-    (\nameref*{ch:05:paper_ecs}) including their \acf{ECS} and \xaxis{} values
-    for the emergent constraints BRI, COX, LIP, SHD, SHL, and SHS. More details
-    on the emergent constraints are given in
+    All \acs{CMIP}5 models used in \cref{ch:05:paper_ecs} including their
+    \acf{ECS} and \xaxis{} values for the emergent constraints BRI, COX, LIP,
+    SHD, SHL, and SHS. More details on the emergent constraints are given in
     \cref{tab:05:overview_emergent_constraints}. The specified index
@@
     List of the seven \acs{CMIP}5 models used in \cref{ch:06:paper_gpp}
-    (\nameref*{ch:06:paper_gpp}) alongside the main references. More details
-    are given by \textcite{Anav2013}. We chose all \acs{CMIP}5 models which
-    provide all necessary variables (\emph{co2}, \emph{gpp}, \emph{lai},
-    \emph{pr}, \emph{rsds}, and \emph{tas}) for all used experiments
+    alongside the main references. More details are given by
+    \textcite{Anav2013}. We chose all \acs{CMIP}5 models which provide all
+    necessary variables (\emph{co2}, \emph{gpp}, \emph{lai}, \emph{pr},
+    \emph{rsds}, and \emph{tas}) for all used experiments
     (\emph{esmHistorical}, \emph{esmrcp85}, and \emph{esmFixClim1}). For all

### 2021-04-09 ch03_esmvaltool/esmvaltool.tex
Capitalized manual references of figures, tables, chapters, sections and versions

@@
 (\url{https://github.com/ESMValGroup}). It is released under the Apache
-License, version 2.0. The source code of the latest released version of the
+License, Version 2.0. The source code of the latest released version of the
 \ac{ESMValCore} package, which includes the core functionalities of the
@@
     \midrule
-    Recipe & \emph{bock20jgr} (figures 8 and 10) & \autocite{Bock2020} \\
+    Recipe & \emph{bock20jgr} (Figures 8 and 10) & \autocite{Bock2020} \\
     & \emph{cox18nature} & \autocite{Cox2018} \\
     & \emph{ecs} & \autocite{Gregory2004} \\
-    & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
+    & \emph{flato13ipcc} (Figure 9.42) & \autocite{Flato2013} \\
     & \emph{meehl20sciadv} & \autocite{Meehl2020} \\
@@
 
-Due to an exhaustive overhaul of the \ac{ESMValTool} version 2, older recipes
-from version 1 \autocite{Eyring2016a} cannot be used directly in the latest
+Due to an exhaustive overhaul of the \ac{ESMValTool} Version 2, older recipes
+from Version 1 \autocite{Eyring2016a} cannot be used directly in the latest
 release. Porting recipes from the old version to the current version requires
@@
 climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which
-reproduces large parts of the climate model evaluation performed in chapter 9
+reproduces large parts of the climate model evaluation performed in Chapter 9
 of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
-recipe, the two-panel figure 9.42 has been added, which shows the relationship
-between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
+recipe, the Figure 9.42 of \textcite{Flato2013} has been added, which shows the
+relationship between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
 (\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS}

### 2021-04-09 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Capitalized manual references of figures, tables, chapters, sections and versions

@@
 \textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
-author of this thesis contributed their figure 8 (a bar chart showing the
+author of this thesis contributed their Figure 8 (a bar chart showing the
 \ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their
-figure 10 (map plots showing cloud-related feedback parameters for the
+Figure 10 (map plots showing cloud-related feedback parameters for the
 \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
@@
 interactions that are relevant in the context of high climate sensitivity in
-\acs{CMIP}6 are found for example in the \ac{HadGEM} version 3. The new aerosol
+\acs{CMIP}6 are found for example in the \ac{HadGEM} Version 3. The new aerosol
 scheme in this model suppresses a strong negative cloud feedback over the

### 2021-04-09 utils/search_in_tex.py

@@ -31,6 +31,8 @@ python search_in_tex.py '.{20}\s+[Aa][n]?\s+\\[Aa]c.{20}'
 
 python search_in_tex.py '\\.{0,7}ref.?\{' | sort -u
 
+python search_in_tex.py '.{50}figure[s]?[^\}A-Za-z]+.{50}'
+
 """
 
 

### 2021-04-09 ch06_paper_gpp/paper_gpp.tex
Fixed capitalization for step XY

@@
 constraint on \ac{CO2} fertilization from \textcite{Wenzel2016} that we apply
-here to our study to rescale the climate model output (step 1). This includes a
+here to our study to rescale the climate model output (Step 1). This includes a
 discussion of the caveats that are associated with this emergent constraint
@@
   Piao2018, Zhao2016}. In \cref{subsec:06:step_2} the \ac{ML} technique used in
-step 2 to constrain gridded \ac{GPP} projections is presented.
+Step 2 to constrain gridded \ac{GPP} projections is presented.
 \Cref{fig:06:schematic_steps} shows an overview of our two-step approach.
@@
 study is given in \cref{tab:app:b:cmip5_models}. To apply the emergent
-constraint in step 1, we need emission-driven \acs{CMIP}5 simulations. The
+constraint in Step 1, we need emission-driven \acs{CMIP}5 simulations. The
 number of \acp{ESM} included is therefore seven, as in \textcite{Wenzel2016}.
@@
   ]{
-    Schematic illustration of our two-step approach. In step 1 an emergent
+    Schematic illustration of our two-step approach. In Step 1 an emergent
     constraint by \textcite{Wenzel2016} is used to constrain the global mean
@@
     absolute \acs{GPP} at the end of the \nth{21} century and fractional
-    \acs{GPP} change over the \nth{21} century. In step 2, a \acl{ML} model is
-    used to constrain these two target variables (step 2a: absolute \acs{GPP};
-    step 2b: fractional \acs{GPP} change) in two independent approaches.
+    \acs{GPP} change over the \nth{21} century. In Step 2, a \acl{ML} model is
+    used to constrain these two target variables (Step 2a: absolute \acs{GPP};
+    Step 2b: fractional \acs{GPP} change) in two independent approaches.
     \AdaptedFrom{Schlund2020}.
@@
 
-Therefore, our target variable in step 1 is the global mean fractional
+Therefore, our target variable in Step 1 is the global mean fractional
 \acs{GPP} change over the \nth{21} century $f$ in the \acs{RCP}8.5 scenario,
@@
 \ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
-\acs{RCP}8.5 scenario (see step 2a) and the fractional \acs{GPP} change over
-the \nth{21} century as defined in \cref{eq:06:f} (see step 2b). The global
+\acs{RCP}8.5 scenario (see Step 2a) and the fractional \acs{GPP} change over
+the \nth{21} century as defined in \cref{eq:06:f} (see Step 2b). The global
 rescaling of the gridded target variables works as follows: Let $f_m$ be the
@@
 where the index $i$ counts over all grid cells (and months when absolute
-\ac{GPP} is used as target variable in step 2a) and $m$ refers to the \ac{ESM}
+\ac{GPP} is used as target variable in Step 2a) and $m$ refers to the \ac{ESM}
 ($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
@@
 vice versa. The rescaled target variables are then fed into the \ac{ML} model
-(step 2) presented in the following \lcnamecref{subsec:06:step_2}. We note that
+(Step 2) presented in the following \lcnamecref{subsec:06:step_2}. We note that
 this global rescaling assumes that the relative biases in the fractional
@@
 
-For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
+For Step 2 we use the \ac{GBRT} algorithm to find an optimal regression
 function $F$ relating the target variable $y$ to a set of $K$ predictors
@@
     Schematic illustration of our \acl{ML} approach to constrain projected
-    absolute \acf{GPP} at the end of the \nth{21} century in step 2a.
+    absolute \acf{GPP} at the end of the \nth{21} century in Step 2a.
   ]{
     Schematic illustration of our \acf{ML} approach to constrain projected
-    absolute \acf{GPP} in step 2a. (a) In the training phase of the algorithm,
+    absolute \acf{GPP} in Step 2a. (a) In the training phase of the algorithm,
     the model is fitted to the training data interpolating the empirical
@@
     leave-one-model-out \acl{CV} approach (see \cref{subsec:02:model_weighting}
-    for details). The schematic illustration of step 2b differs only in the
+    for details). The schematic illustration of Step 2b differs only in the
     target variable used (fractional \acs{GPP} change instead of absolute
@@
     \emph{features}) used in the \acf{GBRT} model to predict the target
-    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
+    variables. For Step 2a (target variable: absolute \acs{GPP}), all listed
     variables are monthly climatologies of the specified time ranges in the
-    historical climate. For step 2b (target variable: fractional \acs{GPP}
+    historical climate. For Step 2b (target variable: fractional \acs{GPP}
     change), the temporal mean over the specified time ranges is calculated for
@@
 two different target variables in this study: the monthly climatology of
-absolute \ac{GPP} at the end of the \nth{21} century (step 2a) and the
-fractional \ac{GPP} change over the \nth{21} century (step 2b). A schematic
-representation of step 2a is shown in \cref{fig:06:mlr_concept}. The target
+absolute \ac{GPP} at the end of the \nth{21} century (Step 2a) and the
+fractional \ac{GPP} change over the \nth{21} century (Step 2b). A schematic
+representation of Step 2a is shown in \cref{fig:06:mlr_concept}. The target
 variable $y$ for the \ac{ML} model is the rescaled monthly climatology of
@@
   a} and \labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). Step 2b differs
-from step 2a only in the target variable used. Both steps are independent
+from Step 2a only in the target variable used. Both steps are independent
 approaches and are evaluated separately.
@@
 particular, increases greater than $300 \unit{\%}$ in the fractional \ac{GPP}
-change that is used as target variable in step 2b are removed. These
+change that is used as target variable in Step 2b are removed. These
 unrealistically high values (making up about $5.9 \unit{\%}$ of the data) occur
@@
 Second, the uncertainty in the rescaling is simply assessed as the uncertainty
-given by the emergent constraint from step 1. Finally, the uncertainty in the
+given by the emergent constraint from Step 1. Finally, the uncertainty in the
 observation-based data (\ie{} uncertainty in the position of the black points
@@
 \cref{eq:06:y}. The following two \lcnamecrefs{subsec:06:results_step2a} show
-the results for the rescaled absolute future \ac{GPP} (step 2a) and the
-rescaled fractional \ac{GPP} change (step 2b) further constrained by the
+the results for the rescaled absolute future \ac{GPP} (Step 2a) and the
+rescaled fractional \ac{GPP} change (Step 2b) further constrained by the
 \ac{GBRT} model.
@@
     importance of our \acl{ML} approach to constrain projected absolute
-    \acf{GPP} at the end of the \nth{21} century in step 2a.
+    \acf{GPP} at the end of the \nth{21} century in Step 2a.
   ]{
     (a) Box plot of the \acf{RMSEP} distributions for six different statistical
-    models used to predict future absolute \acf{GPP} in step 2a using a
+    models used to predict future absolute \acf{GPP} in Step 2a using a
     leave-one-model-out \acl{CV} approach. The distribution for each
@@
     Relative global feature importance for the different predictors used in the
-    \acs{GBRT} model to predict future absolute \acs{GPP} (step 2a). The red
+    \acs{GBRT} model to predict future absolute \acs{GPP} (Step 2a). The red
     bars correspond to positive Pearson correlation coefficients between all
@@
 
-The result of step 1 on the gridded data is illustrated in
+The result of Step 1 on the gridded data is illustrated in
 \cref{fig:06:step2a_results:a}, which shows the ratio between r\acs{MMM} and
@@
     Illustration of our \acl{ML} approach to constrain projected absolute
-    \acf{GPP} at the end of the \nth{21} century with observations in step 2a.
+    \acf{GPP} at the end of the \nth{21} century with observations in Step 2a.
   ]{
@@
     constrained global mean \acs{GPP} change over the \nth{21} century from
-    step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
+    Step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
     \acs{GPP} change ($43 \unit{\%}$). All values close to $0$ for the dataset
@@
     between r\acs{MMM} and our \acf{GBRT} prediction for the end of the
-    \nth{21} century. This corresponds to step 2a of our approach. (c) Bias
+    \nth{21} century. This corresponds to Step 2a of our approach. (c) Bias
     between the modeled \acs{GPP} in the \acs{CMIP}5 \acs{MMM} of the
@@
     color). Panels (b) and (c) show similar bias patterns (pattern correlation
-    of $R^2 = 0.88$). Thus, the \acs{GBRT} prediction in step 2a is able to
+    of $R^2 = 0.88$). Thus, the \acs{GBRT} prediction in Step 2a is able to
     correct the historical bias of the \acs{CMIP}5 ensemble relative to the
@@
     absolute \acf{GPP} at the end of the \nth{21} century with observations in
-    step 2a and comparison to other statistical models.
+    Step 2a and comparison to other statistical models.
   ]{
@@
 
-The spatial result of our \ac{GBRT} model in step 2a can be visualized by
+The spatial result of our \ac{GBRT} model in Step 2a can be visualized by
 comparing r\acs{MMM} and the output of the \acs{GBRT} model.
@@
 end of the century of $169 \unit{\GtCyr{}}$, which is consistent with the
-global constraint from step 1 of $\pmrangeunit{171}{12}{\GtCyr{}}$. In summary,
-our approach first corrects the \acs{CMIP}5 models' response to \ac{CO2} (step
+global constraint from Step 1 of $\pmrangeunit{171}{12}{\GtCyr{}}$. In summary,
+our approach first corrects the \acs{CMIP}5 models' response to \ac{CO2} (Step
 1) and second corrects the historical bias of the \acp{ESM} relative to
-observations (step 2a).
+observations (Step 2a).
 
@@
 
-In step 2b, we constrain the gridded fractional \ac{GPP} change over the
+In Step 2b, we constrain the gridded fractional \ac{GPP} change over the
 \nth{21} century (2100 versus 2000) in the emission-driven \acs{RCP}8.5
-scenario (rescaled using \cref{eq:06:y}). This step is independent from step 2a
+scenario (rescaled using \cref{eq:06:y}). This step is independent from Step 2a
 and only differs in the target variable (\ac{GPP} change instead of absolute
@@
 \ac{GPP} change, its rescaled version using \cref{eq:06:y} (r\acs{MMM}), a
-linear \ac{LASSO} regression model, and the \ac{GBRT} model. Similar to step
+linear \ac{LASSO} regression model, and the \ac{GBRT} model. Similar to Step
 2a, the \ac{GBRT} approach shows the smallest mean \ac{RMSEP} values of all
@@
 \ac{GBRT} model also outperforms the linear \ac{LASSO} model (mean reduced by
-$3 \unit{\%}$). In contrast to step 2a, there is not a single predictor which
+$3 \unit{\%}$). In contrast to Step 2a, there is not a single predictor which
 heavily dominates the feature importance of the \ac{GBRT} model. Thus, using
 single-predictor models (\acs{GBRT}-1D and \acs{LASSO}-1D) is not necessary in
-step 2b. \Cref{fig:06:step2b_ml_inference:b,fig:06:step2b_ml_inference:d} show
+Step 2b. \Cref{fig:06:step2b_ml_inference:b,fig:06:step2b_ml_inference:d} show
 the global feature importance for the \ac{LASSO} and \ac{GBRT} model,
@@
     importance of our \acl{ML} approach to constrain the fractional change in
-    \acf{GPP} over the \nth{21} century in step 2b.
+    \acf{GPP} over the \nth{21} century in Step 2b.
   ]{
@@
     statistical models used to predict the fractional change in \acf{GPP} over
-    the \nth{21} century (step 2b) using the leave-one-model-out \acl{CV}
+    the \nth{21} century (Step 2b) using the leave-one-model-out \acl{CV}
     approach. A detailed description of the representation of these
@@
     different features used in the \acs{LASSO} model to predict the fractional
-    \acs{GPP} change (step 2b). The feature importance for the \acs{LASSO}
+    \acs{GPP} change (Step 2b). The feature importance for the \acs{LASSO}
     model is given by the normalized linear coefficients of the model. (c)
     Local feature importance for the \acs{GBRT} model used to predict
-    fractional \acs{GPP} change (step 2b) calculated using the \acf{LIME}
+    fractional \acs{GPP} change (Step 2b) calculated using the \acf{LIME}
     approach \autocite{Ribeiro2016} for the three dominant features \acf{T},
@@
     features used in the \acs{GBRT} model to predict the fractional \acs{GPP}
-    change (step 2b). The blue bars for the two global feature importance plots
+    change (Step 2b). The blue bars for the two global feature importance plots
     (b) and (d) correspond to negative Pearson correlation coefficients between
@@
 \ac{MMM}, its rescaled version (r\acs{MMM}) shows a slightly lower \ac{GPP}
-increase, which is consistent with the emergent constraint approach from step 1
+increase, which is consistent with the emergent constraint approach from Step 1
 that shows a lower global mean \ac{GPP} increase over the \nth{21} century ($39
@@
 consistent with the ratio $39 \unit{\%} / 43 \unit{\%} \approx 0.91$. Similar
-to step 2a, this can be interpreted as a global correction of the \acp{ESM}'
+to Step 2a, this can be interpreted as a global correction of the \acp{ESM}'
 response of future \ac{GPP} to changes in the atmospheric \ac{CO2}
@@
     Illustration of our \acl{ML} approach to constrain the fractional change in
-    \acf{GPP} over the \nth{21} century with observations in step 2b.
+    \acf{GPP} over the \nth{21} century with observations in Step 2b.
   ]{
     Fractional change in \acf{GPP} over the \nth{21} century (2100 versus 2000)
-    in the \acs{RCP}8.5 scenario for different statistical models (step 2b):
+    in the \acs{RCP}8.5 scenario for different statistical models (Step 2b):
     (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change, (b) rescaled
@@
     Further illustrations of our \acl{ML} approach to constrain the fractional
-    change in \acf{GPP} over the \nth{21} century with observations in step 2b.
+    change in \acf{GPP} over the \nth{21} century with observations in Step 2b.
   ]{
@@
     corresponds to the ratio of the constrained global mean \acs{GPP} change
-    over the \nth{21} century from step 1 ($39 \unit{\%}$) and the \acs{CMIP}5
+    over the \nth{21} century from Step 1 ($39 \unit{\%}$) and the \acs{CMIP}5
     ensemble mean global mean GPP change ($43 \unit{\%}$). All values close to
@@
 information that is added to the target variable by the observations in the
-predictors in step 2b. Overall, the absolute difference
+predictors in Step 2b. Overall, the absolute difference
 (\cref{fig:06:step2b_results_biases:b}) is small over the whole globe,
@@
 \cref{fig:06:step2b_ml_inference:a}, which only covers the uncertainty in the
-statistical models). Similar to step 2a, the local standard errors are large,
+statistical models). Similar to Step 2a, the local standard errors are large,
 even for the \ac{GBRT} model (that performs better than any other statistical
@@
 (\range{2091}{2100}) from the fractional \ac{GPP} change given by the \ac{GBRT}
-model in step 2b. This can be directly compared to the output of step 2a and
+model in Step 2b. This can be directly compared to the output of Step 2a and
 serves as a sanity check between steps 2a and 2b. As shown by
@@
 between the two approaches is in the same order of magnitude as the difference
-between the different statistical models used in step 2a (see
+between the different statistical models used in Step 2a (see
 \cref{fig:06:step2a_results_biases:c,fig:06:step2a_results_biases:d}) and
@@
 approaches (see \cref{fig:06:step2a_results:b}). Moreover, the globally
-aggregated results ($169 \unit{\GtCyr{}}$ for step 2a and $175 \unit{\GtCyr{}}$
-for step 2b) are both compatible with the global value of
+aggregated results ($169 \unit{\GtCyr{}}$ for Step 2a and $175 \unit{\GtCyr{}}$
+for Step 2b) are both compatible with the global value of
 $\pmrangeunit{171}{12}{\GtCyr{}}$ given by the global emergent constraint in
-step 1.
+Step 1.
 
@@
   \caption[
-    Comparison of our two independent approaches from step 2a and step 2b.
+    Comparison of our two independent approaches from Step 2a and Step 2b.
   ]{
     Top row: Absolute future \acf{GPP} at the end of the \nth{21} century
-    (\range{2091}{2100}) calculated using step 2a (a) and step 2b (b). Both
+    (\range{2091}{2100}) calculated using Step 2a (a) and Step 2b (b). Both
     approaches give similar results with global averages of $169
     \unit{\GtCyr{}}$ (a) and $175 \unit{\GtCyr{}}$ (b), which are both
-    consistent with the global result of step 1 of
+    consistent with the global result of Step 1 of
     $\pmrangeunit{171}{12}{\GtCyr{}}$. The pattern correlation between both
@@
 further constrain the gridded \ac{GPP} based on process-based present-day
-predictors. We consider two target variables: first (step 2a) the gridded
+predictors. We consider two target variables: first (Step 2a) the gridded
 monthly climatologies of absolute \ac{GPP} (\range{2091}{2100}) and second
-(step 2b) the gridded fractional \ac{GPP} change over the \nth{21} century
+(Step 2b) the gridded fractional \ac{GPP} change over the \nth{21} century
 (years \range{2091}{2100} versus years \range{1991}{2000}; see \cref{eq:06:f}).
@@
 statistical models. Compared to the unweighted \acs{CMIP}5 \ac{MMM} the mean
-prediction error is reduced by $48 \unit{\%}$ (step 2a) and $16 \unit{\%}$
-(step 2b); compared to a linear \ac{LASSO} regression model the mean prediction
+prediction error is reduced by $48 \unit{\%}$ (Step 2a) and $16 \unit{\%}$
+(Step 2b); compared to a linear \ac{LASSO} regression model the mean prediction
 error is reduced by approximately $3 \unit{\%}$ for both cases. However, local
@@
 expands this bias correction by taking more predictors than the historical
-\ac{GPP} into account. A similar result is provided by step 2b:
+\ac{GPP} into account. A similar result is provided by Step 2b:
 \cref{fig:06:step2b_results_biases:b,fig:06:step2b_results_biases:c} show the
@@
 target variable (fractional \ac{GPP} change) is already normalized with the
-historical \ac{GPP}. Therefore, \ac{T} (second most important feature in step
+historical \ac{GPP}. Therefore, \ac{T} (second most important feature in Step
 2a) is now the most important predictor instead of historical \ac{GPP} (see
@@
 
-Our \ac{GBRT} approach (step 2) is mathematically similar to \ac{MDER}
+Our \ac{GBRT} approach (Step 2) is mathematically similar to \ac{MDER}
 \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}: We establish a
@@
 The number of points establishing the relation is dramatically increased from
-about 50 points for classical emergent constraints to 237852 points (step 2a)
-and 16503 (step 2b) in our approach. We validate the relationship by extensive
+about 50 points for classical emergent constraints to 237852 points (Step 2a)
+and 16503 (Step 2b) in our approach. We validate the relationship by extensive
 out-of-sample testing using a leave-one-model-out \ac{CV} approach and randomly
@@
 Yet, like emergent constraints and performance weighting techniques, our
-\ac{GBRT} approach in step 2 hinges on the assumption that the climate models
+\ac{GBRT} approach in Step 2 hinges on the assumption that the climate models
 reflect the real world. This is certainly only partially true, since there are
@@
 implicitly incorporated in our approach by the target period we are using and
-the global rescaling using the emergent constraint in step 1. Thus, our
+the global rescaling using the emergent constraint in Step 1. Thus, our
 reduction of uncertainty only applies to the specific setup we are considering

### 2021-04-09 end_appendix/appendix.tex
Fixed capitalization for step XY

@@
 observation-driven dataset is regridded to a $\ang{2} \times \ang{2}$ grid and
-masked with a common mask to remove all oceans and Antarctica. For step 2a
+masked with a common mask to remove all oceans and Antarctica. For Step 2a
 (target variable: absolute \acf{GPP} at the end of the \nth{21} century),
 monthly climatologies are calculated for every dataset by averaging over all
-available years for every month. For step 2b (target variable: fractional
+available years for every month. For Step 2b (target variable: fractional
 \ac{GPP} change over the \nth{21} century), temporal means are calculated for
 every dataset by averaging over the full time dimension. In addition, values
-greater than $300 \unit{\%}$ in the target variable in step 2b (fractional
+greater than $300 \unit{\%}$ in the target variable in Step 2b (fractional
 \ac{GPP} change over the \nth{21} century) are masked to avoid numerical
@@
 test data points, and 46344 (3727) points for the prediction are used in the
-machine learning model for step 2a (step 2b).
+machine learning model for Step 2a (Step 2b).
 
@@
 A second source of uncertainty is the error in the rescaling of the target
-variable in step 1 of our approach. Analogous to \cref{eq:06:y}, we estimate
+variable in Step 1 of our approach. Analogous to \cref{eq:06:y}, we estimate
 this error as
@@
 $\bar{y}_i$ is the \acs{CMIP}5 \acf{MMM} of the target variable at prediction
-input point $i$ (step 2a: absolute \ac{GPP} at the end of the \nth{21} century;
-step 2b: fractional \ac{GPP} change over the \nth{21} century), $\sigma_{f'}$
+input point $i$ (Step 2a: absolute \ac{GPP} at the end of the \nth{21} century;
+Step 2b: fractional \ac{GPP} change over the \nth{21} century), $\sigma_{f'}$
 the standard error in the global mean fractional \acs{GPP} change given by the
-emergent constraint from step 1, and $\bar{f}$ the \acs{CMIP}5 \ac{MMM} global
+emergent constraint from Step 1, and $\bar{f}$ the \acs{CMIP}5 \ac{MMM} global
 mean fractional \ac{GPP} change over the \nth{21} century.
@@
     which is used as predictor for the emergent constraint step of our approach
-    (step 1). \AdaptedFrom{Schlund2020}.
+    (Step 1). \AdaptedFrom{Schlund2020}.
   }
@@
     Distribution of the residuals of the \acf{GBRT} model for the two different
-    target variables used in step 2a and step 2b.
+    target variables used in Step 2a and Step 2b.
   ]{
     Distribution of the residuals of the \acf{GBRT} model for the two different
-    target variables used in step 2a (absolute \acf{GPP} at the end of the
-    \nth{21} century) and step 2b (fractional \acs{GPP} change over the
+    target variables used in Step 2a (absolute \acf{GPP} at the end of the
+    \nth{21} century) and Step 2b (fractional \acs{GPP} change over the
     \nth{21} century). The distributions are derived by kernel density
@@
     Geographical distributions of the absolute \acf{GPP} at the end of the
-    \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for different
+    \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for different
     statistical models.
@@
     Geographical distributions of the absolute \acf{GPP} at the end of the
-    \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for different
+    \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for different
     statistical models. (a) \acs{CMIP}5 \acf{MMM}. (b) Rescaled \acs{CMIP}5
@@
     Geographical distributions of the \aclp{SPE} of the absolute \acf{GPP} at
-    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for
+    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for
     different statistical models.
@@
     Geographical distributions of the \acfp{SPE} of the absolute \acf{GPP} at
-    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (step 2a) for
+    the end of the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2a) for
     different statistical models. Details on the calculation of the \acs{SPE}
@@
     Geographical distributions of the \aclp{SPE} of the fractional change in
-    \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario (step 2b)
+    \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2b)
     for different statistical models.
@@
     Geographical distributions of the \acfp{SPE} of the fractional change in
-    \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario (step 2b)
+    \acf{GPP} over the \nth{21} century in the \acs{RCP}8.5 scenario (Step 2b)
     for different statistical models. Details on the calculation of the

### 2021-04-09 ch02_scientific_background/scientific_background.tex
Capitalized manual references of figures, tables, chapters, sections and versions
Spell- and grammer-checked remaining chapters

@@
     added to the coupled system, forming the modern \acfp{ESM}.
-    \IPCCReproducedFrom{Cubasch2013}{their figure 1.13}.
+    \IPCCReproducedFrom{Cubasch2013}{their Figure 1.13}.
   }
@@
     improvements and observational constraints.
-    \IPCCReproducedFrom{Cubasch2013}{their FAQ 1.1, figure 1}.
+    \IPCCReproducedFrom{Cubasch2013}{their FAQ 1.1, Figure 1}.
   }
@@
     large differences in time scales for the various feedbacks.
-    \IPCCReproducedFrom{Cubasch2013}{their figure 1.2}.
+    \IPCCReproducedFrom{Cubasch2013}{their Figure 1.2}.
   }
@@
     (atmosphere, land surface, and ocean). \IPCCReproducedFrom{Ciais2013}{their
-      FAQ 6.2, figure 1}.
+      FAQ 6.2, Figure 1}.
   }
@@
 be parameterized in \acp{ESM} since they occur on spatial scales much smaller
-than the typical grid cell sizes (see
+than the typical grid-cell sizes (see
 \cref{subsec:02:numerical_climate_modeling}). For the terrestrial carbon cycle,
@@
 ordinary least squares linear regression model. This is not always appropriate,
-for example when a non-linear emergent relationship is expected
+for example, when a non-linear emergent relationship is expected
 \autocite{Nijsse2020} or when additional physical considerations further

### 2021-04-09 ch03_esmvaltool/esmvaltool.tex
Spell- and grammer-checked remaining chapters

@@
 in \cref{fig:03:esmvaltool_schematic}). This ensures that the \ac{ESMValTool}
-can process arbitrary observational datasets, which can for example be used to
-assess the skill of climate model simulations. More details on the structure of
-the \ac{ESMValTool} are given in its extensive documentation
+can process arbitrary observational datasets, which can, for example, be used
+to assess the skill of climate model simulations. More details on the structure
+of the \ac{ESMValTool} are given in its extensive documentation
 (\url{https://docs.esmvaltool.org}).
@@
 \ac{CMIP} products by a broad research community and by national and
-international climate assessments. For this reason, the \ac{ESMValTool} is for
-example used to evaluate climate model output and observational data in several
-chapters of the upcoming Sixth \acl{AR} (\acs{AR}6) of the \ac{IPCC}.
+international climate assessments. For this reason, the \ac{ESMValTool} is used
+to evaluate climate model output and observational data in several chapters of
+the upcoming Sixth \acl{AR} (\acs{AR}6) of the \ac{IPCC}.
 
@@
 Similarly, the source code of the latest released version of the
-\ac{ESMValTool} package, which includes the recipe and diagnostic scripts that
-can for example be used to reproduce scientific assessments, is also available
-at Zenodo \autocite{Andela2020}.
+\ac{ESMValTool} package, which includes all publicly available recipes and
+diagnostic scripts, is also available at Zenodo \autocite{Andela2020}.
 
@@
 preprocessor is the computation of a variable's diurnal or seasonal cycle
-amplitude, which is for example necessary for the calculation of the predictor
-variable (\ac{CO2} seasonal cycle amplitude) of the emergent constraint on
-\ac{CO2} fertilization by \textcite{Wenzel2016} that is also used in
+amplitude. For instance, this is necessary to calculate the predictor variable
+(\ac{CO2} seasonal cycle amplitude) of the emergent constraint on \ac{CO2}
+fertilization by \textcite{Wenzel2016} that is also used in
 \cref{ch:06:paper_gpp} to constrain the global mean \ac{GPP} at the end of the
@@
 are reported in units of \enquote{per square meter of land/sea} and not in
-\enquote{per square meter of grid cell}. After the weighting, the grid cell
+\enquote{per square meter of grid cell}. After the weighting, the grid-cell
 areas can be used to integrate the flux-related variable over any desired

### 2021-04-09 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Spell- and grammer-checked remaining chapters

@@
 interactions that are relevant in the context of high climate sensitivity in
-\acs{CMIP}6 are found for example in the \ac{HadGEM} Version 3. The new aerosol
-scheme in this model suppresses a strong negative cloud feedback over the
-Southern Ocean that is found in the model's predecessor versions
-\autocite{BodasSalcedo2019}. The Southern Ocean is one of the most pristine
-aerosol regions in the world \autocite{Hamilton2014}, where aerosols
-predominantly originate from natural sources like marine emissions of gaseous
-dimethyl sulfide. Earlier \ac{HadGEM} versions show a large reduction in cloud
-droplet size with warming over the Southern Ocean, which increases the cloud
-albedo and thus produces a negative shortwave cloud feedback. On the contrary,
-in \acs{HadGEM}3, this negative feedback is completely suppressed or even
-positive due to a small increase of the cloud droplet size with warming, which
-eventually increases \ac{ECS} due to an increased net cloud feedback
-\autocite{BodasSalcedo2019}.
+\acs{CMIP}6 are found in the \ac{HadGEM} Version 3. The new aerosol scheme in
+this model suppresses a strong negative cloud feedback over the Southern Ocean
+that is found in the model's predecessor versions \autocite{BodasSalcedo2019}.
+The Southern Ocean is one of the most pristine aerosol regions in the world
+\autocite{Hamilton2014}, where aerosols predominantly originate from natural
+sources like marine emissions of gaseous dimethyl sulfide. Earlier \ac{HadGEM}
+versions show a large reduction in cloud droplet size with warming over the
+Southern Ocean, which increases the cloud albedo and thus produces a negative
+shortwave cloud feedback. On the contrary, in \acs{HadGEM}3, this negative
+feedback is completely suppressed or even positive due to a small increase of
+the cloud droplet size with warming, which eventually increases \ac{ECS} due to
+an increased net cloud feedback \autocite{BodasSalcedo2019}.
 

### 2021-04-09 ch05_paper_ecs/paper_ecs.tex
Spell- and grammer-checked remaining chapters

@@
 of the \ac{IPCC} (\acs{AR}6), this is not only important for the scientific
-community, but also directly affects policymakers and society. Due to the many
-new and improved processes included to the models, the \acs{CMIP}6 ensemble is
+community but also directly affects policymakers and society. Due to the many
+new and improved processes included in the models, the \acs{CMIP}6 ensemble is
 arguably more realistic than its predecessor generations. However, does this
@@
 this question is \enquote{not necessarily}. Even the \acs{CMIP}6 models are far
-from fully simulating all aspects and processes of the full vastly complex
-Earth system. Therefore, the addition of new processes to these models could
+from fully simulating all aspects and processes of the vastly complex Earth
+system. Therefore, the addition of new processes to these models could
 introduce new biases which have not been present in earlier \ac{CMIP} models.
@@
 models, \ie{} whether they have skill to reduce uncertainties in \ac{ECS} in
-the \acs{CMIP}6 ensemble. In total we assess 11 emergent constraints on
+the \acs{CMIP}6 ensemble. In total, we assess 11 emergent constraints on
 \ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out of which the
@@
 
-In this study we use the output from climate models participating in
+In this study, we use the output from climate models participating in
 \acs{CMIP}5 and \acs{CMIP}6 as shown in
 \cref{tab:app:a:cmip5_models,tab:app:a:cmip6_models}, respectively. In
-addition, for each emergent constraint at least one observational dataset is
+addition, for each emergent constraint, at least one observational dataset is
 used to calculate the observational constraint. A complete list of these
-datasets is given in \cref{tab:05:overview_emergent_constraints} and in
+datasets is given in \cref{tab:05:overview_emergent_constraints} and
 \cref{tab:app:a:observations}. Following other similar studies, we calculate
@@
 
-In this \lcnamecref{sec:05:comparison_of_emergent_constraints} we describe and
+In this \lcnamecref{sec:05:comparison_of_emergent_constraints}, we describe and
 discuss the 11 emergent constraints on \ac{ECS} summarized in
@@
 cloud radiative forcing and solar insolation averaged over the \ac{TLC} region.
-The regression coefficients of de-seasonalized variations of \ac{TLC} shortwave
+The regression coefficients of deseasonalized variations of \ac{TLC} shortwave
 albedo and \ac{SST} (in $\si{\pctK{}}$) are then used as an emergent constraint
@@
 \autocite{Dee2011}, and CERES-EBAF \autocite{Loeb2018} for the \ac{TOA}
-radiative fluxes over the time period \range{2001}{2005}. In the original
+radiative fluxes over the period \range{2001}{2005}. In the original
 publication, \textcite{Brient2016} use similar observation-based datasets with
@@
 observational data, here we use the HadCRUT4 dataset \autocite{Morice2012} over
-the time period \range{1880}{2014}. Under the COX constraint we assess a $66
+the period \range{1880}{2014}. Under the COX constraint we assess a $66
 \unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.03}{0.73}{K}$ for \acs{CMIP}5
@@
 $\pmrangeunit{2.8}{0.6}{K}$ from a different subset of \acs{CMIP}5 models but
-the same observations. When moving from \acs{CMIP}5 to \acs{CMIP}6, the
+with the same observations. When moving from \acs{CMIP}5 to \acs{CMIP}6, the
 significance of the emergent relation drops massively from $p = 0.0032$ to $p =
@@
 reanalysis data from ERA-Interim \autocite{Dee2011} for the meridional wind
-speed over the time period \range{1980}{2005}. Our application of this emergent
+speed over the period \range{1980}{2005}. Our application of this emergent
 constraint gives \ac{ECS} $66 \unit{\%}$ ranges of
@@
 observation-based data, we use vertical velocities from ERA-Interim
-\autocite{Dee2011} over the time period \range{1989}{1998} similar to the
-original publication. We derive \ac{ECS} $66 \unit{\%}$ confidence ranges of
+\autocite{Dee2011} over the period \range{1989}{1998} similar to the original
+publication. We derive \ac{ECS} $66 \unit{\%}$ confidence ranges of
 $\pmrangeunit{3.65}{0.64}{K}$ for \acs{CMIP}5 ($R^2 = 0.28$) and
@@
 humidity against \ac{ECS}. In contrast to \textcite{Su2014}, they do not use
-the entire tropics, but identify two distinct regions with the largest
+the entire tropics but identify two distinct regions with the largest
 correlation.
@@
 \ac{ECS}. Here, we use humidity observations from AIRS \autocite{Aumann2003}
-over the time period \range{2003}{2005} as the reference dataset. We assess a
-$66 \unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5
+over the period \range{2003}{2005} as the reference dataset. We assess a $66
+\unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5
 ($R^2 = 0.24$) and $\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 =
@@
     the use of a different subset of climate models (in this study, we use
-    output from all \acs{CMIP} models that is publicly available; see
+    output from all \acs{CMIP} models that are publicly available; see
     \cref{tab:app:a:cmip5_emergent_constraints_part1,%
@@
     coefficient between the predictor and \acs{ECS} is zero. Smaller
-    $p$-values point to higher significance and vice versa (see
+    $p$-values point to a higher significance and vice versa (see
     \cref{subsec:02:emergent_constraints} for details).
@@
     almost all emergent constraints, the constrained best estimate and range of
-    \acs{ECS} is higher in the \acs{CMIP}6 ensemble.
+    \acs{ECS} are higher in the \acs{CMIP}6 ensemble.
   }
@@
 performance of the emergent constraint to be reduced on out-of-sample data
-(like the new \acs{CMIP}6 ensemble), since the emergent relationship appeared
+(like the new \acs{CMIP}6 ensemble) since the emergent relationship appeared
 just by chance and not because of a physically based mechanism.
@@
 which leads to weaker emergent relationships. On the other hand, emergent
-constraints on the less complex \acs{CMIP}3 and \acs{CMIP}5 ensemble may be
+constraints on the less complex \acs{CMIP}3 and \acs{CMIP}5 ensembles may be
 overconfident.
@@
 which also results in a weaker emergent constraint compared with \acs{CMIP}5
-models, as most of them constrained low-cloud feedbacks. They find that on
+models, as most of them constrain low-cloud feedbacks. They find that on
 average, the shortwave low-cloud feedback is larger in \acs{CMIP}6 than in
@@
 fraction) in mixed-phase clouds resulting in less pronounced increases in
-low-level cloud cover and water content with warmer \acp{SST} particularly in
+low-level cloud cover and water content with warmer \acp{SST}, particularly in
 midlatitudes.
@@
 demonstrated by the statistical significance, which is lower in \acs{CMIP}6
-than in \acs{CMIP}5 for all emergent constraint except for ZHA and SHS. As
+than in \acs{CMIP}5 for all emergent constraints except for ZHA and SHS. As
 described in \cref{subsec:02:emergent_constraints}, our test for statistical
@@
 constraint is very sensitive to outliers and the subset of the climate model
-ensemble used to fit the emergent relationship. Such a behavior might not be
+ensemble used to fit the emergent relationship. Such behavior might not be
 unique to the ZHA constraint but could apply to other emergent constraints as
@@
 
-Our analysis makes a number of simplifying assumptions common to other studies,
+Our analysis makes several simplifying assumptions common to other studies,
 such as model independence, discussed in detail in
@@
 constraint of \ac{ECS} than would be achieved in more complex models. As a
-conclusion, we argue that to constrain \ac{ECS} in a more robust way, it might
+conclusion, we argue that to constrain \ac{ECS} more robustly, it might
 be beneficial to apply multivariate approaches that are able to consider
@@
 ensemble mean similar to the participating individual emergent constraints, but
-with lower uncertainty range. Moreover, new \ac{ML} techniques are a promising
-avenue forward for such multivariate approaches and for constraining
+with a lower uncertainty range. Moreover, new \ac{ML} techniques are a
+promising avenue forward for such multivariate approaches and for constraining
 uncertainties in multi-model projections \commentcite{Schlund2020}{see

### 2021-04-09 ch06_paper_gpp/paper_gpp.tex
Spell- and grammer-checked remaining chapters

@@
 is a major source of uncertainty in climate projections \autocite{Bodman2013,
-  Booth2012, Collins2013}. This applies for example to multi-model projections
-of future \ac{GPP} of the \acs{CMIP}5 ensemble. Thus, the aim of this study is
-to reduce uncertainties in multi-model projections of future \ac{GPP} in the
-\acs{CMIP}5 \acs{RCP}8.5 scenario at the end of the \nth{21} century with a new
+  Booth2012, Collins2013}. For example, this applies to multi-model projections
+of future \ac{GPP} of the \acs{CMIP}5 ensemble. Thus, this study aims to reduce
+uncertainties in multi-model projections of future \ac{GPP} in the \acs{CMIP}5
+\acs{RCP}8.5 scenario at the end of the \nth{21} century with a new
 \acs{ML}-based weighting technique and appropriate observations. The new
@@
 concentration, yet other studies provide indications that part of this increase
-is due to climate and land use changes \autocite{Bastos2019, Forkel2016,
+is due to climate and land-use changes \autocite{Bastos2019, Forkel2016,
   Piao2018, Zhao2016}. In \cref{subsec:06:step_2} the \ac{ML} technique used in
@@
 details). The emergent constraint is physically motivated by the hypothesis
-that increasing terrestrial \ac{GPP} is a main driver for the observed changes
-in the \ac{CO2} seasonal cycle amplitude \autocite{Gray2014, Keeling1996,
-  Zhao2014}. This hypothesis is based on the fact that the seasonal cycle in
-atmospheric \ac{CO2} concentration originates from photosynthesis and
-decomposition processes: in summer, photosynthesis removes \ac{CO2} from the
-atmosphere; in winter, additional \ac{CO2} is added to the atmosphere due to
-the decomposition of organic matter \autocite{Keeling1995}. Thus, \ac{CO2}
-fertilization leads to an increase of the \acs{CO2} seasonal cycle amplitude
-due to increased \acs{CO2} removal from the atmosphere. We note that this
-emergent constraint implicitly assumes that there is no temperature-driven
-increase in respiration over time. Therefore, this approach most likely
-overestimates the true sensitivity of \ac{GPP} to an increase in the
-atmospheric \ac{CO2} concentration.
+that increasing terrestrial \ac{GPP} is the main driver for the observed
+changes in the \ac{CO2} seasonal cycle amplitude \autocite{Gray2014,
+  Keeling1996, Zhao2014}. This hypothesis is based on the fact that the
+seasonal cycle in atmospheric \ac{CO2} concentration originates from
+photosynthesis and decomposition processes: in summer, photosynthesis removes
+\ac{CO2} from the atmosphere; in winter, additional \ac{CO2} is added to the
+atmosphere due to the decomposition of organic matter \autocite{Keeling1995}.
+Thus, \ac{CO2} fertilization leads to an increase of the \acs{CO2} seasonal
+cycle amplitude due to increased \acs{CO2} removal from the atmosphere. We note
+that this emergent constraint implicitly assumes that there is no
+temperature-driven increase in respiration over time. Therefore, this approach
+most likely overestimates the true sensitivity of \ac{GPP} to an increase in
+the atmospheric \ac{CO2} concentration.
 
@@
 high latitudes with additional smaller contributions from air-sea carbon fluxes
-and land use changes. For their analysis, \textcite{Piao2018} use \ac{CO2}
+and land-use changes. For their analysis, \textcite{Piao2018} use \ac{CO2}
 records from 26 stations and nine terrestrial ecosystem models. An evaluation
@@
 the \ac{CO2} seasonal cycle amplitude increase in northern high latitudes with
-negligible contributions from land use changes \autocite{Bastos2019}. A further
+negligible contributions from land-use changes \autocite{Bastos2019}. A further
 study by \textcite{Zhao2016} finds that for seven out of nine dynamic
@@
 two remaining models equally attribute this to \ac{CO2} fertilization, climate
-change, and land use changes. Moreover, \textcite{Winkler2019a} discuss
+change, and land-use changes. Moreover, \textcite{Winkler2019a} discuss
 limitations of carbon cycle emergent constraints related to the calculation of
@@
 mixing of \ac{CO2} in the atmosphere, the change in the \ac{CO2} seasonal cycle
-amplitude at \ac{KUM} is representative for the whole globe. Thus, we apply
-this emergent constraint to the global mean \ac{GPP} change. However, we note
-that this approach might introduce errors since the \ac{CO2} fertilization
-effect could be different for tropical and extratropical ecosystems due to
-nutrient limitations and the temperature dependence of \ac{RuBisCO} kinetics
+amplitude at \ac{KUM} is representative of the whole globe. Thus, we apply this
+emergent constraint to the global mean \ac{GPP} change. However, we note that
+this approach might introduce errors since the \ac{CO2} fertilization effect
+could be different for tropical and extratropical ecosystems due to nutrient
+limitations and the temperature dependence of \ac{RuBisCO} kinetics
 \autocite{CraftsBrandner2000} in the tropics. Nevertheless, we argue that we
@@
 range for the observed \ac{CO2} record at \ac{KUM} \autocite{Keeling2005}.
-Similar \textcite{Wenzel2016}, we extract the observed \ac{CO2} amplitude from
-smoothed atmospheric \ac{CO2} concentrations (using a stiff cubic spline
-function plus a four-harmonics functions with linear gain) as provided by
+Similar to \textcite{Wenzel2016}, we extract the observed \ac{CO2} amplitude
+from smoothed atmospheric \ac{CO2} concentrations (using a stiff cubic spline
+function plus a four-harmonics function with linear gain) as provided by
 \textcite{Keeling2005}. For the \acs{CMIP}5 models, the respective grid cell
@@
 \ac{ML} approach which operates on gridded values (see
-\cref{subsec:06:step_2}), we use it as first step to globally rescale the
+\cref{subsec:06:step_2}), we use it as a first step to globally rescale the
 \ac{ESM} output. This important first step allows us to constrain the large
@@
 parameter $\nu \ll 1$ called \emph{learning rate}) and the model is fitted
-again in the same manner. This ensures that the prediction for poorly modeled
-training points in the beginning will gradually improve during the training.
-The whole procedure is repeated until the loss function does not improve
-anymore, as evaluated on an independent validation dataset. Relevant
+again in the same manner. This ensures that the prediction for training points
+that are poorly modeled in the beginning will gradually improve during the
+training. The whole procedure is repeated until the loss function does not
+improve anymore, as evaluated on an independent validation dataset. Relevant
 hyperparameters of this algorithm (maximum number of decision trees, maximum
@@
     values of the diagnostic (black points) are fed into the trained \acs{ML}
-    model to constrain \acs{GPP} for every grid cell/month to a value which
-    best agrees with the observations. (c) For an independent validation of our
-    method, we use an out-of-sample testing setup based on a
-    leave-one-model-out \acl{CV} approach (see \cref{subsec:02:model_weighting}
-    for details). The schematic illustration of Step 2b differs only in the
-    target variable used (fractional \acs{GPP} change instead of absolute
-    \acs{GPP}). \AdaptedFrom{Schlund2020}.
+    model to constrain \acs{GPP} for every grid cell/month to a value that best
+    agrees with the observations. (c) For independent validation of our method,
+    we use an out-of-sample testing setup based on a leave-one-model-out
+    \acl{CV} approach (see \cref{subsec:02:model_weighting} for details). The
+    schematic illustration of Step 2b differs only in the target variable used
+    (fractional \acs{GPP} change instead of absolute \acs{GPP}).
+    \AdaptedFrom{Schlund2020}.
   }
@@
 particular, increases greater than $300 \unit{\%}$ in the fractional \ac{GPP}
-change that is used as target variable in Step 2b are removed. These
+change that is used as the target variable in Step 2b are removed. These
 unrealistically high values (making up about $5.9 \unit{\%}$ of the data) occur
@@
 the training data and use it as hold-out test data. This dataset is excluded
-from the training and can therefore be used for an independent validation of
-our statistical model.
+from the training and can therefore be used for independent validation of our
+statistical model.
 
@@
     \textcite{Wenzel2016}: The global mean fractional \acs{GPP} change after
-    \acs{CO2} doubling in a \acf{1BGC} defined as 10-year mean of the
+    \acs{CO2} doubling in a \acf{1BGC} defined as the 10-year mean of the
     \acs{1BGC} run centered at the time of \acs{CO2} doubling relative to the
@@
 error), which is at the lower end of the original \acs{CMIP}5 range
-($\rangeunit{31}{57}{\%}$). Moreover, the best estimate $39 \unit{\%}$ is also
-slightly lower than the \acs{CMIP}5 \ac{MMM} of $43 \unit{\%}$. The observed
-\ac{GPP} averaged over the years \range{1991}{2000} from the FLUXNET-MTE
-product is $\pmrangeunit{123}{6}{\GtCyr{}}$ \autocite{Jung2011}. Assuming
-independent errors and using Gaussian error propagation, the constrained
-fractional change corresponds to a global mean \ac{GPP} of
+($\rangeunit{31}{57}{\%}$). Moreover, the best estimate of $39 \unit{\%}$ is
+also slightly lower than the \acs{CMIP}5 \ac{MMM} of $43 \unit{\%}$. The
+observed \ac{GPP} averaged over the years \range{1991}{2000} from the
+FLUXNET-MTE product is $\pmrangeunit{123}{6}{\GtCyr{}}$ \autocite{Jung2011}.
+Assuming independent errors and using Gaussian error propagation, the
+constrained fractional change corresponds to a global mean \ac{GPP} of
 $\pmrangeunit{171}{12}{\GtCyr{}}$ (standard error) at the end of the \nth{21}
@@
 absolute \ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) as
-target variable, rescaled using \cref{eq:06:y} to reduce the large
+the target variable, rescaled using \cref{eq:06:y} to reduce the large
 uncertainties in the \acs{CO2} fertilization effect in the models. In the first
@@
 (\acs{LASSO}-1D), and a single-predictor \ac{GBRT} model (\acs{GBRT}-1D). Both
-single-predictor models use the historical \ac{GPP} as only feature. The
+single-predictor models use the historical \ac{GPP} as the only feature. The
 predictive power (in terms of \ac{RMSEP}) of each statistical model is assessed
@@
 algorithm is not overfitting the training data and that the prediction errors
-on the whole are approximately unbiased (see \cref{fig:app:b:residuals:a}).
+are approximately unbiased (see \cref{fig:app:b:residuals:a}).
 
@@
 increase over the \nth{21} century ($39 \unit{\%}$) than the unweighted
-\acs{CMIP}5 ensemble mean ($43 \unit{\%}$). The corresponding ratio $39
+\acs{CMIP}5 ensemble mean ($43 \unit{\%}$). The corresponding ratio of $39
 \unit{\%} / 43 \unit{\%} \approx 0.91$ is approximately equal to the mean value
@@
   ]{
-    Difference (a) and ratio (b) of the both biases shown in
+    Difference (a) and ratio (b) of the biases shown in
     \cref{fig:06:step2a_results:b,fig:06:step2a_results:c}. For panel (b), all
@@
     full \acs{GBRT} model using all predictors and the \acs{GBRT} model using
-    only historical \acs{GPP} as single predictor. (d) Comparison of the
+    only historical \acs{GPP} as a single predictor. (d) Comparison of the
     \acs{GBRT} versus the \acf{LASSO} projections of future \acs{GPP}. This
@@
   fig:app:b:step2a_results_errors}. Analogous to the results from the
-leave-one-model-out \ac{CV} approach, the uncertainties on grid cell level
+leave-one-model-out \ac{CV} approach, the uncertainties on grid-cell level
 are significantly reduced for the \ac{GBRT} prediction compared to the
@@
 supported by \cref{fig:06:step2a_ml_inference:b}, which shows historical
-\ac{GPP} as by far most important feature. On the contrary, second-order
+\ac{GPP} as the by far most important feature. On the contrary, second-order
 corrections originate from the observations in the remaining predictors. These
@@
 \nth{21} century (2100 versus 2000) in the emission-driven \acs{RCP}8.5
-scenario (rescaled using \cref{eq:06:y}). This step is independent from Step 2a
+scenario (rescaled using \cref{eq:06:y}). This step is independent of Step 2a
 and only differs in the target variable (\ac{GPP} change instead of absolute
@@
 heavily dominates the feature importance of the \ac{GBRT} model. Thus, using
-single-predictor models (\acs{GBRT}-1D and \acs{LASSO}-1D) is not necessary in
+single-predictor models (\acs{GBRT}-1D and \acs{LASSO}-1D) is not necessary for
 Step 2b. \Cref{fig:06:step2b_ml_inference:b,fig:06:step2b_ml_inference:d} show
@@
     (ignoring all other features) is color-coded according to the triangle in
-    the lower left corner. Over large parts of the globe, \acs{T} is the
+    the lower-left corner. Over large parts of the globe, \acs{T} is the
     dominant feature. (d) Relative global feature importance for the different
@@
 
-All four patterns of \cref{fig:06:step2b_results} show very similar
+All four patterns in \cref{fig:06:step2b_results} show very similar
 geographical patterns. Obvious exceptions are the Sahara Desert and the Arabian
@@
 the greening) in the high-latitude regions with high \ac{GPP} changes. On the
-contrary, in the tropical ecosystems the growing season already covers the
-whole year. Thus, a further climate change--induced extension is not possible,
+contrary, in tropical ecosystems, the growing season already covers the whole
+year. Thus, a further climate change--induced extension is not possible,
 leading to a smaller fractional \ac{GPP} change in regions closer to the
@@
 
-To get further insights in the \ac{GBRT} algorithm, we compare it directly to
+To get further insights into the \ac{GBRT} algorithm, we compare it directly to
 the rescaled \acs{CMIP}5 \ac{MMM} (r\acs{MMM}), which is illustrated in
@@
 latitudes as supported by the feature map in
-\cref{fig:06:step2b_ml_inference:c} that shows \ac{LAI} as dominant predictor
-in this region. However, in the absence of a robust physical mechanism this
-connection cannot be proven. For both target variables, the \ac{GBRT} algorithm
-is superior to all other considered statistical models in terms of prediction
-uncertainty evaluated in a leave-one-model-out \ac{CV} approach among different
-statistical models. Compared to the unweighted \acs{CMIP}5 \ac{MMM} the mean
-prediction error is reduced by $48 \unit{\%}$ (Step 2a) and $16 \unit{\%}$
-(Step 2b); compared to a linear \ac{LASSO} regression model the mean prediction
-error is reduced by approximately $3 \unit{\%}$ for both cases. However, local
-standard errors are still large for both target variables, even for the
-\ac{GBRT} model. Due to the unknown covariance structure a global aggregation
-of these errors is not possible. Step 2a mainly corrects the bias of the
-simulated absolute historical \ac{GPP} relative to observations
-\autocite{Anav2013}. Consequently, the historical \ac{GPP} is by far the most
-important predictor for future absolute \ac{GPP} (regions with high \ac{GPP} in
-the past are likely to have high \ac{GPP} in the future and vice versa).
-However, as we show in \cref{fig:06:step2a_results_biases}, the \ac{GBRT} model
-expands this bias correction by taking more predictors than the historical
-\ac{GPP} into account. A similar result is provided by Step 2b:
+\cref{fig:06:step2b_ml_inference:c} that shows \ac{LAI} as the dominant
+predictor in this region. However, in the absence of a robust physical
+mechanism, this connection cannot be proven. For both target variables, the
+\ac{GBRT} algorithm is superior to all other considered statistical models in
+terms of prediction uncertainty evaluated in a leave-one-model-out \ac{CV}
+approach among different statistical models. Compared to the unweighted
+\acs{CMIP}5 \ac{MMM} the mean prediction error is reduced by $48 \unit{\%}$
+(Step 2a) and $16 \unit{\%}$ (Step 2b); compared to a linear \ac{LASSO}
+regression model the mean prediction error is reduced by approximately $3
+\unit{\%}$ for both cases. However, local standard errors are still large for
+both target variables, even for the \ac{GBRT} model. Due to the unknown
+covariance structure, a global aggregation of these errors is not possible.
+Step 2a mainly corrects the bias of the simulated absolute historical \ac{GPP}
+relative to observations \autocite{Anav2013}. Consequently, the historical
+\ac{GPP} is by far the most important predictor for future absolute \ac{GPP}
+(regions with high \ac{GPP} in the past are likely to have high \ac{GPP} in the
+future and vice versa). However, as we show in
+\cref{fig:06:step2a_results_biases}, the \ac{GBRT} model expands this bias
+correction by taking more predictors than the historical \ac{GPP} into account.
+A similar result is provided by Step 2b:
 \cref{fig:06:step2b_results_biases:b,fig:06:step2b_results_biases:c} show the
@@
 target variable (fractional \ac{GPP} change) is already normalized with the
-historical \ac{GPP}. Therefore, \ac{T} (second most important feature in Step
-2a) is now the most important predictor instead of historical \ac{GPP} (see
-\cref{fig:06:step2b_ml_inference:d}). As shown in
+historical \ac{GPP}. Therefore, \ac{T} (the second most important feature in
+Step 2a) is now the most important predictor instead of historical \ac{GPP}
+(see \cref{fig:06:step2b_ml_inference:d}). As shown in
 \cref{fig:06:step2b_ml_inference:c}, the different features have also different
@@
 an implicit performance weighting: The \ac{GBRT} model creates predictions
-based on today's conditions (in form of variables which are physically
-connected to \ac{GPP}). However, unlike other performance weighting techniques
+based on today's conditions (in form of variables that are physically connected
+to \ac{GPP}). However, unlike other performance weighting techniques
 \autocite{Knutti2017a, Sanderson2017}, we do not assign constant weights to the
-different climate models. On the contrary, since the algorithm works on grid
-cell level, it is able to adapt to regional climate characteristics and thus
-applies an implicit local and climate-dependent weighting. Moreover, the
+different climate models. On the contrary, since the algorithm works on
+grid-cell level, it is able to adapt to regional climate characteristics and
+thus applies an implicit local and climate-dependent weighting. Moreover, the
 algorithm adds another level of weighting by assigning a relative relevance
 score to the different predictors. We emphasize that this is only an implicit
-weighting, since it is not possible to extract specific values for the
+weighting since it is not possible to extract specific values for the
 individual weights due to the complex structure of the \ac{GBRT} model.
@@
 \ac{GBRT} approach in Step 2 hinges on the assumption that the climate models
-reflect the real world. This is certainly only partially true, since there are
-processes like \ac{CO2} fertilization, land use and land cover changes,
+reflect the real world. This is certainly only partially true since there are
+processes like \ac{CO2} fertilization, land-use and land-cover changes,
 nutrient cycles and limitations \autocite{Du2020,Fleischer2019}, disturbances,
-and induced vegetation dynamics which may strongly alter the future
-trajectories of \ac{GPP} without this being fully encoded in today's
-observational products. Moreover, all \acs{CO2}-related effects are only
-implicitly incorporated in our approach by the target period we are using and
-the global rescaling using the emergent constraint in Step 1. Thus, our
-reduction of uncertainty only applies to the specific setup we are considering
-here: the future \ac{GPP} in the \acs{CMIP}5 \acs{RCP}8.5 scenario at the end
-of the \nth{21} century. It has to be viewed as a nominal reduction, while the
-real uncertainty remains unknown and is likely larger than the nominal one.
-These issues could be tested in the future by including other (even offline)
-models, which account for the aforementioned processes. A true validation will
-only be possible once we experience the changed condition or could be based on
-paleo data. In summary, our approach does not address the coupled system
-directly but hinges on the climate predictions of the individual \acp{ESM} and
-hence is not able to implicitly correct respective biases of the \acs{CMIP}5
-ensemble.
+and induced vegetation dynamics that may strongly alter the future trajectories
+of \ac{GPP} without this being fully encoded in today's observational products.
+Moreover, all \acs{CO2}-related effects are only implicitly incorporated in our
+approach by the target period we are using and the global rescaling using the
+emergent constraint in Step 1. Thus, our reduction of uncertainty only applies
+to the specific setup we are considering here: the future \ac{GPP} in the
+\acs{CMIP}5 \acs{RCP}8.5 scenario at the end of the \nth{21} century. It has to
+be viewed as a nominal reduction, while the real uncertainty remains unknown
+and is likely larger than the nominal one. These issues could be tested in the
+future by including other (even offline) models, which account for the
+aforementioned processes. A true validation will only be possible once we
+experience the changed condition or could be based on paleo data. In summary,
+our approach does not address the coupled system directly but hinges on the
+climate predictions of the individual \acp{ESM} and hence is not able to
+implicitly correct respective biases of the \acs{CMIP}5 ensemble.
 

### 2021-04-09 ch07_conclusion/conclusion.tex
Spell- and grammer-checked remaining chapters

@@
 like the allowable fossil fuel emissions to meet particular warming targets,
-for example the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
+for example, the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
 \autocite{UNFCCC2015}, research in this field of science is not only relevant
-to climate scientists but to policymakers and human society as a whole. In the
+to climate scientists but policymakers and human society as a whole. In the
 light of the large spread of climate sensitivity in the most recent generation
-of climate models from \acs{CMIP}6, that has been demonstrated in
+of climate models from \acs{CMIP}6, which has been demonstrated in
 \cref{ch:04:papers_ecs_tcr_assessment}, a careful statistical evaluation and
@@
 forcing is required to correctly reproduce the observed historical temperature
-trend, resulting in a higher climate sensitivity in the corresponding models. A
+trend, resulting in higher climate sensitivity in the corresponding models. A
 further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in the
@@
 decrease of the coefficient of determination $R^2$ of the emergent relationship
-and a decrease of the statistical significance using the null hypothesis that
+and a decrease in the statistical significance using the null hypothesis that
 there is no correlation between the predictor and \ac{ECS}. Moreover, the
@@
 first step corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2},
-which is primarily driven by the \ac{CO2} fertilization. In the second step, an
+which is primarily driven by \ac{CO2} fertilization. In the second step, an
 \acs{ML}-based climate model weighting approach is used to further constrain
@@
 be interpreted as an implicit performance weighting. However, due to the
-complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
+complex structure of the used \ac{ML} algorithm (\ac{GBRT}), it is not possible
 to extract specific values for the individual weights. Two target variables are
@@
 The massive progress in \ac{AI} and \ac{ML} in recent years in combination with
-ever increasing computational power and resources provided by modern
+ever-increasing computational power and resources provided by modern
 supercomputers has the potential for enormous improvements in climate modeling
 and analysis \autocite{Reichstein2019}. Apart from the presented \acs{ML}-based
-weighting approach, further possible applications of \ac{AI} and \ac{ML} in
-climate science involve for example \acs{ML}-based parameterizations that are
-learned from high-resolution climate models \autocite{Gentine2018, Rasp2018},
-\acs{ML}-based analysis and prediction of forcing patterns
-\autocite{Barnes2019, Mansfield2020}, or learning of entire \acp{ESM} from
-observational products \autocite{Geer2021}. Therefore, the foundation for
-innovative and groundbreaking research in climate science for the near future
-has been laid, which will help to further understand the Earth system and fight
-one of the greatest challenges for humankind today: climate change.
+weighting approach, further examples of applications of \ac{AI} and \ac{ML} in
+climate science involve \acs{ML}-based parameterizations that are learned from
+high-resolution climate models \autocite{Gentine2018, Rasp2018}, \acs{ML}-based
+analysis and prediction of forcing patterns \autocite{Barnes2019,
+  Mansfield2020}, or learning of entire \acp{ESM} from observational products
+\autocite{Geer2021}. Therefore, the foundation for innovative and
+groundbreaking research in climate science for the near future has been laid,
+which will help to further understand the Earth system and fight one of the
+greatest challenges for humankind today: climate change.

### 2021-04-09 end_appendix/appendix.tex
Spell- and grammer-checked remaining chapters

@@
 The minimization procedure of the loss function (mean square error with
-additional sample weights determined by the grid cell areas) is called
+additional sample weights determined by the grid-cell areas) is called
 \emph{stochastic gradient boosting} and is explained in detail by
@@
 the learning procedure are the total number of decision trees $H$, the
-complexity of the individual trees (for example measured by the maximum tree
+complexity of the individual trees (for example, measured by the maximum tree
 depth), and the learning rate $\nu \ll 1$. The latter parameter is used for
@@
 \end{equation}
-The specified error ranges for the \ac{MMM} approaches are calculated in a
-similar way: The constant \ac{SPE} per grid cell is estimated by the mean
+The specified error ranges for the \ac{MMM} approaches are calculated
+similarly: The constant \ac{SPE} per grid cell is estimated by the mean
 \ac{RMSEP} given in the leave-one-model-out \ac{CV} experiment (see
@@
 unbiased (zero mean) for the training and the test dataset. This justifies the
-use of the \ac{RMSEP} to estimate the \ac{SPE}, since for unbiased residuals
+use of the \ac{RMSEP} to estimate the \ac{SPE} since for unbiased residuals
 the \ac{RMSEP} is equal to the standard deviation of the residuals.
@@
     \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO} model using only the
-    historical \acs{GPP} as single predictor. (d) \Acf{GBRT} model using only
-    the historical \acs{GPP} as single predictor. (e) \acs{LASSO} model using
+    historical \acs{GPP} as a single predictor. (d) \Acf{GBRT} model using only
+    the historical \acs{GPP} as a single predictor. (e) \acs{LASSO} model using
     all predictors. (f) \acs{GBRT} model using all predictors.
@@
     (b) Rescaled \acs{CMIP}5 \acs{MMM} using \cref{eq:06:y}. (c) \Acf{LASSO}
-    model using only the historical \acs{GPP} as single predictor. (d)
-    \Acf{GBRT} model using only the historical \acs{GPP} as single predictor.
+    model using only the historical \acs{GPP} as a single predictor. (d)
+    \Acf{GBRT} model using only the historical \acs{GPP} as a single predictor.
     (e) \acs{LASSO} model using all predictors. (f) \acs{GBRT} model using all

### 2021-04-09 ch03_esmvaltool/esmvaltool.tex
Further style optimization

@@
 in \cref{fig:03:esmvaltool_schematic}). This ensures that the \ac{ESMValTool}
-can process arbitrary observational datasets, which can, for example, be used
+can process arbitrary observational datasets, which can be used, for example,
 to assess the skill of climate model simulations. More details on the structure

### 2021-04-09 ch01_introduction/introduction.tex
Further style optimization

@@
 relevant for climate change \autocite{Collins2013, Flato2013}. A crucial and
-policy-relevant example for this is climate sensitivity, which refers to the
+policy-relevant example of this is climate sensitivity, which refers to the
 change in the \ac{GSAT} that results from a change in the radiative forcing.

### 2021-04-09 ch02_scientific_background/scientific_background.tex
Further style optimization

@@
   \cref{fig:02:sources_of_uncertainty}{\color{BrownRed} b}}. However, such an
-increase corresponds to a previously unmeasured uncertainty. An example for
-this has recently happened in \acs{CMIP}6: most likely due to changes in the
-cloud representation of the models the spread in the projected \ac{GSAT} caused
-by a doubling of the atmospheric \ac{CO2} concentration has substantially
-increased in \acs{CMIP}6 compared to older \ac{CMIP} generations
-\autocite{Zelinka2020}. On the other hand, the climate response uncertainty can
-decrease with additional model improvements and a better understanding of the
-Earth system (see \cref{fig:02:sources_of_uncertainty}{\color{BrownRed} c}).
-Moreover, it can also be reduced by observational constraints, which is the
-main topic of this thesis.
+increase corresponds to a previously unmeasured uncertainty. An example of this
+has recently happened in \acs{CMIP}6: most likely due to changes in the cloud
+representation of the models the spread in the projected \ac{GSAT} caused by a
+doubling of the atmospheric \ac{CO2} concentration has substantially increased
+in \acs{CMIP}6 compared to older \ac{CMIP} generations \autocite{Zelinka2020}.
+On the other hand, the climate response uncertainty can decrease with
+additional model improvements and a better understanding of the Earth system
+(see \cref{fig:02:sources_of_uncertainty}{\color{BrownRed} c}). Moreover, it
+can also be reduced by observational constraints, which is the main topic of
+this thesis.
 
@@
 sensitivity metrics \ac{ECS} and \ac{TCR} introduced in the previous
-\lcnamecref{sec:02:climate_sensitivity} are by definition independent from
-carbon cycle--related effects, since both of them are defined in terms of a
-doubling of the atmospheric \ac{CO2} concentration, not in terms of \ac{CO2}
-emissions. Nevertheless, to assess policy-relevant metrics like the allowable
-fossil fuel emissions to meet particular warming targets, for example, the $1.5
+\lcnamecref{sec:02:climate_sensitivity} are by definition independent of carbon
+cycle--related effects, since both of them are defined in terms of a doubling
+of the atmospheric \ac{CO2} concentration, not in terms of \ac{CO2} emissions.
+Nevertheless, to assess policy-relevant metrics like the allowable fossil fuel
+emissions to meet particular warming targets, for example, the $1.5
 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015}, it is
@@
 climate models are independent, equally plausible, distributed around reality,
-and that the projected multi-model range is representative for the uncertainty
+and that the projected multi-model range is representative of the uncertainty
 in the projected quantity \autocite{Knutti2017a}. However, since the \ac{CMIP}
@@
   Sanderson2017}. A further limitation of this approach is the use of an
-ordinary least squares linear regression model. This is not always appropriate,
-for example, when a non-linear emergent relationship is expected
+ordinary least squares linear regression model. Sometimes, this is not
+appropriate; for example, when a non-linear emergent relationship is expected
 \autocite{Nijsse2020} or when additional physical considerations further

### 2021-04-09 ch07_conclusion/conclusion.tex
Further style optimization

@@
 \autocite{UNFCCC2015}, research in this field of science is not only relevant
-to climate scientists but policymakers and human society as a whole. In the
-light of the large spread of climate sensitivity in the most recent generation
-of climate models from \acs{CMIP}6, which has been demonstrated in
+to climate scientists but also to policymakers and human society as a whole. In
+the light of the large spread of climate sensitivity in the most recent
+generation of climate models from \acs{CMIP}6, which has been demonstrated in
 \cref{ch:04:papers_ecs_tcr_assessment}, a careful statistical evaluation and
@@
 the Earth system and a target variable to reduce uncertainties in the target
-variable with observations \autocite{Allen2002}. In total 11 emergent
+variable with observations \autocite{Allen2002}. In total, 11 emergent
 constraints on \ac{ECS} are assessed, which are mostly related to cloud

### 2021-04-09 ch02_scientific_background/scientific_background.tex
Replaced appearances of 'circle' with 'point'

@@
     with an abrupt quadrupling of the atmospheric \acs{CO2} concentration
-    (\nxcotwo{4}) for the \acs{CMIP}5 \acl{MMM} (blue circles). To account for
+    (\nxcotwo{4}) for the \acs{CMIP}5 \acl{MMM} (blue points). To account for
     energy leakage and model drift, a linear fit of the corresponding
@@
     \acs{CO2} concentration (\onepctcotwo{}) for the \acf{CESM} of the
-    \acf{NCAR} (blue circles). The temperature change is calculated relative to
+    \acf{NCAR} (blue points). The temperature change is calculated relative to
     a corresponding pre-industrial control simulation smoothed with a linear
@@
 or present-day climate. The basis for the relationship is output from the
-different climate models of a multi-model ensemble (blue circles in
+different climate models of a multi-model ensemble (blue points in
 \cref{fig:02:emergent_constraint}). Using an observation of $x$, this
@@
     (\eg{} past or present-day trend or variation) for the climate models of a
-    multi-model ensemble (blue circles). With an observation of $x$,
+    multi-model ensemble (blue points). With an observation of $x$,
     uncertainties in $y$ in the multi-model ensemble, illustrated by the yellow
@@
 schematic that illustrates this linear relationship (gray surface) for the
-different climate models (colored circles) for two diagnostics ($K = 2$). Using
+different climate models (colored points) for two diagnostics ($K = 2$). Using
 ordinary least squares regression, the estimated linear coefficients
@@
     estimate of the target variable $\hat{y}_0$ following
-    \cref{eq:02:mder_best_estimate}. The black circle indicates the best
+    \cref{eq:02:mder_best_estimate}. The black point indicates the best
     estimate for the target variable $y$ given by the observed values of
-    $x^{(1)}$ and $x^{(2)}$. Each of the remaining colored circles represents a
+    $x^{(1)}$ and $x^{(2)}$. Each of the remaining colored points represents a
     single climate model of the multi-model ensemble. The vertical dashed lines
@@
 of the predictors $\bm{x}_0$ are illustrated with horizontal black lines, and
-the best estimate $\hat{y}_0$ is shown as a black circle. By combining
+the best estimate $\hat{y}_0$ is shown as a black point. By combining
 \cref{eq:02:linear_coefficients,eq:02:mder_best_estimate} and comparing this to

### 2021-04-09 ch03_esmvaltool/esmvaltool.tex
Replaced appearances of 'circle' with 'point'

@@
     simulations from \acs{CMIP}5 models (colored lines) and observations from
-    the HadCRUT4 dataset (black circles) \autocite{Morice2012}. (b) Emergent
+    the HadCRUT4 dataset (black points) \autocite{Morice2012}. (b) Emergent
     relationship between \acs{ECS} and the temperature variability metric

### 2021-04-09 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Further style optimization
Replaced appearances of 'circle' with 'point'

--- 2021-04-09 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
@@
     \acs{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}6 \acs{MMM}
-    (circles). To account for energy leakage and model drift, a linear fit of
+    (points). To account for energy leakage and model drift, a linear fit of
     the corresponding pre-industrial control run is subtracted from the
@@
     \cref{subsec:02:ecs} and \cref{fig:02:gregory_regression_different_years}):
-    Using only the first 20 years of the simulation (blue circles and line)
+    Using only the first 20 years of the simulation (blue points and line)
     yields a considerably smaller \acs{ECS} than using only the last 130 years
-    of the simulation (orange circles and line). \AdaptedFrom{Meehl2020}.
+    of the simulation (orange points and line). \AdaptedFrom{Meehl2020}.
   }
@@
     with the corresponding \acf{MMM} illustrated with the horizontal black
-    lines (the green circle represents the output from a single climate
-    model). The numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6
-    models (see \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}).
+    lines (the green point represents the output from a single climate model).
+    The numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models
+    (see \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}).
     \AdaptedFrom{Meehl2020}.
@@
 1990, in which it is estimated with $2.3 \unit{K}$ using only a single climate
-model \commentcite{Bretherton1990}{see single green circle in
+model \commentcite{Bretherton1990}{see single green point in
   \cref{fig:04:historical_ecs_tcr_ranges}}. Over the years, the modeled range
@@
 interactions that are relevant in the context of high climate sensitivity in
-\acs{CMIP}6 are found for example in the \ac{HadGEM} Version 3. The new aerosol
-scheme in this model suppresses a strong negative cloud feedback over the
-Southern Ocean that is found in the model's predecessor versions
+\acs{CMIP}6 can be found, for example, in the \ac{HadGEM} Version 3. The new
+aerosol scheme in this model suppresses a strong negative cloud feedback over
+the Southern Ocean that is found in the model's predecessor versions
 \autocite{BodasSalcedo2019}. The Southern Ocean is one of the most pristine

### 2021-04-09 ch05_paper_ecs/paper_ecs.tex
Replaced appearances of 'circle' with 'point'

@@
     standard error (gray shaded area). The horizontal dashed lines show the
-    best estimates of the constrained \acf{ECS}, and the colored circles mark
+    best estimates of the constrained \acf{ECS}, and the colored points mark
     the \aclp{MMM} for \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange). (b), (d),
@@
     Emergent relationship ZHA \autocite{Zhai2015} for different subsets of
-    \acs{CMIP}5 models. Blue circles show the 15 \acs{CMIP}5 models used in the
+    \acs{CMIP}5 models. Blue points show the 15 \acs{CMIP}5 models used in the
     original publication (except for \acs{CESM}1-CAM5). The solid blue line and
@@
     including the uncertainty range. In our study, we added 11 more \acs{CMIP}5
-    models (red circles). The corresponding emergent relationship that
-    considers all available \acs{CMIP}5 models is shown in red colors. This
-    relationship shows a considerably lower coefficient of determination
-    ($R^2$) and higher $p$-value than the relationship using the original
-    subset of \acs{CMIP}5 models. The vertical dashed line and shaded area
-    correspond to the observational reference, and the horizontal dashed lines
-    show the corresponding constraints on the \acf{ECS} using this observation.
+    models (red points). The corresponding emergent relationship that considers
+    all available \acs{CMIP}5 models is shown in red colors. This relationship
+    shows a considerably lower coefficient of determination ($R^2$) and higher
+    $p$-value than the relationship using the original subset of \acs{CMIP}5
+    models. The vertical dashed line and shaded area correspond to the
+    observational reference, and the horizontal dashed lines show the
+    corresponding constraints on the \acf{ECS} using this observation.
     \AdaptedFrom{Schlund2020a}.
@@
 \cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
-(colored circles), in which the connecting line between the \acs{CMIP}5 and
+(colored points), in which the connecting line between the \acs{CMIP}5 and
 \acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships

### 2021-04-12 ch01_introduction/introduction.tex
Added acknowledgements

@@
 radiation, the outgoing infrared (longwave) radiation that is reflected from
-the planet's surface is partly absorbed by the corresponding \ac{GHG}
-molecules through excitation of their corresponding vibrational modes and
-re-emitted isotropically. This process traps energy near the surface and
-leads to a warming of the Earth's surface and the lower atmosphere. The
-individual impact of the different drivers of climate change is measured with
-the so-called \emph{radiative forcing}, which quantifies the change in energy
+the planet's surface is partly absorbed by \ac{GHG} molecules through
+excitation of their corresponding vibrational modes and re-emitted
+isotropically. This process traps energy near the surface and leads to a
+warming of the Earth's surface and the lower atmosphere. The individual
+impact of the different drivers of climate change is measured with the
+so-called \emph{radiative forcing}, which quantifies the change in energy
 fluxes caused by changes in these drivers relative to pre-industrial

### 2021-04-12 ch03_esmvaltool/esmvaltool.tex
Added acknowledgements

@@
 latent heat vaporization constant. The longwave/shortwave clear-sky net
-\ac{TOA} radiations (\emph{rlntcs}/\emph{rsntcs}) is defined as the difference
+\ac{TOA} radiation (\emph{rlntcs}/\emph{rsntcs}) is defined as the difference
 between the longwave/shortwave incoming \ac{TOA} radiation assuming clear-sky

### 2021-04-12 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added acknowledgements

@@
 than using only the last 130 years of the simulation, which gives \iac{ECS} of
-$4.05 \unit{K}$. As thoroughly describes in \cref{subsec:02:ecs}, the reason
+$4.05 \unit{K}$. As thoroughly described in \cref{subsec:02:ecs}, the reason
 for this is the state and time dependence of the climate feedback parameter,
@@
 assessed for all \acs{CMIP}5 and \acs{CMIP}6 models where the necessary
-temperature and radiation data are available (as of \TheMonth{}).
+temperature and radiation data are available.
 \Cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6} give an overview of the
@@
 model ensembles is notably smaller with about $10 \unit{\%}$ ($0.19 \unit{K}$).
-The spread in the multi-model ensembles (expressed as the multi-model standard
-deviation) shows an even larger increase in the \acs{CMIP}6 ensemble: For
-\ac{ECS}, the relative difference is about $49 \unit{\%}$ ($0.72 \unit{K}$ in
-\acs{CMIP}5 to $1.07 \unit{K}$ in \acs{CMIP}6), and for \ac{TCR}, the relative
-difference is about $17 \unit{\%}$ ($0.36 \unit{K}$ in \acs{CMIP}5 to $0.42
-\unit{K}$ in \acs{CMIP}6). The main reason for the increased \ac{MMM} and
-spread in the \acs{CMIP}6 ensemble is the existence of several models with very
-high values of \ac{ECS} and \ac{TCR}. In addition to that, there also exist
-some models with very low values of \ac{ECS}. Therefore, the \acs{CMIP}6 model
-ranges of \ac{ECS} and \ac{TCR} are well outside the corresponding assessed
-ranges given by the latest published \ac{AR} of the \ac{IPCC} from 2013
+The multi-model spread (expressed as the multi-model standard deviation) shows
+an even larger increase in the \acs{CMIP}6 ensemble: For \ac{ECS}, the relative
+difference is about $49 \unit{\%}$ ($0.72 \unit{K}$ in \acs{CMIP}5 to $1.07
+\unit{K}$ in \acs{CMIP}6), and for \ac{TCR}, the relative difference is about
+$17 \unit{\%}$ ($0.36 \unit{K}$ in \acs{CMIP}5 to $0.42 \unit{K}$ in
+\acs{CMIP}6). The main reason for the increased \ac{MMM} and spread in the
+\acs{CMIP}6 ensemble is the existence of several models with very high values
+of \ac{ECS} and \ac{TCR}. In addition to that, there also exist some models
+with very low values of \ac{ECS}. Therefore, the \acs{CMIP}6 model ranges of
+\ac{ECS} and \ac{TCR} are well outside the corresponding assessed ranges given
+by the latest published \ac{AR} of the \ac{IPCC} from 2013
 \autocite{Stocker2013} with $\rangeunit{1.8}{5.6}{K}$ for \ac{ECS} (\acs{AR}5:
@@
 To illustrate the results of the previous
-\lcnamecref{sec:04:evaluation_ecs_and_tcr} and put them into historical
+\lcnamecref{sec:04:evaluation_ecs_and_tcr} and to put them into historical
 context, \cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS}
@@
 bound are more extreme than in any previous climate model generation, resulting
-in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$ (as of
-\TheMonth{}). This increase in range is particularly relevant for the upper
-bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
-$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that exceed
-this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third (14
-models) have \iac{ECS} above $4.5 \unit{K}$. In addition to the analysis
-presented here, several modeling centers independently confirmed high \ac{ECS}
-values in their models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019,
-  Wyser2020}. A list of possible reasons for the increased climate sensitivity in
-the \acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in
-the shortwave cloud feedback over the Southern Ocean is given in
+in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$. This
+increase in range is particularly relevant for the upper bound, which is about
+$1.1 \unit{K}$ higher than the assessed upper bound of $4.5 \unit{K}$.
+Moreover, it is not only a small fraction of models that exceed this upper
+bound: out of the 42 analyzed \acs{CMIP}6 models, a third (14 models) have
+\iac{ECS} above $4.5 \unit{K}$. In addition to the analysis presented here,
+several modeling centers independently confirmed high \ac{ECS} values in their
+models \autocite{Andrews2019, BodasSalcedo2019, Gettelman2019, Wyser2020}. A
+list of possible reasons for the increased climate sensitivity in the
+\acs{CMIP}6 ensemble including aerosol-cloud interactions and changes in the
+shortwave cloud feedback over the Southern Ocean is given in
 \cref{sec:04:possible_reasons_high_ecs_cmip6}.

### 2021-04-12 end_acknowledgments/acknowledgments.tex
Fixed use of acronyms followed by non-whitespace
Added acknowledgements

@@
 \emph{Coordinated Research in Earth Systems and Climate: Experiments,
-  kNowledge, Dissemination and Outreach (CRESCENDO)} under grant agreement
-number 641816, the European Union's Horizon 2020 project \emph{Climate-Carbon
-  Interactions in the Coming Century (4C)} under grant agreement number 821003,
+  kNowledge, Dissemination and Outreach (CRESCENDO)} under Grant Agreement
+Number 641816, the European Union's Horizon 2020 project \emph{Climate-Carbon
+  Interactions in the Coming Century (4C)} under Grant Agreement Number 821003,
 the European Research Council Synergy Grant \emph{Understanding and Modelling
-  the Earth System with Machine Learning (USMILE)} under grant agreement number
+  the Earth System with Machine Learning (USMILE)} under Grant Agreement Number
 855187, and the \emph{Advanced Earth System Model Evaluation for CMIP
@@
 funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational
-resources of the Deutsches Klima Rechenzentrum (DKRZ, Hamburg, Germany) where
+resources of the German Climate Computing Center (DKRZ; Hamburg, Germany) where
 the \ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
@@
 
-Personal acknowledgments...
+First of all, I would like to express my sincere gratitude to my supervisor and
+primary examiner, \Prof{} \Dr{} Veronika Eyring. Without her continuous support
+and advice, patience, motivation, and expertise, this \PhD{} project would not
+have been possible. I further thank her for providing me the opportunity to
+visit many international conferences, which allowed me to meet amazing people
+and collaborate with colleagues from all over the world.
+
+I am grateful to my second examiner, \Prof{} \Dr{} Pierre Gentine, for his
+great advice and comments on my work.
+
+My sincere thanks go to all my co-authors of my first-author studies, namely
+\Prof{} \Dr{} Gustau Camps-Valls, \Prof{} \Dr{} Veronika Eyring, \Prof{} \Dr{}
+Pierre Friedlingstein, \Prof{} \Dr{} Pierre Gentine, \Dr{} Axel Lauer, \Prof{}
+\Dr{} Markus Reichstein, and \Prof{} \Dr{} Steven Sherwood, for their advice,
+feedback, and collaboration.
+
+I would like to thank \Prof{} \Dr{} Gustau Camps-Valls and his Image and Signal
+Processing (ISP) group for hosting me during my research stay at the University
+of Valencia. I really enjoyed the time in Valencia, where I had the opportunity
+to meet great people and expand my knowledge about \acl{ML}.
+
+Many thanks to \Dr{} Axel Lauer for being an amazing second author and for the
+endless patience he showed whenever I had questions for him. I am also very
+grateful to \Dr{} Birgit Hassler for her advice in all circumstances and for
+providing the urgently-needed chocolate and cookies during long hours in the
+office. A warm thank you to all my colleagues at DLR-EVA and IUP for their
+professional advice, helpful discussions, and the great time we spent together
+inside and outside of the office.
+
+A special thank you to the \ac{ESMValTool} core development team for providing
+me useful insights into modern software development and for the great workshops
+and beers we shared together.
+
+I am deeply grateful to my friends for all the amazing time we spent together
+in the many years of our friendship, which allowed me to stay focused on my
+work.
+
+A very special thanks to the (extended) Langer family for their advice and
+support in all situations of life.
+
+I am forever grateful to my parents and my sister for their everlasting support
+and their unconditional love, without which this work would not have been
+possible.
+
+Finally, I would like to thank Hannah with all my heart, for everything.

### 2021-04-12 utils/get_line_lengths.py
Added script to check line lengths in files

@@ -1,0 +1,59 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+# This file is used to produce materials for the following work:
+
+# Doctoral Dissertation
+# (c) by Manuel Schlund
+
+# This work is licensed under a
+# Creative Commons Attribution 4.0 International License.
+
+# You should have received a copy of the license along with this
+# work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
+
+"""Get maximum line length of all ``tex`` files."""
+
+from pathlib import Path
+
+
+ROOT = Path(__file__).parents[1].resolve()
+
+
+def get_file_as_str(filename):
+    """Read file and return ``str``."""
+    with open(filename, 'r') as f_in:
+        file_contents = f_in.read()
+    return file_contents
+
+
+def check_file(filename):
+    """Check single file."""
+    with open(filename, 'r') as f_in:
+        lines = f_in.readlines()
+    max_length = 0
+    max_length_idx = 0
+    for (idx, line) in enumerate(lines):
+        idx += 1
+        if len(line) > max_length:
+            max_length = len(line)
+            max_length_idx = idx
+    print(f"File '{filename}':")
+    print(f"    Maximum line length is {max_length:4d}, first appearance in "
+          f"line {max_length_idx:5d}")
+    print()
+
+
+def main():
+    """Main function."""
+    print(f"Using root path '{ROOT}'")
+    print()
+
+    # Iterate over all tex files
+    tex_files = ROOT.rglob('*.tex')
+    for filename in tex_files:
+        check_file(filename)
+
+
+if __name__ == '__main__':
+    main()

### 2021-04-13 02_abstract/abstract.tex
Updated lines of code change in ESMValTool

@@
 the \acs{CO2} seasonal cycle at \acl{KUM} are used to constrain the global mean
-\acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in \acl{RCP}
-8.5 simulations with \acsp{ESM} participating in \acs{CMIP}5 to
-$\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the unconstrained model range
-$\rangeunit{156}{247}{\GtCyr{}}$. In a second step, an \acs{ML} model is used
-to constrain gridded future absolute \acs{GPP} and gridded fractional \acs{GPP}
-change in two independent approaches. For this, observational data is fed into
-the \acs{ML} algorithm that has been trained on \acs{CMIP}5 data to learn
-relationships between present-day physically relevant diagnostics and the
+\acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
+\acfAbstract{RCP} 8.5 simulations with \acsp{ESM} participating in \acs{CMIP}5
+to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to the unconstrained model range
+of $\rangeunit{156}{247}{\GtCyr{}}$. In a second step, an \acs{ML} model is
+used to constrain gridded future absolute \acs{GPP} and gridded fractional
+\acs{GPP} change in two independent approaches. For this, observational data is
+fed into the \acs{ML} algorithm that has been trained on \acs{CMIP}5 data to
+learn relationships between present-day physically relevant diagnostics and the
 target variable. In a leave-one-model-out \acl{CV} approach, the \acs{ML} model

### 2021-04-13 ch03_esmvaltool/esmvaltool.tex
Updated lines of code change in ESMValTool

@@
 
-As of \TheMonth{}, 40408 lines of code have been added and 21050 lines of code
+As of \TheMonth{}, 41780 lines of code have been added and 21713 lines of code
 have been removed from the GitHub repository of \ac{ESMValCore} by the author
@@
 and additions have been implemented to improve the code base and/or to fix
-bugs. In total, the author of this study has added 88668 lines of code and
-removed 47816 lines of code from the \ac{ESMValTool} repository (as of
+bugs. In total, the author of this study has added 98662 lines of code and
+removed 49823 lines of code from the \ac{ESMValTool} repository (as of
 \TheMonth{}). Parts of these implementations are already published in the

### 2021-04-13 end_acknowledgments/acknowledgments.tex
Updated lines of code change in ESMValTool

@@
 First of all, I would like to express my sincere gratitude to my supervisor and
-primary examiner, \Prof{} \Dr{} Veronika Eyring. Without her continuous support
-and advice, patience, motivation, and expertise, this \PhD{} project would not
-have been possible. I further thank her for providing me the opportunity to
+primary examiner, \Prof{} \Dr{} Veronika Eyring. Without her continuous
+support, advice, enthusiasm, expertise, and patience, this \PhD{} project would
+not have been possible. I further thank her for providing me the opportunity to
 visit many international conferences, which allowed me to meet amazing people
-and collaborate with colleagues from all over the world.
+and collaborate with colleagues from all around the world.
 
 I am grateful to my second examiner, \Prof{} \Dr{} Pierre Gentine, for his
-great advice and comments on my work.
+great advice and feedback on my work.
 
-My sincere thanks go to all my co-authors of my first-author studies, namely
-\Prof{} \Dr{} Gustau Camps-Valls, \Prof{} \Dr{} Veronika Eyring, \Prof{} \Dr{}
-Pierre Friedlingstein, \Prof{} \Dr{} Pierre Gentine, \Dr{} Axel Lauer, \Prof{}
-\Dr{} Markus Reichstein, and \Prof{} \Dr{} Steven Sherwood, for their advice,
-feedback, and collaboration.
+My sincere thanks go to all my co-authors, namely \Prof{} \Dr{} Gustau
+Camps-Valls, \Prof{} \Dr{} Veronika Eyring, \Prof{} \Dr{} Pierre
+Friedlingstein, \Prof{} \Dr{} Pierre Gentine, \Dr{} Axel Lauer, \Prof{} \Dr{}
+Markus Reichstein, and \Prof{} \Dr{} Steven Sherwood, for sharing their
+expertise, their advice, and helpful comments on my studies.
 
@@
 Processing (ISP) group for hosting me during my research stay at the University
-of Valencia. I really enjoyed the time in Valencia, where I had the opportunity
-to meet great people and expand my knowledge about \acl{ML}.
+of Valencia. I really enjoyed the time there, where I had the opportunity to
+meet great people and expand my knowledge about \acl{ML}.
 
@@
 office. A warm thank you to all my colleagues at DLR-EVA and IUP for their
-professional advice, helpful discussions, and the great time we spent together
-inside and outside of the office.
+professional advice, helpful discussions, and simply the great time we spent
+together inside and outside of the office.
 
@@
 me useful insights into modern software development and for the great workshops
-and beers we shared together.
+and beers we shared.
 
@@
 
-A very special thanks to the (extended) Langer family for their advice and
-support in all situations of life.
+A very special thanks to the (extended) Langer family for their endless support
+and advice in all situations of life.
 

### 2021-04-13 ch01_introduction/introduction.tex
Fixed formatting of long numbers and added missing page information for refs

@@
 led to atmospheric concentrations of \ac{CO2}, \ac{CH4}, and \ac{N2O} that
-are \enquote{unprecedented in at least the last 800000 years}
+are \enquote{unprecedented in at least the last 800,000 years}
 \autocite{IPCC2014}. The effects of these \ac{GHG} emissions and other

### 2021-04-13 ch02_scientific_background/scientific_background.tex
Fixed formatting of long numbers and added missing page information for refs

@@
 domain. The slow domain with turnover times (reservoir mass of carbon divided
-by exchange flux) of more than 10000 years consists of the large carbon stores
+by exchange flux) of more than 10,000 years consists of the large carbon stores
 in rocks and sediments which are connected to the rapid domain of the carbon
@@
 component of the oceanic carbon reservoir is dissolved inorganic carbon
-(carbonic acid, bicarbonate ions, and carbonate ions) with about $38000
-\unit{GtC}$. Further carbon is stored as dissolved organic carbon (about $700
-\unit{GtC}$), in surface sediments (about $1750 \unit{GtC}$), and in marine
-biota (about $3 \unit{GtC}$, predominantly phytoplankton and other
-microorganisms) \autocite{Ciais2013, Friedlingstein2020}.
+(carbonic acid, bicarbonate ions, and carbonate ions) with about
+$\mathtxt{38,000} \unit{GtC}$. Further carbon is stored as dissolved organic
+carbon (about $700 \unit{GtC}$), in surface sediments (about $1750
+\unit{GtC}$), and in marine biota (about $3 \unit{GtC}$, predominantly
+phytoplankton and other microorganisms) \autocite{Ciais2013,
+  Friedlingstein2020}.
 

### 2021-04-13 ch03_esmvaltool/esmvaltool.tex
Fixed formatting of long numbers and added missing page information for refs

@@
 
-As of \TheMonth{}, 41780 lines of code have been added and 21713 lines of code
-have been removed from the GitHub repository of \ac{ESMValCore} by the author
-of this thesis. Apart from general improvements of the code base, these changes
-and additions mainly include new preprocessor functions that can be applied to
-the input data and derivation scripts for new non-\ac{CMOR} variables.
-\Cref{tab:03:changes_to_esmvalcore} shows a summary of these main
+As of \TheMonth{}, 41,780 lines of code have been added and 21,713 lines of
+code have been removed from the GitHub repository of \ac{ESMValCore} by the
+author of this thesis. Apart from general improvements of the code base, these
+changes and additions mainly include new preprocessor functions that can be
+applied to the input data and derivation scripts for new non-\ac{CMOR}
+variables. \Cref{tab:03:changes_to_esmvalcore} shows a summary of these main
 contributions, which are partly published in the scientific documentation of
@@
 and additions have been implemented to improve the code base and/or to fix
-bugs. In total, the author of this study has added 98662 lines of code and
-removed 49823 lines of code from the \ac{ESMValTool} repository (as of
+bugs. In total, the author of this study has added 98,662 lines of code and
+removed 49,823 lines of code from the \ac{ESMValTool} repository (as of
 \TheMonth{}). Parts of these implementations are already published in the

### 2021-04-13 ch06_paper_gpp/paper_gpp.tex
Fixed formatting of long numbers and added missing page information for refs

@@
 The number of points establishing the relation is dramatically increased from
-about 50 points for classical emergent constraints to 237852 points (Step 2a)
-and 16503 (Step 2b) in our approach. We validate the relationship by extensive
+about 50 points for classical emergent constraints to 237,852 points (Step 2a)
+and 16,503 (Step 2b) in our approach. We validate the relationship by extensive
 out-of-sample testing using a leave-one-model-out \ac{CV} approach and randomly

### 2021-04-13 end_appendix/appendix.tex
Added acknowledgements
Fixed formatting of long numbers and added missing page information for refs

@@
     given in \cref{tab:05:overview_emergent_constraints}, which lists the
-    variables used for each emergent constraint.
+    individual variables used for each emergent constraint.
   }
@@
 means and standard deviations so that they have a mean of zero and unit
-variance. In total, 237852 (16503) training data points, 79284 (5501) hold-out
-test data points, and 46344 (3727) points for the prediction are used in the
-machine learning model for Step 2a (Step 2b).
+variance. In total, 237,852 (16,503) training data points, 79,284 (5501)
+hold-out test data points, and 46,344 (3727) points for the prediction are used
+in the machine learning model for Step 2a (Step 2b).
 

### 2021-04-13 utils/search_in_tex.py

@@ -33,6 +33,8 @@ python search_in_tex.py '\\.{0,7}ref.?\{' | sort -u
 
 python search_in_tex.py '.{50}figure[s]?[^\}A-Za-z]+.{50}'
 
+python search_in_tex.py '\d{5,}'
+
 """
 
 

### 2021-04-13 ch01_introduction/introduction.tex
Small optimizations in introduction

@@
 radiation, the outgoing infrared (longwave) radiation that is reflected from
-the planet's surface is partly absorbed by \ac{GHG} molecules through
+the planet's surface is partly absorbed by the \ac{GHG} molecules through
 excitation of their corresponding vibrational modes and re-emitted
@@
 in today's climate \autocite{Friedlingstein2020}, this process substantially
-slows down global warming and directly contributes to the magnitude of
-climate sensitivity. Other important anthropogenic drivers of climate change
-are the emission of aerosols and land-use/land-cover changes. Apart from
-anthropogenic drivers, there are also natural processes that impact the
+slows down global warming. Other important anthropogenic drivers of climate
+change are the emission of aerosols, and land-use/land-cover changes. Apart
+from anthropogenic drivers, there are also natural processes that impact the
 climate system like changes in solar activity or volcanic eruptions. However,

### 2021-04-13 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Optimized conclusions

@@
 shown in \cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper
-bound are more extreme than in any previous climate model generation, resulting
-in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$. This
-increase in range is particularly relevant for the upper bound, which is about
-$1.1 \unit{K}$ higher than the assessed upper bound of $4.5 \unit{K}$.
+bounds are more extreme than in any previous climate model generation,
+resulting in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$.
+This increase in range is particularly relevant for the upper bound, which is
+about $1.1 \unit{K}$ higher than the assessed upper bound of $4.5 \unit{K}$.
 Moreover, it is not only a small fraction of models that exceed this upper

### 2021-04-13 ch05_paper_ecs/paper_ecs.tex
Optimized conclusions

@@
 correlated with the \ac{TLC} albedo, \ie{} using the covariance of clouds with
-changes in \acp{SST}. Differences in the \ac{TLC} albedo account for more than
+changes in \acfp{SST}. Differences in the \ac{TLC} albedo account for more than
 half of the variance of \ac{ECS} in the \acs{CMIP}5 ensemble. Following

### 2021-04-13 ch07_conclusion/conclusion.tex
Optimized conclusions

@@
 \Cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
-\acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
-on the upper and lower bound than any model range from previous \ac{CMIP}
-generations. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of \ac{ECS}
-is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus $3.23
-\unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
-\acs{CMIP}6 model range is considerably larger than in the assessed range of
-$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
+\acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is more
+extreme on the lower and upper bound than any model range from previous
+\ac{CMIP} generations. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM}
+of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6
+versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of
+the \acs{CMIP}6 model range is considerably larger than in the assessed range
+of $\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
 \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
@@
 \textcite{Schlund2020a}, a substantial reduction of skill for the majority of
-emergent constraints is found when applied to the \acs{CMIP}6 ensemble in
-comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
-decrease of the coefficient of determination $R^2$ of the emergent relationship
-and a decrease in the statistical significance using the null hypothesis that
-there is no correlation between the predictor and \ac{ECS}. Moreover, the
-corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
-\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
-constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
-(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
+emergent constraints is found when they are applied to the \acs{CMIP}6 ensemble
+in comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
+decrease of the coefficients of determination $R^2$ of the emergent
+relationships and a decrease in the statistical significances using the null
+hypothesis that there is no correlation between the predictors and \ac{ECS}.
+Moreover, the corresponding \acp{PDF} for the emergent constraints show wider
+constrained \ac{ECS} ranges for almost all emergent constraints, resulting in
+sizes of the $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
+confidence) of $1.32 \unit{K}$ to $2.70 \unit{K}$ for \acs{CMIP}6, compared to
+$1.16 \unit{K}$ to $1.75 \unit{K}$ in \acs{CMIP}5. Averaged over all emergent
 constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
-\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
-$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
-\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
-emergent constraints. Thus, \Cref{enum:01:question_2}
+\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of $2.97
+\unit{K}$ to $3.88 \unit{K}$ in \acs{CMIP}5 and $3.48 \unit{K}$ to $4.32
+\unit{K}$ in \acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$
+averaged over all emergent constraints. Thus, \Cref{enum:01:question_2}
 (\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{not
@@
 last decades. In the latest generation of climate models contributing to
-\acs{CMIP}6, the range has even increased \autocite{Meehl2020}.
-\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
-changes in the aerosol-cloud interaction and changes in the shortwave cloud
-phase change feedback over the Southern Ocean. While these are likely
-explanations, there might be additional relevant aspects that have not been
-analyzed in detail yet. This could include additional feedback processes that
-are not present in older \ac{CMIP} model generations. Obvious candidates for
-these feedback processes are cloud-related feedbacks, which pose a major source
-of uncertainty in climate sensitivity in modern-day \acp{ESM}
+\acs{CMIP}6, the range has even increased (see
+\Cref{ch:04:papers_ecs_tcr_assessment}). This thesis presents two possible
+reasons for this: changes in the aerosol-cloud interaction and changes in the
+shortwave cloud phase change feedback over the Southern Ocean. While these are
+likely explanations, there might be additional relevant aspects that have not
+been analyzed in detail yet. This could include additional feedback processes
+that are not present in older \ac{CMIP} model generations. Obvious candidates
+for these feedback processes are cloud-related feedbacks, which pose a major
+source of uncertainty in climate sensitivity in modern-day \acp{ESM}
 \autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
@@
 \acl{T} over the \nth{21} century based on observable process-based predictors
-of today's climate and which could serve as a possible baseline for the new
-\acs{ML}-based weighting approach. An exciting and valuable addition to the
+of today's climate. This paper could serve as a possible baseline for the new
+\acs{ML}-based weighting approach. An exciting and valuable addition to the new
 method could be causal inference \autocite{Nowack2020, Runge2019}. In the

### 2021-04-13 end_acknowledgments/acknowledgments.tex
Standardized vspaces

@@
 funding agencies who support \ac{CMIP} and the \ac{ESGF}. The computational
-resources of the German Climate Computing Center (DKRZ; Hamburg, Germany) where
-the \ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
+resources of the German Climate Computing Center (DKRZ) where the
+\ac{ESMValTool} is fully integrated into the \ac{ESGF} infrastructure are
 kindly acknowledged.
+
+\vspace{\baselineskip}
 

### 2021-04-13 ch05_paper_ecs/paper_ecs.tex
Fixed list of figures

@@
 constraints on \acs{ECS} and evaluate them on the \acs{CMIP}5 and \acs{CMIP}6
-ensemble (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
+ensembles (\cref{sec:05:comparison_of_emergent_constraints}). Finally, we
 provide a discussion (\cref{sec:05:discussion}) and a summary of the results
@@
     Emergent constraints BRI, COX, and LIP applied to the \acs{CMIP}5 and
-    \acs{CMIP}6 ensemble.
+    \acs{CMIP}6 ensembles.
   ]{
     Emergent constraints BRI, COX, and LIP applied to the \acs{CMIP}5 (blue)
-    and \acs{CMIP}6 ensemble (orange). (a), (c), (e) Emergent relationships
+    and \acs{CMIP}6 ensembles (orange). (a), (c), (e) Emergent relationships
     (solid lines) for the \acs{CMIP} models (numbers; see
@@
     Overview of the constrained \acf{ECS} ranges given by the 11 emergent
-    constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensemble.
+    constraints using the \acs{CMIP}5 and \acs{CMIP}6 ensembles.
   ]{
@@
     emergent constraints using the \acs{CMIP}5 (blue) and \acs{CMIP}6 (orange)
-    ensemble. The bars correspond to the $66 \unit{\%}$ confidence intervals
+    ensembles. The bars correspond to the $66 \unit{\%}$ confidence intervals
     ($\rangeunit{17}{83}{\%}$) calculated from the constrained \acl{PDF} of

### 2021-04-13 02_abstract/abstract.tex
Small optimizations of abstract

@@
 application of the emergent constraints to \acs{CMIP}6 data shows a decrease in
-skill and statistical significance of the emergent relationship for nearly all
+skill and statistical significance of the emergent relationships for nearly all
 constraints, with this decrease being large in many cases. Consequently, the
-size of the constrained \acs{ECS} ranges ($66 \unit{\%}$ confidence intervals)
-widens by $51 \unit{\%}$ on average in \acs{CMIP}6 compared to \acs{CMIP}5.
-This is likely related to the increased multi-model spread of \acs{ECS} in
+sizes of the constrained \acs{ECS} ranges ($66 \unit{\%}$ confidence intervals)
+widen by $51 \unit{\%}$ on average in \acs{CMIP}6 compared to \acs{CMIP}5. This
+is likely related to the increased multi-model spread of \acs{ECS} in
 \acs{CMIP}6, but may in some cases also be due to spurious statistical

### 2021-08-03 ch02_scientific_background/scientific_background.tex
Optimized conclusions
Prepared document for publication

@@
 form the output of climate models, the corresponding input is mainly given by
-the radiative forcing.
+external forcings, which are incorporated as boundary conditions into the
+climate models. For example, the major driver of anthropogenic climate change,
+the \ac{GHG} \ac{CO2}, is either included with prescribed atmospheric \ac{CO2}
+concentrations (\emph{concentration-driven runs}) or prescribed \ac{CO2}
+emissions (\emph{emission-driven runs}).
+
+Many processes in the Earth system occur at spatial scales much smaller than
+the size of a typical grid cell. Illustrative examples for this are clouds,
+which are usually smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
+cell, but still play an important role in the overall climate system by
+reflecting incoming and outgoing radiation \autocite{Boucher2013}. To
+reasonably approximate these subgrid-scale processes, a concept called
+\emph{parameterization} is used. Instead of simulating a process exactly,
+parameterizations aim to represent the effect of that process at the grid scale
+of the climate model by generating the appropriate forcing terms for the rest
+of the system and the rest of the processes \autocite{Gettelman2016}.
+Parameterizations are necessary to simulate many processes of the Earth system,
+for example, surface heat and moisture fluxes, moist convection, turbulent
+mixing, and radiation \autocite{Holton2004}.
 
@@
 \end{figure}
-
-Many processes in the Earth system occur at spatial scales much smaller than
-the size of a typical grid cell. Illustrative examples for this are clouds,
-which are usually smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
-cell, but still play an important role in the overall climate system by
-reflecting incoming and outgoing radiation \autocite{Boucher2013}. To
-reasonably approximate these subgrid-scale processes, a concept called
-\emph{parameterization} is used. Instead of simulating a process exactly,
-parameterizations aim to represent the effect of that process at the grid scale
-of the climate model by generating the appropriate forcing terms for the rest
-of the system and the rest of the processes \autocite{Gettelman2016}.
-Parameterizations are necessary to simulate many processes of the Earth system,
-for example, surface heat and moisture fluxes, moist convection, turbulent
-mixing, and radiation \autocite{Holton2004}.
 
@@
 world to obtain comparable output and to facilitate model analysis. For this
-reason, the \ac{WGCM} of the \ac{WCRP} initiated the \acf{CMIP} in 1995 to
+reason, the \ac{WGCM} of the \ac{WCRP} initiated the \ac{CMIP} in 1995 to
 \enquote{better understand past, present and future climate changes arising
@@
 
-\begin{figure}[!t]
-  \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
-    ch02_scientific_background/figs/feedbacks_overview.pdf}
-  \caption[
-    Climate feedbacks and corresponding time scales.
-  ]{
-    Climate feedbacks and corresponding time scales. The different signs refer
-    to the signs of the different feedbacks. Positive feedbacks amplify the
-    effect of the external forcing (\eg{} the water vapor feedback) and
-    negative feedbacks diminish the effect of the external forcing (\eg{} the
-    longwave radiation feedback). An example of a feedback that might be either
-    positive or negative is the cloud feedback. The smaller box highlights the
-    large differences in time scales for the various feedbacks.
-    \IPCCReproducedFrom{Cubasch2013}{their Figure 1.2}.
-  }
-  \label{fig:02:feedbacks_overview}
-\end{figure}
-
 As already described in \cref{subsec:02:source_of_uncertainties}, the effects
@@
 
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=\LargeFigureWidth{}]{
+    ch02_scientific_background/figs/feedbacks_overview.pdf}
+  \caption[
+    Climate feedbacks and corresponding time scales.
+  ]{
+    Climate feedbacks and corresponding time scales. The different signs refer
+    to the signs of the different feedbacks. Positive feedbacks amplify the
+    effect of the external forcing (\eg{} the water vapor feedback) and
+    negative feedbacks diminish the effect of the external forcing (\eg{} the
+    longwave radiation feedback). An example of a feedback that might be either
+    positive or negative is the cloud feedback. The smaller box highlights the
+    large differences in time scales for the various feedbacks.
+    \IPCCReproducedFrom{Cubasch2013}{their Figure 1.2}.
+  }
+  \label{fig:02:feedbacks_overview}
+\end{figure}
+
 In contrast to positive feedbacks, negative feedbacks diminish the effect of an
@@
 thickness, cloud top and cloud base height, vertical extent, and the
-geographical distribution of clouds. Examples for robust cloud feedback
+geographical distribution of clouds. Examples for possible cloud feedback
 processes are the increase in cloud top height of high-level clouds in a
@@
 \autocite{Roe2009}. Usually, the idealization of a blackbody Earth without an
-atmosphere is used for that: In equilibrium, the incoming solar irradiance is
-balanced with an outgoing thermal irradiance $J_0$ that solely depends on the
-global mean surface temperature $T_0$ following the Stefan-Boltzmann law
+atmosphere is used for that: In equilibrium, the absorbed incoming solar
+irradiance (after accounting for reflection) is balanced with an outgoing
+thermal irradiance $J_0$ that solely depends on the global mean surface
+temperature $T_0$ following the Stefan-Boltzmann law
 \begin{equation}
@@
 \end{equation}
-$\sigma \approx 5.67 \unit{W m^{-2} K^{-4}}$ is the Stefan-Boltzmann constant.
-To answer an external forcing $F$, the climate system reacts with a response
-$R$ expressed by a change in the global mean surface temperature $\Delta T$:
+$\sigma \approx 5.67 \cdot 10^{-8} \unit{W m^{-2} K^{-4}}$ is the
+Stefan-Boltzmann constant. To answer an external forcing $F$, the climate
+system reacts with a response $R$ expressed by a change in the global mean
+surface temperature $\Delta T$:
 \begin{equation}
@@
 \autocite{Bindoff2013}. Being already used in one of the first assessments of
-the anthropogenic climate change, the \emph{Charney Report} from 1979
+anthropogenic climate change, the \emph{Charney Report} from 1979
 \autocite{Charney1979}, the equilibrium climate sensitivity is one of the
@@
 of the equilibrium climate sensitivity. With the help of climate models,
-\textcite{Sherwood2020} showed that the \ac{ECS} is only about $6 \unit{\%}$
+\textcite{Sherwood2020} show that the \ac{ECS} is only about $6 \unit{\%}$
 lower than the best estimate of the true equilibrium warming obtained from
@@
   \centering
-  \includegraphics[width=\FigureWidth{}]{
+  \includegraphics[width=\LargerFigureWidth{}]{
     ch02_scientific_background/figs/carbon_cycle_schematic.png}
@@
 \label{subsec:02:carbon_cycle_perturbations}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\EvenLargerFigureWidth{}]{
+    ch02_scientific_background/figs/keeling_curve.pdf}
+  \caption[
+    The Keeling Curve.
+  ]{
+    The Keeling Curve: monthly mean atmospheric \acs{CO2} concentration at the
+    Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
+    \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
+    \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
+    concentration is superimposed with a pronounced seasonal oscillation caused
+    by the seasonal \acs{CO2} cycle (see
+    \cref{subsec:02:carbon_cycle_overview}).
+  }
+  \label{fig:02:keeling_curve}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\EvenLargerFigureWidth{}]{
-    ch02_scientific_background/figs/keeling_curve.pdf}
-  \caption[
-    The Keeling Curve.
-  ]{
-    The Keeling Curve: monthly mean atmospheric \acs{CO2} concentration at the
-    Mauna Loa Observatory, Hawaii ($19.5 \unit{\degree N}$, $155.6
-    \unit{\degree W}$; elevation: $3397 \unit{m}$) from 1958 to 2019
-    \autocite{Keeling2005}. The steady increase of the atmospheric \acs{CO2}
-    concentration is superimposed with a pronounced seasonal oscillation caused
-    by the seasonal \acs{CO2} cycle (see
-    \cref{subsec:02:carbon_cycle_overview}).
-  }
-  \label{fig:02:keeling_curve}
-\end{figure}
-
 Apart from warming the Earth by altering its radiation budget, the
@@
 corresponding chemical reactions. Both parameterizations utilize variables that
-are known to affect photosynthesis, such as the availability of light, the
-\ac{CO2} concentration, the soil moisture, and the temperature.
+are known to affect photosynthesis, such as the availability of light, \ac{CO2}
+concentration, soil moisture, and temperature.
 
@@
   \centering
-  \includegraphics[width=\LargeFigureWidth{}]{
+  \includegraphics[width=\EvenLargerFigureWidth{}]{
     ch02_scientific_background/figs/carbon_cycle_processes.pdf}
@@
 \begin{equation}
-  \hat{y}_0 = \bm{x}_0^T \hat{\bm{b}}
+  \hat{y}_0 = \bm{x}_0^T \hat{\bm{b}}.
   \label{eq:02:mder_best_estimate}
@@
 climate to span the desired relationship (similar to emergent constraints).
-However, a much weaker assumption is made traditional in weighting approaches
+However, a much weaker assumption is made in traditional weighting approaches
 which assume that climate models that are better in simulating the past or

### 2021-08-03 ch03_esmvaltool/esmvaltool.tex
Fixed list of figures
Prepared document for publication

--- 2021-04-13 ch03_esmvaltool/esmvaltool.tex
@@
 the scientific documentation of the \ac{ESMValTool}, which is published in four
-peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}.
+peer-reviewed studies \autocite{Eyring2020, Lauer2020, Righi2020, Weigel2021}.
 After a brief overview of the \ac{ESMValTool}'s structure
@@
 
-As of \TheMonth{}, 41780 lines of code have been added and 21713 lines of code
-have been removed from the GitHub repository of \ac{ESMValCore} by the author
-of this thesis. Apart from general improvements of the code base, these changes
-and additions mainly include new preprocessor functions that can be applied to
-the input data and derivation scripts for new non-\ac{CMOR} variables.
-\Cref{tab:03:changes_to_esmvalcore} shows a summary of these main
+As of \TheMonth{}, 43,853 lines of code have been added and 22,494 lines of
+code have been removed from the GitHub repository of \ac{ESMValCore} by the
+author of this thesis. Apart from general improvements of the code base, these
+changes and additions mainly include new preprocessor functions that can be
+applied to the input data and derivation scripts for new non-\ac{CMOR}
+variables. \Cref{tab:03:changes_to_esmvalcore} shows a summary of these main
 contributions, which are partly published in the scientific documentation of
@@
 and additions have been implemented to improve the code base and/or to fix
-bugs. In total, the author of this study has added 98662 lines of code and
-removed 49823 lines of code from the \ac{ESMValTool} repository (as of
+bugs. In total, the author of this study has added 98,680 lines of code and
+removed 49,850 lines of code from the \ac{ESMValTool} repository (as of
 \TheMonth{}). Parts of these implementations are already published in the
@@
 future projections \autocite{Lauer2020}, and diagnostics for extreme events and
-regional evaluation \autocite{Weigel2020}.
+regional evaluation \autocite{Weigel2021}.
 
@@
 constraints on \ac{ECS}, analyzed in great detail for the \acs{CMIP}5 and
-\acs{CMIP}6 ensemble in \cref{ch:05:paper_ecs}, which also includes a
+\acs{CMIP}6 ensembles in \cref{ch:05:paper_ecs}, which also includes a
 description of the calculation of $\Psi$ in \cref{subsec:05:cox}.

### 2021-08-03 ch07_conclusion/conclusion.tex
Prepared document for publication

@@
 to climate scientists but also to policymakers and human society as a whole. In
-the light of the large spread of climate sensitivity in the most recent
-generation of climate models from \acs{CMIP}6, which has been demonstrated in
+light of the large spread of climate sensitivity in the most recent generation
+of climate models from \acs{CMIP}6, which has been demonstrated in
 \cref{ch:04:papers_ecs_tcr_assessment}, a careful statistical evaluation and
@@
 in the technical and scientific documentation of the \ac{ESMValTool}
-\autocite{Eyring2020, Lauer2020, Righi2020, Weigel2020}. Since the open-source
+\autocite{Eyring2020, Lauer2020, Righi2020, Weigel2021}. Since the open-source
 tool is freely available, the code that has been implemented as part of this

### 2021-08-03 end_acknowledgments/acknowledgments.tex
Prepared document for publication

@@
 First of all, I would like to express my sincere gratitude to my supervisor and
-primary examiner, \Prof{} \Dr{} Veronika Eyring. Without her continuous
-support, advice, enthusiasm, expertise, and patience, this \PhD{} project would
-not have been possible. I further thank her for providing me the opportunity to
-visit many international conferences, which allowed me to meet amazing people
-and collaborate with colleagues from all around the world.
+primary examiner, \ThePrimaryExaminer{}. Without her continuous support,
+advice, enthusiasm, expertise, and patience, this \PhD{} project would not have
+been possible. I further thank her for providing me the opportunity to visit
+many international conferences, which allowed me to meet amazing people and
+collaborate with colleagues from all around the world.
 
-I am grateful to my second examiner, \Prof{} \Dr{} Pierre Gentine, for his
-great advice and feedback on my work.
+I am grateful to my secondary examiner, \TheSecondaryExaminer{}, for his great
+advice and feedback on my work.
 
 My sincere thanks go to all my co-authors, namely \Prof{} \Dr{} Gustau
-Camps-Valls, \Prof{} \Dr{} Veronika Eyring, \Prof{} \Dr{} Pierre
-Friedlingstein, \Prof{} \Dr{} Pierre Gentine, \Dr{} Axel Lauer, \Prof{} \Dr{}
-Markus Reichstein, and \Prof{} \Dr{} Steven Sherwood, for sharing their
-expertise, their advice, and helpful comments on my studies.
+Camps-Valls, \ThePrimaryExaminer{}, \Prof{} \Dr{} Pierre Friedlingstein,
+\TheSecondaryExaminer{}, \Dr{} Axel Lauer, \Prof{} \Dr{} Markus Reichstein, and
+\Prof{} \Dr{} Steven Sherwood, for sharing their expertise, their advice, and
+helpful comments on my studies.
 

### 2021-08-04 00_header/metadata.tex
Use PDF/A-2b format for publication

%=============================================================================%
%                            Doctoral Dissertation                            %
%                            (c) by Manuel Schlund                            %
%=============================================================================%

%=============================================================================%
% This work is licensed under a
% Creative Commons Attribution 4.0 International License.
%
% You should have received a copy of the license along with this
% work. If not, see <http://creativecommons.org/licenses/by/4.0/>.
%=============================================================================%

% Titles
\newcommand*{\Prof}{Prof.\@}
\newcommand*{\Dr}{Dr.\@}
\newcommand*{\PhD}{Ph.D.\@}

% Dissertation information
\newcommand*{\TheAuthor}{Manuel Schlund}
\newcommand*{\TheColloquiumDate}{2 August 2021}
\newcommand*{\TheDegree}{\foreignlanguage{ngerman}{%
    Doktor der Naturwissenschaften (\Dr{} rer.\@ nat.\@)}%
}
\newcommand*{\TheInstitute}{Institute of Environmental Physics (IUP)}
\newcommand*{\TheKeywords}{%
  Climate Modeling\sep Climate Model Weighting\sep Climate Projections\sep %
  Climate Sensitivity\sep Emergent Constraints\sep Machine Learning%
}
\newcommand*{\TheLanguage}{en-US}
\newcommand*{\TheLicense}{%
  Creative Commons Attribution 4.0 International License%
}
\newcommand*{\TheLicenseLink}{https://creativecommons.org/licenses/by/4.0/}
\newcommand*{\TheMonth}{August 2021}
\newcommand*{\ThePrimaryExaminer}{\Prof{} \Dr{} Veronika Eyring}
\newcommand*{\TheSecondaryExaminer}{\Prof{} \Dr{} Pierre Gentine}
\newcommand*{\TheSubmissionDate}{13 April 2021}
\newcommand*{\TheThesis}{Doctoral Dissertation}
\newcommand*{\TheTitle}{%
  Constraining Uncertainties in Multi-Model Projections of the Future Climate
  with Observations%
}
\newcommand*{\TheUniversity}{University of Bremen}

### 2021-08-05 ch02_scientific_background/scientific_background.tex
Small linguistic optimizations

@@
 climate models. For each grid cell and time step, a single value per model
-variable is usually given. Examples for such variables provided by a climate
+variable is usually given. Examples of such variables provided by a climate
 model are the atmospheric prognostic variables velocity (horizontal and
@@
 Many processes in the Earth system occur at spatial scales much smaller than
-the size of a typical grid cell. Illustrative examples for this are clouds,
+the size of a typical grid cell. Illustrative examples of this are clouds,
 which are usually smaller than a $100 \unit{km} \times 100 \unit{km}$ grid
@@
 of future forcing (\acp{GHG}, aerosols, land-use changes, \etc{}) humankind
-might take. Examples for these are the \acs{SSP}-based experiments given by
+might take. Examples of these are the \acs{SSP}-based experiments given by
 \ac{ScenarioMIP} that include a variety of different scenarios from a
@@
 thickness, cloud top and cloud base height, vertical extent, and the
-geographical distribution of clouds. Examples for possible cloud feedback
+geographical distribution of clouds. Examples of possible cloud feedback
 processes are the increase in cloud top height of high-level clouds in a
@@
 the reduction in mid- and low-level cloud cover which diminishes the reflection
-of incoming solar radiation and also increases the surface warming
+of incoming solar radiation and also increases surface warming
 \autocite{Boucher2013}. In global climate model ensembles, the overall cloud
@@
 variety of different parameterizations for the different processes
-\autocite{Arora2020}. Examples for photosynthesis (\ac{GPP}) are given by
+\autocite{Arora2020}. Examples of photosynthesis (\ac{GPP}) are given by
 \textcite{Farquhar1980} for \ch{C3} plants and \textcite{Collatz1992} for

### 2021-08-05 ch03_esmvaltool/esmvaltool.tex
Small linguistic optimizations

@@
 applied to the input data. There are two classes of preprocessors: non-optional
-and optional preprocessors. Examples for non-optional preprocessors are the
+and optional preprocessors. Examples of non-optional preprocessors are the
 input/output operations \emph{load} (loading input data from climate models and

### 2021-08-05 00_header/metadata.tex
Expanded PDF metadata

@@
 }
-\newcommand*{\TheLicenseLink}{https://creativecommons.org/licenses/by/4.0/}
+\newcommand*{\TheLicenseURL}{https://creativecommons.org/licenses/by/4.0/}
 \newcommand*{\TheMonth}{August 2021}
@@
 \newcommand*{\TheUniversity}{University of Bremen}
+
+% Plain text of the abstract
+% Note: needs to be changed if abstract.tex changes!
+\newcommand*{\TheAbstract}{%
+  A precise quantification of climate change is crucial to assess optimal
+  mitigation and adaptation strategies. Earth system models (ESMs), which are
+  state-of-the-art climate models that allow numerical simulations of the
+  complex physical, biological, and chemical processes of the Earth system, are
+  common tools to understand and project climate change. Due to the chaotic
+  nature of the climate system, unknowns in future emission pathways, and
+  uncertainties in the climate models, projections of the future climate are
+  associated with large uncertainties. The main focus of this thesis is the
+  analysis of future climate projections from ESMs participating in the Coupled
+  Model Intercomparison Project (CMIP) with the aim to reduce uncertainties in
+  climate projections with observations. In a first step, climate sensitivity
+  (i.e., the temperature response of the climate system to an external forcing)
+  is evaluated in the latest generation of ESMs from CMIP6. For the effective
+  climate sensitivity (ECS), which is an estimation of the equilibrium
+  temperature response that follows a doubling of the atmospheric carbon
+  dioxide (CO2) concentration, a multimodel mean (MMM) of 3.74 K and a
+  multi-model range of 1.8–5.6 K are found. These values are higher than in any
+  previous CMIP ensemble before. Moreover, a third of the analyzed CMIP6 models
+  exceed the upper bound of the likely ECS range of 1.5–4.5 K assessed by the
+  Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment Report
+  (AR5) from 2013. Similarly, the transient response of the climate system to a
+  doubling of CO2, also known as transient climate response (TCR), shows an
+  inter-model range of 1.3–3.0 K with an upper bound again higher than the
+  likely range assessed in AR5 of 1.0–2.5 K. Possible reasons for the increased
+  climate sensitivity in many CMIP6 models are the addition of new prognostic
+  aerosol schemes that include aerosol-cloud interactions and changes in the
+  microphysical representation of mixed-phase clouds. The changes in the
+  mixed-cloud representation reduce the strong negative shortwave cloud phase
+  change feedback over the Southern Ocean that is present in climate models
+  from previous CMIP generations. To reduce uncertainties in ECS projected by
+  the CMIP6 models, 11 published emergent constraints on ECS (mostly derived
+  from models participating in CMIP5, the predecessor generation of CMIP6) are
+  systematically analyzed. Emergent constraints are potentially promising
+  approaches to reduce uncertainties in climate model projections by combining
+  observations and output from ESMs. The focus of this analysis is on testing
+  if these emergent constraints hold for ESMs participating in CMIP6. Since
+  none of the emergent constraints considered here have been derived from the
+  CMIP6 ensemble, the CMIP6 models can be used for crosschecking the emergent
+  constraints on a new model ensemble. The application of the emergent
+  constraints to CMIP6 data shows a decrease in skill and statistical
+  significance of the emergent relationships for nearly all constraints, with
+  this decrease being large in many cases. Consequently, the sizes of the
+  constrained ECS ranges (66 \% confidence intervals) widen by 51 \% on average
+  in CMIP6 compared to CMIP5. This is likely related to the increased
+  multi-model spread of ECS in CMIP6, but may in some cases also be due to
+  spurious statistical relationships or a too-small number of models in the
+  ensemble that the emergent constraint was originally derived from. The
+  corresponding best estimates of ECS given by the emergent constraints also
+  increase from CMIP5 to CMIP6 by 12 \% on average. This can at least be partly
+  explained by the increased number of high-ECS models in CMIP6 without a
+  corresponding change in the constraint predictors, suggesting the emergence
+  of new feedback processes rather than changes in the strength of those
+  previously dominant. The results support previous studies concluding that
+  emergent constraints should be based on an independently verifiable physical
+  mechanism and that process-based emergent constraints on ECS should rather be
+  thought of as constraints for the process or feedback they are actually
+  targeting. To overcome these issues of single-process-oriented emergent
+  constraints, an alternative approach based on machine learning (ML) is
+  introduced. Since this new technique relies on a large number of data points
+  to train the ML algorithm, the scalar climate sensitivity expressed as ECS or
+  TCR is not an appropriate target variable. Therefore, gross primary
+  production (GPP) as a process that contributes to climate sensitivity is
+  studied as an alternative. GPP is the largest flux of the terrestrial carbon
+  uptake and slows down global warming by removing CO2 from the atmosphere. In
+  this analysis, an existing emergent constraint on CO2 fertilization is
+  combined with an ML approach to constrain the spatial variations of
+  multi-model GPP projections. In the first step of the two-step approach,
+  observed changes in the CO2 seasonal cycle at Cape Kumukahi, Hawaii are used
+  to constrain the global mean GPP at the end of the 21st century (2091–2100)
+  in Representative Concentration Pathway (RCP) 8.5 simulations with ESMs
+  participating in CMIP5 to (171 ± 12) GtC yr−1 , compared to the unconstrained
+  model range of 156–247 GtC yr−1 . In a second step, an ML model is used to
+  constrain gridded future absolute GPP and gridded fractional GPP change in
+  two independent approaches. For this, observational data is fed into the ML
+  algorithm that has been trained on CMIP5 data to learn relationships between
+  present-day physically relevant diagnostics and the target variable. In a
+  leave-one-model-out cross-validation approach, the ML model shows superior
+  performance to the CMIP5 MMM. The new approach predicts a higher GPP increase
+  in high latitudes and a lower GPP increase in regions closer to the equator.%
+}

### 2021-08-10 00_header/metadata.tex
Optimized PDF metadata

@@
 \newcommand*{\TheKeywords}{%
-  Climate Modeling\sep Climate Model Weighting\sep Climate Projections\sep %
-  Climate Sensitivity\sep Emergent Constraints\sep Machine Learning%
+  Climate Change\sep Climate Modeling\sep Climate Projections\sep
+  Climate Sensitivity\sep Climate Model Weighting\sep Emergent Constraints\sep
+  Machine Learning%
 }
@@
   temperature response that follows a doubling of the atmospheric carbon
-  dioxide (CO2) concentration, a multimodel mean (MMM) of 3.74 K and a
+  dioxide (CO2) concentration, a multi-model mean (MMM) of 3.74 K and a
   multi-model range of 1.8–5.6 K are found. These values are higher than in any
@@
   none of the emergent constraints considered here have been derived from the
-  CMIP6 ensemble, the CMIP6 models can be used for crosschecking the emergent
+  CMIP6 ensemble, the CMIP6 models can be used for cross-checking the emergent
   constraints on a new model ensemble. The application of the emergent
@@
   in Representative Concentration Pathway (RCP) 8.5 simulations with ESMs
-  participating in CMIP5 to (171 ± 12) GtC yr−1 , compared to the unconstrained
-  model range of 156–247 GtC yr−1 . In a second step, an ML model is used to
+  participating in CMIP5 to (171 ± 12) GtC yr−1, compared to the unconstrained
+  model range of 156–247 GtC yr−1. In a second step, an ML model is used to
   constrain gridded future absolute GPP and gridded fractional GPP change in
